### 2012-12-15 RSLC/paper/paper.tex
Start writeup; have abstract

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

%\documentclass[preprint]{sigplanconf}
\documentclass{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathpartir}
\usepackage{url}
\usepackage{graphics}

% ____________________________________________________________
% Listings Package Configuration
% \usepackage[scaled]{beramono}

\renewcommand*\ttdefault{txtt}
\usepackage[T1]{fontenc}

\RequirePackage{listings}
% \lstdefinelanguage{Rust}{
%   morekeywords={fn,let,enum,if,else,return,mut,const,copy,ref,move,match,struct,type,int,while,float,as,fail},
%   sensitive=true,
%   morecomment=[l]{//},
%   morecomment=[s]{/*}{*/},
%   morestring=[b]",
% }
% \lstset{language=Rust,
%         frame=none,%single,
%         basicstyle=\ttfamily\footnotesize,
%         keywordstyle=\bfseries,
%         commentstyle=\footnotesize,
%         tabsize=2,
%         showstringspaces=false,
%         flexiblecolumns=false,
%         mathescape=true,
%         escapeinside={\#}{\#},
%         % numbers=left, numberstyle=\tiny, numbersep=5pt %, stepnumber=2, 
%         % literate={->}{{$\rightarrow$}}1 {<->}{{$\leftarrow$}}1
% }

% This Deep Tex Voodoo is from
%   <http://www.latex-community.org/forum/viewtopic.php?f=5&t=2072>
% It's purpose is to make \lstinline normal size, without affecting
% \lstinputlisting.  It seems to work but I have no idea how or why,
% and I rather hope never to learn.
\makeatletter
\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
\makeatother

\newcommand{\Figref}[1]{Figure \ref{#1}}
\newcommand{\Secref}[1]{Section \ref{#1}}
\newcommand{\x}[1]{\lstinline{#1}}

\input{formaldefs.tex}
\begin{document}

\conferenceinfo{SIGBOVIK '13}{Pittsburgh, PA, USA}
\copyrightyear{2013}
\copyrightdata{[to be supplied]}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{
  The Randomly-Scoped Lambda Calculus
}
% \subtitle{Subtitle Text, if any}

\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}

%\authorinfo{
%  Nicholas D. Matsakis
%  \and Brian Anderson
%  \and Ben Blum \\
%  \and Tim Chevalier 
%  \and Graydon Hoare
%  \and Patrick Walton
%  \and David Herman}
%           {Mozilla Research}
%           {\{nmatsakis, banderson\}@mozilla.com,
%             bblum@cs.cmu.edu, \\
%             \{tchevalier, graydon, pwalton, dherman\}@mozilla.com}

\maketitle

\begin{abstract}

We are inspired by the following travesty, courtesy of Python:
\begin{verbatim}
>>> x
NameError: name 'x' is not defined
>>> [x for x in 1,2,3]
[1, 2, 3]
>>> x
3
>>>
\end{verbatim}

The question of {\em scope} pertains to the set of rules governing which values a program's variables refer to during evaluation.
Prior work has proposed two main approaches: {\em static} (or {\em lexical}) {\em scope}, in which variable bindings are resolved through source code analysis, and {\em dynamic scope}, in which bindings are resolved at run-time using a stack of activation records.

In this work, we present {\em random scope}, a new technique for scoping, as an alternative to the above two.
Random scope affords programmers the flexibility to refer to multiple different binding sites simultaneously; for example, applying the term $(\lambda x. (\lambda x. x))$ to two arguments could evaluate to either the first or the second one.
We ask what kind of evaluation semantics it would take to make this work, and then pretty much stumble onwards from there trying to make sense of the whole damn thing.

We define the Randomly-Scoped Lambda Calculus, and show how it can evaluate random natural numbers as well as more popular deterministic functionalities such as factorial and fibonacci.

\end{abstract}

% \category{D.3.3}{Language Constructs and Features}{Data types and structures}

\keywords
static scope, dynamic scope, lower is better

\section{Introduction}

\bibliographystyle{alpha}
\bibliography{paper}

\end{document}

### 2012-12-16 RSLC/paper/paper.tex
Many modern programming languages [cite cite cite cite cite cite]

@@
 \usepackage{graphics}
+\usepackage[utopia]{mathdesign}
 
@@
 
-We define the Randomly-Scoped Lambda Calculus, and show how it can evaluate random natural numbers as well as more popular deterministic functionalities such as factorial and fibonacci.
+We define the Randomly-Scoped Lambda Calculus, and show how it can be used to compute random natural numbers, while also providing more popular deterministic functionalities such as factorial and fibonacci.
 
@@
 
+Many modern programming languages~\cite{python,lambda,sml,GoLang,RustLang,mla,haskell98,patenttroll,wikiplia}
+
 \bibliographystyle{alpha}

### 2012-12-16 RSLC/paper/paper.tex
finish intro

--- 2012-12-15 RSLC/paper/paper.tex
@@
 %\documentclass[preprint]{sigplanconf}
-\documentclass{sigplanconf}
+\documentclass[10pt]{sigplanconf}
 
@@
 
+\usepackage{yfonts}
 \usepackage{amsmath}
@@
 \usepackage{graphics}
+\usepackage{marvosym}
+\usepackage[utopia]{mathdesign}
 
@@
 \copyrightyear{2013}
-\copyrightdata{[to be supplied]}
+\copyrightdata{}
 
@@
 \title{
-  The Randomly-Scoped Lambda Calculus
+Random Scope for Syntactic Hope
+% Learning to Cope with Random Scope
 }
+\subtitle{\em The Randomly-Scoped Lambda Calculus}
 % \subtitle{Subtitle Text, if any}
@@
 
-We are inspired by the following travesty, courtesy of Python:
+Gaze upon the following travesty, which Python hath wrought upon the world:
 \begin{verbatim}
@@
 
-In this work, we present {\em random scope}, a new technique for scoping, as an alternative to the above two.
+In this work, I present {\em random scope}, a new technique for scoping, as an alternative to the above two.
 Random scope affords programmers the flexibility to refer to multiple different binding sites simultaneously; for example, applying the term $(\lambda x.
@@
 x))$ to two arguments could evaluate to either the first or the second one.
-We ask what kind of evaluation semantics it would take to make this work, and then pretty much stumble onwards from there trying to make sense of the whole damn thing.
+I figure out what kind of evaluation semantics would make this work, and then pretty much stumble onwards from there trying to make sense of the whole damn thing.
 
-We define the Randomly-Scoped Lambda Calculus, and show how it can evaluate random natural numbers as well as more popular deterministic functionalities such as factorial and fibonacci.
+I define the Randomly-Scoped Lambda Calculus, and show how it can be used to compute random natural numbers, while also providing more popular deterministic functionalities such as factorial and fibonacci.
 
@@
 
+Some modern programming languages~\cite{wikiplia,python,lambda,sml,GoLang,RustLang,mla,haskell98,patenttroll} include a mechanism to refer to previously-computed values using more consise names, typically called {\em variables}.
+The question then arises: ``Which values should my variables refer to?''
+In a half-hearted attempt to answer, these languages implement {\em scope} (which I explained in the abstract; go read it), and hope it all works out okay.
+
+Whether resolved statically or dynamically, modern scoping approaches lack the flexibility to refer to multiple binding sites simultaneously.
+I present the Randomly-Scoped Lambda Calculus (RSLC), in which references to shadowed variables can evaluate to any of their binding sites with equal probability.
+
+Consider the following lambda calculus term:
+\[ (\lambda x. (\lambda x. x))~A~B \]
+In ordinary lambda calculus, this always evaluates to $B$ (quite unforgivingly so, in my opinion).
+However, I give the programmer the benefit of the doubt, and assume they named the first argument $x$ as well because they wanted it to have equal opportunity.
+Hence, in RSLC this term can evaluate to either $A$ or $B$.
+(On the other side of the coin, RSLC encourages good programming practice: a programmer who {\em doesn't} want $x$ to evaluate to $A$ should name the first argument something different, to make their intentions more clear.)
+
+In this paper I make the following contributions:
+\begin{enumerate}
+	\item The {\bf Randomly-Scoped Lambda Calculus (RSLC)}, a logical programming language that uses a novel syntactic technique called {\bf random scoping},
+	\item An analysis of two implementation strategies for random scoping, which I call {\bf static random scope (SRS)} and {\bf dynamic random scope (DRS)},
+	\item A new schema for general recursion using DRS (because the Y combinator doesn't work anymore),
+	\item Programming examples in RSLC that employ random scope to compute random natural numbers,
+	\item Programming examples in RSLC that avoid random scope to implement deterministic arithmetic.
+\end{enumerate}
+
+\section{Language Definition}
+
+\begin{figure}[t]
+	\Grammar
+	\caption{RSLC formal grammar.}
+	\label{fig:grammar}
+\end{figure}
+
+RSLC is an extention to the call-by-name un(i)typed lambda calculus with slightly modified substitution rules.
+In the grammar (Figure~\ref{fig:grammar}), a variable $x$ carries around a list of expressions $\rep{e}$ which have been ``tried to have been substituted for $x$ before''.
+When programming in RSLC, we simply write $x$ (or $y$ or \Neptune~or whatever way you swing), because no substitution has happened yet, but after $n$ ``$\lambda x$''s have been $\beta$-reduced, the $\rep{e}$ list will have $n$ expressions in it.
+It might help to think of it like quantum superposition, I guess? I dunno.
+
+Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
+
 \bibliographystyle{alpha}

### 2012-12-16 RSLC/paper/paper.tex
add subst_sully

@@
 	\item The {\bf Randomly-Scoped Lambda Calculus (RSLC)}, a logical programming language that uses a novel syntactic technique called {\bf random scoping},
-	\item An analysis of two implementation strategies for random scoping, which I call {\bf static random scope (SRS)} and {\bf dynamic random scope (DRS)},
-	\item A new schema for general recursion using DRS (because the Y combinator doesn't work anymore),
+	\item A new schema for general recursion in RSLC (because the Y combinator doesn't work anymore),
 	\item Programming examples in RSLC that employ random scope to compute random natural numbers,
@@
 
-RSLC is an extention to the call-by-name un(i)typed lambda calculus with slightly modified substitution rules.
-In the grammar (Figure~\ref{fig:grammar}), a variable $x$ carries around a list of expressions $\rep{e}$ which have been ``tried to have been substituted for $x$ before''.
-When programming in RSLC, we simply write $x$ (or $y$ or \Neptune~or whatever way you swing), because no substitution has happened yet, but after $n$ ``$\lambda x$''s have been $\beta$-reduced, the $\rep{e}$ list will have $n$ expressions in it.
-It might help to think of it like quantum superposition, I guess? I dunno.
+RSLC is an extention to the lazy un(i)typed lambda calculus with slightly modified substitution rules.
+In the grammar (Figure~\ref{fig:grammar}), a variable $x$ carries around a list of expressions $\rep{e}$ which we have ``tried to substitute for it before''.
+When programming in RSLC, we simply write $x$ (or $y$ or \Neptune~or however you happen to swing), because no substitution has happened yet, but after $n$ ``$\lambda x$''s have been $\beta$-reduced, the $\rep{e}$ list will have $n$ expressions in it.
+It might help to think of it like quantum superposition, I guess?
 
 Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
+In the normal $\lambda$-calculus, in the {\sc Subst-Capture} rule, substitution stops when a new binding's name is the same as the one being substituted for, but in RSLC
 

### 2012-12-16 RSLC/RSLC.hs

@@ -40,7 +40,9 @@ instance Show Expr where
     show (Let name e1 e2) =
         "(let " ++ name ++ " = " ++ show e1 ++ " in " ++ show e2 ++ ")"
 
-----
+--------------------------------
+-- Ben's substitution scheme. --
+--------------------------------
 
 -- Called when capturing under another lambda.
 subst_shadow :: Expr -> String -> Expr -> Expr
@@ -94,6 +96,42 @@ subst_pull e0 name (Let name2 e1 e2) =
                   subst_pull e0 name e2
        return $ Let name2 e1' e2'
 
+----------------------------------
+-- Sully's substitution scheme. --
+----------------------------------
+
+-- (in this one we ignore the "es" and just carry around a probability.)
+
+subst_sully :: Int -> Expr -> String -> Expr -> IO Expr
+subst_sully i e0 name (Var name2 es) =
+    if name == name2 then
+        do i <- getStdRandom $ randomR (0,i)
+           return $ if i == 0 then e0 else Var name2 es
+    else return $ Var name2 es
+subst_sully i e0 name (App e1 e2) =
+    do e1' <- subst_sully i e0 name e1
+       e2' <- subst_sully i e0 name e2
+       return $ App e1' e2'
+subst_sully i e0 name (Lam name2 e) =
+    Lam name2 <$> subst_sully (if name == name2 then i+1 else i) e0 name e
+subst_sully i e0 name Zero = return Zero
+subst_sully i e0 name (Suc e) = Suc <$> subst_sully i e0 name e
+subst_sully i e0 name (Natrec e e1 name2 e2) =
+    do e'  <- subst_sully i e0 name e
+       e1' <- subst_sully i e0 name e1
+       e2' <- subst_sully (if name == name2 then i+1 else i) e0 name e2
+       return $ Natrec e' e1' name2 e2'
+subst_sully i e0 name (Let name2 e1 e2) =
+    do e1' <- subst_sully i e0 name e1
+       e2' <- subst_sully (if name == name2 then i+1 else i) e0 name e2
+       return $ Let name2 e1' e2'
+
+-- subst_sully seems to have big issues with multiplication. Stack overflows
+-- for fact(4), swapping to disk on the 3x3 multiplication table. Otherwise
+-- behaves exactly the same.
+-- subst = subst_sully 0
[…]

### 2012-12-19 RSLC/RSLC.hs
clean up a bunch

@@ -49,8 +49,8 @@ subst_shadow :: Expr -> String -> Expr -> Expr
 subst_shadow e0 name (Var name2 es) =
     if name == name2 then
         -- trace ("Adding " ++ show e0 ++ " to " ++ name ++ show es) $
-        Var name2 (e0:es) -- XXX: need to map here, as below?
-    else Var name2 es -- $ map (subst_shadow e0 name) es
+        Var name2 (e0:map (subst_shadow e0 name) es)
+    else Var name2 $ map (subst_shadow e0 name) es
 subst_shadow e0 name (App e1 e2) = App (subst_shadow e0 name e1) (subst_shadow e0 name e2)
 subst_shadow e0 name (Lam name2 e) = Lam name2 $ subst_shadow e0 name e -- capture-seeking substitution
 subst_shadow e0 name Zero = Zero
@@ -65,10 +65,11 @@ subst_shadow e0 name (Let name2 e1 e2) =
 subst_pull :: Expr -> String -> Expr -> IO Expr
 subst_pull e0 name (Var name2 es) =
     if name == name2 then
-        let list = e0:es
-        in do i <- getStdRandom $ randomR (0, length list - 1)
-              return $ list !! i
-    else return $ Var name2 es
+        do es' <- mapM (subst_pull e0 name) es
+           let list = e0:es'
+           i <- getStdRandom $ randomR (0, length list - 1)
+           return $ list !! i
+    else Var name2 <$> mapM (subst_pull e0 name) es
 subst_pull e0 name (App e1 e2) =
     do e1' <- subst_pull e0 name e1
        e2' <- subst_pull e0 name e2
@@ -126,9 +127,7 @@ subst_sully i e0 name (Let name2 e1 e2) =
        e2' <- subst_sully (if name == name2 then i+1 else i) e0 name e2
        return $ Let name2 e1' e2'
 
--- subst_sully seems to have big issues with multiplication. Stack overflows
--- for fact(4), swapping to disk on the 3x3 multiplication table. Otherwise
--- behaves exactly the same.
+-- subst_sully and subst_pull are equivalent, I think.
 -- subst = subst_sully 0
 subst = subst_pull
 
@@ -149,7 +148,6 @@ eval Zero = return Zero
 eval (Suc e) = Suc <$> eval e
 eval (Natrec e e1 name e2) =
     do e' <- eval e
-       -- trace ("Eval'ed 'case " ++ show e ++ "' ...to... " ++ show e') $ return ()
        case e' of Zero -> eval e1
[…]

### 2012-12-19 RSLC/RSLC.hs
Add dice and test out eager eval-app

@@ -140,6 +140,7 @@ eval (Var name es) =
     --    eval $ es !! i
 eval (App e1 e2) =
     do e1' <- eval e1
+       -- e2' <- eval e2 -- CBV works about the same as CBN + eager let.
        case e1' of
            Lam name e1'' -> do e1''' <- subst e2 name e1''; eval e1'''
            _ -> error ("fuck" ++ show (App e1' e2))
@@ -267,7 +268,7 @@ times = -- \mn. case m of 0 => 0 | s(m2) => add n (times m2 n)
                 Natrec (var m) (Lam myself $ Zero)
                        m2 (Lam myself $
                                Let p (y2 "times" $$ var m2 $$ var n $$ var myself) $
-                               add $$ var p $$ var n)
+                               add $$ var n $$ var p)
     in Lam arg1 $ Lam arg2 $ y2 "times" $$ var arg1 $$ var arg2 $$ times_inner
 
 fact = -- \n. case n of 0 => 1 | s(n2) => times n (fact n2)
@@ -281,7 +282,7 @@ fact = -- \n. case n of 0 => 1 | s(n2) => times n (fact n2)
             Lam n $ Natrec (var n) (Lam myself $ Suc Zero)
                            n2 (Lam myself $
                                    Let p (y1 "fact" $$ var n2 $$ var myself) $
-                                   times $$ var p $$ var n)
+                                   times $$ var n $$ var p)
     in Lam arg $ y1 "fact" $$ var arg $$ fact_inner
 
 fib = -- \n. case n of 0 => (0,1) | S(n2) => (\x. (snd x, fst x + snd x))(fib n2)
@@ -306,12 +307,24 @@ fib = -- \n. case n of 0 => (0,1) | S(n2) => (\x. (snd x, fst x + snd x))(fib n2
                             pair b (var n_snd) (add $$ var n_fst $$ var n_snd))
     in Lam arg $ first (y1 "fib" $$ var arg $$ fib_inner)
 
+-- Dice!
+
+coin = rand $$ tonat 1
+d6   = add $$ tonat 1 $$ (rand $$ tonat 5)
+d6d6 = let x = "d6_1"
+           y = "d6_2"
+       in Let x (add $$ (tonat 1) $$ (rand $$ tonat 5)) $
+          Let y (add $$ (tonat 1) $$ (rand $$ tonat 5)) $
[…]

### 2012-12-19 RSLC/paper/paper.tex
Add eval and subst rules to paper

@@
 \usepackage{marvosym}
+\usepackage{stmaryrd}
+\usepackage[usenames,dvipsnames]{xcolor}
 \usepackage[utopia]{mathdesign}
@@
 \title{
-Random Scope for Syntactic Hope
+% Random Scope for Syntactic Hope
 % Learning to Cope with Random Scope
+Randomly-Scoped Lambda Calculus
 }
-\subtitle{\em The Randomly-Scoped Lambda Calculus}
+% \subtitle{\em The Randomly-Scoped Lambda Calculus}
 % \subtitle{Subtitle Text, if any}
@@
 	\item The {\bf Randomly-Scoped Lambda Calculus (RSLC)}, a logical programming language that uses a novel syntactic technique called {\bf random scoping},
+	%\item An analysis of two implementation strategies for random scoping, which I call {\bf static random scope (SRS)} and {\bf dynamic random scope (DRS)},
 	\item A new schema for general recursion in RSLC (because the Y combinator doesn't work anymore),
@@
 Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
-In the normal $\lambda$-calculus, in the {\sc Subst-Capture} rule, substitution stops when a new binding's name is the same as the one being substituted for, but in RSLC
+In the normal $\lambda$-calculus, in the {\sc Subst-Capture} rule, substitution stops when a new binding's name is the same as the one being substituted for, but in RSLC we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var} rule).
+Similarly, when substituting for a variable before switching to {\sc Capture} mode (in the {\sc Subst-Var} rule), we ... hmm, why is two-moded substitution needed with DRS anyway?
+
+\begin{figure}[t]
+	\fbox{\begin{tabular}{ll}$e \eval e'$ \\ $e~\Val$\end{tabular}}
+	\TypeRules{
+		\EvalLam \qquad \EvalApp \qquad \EvalVar \\
+		\EvalZero \qquad \EvalSucA \qquad \EvalSucB \\
+		\EvalCaseA \\ \EvalCaseB \\ \EvalLet
+	}
+	\caption{RSLC small-step evaluation rules. They're exactly the same as for any other call-by-name $\lambda$-calculus, so don't bother reading them. Why am I even including them?}
+	\label{fig:eval}
+\end{figure}
+
+\begin{figure}[t]
+	\fbox{$\subst{e_0}{x}{e}$}
+	\TypeRules{
+		\SubstLamCapture \qquad \SubstLam \\
+		\SubstVarX \qquad \SubstVarY \\
+	}
+	\fbox{$\capture{e_0}{x}{e}$}
+	\TypeRules{
+		\CaptureVarX \qquad \CaptureVarY \\
+	}
+	\caption{RSLC substitution rules. The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice. (The complete set of rules is shown in Figure~\ref{fig:subst-complete}.)}
+	\label{fig:subst}
+\end{figure}
+
+\appendix
+
+\begin{figure*}[t]
+	\fbox{$\subst{e_0}{x}{e}$}
+	\TypeRules{
+		\SubstLamCapture \qquad \SubstLam \qquad
+		\SubstApp \\
+		\SubstVarX \qquad \SubstVarY \\
+		\SubstCaseCapture \\ \SubstCase \\
+		\SubstZero \qquad \SubstSuc \\
+		\SubstLetCapture \qquad \SubstLet
+	}
+	\fbox{$\capture{e_0}{x}{e}$}
+	\TypeRules{
+		\CaptureLam \qquad \CaptureApp \qquad
+		\CaptureVarX \qquad \CaptureVarY \\
+		\CaptureCase \\
+		\CaptureZero \qquad \CaptureSuc \qquad \CaptureLet
+	}
+	\caption{Complete RSLC substitution and capture rules.}
+	\label{fig:subst-complete}
+\end{figure*}
 

### 2012-12-19 RSLC/paper/paper.tex
write more

@@
 \usepackage{graphics}
+\usepackage{graphicx}
 \usepackage{marvosym}
 \usepackage{stmaryrd}
+\usepackage{epsdice}
 \usepackage[usenames,dvipsnames]{xcolor}
@@
 Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
-In the normal $\lambda$-calculus, in the {\sc Subst-Capture} rule, substitution stops when a new binding's name is the same as the one being substituted for, but in RSLC we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var} rule).
-Similarly, when substituting for a variable before switching to {\sc Capture} mode (in the {\sc Subst-Var} rule), we ... hmm, why is two-moded substitution needed with DRS anyway?
+In the normal $\lambda$-calculus, substitution would stop when a new binding's name is the same as the one being substituted for.
+
+However, in RSLC, as shown in {\sc Subst-Lam-Capture}, we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var-X} rule).
+Similarly, when substituting for $x$ before switching to {\sc Capture} mode (in the {\sc Subst-Var-X} rule), we nondeterministically select an expression from $x$'s $\rep{e}$ list to replace $x$ with (with uniform randomness).
 
@@
 	}
-	\caption{RSLC substitution rules. The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice. (The complete set of rules is shown in Figure~\ref{fig:subst-complete}.)}
+	\caption{RSLC substitution rules.
+The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice.
+(Only important rules are shown here; Figure~\ref{fig:subst-complete} shows the complete set.)}
 	\label{fig:subst}
@@
 
+\subsection{Evaluation Example}
+
+As an example, I present the steps for evaluating the $(\lambda x. (\lambda x. x))~A~B$ from the introduction:
+
+\begin{eqnarray*}
+	& & (\lambda x. (\lambda x. x[]))~A~B \\
+	& \mapsto & (\subst{A}{x} (\lambda x. x[])) B \\
+	& = & (\lambda x. \capture{A}{x} x[]) B \\
+	& = & (\lambda x. x[A]) B \\
+	& \mapsto & \subst{B}{x} x[A] \\
+	& = & \text{\em (any element of the list $[B,A]$)}
+\end{eqnarray*}
+
+\section{Random Programming with RSLC}
+
+Already we have a system with which we can implement simple random number generators, such as a simple 6-sided die:
+\[
+(\lambda x.
+(\lambda x.
+(\lambda x.
+(\lambda x.
+(\lambda x.
+(\lambda x. x))))))
+~\epsdice{1}
+~\epsdice{2}
+~\epsdice{3}
+~\epsdice{4}
+~\epsdice{5}
+~\epsdice{6}
+\]
+
+However,
+
+\newpage
+
+\section{Evaluation}
+
+We computed some random numbers woo
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{rslc-rdr.pdf}
+	\end{center}
+	\caption{We show that RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
+	\label{fig:rdr}
+\end{figure}
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.40\textwidth]{rslc-d20.pdf}
+	\end{center}
+	\caption{RSLC can also be used to implement popular role-playing games.}
+	\label{fig:rdr}
+\end{figure}
+
+\bibliographystyle{alpha}
+\bibliography{paper}
+
 \appendix
 
+\section{Discussion of Substitution Rules}
 \begin{figure*}[t]
@@
 
-\bibliographystyle{alpha}
-\bibliography{paper}
-
 \end{document}

### 2012-12-20 RSLC/paper/paper.tex
write a bunch more

@@
 \usepackage{epsdice}
+\usepackage{multirow}
 \usepackage[usenames,dvipsnames]{xcolor}
@@
 
-However, in RSLC, as shown in {\sc Subst-Lam-Capture}, we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var-X} rule).
+However, in RSLC, as shown in {\sc Subst-Capture}, we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var-X} rule).
 Similarly, when substituting for $x$ before switching to {\sc Capture} mode (in the {\sc Subst-Var-X} rule), we nondeterministically select an expression from $x$'s $\rep{e}$ list to replace $x$ with (with uniform randomness).
@@
 	\TypeRules{
-		\SubstLamCapture \qquad \SubstLam \\
+		\SubstCapture \qquad \SubstLam \\
 		\SubstVarX \qquad \SubstVarY \\
@@
 	}
-	\caption{RSLC substitution rules.
-The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice.
-(Only important rules are shown here; Figure~\ref{fig:subst-complete} shows the complete set.)}
+	\caption{Selected RSLC substitution rules. The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice.}
 	\label{fig:subst}
@@
 
-As an example, I present the steps for evaluating the $(\lambda x. (\lambda x. x))~A~B$ from the introduction:
+To demonstrate, here's the evaluation of $(\lambda x. (\lambda x. x))~A~B$ from the introduction:
 
-\begin{eqnarray*}
-	& & (\lambda x. (\lambda x. x[]))~A~B \\
-	& \mapsto & (\subst{A}{x} (\lambda x. x[])) B \\
-	& = & (\lambda x. \capture{A}{x} x[]) B \\
-	& = & (\lambda x. x[A]) B \\
-	& \mapsto & \subst{B}{x} x[A] \\
-	& = & \text{\em (any element of the list $[B,A]$)}
-\end{eqnarray*}
+\begin{center}
+\begin{tabular}{cll}
+	& $(\lambda x. (\lambda x. x[]))~A~B$ & \\
+	$\mapsto$ & $(\subst{A}{x} (\lambda x. x[]))~B$ & {\sc \small Eval-App} \\
+	\substeq & $(\lambda x. \capture{A}{x} x[])~B$ & {\sc \small Subst-Capture} \\
+	\captureeq & $(\lambda x. x[A])~B$ & {\sc \small Capture-Var-X} \\
+	$\mapsto$ & $\subst{B}{x} x[A]$ & {\sc \small Eval-App} \\
+	\substeq & {\em (any element of the list $[B,A]$)} & {\sc \small Subst-Var-X}
+\end{tabular}
+\end{center}
+Now the fairness to $A$ we wanted in the intro is restored.
 
@@
 
-Already we have a system with which we can implement simple random number generators, such as a simple 6-sided die:
+\newcommand{\Y}{\textsf{Y}}
+\newcommand{\Ya}{\ensuremath{\mathsf{Y_1}}}
+\newcommand{\Yab}{\ensuremath{\mathsf{Y_2}}}
+\newcommand{\Yabc}{\ensuremath{\mathsf{Y_3}}}
+
+Already we have a system with which we can implement simple random number generators, such as a simple 6-sided die:\footnote{Really I just wanted an excuse to typeset \LaTeX~dice.}
 \[
@@
 
-However,
+However, this isn't really ``programming'', and Sully thought for sure I wouldn't be able to do anything useful with this language, which of course I wasn't going to stand by.
+The first thing I wanted to do was to recursively generate a random natural number, something like this:
+\begin{equation}
+\label{eq:randomnat}
+\Y~(\lambda f. \lambda n. (\lambda x. \lambda x. x)~n~(f~\Suc{n}))~\Zero
+\end{equation}
+This term would randomly choose whether to output $n$, its argument, or to call itself with $\Suc{n}$.
+However, the conventional Y-combinator doesn't work as intended in RSLC, because once a function is substituted into itself for its first argument, subsequent arguments will be captured in the substituted version.
+Even ignoring the internal workings of the combinator, and supposing $\Y$ satisfies $\Y~g \mapsto^* g~(\Y~g)$, here's what happens:
 
-\newpage
+\begin{center}
+\begin{tabular}{ll}
+\hspace{-1em}
+& $\Y~(\lambda f. \lambda n. (\lambda x. \lambda x. x)~n~(f~\Suc{n}))~\Zero$ \\
+\hspace{-1em}
+$\mapsto^*$ & $(\lambda n. (\lambda x. \lambda x. x)~n~(\Y~(\lambda f.\lambda n. \dots n \dots)~\Suc{n}))~\Zero$ \\
+\hspace{-1em}
+$\mapsto$ & $\subst{\Zero}{n}((\lambda x. \lambda x. x)~n~(\Y~(\lambda f.\lambda n. \dots n \dots)~\Suc{n}))$ \\
+\hspace{-1em}
+= & $((\lambda x. \lambda x. x)~\subst{\Zero}{n}n~(\Y~(\lambda f.\lambda n. \dots \capture{\Zero}{n}n \dots)~\Suc{\subst{\Zero}{n}n}))$ \\
+\end{tabular}
+\end{center}
+
+But wait -- I only wanted random-capture on $x$, not on $n$!
+As a result, though we intended $0$ to be output with half probability, and $1$ with quarter probability, and so on, actually $0$ gets output substantially more often (5/8, I think).
+
+\subsection{The new Y combinator(s)}
+In order to properly recurse in RSLC, we need a new convention in which recursive functions get themselves substituted into them {\em last}, after all other arguments have been applied.
+The desired identity is $\Ya~g~a \mapsto^* g~a~(\Ya g)$ for one-argument functions (and $\Yab~g~a~b \mapsto^* g~a~b~(\Yab~g)$ for two-argument functions, and so on).
+
+Unfortunately, I wasn't even able to find an RSLC term that satisfied that identity.
+I had to also alter the way recursive functions call themselves -- by referring to the combinator itself, passing it the recursive arguments, and {\em then} the substituted version of themselves.
+Since that makes no sense in words, just have a look:
+\begin{eqnarray*}
+\Ya &=& \lambda ag. gag \\
+\Yab &=& \lambda abg. gabg \\
+\Yabc &=& \lambda abcg. gabcg \\
+&\dots&
+\end{eqnarray*}
+Here's a sort of handwavy syntactic transformation example for $\mathsf{fix}$ constructs that supports one, two, etc arguments:
+\begin{eqnarray*}
+\mathsf{fix}_1(f,a.e) &=& \lambda a. \Ya~a~(\lambda af. [^{(\Ya~x~f)}/_{f(x)}]e) \\
+\mathsf{fix}_2(f,a,b.e) &=& \lambda ab. \Ya~a~b~(\lambda abf. [^{(\Ya~x~y~f)}/_{f(x)(y)}]e) \\
+&\dots&
+\end{eqnarray*}
+
+Note that this transformation doesn't just substitute for the variable $f$ in $e$ but also reorders the arguments it's applied to. So now the term from Equation~\ref{eq:randomnat} becomes:
+\begin{equation}
+\label{eq:randomnat-fixed}
+\Ya~\Zero~(\lambda n. \lambda f. (\lambda x. \lambda x. x)~n~(\Ya~\Suc{n}~f))
+\end{equation}
+
+And evaluates as intended:
+
+\begin{center}
+\begin{tabular}{ll}
+$\mapsto^*$ & $(\lambda n. \lambda f. (\lambda x. \lambda x. x)~n~(\Ya~\Suc{n}~f))~\Zero~(\lambda n. \lambda f. \dots)$ \\
+$\mapsto$ & $(\subst{\Zero}{n}\lambda f. (\lambda x. \lambda x. x)~n~(\Ya~\Suc{n}~f))~(\lambda n. \lambda f. \dots)$ \\
+= & $(\lambda f. (\lambda x. \lambda x. x)~\Zero~(\Ya~\Suc{\Zero}~f))~(\lambda n. \lambda f. \dots)$ \\
+$\mapsto$ & $(\lambda x. \lambda x. x)~\Zero~(\Ya~\Suc{\Zero}~(\lambda n. \lambda f. \dots))$
+\end{tabular}
+\end{center}
+
+Figure~\ref{fig:nats} shows the experimental results of evaluating this term 10,000 times.
+
+\begin{figure}[h!]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{rslc-naturals.pdf}
+	\end{center}
+	\caption{Computing random numbers with the term in Equation~\ref{eq:randomnat-fixed}. Log scale, lower is better. (I couldn't show $13$ because it showed up 0 times, and $\mathsf{log}~0 = -\infty$.)}
+	\label{fig:nats}
+\end{figure}
+
+\subsection{Uniform random number generation}
+
+\newcommand{\thing}[1]{\textcolor{Fuchsia}{#1}}
+
+The next thing I wanted to do was compute random numbers with a uniform distribution.
+Previously we were just selecting between two possible bindings for $x$, statically defined in the program text.
+I wondered, ``can I dynamically generate a term with $n$ shadow-bindings\footnote{Shadow binding: Sor/Wiz 3, Complete Arcane.
+Will save negates.} that causes a variable to have $n$ expressions in its $\rep{e}$ list?'' Something like:
+\[
+\thing{(\lambda x. (\lambda x. \dots (\lambda x.x)))}~0~1~2~\dots~n
+\]
+Of course this was where I had to iron out all the crazy parts in the previous sections, but since I already presented it in a manner consistent with it ain't being no thang, I'm going to make this part look easy:
+% Y2 (\x.x) N (\tn. case n of 0 => (\f. t 0) | S(n2) => (\f. Y2 ((\x.t)n) n2 f))
+
+\begin{center}
+\begin{tabular}{rcl}
+\hspace{-1em}
+$\rho$ & = &
+\hspace{-1em}\multirow{4}{*}{
+\begin{tabular}{lllll}
+\multicolumn{5}{l}{$\lambda \thing{\bf t}. \lambda n.$} \\
+& \multicolumn{4}{l}{$\mathsf{case}~n~\mathsf{of}$} \\
+&& $\Zero$ & $\Rightarrow$ & $\lambda f.~\thing{\bf t}~\Zero$ \\
+&& $\Suc{n_2}$ &$\Rightarrow$& $\lambda f.~\Yab~\thing{((\lambda x.{\bf t})~n)}~n_2~f$
+\end{tabular}
+}
+\\
+&&\\
+&&\\
+&&\\
+&&\\
+\hspace{-1em}
+$\mathsf{rand}[0-N]$ &=& $\Yab~\thing{(\lambda x.x)}~N~\rho$
+\end{tabular}
+\end{center}
+
+The variable $\thing{\bf t}$ accumulates ``$\lambda x$''s as $\rho$ recurses (in the $\Suc{n_2}$ case).
+You might ask: why does the $\lambda f$ need to be inside the $\mathsf{case}$ statement?
+Because otherwise $n_2$ would get shadowed in the thing substituted for $f$.
+
+A brief, very condensed, evaluation example:
+
+\begin{center}
+\begin{tabular}{ll}
+& $\mathsf{rand}[0-2]$ \\
+= & $\Yab~\thing{(\lambda x.x)}~2~\rho$ \\
+$\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.x))~2)}~1~\rho$ \\
+$\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.(\lambda x.x))~2)~1)}~0~\rho$ \\
+$\mapsto^*$ & $((\lambda x.(\lambda x.(\lambda x.x))~2)~1)~0$ \\
+$\mapsto^*$ & $\subst{2}{x}x[1,0]$
+\end{tabular}
+\end{center}
+
+\section{Deterministic Programming with RSLC}
 

### 2012-12-20 RSLC/paper/paper.tex
write more

@@
 
-\begin{figure}[h!]
+\begin{figure}[h]
 	\begin{center}
@@
 \begin{tabular}{rcl}
-\hspace{-1em}
 $\rho$ & = &
@@
 &&\\
-\hspace{-1em}
-$\mathsf{rand}[0-N]$ &=& $\Yab~\thing{(\lambda x.x)}~N~\rho$
+$\mathsf{rand}[0,N]$ &=& $\Yab~\thing{(\lambda x.x)}~N~\rho$
 \end{tabular}
@@
 
-The variable $\thing{\bf t}$ accumulates ``$\lambda x$''s as $\rho$ recurses (in the $\Suc{n_2}$ case).
-You might ask: why does the $\lambda f$ need to be inside the $\mathsf{case}$ statement?
-Because otherwise $n_2$ would get shadowed in the thing substituted for $f$.
+The variable $\thing{\bf t}$ accumulates ``$\lambda x$''s as $\rho$ recurses (in the $\Suc{n_2}$ case). You might ask: why does the $\lambda f$ need to be inside the $\mathsf{case}$ statement?
+Because otherwise in {\sc Eval-Case-2}$n_2$ would get captured in the thing substituted for $f$.
 
-A brief, very condensed, evaluation example:
+A brief, very condensed, evaluation run-through:
 
 \begin{center}
-\begin{tabular}{ll}
-& $\mathsf{rand}[0-2]$ \\
+\begin{tabular}{lll}
+$\mathsf{rand}[0,2]$ &
 = & $\Yab~\thing{(\lambda x.x)}~2~\rho$ \\
-$\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.x))~2)}~1~\rho$ \\
-$\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.(\lambda x.x))~2)~1)}~0~\rho$ \\
-$\mapsto^*$ & $((\lambda x.(\lambda x.(\lambda x.x))~2)~1)~0$ \\
-$\mapsto^*$ & $\subst{2}{x}x[1,0]$
+& $\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.x))~2)}~1~\rho$ \\
+& $\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.(\lambda x.x))~2)~1)}~0~\rho$ \\
+& $\mapsto^*$ & $((\lambda x.(\lambda x.(\lambda x.x))~2)~1)~0$ \\
+& $\mapsto^*$ & $\subst{0}{x}x[1,2]$
 \end{tabular}
@@
 \section{Deterministic Programming with RSLC}
+
+OK, great; got here.
+Up until now I showed how random scope adds inherent nondeterminism that can be harnessed to simulate dice anda make silly graphs.
+The other thing Sully didn't believe would be possible was writing deterministic RSLC programs that {\em avoid} using ``the feature''.
+
+In the last section I already introduced the capture-free Y combinator(s). Implementing addition using that is no trouble:
+
+\newcommand{\addition}{\maltese}
+\begin{center}
+\begin{tabular}{rcl}
+$\addition$ & = &
+\hspace{-1em}\multirow{4}{*}{
+\begin{tabular}{lllll}
+\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
+& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
+&& $\Zero$ & $\Rightarrow$ & $\lambda f.~n$ \\
+&& $\Suc{m_2}$ &$\Rightarrow$& $\lambda f.~\Suc{\Yab~m_2~n~f}$
+\end{tabular}
+}
+\\
+&&\\
+&&\\
+&&\\
+&&\\
+$\mathsf{add}~M~N$ &=& $\Yab~M~N~\addition$
+\end{tabular}
+\end{center}
+
+But it turns out that multiplication, which uses addition, winds up needing the eager \textsf{let} construct.
+Among the several ways to implement it fully lazily, some introduce accidental randomness by letting some of \textsf{times}'s variables get captured, and others simply cause my implementation to run out of stack space or start swapping to disk when evaluating.
+Uh, anyway, here it is.
+
+\newcommand{\multiplication}{\bigstar}
+\begin{center}
+\begin{tabular}{rcl}
+$\multiplication$ & = &
+\hspace{-1em}\multirow{6}{*}{
+\begin{tabular}{lllll}
+\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
+& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
+&& $\Zero$ & $\Rightarrow$ & $\lambda f.~\Zero$ \\
+&& $\Suc{m_2}$ &$\Rightarrow$& $\lambda f.$ \\
+&&
+\multicolumn{3}{l}{
+\begin{tabular}{lll}
+& \textsf{let} & $x~=~\Yab~m_2~n~f$ \\
+& \textsf{in} & $\mathsf{add}~x~n$
+\end{tabular}
+}
+\end{tabular}
+}
+\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+$\mathsf{times}~M~N$ &=& $\Yab~M~N~\multiplication$
+\end{tabular}
+\end{center}
 

### 2012-12-20 RSLC/paper/paper.tex
write sect. 4

@@
 \begin{eqnarray*}
-\Ya &=& \lambda ag. gag \\
-\Yab &=& \lambda abg. gabg \\
-\Yabc &=& \lambda abcg. gabcg \\
+\Ya &=& \lambda ag.~gag \\
+\Yab &=& \lambda abg.~gabg \\
+\Yabc &=& \lambda abcg.~gabcg \\
 &\dots&
@@
 && $\Zero$ & $\Rightarrow$ & $\lambda f.~n$ \\
-&& $\Suc{m_2}$ &$\Rightarrow$& $\lambda f.~\Suc{\Yab~m_2~n~f}$
+&& $\Suc{m'}$ &$\Rightarrow$& $\lambda f.~\Suc{\Yab~m'~n~f}$
 \end{tabular}
@@
 \newcommand{\multiplication}{\bigstar}
+\newcommand{\multx}{\hat{x}}
+\newcommand{\multm}{\hat{m}}
+\newcommand{\multn}{\hat{n}}
+\newcommand{\multf}{\hat{f}}
+\newcommand{\multmm}{\hat{m}'}
 \begin{center}
@@
 \begin{tabular}{lllll}
-\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
-& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
-&& $\Zero$ & $\Rightarrow$ & $\lambda f.~\Zero$ \\
-&& $\Suc{m_2}$ &$\Rightarrow$& $\lambda f.$ \\
+\multicolumn{5}{l}{$\lambda \multm. \lambda \multn.$} \\
+& \multicolumn{4}{l}{$\mathsf{case}~\multm~\mathsf{of}$} \\
+&& $\Zero$ & $\Rightarrow$ & $\lambda \multf.~\Zero$ \\
+&& $\Suc{\multmm}$ &$\Rightarrow$& $\lambda \multf.$ \\
 &&
@@
 \begin{tabular}{lll}
-& \textsf{let} & $x~=~\Yab~m_2~n~f$ \\
-& \textsf{in} & $\mathsf{add}~x~n$
+& \textsf{let} & $\multx~=~\Yab~\multmm~\multn~\multf$ \\
+& \textsf{in} & $\mathsf{add}~\multx~\multn$
 \end{tabular}
@@
 \end{center}
+Note that since $\mathsf{times}$ refers to $\mathsf{add}$, $\mathsf{times}$'s variable names must have different names. I put hats on them.
 
-\section{Evaluation}
+Here's the factorial function:
+\newcommand{\factorial}{\textbf{!}}
+\newcommand{\factx}{\mathring{x}}
+\newcommand{\factn}{\mathring{n}}
+\newcommand{\factf}{\mathring{f}}
+\newcommand{\factnn}{\mathring{n}'}
+\begin{center}
+\begin{tabular}{rcl}
+$\factorial$ & = &
+\hspace{-1em}\multirow{6}{*}{
+\begin{tabular}{lllll}
+\multicolumn{5}{l}{$\lambda \factn.$} \\
+& \multicolumn{4}{l}{$\mathsf{case}~\factn~\mathsf{of}$} \\
+&& $\Zero$ & $\Rightarrow$ & $\lambda \factf.~\Suc{\Zero}$ \\
+&& $\Suc{\factnn}$ &$\Rightarrow$& $\lambda \factf.$ \\
+&&
+\multicolumn{3}{l}{
+\begin{tabular}{lll}
+& \textsf{let} & $\factx~=~\Ya~\factnn~\factf$ \\
+& \textsf{in} & $\mathsf{times}~\factn~\factx$
+\end{tabular}
+}
+\end{tabular}
+}
+\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+$\mathsf{fact}~N$ &=& $\Ya~N~\factorial$
+\end{tabular}
+\end{center}
 
-We computed some random numbers woo
+\newcommand{\fst}{\mathsf{fst}}
+\newcommand{\snd}{\mathsf{snd}}
+To compute fibonacci numbers, I encoded pairs:
+\begin{eqnarray*}
+(e_1, e_2) &=& \lambda b.~b~e_1~e_2 \\
+\fst~e &=& e (\lambda x.\lambda y. x) \\
+\snd~e &=& e (\lambda x.\lambda y. y) \\
+\end{eqnarray*}
+Hence:
+\newcommand{\fib}{\Phi}
+\newcommand{\fibxa}{\tilde{x}_1}
+\newcommand{\fibxb}{\tilde{x}_2}
+\newcommand{\fibn}{\tilde{n}}
+\newcommand{\fibf}{\tilde{f}}
+\newcommand{\fibp}{\tilde{p}}
+\newcommand{\fibnn}{\tilde{n}'}
+\begin{center}
+\begin{tabular}{rcl}
+$\fib$ & = &
+\hspace{-1em}\multirow{6}{*}{
+\begin{tabular}{lllll}
+\multicolumn{5}{l}{$\lambda \fibn.$} \\
+& \multicolumn{4}{l}{$\mathsf{case}~\fibn~\mathsf{of}$} \\
+&& $\Zero$ & $\Rightarrow$ & $\lambda \fibf.~(\Zero,~\Suc{\Zero})$ \\
+&& $\Suc{\fibnn}$ &$\Rightarrow$& $\lambda \fibf.$ \\
+&&
+\multicolumn{3}{l}{
+\begin{tabular}{lll}
+& \textsf{let} & $\fibp~=~\Ya~\fibnn~\fibf$ \\
+& & $\fibxa~=~\fst~\fibp$ \\
+& & $\fibxb~=~\snd~\fibp$ \\
+& \textsf{in} & $(\fibxb,~\mathsf{add}~\fibxa~\fibxb)$
+\end{tabular}
+}
+\end{tabular}
+}
+\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+&&\\
+$\mathsf{fib}~N$ &=& $\Ya~N~\fib$
+\end{tabular}
+\end{center}
+
+\section{Applications}
+
+I computed some random numbers woo
 
@@
 	\end{center}
-	\caption{We show that RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
+	\caption{RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
 	\label{fig:rdr}
@@
 	\caption{RSLC can also be used to implement popular role-playing games.}
-	\label{fig:rdr}
+	\label{fig:d20}
 \end{figure}

### 2013-01-18 RSLC/paper/paper.tex
Write applications and conclusion

@@
 \subsection{Uniform random number generation}
+\label{sec:uniform}
 
@@
 
-OK, great; got here.
+Got here.
 Up until now I showed how random scope adds inherent nondeterminism that can be harnessed to simulate dice anda make silly graphs.
@@
 \newcommand{\snd}{\mathsf{snd}}
-To compute fibonacci numbers, I encoded pairs:
+To compute fibonacci numbers, I used the standard Church encoding for pairs:
 \begin{eqnarray*}
@@
 
+At this point I could include some addition and multiplication tables like the kind you saw in 3rd grade, but it would be no surprise.
+Suffice to say that it works.
+
 \section{Applications}
-
-I computed some random numbers woo
 
@@
 
+Now that I have shown in an abstract sense how RSLC can be used to express the same deterministic computations from conventional lambda calculus as well as some novel random programming techniques, let's do a real-world application.
+Using the uniform randomness combinator from Section~\ref{sec:uniform}, we can implement any number of standard dice for any nonzero number of sides:
+
+\begin{eqnarray*}
+	\mathsf{d}(\Suc{N}) &=& \mathsf{add}~1~\mathsf{rand}[0,N] \\
+       2\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
+	&& \mathsf{in}~\mathsf{add}~2~x \\
+       3\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
+	&& \mathsf{in}~\mathsf{let}~y=\mathsf{add}~x~\mathsf{rand}[0,N] \\
+	&& \mathsf{in}~\mathsf{add}~3~y
+\end{eqnarray*}
+and so on.
+
+For $N = 5$, using the $2\mathsf{d}(\Suc{N})$ combinator (more commonly known as $2\mathsf{d}6$), we obtain a random distribution equivalent to the one employed in the Random Distance Run~\cite{rdr}, as shown in Figure~\ref{fig:rdr}.
+
+The value $N = 19$ yields the $\mathsf{d}20$ combinator, which generates the random distribution shown in Figure~\ref{fig:d20}.
+Prior work has shown this combinator to be quite useful in tabletop roleplaying games~\cite{dnd-mm,dnd-phb,dnd-dmg}.
+
+\section{Theoretical Concerns}
+
+Some programming language theoreticians will speak of the so-called ``Frame Rule'', which expresses abstraction: if a property $P$ holds for a certain function $f$, then in another function $g$ which invokes $x$, if $P(x)$ implies $P(g)$, then $P([f/x]g)$.
+
+In the context of RSLC, we scoff at this rule.
+
+\section{Concluding Remarks}
+
+RSLC is an extension to standard un(i)typed lambda calculus in which substitution is essentially completely broken.
+However, since substitution is broken in a particularly careful way, it is possible to write interesting programs that make use of the language's built-in nondeterminism.
+Furthermore, I showed that it is still possible to write a bunch of deterministic programs that are commonly used as examples/proofs-of-concept, as long as you put silly hats on some of your variables.
+
+I believe you could extend RSLC's substitution schema to a typed lambda calculus, and furthermore if it only substitutes for variables of the same type, it would even be type-safe.
+
 \bibliographystyle{alpha}
 \bibliography{paper}
+
+\newpage
 

### 2013-01-19 byzantine/byzantine.tex
byzantine: latexify and finish page 2

\documentclass{article}
\usepackage{fullpage}
\begin{document}

\title{The Problem of Heads of a Fighting Force from Long Ago}
\author{Leslie Lamport \and Robert Shostak \and Marshall Pease}
\date{1982}
\maketitle

\abstract{
Groups of computers that you can trust must handle broken parts that give facts
that don't agree to different parts of the group. This situation can be talked
about by pretending about the heads of a fighting force from long ago, who are
waiting with their people around a bad guy city. Talking to each other
only by a person who carries a letter, the people must agree upon a shared
fighting plan. However, one or more of them may be bad guys who will try to
confuse the others.
The problem is to find a plan to make sure that the good guys will agree on
something.  It is shown that, using only spoken letters, this problem is
possible if and only if more than two thirds of the heads are good guys; so a
single bad guy can confuse two good guys.  With written letters which can't be
pretended, the problem is possible for any number of heads and possible bad
guys. Ways to use the plans that are answers are then talked about.
}

Words about what this paper is about: C.2.4. [Computer-Talking Groups]: Groups of many computers--groups of brains of computers; D.4.4 [Computer Brains]: Managing Talking--talking in groups; D.4.5 [Computer Brains]: Being Trusted--accepting faults

Big Picture Words: Plans, Being Trusted

More Key Words and Groups of Words: Agreeing with yourself

% This work was helped in part by the People Who Go To Space under plan NAS1-15428 Change 3, the Group for Being Safe Against Flying Fire Guns under plan DASG60-78-C-0046, and the Fighting Force Work Office under plan DAAG29-79-C-0102.
% 
% Where the people who wrote this live: Computer Study Work Place, SRI In Many Parts of the World, 333 Flying Animal's Wood Street, Park with a Person's Name, CA 94025.
% 
% Being allowed to take without paying all or part of this paper is given as long as the papers are not made or given for making money, the notice that ACM owns this and the name of the paper and its date appear, and notice is given that taking it is by being allowed by the Agreeing-group for Computer Making. To take in another way, or to put out again, needs paying and/or being more allowed.
% 
% © 1982 ACM 0164-0925/82/0700-0382 $00.75 ACM Talking about Ways to Speak to Computers and Groups, Part 4, No. 3, Month Seven 1982, Pieces of paper 382-401.

\section{Opening Words}

A group of computers that can be trusted must be able to live with the breaking of one or more of its parts.
A broken part may act in a way that people often don't think about -- that is, sending facts that don't agree to different parts of the group.
The problem of living with this type of breaking is talked about with make-believe as the Problem of Heads of a Fighting Force from Long Ago. We spend the biggest part of the paper to talking about this make-believe problem and finish by showing how our answers can be used in building a group of computers that can be trusted.

We imagine that several parts of the fighting force from long ago are waiting outside a bad guy city, each group controlled by its own head.
The heads can talk with one another only by a person who carries a letter.
After watching the bad guys, they must decide upon a shared plan of attack.
However, some of the heads may be bad guys, trying to stop the good guys heads from agreeing. The heads must have an plan to make sure that

1. All good guys decide upon the same plan of how to act.

The good guy heads will all do what the plan says they should, but the bad guy heads may do anything they wish. The plan must make sure of fact 1 no matter what the bad guys do.

The good guys should not only agree, but should agree upon a plan that makes sense. So, we also want to make sure that

2. A small number of bad guy heads can't cause the good guy heads to take a bad plan.

Fact 2 is hard to make clear, since it needs you to say exactly what a bad plan is, and we do not attempt to do so. Instead, we consider how the heads reach a way to agree. Each head watches the bad guy city and tells what he or she sees to the others. Let $a(i)$ be the facts told by the head with number $i$. Each head uses some way to put together the facts $a(1) \dots a(number)$ into a single plan of what to do, where $number$ is the number of heads. We can make Fact 1 happen by having all heads use the same way for putting together the facts, and make Fact 2 happen by using a way that can be trusted. Think about this case: if the only thing to agree on is whether to attack or run away, then $a(i)$ can be Head $i$'s thought of which way is best, and the real plan can come from which way has more heads that want to do it. A small number of bad guys can change the way only if there were almost the same number of good guys who wanted to do each way, in which case not either way could be called bad.

While this approach may not be the only way to make Facts 1 and 2 happen, it is the only one we know of.
It needs a way for the heads to tell their facts $a(i)$ to one another. The most clear way is for the head number $i$ to send $a(i)$ by a person who carries a letter to each other head. However, this does not work, because making Fact 1 happen needs every good guy head to read the same facts $a(1) \dots a(number)$, and a bad guy head may send different facts to different heads. For Fact 1 to happen, the following must be true:

3. Every good guy head must read the same facts $a(1) \dots a(number)$.

Fact 3 means that a head can't in all cases use a fact of $a(i)$ read straight from the head number $i$, since a bad guy head number $i$ may send different facts to different heads.
This means that without care, in meeting Fact 3 we might make it possible for the heads to use a fact for $a(i)$ different from the one sent by head number $i$--even though that head is good.
We must not allow this to happen if Fact 2 is to be met. Think about this case: we can't allow a few bad guys to cause the good guys to act as if the facts were "run away" , . . . , "run away" if every good guy said "attack".
So, we need the following new fact for each $i$:

4. If head number $i$ is good, then the fact that he or she sends must be used by every good guy head as the fact for $a(i)$.

\end{document}

### 2013-01-19 RSLC/paper/paper.tex
Polish up rslc paper

@@
 Hence, in RSLC this term can evaluate to either $A$ or $B$.
-(On the other side of the coin, RSLC encourages good programming practice: a programmer who {\em doesn't} want $x$ to evaluate to $A$ should name the first argument something different, to make their intentions more clear.)
+(On the other side of the coin~\cite{twoface}, RSLC encourages good programming practice: a programmer who {\em doesn't} want $x$ to evaluate to $A$ should name the first argument something different, to make their intentions more clear.)
 
@@
 
-Now that I have shown in an abstract sense how RSLC can be used to express the same deterministic computations from conventional lambda calculus as well as some novel random programming techniques, let's do a real-world application.
-Using the uniform randomness combinator from Section~\ref{sec:uniform}, we can implement any number of standard dice for any nonzero number of sides:
+Now that I showed in an abstract sense how RSLC can be used to express popular computations from conventional lambda calculus as well as some novel random programming techniques, let's do a real-world application.
+Using the uniform combinator from Section~\ref{sec:uniform}, we can make standard dice for any nonzero number of sides:
 
@@
 \end{eqnarray*}
-and so on.
 
@@
 I believe you could extend RSLC's substitution schema to a typed lambda calculus, and furthermore if it only substitutes for variables of the same type, it would even be type-safe.
+
+My implementation of RSLC in Haskell is available at \url{https://github.com/bblum/sigbovik/blob/master/RSLC/RSLC.hs}.
 

### 2013-01-26 RSLC/paper/paper.tex
Finish RSLC paper, I think

@@
 \section{Language Definition}
+\label{sec:definition}
 
@@
 
-The value $N = 19$ yields the $\mathsf{d}20$ combinator, which generates the random distribution shown in Figure~\ref{fig:d20}.
+The value $N = 19$ yields the $\mathsf{d}20$ combinator, which generates the random distribution shown in Figure~\ref{fig:d20} above.
 Prior work has shown this combinator to be quite useful in tabletop roleplaying games~\cite{dnd-mm,dnd-phb,dnd-dmg}.
@@
 
+\section*{Acknowledgement}
+
+Michael Sullivan is, of course, the ``Sully'' that I mentioned (and will in the appendix continue to mention).
+
 \bibliographystyle{alpha}
@@
 
-\newpage
+\onecolumn
 
@@
 \section{Discussion of Substitution Rules}
-\begin{figure*}[t]
+
+Section~\ref{sec:definition} shows some ``most important'' example substitution rules, saving the complete set of rules for the appendix, which is where we are now, hence Figure~\ref{fig:subst-complete} below.
+
+Now, Sully thought my two-judgement scheme was gross, and proposed his own, which I think is operationally equivalent.
+Instead of switching to a different judgement the first time you traverse a matching binder (and annotating variables with expression-lists), in you always carry around a counter, which you increment when crossing matching binders and which represents a ``substitution probability'' (really the inverse thereof).
+
+Figure~\ref{fig:sully} shows rules for this alternate substitution scheme.
+In these rules, the notation $\sully{e_0}{x}{p}{e}$ represents substituting with probability $\sullycolor{1/p}$, and the default substitution used by the evaluation rules, $\subst{e_0}{x}{e}$, is equivalent to substituting with probability $1$, i.e., $\sully{e_0}{x}{1}{e}$.
+Also, of course, variables don't have lists attached; they are $x$ rather than $x_{\rep{e}}$.
+While this alternate system eliminates the clunkiness of annotating variables, it has its own clunkiness because the rules \textsc{Subst2-Var-X-Yes} and \textsc{Subst2-Var-X-No} must have explicit probabilities for which of the two of them will apply.
+
+\begin{figure*}[h!]
 	\fbox{$\subst{e_0}{x}{e}$}
@@
 
+\begin{figure*}[t]
+	\fbox{$\sully{e_0}{x}{p}{e}$}
+	\TypeRules{
+		\SullyLamCapture \qquad \SullyLam \qquad \SullyApp \\
+		\SullyVarXA \qquad \SullyVarXB \qquad \SullyVarY \\
+		\SullyCaseCapture \\ \SullyCase \\
+		\SullyZero \qquad \SullySuc \\
+		\SullyLetCapture \qquad \SullyLet
+	}
+	\caption{Alternate substitution rules for RSLC.
+These maintain a counter for the number of binders traversed and uses explicit probabilities for randomness rather than selecting uniformly from a list.}
+	\label{fig:sully}
+\end{figure*}
+
 \end{document}

### 2013-01-27 RSLC/paper/paper.tex
minor changes to rslc paper
rslc typoes

@@
 However, the conventional Y-combinator doesn't work as intended in RSLC, because once a function is substituted into itself for its first argument, subsequent arguments will be captured in the substituted version.
-Even ignoring the internal workings of the combinator, and supposing $\Y$ satisfies $\Y~g \mapsto^* g~(\Y~g)$, here's what happens:
+Even ignoring the internal workings of the combinator, and supposing some $\Y$ that satisfies $\Y~g \mapsto^* g~(\Y~g)$, here's what happens:
 
@@
 The variable $\thing{\bf t}$ accumulates ``$\lambda x$''s as $\rho$ recurses (in the $\Suc{n_2}$ case). You might ask: why does the $\lambda f$ need to be inside the $\mathsf{case}$ statement?
-Because otherwise in {\sc Eval-Case-2}$n_2$ would get captured in the thing substituted for $f$.
+Because otherwise, in {\sc Eval-Case-2}, $n_2$ would get captured in the thing substituted for $f$.
 
@@
 Got here.
-Up until now I showed how random scope adds inherent nondeterminism that can be harnessed to simulate dice anda make silly graphs.
+Up until now I showed how random scope adds inherent nondeterminism that can be harnessed to simulate dice and make silly graphs.
 The other thing Sully didn't believe would be possible was writing deterministic RSLC programs that {\em avoid} using ``the feature''.
@@
 \begin{tabular}{lll}
-& \textsf{let} & $\fibp~=~\Ya~\fibnn~\fibf$ \\
+& \textsf{let} & $\fibp~~=~\Ya~\fibnn~\fibf$ \\
 & & $\fibxa~=~\fst~\fibp$ \\

### 2013-02-10 byzantine/byzantine.tex

@@
-\documentclass{article}
-\usepackage{fullpage}
-\begin{document}
+% v2-acmlarge-sample.tex, dated March 6 2012
+% This is a sample file for ACM large trim journals
+%
+% Compilation using 'acmlarge.cls' - version 1.3, Aptara Inc.
+% (c) 2011 Association for Computing Machinery (ACM)
+%
+% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
+% Users can also go through the FAQs available on the journal's submission webpage.
+%
+% Steps to compile: latex, bibtex, latex latex
+%
+\documentclass[prodmode,acmtap]{acmlarge}
 
+% Metadata Information
+\acmVolume{4}
+\acmNumber{3}
+\acmArticle{} % value doesn't matter
+\articleSeq{34} % minimum value to make the right margin black box disappear
+\acmYear{1982}
+\acmMonth{7}
+
+% Package to generate and customize Algorithm as per ACM style
+\usepackage[ruled]{algorithm2e}
+\SetAlFnt{\algofont}
+\SetAlCapFnt{\algofont}
+\SetAlCapNameFnt{\algofont}
+\SetAlCapHSkip{0pt}
+\IncMargin{-\parindent}
+\renewcommand{\algorithmcfname}{ALGORITHM}
+
+% Page heads
+\markboth{L. Lamport, R. Shostak, and M. Pease}{The Problem of Heads of a Fighting Force from Long Ago}
+
+% Title portion
 \title{The Problem of Heads of a Fighting Force from Long Ago}
-\author{Leslie Lamport \and Robert Shostak \and Marshall Pease}
-\date{1982}
-\maketitle
+\author{LESLIE LAMPORT, ROBERT SHOSTAK, and MARSHALL PEASE \\
+SRI In Many Lands
+}
+% NOTE! Affiliations placed here should be for the institution where the
+%       BULK of the research was done. If the author has gone to a new
+%       institution, before publication, the (above) affiliation should NOT be changed.
+%       The authors 'current' address may be given in the "Author's addresses:" block (below).
+%       So for example, Mr. Fogarty, the bulk of the research was done at UIUC, and he is
+%       currently affiliated with NASA.
 
-\abstract{
+\begin{abstract}
 Groups of computers that you can trust must handle broken parts that give facts
@@
 guys. Ways to use the plans that are answers are then talked about.
-}
+\end{abstract}
 
-Words about what this paper is about: C.2.4.
-[Computer-Talking Groups]: Groups of many computers--groups of brains of computers; D.4.4 [Computer Brains]: Managing Talking--talking in groups; D.4.5 [Computer Brains]: Being Trusted--accepting faults
+% Words about what this paper is about:
+\category{C.2.4.}{Computer-Talking Groups}{Groups of many computers}[groups of brains of computers]
+\category{D.4.4}{Computer Brains}{Managing Talking}[talking in groups]
+\category{D.4.5}{Computer Brains}{Being Trusted}[accepting faults]
 
-Big Picture Words: Plans, Being Trusted
+%Big Picture Words
+\terms{Plans, Being Trusted}
 
-More Key Words and Groups of Words: Agreeing with yourself
+%More Key Words and Groups of Words
+\keywords{Friends agreeing with themselves}
 
-% This work was helped in part by the People Who Go To Space under plan NAS1-15428 Change 3, the Group for Being Safe Against Flying Fire Guns under plan DASG60-78-C-0046, and the Fighting Force Work Office under plan DAAG29-79-C-0102.
-% 
-% Where the people who wrote this live: Computer Study Work Place, SRI In Many Parts of the World, 333 Flying Animal's Wood Street, Park with a Person's Name, CA 94025.
-% 
-% Being allowed to take without paying all or part of this paper is given as long as the papers are not made or given for making money, the notice that ACM owns this and the name of the paper and its date appear, and notice is given that taking it is by being allowed by the Agreeing-group for Computer Making.
-To take in another way, or to put out again, needs paying and/or being more allowed.
-% 
-% © 1982 ACM 0164-0925/82/0700-0382 $00.75 ACM Talking about Ways to Speak to Computers and Groups, Part 4, No. 3, Month Seven 1982, Pieces of paper 382-401.
+%\acmformat{Daniel Pineo, Colin Ware, and Sean Fogarty. 2009. Neural Modeling of Flow Rendering Effectiveness.}
+% At a minimum you need to supply the author names, year and a title.
+% IMPORTANT:
+% Full first names whenever they are known, surname last, followed by a period.
+% In the case of two authors, 'and' is placed between them.
+% In the case of three or more authors, the serial comma is used, that is, all author names
+% except the last one but including the penultimate author's name are followed by a comma,
+% and then 'and' is placed before the final author's name.
+% If only first and middle initials are known, then each initial
+% is followed by a period and they are separated by a space.
+% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.
 
+\begin{document}
+\setcounter{page}{382}
+
+\begin{bottomstuff}
+This work was helped in part by the People Who Go To Space under plan NAS1-15428 Change 3, the Group for Being Safe Against Flying Fire Guns under plan DASG60-78-C-0046, and the Fighting Force Work Office under plan DAAG29-79-C-0102.
+
+Where the people who wrote this live: Computer Study Work Place, SRI In Many Parts of the World, 333 Flying Animal's Wood Street, Park with a Person's Name, CA 94025.
+\end{bottomstuff}
+
+\maketitle
+
+% Head 1
 \section{Opening Words}
@@
 After watching the bad guys, they must decide upon a shared plan of attack.
-However, some of the heads may be bad guys, trying to stop the good guys heads from agreeing. The heads must have an plan to make sure that
+However, some of the heads may be bad guys, trying to stop the good guys heads from agreeing. The heads must have an plan to make sure that \\
+\\
+A. All good guys decide upon the same plan of how to act. \\
 
-1. All good guys decide upon the same plan of how to act.
+The good guy heads will all do what the plan says they should, but the bad guy heads may do anything they wish. The plan must make sure of situation A no matter what the bad guys do.
 
-The good guy heads will all do what the plan says they should, but the bad guy heads may do anything they wish. The plan must make sure of fact 1 no matter what the bad guys do.
+\newcommand\fact{\ensuremath{\mathsf{fact}}}
 
-The good guys should not only agree, but should agree upon a plan that makes sense. So, we also want to make sure that
+The good guys should not only agree, but should agree upon a plan that makes sense. So, we also want to make sure that \\
+\\
+B. A small number of bad guy heads can't cause the good guy heads to take a bad plan. \\
 
-2. A small number of bad guy heads can't cause the good guy heads to take a bad plan.
-
-Fact 2 is hard to make clear, since it needs you to say exactly what a bad plan is, and we do not attempt to do so.
+Situation B is hard to make clear, since it needs you to say exactly what a bad plan is, and we do not attempt to do so.
 Instead, we consider how the heads reach a way to agree.
 Each head watches the bad guy city and tells what he or she sees to the others.
-Let $a(i)$ be the facts told by the head with number $i$.
-Each head uses some way to put together the facts $a(1) \dots a(number)$ into a single plan of what to do, where $number$ is the number of heads.
-We can make Fact 1 happen by having all heads use the same way for putting together the facts, and make Fact 2 happen by using a way that can be trusted.
-Think about this case: if the only thing to agree on is whether to attack or run away, then $a(i)$ can be Head $i$'s thought of which way is best, and the real plan can come from which way has more heads that want to do it.
+Let $\fact(i)$ be the facts told by the head with number $i$.
+Each head uses some way to put together the facts $\fact(1) \dots \fact(n)$ into a single plan of what to do, where $n$ is the number of heads.
+We can make Situation A happen by having all heads use the same way for putting together the facts, and make Situation B happen by using a way that can be trusted.
+Think about this case: if the only thing to agree on is whether to attack or run away, then $\fact(i)$ can be Head $i$'s thought of which way is best, and the real plan can come from which way has more heads that want to do it.
 A small number of bad guys can change the way only if there were almost the same number of good guys who wanted to do each way, in which case not either way could be called bad.
 
-While this approach may not be the only way to make Facts 1 and 2 happen, it is the only one we know of.
-It needs a way for the heads to tell their facts $a(i)$ to one another.
-The most clear way is for the head number $i$ to send $a(i)$ by a person who carries a letter to each other head.
-However, this does not work, because making Fact 1 happen needs every good guy head to read the same facts $a(1) \dots a(number)$, and a bad guy head may send different facts to different heads.
-For Fact 1 to happen, the following must be true:
+While this approach may not be the only way to make situations A and B happen, it is the only one we know of.
+It needs a way for the heads to tell their facts $\fact(i)$ to one another.
+The most clear way is for the head number $i$ to send $\fact(i)$ by a person who carries a letter to each other head.
+However, this does not work, because making situation A happen needs every good guy head to read the same facts $\fact(1) \dots \fact(n)$, and a bad guy head may send different facts to different heads.
+For situation A to happen, the following must be true: \\
+\\
+1. Every good guy head must read the same facts $\fact(1) \dots \fact(n)$. \\
 
-3. Every good guy head must read the same facts $a(1) \dots a(number)$.
+Situation 1 means that a head can't in all cases use a fact of $\fact(i)$ read straight from the head number $i$, since a bad guy head number $i$ may send different facts to different heads.
+This means that without care, in meeting situation 1 we might make it possible for the heads to use a fact for $\fact(i)$ different from the one sent by head number $i$--even though that head is good.
+We must not allow this to happen if situation B is to be met.
+Think about this case: we can't allow a few bad guys to cause the good guys to act as if the facts were ``run away'' , \dots, ``run away'' if every good guy said ``attack''.
+So, we need the following thing for each $i$: \\
+\\
+2. If head number $i$ is good, then the fact that he or she sends must be used by every good guy head as the fact for $\fact(i)$. \\
 
-Fact 3 means that a head can't in all cases use a fact of $a(i)$ read straight from the head number $i$, since a bad guy head number $i$ may send different facts to different heads.
-This means that without care, in meeting Fact 3 we might make it possible for the heads to use a fact for $a(i)$ different from the one sent by head number $i$--even though that head is good.
-We must not allow this to happen if Fact 2 is to be met.
-Think about this case: we can't allow a few bad guys to cause the good guys to act as if the facts were "run away" , .
-.
-.
-, "run away" if every good guy said "attack".
-So, we need the following new fact for each $i$:
+We can say situation 1 another way by saying that for every $i$ (whether or not the head number $i$ is a good guy), \\
+\\
+1'. Any two good guy heads use the same meaning of $\fact(i)$. \\
 
-4. If head number $i$ is good, then the fact that he or she sends must be used by every good guy head as the fact for $a(i)$.
+Situations 1' and 2 both use just the single fact that was sent by head number $i$.
+Because of this, we can think about the smaller problem of how a single head sends their fact to the others.
+We'll talk about this by talking about a controlling head sending a letter to her friends, causing the following problem.
+\\
+
+{\em The Problem of Heads of a Fighting Force from Long Ago.} A controlling head must send an order to her n-1 friends such that \\
+FA1. All good-guy friends listen to the same order. \\
+FA2. If the controlling head is a good guy, then every good guy friend listens to the order she sends. \\
+
+These situations, FA1 and FA2, are called the {\em friends agreeing} situations.
+Note that if the controlling head is a good guy, then FA1 follows from FA2.
+However, the controlling head need not be good.
+
+To fix our first problem, head number i sends their fact of $\fact(i)$ by using an answer to the Fighting Force Heads Problem to send the order ``use $\fact(i)$ as my fact'', with the other heads acting as the helping friends.
+
+\section{Showing What Isn't Possible}
+
+% Bibliography
+\bibliographystyle{ACM-Reference-Format-Journals}
+\bibliography{citations}
+
+% History dates
+\received{Month Four 1980}{Month Ten-and-One 1981}{Month Four 2013}
+
+% \elecappendix
+% \section{Analysis of Invalid Trials}
 

### 2013-02-10 byzantine/byzantine.tex
byzantine: rewrite abstract

@@
 \begin{abstract}
-Groups of computers that you can trust must handle broken parts that give facts
-that don't agree to different parts of the group. This situation can be talked
-about by pretending about the heads of a fighting force from long ago, who are
-waiting with their people around a bad guy city. Talking to each other
-only by a person who carries a letter, the people must agree upon a shared
-fighting plan. However, one or more of them may be bad guys who will try to
-confuse the others.
-The problem is to find a plan to make sure that the good guys will agree on
-something.  It is shown that, using only spoken letters, this problem is
-possible if and only if more than two thirds of the heads are good guys; so a
-single bad guy can confuse two good guys.  With written letters which can't be
-pretended, the problem is possible for any number of heads and possible bad
-guys. Ways to use the plans that are answers are then talked about.
+Sometimes, in a group of computers, there will be a broken part that will tell
+confusing facts to the rest of the computers in the group. If we want to be
+able to trust the entire group, it must be able to ignore these confusing facts
+when deciding what to do. We can talk about this situation by playing
+make-believe about the heads of a fighting force from long ago, who are waiting
+with their people around a bad guy city. The people must agree upon a shared
+fighting plan, but can only talk to each other by sending a person who carries
+a letter. However, one or more of the heads may be bad guys who will try to
+confuse the others. The problem is to find a way for the good guys to talk to
+each other, without knowing who the bad guys are, to make sure they can agree
+on what to do anyway. We will show that, if the good guys use only
+word-of-mouth, this problem is possible if and only if more than two-thirds of
+the heads are good guys. This means that a single bad guy could confuse two
+good guys. If the good guys use signed written letters, which means a bad guy
+couldn't pretend their letter came from someone else, the problem is possible
+for any number of heads and possible bad guys.
+Finally we will talk about ways to use the plans that are answers to this
+problem in different situations in the real world.
 \end{abstract}

### 2013-03-03 bikes/bikes.tex
Add bikes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Large Colored Title Article
% LaTeX Template
% Version 1.1 (25/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[DIV=calc, paper=a4, fontsize=11pt, twocolumn]{scrartcl}	 % A4 paper and 11pt font size

\usepackage{graphicx}
\usepackage[english]{babel} % English language/hyphenation
\usepackage[protrusion=true,expansion=true]{microtype} % Better typography
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{fix-cm}	 % Custom font sizes - used for the initial letter in the document

\usepackage{sectsty} % Enables custom section titles
\allsectionsfont{\usefont{OT1}{phv}{b}{n}} % Change the font of all section commands

\usepackage{fancyhdr} % Needed to define custom headers/footers
\pagestyle{fancy} % Enables the custom headers/footers
\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")

% Headers - all currently empty
\lhead{}
\chead{}
\rhead{}

% Footers
\lfoot{}
\cfoot{}
\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"

\renewcommand{\headrulewidth}{0.0pt} % No header rule
\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule

\usepackage{lettrine} % Package to accentuate the first letter of the text
\newcommand{\initial}[1]{ % Defines the command and style for the first letter
\lettrine[lines=3,lhang=0.3,nindent=0em]{
\color{DarkGoldenrod}
{\textsf{#1}}}{}}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\usepackage{titling} % Allows custom title configuration

\newcommand\prob{$n$-People $k$-Bikes Problem}
\newcommand\Prob{\prob}

\newcommand{\HorRule}{\color{DarkGoldenrod} \rule{\linewidth}{1pt}} % Defines the gold horizontal rule around the title

\pretitle{\vspace{-30pt} \begin{flushleft} \HorRule \fontsize{32}{32} \usefont{OT1}{phv}{b}{n} \color{DarkRed} \selectfont} % Horizontal rule before the title

\title{The \Prob} % Your article title

\posttitle{\par\end{flushleft}\vskip 0.5em} % Whitespace under the title

\preauthor{\begin{flushleft}\large \lineskip 0.5em \usefont{OT1}{phv}{b}{sl} \color{DarkRed}} % Author font configuration

\author{Lancer \r{A}ngstr\"{o}m, } % Your name

\postauthor{\footnotesize \usefont{OT1}{phv}{m}{sl} \color{Black} % Configuration for the institution name
Dept. for Biologically-Impelled Konveyance and Excursion Research Studies% Your institution

\par\end{flushleft}\HorRule} % Horizontal rule after the title

\date{} % Add a date here if you would like one to appear underneath the title block

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Print the title

\thispagestyle{fancy} % Enabling the custom headers/footers for the first page

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% The first character should be within \initial{}
%\initial{H}\textbf{ere is some sample text to show the initial in the introductory paragraph of this template article. The color and lineheight of the initial can be modified in the preamble of this document.}

\initial{W}e present and solve the {\em \prob}. This problem concerns a group of $n$ people who possess among them $k$ bikes, with $k<n$, and wish to travel to a destination $d$ in a way that minimizes the time it takes the last member among them to arrive. We show that $n$ people can, indeed, use $k$ bikes to accelerate their travel beyond walking speed, and give an algorithm for planning bicycle allocation for general $n$ and $k$.

The \prob~has far-reaching implications in many fields, including comestibility theory, deep-space navigation, maximum-jerk motion planning, and pop-math brainteasers.

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------
\section{Introduction}

\initial{Y}ou're hungry. Your $n-1$ friends, some of whom are cyclists, are also hungry. Collectively, you decide to venture out from the comfort of your respective offices on campus, through the perilously cold (and often wet) outdoors of the Pittsburgh winter, to your favorite dining establishment $d$.\footnote{
We have only experimentally verified our research in the specific case of $d =$ {\sf Chipotle}, though we expect the results to generalize to all $d$.}
During the journey, however, the advancing hunger and tormenting weather overcome your group's desire to amble idly at the leisurely walking speed of $w$.
You glance over at your cyclist friends, who are pushing along their $k$ bikes beside themselves, and wonder: ``Might there be a way to exploit the bikes, so that all $n$ of us can get to $d$ sooner?''

Such is the {\em \prob}.

\section{Problem Statement}

% assumptions: free dismount/mount

% restrictions: no handlebar riding

\begin{figure}[h]
	\begin{center}
	\includegraphics[width=0.48\textwidth]{conference.jpg}
	\end{center}
	\label{fig:conference}
	\caption{Not like this.}
\end{figure}
\begin{figure}[h]
	\includegraphics[width=0.48\textwidth]{suicideandmurder.jpg}
	\label{fig:conference}
	\caption{Stripey kid demonstrates correct bicycle allocation protocol.}
\end{figure}

\section{Solution}

\initial{T}o appreciate the method by which $n$ people can ride $k$ bikes to arrive with average speed greater than walking speed $w$, it is best to first consider the minimum nontrivial instantiation of the problem.

\subsection{Specific Case: $n,k = 2,1$}

\subsection{Generalized It}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}

### 2013-03-04 bikes/bikes.tex
bikes: Specific solution

@@
 
-\usepackage{fancyhdr} % Needed to define custom headers/footers
-\pagestyle{fancy} % Enables the custom headers/footers
-\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")
+%\usepackage{fancyhdr} % Needed to define custom headers/footers
+%\pagestyle{fancy} % Enables the custom headers/footers
+%\usepackage{lastpage} % Used to determine the number of pages in the document (for "Page X of Total")
 
 % Headers - all currently empty
-\lhead{}
-\chead{}
-\rhead{}
+%\lhead{}
+%\chead{}
+%\rhead{}
 
 % Footers
-\lfoot{}
-\cfoot{}
-\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"
+%\lfoot{}
+%\cfoot{}
+%\rfoot{\footnotesize Page \thepage\ of \pageref{LastPage}} % "Page 1 of 2"
 
-\renewcommand{\headrulewidth}{0.0pt} % No header rule
-\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule
+%\renewcommand{\headrulewidth}{0.0pt} % No header rule
+%\renewcommand{\footrulewidth}{0.4pt} % Thin footer rule
 
@@
 
-\thispagestyle{fancy} % Enabling the custom headers/footers for the first page
+%\thispagestyle{fancy} % Enabling the custom headers/footers for the first page
 
@@
 During the journey, however, the advancing hunger and tormenting weather overcome your group's desire to amble idly at the leisurely walking speed of $w$.
-You glance over at your cyclist friends, who are pushing along their $k$ bikes beside themselves, and wonder: ``Might there be a way to exploit the bikes, so that all $n$ of us can get to $d$ sooner?''
+You glance over at your cyclist friends, who are pushing along their $k$ bikes beside themselves, and wonder: ``Might there be a way to exploit the bikes so that all $n$ of us can get to $d$ sooner?''
 
@@
 
-% assumptions: free dismount/mount
-
-% restrictions: no handlebar riding
-
 \begin{figure}[h]
 	\begin{center}
-	\includegraphics[width=0.48\textwidth]{conference.jpg}
+	\includegraphics[width=0.5\textwidth]{conference.jpg}
 	\end{center}
+	\caption{Not like this.}
 	\label{fig:conference}
-	\caption{Not like this.}
 \end{figure}
 \begin{figure}[h]
-	\includegraphics[width=0.48\textwidth]{suicideandmurder.jpg}
-	\label{fig:conference}
-	\caption{Stripey kid demonstrates correct bicycle allocation protocol.}
+	\begin{center}
+	\includegraphics[width=0.4\textwidth]{suicideandmurder.jpg}
+	\end{center}
+	\caption{Stripe-Boy demonstrates correct bicycle allocation protocol. (Helmets are apparently beyond the scope of his work.)}
+	\label{fig:stripey}
 \end{figure}
+
+{T}he precise formulation and constraints of the problem are as follows.
+\initial{G}iven $n$ people and $k$ bikes, $n<k$, all colocated, where:
+\begin{enumerate}
+	\setlength{\itemsep}{-0.5em}
+	\item each person can travel on foot, at walking speed $w$, or bike at speed $b$, with $w<b$,
+	\item mounting and dismounting a bike takes a negligible amount of time $\epsilon$,
+	\item at most one person can ride a given bike at a time (see Figures~\ref{fig:conference} and \ref{fig:stripey}), and
+	\item bikes are stationary when not ridden,
+\end{enumerate}
+
+show how all $n$ people can arrive at a destination $d$ in some time $t$ {\em less} than that afforded by walking speed $w$.\footnote{So ``$k$ people bike the entire distance, and wait while $n-k$ people walk'' is not a solution.
+That would be a different kind of ``maximum-jerk'' motion planning, which has already been thoroughly investigated in prior work~\cite{jerk}.}
+
+{W}e now insert a page break so you can figure it out on your own if you want before moving on.
+
+\newpage
 
@@
 
-\initial{T}o appreciate the method by which $n$ people can ride $k$ bikes to arrive with average speed greater than walking speed $w$, it is best to first consider the minimum nontrivial instantiation of the problem.
+\initial{T}he key insight to solving the problem is that bikes can be left unattended for a time, until a walker catches up to its location.
+In fact in any optimal solution every bike must idle at some point.
+
+To appreciate the method by which $n$ people can ride $k$ bikes to arrive with average speed greater than walking speed $w$, it is best to first consider the minimum nontrivial instantiation of the problem.
 
@@
 
+\begin{figure}[h]
+	\bf d=0\space d=0.5\space d=1\space d=1.5\space d=2
+	\includegraphics[width=0.5\textwidth]{x0.png}\\
+	\includegraphics[width=0.5\textwidth]{x1.png}\\
+	\includegraphics[width=0.5\textwidth]{x2.png}\\
+	\includegraphics[width=0.5\textwidth]{x3.png}\\
+	\includegraphics[width=0.5\textwidth]{x4.png}\\
+	\caption{The solution for $n,k=2,1$.}
+	\label{fig:sol}
+\end{figure}
+
+Figure~\ref{fig:sol} shows the solution for 2 people sharing 1 bike.
+The frames show progress after $0$, $1$, $1.5$, $2$, and $3$ units of time have elapsed, assuming $w=0.5$ and $b=1$ (i.e., biking is twice as fast as walking).
+Note that the bike is idle between times $1$ and $2$ -- for a full third of the transit time!
+-- yet the stick figures arrive in $3$ time units, a $25\%$ improvement over walking speed, and only $50\%$ slower than both participants having a bike of their own.
+
 \subsection{Generalized It}
+
+Without loss of generality we say that $d$ is $n$ units of distance away.
 

### 2013-03-04 bikes/bikes.tex
bikes: discuss optimality

@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\bf d=0\space d=0.5\space d=1\space d=1.5\space d=2
@@
 
+Formally we say that the {\em bicycle allocation} for this case is that person $0$ (the blue one, if you're reading this electronically or printed in color) rides bike $0$ for distances between $0$ and $1$, and person $1$ (the red one, if same) rides it for distances between $1$ and $2$.
+For short, we write:
+\begin{eqnarray*}
+	p_0 &=& b_0[0,1] \\
+	p_1 &=& b_0[1,2]
+\end{eqnarray*}
+
 \subsection{Generalized It}
 
-Without loss of generality we say that $d$ is $n$ units of distance away.
+Without loss of generality, let us say that $d$ is $n$ units of distance away (i.e., the same as the number of people).
+Before presenting a general allocation policy, some observations about the problem:
+\begin{enumerate}
+	\setlength{\itemsep}{-0.5em}
+	\item A total of $n^2$ distances will be traveled. A total of $nk$ distances will be biked. It is pointless for bikes to go backwards.
+	\item By symmetry, in an optimal solution, each person will bike $k$ distance and walk $n-k$ distance.
+This also means that fractional distances (less than $1/n$ of the total distance) need not be considered as places to mount/dismount.
+	\item No more than $k$ people can ride for any given distance unit $[i,i+1]$.
+If any fewer than $k$ people rode this way, the bikes would be under-utilised.
+This allows us to check both validity and optimality.
+	\item The optimal (minimum) total travel time is $k/b + (n-k)/w$. The optimal (maximum) average speed is $nbw/(kw+bn-bk)$.
+	% TODO: Discuss epsilon.
+\end{enumerate}
 

### 2013-03-04 bikes/bikes.tex
bikes: add allocation tables

@@
 
+\usepackage{multirow}
 \usepackage{graphicx}
@@
 \usepackage{amsmath,amsfonts,amsthm} % Math packages
-\usepackage[svgnames]{xcolor} % Enabling colors by their 'svgnames'
+\usepackage[svgnames,table]{xcolor} % Enabling colors by their 'svgnames'
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
+\usepackage{subcaption}
 \usepackage{booktabs} % Horizontal rules in tables
@@
 	\begin{center}
-	\includegraphics[width=0.4\textwidth]{suicideandmurder.jpg}
+	\includegraphics[width=0.5\textwidth]{suicideandmurder.jpg}
 	\end{center}
@@
 
+While Figure~\ref{fig:sol} plots the participants' positions at (time,distance) coordinates, for the general instance of the problem it is more useful to plot {\em transportation mode} (i.e., whether a person walks or rides bike $i$) at (person,distance) coordinates.
+For example, in Figure~\ref{fig:grids} we show $n \times n$ grids for the $2,1$; $3,2$; and $5,3$ instances of the problem.
+
+\newcommand\BA{\cellcolor{red}}
+\newcommand\BB{\cellcolor{blue}}
+\newcommand\BC{\cellcolor{green}}
+\begin{figure*}[t]
+	\begin{subfigure}[b]{0.25\textwidth}
+		\begin{tabular}{rp{1.5em}|p{1.5em}|}
+		& \multicolumn{2}{c}{distance} \\
+		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} \\
+		\hline
+		\multicolumn{1}{r|}{0} & \BA & \\
+		\cline{2-3}
+		\multicolumn{1}{r|}{1} & & \BA \\
+		\cline{2-3}
+		\multicolumn{3}{c}{}\\
+		\multicolumn{3}{c}{}\\
+		\multicolumn{3}{c}{}\\
+		\end{tabular}
+		\caption{$2$ people, $1$ bike.}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.3\textwidth}
+		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|}
+		& \multicolumn{3}{c}{distance} \\
+		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} \\
+		\hline
+		\multicolumn{1}{r|}{0} & \BA & \BA & \\
+		\cline{2-4}
+		\multicolumn{1}{r|}{1} & & \BB & \BB \\
+		\cline{2-4}
+		\multicolumn{1}{r|}{2} & \BB & & \BA \\
+		\cline{2-4}
+		\multicolumn{4}{c}{}\\
+		\multicolumn{4}{c}{}\\
+		\end{tabular}
+		\caption{$3$ people, $2$ bikes.}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.35\textwidth}
+		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+		& \multicolumn{5}{c}{distance} \\
+		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} \\
+		\hline
+		\multicolumn{1}{r|}{0} &\BA&\BA&\BA&   & \\
+		\cline{2-6}
+		\multicolumn{1}{r|}{1} &   &\BB&\BB&\BB& \\
+		\cline{2-6}
+		\multicolumn{1}{r|}{2} &   &   &\BC&\BC&\BC\\
+		\cline{2-6}
+		\multicolumn{1}{r|}{3} &\BB&   &   &\BA&\BA\\
+		\cline{2-6}
+		\multicolumn{1}{r|}{4} &\BC&\BC&   &   &\BB\\
+		\cline{2-6}
+		\end{tabular}
+		\caption{$5$ people, $3$ bikes.}
+	\end{subfigure}
+	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.}
+	\label{fig:grids}
+\end{figure*}
+
 %----------------------------------------------------------------------------------------

### 2013-03-04 bikes/bikes.tex
bikes: make tables better

@@
 \usepackage{amsmath,amsfonts,amsthm} % Math packages
-\usepackage[svgnames,table]{xcolor} % Enabling colors by their 'svgnames'
+\usepackage[dvipsnames, svgnames,table]{xcolor} % Enabling colors by their 'svgnames'
 \usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
@@
 
-Formally we say that the {\em bicycle allocation} for this case is that person $0$ (the blue one, if you're reading this electronically or printed in color) rides bike $0$ for distances between $0$ and $1$, and person $1$ (the red one, if same) rides it for distances between $1$ and $2$.
+\newcommand\PA{\ensuremath{\mathsf{i}}}
+\newcommand\PB{\ensuremath{\mathsf{ii}}}
+\newcommand\PC{\ensuremath{\mathsf{iii}}}
+\newcommand\PD{\ensuremath{\mathsf{iv}}}
+\newcommand\PE{\ensuremath{\mathsf{v}}}
+Formally we say that the {\em bicycle allocation} for this case is that person \PA~(the blue one, if you're reading this electronically or printed in color) rides bike $0$ for distances between $0$ and $1$, and person \PB~(the red one, if same) rides it for distances between $1$ and $2$.
 For short, we write:
 \begin{eqnarray*}
-	p_0 &=& b_0[0,1] \\
-	p_1 &=& b_0[1,2]
+	p_\PA &=& b_0[0,1] \\
+	p_\PB &=& b_0[1,2]
 \end{eqnarray*}
@@
 
-\newcommand\BA{\cellcolor{red}}
-\newcommand\BB{\cellcolor{blue}}
-\newcommand\BC{\cellcolor{green}}
+\newcommand\BA{\cellcolor{WildStrawberry}~\textcolor{White}{$b_0$}}
+\newcommand\BB{\cellcolor{NavyBlue}      ~\textcolor{White}{$b_1$}}
+\newcommand\BC{\cellcolor{OliveGreen}   ~\textcolor{White}{$b_2$}}
 \begin{figure*}[t]
+	\begin{center}
 	\begin{subfigure}[b]{0.25\textwidth}
@@
 		\hline
-		\multicolumn{1}{r|}{0} & \BA & \\
+		\multicolumn{1}{r|}{\PA} & \BA & \\
 		\cline{2-3}
-		\multicolumn{1}{r|}{1} & & \BA \\
+		\multicolumn{1}{r|}{\PB} & & \BA \\
 		\cline{2-3}
@@
 		\hline
-		\multicolumn{1}{r|}{0} & \BA & \BA & \\
+		\multicolumn{1}{r|}{\PA} & \BA & \BA & \\
 		\cline{2-4}
-		\multicolumn{1}{r|}{1} & & \BB & \BB \\
+		\multicolumn{1}{r|}{\PB} & & \BB & \BB \\
 		\cline{2-4}
-		\multicolumn{1}{r|}{2} & \BB & & \BA \\
+		\multicolumn{1}{r|}{\PC} & \BB & & \BA \\
 		\cline{2-4}
@@
 		\hline
-		\multicolumn{1}{r|}{0} &\BA&\BA&\BA&   & \\
+		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&   & \\
 		\cline{2-6}
-		\multicolumn{1}{r|}{1} &   &\BB&\BB&\BB& \\
+		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB& \\
 		\cline{2-6}
-		\multicolumn{1}{r|}{2} &   &   &\BC&\BC&\BC\\
+		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC\\
 		\cline{2-6}
-		\multicolumn{1}{r|}{3} &\BB&   &   &\BA&\BA\\
+		\multicolumn{1}{r|}{\PD} &\BB&   &   &\BA&\BA\\
 		\cline{2-6}
-		\multicolumn{1}{r|}{4} &\BC&\BC&   &   &\BB\\
+		\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BB\\
 		\cline{2-6}
@@
 	\end{subfigure}
+	\end{center}
 	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.}

### 2013-03-05 bikes/bikes.tex
a little more

@@
 \begin{figure*}[t]
-	\begin{center}
-	\begin{subfigure}[b]{0.25\textwidth}
+	\begin{subfigure}[b]{0.28\textwidth}
 		\begin{tabular}{rp{1.5em}|p{1.5em}|}
@@
 	\end{subfigure}
-	\begin{subfigure}[b]{0.3\textwidth}
+	\begin{subfigure}[b]{0.33\textwidth}
 		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|}
@@
 	\end{subfigure}
-	\begin{subfigure}[b]{0.35\textwidth}
+	\begin{subfigure}[b]{0.38\textwidth}
 		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
@@
 	\end{subfigure}
-	\end{center}
 	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.}
@@
 \end{figure*}
+
+The pattern is now evident: Person $i$ picks up a bike at distance $i$ and drops it off at distance $i+k~\mathsf{mod}~n$.
+%TODO
+%In the notation from the previous section:
+%\begin{eqnarray*}
+%	p_i  &=&
+%\end{eqnarray*}
+
+We now show that this strategy is both possible and optimal.
+
+\newtheorem{theorem}{Theorem}
+
+\begin{theorem}[Bicycle Allocation]
+The \prob~is solved optimally when $p_i$ rides a bike between distance $i$ and distance $i+k~\mathsf{mod}~n$.
+\end{theorem}
+\begin{proof}
+By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride.
+% TODO
+\end{proof}
 

### 2013-03-05 bikes/bikes.tex
bikes: add a graph

@@
 
-Figure~\ref{fig:sol} shows the solution for 2 people sharing 1 bike.
+\initial{F}igure~\ref{fig:sol} shows the solution for 2 people sharing 1 bike.
 The frames show progress after $0$, $1$, $1.5$, $2$, and $3$ units of time have elapsed, assuming $w=0.5$ and $b=1$ (i.e., biking is twice as fast as walking).
@@
 \begin{eqnarray*}
-	p_\PA &=& b_0[0,1] \\
-	p_\PB &=& b_0[1,2]
+	p_\PA &=& b_0@0 \\
+	p_\PB &=& b_0@1
 \end{eqnarray*}
@@
 
-Without loss of generality, let us say that $d$ is $n$ units of distance away (i.e., the same as the number of people).
-Before presenting a general allocation policy, some observations about the problem:
+\initial{W}ithout loss of generality, let us say that $d$ is $n$ units of distance away (i.e., the same as the number of people).
+First, some observations about the problem:
 \begin{enumerate}
@@
 		\hline
-		\multicolumn{1}{r|}{\PA} & \BA & \BA & \\
+		\multicolumn{1}{r|}{\PA} & \BA & \BB & \\
 		\cline{2-4}
-		\multicolumn{1}{r|}{\PB} & & \BB & \BB \\
+		\multicolumn{1}{r|}{\PB} & & \BA & \BB \\
 		\cline{2-4}
@@
 	\begin{subfigure}[b]{0.38\textwidth}
+		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+		& \multicolumn{5}{c}{distance} \\
+		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} \\
+		\hline
+		\multicolumn{1}{r|}{\PA} &\BA&\BB&\BC&   & \\
+		\cline{2-6}
+		\multicolumn{1}{r|}{\PB} &   &\BA&\BB&\BC& \\
+		\cline{2-6}
+		\multicolumn{1}{r|}{\PC} &   &   &\BA&\BB&\BC\\
+		\cline{2-6}
+		\multicolumn{1}{r|}{\PD} &\BC&   &   &\BA&\BB\\
+		\cline{2-6}
+		\multicolumn{1}{r|}{\PE} &\BB&\BC&   &   &\BA\\
+		\cline{2-6}
+		\end{tabular}
+		\caption{$5$ people, $3$ bikes.}
+	\end{subfigure}
+	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.
+		Can you spot the French flag?}
+	\label{fig:grids}
+\end{figure*}
+
+The pattern is now evident: Person $i$ rides a bike between distance $i$ and distance $i+k~\mathsf{mod}~n$.
+To express this in the notation from the previous section, we can say: $\forall i,j$ with $0\le i < n$ and $0 \le j < k$:
+\begin{eqnarray*}
+	p_i  &=& b_j @ (i+j~\mathsf{mod}~n)
+\end{eqnarray*}
+
+We now show that this strategy is both possible and optimal.
+
+\newtheorem{theorem}{Theorem}
+
+\begin{theorem}[Optimality]
+The \prob~is solved optimally when $p_i$ rides between distance $i$ and distance $i+k~\mathsf{mod}~n$.
+\end{theorem}
+\begin{proof}
+By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride.
+% TODO
+\end{proof}
+
+\subsection{Minimizing Mounts/Dismounts}
+
+\initial{T}he astute reader will note that, while the allocations in the above tables successfully minimize total travel time, they {\em maximize} the number of times each rider must mount and dismount a bicycle.
+If we let $\epsilon$ (the time to mount or dismount) be non-negligible, we see that the travel time achieved above is actually: $k/b+(n-k)/w + 2k\epsilon$.
+
+We might instead attempt to allocate bicycles contiguously, like so:
+		\begin{center}
 		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
@@
 		\end{tabular}
-		\caption{$5$ people, $3$ bikes.}
-	\end{subfigure}
-	\caption{Bicycle allocation policies for selected instances of the \prob. White cells represent walking; cells of different colours represent different bikes.}
-	\label{fig:grids}
-\end{figure*}
+		\end{center}
 
-The pattern is now evident: Person $i$ picks up a bike at distance $i$ and drops it off at distance $i+k~\mathsf{mod}~n$.
-%TODO
-%In the notation from the previous section:
-%\begin{eqnarray*}
-%	p_i  &=&
-%\end{eqnarray*}
+Here, $n-k+1$ riders can ride ``monogamously'', though $k-1$ people will have two riding  sessions with a walking intermission.
+This gives us a better travel time of $k/b+(n-k)/w + 4\epsilon$, which is a strict improvement whenever $k>2$.
 
-We now show that this strategy is both possible and optimal.
-
-\newtheorem{theorem}{Theorem}
-
-\begin{theorem}[Bicycle Allocation]
-The \prob~is solved optimally when $p_i$ rides a bike between distance $i$ and distance $i+k~\mathsf{mod}~n$.
+\begin{theorem}[Contiguous Allocation]
+woo
 \end{theorem}
 \begin{proof}
-By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride.
-% TODO
+woo
 \end{proof}
 
+\section{Evaluation}
+
+\initial{I}n Figure~\ref{fig:graph} we visualize our findings.
+The horizontal line on top represents the state of the art, in which some participants walk the entire distance.
+The bottom line represents the ideal case, in which there are enough bikes for everyone to ride together.
+
+Here we have plotted the travel time for various values of $n,k$ against different configurations of the $b/w$ ratio; i.e., how much faster it is to bike than to walk.
+We hope this presentation will be useful in many different terrains; for example, if the destination is downhill, $b/w$ may be very high indeed.
+
+\begin{figure}[h]
+	\includegraphics[width=0.5\textwidth]{graph.pdf}
+	\caption{Speedup achieved in the \prob. As always, lower is better.}
+	\label{fig:graph}
+\end{figure}
 %----------------------------------------------------------------------------------------

### 2013-03-05 bikes/bikes.tex
bikes: adjust figure
bikes: prove thm 1

@@
 \newcommand\BC{\cellcolor{OliveGreen}   ~\textcolor{White}{$b_2$}}
+\newcommand\BD{\cellcolor{Dandelion}    ~\textcolor{White}{$b_3$}}
 \begin{figure*}[t]
@@
 \begin{proof}
-By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride.
-% TODO
+By condition (3) above, it suffices to show that, at each distance unit, exactly $k$ people ride. A simple induction on $k$ will do nicely.
 \end{proof}
@@
 
-We might instead attempt to allocate bicycles contiguously, like so:
+\begin{figure}[b]
 		\begin{center}
@@
 		\end{center}
+		\caption{Each person mounts/dismounts 2 or 4 times.}
+		\label{fig:continuous}
+\end{figure}
 
+We might instead attempt to allocate bicycles contiguously, like in Figure~\ref{fig:continuous} below.
 Here, $n-k+1$ riders can ride ``monogamously'', though $k-1$ people will have two riding  sessions with a walking intermission.
@@
 woo
+% TODO
 \end{theorem}
@@
 woo
+%TODO
 \end{proof}
+%\begin{figure}[b]
+%		\begin{center}
+%			\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+%		& \multicolumn{6}{c}{distance} \\
+%		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} & \multicolumn{1}{c|}{5} \\
+%		\hline
+%		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&   &   \\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&   \\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&\BC\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PD} &\BB&   &   &\BD&\BD&\BD\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BA&\BA\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{fff} &\BD&\BD&\BD&   &   &\BB\\
+%		\cline{2-7}
+%		\end{tabular}
+%		\end{center}
+%		\caption{Scratch work}
+%\end{figure}
 
@@
 
-\initial{I}n Figure~\ref{fig:graph} we visualize our findings.
+\initial{I}n Figure~\ref{fig:graph} we find our visualizings.
 The horizontal line on top represents the state of the art, in which some participants walk the entire distance.
@@
 
-\begin{figure}[h]
-	\includegraphics[width=0.5\textwidth]{graph.pdf}
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=0.75\textwidth]{graph.pdf}
 	\caption{Speedup achieved in the \prob. As always, lower is better.}
 	\label{fig:graph}
-\end{figure}
+\end{figure*}
 %----------------------------------------------------------------------------------------

### 2013-03-05 byzantine/byzantine.tex
byz: more text

@@
 
+\newcommand\prob{Problem of Heads of a Fighting Force from Long Ago}
 % Page heads
-\markboth{L. Lamport, R. Shostak, and M. Pease}{The Problem of Heads of a Fighting Force from Long Ago}
+\markboth{L. Lamport, R. Shostak, and M. Pease}{The \prob}
 
 % Title portion
-\title{The Problem of Heads of a Fighting Force from Long Ago}
+\title{The \prob}
 \author{LESLIE LAMPORT, ROBERT SHOSTAK, and MARSHALL PEASE \\
@@
 A broken part may act in a way that people often don't think about -- that is, sending facts that don't agree to different parts of the group.
-The problem of living with this type of breaking is talked about with make-believe as the Problem of Heads of a Fighting Force from Long Ago.
+The problem of living with this type of breaking is talked about with make-believe as the \prob.
 We spend the biggest part of the paper to talking about this make-believe problem and finish by showing how our answers can be used in building a group of computers that can be trusted.
@@
 
-{\em The Problem of Heads of a Fighting Force from Long Ago.} A controlling head must send an order to her n-1 friends such that \\
+{\em The \prob.} A controlling head must send an order to her n-1 friends such that \\
 FA1. All good-guy friends listen to the same order. \\
@@
 \section{Showing What Isn't Possible}
+
+The Problem of Heads of a Fighting Force from Long Ago seems simple, but it is actually not.
+It is hard because of the surprising fact that if the heads can send only talk to each other out loud, then any plan can only work if more than two-thirds of the heads are good guys.
+Now, when we say ``talking out loud'', we mean that the words are completely under the control of the person saying them, so a bad guy could say anything he or she wanted to, even ``Your good guy friend said to do such-and-such!''
+This sort of speaking is the same as how computers usually speak to one another. In Part 4, we consider signed, written letters, for which this is not true.
+
+We now show that with spoken words, no plan for three people can handle a single bad guy.
+To keep things simple, we consider the case in which the only things to do are ``attack'' or ``run away''.
+Let us first think about the situation pictured in Figure 1 in which the first head is a good guy, and sends an "attack" order, but Friend 2 is a bad guy and tells Friend 1 that he heard a ``run away'' order.
+For FA2 to happen, Friend 1 must listen to the order to attack.
+
+Now consider another situation, shown in Figure 2, in which the first head is actually a bad guy, and sends an ``attack'' order to Friend 1 and a ``run away'' order to Friend 2.
+Friend 1 does not know who the bad guy is, and she can't tell what the first head actually said to Friend 2.
+So, the situations in these two pictures appear exactly the same to Friend 1.
+If the bad guy lies all the time, then there is no way for Friend 1 to know which situation is happening, so she must listen to the ``attack'' order in both of them.
+So, any time Friend 1 hears an ``attack'' order from the first head, she must listen to it.
+
+However, in the same way we could show that if Friend 2 hears a ``run away'' order from the first head then he must listen to it even if Friend 1 tells him that the head said "attack".
+Because of this, in Figure 2, Friend 2 must listen to the ``run away'' order while Friend 1 listens to the ``attack'' order, which means situation FA1 doesn't happen.
+So, there is no way for three heads to agree if one of them is a bad guy.
+
+This way of thinking about it may appear right, but you should think hard before believing such hand-waving reasoning.
+Although what we said is right after all, we have seen other ``ways of thinking about it'' that are totally wrong.
+We know of no area in the study of computers or the study of numbers in which hand-waving reasoning could more easily lead to wrong answers than in the study of this type of problem.
+For a stronger way of thinking about why this isn't possible, you should read [3].
+% TODO cite
+Using this answer, we can show that no plan can work for a given number of heads if one third of them are bad guys.
+% Omitted: Albanian generals contradiction proof.
+
+You might think that it's so hard to fix the \prob because the heads need to agree with each other completely.
+Actually, this is not the case, and only-sort-of agreeing with each other is just as hard.
+Suppose that instead of trying to agree on a complete plan to attack, the heads must agree only upon a set of times, during which the attack should happen.
+We will say the first head needs to order the time of the attack, and the following two situations need to happen: \\
+\\
+FA1'. All good guys attack within 10 minutes of one another. \\
+FA2'. If the first head is a good guy, then every other good guy attacks within 10 minutes of the time given in the first head's order. \\
+\\
+(We are supposing that the orders are given and thought about the day before the attack and that the time at which an order is heard doesn't matter--only the attack time given in the order matters.)
 

### 2013-03-07 bikes/bikes.tex
bikes: contiguous and conclusion

@@
 
-\newcommand\PA{\ensuremath{\mathsf{i}}}
-\newcommand\PB{\ensuremath{\mathsf{ii}}}
-\newcommand\PC{\ensuremath{\mathsf{iii}}}
-\newcommand\PD{\ensuremath{\mathsf{iv}}}
-\newcommand\PE{\ensuremath{\mathsf{v}}}
+\newcommand\PA{\ensuremath{\mathtt{0}}}
+\newcommand\PB{\ensuremath{\mathtt{1}}}
+\newcommand\PC{\ensuremath{\mathtt{2}}}
+\newcommand\PD{\ensuremath{\mathtt{3}}}
+\newcommand\PE{\ensuremath{\mathtt{4}}}
+\newcommand\distance[1]{\ensuremath{\mathsf{#1}}}
 Formally we say that the {\em bicycle allocation} for this case is that person \PA~(the blue one, if you're reading this electronically or printed in color) rides bike $0$ for distances between $0$ and $1$, and person \PB~(the red one, if same) rides it for distances between $1$ and $2$.
@@
 \newcommand\BD{\cellcolor{Dandelion}    ~\textcolor{White}{$b_3$}}
+\newcommand\BE{\cellcolor{ProcessBlue}       ~\textcolor{White}{$b_3$}}
 \begin{figure*}[t]
@@
 		& \multicolumn{2}{c}{distance} \\
-		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} \\
+		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} \\
 		\hline
@@
 		& \multicolumn{3}{c}{distance} \\
-		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} \\
+		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} \\
 		\hline
@@
 		& \multicolumn{5}{c}{distance} \\
-		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} \\
+		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} \\
 		\hline
@@
 \begin{theorem}[Contiguous Allocation]
-woo
-% TODO
+For all $n$ and $k$, an optimal bicycle allocation exists in which no rider mounts and dismounts more than four times, and is given as follows:
+\newcommand\open{\{}
+\newcommand\close{\}}
+\begin{center}
+\begin{tabular}{rclll}
+	$p_i$ & $=$ & $b_{(i~\mathsf{mod}~k)}@ \open i \rightsquigarrow i+k \close$ && $i \le n-k$ \\
+	\\
+	\multirow{2}{*}{$p_i$} & \multirow{2}{*}{$=$} & $b_{(i~\mathsf{mod}~k)}@ \open i\rightsquigarrow n \close,$ && \multirow{2}{*}{$i > n-k$} \\
+	& & $b_{(i+k-n)}\hspace{0.5em}@ \open 0 \rightsquigarrow i+k-n \close$ \\
+	\\
+\end{tabular}
+\end{center}
+where $b_x@\open y\rightsquigarrow z\close$ expands to $b_x@(y)$, $b_x@(y+1)$, \dots $b_x@(z-1)$.
+
 \end{theorem}
 \begin{proof}
-woo
-%TODO
+By induction on $k$.
 \end{proof}
+
 %\begin{figure}[b]
@@
 %		& \multicolumn{6}{c}{distance} \\
-%		person & \multicolumn{1}{|c|}{0} & \multicolumn{1}{c|}{1} & \multicolumn{1}{c|}{2} & \multicolumn{1}{c|}{3} & \multicolumn{1}{c|}{4} & \multicolumn{1}{c|}{5} \\
+%		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} \\
 %		\hline
@@
 %		\end{tabular}
+%		% 6/4
+%			\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+%		& \multicolumn{6}{c}{distance} \\
+%		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} \\
+%		\hline
+%		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&\BA&   \\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&\BB\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PC} &\BB&   &\BC&\BC&\BC&\BC\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PD} &\BC&\BC&   &\BD&\BD&\BD\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{\PE} &\BD&\BD&\BD&   &\BE&\BE\\
+%		\cline{2-7}
+%		\multicolumn{1}{r|}{fff} &\BE&\BE&\BE&\BE&   &\BA\\
+%		\cline{2-7}
+%		\end{tabular}
+%		% 7/3
+%		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+%		& \multicolumn{7}{c}{distance} \\
+%		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} & \multicolumn{1}{c|}{\distance{6}}\\
+%		\hline
+%		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&   &   &   &   \\
+%		\cline{2-8}
+%		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&   &   &   \\
+%		\cline{2-8}
+%		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&   &   \\
+%		\cline{2-8}
+%		\multicolumn{1}{r|}{\PD} &   &   &   &\BA&\BA&\BA&   \\
+%		\cline{2-8}
+%		\multicolumn{1}{r|}{\PE} &   &   &   &   &\BB&\BB&\BB\\
+%		\cline{2-8}
+%		\multicolumn{1}{r|}{fff} &\BB&   &   &   &   &\BC&\BC\\
+%		\cline{2-8}
+%		\multicolumn{1}{r|}{ggg} &\BC&\BC&   &   &   &   &\BA\\
+%		\cline{2-8}
+%		\end{tabular}
 %		\end{center}
@@
 
+\section{Conclusion}
+
+\initial{Y}ou and your friends have arrived at $d$ at the unprecedentedly fast speed of $nbw/(kw+bn-bk)$, and are warming up indoors and enjoying dinner.
+``Wow,'' you say, ``I sure am glad everyone thought to make $n$ copies of their bike-lock keys!''
+
 \begin{figure*}[t]
 	\centering
-	\includegraphics[width=0.75\textwidth]{graph.pdf}
+	\includegraphics[width=0.85\textwidth]{graph.pdf}
 	\caption{Speedup achieved in the \prob. As always, lower is better.}

### 2013-03-07 bikes/bikes.tex
Finish bikes

@@
 
-The \prob~has far-reaching implications in many fields, including comestibility theory, deep-space navigation, maximum-jerk motion planning, and pop-math brainteasers.
+The \prob~has far-reaching implications in many fields, including comestibility theory~\cite{comestibility}, deep-space navigation~\cite{arkanoid}, maximum-jerk motion planning~\cite{jerk,jerk2}, and pop-math brainteasers.
 
@@
 \newcommand\PE{\ensuremath{\mathtt{4}}}
+\newcommand\PF{\ensuremath{\mathtt{5}}}
+\newcommand\PG{\ensuremath{\mathtt{6}}}
 \newcommand\distance[1]{\ensuremath{\mathsf{#1}}}
@@
 \newcommand\BD{\cellcolor{Dandelion}    ~\textcolor{White}{$b_3$}}
-\newcommand\BE{\cellcolor{ProcessBlue}       ~\textcolor{White}{$b_3$}}
+\newcommand\BE{\cellcolor{ProcessBlue}       ~\textcolor{White}{$b_4$}}
 \begin{figure*}[t]
@@
 \begin{theorem}[Contiguous Allocation]
+\label{thm:cont}
 For all $n$ and $k$, an optimal bicycle allocation exists in which no rider mounts and dismounts more than four times, and is given as follows:
@@
 
-%\begin{figure}[b]
-%		\begin{center}
-%			\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
-%		& \multicolumn{6}{c}{distance} \\
-%		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} \\
-%		\hline
-%		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&   &   \\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&   \\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&\BC\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PD} &\BB&   &   &\BD&\BD&\BD\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BA&\BA\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{fff} &\BD&\BD&\BD&   &   &\BB\\
-%		\cline{2-7}
-%		\end{tabular}
-%		% 6/4
-%			\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
-%		& \multicolumn{6}{c}{distance} \\
-%		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} \\
-%		\hline
-%		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&\BA&   \\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&\BB\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PC} &\BB&   &\BC&\BC&\BC&\BC\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PD} &\BC&\BC&   &\BD&\BD&\BD\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{\PE} &\BD&\BD&\BD&   &\BE&\BE\\
-%		\cline{2-7}
-%		\multicolumn{1}{r|}{fff} &\BE&\BE&\BE&\BE&   &\BA\\
-%		\cline{2-7}
-%		\end{tabular}
-%		% 7/3
-%		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
-%		& \multicolumn{7}{c}{distance} \\
-%		person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} & \multicolumn{1}{c|}{\distance{6}}\\
-%		\hline
-%		\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&   &   &   &   \\
-%		\cline{2-8}
-%		\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&   &   &   \\
-%		\cline{2-8}
-%		\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&   &   \\
-%		\cline{2-8}
-%		\multicolumn{1}{r|}{\PD} &   &   &   &\BA&\BA&\BA&   \\
-%		\cline{2-8}
-%		\multicolumn{1}{r|}{\PE} &   &   &   &   &\BB&\BB&\BB\\
-%		\cline{2-8}
-%		\multicolumn{1}{r|}{fff} &\BB&   &   &   &   &\BC&\BC\\
-%		\cline{2-8}
-%		\multicolumn{1}{r|}{ggg} &\BC&\BC&   &   &   &   &\BA\\
-%		\cline{2-8}
-%		\end{tabular}
-%		\end{center}
-%		\caption{Scratch work}
-%\end{figure}
-
 \section{Evaluation}
@@
 \initial{I}n Figure~\ref{fig:graph} we find our visualizings.
-The horizontal line on top represents the state of the art, in which some participants walk the entire distance.
+The horizontal line on top represents the previous state of the art, in which some participants walk the entire distance.
 The bottom line represents the ideal case, in which there are enough bikes for everyone to ride together.
@@
 
-\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template
-
-\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
-Figueredo, A.~J. and Wolf, P. S.~A. (2009).
-\newblock Assortative pairing and life history strategy - a cross-cultural
-  study.
-\newblock {\em Human Nature}, 20:317--330.
-
-\end{thebibliography}
+\bibliographystyle{abbrv}
+\bibliography{bikes}
 
@@
 
+\newpage
+\appendix
+\section{Scratch Work}
+
+These are some more tables we drew when deciding whether Theorem~\ref{thm:cont} was really true, and they were too colorful not to include.
+
+\begin{figure}[h]
+		\begin{center}
+		%	\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+		%& \multicolumn{6}{c}{distance} \\
+		%person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} \\
+		%\hline
+		%\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&   &   \\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&   \\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&\BC\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PD} &\BB&   &   &\BD&\BD&\BD\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BA&\BA\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{fff} &\BD&\BD&\BD&   &   &\BB\\
+		%\cline{2-7}
+		%\end{tabular}
+		%% 6/4
+		%	\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+		%& \multicolumn{6}{c}{distance} \\
+		%person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} \\
+		%\hline
+		%\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&\BA&   \\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&\BB\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PC} &\BB&   &\BC&\BC&\BC&\BC\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PD} &\BC&\BC&   &\BD&\BD&\BD\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{\PE} &\BD&\BD&\BD&   &\BE&\BE\\
+		%\cline{2-7}
+		%\multicolumn{1}{r|}{fff} &\BE&\BE&\BE&\BE&   &\BA\\
+		%\cline{2-7}
+		%\end{tabular}
+		% 7/3
+		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+			& \multicolumn{7}{c}{distance} \\
+			person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} & \multicolumn{1}{c|}{\distance{6}}\\
+			\hline
+			\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&   &   &   &   \\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&   &   &   \\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&   &   \\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PD} &   &   &   &\BA&\BA&\BA&   \\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PE} &   &   &   &   &\BB&\BB&\BB\\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PF} &\BB&   &   &   &   &\BC&\BC\\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PG} &\BC&\BC&   &   &   &   &\BA\\
+			\cline{2-8}
+		\end{tabular}
+		\vspace{1em}
+
+		\begin{tabular}{rp{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|p{1.5em}|}
+			& \multicolumn{7}{c}{distance} \\
+			person & \multicolumn{1}{|c|}{\distance{0}} & \multicolumn{1}{c|}{\distance{1}} & \multicolumn{1}{c|}{\distance{2}} & \multicolumn{1}{c|}{\distance{3}} & \multicolumn{1}{c|}{\distance{4}} & \multicolumn{1}{c|}{\distance{5}} & \multicolumn{1}{c|}{\distance{6}}\\
+			\hline
+			\multicolumn{1}{r|}{\PA} &\BA&\BA&\BA&\BA&\BA&   &   \\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PB} &   &\BB&\BB&\BB&\BB&\BB&   \\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PC} &   &   &\BC&\BC&\BC&\BC&\BC\\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PD} &\BB&   &   &\BD&\BD&\BD&\BD\\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PE} &\BC&\BC&   &   &\BE&\BE&\BE\\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PF} &\BD&\BD&\BD&   &   &\BA&\BA\\
+			\cline{2-8}
+			\multicolumn{1}{r|}{\PG} &\BE&\BE&\BE&\BE&   &   &\BB\\
+			\cline{2-8}
+		\end{tabular}
+		\end{center}
+		\caption{
+To those who printed in black-and-white and find these illegible, we say: stop reinforcing the color binary!}
+\end{figure}
+
 \end{document}

### 2013-03-07 bikes/bikes.tex
bikes touchups

@@
 
-The \prob~has far-reaching implications in many fields, including comestibility theory~\cite{comestibility}, deep-space navigation~\cite{arkanoid}, maximum-jerk motion planning~\cite{jerk,jerk2}, and pop-math brainteasers.
+The \prob~has far-reaching implications in many fields, including comestibility theory~\cite{comestibility}, deep-space navigation~\cite{arkanoid}, maximum-jerk motion planning~\cite{jerk2}, and pop-math brainteasers.
 
@@
 
-These are some more tables we drew when deciding whether Theorem~\ref{thm:cont} was really true, and they were too colorful not to include.
+These are some more tables we drew when deciding whether Theorem~\ref{thm:cont} was even true at all, and they were too colorful not to include.
 

### 2013-03-10 byzantine/byzantine.tex
more byzantine

@@
 Because of this, in Figure 2, Friend 2 must listen to the ``run away'' order while Friend 1 listens to the ``attack'' order, which means situation FA1 doesn't happen.
-So, there is no way for three heads to agree if one of them is a bad guy.
+So, there is no way for three people to agree if one of them is a bad guy.
 
@@
 
-You might think that it's so hard to fix the \prob because the heads need to agree with each other completely.
+You might think that it's so hard to fix the \prob~because the heads need to agree with each other completely.
 Actually, this is not the case, and only-sort-of agreeing with each other is just as hard.
@@
 (We are supposing that the orders are given and thought about the day before the attack and that the time at which an order is heard doesn't matter--only the attack time given in the order matters.)
+
+Like the \prob, there is no answer to this problem except when more than two-thirds of the heads are good guys.
+We make sure this is true by first showing that if there were an answer for three people that were okay with one bad guy, then we could answer the first problem, which we already showed was not possible.
+Suppose the first head wishes to send an ``attack'' or ``run away'' order.
+We'll say that he can show he wants to attack by ordering an attack time of 1:00, or he could show he wants to run away by ordering an attack time of 2:00.
+ Each of his friends, that is, the other heads, will listen to his order in the following way.
+
+\begin{enumerate}
+	\item After hearing the attack time, a head does one of the following:
+	\begin{enumerate}
+		\item If the time is 1:10 or earlier, then attack.
+		\item If the time is 1:50 or later, then run away.
+		\item Or else, continue to step (2).
+	\end{enumerate}
+	\item Ask the other not-first head what they did in step (1).
+	\begin{enumerate}
+		\item If they did either of the first two things, then do the same thing they did.
+		\item Or else, run away.
+	\end{enumerate}
+\end{enumerate}
+
+It follows from FA2' that if the first head is a good guy, then her good guy friends will hear the right order in step (1), so situation FA2 happens.
+Also, FA1 follows from FA2, so we only need to make sure of FA1 when we suppose the first head is a bad guy.
+Since there is at most one bad guy, this means that both of her friends are good guys.
+It follows from FA1' that if one friend decides to attack in step (1), then the other can't decide to run away in step (1).
+So, either they will both do the same thing in step (1) or at least one of them will wait until step (2).
+In this case, it is easy to see that they both decide to do the same thing, so FA1 happens. This means we just built a three-person answer for the first problem, which is not possible.
+So, we can't have a plan for three people that makes FA1' and FA2' happen if one of them is a bad guy.
+
+Notice how we had one of the heads pretend to be $m$ other heads at the same time.
+We can now do that again to make sure that no answer with three times $m$ (or fewer) people can ever be okay with $m$ bad guys.
+The way of making sure that's true is like the one for the first problem, and is left for the person reading this paper to do on their own.
 

### 2013-03-18 byzantine/byzantine.tex
byz: section 3

@@
 To keep things simple, we consider the case in which the only things to do are ``attack'' or ``run away''.
-Let us first think about the situation pictured in Figure 1 in which the first head is a good guy, and sends an "attack" order, but Friend 2 is a bad guy and tells Friend 1 that he heard a ``run away'' order.
-For FA2 to happen, Friend 1 must listen to the order to attack.
+Let us first think about the situation pictured in Figure~\ref{fig:1} in which the first head is a good guy, and sends an "attack" order, but the second friend is a bad guy and tells the first friend that he heard a ``run away'' order.
+For FA2 to happen, the first friend must listen to the order to attack.
 
-Now consider another situation, shown in Figure 2, in which the first head is actually a bad guy, and sends an ``attack'' order to Friend 1 and a ``run away'' order to Friend 2.
-Friend 1 does not know who the bad guy is, and she can't tell what the first head actually said to Friend 2.
-So, the situations in these two pictures appear exactly the same to Friend 1.
-If the bad guy lies all the time, then there is no way for Friend 1 to know which situation is happening, so she must listen to the ``attack'' order in both of them.
-So, any time Friend 1 hears an ``attack'' order from the first head, she must listen to it.
+Now consider another situation, shown in Figure~\ref{fig:2}, in which the first head is actually a bad guy, and sends an ``attack'' order to the first friend and a ``run away'' order to the second friend.
+The first friend does not know who the bad guy is, and she can't tell what the first head actually said to the second head.
+So, the situations in these two pictures appear exactly the same to the first friend.
+If the bad guy lies all the time, then there is no way for the first friend to know which situation is happening, so she must listen to the ``attack'' order in both of them.
+So, any time the first friend hears an ``attack'' order from the first head, she must listen to it.
 
-However, in the same way we could show that if Friend 2 hears a ``run away'' order from the first head then he must listen to it even if Friend 1 tells him that the head said "attack".
-Because of this, in Figure 2, Friend 2 must listen to the ``run away'' order while Friend 1 listens to the ``attack'' order, which means situation FA1 doesn't happen.
+\begin{figure}[h]
+	TODO
+	\caption{The second friend is a bad guy.}
+	\label{fig:1}
+\end{figure}
+\begin{figure}[h]
+	TODO
+	\caption{The first head is a bad guy.}
+	\label{fig:2}
+\end{figure}
+
+However, in the same way we could show that if the second friend hears a ``run away'' order from the first head then he must listen to it even if the first friend tells him that the head said "attack".
+Because of this, in Figure 2, the second friend must listen to the ``run away'' order while the first friend listens to the ``attack'' order, which means situation FA1 doesn't happen.
 So, there is no way for three people to agree if one of them is a bad guy.
@@
 
+\section{An answer for when the heads use spoken words}
+
+We showed above that for an answer to the \prob~using spoken words to be okay with with $m$ bad guys, there must be more than three times $m$ heads in total.
+We will now give an answer that works for more than three times $m$ heads.
+However, we first want to make clear exactly what we mean by ``spoken words''.
+Each head is supposed to follow some plan that tells them to send some facts to the other heads, and we suppose that a good guy will actually follow the plan.
+When we say ``spoken words'', we mean that we're supposing the following things about the way the heads talk to each other: \\
+\\
+A1. Every word that is said can be heard by the person who is meant to hear it. \\
+A2. Any person who hears something knows who said it. \\
+A3. If someone tried to say something, but their words went missing, you can know that that happened. \\
+
+By supposing A1 and A2, we make sure a bad guy can't cause two other people to not be able to talk each other or to think they said things that they didn't.
+By supposing A3, we make sure a bad guy can't make people not be able to decide something by not saying anything at all.
+
+The answers that we give in this part of the paper and in the following one only work if each head is able to speak to each other head without needing anyone in between to help.
+In the part of this paper number five, we talk about answers which do not need this to happen.
+
+A bad guy first-head may decide not to send any order.
+Since the other heads must decide to follow some order, they need something to do if no-one tells them to do anything.
+We'll say RUN AWAY will be this order.
+
+We will show how to run a Spoken Words plan, which we call SW($m$), for all numbers $m$, by which a first-head sends an order to the rest of the heads.
+We'll talk about this plan by a way of supposing that, as long as the plan works for some number of heads, it can also work for one more than that many heads.
+So if we can show that, then we know it works for all possible numbers of heads.
+
+We show that SW($m$) is an answer to the \prob for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.
+
+The Spoken Words plan where $m$ is nothing at all (which means there are no bad guys): \begin{enumerate}
+	\item The first head says what he decided to do to every other head.
+	\item Each other head does the thing they heard the first head say, or does RUN AWAY if they didn't hear anything.
+\end{enumerate}
+
+The Spoken Words plan where $m$ is one more than some other number $j$:
+\begin{enumerate}
+	\item The first head says what he decided to do to every other head.
+	\item For each number $i$, let $\fact_i$ be the fact that head number $i$ heard from the first head, or else be RUN AWAY if he hears nothing.
+Head number $i$ acts as the first-head in the Spoken Word plan for one fewer than $m$, to say $\fact_i$ to the other heads.
+	\item For each number $i$, and each number $j$ not the same as $i$, let $\fact_j$ be the fact that head number $i$ heard from head number $j$ in step (2) (using the Spoken Word plan for one fewer than $m$), or else RUN AWAY if they heard nothing.
+Head number $i$ will do the thing that they heard most often, among all $\fact_j$.
+\end{enumerate}
+
+To understand how this plan works, we consider the case where $m$ is one and there are four total heads.
+Figure~\ref{fig:3} shows what the second head hears when the first head says some fact $f$, and the third head is a bad guy.
+The third head says some other fact $g$. In step 3, the second head then heard $f$ two times and $g$ once, so decides to do the right thing, which is $f$.
+
+Next, we see what happens if the first head is a bad guy.
+Figure~\ref{fig:4} shows what the other heads hear if the first head says whatever he wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
+Each other head hears all three different orders, so they all decide to do the same thing in step (3), no matter if $x$, $y$, and $z$ are the same.
+
+The bigger Spoken Word plan, where there are two or more bad guys, uses the smaller plan many different times.
+This means that for more than one bad guy, each head will have to talk to the other heads many different times.
+But there has to be a way for each head to tell which facts are which.
+You can make sure there is a way to do this by having each head, just before saying some fact $\fact_i$, also says the number $i$, in step (2).
+
+\begin{figure}[h]
+	TODO
+	\caption{Plan SW(1), where the third head is a bad guy.}
+	\label{fig:3}
+\end{figure}
+\begin{figure}[h]
+	TODO
+	\caption{Plan SW(1), where the first head is a bad guy.}
+	\label{fig:4}
+\end{figure}
+
+\section{A plan for when bad guys can't pretend their words came from someone else}
+
 % Bibliography

### 2013-03-20 byzantine/byzantine.tex
byz: signed messages section

@@
 
-We imagine that several parts of the fighting force from long ago are waiting outside a bad guy city, each group controlled by its own head.
+We imagine that several parts of the fighting force from long ago are waiting outside a bad guy city, each group controlled by its own head person.
 The heads can talk with one another only by a person who carries a letter.
@@
 \\
-1. Every good guy head must read the same facts $\fact(1) \dots \fact(n)$. \\
+1. Every good guy head must hear the same facts $\fact(1) \dots \fact(n)$. \\
 
@@
 FA1. All good-guy friends listen to the same order. \\
-FA2. If the controlling head is a good guy, then every good guy friend listens to the order she sends. \\
+FA2. If the controlling head is a good guy, then every good guy friend listens to the order he sends. \\
 
@@
 To keep things simple, we consider the case in which the only things to do are ``attack'' or ``run away''.
-Let us first think about the situation pictured in Figure~\ref{fig:1} in which the first head is a good guy, and sends an "attack" order, but the second friend is a bad guy and tells the first friend that he heard a ``run away'' order.
+Let us first think about the situation pictured in Figure~\ref{fig:1} in which the controlling head is a good guy, and sends an "attack" order, but the second friend is a bad guy and tells the first friend that he heard a ``run away'' order.
 For FA2 to happen, the first friend must listen to the order to attack.
 
-Now consider another situation, shown in Figure~\ref{fig:2}, in which the first head is actually a bad guy, and sends an ``attack'' order to the first friend and a ``run away'' order to the second friend.
-The first friend does not know who the bad guy is, and she can't tell what the first head actually said to the second head.
+Now consider another situation, shown in Figure~\ref{fig:2}, in which the controlling head is actually a bad guy, and sends an ``attack'' order to the first friend and a ``run away'' order to the second friend.
+The first friend does not know who the bad guy is, and she can't tell what the controlling head actually said to the second head.
 So, the situations in these two pictures appear exactly the same to the first friend.
 If the bad guy lies all the time, then there is no way for the first friend to know which situation is happening, so she must listen to the ``attack'' order in both of them.
-So, any time the first friend hears an ``attack'' order from the first head, she must listen to it.
+So, any time the first friend hears an ``attack'' order from the controlling head, she must listen to it.
 
@@
 	TODO
-	\caption{The first head is a bad guy.}
+	\caption{The controlling head is a bad guy.}
 	\label{fig:2}
@@
 
-However, in the same way we could show that if the second friend hears a ``run away'' order from the first head then he must listen to it even if the first friend tells him that the head said "attack".
+However, in the same way we could show that if the second friend hears a ``run away'' order from the controlling head then he must listen to it even if the first friend tells him that the head said "attack".
 Because of this, in Figure 2, the second friend must listen to the ``run away'' order while the first friend listens to the ``attack'' order, which means situation FA1 doesn't happen.
@@
 Suppose that instead of trying to agree on a complete plan to attack, the heads must agree only upon a set of times, during which the attack should happen.
-We will say the first head needs to order the time of the attack, and the following two situations need to happen: \\
+We will say the controlling head needs to order the time of the attack, and the following two situations need to happen: \\
 \\
 FA1'. All good guys attack within 10 minutes of one another. \\
-FA2'. If the first head is a good guy, then every other good guy attacks within 10 minutes of the time given in the first head's order. \\
+FA2'. If the controlling head is a good guy, then every other good guy attacks within 10 minutes of the time given in the controlling head's order. \\
 \\
@@
 We make sure this is true by first showing that if there were an answer for three people that were okay with one bad guy, then we could answer the first problem, which we already showed was not possible.
-Suppose the first head wishes to send an ``attack'' or ``run away'' order.
+Suppose the controlling head wishes to send an ``attack'' or ``run away'' order.
 We'll say that he can show he wants to attack by ordering an attack time of 1:00, or he could show he wants to run away by ordering an attack time of 2:00.
@@
 	\end{enumerate}
-	\item Ask the other not-first head what they did in step (1).
+	\item Ask the other not-controlling head what they did in step (1).
 	\begin{enumerate}
@@
 
-It follows from FA2' that if the first head is a good guy, then her good guy friends will hear the right order in step (1), so situation FA2 happens.
-Also, FA1 follows from FA2, so we only need to make sure of FA1 when we suppose the first head is a bad guy.
+It follows from FA2' that if the controlling head is a good guy, then his good guy friends will hear the right order in step (1), so situation FA2 happens.
+Also, FA1 follows from FA2, so we only need to make sure of FA1 when we suppose the controlling head is a bad guy.
 Since there is at most one bad guy, this means that both of her friends are good guys.
@@
 The answers that we give in this part of the paper and in the following one only work if each head is able to speak to each other head without needing anyone in between to help.
-In the part of this paper number five, we talk about answers which do not need this to happen.
+% TODO: Decide whether to write this paper
+We would talk about answers which do not need this to happen, but we are leaving that part out of our paper so the paper does not get too long.
+% In the part of this paper number five, we talk about answers which do not need this to happen.
 
-A bad guy first-head may decide not to send any order.
+A bad guy controlling-head may decide not to send any order.
 Since the other heads must decide to follow some order, they need something to do if no-one tells them to do anything.
@@
 
-We will show how to run a Spoken Words plan, which we call SW($m$), for all numbers $m$, by which a first-head sends an order to the rest of the heads.
+We will show how to run a Spoken Words plan by which a controlling-head sends an order to the rest of the heads.
 We'll talk about this plan by a way of supposing that, as long as the plan works for some number of heads, it can also work for one more than that many heads.
@@
 
-We show that SW($m$) is an answer to the \prob for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.
+We show that the Spoken Words plan, with some number $m$, is an answer to the \prob for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.
 
 The Spoken Words plan where $m$ is nothing at all (which means there are no bad guys): \begin{enumerate}
-	\item The first head says what he decided to do to every other head.
-	\item Each other head does the thing they heard the first head say, or does RUN AWAY if they didn't hear anything.
+	\item The controlling head says what he decided to do to every other head.
+	\item Each other head does the thing they heard the controlling head say, or does RUN AWAY if they didn't hear anything.
 \end{enumerate}
@@
 \begin{enumerate}
-	\item The first head says what he decided to do to every other head.
-	\item For each number $i$, let $\fact_i$ be the fact that head number $i$ heard from the first head, or else be RUN AWAY if he hears nothing.
-Head number $i$ acts as the first-head in the Spoken Word plan for one fewer than $m$, to say $\fact_i$ to the other heads.
+	\item The controlling head says what he decided to do to every other head.
+	\item For each number $i$, let $\fact_i$ be the fact that head number $i$ heard from the controlling head, or else be RUN AWAY if he or she hears nothing.
+Head number $i$ acts as the controlling-head in the Spoken Word plan for one fewer than $m$, to say $\fact_i$ to the other heads.
 	\item For each number $i$, and each number $j$ not the same as $i$, let $\fact_j$ be the fact that head number $i$ heard from head number $j$ in step (2) (using the Spoken Word plan for one fewer than $m$), or else RUN AWAY if they heard nothing.
@@
 To understand how this plan works, we consider the case where $m$ is one and there are four total heads.
-Figure~\ref{fig:3} shows what the second head hears when the first head says some fact $f$, and the third head is a bad guy.
+Figure~\ref{fig:3} shows what the second head hears when the controlling head says some fact $f$, and the third head is a bad guy.
 The third head says some other fact $g$. In step 3, the second head then heard $f$ two times and $g$ once, so decides to do the right thing, which is $f$.
 
-Next, we see what happens if the first head is a bad guy.
-Figure~\ref{fig:4} shows what the other heads hear if the first head says whatever he wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
+Next, we see what happens if the controlling head is a bad guy.
+Figure~\ref{fig:4} shows what the other heads hear if the controlling head says whatever he wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
 Each other head hears all three different orders, so they all decide to do the same thing in step (3), no matter if $x$, $y$, and $z$ are the same.
@@
 	TODO
-	\caption{Plan SW(1), where the third head is a bad guy.}
+	\caption{Spoken Words plan, where the third head is a bad guy.}
 	\label{fig:3}
@@
 	TODO
-	\caption{Plan SW(1), where the first head is a bad guy.}
+	\caption{Spoken Words plan, where the controlling head is a bad guy.}
 	\label{fig:4}
@@
 \section{A plan for when bad guys can't pretend their words came from someone else}
+
+As we saw from the situations in Figures 1 and 2, it is the fact that bad guys can lie that makes the \prob so hard.
+The problem becomes easier to find an answer for if we can make sure they don't lie.
+One way to do this is to let the heads send signed letters to each other.
+Using computers, there are ways to sign your letters where you use some very big numbers that make it very hard for a bad guy to pretend a signed letter was written by someone else or says something different.
+We'll suppose the heads use this way of signing letters.
+So, already supposing A1, A2, and A3, we will also suppose the following: \\
+\\
+\begin{tabular}{p{0.8em}p{0.8em}p{0.9\textwidth}}
+A4 & (a) &
+A bad guy can't write a signed letter and pretend that a good guy wrote it, and also can't change a signed
+letter that was already written to say something else. \\
+& (b) & Anyone can make sure that a signed letter came from the person who signed it.
+\end{tabular} \\
+
+Note that we don't suppose anything about letters that a bad guy signed.
+One important thing this means is that bad guys can write signed letters and pretend they came from other bad guys.
+So we are even letting bad guys work together instead of acting alone.
+
+Now that we can write signed letters, the thing we said earlier about needing four or more heads in order to be okay with one bad guy is no longer true.
+In fact, there is an answer for three heads now.
+We now give an answer where any number of heads are okay with any number of bad guys (though the problem doesn't make any sense if there are fewer than two good guys).
+
+Usually, when signing a letter, you write your name at the bottom of the letter.
+But since we are using computers and very big numbers for this stronger sort of signing, the thing you write after the letter is numbers, instead of your name.
+We will call these ``signing-numbers''.
+
+In our answer, the controlling head sends a signed order to each of his friends.
+Each friend then adds his or her signing-numbers to that order and sends it to the other friends, who add their signing-numbers to that order and send it to everyone else, and so on.
+This means that each head must be given one signed letter and sign it and send it to several other people.
+% omitted discussion of what it means to copy and choice(V)
+
+In the following answer, when we say $f:i$, we mean the fact $f$ signed by head number $i$.
+So, $f:j:i$ means the fact $f$ signed by head $j$, and then that fact-and-signing-numbers signed again by head $i$.
+We will say Head One is the controlling-head, who gives the orders.
+In this plan, each head remembers a set $V_i$ which is made of all (not pretended) signed orders he or she has been given so far.
+(If Head One is a good guy, then this set should never have more than one thing in it.)
+
+Do not confuse $V_i$, the set of orders that a head was given, with the set of words that were said to them. There may be many different words spoken about the same order. \\
+
+{\em Spoken Words Plan for $m$ bad guys.} \\
+\\
+At the start, each $V_i$ is empty.
+\begin{enumerate}
+	\item The controlling head signs and sends his order to every friend head.
+	\item For each $i$:
+	\begin{enumerate}
+		\item If head number $i$ is given a letter of the form $f : 1$ from the controlling head and she has not yet heard any order, then
+		\begin{enumerate}
+			\item she lets $V_i$ become just $\{v\}$;
+			\item she sends the signed letter $v: 1 : i$ to every other head.
+		\end{enumerate}
+		\item If head number $i$ gets a letter of the form $f : 1 :j_1 : \dots : j_k$ and $v$ is not in the set $V_i$, then
+		\begin{enumerate}
+			\item she adds $v$ to $V_i$;
+			\item if $k$ is less than $m$ (the number of bad guys), then she sends the letter $f:1 :j_i1 : \dots :j_k : i$ to every head besides the ones with numbers $j_1$ through $j_k$.
+		\end{enumerate}
+	\end{enumerate}
+	\item For each $i$: When head number $i$ will get no more letters, he or she decides to any order in the set $V_i$, or RUN AWAY if the set is empty.
+\end{enumerate}
+
+Note that in step (2), the heads ignore any letter that talks about an order already in the set $V_i$.
+
+% omitted: how to decide when no more messages are coming
+
+\begin{figure}[h]
+	TODO
+	\caption{Signed Letters plan, with the controlling head as a bad guy}
+	\label{fig:5}
+\end{figure}
+
+Figure~\ref{fig:5} shows the Signed Letters plan for the case of three heads when the controlling head is a bad guy.
+The controlling head sends an ``attack'' order to one head and a ``run away'' order to another.
+Both
+friend-heads get the same two orders in step (2), so after step (2) $V_2$ and $V_3$ are both \{``attack'', ``run away''\}, and they both decide what to do the same way.
+Notice that here, which is not like the situation in Figure~\ref{fig:2}, the friends know that the controlling head is a bad guy because he signed two different orders, and A4 states that only he could have made those signing-numbers.
+
+% omitted: correctness proof
+
+% omitted: missing communication paths
+
+\section{Groups of computers that you can trust}
 

### 2013-03-20 byzantine/byzantine.tex
byz: conclusion and references

@@
 for any number of heads and possible bad guys.
-Finally we will talk about ways to use the plans that are answers to this
-problem in different situations in the real world.
+% Finally we will talk about ways to use the plans that are answers to this problem in different situations in the real world.
 \end{abstract}
@@
 For a stronger way of thinking about why this isn't possible, you should read [3].
-% TODO cite
 Using this answer, we can show that no plan can work for a given number of heads if one third of them are bad guys.
@@
 
-\section{Groups of computers that you can trust}
+% omitted: \section{Groups of computers that you can trust}
+
+\section{Closing Words}
+
+We have presented several answers to the \prob, in several different situations.
+The plans in these answers can take a lot of time to run and can also need a lot of letters to be sent.
+Both the Spoken Words plan and the Signed Letters plan might need each order to be sent more than $m$ times, where $m$ is the number of bad guys.
+
+In other words, each head may have to wait for orders that came from the controlling head and were then passed on by $m$ other heads.
+Other people who study computers
+have shown that this must be true for any answer that can be okay with with $m$ bad guys, so our answers are as good as possible.
+
+The Spoken Words plan and the Signed Letters plan need you to send up to $(n-1)$ times $(n-2)$ times $(n-m-1)$ letters or words.
+You can make the number of different times you need to send lower by putting letters or words together and sending them at the same time.
+It may also be possible to lower how much you actually need to say at all, but we have not studied this enough to be sure.
+However, we expect that a large number of letters will still be needed.
+
+Making a group of computers that you can trust even when there might be bad guys that try to confuse you in whatever way they want is a hard problem.
+It seems like any plan to do this will need to take a lot of time or space.
+The only way to make it need less is to start supposing things about the way things might go wrong.
+One such way is to suppose that a computer may stop responding but will never respond with something wrong.
+However, when you need to trust your computers very very much, you can't suppose that sort of thing, and you need to spend all time and space it takes to solve the \prob.
 
 % Bibliography
-\bibliographystyle{ACM-Reference-Format-Journals}
-\bibliography{citations}
+\section*{PAST THINGS THAT WE BUILT ON}
+
+{\small
+\begin{tabular}{p{0.3em}p{0.95\textwidth}}
+1.& {\sc Monroe, R.} Up Goer Five. {\em XKCD Computer Funny Pictures,} \url{http://xkcd.com/1133/}, 2013.\\
+2.& {\sc Sanderson, T.} The Up-Goer Five Word-Writing Computer Game. \url{http://splasho.com/upgoer5/}, 2013.\\
+3.& {\sc Pease, M., Shostak, R., and Lamport, L.} Reaching a Way to Agree Even When There are Faults. {\em J. ACM 27,} 2 (Month Four 1980), 228-234.
+\end{tabular}
+}
 

### 2013-03-20 byzantine/byzantine.tex
byz: small fixes

@@
 
-We show that the Spoken Words plan, with some number $m$, is an answer to the \prob for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.
+We show that the Spoken Words plan, with some number $m$, is an answer to the \prob~for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.
 
@@
 
-As we saw from the situations in Figures 1 and 2, it is the fact that bad guys can lie that makes the \prob so hard.
+As we saw from the situations in Figures 1 and 2, it is the fact that bad guys can lie that makes the \prob~so hard.
 The problem becomes easier to find an answer for if we can make sure they don't lie.

### 2013-03-20 byzantine/byzantine.tex
byz: add figures

@@
 % Package to generate and customize Algorithm as per ACM style
+\usepackage{multirow}
+\usepackage{rotating}
 \usepackage[ruled]{algorithm2e}
@@
 \begin{figure}[h]
-	TODO
+	\begin{center}
+		\includegraphics[width=0.7\textwidth]{fig1.pdf}
+	\end{center}
 	\caption{The second friend is a bad guy.}
@@
 \begin{figure}[h]
-	TODO
+	\begin{center}
+		\includegraphics[width=0.75\textwidth]{fig2.pdf}
+	\end{center}
 	\caption{The controlling head is a bad guy.}
@@
 
-We show that the Spoken Words plan, with some number $m$, is an answer to the \prob~for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.
+We show that the Spoken Words plan, with some number $m$, is an answer to the \prob~for more than three times $m$ many heads, as long as there are $m$ or fewer bad guys.\\
 
-The Spoken Words plan where $m$ is nothing at all (which means there are no bad guys): \begin{enumerate}
+{\em The Spoken Words plan where $m$ is nothing at all (which means there are no bad guys):}\begin{enumerate}
 	\item The controlling head says what he decided to do to every other head.
@@
 
-The Spoken Words plan where $m$ is one more than some other number $j$:
-\begin{enumerate}
+{\em The Spoken Words plan where $m$ is one more than some other number $j$:}\begin{enumerate}
 	\item The controlling head says what he decided to do to every other head.
@@
 
-To understand how this plan works, we consider the case where $m$ is one and there are four total heads.
-Figure~\ref{fig:3} shows what the second head hears when the controlling head says some fact $f$, and the third head is a bad guy.
+To understand how this plan works, we consider the case where $m$ is one and there are four total heads. Figure
+%\ref{fig:3}
+3
+shows what the second head hears when the controlling head says some fact $f$, and the third head is a bad guy.
 The third head says some other fact $g$. In step 3, the second head then heard $f$ two times and $g$ once, so decides to do the right thing, which is $f$.
 
-Next, we see what happens if the controlling head is a bad guy.
-Figure~\ref{fig:4} shows what the other heads hear if the controlling head says whatever he wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
+Next, we see what happens if the controlling head is a bad guy. Figure
+%\ref{fig:4}
+4
+shows what the other heads hear if the controlling head says whatever he wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
 Each other head hears all three different orders, so they all decide to do the same thing in step (3), no matter if $x$, $y$, and $z$ are the same.
@@
 
-\begin{figure}[h]
-	TODO
-	\caption{Spoken Words plan, where the third head is a bad guy.}
-	\label{fig:3}
-\end{figure}
-\begin{figure}[h]
-	TODO
-	\caption{Spoken Words plan, where the controlling head is a bad guy.}
-	\label{fig:4}
-\end{figure}
+\pagebreak
+\thispagestyle{empty}
+\begin{sideways}
+	\begin{tabular}{p{2in}p{9in}p{2in}}
+	\vspace{2em}
+	\\
+	\multicolumn{2}{c}{
+		\includegraphics[width=0.9\textwidth]{fig3.pdf}
+	}
+	&
+	\vspace{-8em}
+	{\footnotesize Fig. 3. Spoken Words plan, where the third head is a bad guy.}
+	\\
+	\vspace{2em}
+	\\
+	\vspace{-10em}
+	{\footnotesize Fig. 4. Spoken Words plan, where the controlling head is a bad guy.}
+	&
+	\multicolumn{2}{c}{
+		\includegraphics[width=0.9\textwidth]{fig4.pdf}
+	}
+	\end{tabular}
+\end{sideways}
+\pagebreak
+\setcounter{figure}{4}
+
+%\begin{figure}[h]
+%	%\begin{sideways}\begin{minipage}{\textwidth}
+%	\begin{center}
+%		\includegraphics[width=0.75\textwidth]{fig3.pdf}
+%	\end{center}
+%	\caption{Spoken Words plan, where the third head is a bad guy.}
+%	\label{fig:3}
+%	%\end{minipage}\end{sideways}
+%\end{figure}
+%\begin{figure}[h]
+%	%\begin{sideways}\begin{minipage}{\textwidth}
+%	\begin{center}
+%		\includegraphics[width=0.75\textwidth]{fig4.pdf}
+%	\end{center}
+%	\caption{Spoken Words plan, where the controlling head is a bad guy.}
+%	\label{fig:4}
+%	%\end{minipage}\end{sideways}
+%\end{figure}
 
@@
 
-\begin{figure}[h]
-	TODO
+\begin{figure}[t]
+	\begin{center}
+		\includegraphics[width=0.7\textwidth]{fig5.pdf}
+	\end{center}
 	\caption{Signed Letters plan, with the controlling head as a bad guy}

### 2013-03-20 byzantine/byzantine.tex
byz: final(?) proofreads

@@
 make-believe about the heads of a fighting force from long ago, who are waiting
-with their people around a bad guy city. The people must agree upon a shared
+with their people around a bad-guy city. The people must agree upon a shared
 fighting plan, but can only talk to each other by sending a person who carries
-a letter. However, one or more of the heads may be bad guys who will try to
-confuse the others. The problem is to find a way for the good guys to talk to
-each other, without knowing who the bad guys are, to make sure they can agree
-on what to do anyway. We will show that, if the good guys use only
-word-of-mouth, this problem is possible if and only if more than two-thirds of
-the heads are good guys. This means that a single bad guy could confuse two
-good guys. If the good guys use signed written letters, which means a bad guy
-couldn't pretend their letter came from someone else, the problem is possible
-for any number of heads and possible bad guys.
+a letter with orders written on it. However, one or more of the heads may be
+bad guys who will try to confuse the others. The problem is to find a way for
+the good guys to talk to each other, without knowing who the bad guys are, to
+make sure they can agree on what to do anyway. We will show that, if the good
+guys use only word-of-mouth, this problem is possible if and only if more than
+two-thirds of the heads are good guys. This means that a single bad guy could
+confuse two good guys. If the good guys use signed written letters, which means
+a bad guy couldn't pretend their letter came from someone else, the problem is
+possible for any number of heads and possible bad guys.
 % Finally we will talk about ways to use the plans that are answers to this problem in different situations in the real world.
@@
 
-A group of computers that can be trusted must be able to live with the breaking of one or more of its parts.
-A broken part may act in a way that people often don't think about -- that is, sending facts that don't agree to different parts of the group.
-The problem of living with this type of breaking is talked about with make-believe as the \prob.
-We spend the biggest part of the paper to talking about this make-believe problem and finish by showing how our answers can be used in building a group of computers that can be trusted.
+If you have a group of computers that you want to be able to trust, it must be able to live even if one of its parts breaks.
+A broken part may act in a way that people often don't think about -- that is, sending confusing facts to different parts of the computer-group.
+We will talk about this kind of breaking-problem, using make-believe, as the \prob.
+%We spend the biggest part of the paper to talking about this make-believe problem and finish by showing how our answers can be used in building a group of computers that can be trusted.
 
-We imagine that several parts of the fighting force from long ago are waiting outside a bad guy city, each group controlled by its own head person.
-The heads can talk with one another only by a person who carries a letter.
+We imagine that several parts of the fighting force from long ago are waiting outside a bad-guy city, each group controlled by its own head person.
+The heads can talk with one another only by sending a person who carries a letter with the orders written on it.
 After watching the bad guys, they must decide upon a shared plan of attack.
-However, some of the heads may be bad guys, trying to stop the good guys heads from agreeing. The heads must have an plan to make sure that \\
+However, some of the heads may be bad guys, trying to stop the good guys heads from agreeing.
+The heads must have an plan to make sure that: \\
 \\
@@
 
-The good guy heads will all do what the plan says they should, but the bad guy heads may do anything they wish. The plan must make sure of situation A no matter what the bad guys do.
+The good guy heads will all do what the plan says they should, but the bad guy heads may do anything they wish. The plan must make sure that situation A happens no matter what the bad guys do.
 
@@
 
-Situation B is hard to make clear, since it needs you to say exactly what a bad plan is, and we do not attempt to do so.
+It is hard to make clear what we mean in Situation B, since it needs you to say exactly what a bad plan is or isn't, and we do not attempt to do so.
 Instead, we consider how the heads reach a way to agree.
-Each head watches the bad guy city and tells what he or she sees to the others.
+Each head watches the bad-guy city and tells what he or she sees to the others.
 Let $\fact(i)$ be the facts told by the head with number $i$.
@@
 FA1. All good-guy friends listen to the same order. \\
-FA2. If the controlling head is a good guy, then every good guy friend listens to the order he sends. \\
+FA2. If the controlling head is a good guy, then every good guy friend listens to the order she sends. \\
 
@@
 
-To fix our first problem, head number i sends their fact of $\fact(i)$ by using an answer to the Fighting Force Heads Problem to send the order ``use $\fact(i)$ as my fact'', with the other heads acting as the helping friends.
+\newcommand\shortprob{Fighting Force Heads Problem}
+To fix our first problem, head number i sends their fact of $\fact(i)$ by using an answer to the \shortprob~to send the order ``use $\fact(i)$ as my fact'', with the other heads acting as the helping friends.
 
@@
 It is hard because of the surprising fact that if the heads can send only talk to each other out loud, then any plan can only work if more than two-thirds of the heads are good guys.
-Now, when we say ``talking out loud'', we mean that the words are completely under the control of the person saying them, so a bad guy could say anything he or she wanted to, even ``Your good guy friend said to do such-and-such!''
-This sort of speaking is the same as how computers usually speak to one another. In Part 4, we consider signed, written letters, for which this is not true.
+Now, when we say ``talking out loud'', we mean that the words are completely under the control of the person saying them, so a bad guy could say anything he or she wanted to, even ``Your good-guy friend said to do such-and-such!''
+This sort of speaking is the same as how computers usually speak to one another. In Part~\ref{sec:4}, we consider signed, written letters, for which this is not true.
 
@@
 To keep things simple, we consider the case in which the only things to do are ``attack'' or ``run away''.
-Let us first think about the situation pictured in Figure~\ref{fig:1} in which the controlling head is a good guy, and sends an "attack" order, but the second friend is a bad guy and tells the first friend that he heard a ``run away'' order.
+Let us first think about the situation pictured in Figure~\ref{fig:1} in which the controlling head is a good guy, and sends an ``attack'' order, but the second friend is a bad guy and tells the first friend that he heard a ``run away'' order.
 For FA2 to happen, the first friend must listen to the order to attack.
@@
 
-However, in the same way we could show that if the second friend hears a ``run away'' order from the controlling head then he must listen to it even if the first friend tells him that the head said "attack".
+However, in the same way we could show that if the second friend hears a ``run away'' order from the controlling head then he must listen to it even if the first friend tells him that the head said ``attack''.
 Because of this, in Figure 2, the second friend must listen to the ``run away'' order while the first friend listens to the ``attack'' order, which means situation FA1 doesn't happen.
@@
 
-You might think that it's so hard to fix the \prob~because the heads need to agree with each other completely.
+You might think that it's so hard to fix the \shortprob~because the heads need to agree with each other completely.
 Actually, this is not the case, and only-sort-of agreeing with each other is just as hard.
@@
 Suppose the controlling head wishes to send an ``attack'' or ``run away'' order.
-We'll say that he can show he wants to attack by ordering an attack time of 1:00, or he could show he wants to run away by ordering an attack time of 2:00.
- Each of his friends, that is, the other heads, will listen to his order in the following way.
+We'll say that she can show she wants to attack by ordering an attack time of 1:00, or he could show he wants to run away by ordering an attack time of 2:00.
+ Each of her friends, that is, the other heads, will listen to her order in the following way.
 
@@
 
-It follows from FA2' that if the controlling head is a good guy, then his good guy friends will hear the right order in step (1), so situation FA2 happens.
+It follows from FA2' that if the controlling head is a good guy, then her good guy friends will hear the right order in step (1), so situation FA2 happens.
 Also, FA1 follows from FA2, so we only need to make sure of FA1 when we suppose the controlling head is a bad guy.
@@
 The answers that we give in this part of the paper and in the following one only work if each head is able to speak to each other head without needing anyone in between to help.
-% TODO: Decide whether to write this paper
+% Decide whether to write this part -- prob. not
 We would talk about answers which do not need this to happen, but we are leaving that part out of our paper so the paper does not get too long.
@@
 
-A bad guy controlling-head may decide not to send any order.
+A bad-guy controlling-head may decide not to send any order.
 Since the other heads must decide to follow some order, they need something to do if no-one tells them to do anything.
@@
 {\em The Spoken Words plan where $m$ is nothing at all (which means there are no bad guys):}\begin{enumerate}
-	\item The controlling head says what he decided to do to every other head.
+	\item The controlling head says what she decided to do to every other head.
 	\item Each other head does the thing they heard the controlling head say, or does RUN AWAY if they didn't hear anything.
@@
 {\em The Spoken Words plan where $m$ is one more than some other number $j$:}\begin{enumerate}
-	\item The controlling head says what he decided to do to every other head.
-	\item For each number $i$, let $\fact_i$ be the fact that head number $i$ heard from the controlling head, or else be RUN AWAY if he or she hears nothing.
+	\item The controlling head says what she decided to do to every other head.
+	\item For each number $i$, let $\fact_i$ be the fact that head number $i$ heard from the controlling head, or else be RUN AWAY if they hear nothing.
 Head number $i$ acts as the controlling-head in the Spoken Word plan for one fewer than $m$, to say $\fact_i$ to the other heads.
 	\item For each number $i$, and each number $j$ not the same as $i$, let $\fact_j$ be the fact that head number $i$ heard from head number $j$ in step (2) (using the Spoken Word plan for one fewer than $m$), or else RUN AWAY if they heard nothing.
-Head number $i$ will do the thing that they heard most often, among all $\fact_j$.
+Head number $i$ will do the thing that they heard most often, from all the $\fact_j$.
 \end{enumerate}
@@
 4
-shows what the other heads hear if the controlling head says whatever he wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
+shows what the other heads hear if the controlling head says whatever she wants, which we'll say are $x$, $y$, and $z$, because they can all be different.
 Each other head hears all three different orders, so they all decide to do the same thing in step (3), no matter if $x$, $y$, and $z$ are the same.
@@
 You can make sure there is a way to do this by having each head, just before saying some fact $\fact_i$, also says the number $i$, in step (2).
+
+To make sure the Spoken Words plan is right for any number of bad guys, we first make sure the following expression is true.
+
+\newtheorem{mylemma}{Expression}
+\begin{mylemma}
+	\label{lemma:1}
+	For any numbers $m$ and $k$, the Spoken Words plan makes situation FA2 happen if there are more than $2k+m$ heads and at most $k$ bad guys.
+\end{mylemma}
+
+{\sc Reason That is True.}
+We can show the reason this is true by thinking about it in a way of supposing it is true for a given number then showing it must also be true for one more than that number.
+In this case the number we will pay attention to is $m$.
+FA2 only talks about what must happen if the
+controlling head is a good guy. First, using A1, it is easy to see that the most simple plan (the Spoken Words plan for no bad guys at all), so the expression is true for the starting case.
+We now suppose the expression is true for one fewer than some number $m$, and show it is true for $m$.
 
@@
 
+% TODO: If you rewrite the introduction, readjust this
+In step (1), the good-guy controlling head sends an order to all $n - 1$ of their friends. In step (2),
+each good-guy friend runs the Spoken Words plan for $m-1$ bad guys with $n-1$ heads.
+Since we supposed $n$ is bigger than $2k+m$, we know $n-1$ is bigger than $2k+(m-1)$, so we can use the thing we supposed to show that every good-guy friend decides to do what the controlling head told them.
+Since there are at most $k$ bad guys, and $n-1$ is bigger than $2k+(m-1)$, which in turn is bigger than two times $k$, then more than half of the friend-heads are good guys.
+So, each good-guy friend decides to do what the controlling head told them in the third step, which makes sure that FA2 happens.
+\qed
+
+The following expression says that the Spoken Words plan is a right answer to the \prob.
+
+\begin{mylemma}
+For any number of bad guys, the Spoken Words plan makes sure both Friends-Agreeing situations happen if there are more than three times as many people as bad guys.
+\end{mylemma}
+
+{\sc Reason That is True.}
+We show the reason this is true by the same way of supposing as before.
+If there are no bad guys, then it is easy to see that the easiest Spoken Words plan makes situations FA1 and FA2 happen.
+So, we suppose that the expression is true for the Spoken Words plan for $m-1$ bad guys, and show it is true for $m$ bad guys.
+
+We first consider the case in which the controlling head is good. By letting $k$ be the same as $m$ in Expression~\ref{lemma:1},
+we see that the Spoken Words plan for $m$ bad guys makes FA2 happen.
+FA1 follows from FA2 if the controlling head is good, so we need only make sure FA1 happens in the case that the controlling head is a bad guy.
+
+There are at most $m$ bad guys, and the controlling head is one of them, so at most one fewer than $m$ of their friends are also bad guys.
+Since there are more than three times $m$ people in total, there are more than $3m-1$ friends, and one fewer than three times $m$ is bigger than three times one fewer than $m$.
+So we can use the thing we supposed to show that the Spoken Words plan for $m-1$ bad guys makes FA1 and FA2 happen.
+Because of that, any two good-guy friend-heads decide to do the same thing in the third step of the plan, which makes FA1 happen.
+\qed
+
 \section{A plan for when bad guys can't pretend their words came from someone else}
+\label{sec:4}
 
-As we saw from the situations in Figures 1 and 2, it is the fact that bad guys can lie that makes the \prob~so hard.
+As we saw from the situations in Figures 1 and 2, it is the fact that bad guys can lie that makes the \shortprob~so hard.
 The problem becomes easier to find an answer for if we can make sure they don't lie.
@@
 
-In our answer, the controlling head sends a signed order to each of his friends.
+In our answer, the controlling head sends a signed order to each of her friends.
 Each friend then adds his or her signing-numbers to that order and sends it to the other friends, who add their signing-numbers to that order and send it to everyone else, and so on.
@@
 \begin{enumerate}
-	\item The controlling head signs and sends his order to every friend head.
+	\item The controlling head signs and sends her order to every friend head.
 	\item For each $i$:
 	\begin{enumerate}
-		\item If head number $i$ is given a letter of the form $f : 1$ from the controlling head and she has not yet heard any order, then
+		\item If head number $i$ is given a letter of the form $f : 1$ from the controlling head and he has not yet heard any order, then
 		\begin{enumerate}
-			\item she lets $V_i$ become just $\{v\}$;
-			\item she sends the signed letter $v: 1 : i$ to every other head.
+			\item he lets $V_i$ become just $\{v\}$;
+			\item he sends the signed letter $v: 1 : i$ to every other head.
 		\end{enumerate}
@@
 		\begin{enumerate}
-			\item she adds $v$ to $V_i$;
-			\item if $k$ is less than $m$ (the number of bad guys), then she sends the letter $f:1 :j_i1 : \dots :j_k : i$ to every head besides the ones with numbers $j_1$ through $j_k$.
+			\item he adds $v$ to $V_i$;
+			\item if $k$ is less than $m$ (the number of bad guys), then he sends the letter $f:1 :j_i1 : \dots :j_k : i$ to every head besides the ones with numbers $j_1$ through $j_k$.
 		\end{enumerate}
 	\end{enumerate}
-	\item For each $i$: When head number $i$ will get no more letters, he or she decides to any order in the set $V_i$, or RUN AWAY if the set is empty.
+	\item For each $i$: When head number $i$ will get no more letters, he or she decides to follow any order in the set $V_i$, or RUN AWAY if the set is empty.
 \end{enumerate}
@@
 
-\begin{figure}[t]
+\begin{figure}[ht]
 	\begin{center}
@@
 friend-heads get the same two orders in step (2), so after step (2) $V_2$ and $V_3$ are both \{``attack'', ``run away''\}, and they both decide what to do the same way.
-Notice that here, which is not like the situation in Figure~\ref{fig:2}, the friends know that the controlling head is a bad guy because he signed two different orders, and A4 states that only he could have made those signing-numbers.
+Notice that here, which is not like the situation in Figure~\ref{fig:2}, the friends know that the controlling head is a bad guy because she signed two different orders, and A4 states that only she could have made those signing-numbers.
 
@@
 
-The Spoken Words plan and the Signed Letters plan need you to send up to $(n-1)$ times $(n-2)$ times $(n-m-1)$ letters or words.
+The Spoken Words plan and the Signed Letters plan need you to send up to $(n-1)$ times $(n-2)$ times (and so on\dots) times $(n-m-1)$ letters or words.
 You can make the number of different times you need to send lower by putting letters or words together and sending them at the same time.
@@
 The only way to make it need less is to start supposing things about the way things might go wrong.
-One such way is to suppose that a computer may stop responding but will never respond with something wrong.
-However, when you need to trust your computers very very much, you can't suppose that sort of thing, and you need to spend all time and space it takes to solve the \prob.
+One such way is to suppose that a computer may stop responding but will never respond with something confusing.
+However, when you need to trust your computers very very much, you can't suppose that sort of thing, and you need to spend all time and space it takes to find an answer for the \shortprob.
 

### 2013-03-21 tracks/tracks.tex
Update tracks to 2013

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}

\usepackage[lf]{electrum}
\usepackage[T1]{fontenc}

\newcommand\track[2]{\begin{center} {\Large \bf \em \sc Track #1} \vspace{0.2in}

{\huge \bf #2} \vspace{0.15in}
\end{center}
\thispagestyle{empty}}

\newcommand\asdf[2]{\vspace{0.15in} \item {\Large \bf \sf #1}

{\large \em #2}}
\newcommand\keywords[1]{

{\small {\bf Keywords:} #1}}

\newcommand\intermission[2]{}
%{\vspace{4em}\begin{tabular}{rl}
%	\Large \sc Intermission: & \Large \bf \sf #1 \\
%				 & \keywords{#2}
%\end{tabular}}
\begin{document}

\track{1}{Contemporary Issues in Politics and Law}

\begin{enumerate}
\asdf{SIGBOVIK License Agreement (excerpt)}
{Ix Tchow, Ph.D.}
\asdf{Redistributive Version Control Systems}
{Karlo Angiuli and Frederick Engels}
\keywords{redistribution, communism, version control}
\asdf{Psychology in Response to Presidential Poles}
{Timothy Broman}
\keywords{statistics, politics, election, cognitive dissonance, confirmation bias}
\end{enumerate}
\intermission{P-Rotip}{dreck, doggerel, puns}

\newpage

\track{2}{Harry Bovik's Research Calculus}

\begin{enumerate}
\asdf{The Randomly-Scoped Lambda Calculus}
{Ben Blum}
\keywords{static scope, dynamic scope, lower is better}
\asdf{Recovered Mathematical Journal}
{James McCann}
\asdf{MacLeod Computing: A New Paradigm for Immortal Distribution}
{Taus Brock-Nannestad, Gian Perrone, and Dr. Tom Murphy VII Ph.D.}
\keywords{cloude, compute, scottt}
\asdf{The ($\infty$,1)-Accidentopos Model of Unintentional Type Theory \\(extended abstract)}
{Carlo Angiuli}
\keywords{mistaken identity, accidentopos, object misclassifiers}
\end{enumerate}
\intermission{Pr-Otip}{art, artifice, artificial}

\newpage

\track{3}{Artificial Stupidity}

\begin{enumerate}
\asdf{You Only Learn Once: A Stochastically weighted AGGRegation Approach to Online Regret Minimization}
{danny ``vato loco'' m4turana and d4vid ``the H is silent'' f0uhey}
\keywords{yolo, swag, swagger, dat, bound, machine learning, ICML, brogrammerproblems, noregrets, belieber}
\asdf{I Lost The Game, and So Will You: Implications of Mindvirus Circulation in a Post-Singularity World}
{Shomir Wilson}
\keywords{singularity, game, experiment}
\asdf{Fandomized Algorithms and Fandom Number Generation}
{Lindsey Kuper and Alex Rudnick}
\keywords{fandomized algorithms, fandom numbers, fentropy}
\end{enumerate}
\intermission{pro-tip}{pro, tip, cycle}

\newpage

\track{4}{Computer Vision(aries)}

\begin{enumerate}
\asdf{Optimal Coupling and Gaybies}
{Nicolas Feltman}
\keywords{coupling, gaybies, neil patrick harris}
\asdf{Cat Basis Purrsuit}
{Daniel Caturana and David Furry}
\keywords{Meow, mew, miao, compressed sensing}
\asdf{A Spectral Approach to Ghost Detection}
{Daniel Maturana and David Fouhey}
\keywords{spectral and astral methods, trans-dimensional group lasso, ghosts, occult, hauntology, crystal energy, supernatural, paranormal}
\end{enumerate}
\intermission{prot-ip}{bad, twitter, joke}

\newpage

\track{5}{Productivity and Meta-Productivity \\ \large (or ``Synergistic Hyperparadigmatism'', as they say in the literature)}
\begin{enumerate}
\asdf{Really Amazing New Idea}
{Basquet Kase, Nought Job, and Ayn San Ety}
\keywords{ideas, doggerel, dreck}
\asdf{METHOD AND APPARATUS FOR PRESSING SPACEBAR}
{Joshua A. Wise and Jacob D. Potter}
\keywords{patent, spacebar, microcontrollers}
\asdf{METHOD AND APPARATUS FOR PUSHING SPACEBAR}
{Joshua A. Wise and Jacob D. Potter}
\keywords{patent, lego, spacebar}
\asdf{Find a Separating Hyperplane With This One Weird Kernel Trick \\(sponsored contribution)}
{Daniel Maturana and David F. Fouhey}
\keywords{kernel methods, fast money, overfitting, bayesian monopoly}
\asdf{On $n$-Dimensional Polytope Schemes}
{David F. Fouhey and Daniel Maturana}
\keywords{not pyramid scheme, polytope schemes, get rich fast, algebraic geometry, elimination ideals}
\asdf{DUI: A Fast Probabilistic Paper Evaluation Tool}
{Ivan Ruchkin and Ashwini Rao}
\asdf{Paper and Pencil: a Lightweight WYSIWYG Typesetting System}
{Paul Stansifer}
\end{enumerate}
\intermission{proti-p}{joke, got, old}

\newpage

\track{6}{Time Travel, Space Travel, and Other Fun Games for Children}
\begin{enumerate}
\asdf{A Proposal for Overhead-Free Dependency Management with Temporally Distributed Virtualization}
{Peter chapman, Deby Katz, and Stefan Muller}
\keywords{virtual time machine, time machine, virtualization}
\asdf{The $n$-People $k$-Bikes Problem}
{Lancer \r{A}ngstr\"{o}m}
\keywords{bikes, bixe, biologically-impelled konveyance, motion planning, contiguous allocation algorithms}
\asdf{The Problem of Heads of a Fighting Force from Long Ago}
{Leslie Lamport, Robert Shostak, and Marshall Pease}
\keywords{groups of brains of computers, talking in groups, accepting faults, friends agreeing with themselves, people who go to space, agreeing-group for computer making, ten hundred most used words, sorry to the people who wrote this paper the first time}
\asdf{duoludo: a game whose purpose is games}
{David Renshaw}
\keywords{game, games, gameses}
\asdf{The First Level of Super Mario Bros. is Easy with Lexicographic Orderings and Time Travel {\normalsize\dots after that it gets a little tricky.}}
{Dr. Tom Murphy VII Ph.D.*}
\keywords{computational super mario brothers, memory inspection, lexicographic induction, networked entertainment systems, pit-jumping}
\end{enumerate}

\end{document}

### 2014-03-19 proceedings/copyright-page.tex
add a bunch of sigbovik proceedings stuff

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

% TODO: This is the only thing you need to change here each year!
\newcommand\thisyear{2014}

\multirow{3}{*}{\includegraphics[width=\textwidth]{copyright-header.png}}\\
\vspace{10em}

SIGBOVIK

A Record of the Proceedings of SIGBOVIK \thisyear

ISSN 2155-0166

April 1, \thisyear

\vspace{2em}

Copyright is maintained by the individual authors, though obviously this all gets posted to the Internet and stuff, because it's \thisyear.

Permission to make digital or hard copies of portions of this work for personal use is granted; permission to make digital or hard copies of portions of this work for classroom use is also granted, but seems ill-advised.
Abstracting with credit is permitted; abstracting with credit cards seems difficult.

Additional copies of this work may be ordered from Lulu; refer to {\tt http://sigbovik.org} for details.

\end{document}

### 2014-03-19 proceedings/message-from-committee.tex
add a bunch of sigbovik proceedings stuff

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\usepackage[normalem]{ulem}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{3}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 2014}} \\ &\\
& \LARGE{Message from the Organizing Committee} \\
&\\
\hline
\end{tabular}}
\vspace{2em}

{\large \bf \sffamily Multiple-Choice Section}

\begin{enumerate}
\item {\em 5 points.} Please indicate below the emotions with which the SIGBOVIK Organizing Committee presents this document. Circle up to three, but no fewer than four, choices.
\begin{enumerate}
\item Joy
\item Embarrassment
\item Surprise
\item Admiration
\item Arrogance
\item Pride
\item Terror
\end{enumerate}
\item {\em 5 points.} Which is the correct full name of this conference?
\begin{enumerate}
\item The Eighth Annual Intercalary Conference about Symposium on Human Dance Party of Workshop in Celebration of Harry W. Bovik's $2^6$th Birthday.
\item The Eightieth Annual Intracalary Workshop about Workshop on Robot Dance Party of Conference in Mourning of Harry Q. Bovik's $2^6$th Birthday.
\item The Eighth Centennial Intercalary Workshop about Symposium on Robot Karaoke Party of Conference in Celebration of Harry Q. Bovik's $3^6$th Birthday.
\item The Eighth Annual Intercalary Workshop about Symposium on Robot Dance Party of Conference in Celebration of Harry Q. Bovik's $2^6$th Birthday.
\end{enumerate}

\end{enumerate}

{\large \bf \sffamily Reading Comprehension Section}

{\em 25 points.}

This year's SIGBOVIK marks our eighth continuous year of high-quality research in a rich variety of topics, including but not limited to applied phlogistonics, synergistic hyperparadigmatism, and elbow macaroni. Eight is a very special annuality of this conference, as it not only breaks the ``highest annuality'' record set last year, but also is the fourth power of two number of years for which this conference has been in existence. If you're not impressed yet, note that the fourth power of two is a very special power of two, as four is the second power of two, and you know what they say about the second power of two.

As it is a very special year, we have decided to conduct a survey of past SIGBOVIKs, or SIGBOVIX for short, classifying the various papers to get a feel for where future SIGBOVIKs might be headed. This was driven by the realization depicted in Figure~\ref{fig:table}.

\begin{figure}[h]
\begin{center}
\begin{tabular}{p{5em}|p{5em}|p{5em}|}
& Serious treatment & Humorous treatment \\
\hline
Serious idea & \bf Mainstream conferences & \bf SIGBOVIK \\
\hline
Humorous idea & \bf SIGBOVIK & \bf SIGBOVIK \\
\hline

\end{tabular}
\end{center}
\caption{A research paper is comprised of an {\em idea} and a {\em treatment} of that idea. Each of the idea and the treatment may be either serious or humorous.}
\label{fig:table}
\end{figure}

Observe that SIGBOVIK offers a venue for three times as many different types of research as ``mainstream'' conferences. We won't belabour the obvious conclusion that SIGBOVIK is superior, although, just saying, you know; rather, we are interested in distinguishing among SIGBOVIK's three major categories of research.
\begin{enumerate}
\item {\bf Humorous idea, humorous treatment.} The most common, but by no means the most lowly, of SIGBOVIK publications. The research contribution is typically contained entirely within the paper itself; no separate artefact is constructed.
\item {\bf Serious idea, humorous treatment.} The rarest specimen of SIGBOVIK research. Often a retelling of famous events, people, or theoretical results from mainstream computer science.
\item {\bf Humorous idea; serious treatment.} Frequently denoted by independent artefacts accompanying the paper submission, such as a website, compiler implementation, hardware, or proof. Such publications would often be suitable for acceptance at ``mainstream'' conferences, were the core idea not out of scope.
\end{enumerate}

The results of our survey are shown in Figure~\ref{fig:results} \sout{below}on the next page.

\begin{figure}[h]
\begin{center}
\begin{tabular}{r|c|c|c|c|c|c|c|c|}
& \multicolumn{8}{c|}{SIGBOVIK year of incidence} \\
& 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 \\
\hline
\hline
Category 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\hline
Category 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\hline
Category 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\hline
Other & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\hline
\hline
Total publications & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\

\end{tabular}
\end{center}
\caption{Results.}
\label{fig:results}
\end{figure}

\end{document}

### 2014-03-19 proceedings/titlepage.tex
add a bunch of sigbovik proceedings stuff

% NB. xelatex this file.

\documentclass{article}
\usepackage{fontspec}

\begin{document}

\setmainfont{Rumsey}
\setsansfont{Birbaslo}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\begin{titlepage}
\begin{center}

% Upper part of the page. The '~' is needed because \\
% only works if a paragraph has started.
\textsc{\LARGE The Association for Computational Heresy}\\[0.5cm]

\textsc{\large presents}\\[1cm]

\textsc{\Large A Record of the Proceedings of}\\[0.5cm]
% Title
\HRule \\[0.4cm]
{ \Huge \sf SIGBOVIK 2014 \\[0.4cm] }

\HRule \\[2cm]

% Author and supervisor

The eight annual intercalary robot dance party in celebration

of workshop on symposium about Harry Q. Bovik's $2^6$th birthday

\vfill

% Bottom of the page
{\Large Carnegie Mellon University\\[0.25cm]
Pittsburgh, PA\\[0.3cm]
April 1, 2014}

\end{center}
\end{titlepage}

\end{document}

### 2014-03-19 proceedings/tracks.tex
add a bunch of sigbovik proceedings stuff

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}

\newcommand\track[2]{\center {\Large \bf \em \sc Track #1} \vspace{0.2in}

{\huge \bf #2} \vspace{0.15in}
\thispagestyle{empty}}

\newcommand\asdf[2]{\vspace{0.15in} \item {\Large \bf \sf #1}

{\large \em #2}}
\newcommand\keywords[1]{

{\small {\bf Keywords:} #1}}

\begin{document}

\track{1}{Official 2012 Apocalypse Track}

\begin{enumerate}
\asdf{An Eschatological Approach: The Four Horsemen of Computer Science}
{The Reverend Nicolas A. Feltman and Joan E. Chao, Mother Supremum}
\keywords{Eschatology, Horsies, Indulgences}
\asdf{World Domination Through the Use of a Graphical Reduction of the Six Degrees of Separation Concept with Potential Robot Design for Mode of Implementation}
{L. S. Berg and P. J. Barton}
\keywords{Six Degrees of Separation, Kevin Bacon, Turtle Power, World Domination, Bad Cartoon Jokes, Topology, Reduction, Depression, Graphical Model Abuse}
\end{enumerate}

\newpage

\track{2}{Comestibility theory}

\begin{enumerate}
\asdf{Food for Thought: Dining Philosophers}{}
\keywords{hunger-driven devices, morsels, monads}
\asdf{Higher-Order Generalized Algebraic Pizzas}
{Brandon Bohrer and Samir Jindel}
\keywords{Curry-Pepper Isomorphism, Grease Monad, Monoids in the Category of Delicious, $A^\flat$ Mixolydian}
\vspace{-0.1in}
\asdf{Lollipops and Lemma Drops: the sweet, sweet logic of candy}
{William Lovas}
\asdf{Algorithms for k/n Power-Hours}
{Ben Blum, Dr. William Lovas, Ph.D., Chris Martens, and Dr. Tom Murphy VII, Ph.D.}
\keywords{alcohol in computer science, algorithms for the real world, chemically-assisted reasoning, drinking game theory, hyper-driven devices}
\end{enumerate}

\newpage

\track{3}{Brought to you by the letter\dots}

\begin{enumerate}
\setcounter{enumi}{-1}
\asdf{TBD}
{Taus Brock-Nannestad and Gian Perrone}
\keywords{all is well here, send money, love to you and yours}
\asdf{The Letter}
{Frederick J. Mesp\"{u}tchy}
\asdf{Proof of P = NP}
{Samir Jindel and Brandon Bohrer}
\asdf{An Epistolary Reconstruction of the Curry-Howard Correspondence}
{Ben Blum and Michael Sullivan}
\keywords{Corre-Howard-Spondence, Simply Typed Lambada Calculus, Supernatural Deduction}
\asdf{The Kardashian Kernel}
{David F. Fouhey and Daniel Maturana}
\keywords{Kardashian, Kim, Kourtney, Khloe, Kernel, Kuadratic, Konvex, Koncave, Krylov, Kronecker, Kolmogorov, Karush-Kuhn-Tucker, K-Means, K-Armed-Bandit, Kohonen, Karhunen-Loeve, Kriging, Kalman, Kinect, Kovariance, Kurse-of-dimensionality, Kurvature, K-Nearest-Neighbor}
\end{enumerate}

\newpage

\track{4}{Did you bring enough to reshare \\ with the class?}

\begin{enumerate}
\setcounter{enumi}{-1}
\asdf{Implications of Constrained Thought Expression Achieved via a One Hundred-forty Character Message Limitation Applied to Complex Social Netwo}
{Nathan Brooks and Tommy Liu}
\keywords{\#socialmedia, \#sm, \#140}
\asdf{The Spineless Tagless Tweet Machine: Distributed Cloud-Based Social Crowdsourced Lazy Graph Reduction on the Web 2.0}
{Michael Sullivan}
\asdf{SIGBOVIK 2012 Take-Home Midterm Examination}
{James McCann}
\asdf{The National Month Of Pushing Spacebar}
{Tom Murphy VII}
\keywords{C++harles D:{\textbackslash}ickens, fanfic, repetitive stress injury, 2012: A Space Odyssey}
\asdf{What Most Medical Students Know About Computer Science}
{Brian R. Hirshman}
\keywords{computer science, electronic medical records, all-knowing doctors, important stuff, take two kilobytes and see me in the morning, dude we're too busy learning other things}
\end{enumerate}

\newpage

\track{5}{Programming languages research and \\ other games for children}
\begin{enumerate}
\asdf{Modeling Perceived Cuteness}
{Nathan Brooks and Evelyn Yarzebinski}
\keywords{cute, puppy, kitten, turtle, spider, snake, dog, cat, gecko, hamster}
\asdf{i-PHONE app stor : where is my pants}
{Dr. Tom Murphy VII, Ph.D. and Dr. Abigale G. Lade, Ph.D.}
\keywords{pants, where, is my, juvenilia}
\asdf{An Extensible Platform for Upwards and Sidewards Mobility}
{David Renshaw}
\asdf{A modern optimizing compiler for a dynamically typed programming \\ language: Standard ML of New Jersey (preliminary report)}
{Ian Zerny}
\asdf{Programming Language Checklist}
{Colin McMillen, Jason Reed, and Elly Jones}
\end{enumerate}

\newpage

\track{6}{Protect yo' shit}
\begin{enumerate}
\asdf{Cryptographically-Strong Hashing with Physics}
{James McCann and Fake Otherauthor}
\asdf{The Physics-Based Hashes of Twelve Really Good Ideas}
{Patent Troll}
\asdf{A modest proposal for the purity of programming}
{Robert J. Simmons and Tom Murphy VII}
\keywords{the next 700 terrible programming languages, hyper-driven devices, types}
\asdf{Preparation-Hom as the ideal completion of a Hemorrhoidal category}
{Sitsin C{\o}mfort and Minr{\o}v G{\o}rondt}
\keywords{hemorrhoids, fistulads, co-fistulads, push outs}
\asdf{A Randomized Commit Protocol for Adversarial - Nay, Supervillain - \\ Workloads}
{Ben Blum}
\keywords{two-phase, two-face, randomized algorithms, acid}
\asdf{Address space content randomization: exploit mitigation through data \\ randomization}
{Carlo Angiuli}
\keywords{security, optimization, nondeterministic memory, lower is better}
\end{enumerate}

\end{document}

### 2014-03-20 proceedings/message-from-committee.tex
finish message 2014

@@
 {\large \bf \sffamily Multiple-Choice Section}
-
 \begin{enumerate}
@@
 Observe that SIGBOVIK offers a venue for three times as many different types of research as ``mainstream'' conferences.
-We won't belabour the obvious conclusion that SIGBOVIK is superior, although, just saying, you know; rather, we are interested in distinguishing among SIGBOVIK's three major categories of research.
+We won't belabour the obvious conclusion that SIGBOVIK is superior, although, you know, just getting that out there, we were all thinking it; rather, we are interested in distinguishing among SIGBOVIK's three major categories of research.
 \begin{enumerate}
@@
 
-The results of our survey are shown in Figure~\ref{fig:results} \sout{below}on the next page.
+\sout{I} We surveyed the proceedings of past SIGs BOVIK \sout{the night before finalizing the proceedings and sending them off to Lulu} like responsible researchers.
+Our methodology is definitely completely immune to any biases that might be caused by the survey not being blinded in any way or by it being conducted by only one person, and {\em definitely} free from any bias related to the study subject knowing in advance which conclusions would be drawn.
+%Moving right along.
+The results of our survey are shown in Figures~\ref{fig:results} and~\ref{fig:results2} \sout{below} on the next page.
 
@@
 \hline
-Category 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+Humorous treatment of humorous idea & 32 & 20 & 30 & 33 & 16 & 17 & 15 & 16 \\ \hline
+Humorous treatment of serious idea  & 2  & 0  & 1  & 0  & 0  & 1  & 2  & 0 \\ \hline
+Serious treatment of humorous idea  & 2  & 4  & 3  & 0  & 4  & 6  & 6  & 8 \\ \hline
+Other                               & 1  & 1  & 5  & 3  & 0  & 2  & 2  & 0 \\ \hline
 \hline
-Category 2 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-Category 3 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-Other & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-\hline
-\hline
-Total publications & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+Total publications                  & 37 & 25 & 39 & 36 & 20 & 26 & 25 & 24 \\
 
@@
 \end{center}
-\caption{Results.}
+\caption{Results.
+The `other' category comprises submissions that defy my perfect classification scheme, including cryptic iconography, video games, comics, choose your own adventures, and take-home midterm examinations.}
 \label{fig:results}
@@
 
+\begin{figure}[h]
+\begin{center}
+\begin{tabular}{r|c|c|c|c|c|c|c|c|}
+& \multicolumn{8}{c|}{SIGBOVIK year of incidence} \\
+& 2007 & 2008 & 2009 & 2010 & 2011 & 2012 & 2013 & 2014 \\
+\hline
+\hline
+Meta-humor about research
+& 7  & 2  & 8  & 8  & 8  & 1  & 2  & 2  \\ \hline
+Type theory or programming jokes
+& 8  & 3  & 1  & 7  & 2  & 1  & 2  & 4  \\ \hline
+Puns
+& 4  & 4  & 8  & 1  & 0  & 0  & 2  & 3  \\ \hline
+Poop or dick jokes
+& 1  & 0  & 1  & 1  & 0  & 1  & 1  & 2  \\ \hline
+Politics, economics, or patents jokes
+& 0  & 0  & 0  & 0  & 1  & 2  & 2  & 1  \\ \hline
+Social media jokes
+& 0  & 0  & 0  & 2  & 2  & 1  & 0  & 0  \\ \hline
+%\hline
+%Food jokes
+%& 1  & 0  & 2  & 1  & 0  & 3  & 0  & 0  \\ \hline
+%NP completeness jokes
+%& 1  & 1  & 0  & 0  & 0  & 1  & 0  & 0  \\ \hline
+%Graph jokes
+%& 1  & 0  & 0  & 1  & 0  & 0  & 0  & 0  \\ \hline
+%Pop culture jokes
+%& 1  & 1  & 0  & 0  & 0  & 2  & 2  & 0  \\ \hline
+%Computational impossibilities
+%& 0  & 1  & 4  & 4  & 1  & 1  & 0  & 0  \\ \hline
+%Cat jokes
+%& 0  & 1  & 2  & 0  & 1  & 1  & 0  & 0  \\ \hline
+%Images / visual gags
+%& 1  & 1  & 0  & 1  & 0  & 1  & 1  & 0  \\ \hline
+%Robot uprisings, apocalypse, and the supernatural
+%& 3  & 2  & 0  & 3  & 0  & 2  & 1  & 1  \\ \hline
+%Sex jokes
+%& 0  & 0  & 0  & 0  & 1  & 0  & 0  & 0  \\ \hline
+%Pittsburgh local humour
+%& 0  & 0  & 1  & 1  & 0  & 0  & 0  & 3  \\ \hline
+%Other
+%& 4  & 4  & 2  & 1  & 0  & 0  & 0  & 0  \\ \hline
+Other
+& 12 & 11 & 11 & 12 & 3  & 11 & 4  & 4  \\ \hline
+
+\end{tabular}
+\end{center}
+\caption{Breakdown among category-1 publications of subject of humor.
+Among this new `other' category dwell comestibility theory, NP-completeness, cat pictures, computational impossibilities, the supernatural, and robot uprisings and other apocalypse situations.}
+\label{fig:results2}
+\end{figure}
+
+A number of trends are evident.
+First and foremost is the decline over time of the proportion of category-1 submissions and a corresponding increase in submissions accompanied by an actual artefact that must have been put together with blood, sweat, and so forth, rather than just the cushy endeavour of writing down shallow drivel in LaTeX and calling that a ``research paper''.
+Go us!
+Just kidding.
+We value all SIGBOVIK submissions highly, \sout{except for} especially yours.
+But really.
+
+Second, we note the prominence among category-1 submissions of meta-research papers; that is, papers about writing papers, giving talks, or improving productivity.
+It is unsurprising that this is a popular subject for obvious reasons.
+Finally, we note a modest decline in the proportion of programming language papers, which we attribute to SIGBOVIK's audience's interests expanding beyond the core of its type-theoretician founding mothers and fathers, to include such newfangled topics as Twitterers and Bitdollars.
+
+Without further ado, adon't, or adonuts, the proceedings of your SIGBOVIK \#8.
+
 \end{document}

### 2014-03-20 review-2014/review-5.tex
finish mostly tracks and toc

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{2}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 2014 Paper Review}} \\ &\\
& \LARGE{Paper 5: Yet Another Application of Our Pet Technique} \\
&\\
\hline
\end{tabular}}
\vspace{2em}

{\large\bf
\begin{tabular}{l}
Robert Marsh, King Under the Mountain \\
Rating: 1 (strong reject) \\
Confidence: absolute/4 \\
\end{tabular}}
\vspace{1em}

There are several problems with this paper.
The modeling of vehicles is rather poor - at the very least their acceleration in the road's reference frame is limited by relativistic effects.
Furthermore, modern vehicles are curvy, not dumb and boxy (except the Nissan Cube, which isdumb, boxy, and curvy, truly a marvel of modern engineering) and thus would be best modeled as spheres.
In addition, the alternatives they test against don't include the current industry standard dime-an-hour hick, suggesting that they were cherrypicked to appear more favorable.
And besides, they didn't cite \emph{anything} I did, so I don't see why I should let them get published.

\end{document}

### 2014-03-20 review-2014/review_9.tex
finish mostly tracks and toc

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{2}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 2014 Paper Review}} \\ &\\
& \LARGE{Paper $9$: A Simple Category-Theoretic ...} \\
&\\
\hline
\end{tabular}}

{\large\bf
\begin{tabular}{l}
Reviewed by: Ed Morehouse (Carnegie Mellon University) \\
\end{tabular}}

	``A Simple Category-Theoretic Understanding of Category-Theoretic Diagrams'' by ``Stefan Muller''
	is a work of incomprehensible abstract nonsense,
	and as such, constitutes a valuable contribution to the field of category theory.

	The article seems to be about understanding category
	theory through the graphical language of commutative diagrams.
	I say, ``seems to be'', because after gleaning this much from the abstract,
	I decided to test (what I assume to be) the article's central thesis
	by attempting to understand the article itself by just looking at the pictures.

	I infer that the article begins by describing the basic properties
	of morphisms in a category under composition and identity
	 -- although it could be talking about something else entirely, it's kind of hard to say.
%	In figure $3$, I discovered the first major flaw of the paper,
%	here ``Muller'' uses applicative-order notation for composition.
%	It has been well-known since the foundational work of Mac Lane that this is an intellectual dead-end.
	The next bit seems like it might be about constructing a category
	whose objects are themselves commutative diagrams and whose morphisms are affine transformations of such%
	\footnote
	{
		note to self: ask ``Muller'' what the paper is really about, then reject it and steal his idea.
	}.

	The paper ends by constructing the ``Morehouse-Sierpinski $\omega$-category of commuting squares''
	(let's call it).
	Although ``Muller'' has (for all I know) now solved a long-standing open problem in higher dimensional category theory
	by giving a finitary, constructive presentation of this important category,
	there is still currently no type-theoretic interpretation (that I could think of in 5 minutes),
	nor are current \LaTeX \, diagram-description packages
	adequate for drawing $\omega$-dimensional commutative diagrams.
	I'll just assume that ``Muller'' leaves these issues for future work.

\begin{tabular}{l}
{\large\bf Rating:} I give this article the terminal rating in the co-op of the bicategory of journal reviews. \\
{\large\bf Confidence:} What am I supposed to put here? \\
	Why are the instructions not presented in picture-form, like the ones for assembling my bookshelf: \\
\end{tabular}
\vspace{-1em}
\begin{center}
	{\includegraphics[width=2.5in]{instructions.jpg}}
\end{center}
\end{document}

### 2014-03-20 proceedings/table-of-contents.tex

\documentclass[letter]{article}
\special{papersize=8.5in,11in}
\usepackage{fullpage}
\usepackage{fontspec}
\usepackage{amsfonts}

\setmainfont{TragicSans}

\newcommand\track[2]{{\Large Track #1: #2}}

\newcommand\paper[2]{\item \sf #1 \dotfill TODO}

\newcommand\keywords[1]{}

\begin{document}

\begin{center}
{\Huge You Won't Believe This Table of Contents}
\end{center}
\thispagestyle{empty}

\vspace{6em}

\track{8}{Take The Quiz: Which Pointless Time-Wasting Amusement are You?}
%\track{8}{How to Keep Adults Busy and Other Fun Games for Children}

\begin{enumerate}
\paper{How to Keep a Graduate Student Busy}
{Paul Stansifer}
\keywords{all, work, and, no, play, makes, paul, a, dull, boy}
\paper{New Results in $k/n$ Power-Hours}
{Dr. Tom Murphy VII Ph.D.}
\keywords{generalized binge drinking, maths, finite-state automata, abstract interpretation}
\paper{Linear Logic Free}
{Carlo Angiuli}
\keywords{linear logic, free to play, app store}
\paper{It Still Seems That Black Has Hope in These Extremely Unfair Variants of Chess}
{Dr. Tom Murphy VII Ph.D., Ben Blum, and Jim McCann, in italics ...... oh! .... nervous laughter .... no, that's not part of the name still. I stopped saying the name. Why are you still typing? This isn't the name any more. End quote.}
\keywords{chess, stratego, guess who, clue, candy crush saga, game-tree search, combinations}
\end{enumerate}

\vspace{4em}

\track{88}{Epic Local News Stories You'll Be Sad You Missed}
%\track{88}{Concerning Local Affairs}

\begin{enumerate}
\paper{PARTIAL TRANSCRIPT OF DEAN'S SPEECH TO THE FACULTY OF THE MELLON COLLEGE OF SCIENCES}
{Clint Falamenos}
\keywords{news, lane center, college of sciences}
\paper{Analysis of the Effects of Substantial Lane Closure During Afternoon Peak Along a Heavily-Traveled Urban Arterial Choke-Point}
{Nicholas Fudala}
\keywords{knowledge, pagers, RATPAG}
\paper{Yet Another Application of Our Pet Technique}
{Chester Francis and Nicholas Fudala}
\keywords{complexity, multiple patents pending, nobel peace prize}
\paper{Please Don't Let Open House Destroy the Universe}
{Jenn Landefeld}
\keywords{open house, destruction of universe, herding cats}
\end{enumerate}

\vspace{4em}

\track{888}{Four Unbelievable New Theoretical Perspectives}
%\track{888}{New Theoretical Perspectives}

\begin{enumerate}
%\setcounter{enumi}{-1}
\paper{Heterotopy Type Theory: A Defense of the Traditional Foundations of Mathematics}
{Thomas Cramer and Sam Yeager}
\keywords{heterotopy type theory, identity type, freedom of expression axiom}
\paper{The Dumping Lemma - Assessing Regularity}
{Naomi Saphra}
\keywords{computational scatology, formal languages, poop jokes}
\paper{Statistical Watch Optimization}
{Bryce Summers}
\keywords{Watch, clock, time, probability, statistics}
\paper{A Simple Category-Theoretic Understanding of Category-Theoretic Diagrams}
{Stefan Muller}
\keywords{category, diagram, functor, diagram diagram}
\end{enumerate}

\vspace{4em}

\track{8,888}{The Three Most Important Ways to Stay Safe This Research Season}
%\track{8,888}{Stay Safe Out There}

\begin{enumerate}
%\setcounter{enumi}{-1}
\paper{Cryptographically-Sound Jokes}
{Jim McCann}
\keywords{jokes, sha, md5}
\paper{DollarCoin: A Cryptocurrency with Proof-of-Dollar}
{Patrick J. Xia}
\keywords{cryptocurrency,
bitcoin,
dollars,
bills,
gamechanger,
disrupt,
awesome,
supercool,
thebest,
hashtag}
\paper{Towards a Completely Autonomous Vehicle}
{Nicholas Fudala and Chester Francis}
\keywords{pain, suffering, frustration}
\end{enumerate}

\vspace{4em}

\track{88,888}{Improve Your Own Research With These Five Weird Tricks}
%\track{88,888}{Easier, Better, Faster, Stronger Research Methodologies}
\begin{enumerate}
\paper{Belief-Sustaining Inference}
{Alex Reinhart and Jerzy Wieczorek}
\keywords{uninformative likelihood,
statistics,
embarrassment,
cognitive dissonance,
inference}
\paper{Solutions to Ley Line Access in Occult Computing}
{Maija Mednieks}
\keywords{ley lines,
occult computing,
miskatonic university,
installing linux on a dead badger,
playing god}
\paper{Measuring Your $\mathbb{P}$-ness: \\ Determining How Like a Probability Distribution a Given Mathematical Object Is}
{John T. Longwood}
\keywords{lowbrow humour,
probability,
statistics}
\paper{a clarification of terminology}
{David Renshaw}
\keywords{did you mean,
search instead for,
not to be confused with}
\paper{What, If Anything, is Epsilon?}
{Dr. Tom Murphy VII Ph.D.}
\keywords{computational archaeology,
epsilon,
very-small and medium-small numbers}
\end{enumerate}

\vspace{4em}

\track{888,888}{These New Ways to Program Will Blow Your Mind}
%\track{888,888}{Novel Programming Methodologies}
\begin{enumerate}
\paper{IRS-1040-Based Computing: The Untapped Computational Power of Bureaucracy}
{William Gunther and Brian Kell}
\keywords{unconventional computation,
taxes,
compilers,
bureaucracy}
\paper{Towards Fooling and Foiling the NSA: \\
Surveying the State of Steganographic Programming Languages}
{Wolfgang Richter and Debjani Biswas}
\keywords{steganography,
esoteric programming language,
esoteric,
programming languages,
obfuscation,
cryptography,
shit}
\paper{Pikachu, Domosaur, and other Monolexical Languages}
{Sarah Allen, Jesse Dodge, and Domosaur Domosaur}
\keywords{natural language processing,
regular languages,
translation}
\paper{Unit-Test-Based Programming}
{Miguel \'{A}. Lech\'{o}n}
\keywords{unit test,
automatic stubbornness,
holy grail}
\end{enumerate}
\thispagestyle{empty}

\end{document}

### 2014-03-20 proceedings/tracks.tex

@@
 \usepackage{fullpage}
+\usepackage{fontspec}
+\usepackage{amsfonts}
+
+\setmainfont{Rumsey}
+\setsansfont{TragicSans}
 
@@
 
-\newcommand\asdf[2]{\vspace{0.15in} \item {\Large \bf \sf #1}
+\newcommand\paper[2]{\vspace{0.15in} \item {\Large \bf \sf #1}
 
-{\large \em #2}}
+{\large \em \sf #2}}
 \newcommand\keywords[1]{
@@
 
-\track{1}{Official 2012 Apocalypse Track}
+%\track{8}{Take The Quiz: Which Pointless Time-Wasting Amusement are You?}
+\track{8}{How to Keep Adults Busy and Other Fun Games for Children}
 
 \begin{enumerate}
-\asdf{An Eschatological Approach: The Four Horsemen of Computer Science}
-{The Reverend Nicolas A. Feltman and Joan E. Chao, Mother Supremum}
-\keywords{Eschatology, Horsies, Indulgences}
-\asdf{World Domination Through the Use of a Graphical Reduction of the Six Degrees of Separation Concept with Potential Robot Design for Mode of Implementation}
-{L. S. Berg and P. J. Barton}
-\keywords{Six Degrees of Separation, Kevin Bacon, Turtle Power, World Domination, Bad Cartoon Jokes, Topology, Reduction, Depression, Graphical Model Abuse}
+\paper{How to Keep a Graduate Student Busy}
+{Paul Stansifer}
+\keywords{all, work, and, no, play, makes, paul, a, dull, boy}
+\paper{New Results in $k/n$ Power-Hours}
+{Dr. Tom Murphy VII Ph.D.}
+\keywords{generalized binge drinking, maths, finite-state automata, abstract interpretation}
+\paper{Linear Logic Free}
+{Carlo Angiuli}
+\keywords{linear logic, free to play, app store}
+\paper{It Still Seems That Black Has Hope in These Extremely Unfair Variants of Chess}
+{Dr.
+Tom Murphy VII Ph.D., Ben Blum, and Jim McCann, in italics ......
+oh!
+....
+nervous laughter ....
+no, that's not part of the name still.
+I stopped saying the name.
+Why are you still typing?
+This isn't the name any more.
+End quote.}
+\keywords{chess, stratego, guess who, clue, candy crush saga, game-tree search, combinations}
 \end{enumerate}
@@
 
-\track{2}{Comestibility theory}
+%\track{88}{Epic Local News You'll Be Sad You Missed}
+\track{88}{Concerning Local Affairs}
 
 \begin{enumerate}
-\asdf{Food for Thought: Dining Philosophers}{}
-\keywords{hunger-driven devices, morsels, monads}
-\asdf{Higher-Order Generalized Algebraic Pizzas}
-{Brandon Bohrer and Samir Jindel}
-\keywords{Curry-Pepper Isomorphism, Grease Monad, Monoids in the Category of Delicious, $A^\flat$ Mixolydian}
-\vspace{-0.1in}
-\asdf{Lollipops and Lemma Drops: the sweet, sweet logic of candy}
-{William Lovas}
-\asdf{Algorithms for k/n Power-Hours}
-{Ben Blum, Dr. William Lovas, Ph.D., Chris Martens, and Dr. Tom Murphy VII, Ph.D.}
-\keywords{alcohol in computer science, algorithms for the real world, chemically-assisted reasoning, drinking game theory, hyper-driven devices}
+\paper{PARTIAL TRANSCRIPT OF DEAN'S SPEECH TO THE FACULTY OF THE MELLON COLLEGE OF SCIENCES}
+{Clint Falamenos}
+\keywords{news, lane center, college of sciences}
+\paper{Analysis of the Effects of Substantial Lane Closure During Afternoon Peak Along a Heavily-Traveled Urban Arterial Choke-Point}
+{Nicholas Fudala}
+\keywords{knowledge, pagers, RATPAG}
+\paper{Yet Another Application of Our Pet Technique}
+{Chester Francis and Nicholas Fudala}
+\keywords{complexity, multiple patents pending, nobel peace prize}
+\paper{Please Don't Let Open House Destroy the Universe}
+{Jenn Landefeld}
+\keywords{open house, destruction of universe, herding cats}
 \end{enumerate}
@@
 
-\track{3}{Brought to you by the letter\dots}
+%\track{888}{Four Unbelievable New Theoretical Perspectives}
+\track{888}{New Theoretical Perspectives}
 
 \begin{enumerate}
-\setcounter{enumi}{-1}
-\asdf{TBD}
-{Taus Brock-Nannestad and Gian Perrone}
-\keywords{all is well here, send money, love to you and yours}
-\asdf{The Letter}
-{Frederick J. Mesp\"{u}tchy}
-\asdf{Proof of P = NP}
-{Samir Jindel and Brandon Bohrer}
-\asdf{An Epistolary Reconstruction of the Curry-Howard Correspondence}
-{Ben Blum and Michael Sullivan}
-\keywords{Corre-Howard-Spondence, Simply Typed Lambada Calculus, Supernatural Deduction}
-\asdf{The Kardashian Kernel}
-{David F. Fouhey and Daniel Maturana}
-\keywords{Kardashian, Kim, Kourtney, Khloe, Kernel, Kuadratic, Konvex, Koncave, Krylov, Kronecker, Kolmogorov, Karush-Kuhn-Tucker, K-Means, K-Armed-Bandit, Kohonen, Karhunen-Loeve, Kriging, Kalman, Kinect, Kovariance, Kurse-of-dimensionality, Kurvature, K-Nearest-Neighbor}
+%\setcounter{enumi}{-1}
+\paper{Heterotopy Type Theory: \\ A Defense of the Traditional Foundations of Mathematics}
+{Thomas Cramer and Sam Yeager}
+\keywords{heterotopy type theory, identity type, freedom of expression axiom}
+\paper{The Dumping Lemma - Assessing Regularity}
+{Naomi Saphra}
+\keywords{computational scatology, formal languages, poop jokes}
+\paper{Statistical Watch Optimization}
+{Bryce Summers}
+\keywords{Watch, clock, time, probability, statistics}
+\paper{A Simple Category-Theoretic Understanding of Category-Theoretic Diagrams}
+{Stefan Muller}
+\keywords{category, diagram, functor, diagram diagram}
 \end{enumerate}
@@
 
-\track{4}{Did you bring enough to reshare \\ with the class?}
+%\track{8,888}{Three Important Ways to Stay Safe This Research Season}
+\track{8,888}{Stay Safe Out There}
 
 \begin{enumerate}
-\setcounter{enumi}{-1}
-\asdf{Implications of Constrained Thought Expression Achieved via a One Hundred-forty Character Message Limitation Applied to Complex Social Netwo}
-{Nathan Brooks and Tommy Liu}
-\keywords{\#socialmedia, \#sm, \#140}
-\asdf{The Spineless Tagless Tweet Machine: Distributed Cloud-Based Social Crowdsourced Lazy Graph Reduction on the Web 2.0}
-{Michael Sullivan}
-\asdf{SIGBOVIK 2012 Take-Home Midterm Examination}
-{James McCann}
-\asdf{The National Month Of Pushing Spacebar}
-{Tom Murphy VII}
-\keywords{C++harles D:{\textbackslash}ickens, fanfic, repetitive stress injury, 2012: A Space Odyssey}
-\asdf{What Most Medical Students Know About Computer Science}
-{Brian R. Hirshman}
-\keywords{computer science, electronic medical records, all-knowing doctors, important stuff, take two kilobytes and see me in the morning, dude we're too busy learning other things}
+%\setcounter{enumi}{-1}
+\paper{Cryptographically-Sound Jokes}
+{Jim McCann}
+\keywords{jokes, sha, md5}
+\paper{DollarCoin: A Cryptocurrency with Proof-of-Dollar}
+{Patrick J. Xia}
+\keywords{cryptocurrency,
+bitcoin,
+dollars,
+bills,
+gamechanger,
+disrupt,
+awesome,
+supercool,
+thebest,
+hashtag}
+\paper{Towards a Completely Autonomous Vehicle}
+{Nicholas Fudala and Chester Francis}
+\keywords{pain, suffering, frustration}
 \end{enumerate}
@@
 
-\track{5}{Programming languages research and \\ other games for children}
+%\track{88,888}{Improve Your Own Research With These Five Weird Tricks}
+\track{88,888}{Easier, Better, Faster, Stronger Research Methodologies}
 \begin{enumerate}
-\asdf{Modeling Perceived Cuteness}
-{Nathan Brooks and Evelyn Yarzebinski}
-\keywords{cute, puppy, kitten, turtle, spider, snake, dog, cat, gecko, hamster}
-\asdf{i-PHONE app stor : where is my pants}
-{Dr. Tom Murphy VII, Ph.D. and Dr. Abigale G. Lade, Ph.D.}
-\keywords{pants, where, is my, juvenilia}
-\asdf{An Extensible Platform for Upwards and Sidewards Mobility}
+\paper{Belief-Sustaining Inference}
+{Alex Reinhart and Jerzy Wieczorek}
+\keywords{uninformative likelihood,
+statistics,
+embarrassment,
+cognitive dissonance,
+inference}
+\paper{Solutions to Ley Line Access in Occult Computing}
+{Maija Mednieks}
+\keywords{ley lines,
+occult computing,
+miskatonic university,
+installing linux on a dead badger,
+playing god}
+\paper{Measuring Your $\mathbb{P}$-ness: Determining How Like a Probability Distribution a Given Mathematical Object Is}
+{John T. Longwood}
+\keywords{lowbrow humour,
+probability,
+statistics}
+\paper{a clarification of terminology}
 {David Renshaw}
-\asdf{A modern optimizing compiler for a dynamically typed programming \\ language: Standard ML of New Jersey (preliminary report)}
-{Ian Zerny}
-\asdf{Programming Language Checklist}
-{Colin McMillen, Jason Reed, and Elly Jones}
+\keywords{did you mean,
+search instead for,
+not to be confused with}
+\paper{What, If Anything, is Epsilon?}
+{Dr. Tom Murphy VII Ph.D.}
+\keywords{computational archaeology,
+epsilon,
+very-small and medium-small numbers}
 \end{enumerate}
@@
 
-\track{6}{Protect yo' shit}
+%\track{888,888}{These New Ways to Program Will Blow Your Mind}
+\track{888,888}{Innovations in Natural and Unnatural Languages}
 \begin{enumerate}
-\asdf{Cryptographically-Strong Hashing with Physics}
-{James McCann and Fake Otherauthor}
-\asdf{The Physics-Based Hashes of Twelve Really Good Ideas}
-{Patent Troll}
-\asdf{A modest proposal for the purity of programming}
-{Robert J. Simmons and Tom Murphy VII}
-\keywords{the next 700 terrible programming languages, hyper-driven devices, types}
-\asdf{Preparation-Hom as the ideal completion of a Hemorrhoidal category}
-{Sitsin C{\o}mfort and Minr{\o}v G{\o}rondt}
-\keywords{hemorrhoids, fistulads, co-fistulads, push outs}
-\asdf{A Randomized Commit Protocol for Adversarial - Nay, Supervillain - \\ Workloads}
-{Ben Blum}
-\keywords{two-phase, two-face, randomized algorithms, acid}
-\asdf{Address space content randomization: exploit mitigation through data \\ randomization}
-{Carlo Angiuli}
-\keywords{security, optimization, nondeterministic memory, lower is better}
+\paper{IRS-1040-Based Computing: The Untapped Computational Power of Bureaucracy}
+{William Gunther and Brian Kell}
+\keywords{unconventional computation,
+taxes,
+compilers,
+bureaucracy}
+\paper{Towards Fooling and Foiling the NSA: \\
+Surveying the State of Steganographic Programming Languages}
+{Wolfgang Richter and Debjani Biswas}
+\keywords{steganography,
+esoteric programming language,
+esoteric,
+programming languages,
+obfuscation,
+cryptography,
+shit}
+\paper{Pikachu, Domosaur, and other Monolexical Languages}
+{Sarah Allen, Jesse Dodge, and Domosaur Domosaur}
+\keywords{natural language processing,
+regular languages,
+translation}
+\paper{Unit-Test-Based Programming}
+{Miguel \'{A}. Lech\'{o}n}
+\keywords{unit test,
+automatic stubbornness,
+holy grail}
 \end{enumerate}

### 2014-03-20 review-2014/review-1.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{fullpage}
\usepackage{times}
\setlength\parindent{0pt}
\setlength\parskip{12pt}

\begin{document}

{\sffamily
\begin{tabular}{ll}
\multirow{2}{*}{\includegraphics[width=1in]{ach.png}}\\
& \textbf{\Huge{SIGBOVIK 2014 Paper Review}} \\ &\\
& \LARGE{Paper 1: New Results in $k/n$ Power-Hours} \\
&\\
\hline
\end{tabular}}
\vspace{2em}

{\large\bf
\begin{tabular}{l}
Robert Marsh, King Under the Mountain \\
Rating: 3 (weak accept) \\
Confidence: 2/4 \\
\end{tabular}}
\vspace{1em}

Man, figure 5 does look really sweet.  like whoa.  you could take that and trim the edges off and use it as a boss in some kind of diagonal-scrolling space shoot-em-up.  no really, that's blowing my mind.  you could, like make an entire space invaders galaga thingy out of these charts.  it would be rad as all get out.

On the other hand, one must consider the broader implications of research of this sort.
In particular, should this conference encourage an attempt at a $k/n$ power hour for any $k$, even a difficult-to-reach one like $\langle 11, 53\rangle$?  
Let us first note that the finer things in life should be savored.
However, there is such a thing as Beer 30 Light, and people are going to attempt such things whether we help them or not, so we might as well ensure they have the tools they need for slightly more responsible drinking.

In sum, this paper's significant contributions to cool fractally things seems to outweigh its potential harm to society.

\end{document}

### 2014-03-20 proceedings/table-of-contents.tex
finish proceedings for this year & write knowledge in readme

@@
 
-\newcommand\paper[2]{\item \sf #1 \dotfill TODO}
+\newcommand\paper[2]{\item \sf #1 \dotfill #2}
 
@@
 \paper{How to Keep a Graduate Student Busy}
-{Paul Stansifer}
+{3}
 \keywords{all, work, and, no, play, makes, paul, a, dull, boy}
 \paper{New Results in $k/n$ Power-Hours}
-{Dr. Tom Murphy VII Ph.D.}
+{5}
 \keywords{generalized binge drinking, maths, finite-state automata, abstract interpretation}
 \paper{Linear Logic Free}
-{Carlo Angiuli}
+{17}
 \keywords{linear logic, free to play, app store}
 \paper{It Still Seems That Black Has Hope in These Extremely Unfair Variants of Chess}
-{Dr.
-Tom Murphy VII Ph.D., Ben Blum, and Jim McCann, in italics ......
-oh!
-....
-nervous laughter ....
-no, that's not part of the name still.
-I stopped saying the name.
-Why are you still typing?
-This isn't the name any more.
-End quote.}
+{21}
 \keywords{chess, stratego, guess who, clue, candy crush saga, game-tree search, combinations}
@@
 \paper{PARTIAL TRANSCRIPT OF DEAN'S SPEECH TO THE FACULTY OF THE MELLON COLLEGE OF SCIENCES}
-{Clint Falamenos}
+{29}
 \keywords{news, lane center, college of sciences}
 \paper{Analysis of the Effects of Substantial Lane Closure During Afternoon Peak Along a Heavily-Traveled Urban Arterial Choke-Point}
-{Nicholas Fudala}
+{33}
 \keywords{knowledge, pagers, RATPAG}
 \paper{Yet Another Application of Our Pet Technique}
-{Chester Francis and Nicholas Fudala}
+{35}
 \keywords{complexity, multiple patents pending, nobel peace prize}
 \paper{Please Don't Let Open House Destroy the Universe}
-{Jenn Landefeld}
+{41}
 \keywords{open house, destruction of universe, herding cats}
@@
 \paper{Heterotopy Type Theory: A Defense of the Traditional Foundations of Mathematics}
-{Thomas Cramer and Sam Yeager}
+{47}
 \keywords{heterotopy type theory, identity type, freedom of expression axiom}
 \paper{The Dumping Lemma - Assessing Regularity}
-{Naomi Saphra}
+{51}
 \keywords{computational scatology, formal languages, poop jokes}
 \paper{Statistical Watch Optimization}
-{Bryce Summers}
+{53}
 \keywords{Watch, clock, time, probability, statistics}
 \paper{A Simple Category-Theoretic Understanding of Category-Theoretic Diagrams}
-{Stefan Muller}
+{57}
 \keywords{category, diagram, functor, diagram diagram}
@@
 \paper{Cryptographically-Sound Jokes}
-{Jim McCann}
+{65}
 \keywords{jokes, sha, md5}
 \paper{DollarCoin: A Cryptocurrency with Proof-of-Dollar}
-{Patrick J. Xia}
+{67}
 \keywords{cryptocurrency,
@@
 \paper{Towards a Completely Autonomous Vehicle}
-{Nicholas Fudala and Chester Francis}
+{71}
 \keywords{pain, suffering, frustration}
@@
 \paper{Belief-Sustaining Inference}
-{Alex Reinhart and Jerzy Wieczorek}
+{77}
 \keywords{uninformative likelihood,
@@
 \paper{Solutions to Ley Line Access in Occult Computing}
-{Maija Mednieks}
+{83}
 \keywords{ley lines,
@@
 \paper{Measuring Your $\mathbb{P}$-ness: \\ Determining How Like a Probability Distribution a Given Mathematical Object Is}
-{John T. Longwood}
+{87}
 \keywords{lowbrow humour,
@@
 \paper{a clarification of terminology}
-{David Renshaw}
+{91}
 \keywords{did you mean,
@@
 \paper{What, If Anything, is Epsilon?}
-{Dr. Tom Murphy VII Ph.D.}
+{93}
 \keywords{computational archaeology,
@@
 \paper{IRS-1040-Based Computing: The Untapped Computational Power of Bureaucracy}
-{William Gunther and Brian Kell}
+{101}
 \keywords{unconventional computation,
@@
 Surveying the State of Steganographic Programming Languages}
-{Wolfgang Richter and Debjani Biswas}
+{103}
 \keywords{steganography,
@@
 \paper{Pikachu, Domosaur, and other Monolexical Languages}
-{Sarah Allen, Jesse Dodge, and Domosaur Domosaur}
+{109}
 \keywords{natural language processing,
@@
 \paper{Unit-Test-Based Programming}
-{Miguel \'{A}. Lech\'{o}n}
+{115}
 \keywords{unit test,

### 2015-03-16 12color/paper.tex
write 12color paper

\documentclass{article}
\special{papersize=8.5in,11in}

\usepackage[lf]{electrum}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{subcaption}

\title{A Proof of the Twelve Color Theorem: \\ Hey, Nobody Bothered Before. \\ {\small Sixth Grade Class Project for Mr. Blum}}
\author{Y. T. Boveck \and Victoria V. Varg-Mack, VI}
\date{April 1$\tfrac{1}{3}$, 2071}

\begin{document}

\maketitle

\thispagestyle{empty}
\begin{center}
\begin{tabular}{p{3.9in}}
\fontfamily{cmr}\selectfont
\textit{Editor's note: Several days ago, the wreckage of a small time machine appeared in the program committee's office,
containing apparently only a copy of the proceedings from SIGBOVIK 2071.
Unfortunately, all papers but one were burnt beyond recognition.
Current speculation holds that the time machine operators forgot to disable the paradox safety interlock, and all the important (potentially causality-violating) papers were destroyed, leaving only this drivel. We're publishing it anyway, sorry.}
\end{tabular}
\end{center}

\begin{abstract}
	Roughly 100 years ago, mathematicians successfully proved the \textit{4-Color Theorem}.
	Widely regarded as one of the most important theorems in graph theory at the time, the 4-Color Theorem states that for all planar graphs there exists an assignment of colors to nobes, such that no two adjacent nobes are assigned the same color, with at most 4 colors used.
	However, modern-day computers are capable of rendering far more diverse color palettes, and the old 4-color limitation is now largely irrelevant.
	In this work we extend the old result to support 12-colorings, offer some thoughts on generalization, and leave a conspicuous hole in our proof to support our future work.
\end{abstract}

\section{Introduction}

Victoria said I had to write this section, even though she's the only one of us who actually knows how the 4-color theorem proof goes. But luckily I found a proof for the 5-color theorem that made sense, and since having more colors is obviously better, that should be even better. It goes like this:

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\begin{lemma}
All planar graphs have at least one vertex with at most 5 neighbors.
\end{lemma}

\begin{proof}
	Let $N(n)$ mean how many neighbors a given nobe $n$ has. Then because each edge connects two nobes, the sum of all nobes' neighbours $\sum_n N(n) = 2e$. Let's assume all nobes have at least six neighbours. Then $6n \le 2e$.

	Next let $E(f)$ mean the number of edges that make up the border of a given face $f$ of the graph. Since each edge borders 2 faces, $\sum_f E(f) = 2e$. Every region is bounded by at least 3 edges so that means $3f \le 2e$.

	If we plug all that into the ``Euler's Forumula'' that Mr. Blum gave us in class, which says that $n-e+f=2$, here's what we get: $n-e+f \le \frac{e}{3} - e + \frac{2e}{3} = 0$. But since $n-e+f = 2$ we end up with $2\le0$. Then we must have been wrong to assume all nobes have at least six neighbours. That's a ``contradiction''!
\end{proof}

\begin{theorem}
All planar graphs can be colored with no more than 5 colors so that no neighbor nobes have the same color.
\label{thm:5c}
\end{theorem}

\begin{proof}
	Let $G$ mean the graph we're trying to prove this for. Using the lemon I just proved, pick the nobe $n$ that has at most five neighbors and call those $n_1$, $n_2$, $n_3$, $n_4$, $n_5$ in whatever order you want. Let $G'$ mean the smaller graph you get by taking out $n$, which of course is still planar. Because $G'$ is smaller, we know with ``induction'' that $G'$ can be 5-colored. (That step seemed kinda weird to me but Victoria said it was legit so please don't take points off.)

	Now we're gonna assume $G$ cannot be 5-colored. For this to be true, then each $n_i$ from $1$ to $5$ must all have different colors, or else you could just pick the missing one for $n$. Let's think about the sub-graph that has only the 2 colors matching $n_1$ and $n_3$, call it $G_{13}$. If the graph is disjoint, meaning $n_1$ and $n_3$ are in two separate parts, then you would be able to re-color the graph to make $n_3$ have color $1$, so you can paint $n$ color $3$. Therefore $G_{13}$ must be connected. You can make the same argument for $G_{24}$ to be connected, but suddenly that doesn't make any sense, because a path that keeps $G_{24}$ connected would have to intersect the connected $G_{13}$!

	Therefore, even if the 5-coloring of $G'$ gives all different colors to $n_1$, $n_2$, $n_3$, $n_4$, and $n_5$, you should be able to re-color one of the $n_i$ so you can have a spare color to give to $n$. Therefore $G$ can be 5-colored.
\end{proof}

\section{The 12-Color Theorem}

\thispagestyle{empty}
In the previous section, Y.T. et al.~outlined the prior work in the field. Now we will make our main contribution; namely, to extend the 5-color theorem reviewed above to allow for 12-colorings of arbitrary planar graphs.

\begin{theorem}
	Given an arbitrary planar graph $G$, there exists an assignment of no more than 12 colors to the set of nobes $N(G)$ such that no two neighbouring nobes are assigned the same color.
\end{theorem}

\begin{proof}
	Using Theorem~\ref{thm:5c}, generate a 5-color assignment for $G$. As 5 is less than 12, this assignment suffices.
\end{proof}

\begin{corollary}
	If $G$ has at least 12 nobes, then all 12 colors may be used.
\end{corollary}

\begin{proof}
	First we will prove the case for a smaller number of colors $c<12$, by induction on $c$. Assume a graph $G$ with $c$ or more nobes, and an existing assignment of $c-1$ colors in which all colors are represented, such that $c<12$. The pigeonhole principle states that there must exist two nobes $n_1$ and $n_2$ with the same color. Simply assign the $c$th color to $n_2$ in the new coloring.

	Next for the case $c=12$. Given a graph $G$ with 12 or more nobes, there must exist an assignment of 11 colors in which all colors are represented. The pigeonhole principle states that there must exist two nobes $n_1$ and $n_2$ with the same color. Assigning the 12th color to $n_2$ causes all 12 colors to be used.
\end{proof}

An example application of our algorithm is depicted in Figure~\ref{fig:usa}.

\section{Conclusions and Future Work}

In this paper we reviewed historical developments in the field, with a particular focus on the proof of the \sout{4-Color Theorem} 5-Color Theorem; and we presented our main contribution, a constructive proof of our new \textit{12-Color Theorem} based on the assumption of existing \sout{4-colorings} 5-colorings of an arbitrary planar graph $G$.

The most significant limitation of our work is that a generalized version of our proof only applies for $C$-colorings when $C \le 12$. In future research, we plan on extending our theorem to support ever greater values of $C$. We hope the need for additional funding to study such advanced open problems is self-evident.

\newpage

\thispagestyle{empty}

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{4color.pdf}
	\caption{Make a 4-color of your graph first. \\ {}}
	\end{subfigure}
	\quad
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{5color.pdf}
	\caption{Choose a nobe and color it different.}
	\end{subfigure}

	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{12color.pdf}
	\caption{Repeat that seven more times to get a 12-coloring! \\ \vspace{3.3em}}
	\end{subfigure}
	\quad
	\begin{subfigure}[b]{0.45\textwidth}
	\includegraphics[width=\textwidth]{50color.pdf}
	\caption{If you had 51 different colors and wanted to use them all, it would look like this. However, a proof of the existence of generalized 51-colorings is left to future work.}
	\end{subfigure}
	\caption{Mr. Blum said we needed to have some pictures in our paper, so we pretended the USA was a graph and colored it like our algorithm says.}
	\label{fig:usa}
\end{figure}

\end{document}

### 2015-03-17 koenigs/koenigs-pittsburgh.tex
Write koenigs pittsburgh paper

\documentclass[twocolumn]{article}
\usepackage{fullpage}
%%\usepackage{amsmath}
%%\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{hyperref}
\usepackage{algpseudocode}
%%\usepackage{ulem}
%%\usepackage{setspace}
\usepackage{graphicx}
\usepackage{subcaption}
%%\usepackage{phonetic}
%%\usepackage{synttree}

%\setlength{\columnsep}{14pt}

\usepackage{mathspec}
%\usepackage{fouriernc}
%\usepackage[T1]{fontenc}
%\usepackage[veryoldstyle]{kpfonts}

\newfontfamily\oldestyle{Slavkappen}

\setallmainfonts{OldNewspaperTypes}

\usepackage{titlesec}
\titleformat{\section}
  {\normalfont\oldestyle\Large\bfseries}
  {\thesection}{1em}{}

\renewcommand\exists{\reflectbox{E}}
\renewcommand\forall{\raisebox{\depth}{\rotatebox{180}{A}}}

%\newcommand{\longs}{\hspace{-1.5pt}$\int$\hspace{-2.5pt}}
\theoremstyle{plain}  % Italic body text
  \newtheorem{defn}{Definition}
  \newtheorem{thm}{Theorem}
  \newtheorem{lemma}{Lemma}
\theoremstyle{definition}  % Roman body text
  \newtheorem{proc}{Procedure}
\theoremstyle{remark}  % Italic header, roman body text
  \newtheorem{rem}{Remark}
  \newtheorem{cond}{Condition}

  \title{\em A CONSTRVCTIVE SOLVTION TO THE K\"ONIGS-PITTSBVRGH BRIDGE PROBLEM}
\author{Greg HANNEMAN\\
	Language Enthusiast and Errant Scribbler
	%\footnote{and Salaried Employee now too!}
	\\
        Carnegie Mellon Vniversity
	\and
	Benj. BLVM \\
	Light Cone Sedentarian \\
	Carnegie Mellon Vniversity}
\date{}
\pagestyle{empty}

\begin{document}
%\normalem

\maketitle
\pagestyle{empty}
\thispagestyle{empty}

\begin{abstract}
	We walked across a bunch of bridges one day.
\end{abstract}

% Format this paper to look like an old-style cramped math paper from like the '50s.  All math, no text, very dense.

\section*{\S 1 - Back-Ground}

Figure \ref{fig:koenigs} and Definition \ref{Def-Konigs} state the well-studied K\"onigsberg Bridge Problem \cite{original}. % http://math.dartmouth.edu/~euler/docs/originals/E053.pdf

\begin{defn} \label{Def-Konigs}
	The K\"onigsberg Bridge Problem (KBP). % Refers to ``the figure''

Regiomonti in Borussia esse insulam A {\em der Kneiphof} dictam, fluviumque eam cingentem in duos dividi ramos, quemadmodum ex figura videre licet: ramos uero huius fluvii septem instructos esse pontibus, $a$, $b$, $c$, $d$, $e$, $f$, et $g$.  Circa hos pontes iam ista proponebatur quaestio, num quis cursum ita instituere queat, ut per singulos pontes semel et non plus quam semel transeat.
\end{defn}

According to the best modern scholarship, it is thought that this passage reads as follows.

\begin{quote}
	,,At K\"onigsberg in Prussia there is an island A (Kneiphof), and flowing around it is divided into two branches, as shown in the figure; and the branches of the river, seven bridges, however, of this, \$a\$, \$B\$, \$C\$, \$d\$, \$e\$, \$F\$, and \$G\$. Already presented to him, the question concerning these bridges is, the course to be instructed so to whether a person may be able to, and not more than once in order to pass one by one the bridges honey themselves.''
\end{quote}

\begin{figure}[h]
\centering
	\includegraphics[width=0.48\textwidth]{koenigsburg.pdf}
	\caption{Kaliningrad, Russia (colloq.); better known in the formal literature as K\"onigsberg, Prussia}
	\label{fig:koenigs}
\end{figure}

The criteria expressed in Definition \ref{Def-Konigs} have been generalized to arbitrary geographical topologies. % TODO: cite?
Of extreme local interest is the instance derived from the city of Pittsburgh, Pennsylvania, V.S.A., as Definition \ref{Def-KonigsPgh}.

\begin{defn} \label{Def-KonigsPgh}
	The K\"onigs-Pittsburgh Bridge Problem (KPBP).

Cross every pedestrian-navigable trans-Three-River bridge with at least one endpoint in the city of Pittsburgh, exactly once, to be accomplished during one continuous circuit.
\end{defn}

This article presents a constructive proof for the existence of a solution to KPBP. Furthermore, while most prior work \cite{wikipedia} has largely focused on theoretical analysis of KBP variants based on graph theory, here we additionally offer a more practical approach based on ambulation theory.

\section*{\S 2 - Formalization}

The formal statement of KBP inquires after the existence of an Eulerian path on a graph representing K\"onigsberg, with one node to represent each of the landmasses connected by various edges (bridges). In KPBP we simply change the graph to represent the topology of our home city (Figure \ref{fig:map}).\footnote{The astute observer will note that no edge among our formalism represents the $32^{\text{nd}}$ street bridge. During the progress of our research, this bridge did not exist due to construction, and hence is omitted. If added to the graph (call it KPBP'), it is known that a solution exists, but a constructive proof is left to future work.}

\begin{figure}[t]
\centering
	\includegraphics[width=0.35\textwidth]{1859pittsburgh.jpg}
	\caption{Informal Diagram of the Topology of Pittsburgh, Pennsylvania as Defined by its Bridges and Rivers.}
	\label{fig:map}
\end{figure}

\begin{figure}[t]
\centering
	\includegraphics[width=0.48\textwidth]{topology.pdf}
	\caption{Formalization of Figure \ref{fig:map}.}
	\label{fig:topo}
\end{figure}

Vnlike KBP, in KPBP we require an Eulerian circuit, rather than an Eulerian path.
Recent developments in ambulation theory have shifted the common wisdom towards a preference for the use of circuits over paths in ambulatory proofs, as the circuit presents a more convenient practical application. (Put in more intuitive terms, the Eulerian circuit allows the bridge-walkers to sleep in their own beds the following night, rather than being stranded across a river.) Fortunately, the topology we have selected for KPBP allows for such a solution..
% http://en.wikipedia.org/wiki/Eulerian_path

\section*{\S 3 - Proof}

\begin{thm}
The K\"onigs-Pittsburgh Bridge Problem is solvable.
\end{thm}

We present here a constructive solution to the problem, first given by Hanneman, Vangpat, and Blum (2012, Oct. 27, 5:15 a.m.--10:05 p.m.).  The full solution has length 35 miles and consumes approximately 17 hours of wall-clock time. We depict $R$ in two distinct styles, in a somewhat abbreviated form, in Table \ref{tab:solution} and Figure \ref{fig:solution}.

\begin{table}[t]
	\centering
	\small
	\begin{tabular}{ll}
	\begin{tabular}{|l|l|l|}
		\hline
		Node & Edge & Node \\
		\hline
		EE & HP & NS \\
		NS & 62 & EE \\
		EE & 40 & NS \\
		NS & 16 & EE \\
		EE & RaC & NS \\
		NS & AW & EE \\
		EE & RoC & NS \\
		NS & MKR & SS \\
		SS & WE & NS \\
		\hline
	\end{tabular}
	&
	\begin{tabular}{|l|l|l|}
		\hline
		Node & Edge & Node \\
		\hline
		NS & FtD & EE \\
		EE & FtP & SS \\
		SS & SF & EE \\
		EE & 10 & SS \\
		SS & Lib & EE \\
		EE & BH & SS \\
		SS & HMS & EE \\
		EE & GF & SS \\
		SS & HSG & EE \\
		\hline
	\end{tabular}
	\end{tabular}
	\caption{$|R|$ = 35 mi.}
	\label{tab:solution}
\end{table}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.48\textwidth]{solution.png}
	\caption{A cartographic rendering of the solution route $R$.}
	\label{fig:solution}
\end{figure}

Intermediate details may easily be filled in by following Procedure \ref{Proc-Walk}.

\begin{proc} \label{Proc-Walk}
Let $R$ be the given solution route for KPBP, $p$ be a position along it measured in miles from the start of $R$, and $\hat{r}_x$ be a unit vector in the direction of $R$ at position $x$.
\begin{algorithmic}
\While{$p < |R|$}
    \State $\vec{d} \gets \mathrm{StepSize} \cdot \hat{r}_p$
    \State $\mathrm{Walk}(p, \vec{d})$
    \State $p \gets p + ||\vec{d}||$
\EndWhile
\end{algorithmic}
\end{proc}

\begin{rem}
Some experimental determination will of course be necessary to arrive at the correct value of the StepSize variable.  We suggest a default value of 75 cm, to be adjusted upwards or downwards as personal height and local terrain require.
\end{rem}

\begin{lemma} \label{Lemma-Walk}
Walkability.

A point $p'$ along $R$ can be reached from point $p$ along $R$, following Procedure \ref{Proc-Walk}, assuming $p < p'$.

\end{lemma}
\begin{proof}
	Try it.
\end{proof}

Lemma \ref{Lemma-Walk} thus establishes a ``modus ponens'' result for traversal of the route $R$ from its beginning ($p = 0$) to its end ($p = |R|$).  This is, however, a purely theoretical result.  We now show the practicability of the solution in an operational scenario via a number of satisfiable conditions.

\begin{figure*}[b]
	\begin{tabular}{cc}
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{initialization.png}
	\caption{Condition \ref{cond:init}, initialization.}
	\end{subfigure}
	&
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{mckees.png}
	\caption{An example member of the set $B$.}
	\end{subfigure}
	\\

	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{mini-koenigs.png}
	\caption{$\exists b \in R~\text{s.t.}~b \not\in B$.}
	\end{subfigure}
	&
	\begin{subfigure}[b]{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{finish.png}
	\caption{Condition \ref{cond:term}, successful termination.}
	\end{subfigure}

	\end{tabular}
	\caption{Satisfiability of Conditions. \cite{alanv}}
	\label{fig:photos}
\end{figure*}

\begin{cond}
	\label{cond:init}
Initialization.

Let $A$ be a set of attempting participants.  Then $a$ must be present at $p = 0 \forall~a \in A$.
\end{cond}

\begin{cond}
	\label{cond:prog}
Progression.

Let $B$ be a set of qualifying bridges such that the following are true $\forall~b \in B$:  (1) $b$ satisfies the bridge criteria from Definition \ref{Def-KonigsPgh}; (2) $b$ has defined $0 \leq p_b \leq 35$ on $R$.  Let $b_i$ be the $i$th ordered element of $B$, and let $|B| = n$.  Then $C = \{c_1, ..., c_n\}$, and $c_i = (S_i, b_i)$ iff some $S_i \subseteq A$ crosses $b_i$.
\end{cond}

\begin{cond}
	\label{cond:term}
Termination.

Let the set of ending participants $E \subseteq A = \{ a \in A\,|\, a \in S_i\ \forall~i \}$.  Then the solution is successful iff $E \neq \emptyset$ and $e$ is present at $p = 35\ \forall~e \in E$.
\end{cond}

\begin{lemma}
Satisfiability of Conditions.

Conditions 1--3 are satisfiable.
\label{lem:satisf}

\end{lemma}
\begin{proof}
	Set $A = \{$Alan, Ben, Dan, Greg, Keith$\}$. Note that $\Pr(x~\mathrm{awake~at~5~a.m.}) = \epsilon\ \forall~x \in A$. As long as $\epsilon > 0$, $\epsilon^5 > 0$ \cite{what-is-epsilon}.

Set $B = \{$Highland Park, 61st Street, 40th Street, 16th Street, 9th Street, 7th Street, 6th Street, McKees Rocks, West End, Fort Duquesne, Fort Pitt, Smithfield Street, Liberty, 10th Street, Birmingham, Hot Metal, Glenwood, Homestead Grays$\}$.
The bridge criteria of Definition \ref{Def-KonigsPgh} are satisfied $\forall~b \in B$. Therefore the set $C$ of crossings may be freely constructed following $R$ from Figure \ref{fig:solution}.

Set $E = \{$Alan, Ben, Greg$\}$.
The diagrams in Figure \ref{fig:photos} offer a visualization of these values.

Since $|E| = 3 > 0$, this proves Lemma \ref{lem:satisf} and thus the successful instantiation of Conditions 1--3.  Together with Lemma \ref{Lemma-Walk}, this proves that $R$ is a navigable Eulerian circuit of the city of Pittsburgh.  Therefore a solution to the K\"onigs-Pittsburgh Bridge Problem exists.
\end{proof}

\section*{\S 4 - Conclusion}

Vetere Konigsberg consequat pons (KBP) designat insulae urbem, flumina, pontes, qui in tali figuratione transiens per singulos semel iter unum non potest.
In hoc articulo quaestionis exposuimus accommodatio hodierna die ad oppidum Pittsburgh (KPBP).
Significat adaptationem ad profectum nostrum insignis civitas-in-es super duobus.

Vno modo, sicut KPBP est solubilis, offert mathematici nostram collationem incitamentum ad ire foris, et accipiam ab eo aliquid exercitium, quod hodie magis quam somnium, quod fieri non potest.
Deinde insuper offerebat speculativa ratio nos instruit etiam problema aedificant, qui probi ad ultimum pedes decem milia multa mala!

\bibliographystyle{alpha}
\bibliography{sigbovik}

\end{document}

### 2015-03-17 koenigs/koenigs-pittsburgh.tex
fix stuff in koenigs

@@
 \usepackage{amsthm}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{algpseudocode}
@@
 
+%\usepackage{rotating}
 \usepackage{mathspec}
@@
 
-\usepackage{titlesec}
-\titleformat{\section}
-  {\normalfont\oldestyle\Large\bfseries}
-  {\thesection}{1em}{}
+%\usepackage{titlesec}
+%\titleformat{\section}
+%  {\normalfont\oldestyle\Large\bfseries}
+%  {\thesection}{1em}{}
 
@@
 
-  \title{\em A CONSTRVCTIVE SOLVTION TO THE K\"ONIGS-PITTSBVRGH BRIDGE PROBLEM}
+  \title{\em \huge A CONSTRVCTIVE SOLVTION TO THE K\"ONIGS-PITTSBVRGH BRIDGE PROBLEM}
 \author{Greg HANNEMAN\\
@@
 
-\begin{abstract}
+\section*{ABSTRACT.}
 	We walked across a bunch of bridges one day.
-\end{abstract}
 
@@
 
-\section*{\S 1 - Back-Ground}
+\section*{\S 1..... BACKGROVND.}
 
@@
 	\includegraphics[width=0.48\textwidth]{koenigsburg.pdf}
-	\caption{Kaliningrad, Russia (colloq.); better known in the formal literature as K\"onigsberg, Prussia}
+	\caption{Kaliningrad, Russia (colloq.); better Known in the Formal Literature as K\"onigsberg, Prussia}
 	\label{fig:koenigs}
@@
 
-\section*{\S 2 - Formalization}
+\section*{\S 2..... FORMALIZATION.}
 
@@
 During the progress of our research, this bridge did not exist due to construction, and hence is omitted.
-If added to the graph (call it KPBP'), it is known that a solution exists, but a constructive proof is left to future work.}
+If added to the graph (call it KPBP'), it is known that a non-cyclic solution exists, but a constructive proof is left to future work.}
 
@@
 	\includegraphics[width=0.35\textwidth]{1859pittsburgh.jpg}
-	\caption{Informal Diagram of the Topology of Pittsburgh, Pennsylvania as Defined by its Bridges and Rivers.}
+	\caption{informal Diagram of the Topology of Pittsburgh, Pennsylvania as defined by its Bridges and Rivers.}
 	\label{fig:map}
@@
 
-\section*{\S 3 - Proof}
+\section*{\S 3..... PROOF.}
 
@@
 
-\begin{table}[t]
+\begin{table}[b]
 	\centering
@@
 	\end{tabular}
-	\caption{$|R|$ = 35 mi.}
+	\caption{|R| = 35 mi.}
 	\label{tab:solution}
 \end{table}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.48\textwidth]{solution.png}
-	\caption{A cartographic rendering of the solution route $R$.}
-	\label{fig:solution}
-\end{figure}
 
@@
 Let $R$ be the given solution route for KPBP, $p$ be a position along it measured in miles from the start of $R$, and $\hat{r}_x$ be a unit vector in the direction of $R$ at position $x$.
+\end{proc}
+
+%%%%%%%%%%%%%
+%\newpage
+%\begin{turn}{1}
+%\begin{minipage}{\textwidth}
+%\begin{multicols}{2}
+
 \begin{algorithmic}
-\While{$p < |R|$}
+\While{$p < $|R|}
     \State $\vec{d} \gets \mathrm{StepSize} \cdot \hat{r}_p$
-    \State $\mathrm{Walk}(p, \vec{d})$
-    \State $p \gets p + ||\vec{d}||$
+    \State Walk($p, \vec{d}$)
+    \State p $\gets$ p + ||$\vec{d}$||
 \EndWhile
 \end{algorithmic}
-\end{proc}
 
@@
 
-Lemma \ref{Lemma-Walk} thus establishes a ``modus ponens'' result for traversal of the route $R$ from its beginning ($p = 0$) to its end ($p = |R|$).
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.48\textwidth]{solution.png}
+	\caption{a cartographic Rendering of the Solution Route $R$.}
+	\label{fig:solution}
+\end{figure}
+
+Lemma \ref{Lemma-Walk} thus establishes a ``modus ponens'' result for traversal of the route $R$ from its beginning (p = 0) to its end (p = |R|).
 This is, however, a purely theoretical result.
 We now show the practicability of the solution in an operational scenario via a number of satisfiable conditions.
-
-\begin{figure*}[b]
-	\begin{tabular}{cc}
-	\begin{subfigure}[b]{0.45\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{initialization.png}
-	\caption{Condition \ref{cond:init}, initialization.}
-	\end{subfigure}
-	&
-	\begin{subfigure}[b]{0.45\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{mckees.png}
-	\caption{An example member of the set $B$.}
-	\end{subfigure}
-	\\
-
-	\begin{subfigure}[b]{0.45\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{mini-koenigs.png}
-	\caption{$\exists b \in R~\text{s.t.}~b \not\in B$.}
-	\end{subfigure}
-	&
-	\begin{subfigure}[b]{0.45\textwidth}
-	\centering
-	\includegraphics[width=\textwidth]{finish.png}
-	\caption{Condition \ref{cond:term}, successful termination.}
-	\end{subfigure}
-
-	\end{tabular}
-	\caption{Satisfiability of Conditions. \cite{alanv}}
-	\label{fig:photos}
-\end{figure*}
 
@@
 
-Let $A$ be a set of attempting participants.  Then $a$ must be present at $p = 0 \forall~a \in A$.
+Let $A$ be a set of attempting participants.  Then $a$ must be present at p = 0 \forall{}a $\in$ A.
 \end{cond}
@@
 
-Let $B$ be a set of qualifying bridges such that the following are true $\forall~b \in B$:  (1) $b$ satisfies the bridge criteria from Definition \ref{Def-KonigsPgh}; (2) $b$ has defined $0 \leq p_b \leq 35$ on $R$.
-Let $b_i$ be the $i$th ordered element of $B$, and let $|B| = n$.
-Then $C = \{c_1, ..., c_n\}$, and $c_i = (S_i, b_i)$ iff some $S_i \subseteq A$ crosses $b_i$.
+Let $B$ be a set of qualifying bridges such that the following are true \forall{}b $\in$ B:  (1) $b$ satisfies the bridge criteria from Definition \ref{Def-KonigsPgh}; (2) $b$ has defined 0 $\leq$ $p_b$ $\leq$ 35 on $R$.
+Let $b_i$ be the $i$th ordered element of $B$, and let |B| = n.
+Then C = \{$c_1, ..., c_n$\}, and $c_i$ = ($S_i, b_i$) iff some $S_i \subseteq A$ crosses $b_i$.
 \end{cond}
@@
 
-Let the set of ending participants $E \subseteq A = \{ a \in A\,|\, a \in S_i\ \forall~i \}$.  Then the solution is successful iff $E \neq \emptyset$ and $e$ is present at $p = 35\ \forall~e \in E$.
+Let the set of ending participants $E \subseteq A$ = \{$a \in A\,|\, a \in S_i\ \forall{}i$\}.  Then the solution is successful iff $E \neq \emptyset$ and $e$ is present at p = 35 \forall{}e $\in$ E.
 \end{cond}
@@
 \begin{proof}
-	Set $A = \{$Alan, Ben, Dan, Greg, Keith$\}$. Note that $\Pr(x~\mathrm{awake~at~5~a.m.}) = \epsilon\ \forall~x \in A$. As long as $\epsilon > 0$, $\epsilon^5 > 0$ \cite{what-is-epsilon}.
+	Set A = \{Alan, Ben, Dan, Greg, Keith\}. Note that Pr(x  awake at 5 a.m.) = $\epsilon$ \forall{}x $\in$ A. As long as $\epsilon$ > 0, $\epsilon^5$ > 0. \cite{what-is-epsilon}
 
-Set $B = \{$Highland Park, 61st Street, 40th Street, 16th Street, 9th Street, 7th Street, 6th Street, McKees Rocks, West End, Fort Duquesne, Fort Pitt, Smithfield Street, Liberty, 10th Street, Birmingham, Hot Metal, Glenwood, Homestead Grays$\}$.
-The bridge criteria of Definition \ref{Def-KonigsPgh} are satisfied $\forall~b \in B$. Therefore the set $C$ of crossings may be freely constructed following $R$ from Figure \ref{fig:solution}.
+Set B = \{Highland Park, 61st Street, 40th Street, 16th Street, 9th Street, 7th Street, 6th Street, McKees Rocks, West End, Fort Duquesne, Fort Pitt, Smithfield Street, Liberty, 10th Street, Birmingham, Hot Metal, Glenwood, Homestead Grays\}.
+The bridge criteria of Definition \ref{Def-KonigsPgh} are satisfied $\forall{}b \in B$. Therefore the set $C$ of crossings may be freely constructed following $R$ from Figure \ref{fig:solution}.
 
-Set $E = \{$Alan, Ben, Greg$\}$.
-The diagrams in Figure \ref{fig:photos} offer a visualization of these values.
+Finally, set E = \{Alan, Ben, Greg\}.
+The diagrams in Figure \ref{fig:photos} offer a visualization of these values for A, B, and E.
 
-Since $|E| = 3 > 0$, this proves Lemma \ref{lem:satisf} and thus the successful instantiation of Conditions 1--3.
+Since |E| = 3 > 0, this proves Lemma \ref{lem:satisf} and thus the successful instantiation of Conditions 1--3.
 Together with Lemma \ref{Lemma-Walk}, this proves that $R$ is a navigable Eulerian circuit of the city of Pittsburgh.
@@
 
-\section*{\S 4 - Conclusion}
+\section*{\S 4..... CONCLVSION.}
 
@@
 
+\renewcommand\refname{REFERENCES.}
 \bibliographystyle{alpha}
@@
 
+%%%%%%%%%%%%%
+%\end{multicols}
+%\end{minipage}
+%\end{turn}
+%\newpage
+
+\begin{figure*}[b]
+	\begin{tabular}{cc}
+	\begin{subfigure}[b]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{initialization.png}
+	\caption{Condition \ref{cond:init}, Initialization.}
+	\end{subfigure}
+	&
+	\begin{subfigure}[b]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{mckees.png}
+	\caption{an example Member of the set $B$.}
+	\end{subfigure}
+	\\
+
+	\begin{subfigure}[b]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{mini-koenigs.png}
+	\caption{\exists b $\in$ R s.t. b $\not\in$ B.}
+	\end{subfigure}
+	&
+	\begin{subfigure}[b]{0.45\textwidth}
+	\centering
+	\includegraphics[width=\textwidth]{finish.png}
+	\caption{Condition \ref{cond:term}, successful Termination.}
+	\end{subfigure}
+
+	\end{tabular}
+	\caption{Satisfiability of Conditions. \cite{alanv}}
+	\label{fig:photos}
+\end{figure*}
+
 \end{document}

### 2015-03-17 12color/paper.tex
update 12color

@@
 Victoria said I had to write this section, even though she's the only one of us who actually knows how the 4-color theorem proof goes.
-But luckily I found a proof for the 5-color theorem that made sense, and since having more colors is obviously better, that should be even better.
+But luckily I found a proof for the 5-color theorem that made sense, and since having more colors is obviously better, that should work just as good.
 It goes like this:
@@
 \begin{lemma}
-All planar graphs have at least one vertex with at most 5 neighbors.
+All planar graphs have at least one nobe with at most 5 neighbors.
 \end{lemma}
@@
 	Let $N(n)$ mean how many neighbors a given nobe $n$ has.
-Then because each edge connects two nobes, the sum of all nobes' neighbours $\sum_n N(n) = 2e$.
-Let's assume all nobes have at least six neighbours.
+Then because each edge connects two nobes, the sum of all nobes' neighbors $\sum_n N(n) = 2e$.
+Let's assume all nobes have at least six neighbors.
 Then $6n \le 2e$.
@@
 But since $n-e+f = 2$ we end up with $2\le0$.
-Then we must have been wrong to assume all nobes have at least six neighbours.
+Then we must have been wrong to assume all nobes have at least six neighbors.
 That's a ``contradiction''!
@@
 	Let $G$ mean the graph we're trying to prove this for.
-Using the lemon I just proved, pick the nobe $n$ that has at most five neighbors and call those $n_1$, $n_2$, $n_3$, $n_4$, $n_5$ in whatever order you want.
+Using the lemon I just proved, pick the nobe $n$ that has at most five neighbors and call those $n_1$, $n_2$, $n_3$, $n_4$, $n_5$ in whatever order you want, but remember the order!
 Let $G'$ mean the smaller graph you get by taking out $n$, which of course is still planar.
-Because $G'$ is smaller, we know with ``induction'' that $G'$ can be 5-colored.
+Because $G'$ is smaller, we already know with ``induction'' that $G'$ can be 5-colored.
 (That step seemed kinda weird to me but Victoria said it was legit so please don't take points off.)
 
-	Now we're gonna assume $G$ cannot be 5-colored.
+	Now we're gonna assume $G$ can't be 5-colored.
 For this to be true, then each $n_i$ from $1$ to $5$ must all have different colors, or else you could just pick the missing one for $n$.
 Let's think about the sub-graph that has only the 2 colors matching $n_1$ and $n_3$, call it $G_{13}$.
-If the graph is disjoint, meaning $n_1$ and $n_3$ are in two separate parts, then you would be able to re-color the graph to make $n_3$ have color $1$, so you can paint $n$ color $3$.
+If the graph is disjoint, meaning $n_1$ and $n_3$ are in two separate parts, then you would be able to re-color the graph to make $n_3$ have color $1$, so you can paint $n$ with color $3$.
 Therefore $G_{13}$ must be connected.
@@
 
-	Therefore, even if the 5-coloring of $G'$ gives all different colors to $n_1$, $n_2$, $n_3$, $n_4$, and $n_5$, you should be able to re-color one of the $n_i$ so you can have a spare color to give to $n$.
-Therefore $G$ can be 5-colored.
+	So now, even if the 5-coloring of $G'$ gives all different colors to $n_1$, $n_2$, $n_3$, $n_4$, $n_5$, you should be able to re-color one of the $n_i$s so you can have a spare color to give to $n$.
+That means you can 5-color $G$.
 \end{proof}
+
+Nice! Hey, that wasn't so bad.
 
@@
 	\includegraphics[width=\textwidth]{50color.pdf}
-	\caption{If you had 51 different colors and wanted to use them all, it would look like this. However, a proof of the existence of generalized 51-colorings is left to future work.}
+	\caption{If you had 51 different colors and wanted to use them all, it would look like this. However, we are leaving a proof of the existence of generalized 51-colorings to future work.}
 	\end{subfigure}

### 2016-03-07 ode/ode.tex
Ode to Reviewer Two

\documentclass{article}
\special{papersize=8.5in,11in}

\usepackage[charter]{mathdesign}
\usepackage{fullpage}

\title{Ode to Reviewer Two}
\author{{\em anonymized for double-blind submission}}
\date{April 1, 2016}

\begin{document}

\maketitle
\thispagestyle{empty}

\noindent
My paper submitted, the deadline complete;\\
The product of months of lonely toil,\\
With quality prose and experiments replete\\
Amid insecurities, fears, and other turmoil.\\
Though once I feared a harsh rejection,\\
My advisor assured me that my proofs were sound\\
And my treatment of the work related, fair,\\
So I've come to believe in the paper's perfection.\\
Through this research, new self-confidence found,\\
Though all-nighters have left me exhausted and drowned.\\
Now waiting for judgement from reviewers elsewhere.\\
\\
But alas! Though first and third reviewers were happy,\\
Reviewer the second couldn't seem to accept.\\
He gave reasons he thought my paper was crappy,\\
But I found his attempted critique most inept.\\
His comments betrayed a misunderstanding\\
And nonsense ``suggestions'' were falsely polite,\\
Completely missing the point of my work,\\
I couldn't believe what he was demanding:\\
To re-run experiments, perhaps out of spite;\\
An unrelated paper he asked me to cite!\\
(Probably his own.) What a jerk.\\
\\
With a glimmer of hope, I wrote a rebuttal\\
Appealing to reviewers One and Three impressed,\\
And suggested to Two, hey, you missed something subtle?\\
You'll reconsider, I desp'rately expressed.\\
The final suggestions were naught but derision:\\
``Clearly elaborate!'' was all Two replied,\\
Hiding the true message that he was outwit;\\
But it was too late to change their decision.\\
My paper rejected, my joy and pride,\\
My confidence collapsed in a sudden landslide;\\
Now to find somewhere to resubmit.\\

\end{document}

### 2016-03-07 koenigs/koenigs-pittsburgh.tex
some koenigs changes that were lying around from last year

@@
   \title{\em \huge A CONSTRVCTIVE SOLVTION TO THE K\"ONIGS-PITTSBVRGH BRIDGE PROBLEM}
-\author{Greg HANNEMAN\\
+\author{Greg. HANNEMAN\\
 	Language Enthusiast and Errant Scribbler
@@
 \begin{quote}
-	,,At K\"onigsberg in Prussia there is an island A (Kneiphof), and flowing around it is divided into two branches, as shown in the figure; and the branches of the river, seven bridges, however, of this, \$a\$, \$B\$, \$C\$, \$d\$, \$e\$, \$F\$, and \$G\$.
+	,,At K\"onigsberg in Prussia there is an island A known as Kneiphof, and flowing around it is divided into two branches, as shown in the figure; and the branches of the river, seven bridges, however, of this, \$a\$, \$B\$, \$C\$, \$d\$, \$e\$, \$F\$, and \$G\$.
 Already presented to him, the question concerning these bridges is, the course to be instructed so to whether a person may be able to, and not more than once in order to pass one by one the bridges honey themselves.''

### 2016-03-07 ode/ode.tex
fix ode punctuation

@@
 And nonsense ``suggestions'' were falsely polite,\\
-Completely missing the point of my work,\\
+Completely missing the point of my work.\\
 I couldn't believe what he was demanding:\\

### 2016-03-07 ode/ode.tex
fix meter in ode

@@
 An unrelated paper he asked me to cite!\\
-(Probably his own.) What a jerk.\\
+(Probably his own.) What an arrogant jerk.\\
 \\

### 2016-03-08 ode/ode.tex
fix meter of many lines in ode

--- 2016-03-07 ode/ode.tex
@@
 The product of months of lonely toil,\\
-With quality prose and experiments replete\\
-Amid insecurities, fears, and other turmoil.\\
+With quality prose and exper'ments replete\\
+Amid insecurities and other turmoil.\\
 Though once I feared a harsh rejection,\\
-My advisor assured me that my proofs were sound\\
-And my treatment of the work related, fair,\\
-So I've come to believe in the paper's perfection.\\
-Through this research, new self-confidence found,\\
-Though all-nighters have left me exhausted and drowned.\\
-Now waiting for judgement from reviewers elsewhere.\\
+My advisor assured me my proofs were quite sound\\
+And my treatment of the work related, fair.\\
+So I've come to believe in the paper's perfection;\\
+Though all-nighters have left me exhausted and drowned,\\
+Through this research, new self-esteem found!\\
+Now waiting for judgment from reviewers elsewhere.\\
 \\
-But alas! Though first and third reviewers were happy,\\
-Reviewer the second couldn't seem to accept.\\
-He gave reasons he thought my paper was crappy,\\
+Alas! Though readers first and third were happy,\\
+Reviewer the second couldn't bear to accept.\\
+He gave several reasons my paper seemed crappy,\\
 But I found his attempted critique most inept.\\
 His comments betrayed a misunderstanding\\
-And nonsense ``suggestions'' were falsely polite,\\
+And nonsense `suggestions' were falsely polite,\\
 Completely missing the point of my work.\\
 I couldn't believe what he was demanding:\\
-To re-run experiments, perhaps out of spite;\\
-An unrelated paper he asked me to cite!\\
-(Probably his own.) What a jerk.\\
+To rerun my test suite, perhaps out of spite;\\
+An unr'lated paper he asked me to cite!\\
+(Probably his own.) What an arrogant jerk.\\
 \\
 With a glimmer of hope, I wrote a rebuttal\\
-Appealing to reviewers One and Three impressed,\\
-And suggested to Two, hey, you missed something subtle?\\
-You'll reconsider, I desp'rately expressed.\\
+Appealing to readers One and Three impressed,\\
+And suggested to Two, ``Hey, you missed something subtle?\\
+You'll reconsider,'' I desp'rately expressed.\\
 The final suggestions were naught but derision:\\
 ``Clearly elaborate!'' was all Two replied,\\
-Hiding the true message that he was outwit;\\
-But it was too late to change their decision.\\
-My paper rejected, my joy and pride,\\
-My confidence collapsed in a sudden landslide;\\
-Now to find somewhere to resubmit.\\
+Hiding the plain truth that he'd been outwit.\\
+For it was too late to change their decision:\\
+My paper rejected, my joy and my pride,\\
+My conf'dence collapsed in a sudden landslide.\\
+Now to find somewhere to soon resubmit.\\
 

### 2016-03-11 itg/paper.tex
start writing a paper about itg

--- 2013-01-19 RSLC/paper/paper.tex
@@
 \usepackage{amssymb}
-\usepackage{mathpartir}
+%\usepackage{mathpartir}
 \usepackage{url}
@@
 
-\renewcommand*\ttdefault{txtt}
+%\renewcommand*\ttdefault{txtt}
 \usepackage[T1]{fontenc}
-
-\RequirePackage{listings}
-% \lstdefinelanguage{Rust}{
-%   morekeywords={fn,let,enum,if,else,return,mut,const,copy,ref,move,match,struct,type,int,while,float,as,fail},
-%   sensitive=true,
-%   morecomment=[l]{//},
-%   morecomment=[s]{/*}{*/},
-%   morestring=[b]",
-% }
-% \lstset{language=Rust,
-%         frame=none,%single,
-%         basicstyle=\ttfamily\footnotesize,
-%         keywordstyle=\bfseries,
-%         commentstyle=\footnotesize,
-%         tabsize=2,
-%         showstringspaces=false,
-%         flexiblecolumns=false,
-%         mathescape=true,
-%         escapeinside={\#}{\#},
-%         % numbers=left, numberstyle=\tiny, numbersep=5pt %, stepnumber=2, 
-%         % literate={->}{{$\rightarrow$}}1 {<->}{{$\leftarrow$}}1
-% }
 
@@
 % and I rather hope never to learn.
-\makeatletter
-\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
-\makeatother
+%\makeatletter
+%\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
+%\makeatother
 
-\newcommand{\Figref}[1]{Figure \ref{#1}}
-\newcommand{\Secref}[1]{Section \ref{#1}}
-\newcommand{\x}[1]{\lstinline{#1}}
-
-\input{formaldefs.tex}
 \begin{document}
 
-\conferenceinfo{SIGBOVIK '13}{Pittsburgh, PA, USA}
-\copyrightyear{2013}
+\conferenceinfo{SIGBOVIK '16}{Pittsburgh, PA, USA}
+\copyrightyear{2016}
 \copyrightdata{}
@@
 \title{
-% Random Scope for Syntactic Hope
-% Learning to Cope with Random Scope
-Randomly-Scoped Lambda Calculus
+Which ITG Stepcharts are Turniest?
 }
@@
 
-%\authorinfo{
-%  Nicholas D. Matsakis
-%  \and Brian Anderson
-%  \and Ben Blum \\
-%  \and Tim Chevalier 
-%  \and Graydon Hoare
-%  \and Patrick Walton
-%  \and David Herman}
-%           {Mozilla Research}
-%           {\{nmatsakis, banderson\}@mozilla.com,
-%             bblum@cs.cmu.edu, \\
-%             \{tchevalier, graydon, pwalton, dherman\}@mozilla.com}
-
 \maketitle
@@
 \begin{abstract}
-
-Gaze upon the following travesty, which Python hath wrought upon the world:
-\begin{verbatim}
-    >>> x
-    NameError: name 'x' is not defined
-    >>> [x for x in 1,2,3]
-    [1, 2, 3]
-    >>> x
-    3
-    >>>
-\end{verbatim}
-
-The question of {\em scope} pertains to the set of rules governing which values a program's variables refer to during evaluation.
-Prior work has proposed two main approaches: {\em static} (or {\em lexical}) {\em scope}, in which variable bindings are resolved through source code analysis, and {\em dynamic scope}, in which bindings are resolved at run-time using a stack of activation records.
-
-In this work, I present {\em random scope}, a new technique for scoping, as an alternative to the above two.
-Random scope affords programmers the flexibility to refer to multiple different binding sites simultaneously; for example, applying the term $(\lambda x.
-(\lambda x.
-x))$ to two arguments could evaluate to either the first or the second one.
-I figure out what kind of evaluation semantics would make this work, and then pretty much stumble onwards from there trying to make sense of the whole damn thing.
-
-I define the Randomly-Scoped Lambda Calculus, and show how it can be used to compute random natural numbers, while also providing more popular deterministic functionalities such as factorial and fibonacci.
+	ITG is a popular dance game in which players step on arrows while listening to music.
+The arrow patterns, indicated by a {\em stepchart}, may range among any level of complexity and difficulty.
+Among the many factors contributing to a stepchart's difficulty is how much the player must turn from side to side.
+	Other more obvious factors, such as raw speed, have been well studied in prior work. % TODO cite.
+	This paper presents an analytic study of this {\em turniness} factor.
+	We study the turniness of many existing stepcharts, and present a novel (but unsurprising) approach to automatically generating maximally (or minimally) turny charts.
 
@@
 
-% \category{D.3.3}{Language Constructs and Features}{Data types and structures}
+\category{D.D.R.}{Exercise and Fitness}{Arcade Dance Games}
 
 \keywords
-static scope, dynamic scope, lower is better
 
@@
 
-Some modern programming languages~\cite{wikiplia,python,lambda,sml,GoLang,RustLang,mla,haskell98,patenttroll} include a mechanism to refer to previously-computed values using more consise names, typically called {\em variables}.
-The question then arises: ``Which values should my variables refer to?''
-In a half-hearted attempt to answer, these languages implement {\em scope} (which I explained in the abstract; go read it), and hope it all works out okay.
-
-Whether resolved statically or dynamically, modern scoping approaches lack the flexibility to refer to multiple binding sites simultaneously.
-I present the Randomly-Scoped Lambda Calculus (RSLC), in which references to shadowed variables can evaluate to any of their binding sites with equal probability.
-
-Consider the following lambda calculus term:
-\[ (\lambda x. (\lambda x. x))~A~B \]
-In ordinary lambda calculus, this always evaluates to $B$ (quite unforgivingly so, in my opinion).
-However, I give the programmer the benefit of the doubt, and assume they named the first argument $x$ as well because they wanted it to have equal opportunity.
-Hence, in RSLC this term can evaluate to either $A$ or $B$.
-(On the other side of the coin~\cite{twoface}, RSLC encourages good programming practice: a programmer who {\em doesn't} want $x$ to evaluate to $A$ should name the first argument something different, to make their intentions more clear.)
-
-In this paper I make the following contributions:
-\begin{enumerate}
-	\item The {\bf Randomly-Scoped Lambda Calculus (RSLC)}, a logical programming language that uses a novel syntactic technique called {\bf random scoping},
-	%\item An analysis of two implementation strategies for random scoping, which I call {\bf static random scope (SRS)} and {\bf dynamic random scope (DRS)},
-	\item A new schema for general recursion in RSLC (because the Y combinator doesn't work anymore),
-	\item Programming examples in RSLC that employ random scope to compute random natural numbers,
-	\item Programming examples in RSLC that avoid random scope to implement deterministic arithmetic.
-\end{enumerate}
-
-\section{Language Definition}
-
-\begin{figure}[t]
-	\Grammar
-	\caption{RSLC formal grammar.}
-	\label{fig:grammar}
-\end{figure}
-
-RSLC is an extention to the lazy un(i)typed lambda calculus with slightly modified substitution rules.
-In the grammar (Figure~\ref{fig:grammar}), a variable $x$ carries around a list of expressions $\rep{e}$ which we have ``tried to substitute for it before''.
-When programming in RSLC, we simply write $x$ (or $y$ or \Neptune~or however you happen to swing), because no substitution has happened yet, but after $n$ ``$\lambda x$''s have been $\beta$-reduced, the $\rep{e}$ list will have $n$ expressions in it.
-It might help to think of it like quantum superposition, I guess?
-
-Anyway, Figure~\ref{fig:eval} shows the evaluation rules, which are basically what you'd expect, and Figure~\ref{fig:subst} shows the substitution rules, which are new.
-In the normal $\lambda$-calculus, substitution would stop when a new binding's name is the same as the one being substituted for.
-
-However, in RSLC, as shown in {\sc Subst-Capture}, we switch to a second substitution mode, {\sc Capture}, in which substituted expressions are appended to a variable's $\rep{e}$ list (in the {\sc Capture-Var-X} rule).
-Similarly, when substituting for $x$ before switching to {\sc Capture} mode (in the {\sc Subst-Var-X} rule), we nondeterministically select an expression from $x$'s $\rep{e}$ list to replace $x$ with (with uniform randomness).
-
-\begin{figure}[t]
-	\fbox{\begin{tabular}{ll}$e \eval e'$ \\ $e~\Val$\end{tabular}}
-	\TypeRules{
-		\EvalLam \qquad \EvalApp \qquad \EvalVar \\
-		\EvalZero \qquad \EvalSucA \qquad \EvalSucB \\
-		\EvalCaseA \\ \EvalCaseB \\ \EvalLet
-	}
-	\caption{RSLC small-step evaluation rules. They're exactly the same as for any other call-by-name $\lambda$-calculus, so don't bother reading them. Why am I even including them?}
-	\label{fig:eval}
-\end{figure}
-
-\begin{figure}[t]
-	\fbox{$\subst{e_0}{x}{e}$}
-	\TypeRules{
-		\SubstCapture \qquad \SubstLam \\
-		\SubstVarX \qquad \SubstVarY \\
-	}
-	\fbox{$\capture{e_0}{x}{e}$}
-	\TypeRules{
-		\CaptureVarX \qquad \CaptureVarY \\
-	}
-	\caption{Selected RSLC substitution rules. The ``$e_1 \in ...$'' premise in {\sc Subst-Var-X} represents nondeterministic choice.}
-	\label{fig:subst}
-\end{figure}
-
-\subsection{Evaluation Example}
-
-To demonstrate, here's the evaluation of $(\lambda x. (\lambda x. x))~A~B$ from the introduction:
-
-\begin{center}
-\begin{tabular}{cll}
-	& $(\lambda x. (\lambda x. x[]))~A~B$ & \\
-	$\mapsto$ & $(\subst{A}{x} (\lambda x. x[]))~B$ & {\sc \small Eval-App} \\
-	\substeq & $(\lambda x. \capture{A}{x} x[])~B$ & {\sc \small Subst-Capture} \\
-	\captureeq & $(\lambda x. x[A])~B$ & {\sc \small Capture-Var-X} \\
-	$\mapsto$ & $\subst{B}{x} x[A]$ & {\sc \small Eval-App} \\
-	\substeq & {\em (any element of the list $[B,A]$)} & {\sc \small Subst-Var-X}
-\end{tabular}
-\end{center}
-Now the fairness to $A$ we wanted in the intro is restored.
-
-\section{Random Programming with RSLC}
-
-\newcommand{\Y}{\textsf{Y}}
-\newcommand{\Ya}{\ensuremath{\mathsf{Y_1}}}
-\newcommand{\Yab}{\ensuremath{\mathsf{Y_2}}}
-\newcommand{\Yabc}{\ensuremath{\mathsf{Y_3}}}
-
-Already we have a system with which we can implement simple random number generators, such as a simple 6-sided die:\footnote{Really I just wanted an excuse to typeset \LaTeX~dice.}
-\[
-(\lambda x.
-(\lambda x.
-(\lambda x.
-(\lambda x.
-(\lambda x.
-(\lambda x. x))))))
-~\epsdice{1}
-~\epsdice{2}
-~\epsdice{3}
-~\epsdice{4}
-~\epsdice{5}
-~\epsdice{6}
-\]
-
-However, this isn't really ``programming'', and Sully thought for sure I wouldn't be able to do anything useful with this language, which of course I wasn't going to stand by.
-The first thing I wanted to do was to recursively generate a random natural number, something like this:
-\begin{equation}
-\label{eq:randomnat}
-\Y~(\lambda f. \lambda n. (\lambda x. \lambda x. x)~n~(f~\Suc{n}))~\Zero
-\end{equation}
-This term would randomly choose whether to output $n$, its argument, or to call itself with $\Suc{n}$.
-However, the conventional Y-combinator doesn't work as intended in RSLC, because once a function is substituted into itself for its first argument, subsequent arguments will be captured in the substituted version.
-Even ignoring the internal workings of the combinator, and supposing $\Y$ satisfies $\Y~g \mapsto^* g~(\Y~g)$, here's what happens:
-
-\begin{center}
-\begin{tabular}{ll}
-\hspace{-1em}
-& $\Y~(\lambda f. \lambda n. (\lambda x. \lambda x. x)~n~(f~\Suc{n}))~\Zero$ \\
-\hspace{-1em}
-$\mapsto^*$ & $(\lambda n. (\lambda x. \lambda x. x)~n~(\Y~(\lambda f.\lambda n. \dots n \dots)~\Suc{n}))~\Zero$ \\
-\hspace{-1em}
-$\mapsto$ & $\subst{\Zero}{n}((\lambda x. \lambda x. x)~n~(\Y~(\lambda f.\lambda n. \dots n \dots)~\Suc{n}))$ \\
-\hspace{-1em}
-= & $((\lambda x. \lambda x. x)~\subst{\Zero}{n}n~(\Y~(\lambda f.\lambda n. \dots \capture{\Zero}{n}n \dots)~\Suc{\subst{\Zero}{n}n}))$ \\
-\end{tabular}
-\end{center}
-
-But wait -- I only wanted random-capture on $x$, not on $n$!
-As a result, though we intended $0$ to be output with half probability, and $1$ with quarter probability, and so on, actually $0$ gets output substantially more often (5/8, I think).
-
-\subsection{The new Y combinator(s)}
-In order to properly recurse in RSLC, we need a new convention in which recursive functions get themselves substituted into them {\em last}, after all other arguments have been applied.
-The desired identity is $\Ya~g~a \mapsto^* g~a~(\Ya g)$ for one-argument functions (and $\Yab~g~a~b \mapsto^* g~a~b~(\Yab~g)$ for two-argument functions, and so on).
-
-Unfortunately, I wasn't even able to find an RSLC term that satisfied that identity.
-I had to also alter the way recursive functions call themselves -- by referring to the combinator itself, passing it the recursive arguments, and {\em then} the substituted version of themselves.
-Since that makes no sense in words, just have a look:
-\begin{eqnarray*}
-\Ya &=& \lambda ag.~gag \\
-\Yab &=& \lambda abg.~gabg \\
-\Yabc &=& \lambda abcg.~gabcg \\
-&\dots&
-\end{eqnarray*}
-Here's a sort of handwavy syntactic transformation example for $\mathsf{fix}$ constructs that supports one, two, etc arguments:
-\begin{eqnarray*}
-\mathsf{fix}_1(f,a.e) &=& \lambda a. \Ya~a~(\lambda af. [^{(\Ya~x~f)}/_{f(x)}]e) \\
-\mathsf{fix}_2(f,a,b.e) &=& \lambda ab. \Ya~a~b~(\lambda abf. [^{(\Ya~x~y~f)}/_{f(x)(y)}]e) \\
-&\dots&
-\end{eqnarray*}
-
-Note that this transformation doesn't just substitute for the variable $f$ in $e$ but also reorders the arguments it's applied to. So now the term from Equation~\ref{eq:randomnat} becomes:
-\begin{equation}
-\label{eq:randomnat-fixed}
-\Ya~\Zero~(\lambda n. \lambda f. (\lambda x. \lambda x. x)~n~(\Ya~\Suc{n}~f))
-\end{equation}
-
-And evaluates as intended:
-
-\begin{center}
-\begin{tabular}{ll}
-$\mapsto^*$ & $(\lambda n. \lambda f. (\lambda x. \lambda x. x)~n~(\Ya~\Suc{n}~f))~\Zero~(\lambda n. \lambda f. \dots)$ \\
-$\mapsto$ & $(\subst{\Zero}{n}\lambda f. (\lambda x. \lambda x. x)~n~(\Ya~\Suc{n}~f))~(\lambda n. \lambda f. \dots)$ \\
-= & $(\lambda f. (\lambda x. \lambda x. x)~\Zero~(\Ya~\Suc{\Zero}~f))~(\lambda n. \lambda f. \dots)$ \\
-$\mapsto$ & $(\lambda x. \lambda x. x)~\Zero~(\Ya~\Suc{\Zero}~(\lambda n. \lambda f. \dots))$
-\end{tabular}
-\end{center}
-
-Figure~\ref{fig:nats} shows the experimental results of evaluating this term 10,000 times.
-
-\begin{figure}[h]
-	\begin{center}
-	\includegraphics[width=0.48\textwidth]{rslc-naturals.pdf}
-	\end{center}
-	\caption{Computing random numbers with the term in Equation~\ref{eq:randomnat-fixed}. Log scale, lower is better. (I couldn't show $13$ because it showed up 0 times, and $\mathsf{log}~0 = -\infty$.)}
-	\label{fig:nats}
-\end{figure}
-
-\subsection{Uniform random number generation}
-\label{sec:uniform}
-
-\newcommand{\thing}[1]{\textcolor{Fuchsia}{#1}}
-
-The next thing I wanted to do was compute random numbers with a uniform distribution.
-Previously we were just selecting between two possible bindings for $x$, statically defined in the program text.
-I wondered, ``can I dynamically generate a term with $n$ shadow-bindings\footnote{Shadow binding: Sor/Wiz 3, Complete Arcane.
-Will save negates.} that causes a variable to have $n$ expressions in its $\rep{e}$ list?'' Something like:
-\[
-\thing{(\lambda x. (\lambda x. \dots (\lambda x.x)))}~0~1~2~\dots~n
-\]
-Of course this was where I had to iron out all the crazy parts in the previous sections, but since I already presented it in a manner consistent with it ain't being no thang, I'm going to make this part look easy:
-% Y2 (\x.x) N (\tn. case n of 0 => (\f. t 0) | S(n2) => (\f. Y2 ((\x.t)n) n2 f))
-
-\begin{center}
-\begin{tabular}{rcl}
-$\rho$ & = &
-\hspace{-1em}\multirow{4}{*}{
-\begin{tabular}{lllll}
-\multicolumn{5}{l}{$\lambda \thing{\bf t}. \lambda n.$} \\
-& \multicolumn{4}{l}{$\mathsf{case}~n~\mathsf{of}$} \\
-&& $\Zero$ & $\Rightarrow$ & $\lambda f.~\thing{\bf t}~\Zero$ \\
-&& $\Suc{n_2}$ &$\Rightarrow$& $\lambda f.~\Yab~\thing{((\lambda x.{\bf t})~n)}~n_2~f$
-\end{tabular}
-}
-\\
-&&\\
-&&\\
-&&\\
-&&\\
-$\mathsf{rand}[0,N]$ &=& $\Yab~\thing{(\lambda x.x)}~N~\rho$
-\end{tabular}
-\end{center}
-
-The variable $\thing{\bf t}$ accumulates ``$\lambda x$''s as $\rho$ recurses (in the $\Suc{n_2}$ case). You might ask: why does the $\lambda f$ need to be inside the $\mathsf{case}$ statement?
-Because otherwise in {\sc Eval-Case-2}$n_2$ would get captured in the thing substituted for $f$.
-
-A brief, very condensed, evaluation run-through:
-
-\begin{center}
-\begin{tabular}{lll}
-$\mathsf{rand}[0,2]$ &
-= & $\Yab~\thing{(\lambda x.x)}~2~\rho$ \\
-& $\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.x))~2)}~1~\rho$ \\
-& $\mapsto^*$ & $\Yab~\thing{((\lambda x.(\lambda x.(\lambda x.x))~2)~1)}~0~\rho$ \\
-& $\mapsto^*$ & $((\lambda x.(\lambda x.(\lambda x.x))~2)~1)~0$ \\
-& $\mapsto^*$ & $\subst{0}{x}x[1,2]$
-\end{tabular}
-\end{center}
-
-\section{Deterministic Programming with RSLC}
-
-Got here.
-Up until now I showed how random scope adds inherent nondeterminism that can be harnessed to simulate dice anda make silly graphs.
-The other thing Sully didn't believe would be possible was writing deterministic RSLC programs that {\em avoid} using ``the feature''.
-
-In the last section I already introduced the capture-free Y combinator(s). Implementing addition using that is no trouble:
-
-\newcommand{\addition}{\maltese}
-\begin{center}
-\begin{tabular}{rcl}
-$\addition$ & = &
-\hspace{-1em}\multirow{4}{*}{
-\begin{tabular}{lllll}
-\multicolumn{5}{l}{$\lambda m. \lambda n.$} \\
-& \multicolumn{4}{l}{$\mathsf{case}~m~\mathsf{of}$} \\
-&& $\Zero$ & $\Rightarrow$ & $\lambda f.~n$ \\
-&& $\Suc{m'}$ &$\Rightarrow$& $\lambda f.~\Suc{\Yab~m'~n~f}$
-\end{tabular}
-}
-\\
-&&\\
-&&\\
-&&\\
-&&\\
-$\mathsf{add}~M~N$ &=& $\Yab~M~N~\addition$
-\end{tabular}
-\end{center}
-
-But it turns out that multiplication, which uses addition, winds up needing the eager \textsf{let} construct.
-Among the several ways to implement it fully lazily, some introduce accidental randomness by letting some of \textsf{times}'s variables get captured, and others simply cause my implementation to run out of stack space or start swapping to disk when evaluating.
-Uh, anyway, here it is.
-
-\newcommand{\multiplication}{\bigstar}
-\newcommand{\multx}{\hat{x}}
-\newcommand{\multm}{\hat{m}}
-\newcommand{\multn}{\hat{n}}
-\newcommand{\multf}{\hat{f}}
-\newcommand{\multmm}{\hat{m}'}
-\begin{center}
-\begin{tabular}{rcl}
-$\multiplication$ & = &
-\hspace{-1em}\multirow{6}{*}{
-\begin{tabular}{lllll}
-\multicolumn{5}{l}{$\lambda \multm. \lambda \multn.$} \\
-& \multicolumn{4}{l}{$\mathsf{case}~\multm~\mathsf{of}$} \\
-&& $\Zero$ & $\Rightarrow$ & $\lambda \multf.~\Zero$ \\
-&& $\Suc{\multmm}$ &$\Rightarrow$& $\lambda \multf.$ \\
-&&
-\multicolumn{3}{l}{
-\begin{tabular}{lll}
-& \textsf{let} & $\multx~=~\Yab~\multmm~\multn~\multf$ \\
-& \textsf{in} & $\mathsf{add}~\multx~\multn$
-\end{tabular}
-}
-\end{tabular}
-}
-\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-$\mathsf{times}~M~N$ &=& $\Yab~M~N~\multiplication$
-\end{tabular}
-\end{center}
-Note that since $\mathsf{times}$ refers to $\mathsf{add}$, $\mathsf{times}$'s variable names must have different names. I put hats on them.
-
-Here's the factorial function:
-\newcommand{\factorial}{\textbf{!}}
-\newcommand{\factx}{\mathring{x}}
-\newcommand{\factn}{\mathring{n}}
-\newcommand{\factf}{\mathring{f}}
-\newcommand{\factnn}{\mathring{n}'}
-\begin{center}
-\begin{tabular}{rcl}
-$\factorial$ & = &
-\hspace{-1em}\multirow{6}{*}{
-\begin{tabular}{lllll}
-\multicolumn{5}{l}{$\lambda \factn.$} \\
-& \multicolumn{4}{l}{$\mathsf{case}~\factn~\mathsf{of}$} \\
-&& $\Zero$ & $\Rightarrow$ & $\lambda \factf.~\Suc{\Zero}$ \\
-&& $\Suc{\factnn}$ &$\Rightarrow$& $\lambda \factf.$ \\
-&&
-\multicolumn{3}{l}{
-\begin{tabular}{lll}
-& \textsf{let} & $\factx~=~\Ya~\factnn~\factf$ \\
-& \textsf{in} & $\mathsf{times}~\factn~\factx$
-\end{tabular}
-}
-\end{tabular}
-}
-\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-$\mathsf{fact}~N$ &=& $\Ya~N~\factorial$
-\end{tabular}
-\end{center}
-
-\newcommand{\fst}{\mathsf{fst}}
-\newcommand{\snd}{\mathsf{snd}}
-To compute fibonacci numbers, I used the standard Church encoding for pairs:
-\begin{eqnarray*}
-(e_1, e_2) &=& \lambda b.~b~e_1~e_2 \\
-\fst~e &=& e (\lambda x.\lambda y. x) \\
-\snd~e &=& e (\lambda x.\lambda y. y) \\
-\end{eqnarray*}
-Hence:
-\newcommand{\fib}{\Phi}
-\newcommand{\fibxa}{\tilde{x}_1}
-\newcommand{\fibxb}{\tilde{x}_2}
-\newcommand{\fibn}{\tilde{n}}
-\newcommand{\fibf}{\tilde{f}}
-\newcommand{\fibp}{\tilde{p}}
-\newcommand{\fibnn}{\tilde{n}'}
-\begin{center}
-\begin{tabular}{rcl}
-$\fib$ & = &
-\hspace{-1em}\multirow{6}{*}{
-\begin{tabular}{lllll}
-\multicolumn{5}{l}{$\lambda \fibn.$} \\
-& \multicolumn{4}{l}{$\mathsf{case}~\fibn~\mathsf{of}$} \\
-&& $\Zero$ & $\Rightarrow$ & $\lambda \fibf.~(\Zero,~\Suc{\Zero})$ \\
-&& $\Suc{\fibnn}$ &$\Rightarrow$& $\lambda \fibf.$ \\
-&&
-\multicolumn{3}{l}{
-\begin{tabular}{lll}
-& \textsf{let} & $\fibp~=~\Ya~\fibnn~\fibf$ \\
-& & $\fibxa~=~\fst~\fibp$ \\
-& & $\fibxb~=~\snd~\fibp$ \\
-& \textsf{in} & $(\fibxb,~\mathsf{add}~\fibxa~\fibxb)$
-\end{tabular}
-}
-\end{tabular}
-}
-\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-&&\\
-$\mathsf{fib}~N$ &=& $\Ya~N~\fib$
-\end{tabular}
-\end{center}
-
-At this point I could include some addition and multiplication tables like the kind you saw in 3rd grade, but it would be no surprise.
-Suffice to say that it works.
-
-\section{Applications}
-
-\begin{figure}[t]
-	\begin{center}
-	\includegraphics[width=0.48\textwidth]{rslc-rdr.pdf}
-	\end{center}
-	\caption{RSLC is a generalization of the Random Distance Run~\cite{rdr}.}
-	\label{fig:rdr}
-\end{figure}
-
-\begin{figure}[t]
-	\begin{center}
-	\includegraphics[width=0.40\textwidth]{rslc-d20.pdf}
-	\end{center}
-	\caption{RSLC can also be used to implement popular role-playing games.}
-	\label{fig:d20}
-\end{figure}
-
-Now that I showed in an abstract sense how RSLC can be used to express popular computations from conventional lambda calculus as well as some novel random programming techniques, let's do a real-world application.
-Using the uniform combinator from Section~\ref{sec:uniform}, we can make standard dice for any nonzero number of sides:
-
-\begin{eqnarray*}
-	\mathsf{d}(\Suc{N}) &=& \mathsf{add}~1~\mathsf{rand}[0,N] \\
-       2\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
-	&& \mathsf{in}~\mathsf{add}~2~x \\
-       3\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
-	&& \mathsf{in}~\mathsf{let}~y=\mathsf{add}~x~\mathsf{rand}[0,N] \\
-	&& \mathsf{in}~\mathsf{add}~3~y
-\end{eqnarray*}
-
-For $N = 5$, using the $2\mathsf{d}(\Suc{N})$ combinator (more commonly known as $2\mathsf{d}6$), we obtain a random distribution equivalent to the one employed in the Random Distance Run~\cite{rdr}, as shown in Figure~\ref{fig:rdr}.
-
-The value $N = 19$ yields the $\mathsf{d}20$ combinator, which generates the random distribution shown in Figure~\ref{fig:d20}.
-Prior work has shown this combinator to be quite useful in tabletop roleplaying games~\cite{dnd-mm,dnd-phb,dnd-dmg}.
-
-\section{Theoretical Concerns}
-
-Some programming language theoreticians will speak of the so-called ``Frame Rule'', which expresses abstraction: if a property $P$ holds for a certain function $f$, then in another function $g$ which invokes $x$, if $P(x)$ implies $P(g)$, then $P([f/x]g)$.
-
-In the context of RSLC, we scoff at this rule.
-
-\section{Concluding Remarks}
-
-RSLC is an extension to standard un(i)typed lambda calculus in which substitution is essentially completely broken.
-However, since substitution is broken in a particularly careful way, it is possible to write interesting programs that make use of the language's built-in nondeterminism.
-Furthermore, I showed that it is still possible to write a bunch of deterministic programs that are commonly used as examples/proofs-of-concept, as long as you put silly hats on some of your variables.
-
-I believe you could extend RSLC's substitution schema to a typed lambda calculus, and furthermore if it only substitutes for variables of the same type, it would even be type-safe.
-
-My implementation of RSLC in Haskell is available at \url{https://github.com/bblum/sigbovik/blob/master/RSLC/RSLC.hs}.
+In 2005, Roxor Games, Inc.
+released {\em In The Groove}, a dance rhythm music video arcade fitness game, in which players control a protagonist using their feet to step on floor-mounted directional indicators.
+The protagonist, shown in Figure~\ref{fig:protagonist}, takes the form of any number of arrow-shaped directional receptacles, and must navigate a world of similarly-shaped obstacles (henceforth ``arrows'') by consuming them with the appropriate receptacle.
+Roxor Games, Inc. In The Groove (henceforth ``ITG'')
 
@@
 
-\newpage
-
-\appendix
-
-\section{Discussion of Substitution Rules}
-\begin{figure*}[t]
-	\fbox{$\subst{e_0}{x}{e}$}
-	\TypeRules{
-		\SubstLamCapture \qquad \SubstLam \qquad
-		\SubstApp \\
-		\SubstVarX \qquad \SubstVarY \\
-		\SubstCaseCapture \\ \SubstCase \\
-		\SubstZero \qquad \SubstSuc \\
-		\SubstLetCapture \qquad \SubstLet
-	}
-	\fbox{$\capture{e_0}{x}{e}$}
-	\TypeRules{
-		\CaptureLam \qquad \CaptureApp \qquad
-		\CaptureVarX \qquad \CaptureVarY \\
-		\CaptureCase \\
-		\CaptureZero \qquad \CaptureSuc \qquad \CaptureLet
-	}
-	\caption{Complete RSLC substitution and capture rules.}
-	\label{fig:subst-complete}
-\end{figure*}
+%\onecolumn
+%
+%\appendix
 

### 2016-03-11 itg/code/itg.c

@@ -1,0 +1,196 @@
+#include <stdio.h>
+#include <stdbool.h>
+#include <assert.h>
+
+/*
+   1  2  3
+      U
+   4 L R 5
+      D
+   6  7  8
+
+ Figure 1. The Roxor, Inc. In The Groove Dance Game (henceforth "ITG") Dance
+ Step Direction Indicators (henceforth "Arrows"), UDLR, and the 8 possible
+ directions of facing.
+ */
+
+enum facing {
+	F_UL, F_U, F_UR,
+	F_L,  F_N, F_R,
+	F_DL, F_D, F_DR,
+	NUM_FACINGS,
+};
+
+enum arrow { U, D, L, R, NUM_ARROWS };
+
+enum facing whichway(enum arrow left_foot, enum arrow right_foot) {
+	// Foot switches: future work!
+	assert(left_foot != right_foot);
+	static enum facing way_table[NUM_ARROWS][NUM_ARROWS] = {
+		[U][D] = F_R,
+		[U][L] = F_DR, // xover
+		[U][R] = F_UR,
+		[L][U] = F_UL,
+		[L][D] = F_UR,
+		[L][R] = F_U,
+		[D][U] = F_L,
+		[D][L] = F_DL, // xover
+		[D][R] = F_UL,
+		[R][L] = F_D, // lateral
+		[R][U] = F_DL, // xover
+		[R][D] = F_DR, // xover
+	};
+	return way_table[left_foot][right_foot];
+}
+
+// For any two facings possible with consecutive itg steps (ie, only 1 foot
+// moves), shouldn't return more than 2.
+// Higher turniness is possible in future work using jumps!
+unsigned int turn_distance(enum facing before, enum facing after)
+{
+	assert(before < NUM_FACINGS && after < NUM_FACINGS);
+	static int turn_table[NUM_FACINGS][NUM_FACINGS] = {
+		[F_UL] = { 0, 1, 2,
+			   1, 0, 3,
+			   2, 3, 4, },
+		[F_U]  = { 1, 0, 1,
+			   2, 0, 2,
+			   3, 4, 3, },
+		[F_UR] = { 2, 1, 0,
+			   3, 0, 1,
+			   4, 3, 2, },
+		[F_L]  = { 1, 2, 3,
+			   0, 0, 4,
+			   1, 2, 3, },
+		[F_N]  = { 0, 0, 0,
+			   0, 0, 0,
[…]

### 2016-03-12 itg/paper.tex
write more paper

@@
 The protagonist, shown in Figure~\ref{fig:protagonist}, takes the form of any number of arrow-shaped directional receptacles, and must navigate a world of similarly-shaped obstacles (henceforth ``arrows'') by consuming them with the appropriate receptacle.
-Roxor Games, Inc. In The Groove (henceforth ``ITG'')
+Roxor Games, Inc.
+In The Groove (henceforth ``ITG'') is most commonly played using the ``cabinet'' form factor, shown in Figure~\ref{fig:cab}, which includes two large metal dance pads, each with four directional indicators (henceforth, also, ``arrows'').
+The game includes a library of rhythmic audio accompaniment files (henceforth, ``songs''), each of which is associated with one or more fixed patterns of arrows (henceforth, ``stepcharts'').
+These charts are often, but not always, synchronized to the beat of the song.
+During gameplay, the stepcharts appear on screen and scroll towards the protagonist avatar at a rate either fixed or variable (henceforth, ``BPM'').
+When the position of an arrow in the chart coincides with the avatar, the player must actuate the arrow of the corresponding direction.
+The game will judge the player's timing accuracy, and penalize or reward them accordingly with scores and life bar fill.
+A ``Fantastic'' judgement (as in Figure~\ref{fig:protagonist}) indicates a timing error not exceeding 15 milliseconds.
+Other judgements include Excellent, Great, Decent, Way Off, and Miss.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.25\textwidth]{protagonist.png}
+	\end{center}
+	\caption{ITG gameplay, including score indicator (top), protagonist avatar (mid), directional obstacles (low), and step judgement, life bar, and combo indicator (figure these out for yourself).}
+	\label{fig:protagonist}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.2\textwidth]{itg2cabinet.jpg}
+	\end{center}
+	\caption{An ITG cabinet. RIP in Peace, Roxor.}
+	\label{fig:cab}
+\end{figure}
+
+The game may be played in several modes, the most common of which supports up to two (2) players, each operating their own protagonist using either the left or right set of four arrows.
+This game mode may be played with or without the assist of a curved metal rod mounted behind the arrows (henceforth, ``bar''), as shown in Figure~\ref{fig:bengreg}.
+In other game modes, a single player may operate up to all 8 of the arrows.
+When all 8 are used, the game mode is known as ``Doubles'', as shown in Figure~\ref{fig:jim}, and is often associated with excessive No-Barring, use of hands and knees to operate the arrows, and impressive facial hair.
+In this body of work, we will focus exclusively on the Singles game mode, in which each player controls four arrows: one Up ($U$), one Down ($D$), one Left ($L$), and one Right ($R$).
+Without loss of generality, we further assume that only one player plays at a time, and that she or he will shower immediately afterward.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{bengreg.jpg}
+	\end{center}
+	\caption{ITG can be played ``Bar'' or ``No-Bar''. Bar players must wear sandals (like an idiot; who is that guy anyway?), while No-Barrers must always play on the right.}
+	\label{fig:bengreg}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.25\textwidth]{jim.jpg}
+	\end{center}
+	\caption{Doubles play is beyond the scope of our work. It requires (or perhaps produces?) a magnificent beard.}
+	\label{fig:jim}
+\end{figure}
 

### 2016-03-12 itg/paper.tex
Finish most of itg paper

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{amssymb}
@@
 \usepackage{graphicx}
+\usepackage{wasysym}
+\usepackage{harmony}
 \usepackage{marvosym}
-\usepackage{stmaryrd}
-\usepackage{epsdice}
 \usepackage{multirow}
@@
 \keywords
+in, the, groove
 
@@
 In The Groove (henceforth ``ITG'') is most commonly played using the ``cabinet'' form factor, shown in Figure~\ref{fig:cab}, which includes two large metal dance pads, each with four directional indicators (henceforth, also, ``arrows'').
+
 The game includes a library of rhythmic audio accompaniment files (henceforth, ``songs''), each of which is associated with one or more fixed patterns of arrows (henceforth, ``stepcharts'').
@@
 Other judgements include Excellent, Great, Decent, Way Off, and Miss.
+As a visual assist to the player, notes are coloured according to their beat granularity:
+\color{red}\Vier,
+\color{Purple}\Vier$^3$,
+\color{blue}\Acht
+\color{LimeGreen}\Sech,
+\color{Magenta}\Sech$^3$,
+\color{Dandelion}\Zwdr.
+\color{black}
 
@@
 	\end{center}
-	\caption{ITG gameplay, including score indicator (top), protagonist avatar (mid), directional obstacles (low), and step judgement, life bar, and combo indicator (figure these out for yourself).}
+	\caption{ITG gameplay, including score indicator (top), protagonist avatar (mid), directional obstacles (low), and step judgement, life bar, and combo indicator (figure these out for yourself, I'm getting tired).}
 	\label{fig:protagonist}
@@
 	\end{center}
-	\caption{An ITG cabinet. RIP in Peace, Roxor.}
+	\caption{An ITG cab. RIP in peace, Roxor \cite{konami}.}
 	\label{fig:cab}
@@
 
-The game may be played in several modes, the most common of which supports up to two (2) players, each operating their own protagonist using either the left or right set of four arrows.
+The game may be played in several modes, the most common supporting up to two (2) players, each operating their own protagonist using either the left or right set of four arrows.
 This game mode may be played with or without the assist of a curved metal rod mounted behind the arrows (henceforth, ``bar''), as shown in Figure~\ref{fig:bengreg}.
@@
 In this body of work, we will focus exclusively on the Singles game mode, in which each player controls four arrows: one Up ($U$), one Down ($D$), one Left ($L$), and one Right ($R$).
-Without loss of generality, we further assume that only one player plays at a time, and that she or he will shower immediately afterward.
+Without loss of generality, we further assume that only one player plays at a time, using exactly two feet at a time, and that she will shower immediately afterward.
 
@@
 	\begin{center}
-	\includegraphics[width=0.25\textwidth]{jim.jpg}
+	\includegraphics[width=0.2\textwidth]{jim.jpg}
 	\end{center}
@@
 
-\bibliographystyle{alpha}
-\bibliography{paper}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\onecolumn
-%
-%\appendix
+\section{Properties of Stepcharts}
+
+Being a game, one might assume that players find it fun to step the arrows in the indicated patterns, as opposed to, for example, stepping them in random unrelated patterns, or stepping off the pad entirely to go play Rocket League \cite{rocketleague}.
+Accordingly, step authors take care to produce stepcharts with interesting patterns.
+Typically, players step the arrows alternatingly with either foot.
+Step authors can either reinforce this tendency, by maintaining that left and right arrow targets always receive arrows of different (even or odd) sequential parity;
+or subvert it, forcing the player to step on the left arrow with their right foot (or v.v.).
+The former pattern is called a {\em stream}, and the latter pattern a {\em crossover}.
+Crossovers are loved by some players (Figure~\ref{fig:bengreg}, left), and hated by others (Figure~\ref{fig:bengreg}, right).
+
+Figure~\ref{fig:patterns} shows many common step patterns.
+Among these, note especially the {\em lateral} ($LURLDR$), in which the player briefly faces backwards (left foot on $R$ and right foot on $L$), and the {\em spin} ($LURDLU$), in which, yeah, you get the picture.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{patterns.jpg}
+	\end{center}
+	\caption{Common step patterns \cite{patterns}.}
+	\label{fig:patterns}
+\end{figure}
+
+\subsection{Difficulty}
+
+Stepcharts are rated in difficulty on a scale from 1 to 20 ``feet''\footnote{
+Other dance games \cite{konami}, which shall remain unnamed, feature wimpier stepcharts rated only up to 10, but as ITG players got better, they left those charts in the dust and began creating their own charts to challenge themselves with ridiculous difficulties.
+This paper's author plays ITG at the 13-14 level.}
+To date, difficulty is always judged by a human (typically the step author), and primarily reflects factors such as burstiness (footspeed), prolognedness of streams (stamina), and complexity of jumps, holds, and crossovers (technical).
+These measures of difficulty have been well-studied in prior works \cite{callofthehound,dawgsinthehouse}.
+
+\subsection{Facing}
+
+Regardless of the presence of crossovers, all step patterns will keep the player {\em facing} one way or another, and most include sequences that will change the player's facing back and forth.
+For example, when the player stands with left foot on $L$ and right foot on $U$, they are facing up-left, or $UL$.
+Standing on $D$ and $R$ also produces the same $UL$ facing, unless the player is crossed-over (left foot on $R$), in which case the facing is $DL$.
+Table~\ref{tab:facing} shows all possibilities of facing using two feet on four arrows.
+We exclude the possibility for both feet to be instantaneously on the same arrow (called a ``footswitch''), leaving this pattern to future work.
+
+\begin{table}[h]
+	\begin{center}
+	\begin{tabular}{cc|cccc}
+		& & \multicolumn{4}{c}{Right foot} \\
+		& & $\leftarrow$ & $\downarrow$ & $\uparrow$ & $\rightarrow$ \\
+		\hline
+		\multirow{4}{*}{Left foot}
+		%                 L   D   U   R
+		& $\leftarrow$  & - & $UR$ & $UL$ & $U$ \\
+		& $\downarrow$  & $DL$* & - & $L$ & $UL$ \\
+		& $\uparrow$    & $DR$* & $R$ & - & $UR$ \\
+		& $\rightarrow$ & $D\dagger$ & $DR$* & $DL$* & - \\
+
+	\end{tabular}
+	\end{center}
+	\caption{Facing directions. ``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
+	\label{tab:facing}
+\end{table}
+
+\subsection{Turning}
+
+A stepchart contains a {\em turn} when a sequence of steps changes the player's facing.
+
+\newtheorem{definition}{Definition}
+\begin{definition}[Turniness]
+	The turniness $\mathcal{T}$ of a single step is the angular distance (measured in increments of $\pi/4$) which that step changes the facing compared to the facing from the previous two steps.
+
+	The turniness $\mathcal{T}$ of a stream is the average of each step's turniness (excluding the first two for which it is undefined).
+\end{definition}
+
+Because only one foot may change at a time, it is easy to see that the maximum turniness of a single step is 2.
+These steps always involve one foot moving entirely across the middle of the pad (from $U$ to $D$, from $L$ to $R$, etc).
+By tradition, these are called {\em candle steps}, according to the imagery that if a candle were placed in the middle of the 4 arrows, this step would knock it over.
+I don't know why it couldn't be called a beer step instead, though.
+
+{\bf Turning is exhausting.} Consider the three stream patterns shown in Figure~\ref{fig:stream}.
+In the leftmost stream, the player must barely move her feet at all, never changing facing from $U$.
+In the middle stream, the player faces largely $UR$, with brief $U$ facings ($\mathcal{T}=1$) as she moves between the $L$/$D$ and $U$/$R$ arrows.
+Both of these streams are very easy to step without becoming fatigued.
+However, the rightmost stream features many candle steps (or beer steps), changing facing from $UR$ to $UL$ and back in the span of a single measure.
+ITG players hate this \cite{weirdtrick}!
+However, note that even in this stream, not {\em all} of the steps have $\mathcal{T}=2$: the {\em average turniness} of this stream is only 4/3.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.15\textwidth]{tower.png}
+	\includegraphics[width=0.15\textwidth]{staircase.png}
+	\includegraphics[width=0.15\textwidth]{twistystair.png}
+	\end{center}
+	\caption{Three streams of differing turniness.
+	Note how the color-coding of the notes by beat aids to discern turniness at a glance: the right foot is always on blue, and it moves among 1, 2, or 3 different arrows respectively.}
+	\label{fig:stream}
+\end{figure}
+
+We are interested in the following questions:
+\begin{enumerate}
+	\item What is the turniest possible stepchart?
+	\item How is maximum turniness affected by various constraints to contain no crossovers/laterals/etc?
+	\item How turny are real-world ITG charts?
+\end{enumerate}
+
+\section{Chart Synthesis}
+
+We developed a program for exploring all possible step patterns using a little-known yet powerful computational technique \cite{bruteforce1,bruteforce2}.
+This program computes the average turniness of every such pattern.
+However, while some players enjoy any step pattern including spins \cite{alanv},
+other players may prefer only crossovers and laterals in their charts,
+while still others prefer charts entirely crossover-free.
+Hence, our program is further capable of filtering patterns according to 5 predicates:
+\begin{itemize}
+	\item All patterns allowed
+	\item No spins
+	\item No 270s
+	\item No laterals
+	\item No crossovers
+\end{itemize}
+Note that each predicate captures a strict subset of step patterns compared to the one above it (a lateral is a crossover, and so forth).
+These predicates are implemented as shown in Figure~\ref{fig:eqn}.
+
+\begin{figure*}
+	\begin{eqnarray*}
+		\mathsf{is\_xover}(\phi) &=& \phi \equiv DL \vee \phi \equiv DR \\
+		\mathsf{is\_lateral}(\phi) &=& \phi \equiv D \\
+		\mathsf{is\_270}(\phi, \phi_1, \phi_2) &=& (\phi \equiv DR \wedge (\phi_1 \equiv DL \vee (\phi_1 \equiv D \wedge \phi_2 \equiv DL))) \vee \\
+                                              & & (\phi \equiv DL \wedge (\phi_1 \equiv DR \vee (\phi_1 \equiv D \wedge \phi_2 \equiv DR))) \\
+		\mathsf{is\_spin}(\phi, \phi_1, \phi_2, \phi_3) &=& (\phi \equiv R \vee \phi \equiv UR \vee \phi \equiv L \vee \phi \equiv UL) \wedge \mathsf{is\_270}(\phi_1, \phi_2, \phi_3)
+	\end{eqnarray*}
+	\caption{Formal definitions of crossovers, laterals, 270s, and spins.
+Each predicate takes as arguments the current facing $\phi$, and up to the three most recent previous facings, $\phi_1$, $\phi_2$, $\phi_3$.}
+	\label{fig:eqn}
+\end{figure*}
+
+Deciding on how long of step sequences we should search for is a tradeoff between accuracy, in judging turniness in small fractions, and (drumroll...) exponential explosion.
+To strike a balance we decided to experimentally measure turniness at a granularity of 1/8th, which requires searching for step sequences of length 18 (twice the denominator, plus 2 for the first 2 steps whose turniness is undefined).
+Figure~\ref{fig:graph} shows our results.
+The highest data point clocks in at 40,609,780: there are 40 million unique 18-step sequences, any patterns allowed, that have turniness 0.875.
+
+We see that, with one exception, allowing each new type of step pattern increases the maximum possible turniness of the chart.
+The exception is that 270s do not allow for any additional turniness over laterals, although they do allow for {\em more possible ways} to achieve maximum turniness.
+Despite no additional turniness, this is crucial information for stepchart authors, as ITG players will (usually) get bored of any chart that simply repeats the same pattern over and over.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{turniness.pdf}
+	\end{center}
+	\caption{Distribution of unique step sequences among turninesses, classified by what types of step patterns are allowed in the chart.}
+	\label{fig:graph}
+\end{figure}
+
+{\bf Maximum turniness!}
+Now to the paper's titular question: what step sequences underlie those data points on the far right of each curve?
+Though it would be tedious to transcribe them all into Stepmania's step editor \cite{stepmania}, we manually inspected many of them to select the most representative and/or photogenic ones from each category.
+We show these sequences in Figure~\ref{fig:sequences}.
+These visualizations allow us to manually compute the true turniness of each sequence, imagining the core sequence repeated ad infinitum, freeing ourselves from the 1/8th granularity imposed by our experimental setup.
+We find that the turniest crossovers have $\mathcal{T}=3/2$ and that the turniest laterals have $\mathcal{T}=8/5$.
+
+The case of 270s is irregular: we found no 18-step sequence turnier than the turniest laterals, and most notably,
+none of the sequences included a 270 step in the {\em middle} of the stream.
+Figure~\ref{fig:sequences}(d) shows a stream with crossovers and laterals and no 270 step until the very last one.
+To see why, imagine yourself on the pad here: the only possible 19th step that doesn't result in a spin is to step back on the same arrow your left foot is already on.
+That step has $\mathcal{T}=0$, which would undo any potential turniness gained from going into the 270.
+Nevertheless, 270s retain some real-world application, as demonstrated by \cite{utopia} as the very last step.
+
+\begin{figure*}[t]
+	\begin{center}
+	\begin{tabular}{ccccc}
+		\includegraphics[width=0.17\textwidth]{result-1-3-vanilla.png} &
+		\includegraphics[width=0.17\textwidth]{result-1-5-xover.png} &
+		\includegraphics[width=0.17\textwidth]{result-1-6-lats.png} &
+		\includegraphics[width=0.17\textwidth]{result-irregular-270.png} &
+		\includegraphics[width=0.17\textwidth]{result-2-spin.png} \\
+		\small (a) Vanilla, $\mathcal{T}=4/3$. &
+		\small (b) Crossovers, $\mathcal{T}=3/2$. &
+		\small (c) Laterals, $\mathcal{T}=8/5$. &
+		\small (d) 270s (irregular). &
+		\small (e) Spins, $\mathcal{T}=2$.
+	\end{tabular}
+	\end{center}
+	\caption{The turniest possible stepcharts.}
+	\label{fig:sequences}
+\end{figure*}
+
+\section{Existing Chart Analysis}
+
+TODO % TODO
+
+\section{Conclusion}
+
+In conclusion, LURR DURR!
+
+\bibliographystyle{abbrvnat}
+\bibliography{citations}
+
+\appendix
+
+\section{Most turny patterns}
+
+% TODO: figure out how to uniquify these.
+
+\subsection{Vanilla}
+
+Following are the most turny 18-step patterns that are totally vanilla, with turniness of 1.375. (Manual inspection reveals the true turniness to be 4/3.)
+
+\noindent
+URDULDURDULDURDULD
+DRUDLUDRUDLUDRUDLU
+
+\subsection{Crossovers}
+
+Following are the most turny 18-step patterns that may include crossovers only, with turniness of 1.5.
+
+\noindent
+URDULDURDLDRULURDL
+URDULDRDLUDRULURDL
+URDULDRDLURULDURDL
+URDULDRDLURULDRDLU
+URDLDRUDLUDRULURDL
+URDLDRUDLURULDURDL
+URDLDRUDLURULDRDLU
+URDLDRULURDULDURDL
+URDLDRULURDULDRDLU
+URDLDRULURDLDRUDLU
+URDLDRULURDLDRULUR
+URDLDRULURDLDRURDL
+URDLDRULURDRULURDL
+URDLDRURDLDRULURDL
+URDRULURDLDRULURDL
+DRUDLUDRULURDLDRUL
+DRUDLURULDURDLDRUL
+DRUDLURULDRDLUDRUL
+DRUDLURULDRDLURULD
+DRULURDULDURDLDRUL
+DRULURDULDRDLUDRUL
+DRULURDULDRDLURULD
+DRULURDLDRUDLUDRUL
+DRULURDLDRUDLURULD
+DRULURDLDRULURDULD
+DRULURDLDRULURDLDR
+DRULURDLDRULURDRUL
+DRULURDLDRURDLDRUL
+DRULURDRULURDLDRUL
+DRURDLDRULURDLDRUL
+LUDRULURDLDRULURDL
+LURULDURDLDRULURDL
+LURULDRDLUDRULURDL
+LURULDRDLURULDURDL
+LURULDRDLURULDRDLU
+LDURDLDRULURDLDRUL
+LDRDLUDRULURDLDRUL
+LDRDLURULDURDLDRUL
+LDRDLURULDRDLUDRUL
+LDRDLURULDRDLURULD
+
+\subsection{Laterals}
+
+Following are the most turny 18-step patterns that may include crossovers and laterals, with turniness of 1.625. (Manual inspection reveals the true turniness to be 8/5.)
+
+\noindent
+URDLDRULRDLURLDRUL
+URDLRULDRDLURLDRUL
+URDLRULDRLURDLDRUL
+URDLRULDRLURDLRULD
+DRULURDLRULDRLURDL
+DRULRDLURULDRLURDL
+DRULRDLURLDRULURDL
+DRULRDLURLDRULRDLU
+LURLDRULRDLURLDRUL
+LDRLURDLRULDRLURDL
+
+\subsection{270s}
+
+Following are the most turny 18-step patterns that may include crossovers, laterals, and 270s, with turniness of 1.625.
+(Manual inspection reveals irregularity; these streams exhibit no core pattern that can be repeated to retain the maximum turniness.)
+
+\noindent
+URDULDRLURDLRULDRU
+URDLDRULURDLRULDRU
+URDLDRULRDLURULDRU
+URDLDRULRDLURLDRUL
+URDLRULDURDLRULDRU
+URDLRULDRDLURULDRU
+URDLRULDRDLURLDRUL
+URDLRULDRLURDULDRU
+URDLRULDRLURDLDRUL
+URDLRULDRLURDLRULD
+DRUDLURLDRULRDLURD
+DRULURDLDRULRDLURD
+DRULURDLRULDRDLURD
+DRULURDLRULDRLURDL
+DRULRDLUDRULRDLURD
+DRULRDLURULDRDLURD
+DRULRDLURULDRLURDL
+DRULRDLURLDRUDLURD
+DRULRDLURLDRULURDL
+DRULRDLURLDRULRDLU
+LURULDRLURDLRULDRU
+LURLDRULURDLRULDRU
+LURLDRULRDLURULDRU
+LURLDRULRDLURLDRUL \\
+LDRDLURLDRULRDLURD
+LDRLURDLDRULRDLURD
+LDRLURDLRULDRDLURD
+LDRLURDLRULDRLURDL
+
+\subsection{Spins}
+
+Following are the most turny 18-step patterns that may include any pattern, with turniness of 2.
+
+\noindent
+URDLURDLURDLURDLUR
+DRULDRULDRULDRULDR
+LURDLURDLURDLURDLU
+LDRULDRULDRULDRULD
 

### 2016-03-12 ode/ode.tex
Finish most of itg paper

@@
 I couldn't believe what he was demanding:\\
-To rerun my test suite, perhaps out of spite;\\
+To rerun my trials, perhaps out of spite;\\
 An unr'lated paper he asked me to cite!\\

### 2016-03-12 itg/code/itg.c

@@ -87,14 +87,14 @@ bool is_270(enum facing new_facing, enum facing last_facing, enum facing two_fac
 	if (new_facing == F_DR) {
 		if (last_facing == F_DL) {
 			return true; // LURD
-		} else if (last_facing == F_D && two_facings_ago == F_DR) {
+		} else if (last_facing == F_D && two_facings_ago == F_DL) {
 			return true; // LRDLRD
 		}
 	// symmetric case
 	} else if (new_facing == F_DL) {
 		if (last_facing == F_DR) {
 			return true; // RULD
-		} else if (last_facing == F_D && two_facings_ago == F_DL) {
+		} else if (last_facing == F_D && two_facings_ago == F_DR) {
 			return true; // RLDRLD
 		}
 	}
@@ -103,10 +103,12 @@ bool is_270(enum facing new_facing, enum facing last_facing, enum facing two_fac
 bool is_spin(enum facing new_facing, enum facing last_facing, enum facing two_facings_ago, enum facing three_facings_ago) {
 	// Patterns such as LURD are allowed but you have to twist back around the way you came
 	// from. So check at facings L, R, UR, and UL.
-	return (new_facing == F_R || new_facing == F_UR) && is_270(last_facing, two_facings_ago, three_facings_ago);
+	return (new_facing == F_R || new_facing == F_UR ||
+		new_facing == F_L || new_facing == F_UL) &&
+		is_270(last_facing, two_facings_ago, three_facings_ago);
 }
 
-#define SEARCHDEPTH 16
+#define SEARCHDEPTH 18
 #define MAX(x, y) ({ typeof(x) __x = (x); typeof(y) __y = (y); __x < __y ? __y : __x; })
 
 enum search_rule { NONE = 0, NO_SPINS, NO_270S, NO_LATERALS, NO_XOVERS, NUM_RULES };
@@ -191,6 +193,6 @@ unsigned int search(enum arrow *chart, unsigned int index, unsigned int turnines
 void main()
 {
 	enum arrow chart[SEARCHDEPTH];
-	unsigned int result = search(chart, 0, 0, NO_XOVERS);
+	unsigned int result = search(chart, 0, 0, NONE); // can customize this
 	printf("Best turniness %u\n", result);
 }

### 2016-03-12 itg/paper.tex
minor revisions to itg
ITG: implement .sm file analysis, and discuss some small matters like minimum turniness.

@@
 As a visual assist to the player, notes are coloured according to their beat granularity:
-\color{red}\Vier,
-\color{Purple}\Vier$^3$,
-\color{blue}\Acht
-\color{LimeGreen}\Sech,
-\color{Magenta}\Sech$^3$,
-\color{Dandelion}\Zwdr.
-\color{black}
+\color{red}\Vier\color{black},
+\color{Purple}\Vier$^3$\color{black},
+\color{blue}\Acht\color{black},
+\color{Purple}\Acht$^3$\color{black},
+\color{LimeGreen}\Sech\color{black},
+\color{Magenta}\Sech$^3$\color{black},
+\color{Dandelion}\Zwdr\color{black}.
 
@@
 	\end{center}
-	\caption{Common step patterns \cite{patterns}.}
+	\caption{Common patterns \cite{patterns}. Honk honk!}
 	\label{fig:patterns}
@@
 
-Stepcharts are rated in difficulty on a scale from 1 to 20 ``feet''\footnote{
+Stepcharts are rated in difficulty on a scale from 1 to 20 ``feet'' \cite{simplylove}\footnote{
 Other dance games \cite{konami}, which shall remain unnamed, feature wimpier stepcharts rated only up to 10, but as ITG players got better, they left those charts in the dust and began creating their own charts to challenge themselves with ridiculous difficulties.
-This paper's author plays ITG at the 13-14 level.}
+This paper's author plays ITG at the 13-14 level.}.
 To date, difficulty is always judged by a human (typically the step author), and primarily reflects factors such as burstiness (footspeed), prolognedness of streams (stamina), and complexity of jumps, holds, and crossovers (technical).
@@
 ITG players hate this \cite{weirdtrick}!
-However, note that even in this stream, not {\em all} of the steps have $\mathcal{T}=2$: the {\em average turniness} of this stream is only 4/3.
+However, note that even in this stream, not {\em all} of the steps have $\mathcal{T}=2$: the {\em average} turniness of this stream is only 4/3.
 
@@
 other players may prefer only crossovers and laterals in their charts,
-while still others prefer charts entirely crossover-free.
-Hence, our program is further capable of filtering patterns according to 5 predicates:
+while still others prefer charts completely vanilla.
+Hence, our program is further capable of filtering patterns by 5 predicates:
 \begin{itemize}
@@
 	\item No laterals
-	\item No crossovers
+	\item No crossovers (vanilla)
 \end{itemize}
@@
 		\mathsf{is\_xover}(\phi) &=& \phi \equiv DL \vee \phi \equiv DR \\
-		\mathsf{is\_lateral}(\phi) &=& \phi \equiv D \\
+		\mathsf{is\_lat}(\phi) &=& \phi \equiv D \\
 		\mathsf{is\_270}(\phi, \phi_1, \phi_2) &=& (\phi \equiv DR \wedge (\phi_1 \equiv DL \vee (\phi_1 \equiv D \wedge \phi_2 \equiv DL))) \vee \\
@@
 To strike a balance we decided to experimentally measure turniness at a granularity of 1/8th, which requires searching for step sequences of length 18 (twice the denominator, plus 2 for the first 2 steps whose turniness is undefined).
+Also, our program WLOG restricts streams to always start on the left foot, and never to start crossed over.
 Figure~\ref{fig:graph} shows our results.
-The highest data point clocks in at 40,609,780: there are 40 million unique 18-step sequences, any patterns allowed, that have turniness 0.875.
+The highest data point clocks in at 40,609,780: there are over 40 million unique 18-step sequences, any patterns allowed, that have turniness 0.875.
 
@@
 That step has $\mathcal{T}=0$, which would undo any potential turniness gained from going into the 270.
-Nevertheless, 270s retain some real-world application, as demonstrated by \cite{utopia} as the very last step.
+Nevertheless, 270s retain some real-world application, as demonstrated by \cite{utopia} which similarly employs a 270 as the very last step.
+
+Perhaps predictably, pure spins make for the turniest possible chart. There are only two possible ways to candle every step; one's name is ``clockwise'', and you can guess the other.
+
+{\bf Minimum turniness.}
+Each of the 5 categories shares the same data point at $\mathcal{T}=0$: there are 7 possible streams that never turn at all.
+These are the 7 tower patterns $LULU$, $LRLR$, $LDLD$, $URUR$, $DRDR$, $UDUD$, and $DUDU$.
+We already showed the first of these in Figure~\ref{fig:stream}, left.
+For further treatment of minimally turny charts, we refer the reader to \cite{deltamax}.
 

### 2016-03-12 itg/code/itg.c

@@ -2,6 +2,12 @@
 #include <stdbool.h>
 #include <assert.h>
 
+#define analysis_main main
+
+/******************************************************************************
+ * Common helpers
+ ******************************************************************************/
+
 /*
    1  2  3
       U
@@ -30,15 +36,19 @@ enum facing whichway(enum arrow left_foot, enum arrow right_foot) {
 		[U][D] = F_R,
 		[U][L] = F_DR, // xover
 		[U][R] = F_UR,
+		//[U][U] = F_U, // footswitch
 		[L][U] = F_UL,
 		[L][D] = F_UR,
 		[L][R] = F_U,
+		//[L][L] = F_L, // xover footswitch
 		[D][U] = F_L,
 		[D][L] = F_DL, // xover
 		[D][R] = F_UL,
+		//[D][D] = F_U, // footswitch (to be theoretically beautiful, this would be F_D, a backwards switch, but who the hell does that?)
 		[R][L] = F_D, // lateral
 		[R][U] = F_DL, // xover
 		[R][D] = F_DR, // xover
+		//[R][R] = F_R, // xover footswitch
 	};
 	return way_table[left_foot][right_foot];
 }
@@ -80,6 +90,83 @@ unsigned int turn_distance(enum facing before, enum facing after)
 	};
 	return turn_table[before][after];
 }
+
+/******************************************************************************
+ * Analysis
+ ******************************************************************************/
+
+// 1 = tap; 2 = hold; 3 = hold release; 4 = roll; M = mine
+// IDK what stepmania does for lifts/fakes; I don't know of any chart that uses them anyway.
+#define COULD_BE_NOTE(c) (((c) >= '0' && (c) <= '9') || (c) == 'M' || (c) == 'm')
+#define IS_NOTE(c) ((c) == '1' || (c) == '2' || (c) == '4')
+
+bool get_note(char *buf, enum arrow *note)
+{
+	char l, d, u, r;
+	if (sscanf(buf, "%c%c%c%c\n", &l, &d, &u, &r) == 4) {
+		if (COULD_BE_NOTE(l) && COULD_BE_NOTE(d) && COULD_BE_NOTE(u) && COULD_BE_NOTE(r)) {
+			if (IS_NOTE(l) && !IS_NOTE(d) && !IS_NOTE(u) && !IS_NOTE(r)) {
+				*note = L;
+				return true;
[…]

### 2016-03-13 itg/paper.tex
Finish itg paper(?)

@@
 	We study the turniness of many existing stepcharts, and present a novel (but unsurprising) approach to automatically generating maximally (or minimally) turny charts.
+	Among real-world songs, we find stepcharts with overall turniness ranging from 0\% to 81.33\% of the theoretical maximum.
 
@@
 	\begin{center}
-	\includegraphics[width=0.25\textwidth]{protagonist.png}
+	\includegraphics[width=0.2\textwidth]{protagonist.png}
 	\end{center}
@@
 	\begin{center}
-	\includegraphics[width=0.2\textwidth]{itg2cabinet.jpg}
+	\includegraphics[width=0.18\textwidth]{itg2cabinet.jpg}
 	\end{center}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Existing Chart Analysis}
 
-TODO % TODO
+Of course, composing a stepchart of nothing but the theoretically-turniest patterns would be extremely tiresome (in more ways than one).
+We turn our attention to existing ITG charts which were developed prior to this study, to investigate how much ITG players turn in the real world.
+We processed every stepchart in our collection, which includes 70+ song packs.
+Most well-known pack series from the ITG community are included, such as {\em Cirque}, {\em ITG}, {\em Mute Sims}, {\em Pendulum}, {\em r21*}, {\em SPEEDCOOOORE}, {\em The Legend of Zim}, and {\em Valex's Magical 4-Arrow Adventure}.
+
+{\bf Limitations.}
+Real-world stepcharts include many patterns not accounted for in our analysis.
+We did not expect much trouble from jumps and jacks, opting simply to ignore them and count the turns among the surrounding single-steps.
+Footswitches and doublesteps, however, can foil this strategy, as our analysis will inadvertently begin stepping with feet inverted from where they should be.
+Merely getting turned around is not a problem (computers don't have eyes), but the act of turning around involves some candle-steps, which could cause a footswitch-heavy chart to seem very spinny even though the proper footing would never make the player cross over at all.
+Coping with these limitations is beyond the scope of our work, although we will highlight some experimental false-positives later.
+
+\subsection{Results}
+
+As shown in Figure~\ref{fig:distribution}, the vast majority of real-world stepcharts have average turniness between 0.6 and 1.0.
+We don't really care about those.
+
+\begin{figure}[t]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{realworld.pdf}
+	\end{center}
+	\caption{Distribution of turniness among 8,278 existing ITG charts.}
+	\label{fig:distribution}
+\end{figure}
+
+{\bf The Least Turny Charts.}
+16 charts in our collection have $\mathcal{T}=0$.
+Most of these are beginner charts, rated 2 or 3 feet and with no more than 100 steps.
+\cite{deltamax} stands out, as a 15 with 945 steps.
+God is that chart ever annoying.
+
+Ignoring all beginner-difficulty and turn-free charts, the least turny stepchart is {\em Sick Dubstep Track} \cite{sickdubstep}, a 4-measure joke rated 6,482 feet, the majority of which is shown in Figure~\ref{fig:example}(a).
+
+{\bf The Turniest Charts.}
+No joker has yet dared to make the turniest possible chart (all spins), but one chart comes close.
+Standing alone as the only chart turnier than 1.5 is the Easy 3 for {\em DO ME} \cite{dome}, which is mostly spins punctuated by a few breaks, as shown in Figure~\ref{fig:example}(d).
+\cite{alanv} loves this chart.
+
+Actually, 20 of the 30 turniest charts are all from DDR packs, most of which are Easies.
+Although DDR charts are known for being fairly crossovery in the upper ranges of difficulty, a close inspection of the easier charts will show that very little care was put in while writing them to avoid excessive double-stepping.
+It's no wonder new ITG players, who used to play DDR, always double-step even when they don't need to. Sheesh, Konami.
+
+Beyond these extreme examples, we are interested in some more normal charts people might actually like to play.
+Table~\ref{tab:feets} shows the turniest charts of each level, for the most popular difficulty levels ranging from 7 to 17.
+So, next time you want to work on your laterals or candles, you have me to thank!
+
+Surprisingly, {\em BRILLIANT 2U (Orchestra-Groove)} \cite{brilliant2u} is full of 270s. I didn't think anybody ever put those in real charts.
+Finally, an honourable mention goes to {\em Oedo Hop} \cite{oedo}, the 2nd turniest 7 (and 4th turniest chart overall!) with $\mathcal{T}=1.384$.
+\cite{deltax} really likes that one so we had to mention it.
+
+\begin{figure*}[t]
+	\begin{center}
+	\begin{tabular}{cccc}
+		\hspace{-2em}
+		\begin{tabular}{c}
+		\includegraphics[width=0.18\textwidth]{sick-dubstep-track.png} \\
+		\small
+		\begin{tabular}{l}
+			(a) {\em Sick Dubstep Track}, $\mathcal{T}=0.26$. \\
+		The least turny non-trivial \\
+		chart.\\
+		\\
+		\end{tabular}
+		\end{tabular}
+		&
+		\hspace{-1em}
+		\begin{tabular}{c}
+		\includegraphics[width=0.18\textwidth]{mr-saxobeat.png} \\
+		\small
+		\begin{tabular}{l}
+			(b) {\em Mr. Saxobeat}, $\mathcal{T}=1.04$. \\
+		Look at all those candles! \\
+		No wonder it was so hard. \\
+		\\
+		\end{tabular}
+		\end{tabular}
+		&
+		\hspace{-1em}
+		\begin{tabular}{c}
+		\includegraphics[width=0.185\textwidth]{flames.png} \\
+		\small
+		\begin{tabular}{l}
+			(c) {\em Flames of the Sky}, $\mathcal{T}=1.04$. \\
+		False-positive with foot-\\
+		switches. Actually faces \\
+		$UR$/$U$/$UL$ the whole time.
+		\end{tabular}
+		\end{tabular}
+		&
+		\hspace{-1em}
+		\begin{tabular}{c}
+		\includegraphics[width=0.18\textwidth]{do-me.png} \\
+		\small
+		\begin{tabular}{l}
+			(d) {\em DO ME}, $\mathcal{T}=1.63$. \\
+		I'll pass, thanks.\\
+		\\
+		\\
+		\end{tabular}
+		\end{tabular}
+	\end{tabular}
+	\end{center}
+	\caption{Real-world ITG charts of varying degrees of turniness.}
+	\label{fig:example}
+\end{figure*}
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c}
+		Ft. & Name & Pack & $\mathcal{T}$ \\
+		\hline
+		7 & Trip Machine & DDR 1st & 1.407 \\
+		%7 & Oedo Hop & r2112 & 1.38 \\
+		%  & (honorable mention; way better chart) & & \\
+		8 & Brilliant 2U (O-G) & DDR 2ndMIX & 1.347 \\
+		9 & Hearts of Ice & In The Groove 3 & 1.127 \\
+		10 & Visible Noise & In The Groove 2 & 1.188 \\
+		11 & KKKKK Come On & r2112 & 1.145 \\
+		12 & Winnipeg is Fucking Over & best of r21freak ii & 1.150 \\
+		13 & Mr. Saxobeat (*) & Sexuality Violation 2 & 1.044 \\
+		14 & Fancy Footwork ($\dagger$) & Cirque du Zeppelin & 1.026 \\
+		15 & Flames of the Sky (*,$\dagger$) & Ft. Rapids VII &  1.037 \\
+		16 & {\em (no 16 with $\mathcal{T}\ge 1$)} & - & - \\
+		17 & Superluminal & Ft. Rapids VII & 1.058 \\
+	\end{tabular}
+	\end{center}
+	\caption{The turniest existing stepcharts of each difficulty.
+Songs marked (*) are also shown in Figure~\ref{fig:example}.
+Songs marked ($\dagger$) are false-positives resulting from too many footswitches incorrectly analyzed as spins.}
+	\label{tab:feets}
+\end{table}
+
+{\bf The Turniest Other Things.}
+Although our analysis picked out a good many turny charts, any ITG player will inevitably get tired of playing the same ones over and over.
+Hence, we are also interested in an aggregated turniness profile to show the average scores of many different stepcharts at once.
+We computed average turninesses for stepcharts grouped {\em by pack} and {\em by step artist}.
+Abbreviated results for the former are shown in Table~\ref{tab:pack} and for the latter in Table~\ref{tab:artist}.
+For the sake of players who dislike turniness, we also list the lowest-scoring packs and artists in addition to the highest-scoring ones.
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|c|c}
+		Song Pack & \# of Charts & Avg. $\mathcal{T}$ \\
+		\hline
+		Valex's Magical 4-Arrow Adventure 5 & 150 & 0.964 \\
+		Valex's Magical 4-Arrow Adventure 4 & 170 & 0.956 \\
+		Valex's Magical 4-Arrow Adventure 3 & 155 & 0.950 \\
+		Valex's Magical 4-Arrow Adventure 7 & 150 & 0.937 \\
+		Valex's Magical 4-Arrow Adventure 2 & 130 & 0.929 \\
+		Valex's Magical 4-Arrow Adventure   & 95 & 0.919 \\
+		Valex's Magical 4-Arrow Adventure 6 & 145 & 0.917 \\
+		Subluminal & 16 & 0.906 \\
+		The Legend of Zim 1 & 26 & 0.906 \\
+		FoxyMix 5 & 30 & 0.901 \\
+		FoxyMix 4 - Nuclear Overdrive & 68 & 0.899 \\ 
+		Mute Sims 9 & 48 & 0.886 \\
+		In The Groove Rebirth + & 147 & 0.881 \\
+		StepMania 5 & 5 & 0.869 \\
+		\multicolumn{3}{c}{\normalsize\dots} \\
+		Stephcharts and Richarts & 81 & 0.770 \\
+		Tachyon Delta & 32 & 0.769 \\
+		Tachyon Alpha & 151 & 0.769 \\
+		Sexuality Violation 2 & 219 & 0.766 \\
+		TLOES Chapter 1 & 120 & 0.765 \\
+		Fort Rapids V (12s-16s) r3 Final & 98 & 0.759 \\
+		Noisiastreamz & 20 & 0.758 \\
+		In The Groove & 331 & 0.754 \\
+		Pendulum & 150 & 0.747 \\
+		SPEEDCOOOORE 2 & 70 & 0.731 \\
+		SPEEDCOOOORE & 47 & 0.709 \\
+		The Legend of Zim 3 & 40 & 0.702 \\
+		The Legend of Zim 4 & 44 & 0.701 \\
+		The Legend of Zim 5 & 19 & 0.667 \\
+	\end{tabular}
+	\end{center}
+	\caption{The turniest and least turny stepfile packs. Wow, great job Valex! Zim had a good thing going but has been slacking off recently.}
+	\label{tab:pack}
+\end{table}
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|c|c}
+		Step Artist & \# of Charts & Avg. $\mathcal{T}$ \\
+		\hline
+		M. Emirzian & 25 & 0.981 \\
+		sssmsm & 40 & 0.980 \\
+		M. Calfin & 16 & 0.947 \\
+		J. Berkowitz (Valex) & 1016 & 0.940 \\
+		B. Dinh & 10 & 0.919 \\
+		J. DeGarmo & 24 & 0.915 \\
+		X. Ythar & 17 & 0.914 \\
+		R. McKanna & 59 & 0.895 \\
+		R. Woods & 18 & 0.892 \\
+		Fraxtil & 23 & 0.880 \\
+		\multicolumn{3}{c}{\normalsize\dots} \\
+		I. Pyles & 175 & 0.754 \\
+		P. Shanklin & 29 & 0.752 \\
+		Gazebo & 355 & 0.751 \\
+		zimlord & 199 & 0.750 \\
+		C. Emirzian & 17 & 0.736 \\
+		Mootz & 22 & 0.734 \\
+		B. Abear & 27 & 0.729 \\
+		Insane Steve & 41 & 0.718 \\
+		Blazing & 15 & 0.696 \\
+		C. Foy (\smiley) & 138 & 0.678 \\
+	\end{tabular}
+	\end{center}
+	\caption{The turniest and least turny step authors. Only authors with 10 or more charts are included.}
+	\label{tab:artist}
+\end{table}
 
@@
 
-In conclusion, LURR DURR!
+In conclusion, $LURR$ $DURR$!
 
-\bibliographystyle{abbrvnat}
-\bibliography{citations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \appendix
+
+\section{Linx}
+
+Our analysis program can be found at \url{https://github.com/bblum/sigbovik/tree/master/itg/code}, and our full experimental data-set is available at \url{http://tinyurl.com/turniness}.
 
@@
 
+\bibliographystyle{abbrvnat}
+\bibliography{citations}
+
 \end{document}

### 2016-03-13 itg/code/get-chart.sh

@@ -1,0 +1,60 @@
+#!/bin/sh
+
+FILE="$1"
+if [ ! -f "$FILE" ]; then
+	echo "no such file $file"
+	exit 1
+fi
+
+function process() {
+	FOUND_SINGLES_CHART=0
+	NEXT_LINE_SINGLES_CHART=0 # used for finding step artist
+	FOUND_DIFFICULTY=0
+	NEXT_LINE_DIFFICULTY=0 # used for finding difficulty rating
+	MAYBE_STEP_ARTIST=
+	STEP_ARTIST=
+	NUM_FEET=
+	CHART_CONTENT=
+	while read line; do
+		if [ "$FOUND_DIFFICULTY" = 1 ]; then
+			if [ "$NEXT_LINE_DIFFICULTY" = 1 ]; then
+				# Difficulty category was found immediately previous line; number of feet will be here.
+				NUM_FEET=`echo "$line" | sed 's/^ *//' | sed 's/:$//' | sed 's/:\r$//'`
+				NEXT_LINE_DIFFICULTY=0
+			elif echo "$line" | grep "dance-.*:" >/dev/null; then
+				# End of our chart.
+				break;
+			else
+				# Line in middle of our chart. Print it!
+				CHART_CONTENT="${CHART_CONTENT}${line}\n"
+			fi
+		elif [ "$FOUND_SINGLES_CHART" = 1 ]; then
+			if [ "$NEXT_LINE_SINGLES_CHART" = 1 ]; then
+				# "dance-single:" was found immediately previous line. Artist will be here.
+				MAYBE_STEP_ARTIST=`echo "$line" | sed 's/^ *//' | sed 's/:$//' | sed 's/:\r$//'`
+				NEXT_LINE_SINGLES_CHART=0
+			elif echo "$line" | grep "$DIFF:" >/dev/null; then
+				# "dance-single:" was found a while ago. Look for difficulty category.
+				FOUND_DIFFICULTY=1
+				NEXT_LINE_DIFFICULTY=1
+				STEP_ARTIST="$MAYBE_STEP_ARTIST"
+			fi
+		fi
+		if echo "$line" | grep "dance-single:" >/dev/null; then
+			FOUND_SINGLES_CHART=1
+			NEXT_LINE_SINGLES_CHART=1
+		fi
+
+	done
+	if [ "$FOUND_DIFFICULTY" = 1 ]; then
[…]

### 2016-03-13 itg/code/itg.c

@@ -160,7 +160,7 @@ void analysis_main()
 			notes++;
 		}
 	}
-	printf("%u turniness (x1000) (%u turns / %u notes)\n", (turniness * 1000 / notes), turniness, notes);
+	printf("%u turniness (x10000) (%u turns / %u notes)\n", (turniness * 10000 / notes), turniness, notes);
 }
 
 

### 2016-03-13 itg/paper.tex
fix citations
some more fixes
more fixes

@@
 In the leftmost stream, the player must barely move her feet at all, never changing facing from $U$.
-In the middle stream, the player faces largely $UR$, with brief $U$ facings ($\mathcal{T}=1$) as she moves between the $L$/$D$ and $U$/$R$ arrows.
+In the middle stream, the player faces largely $UR$, with brief $U$ and $R$ facings ($\mathcal{T}=1$) as she moves between the $L$/$D$ and $U$/$R$ arrows.
 Both of these streams are very easy to step without becoming fatigued.
@@
 Most of these are beginner charts, rated 2 or 3 feet and with no more than 100 steps.
-\cite{deltamax} stands out, as a 15 with 945 steps.
+$\Delta$MAX \cite{deltamax} stands out, as a 15-footer with 945 steps.
 God is that chart ever annoying.
@@
 Ignoring all beginner-difficulty and turn-free charts, the least turny stepchart is {\em Sick Dubstep Track} \cite{sickdubstep}, a 4-measure joke rated 6,482 feet, the majority of which is shown in Figure~\ref{fig:example}(a).
+It has $\mathcal{T}=0.263$.
 
@@
 		&
-		\hspace{-1em}
+		\hspace{-1.5em}
 		\begin{tabular}{c}
@@
 		Look at all those candles! \\
-		No wonder it was so hard. \\
-		\\
+		No wonder it was so hard \\
+		for the author to pass.\\
 		\end{tabular}
@@
 			(c) {\em Flames of the Sky}, $\mathcal{T}=1.04$. \\
-		False-positive with foot-\\
-		switches. Actually faces \\
-		$UR$/$U$/$UL$ the whole time.
+			False-positive $\mathcal{T}$ from foot-\\
+		switches. Actually faces $R$/$UR$/\\
+		$U$/$UL$/$L$ the whole time.
 		\end{tabular}
@@
 		Blazing & 15 & 0.696 \\
-		C. Foy (\smiley) & 138 & 0.678 \\
+		C. Foy & 138 & 0.678 \\
 	\end{tabular}

### 2016-03-13 itg/paper.tex
Add more about the least turny charts, and add future work.

@@
 %\usepackage{mathpartir}
+\usepackage{hyperref}
 \usepackage{url}
@@
 \usepackage[utopia]{mathdesign}
+\usepackage{natbib}
 
@@
 Among the many factors contributing to a stepchart's difficulty is how much the player must turn from side to side.
-	Other more obvious factors, such as raw speed, have been well studied in prior work. % TODO cite.
+	Other more obvious factors, such as raw speed, have been well studied in prior work.
 	This paper presents an analytic study of this {\em turniness} factor.
@@
 This paper's author plays ITG at the 13-14 level.}.
-To date, difficulty is always judged by a human (typically the step author), and primarily reflects factors such as burstiness (footspeed), prolognedness of streams (stamina), and complexity of jumps, holds, and crossovers (technical).
+To date, difficulty is always judged by a human (typically the step author), and primarily reflects factors such as burstiness (footspeed), prolongedness of streams (stamina), and complexity of jumps, holds, and crossovers (technical).
 These measures of difficulty have been well-studied in prior works \cite{callofthehound,dawgsinthehouse}.
@@
 Each of the 5 categories shares the same data point at $\mathcal{T}=0$: there are 7 possible streams that never turn at all.
-These are the 7 tower patterns $LULU$, $LRLR$, $LDLD$, $URUR$, $DRDR$, $UDUD$, and $DUDU$.
+These are the 7 ``tower'' patterns $LULU$, $LRLR$, $LDLD$, $URUR$, $DRDR$, $UDUD$, and $DUDU$.
 We already showed the first of these in Figure~\ref{fig:stream}, left.
@@
 
-{\bf The Least Turny Charts.}
-16 charts in our collection have $\mathcal{T}=0$.
-Most of these are beginner charts, rated 2 or 3 feet and with no more than 100 steps.
-$\Delta$MAX \cite{deltamax} stands out, as a 15-footer with 945 steps.
-God is that chart ever annoying.
-
-Ignoring all beginner-difficulty and turn-free charts, the least turny stepchart is {\em Sick Dubstep Track} \cite{sickdubstep}, a 4-measure joke rated 6,482 feet, the majority of which is shown in Figure~\ref{fig:example}(a).
-It has $\mathcal{T}=0.263$.
-
 {\bf The Turniest Charts.}
 No joker has yet dared to make the turniest possible chart (all spins), but one chart comes close.
-Standing alone as the only chart turnier than 1.5 is the Easy 3 for {\em DO ME} \cite{dome}, which is mostly spins punctuated by a few breaks, as shown in Figure~\ref{fig:example}(d).
+Standing alone as the only chart turnier than 1.5 is the Easy 3 for {\em DO ME} \cite{dome}, which is mostly spins punctuated by a few breaks, as shown in Figure~\ref{fig:example}(a).
 \cite{alanv} loves this chart.
@@
 
-Beyond these extreme examples, we are interested in some more normal charts people might actually like to play.
-Table~\ref{tab:feets} shows the turniest charts of each level, for the most popular difficulty levels ranging from 7 to 17.
-So, next time you want to work on your laterals or candles, you have me to thank!
-
-Surprisingly, {\em BRILLIANT 2U (Orchestra-Groove)} \cite{brilliant2u} is full of 270s. I didn't think anybody ever put those in real charts.
-Finally, an honourable mention goes to {\em Oedo Hop} \cite{oedo}, the 2nd turniest 7 (and 4th turniest chart overall!) with $\mathcal{T}=1.384$.
-\cite{deltax} really likes that one so we had to mention it.
-
 \begin{figure*}[t]
@@
 	\begin{tabular}{cccc}
-		\hspace{-2em}
+		\hspace{-1em}
 		\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{sick-dubstep-track.png} \\
+		\includegraphics[width=0.18\textwidth]{do-me.png} \\
 		\small
 		\begin{tabular}{l}
-			(a) {\em Sick Dubstep Track}, $\mathcal{T}=0.26$. \\
-		The least turny non-trivial \\
-		chart.\\
-		\\
+			(a) {\em DO ME}, $\mathcal{T}=1.63$. \\
+			Proportion of non-spinny \\
+			section enlarged to show \\
+			detail. \\
 		\end{tabular}
@@
 		&
-		\hspace{-1.5em}
+		\hspace{-1em}
 		\begin{tabular}{c}
@@
 		&
-		\hspace{-1.5em}
+		\hspace{-2em}
 		\begin{tabular}{c}
@@
 		&
-		\hspace{-1.5em}
+		\hspace{-2em}
 		\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{do-me.png} \\
+		\includegraphics[width=0.18\textwidth]{sick-dubstep-track.png} \\
 		\small
 		\begin{tabular}{l}
-			(d) {\em DO ME}, $\mathcal{T}=1.63$. \\
-		I'll pass, thanks.\\
-		\\
+			(d) {\em Sick Dubstep Track}, $\mathcal{T}=$ \\
+			$0.26$. 
+		The least turny non- \\trivial
+		non-beginner chart.\\
 		\\
@@
 		\hline
-		7 & Trip Machine & DDR 1st & 1.407 \\
+		7 & TRIP MACHINE & DDR 1st & 1.407 \\
 		%7 & Oedo Hop & r2112 & 1.38 \\
 		%  & (honorable mention; way better chart) & & \\
-		8 & Brilliant 2U (O-G) & DDR 2ndMIX & 1.347 \\
+		8 & BRILLIANT 2U (O-G) & DDR 2ndMIX & 1.347 \\
 		9 & Hearts of Ice & In The Groove 3 & 1.127 \\
@@
 		17 & Superluminal & Ft. Rapids VII & 1.058 \\
+		18 & {\em (no 18s+ with $\mathcal{T}\ge 1$)} & - & - \\
 	\end{tabular}
@@
 
-{\bf The Turniest Other Things.}
-Although our analysis picked out a good many turny charts, any ITG player will inevitably get tired of playing the same ones over and over.
-Hence, we are also interested in an aggregated turniness profile to show the average scores of many different stepcharts at once.
-We computed average turninesses for stepcharts grouped {\em by pack} and {\em by step artist}.
-Abbreviated results for the former are shown in Table~\ref{tab:pack} and for the latter in Table~\ref{tab:artist}.
-For the sake of players who dislike turniness, we also list the lowest-scoring packs and artists in addition to the highest-scoring ones.
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c}
+		Ft. & Name & Pack & $\mathcal{T}$ \\
+		\hline
+		7  & King Kong & In The Groove 3 & 0.354 \\
+		8  & Jam Jam Reggae & DDR 2ndMIX & 0.428 \\
+		9  & Fantasia & In The Groove 3 & 0.382 \\
+		10 & Oosanbashi & TLOES Chapter 1 & 0.534 \\
+		11 & Release the Music & Ft. Rapids VII & 0.507 \\
+		12 & Payon (Theme Of) & CuoReNeRo MeGaPacK & 0.396 \\
+		13 & Technodildo & The Legend of Zim 5 & 0.436 \\
+		14 & Gentleman & TLOES Chapter 1 & 0.556 \\
+		15 & $\Delta$MAX & Tachyon Alpha & 0.000 \\
+		16 & Dash Hopes 2 & Tachyon Alpha & 0.327 \\
+		17 & Resist & SPEEDCOOOORE & 0.592\\
+		18 & Baby & Sexuality Violation 2 & 0.524 \\
+		19 & Pedal to the Metal & SPEEDCOOOORE & 0.468 \\
+		%6482 & Sick Dubstep Track & Mute Sims 7 & 0.263 \\
+	\end{tabular}
+	\end{center}
+	\caption{The least turny existing stepcharts of each difficulty (trivial charts excluded).}
+	\label{tab:feets2}
+\end{table}
 
@@
 \end{table}
+Beyond these extreme examples, we are interested in some more normal charts people might actually like to play.
+Table~\ref{tab:feets} shows the turniest charts of each level, for the most popular difficulty levels ranging from 7 to 17.
+So, next time you want to work on your laterals or candles, you have me to thank!
+
+Surprisingly, {\em BRILLIANT 2U (Orchestra-Groove)} \cite{brilliant2u} is full of 270s. I didn't think anybody ever put those in real charts.
+Finally, an honourable mention goes to {\em Oedo Hop} \cite{oedo}, the 2nd turniest 7 (and 4th turniest chart overall!) with $\mathcal{T}=1.384$.
+\cite{deltax} really likes that one so we had to mention it.
+
+{\bf The Least Turny Charts.}
+16 charts in our collection have $\mathcal{T}=0$.
+Most of these are beginner charts, rated 2 or 3 feet and with no more than 100 steps.
+$\Delta$MAX \cite{deltamax} stands out, as a 15-footer with 945 steps.
+God is that chart ever annoying.
+
+Ignoring all beginner-difficulty and turn-free charts, the least turny stepchart is {\em Sick Dubstep Track} \cite{sickdubstep}, a 4-measure joke rated 6,482 feet, the majority of which is shown in Figure~\ref{fig:example}(d).
+It has $\mathcal{T}=0.263$.
+
+For readers who prefer more to tower than to turn, we similarly tabulated the least turny charts of each common difficulty in Table~\ref{tab:feets2}.
+
+{\bf The Turniest Other Things.}
+Although our analysis picked out a good many turny charts, any ITG player will inevitably get tired of playing the same ones over and over.
+Hence, we are also interested in an aggregated turniness profile to show the average scores of many different stepcharts at once.
+We computed average turninesses for stepcharts grouped {\em by pack} and {\em by step artist}.
+Abbreviated results for the former are shown in Table~\ref{tab:pack} and for the latter in Table~\ref{tab:artist}.
+For the sake of players who dislike turniness, we also list the lowest-scoring packs and artists in addition to the highest-scoring ones.
+
+\section{Future Work}
+
+\defcitealias{ode}{Anonymized 2016}
+The most obvious next step for this work is to account for jumps.
+This seems quite easy: compute the two possible facings of the jump, depending which foot lands on which arrow, and use whichever facing produces the least turniness when compared to the previous and next steps in the sequence.
+Honestly, we should have implemented that for this paper, but we only realized how it should work after we finished computing our dataset;
+and owing to an inferior conference's deadline coming up in 10 days \citepalias{ode}, we don't have time for re-running experiments \citepalias{ode}.
+
+More difficult is the matter of deciding whether stepping the same arrow twice in a row constitutes a jack (same foot) or a footswitch.
+As we saw in Figure~\ref{fig:example}, our current approach of assuming it's always a jack can lead to some false-positive $\mathcal{T}$ values.
+Judging the presence of footswitches is a nontrivial problem even for human intuition, so we view this as an exciting area for future research.
+So as not to piss off the player, most stepcharts tend to feature only jacks or only switches, never both.
+Future extensions of our algorithm could exploit this convention to decide with high confidence whether or not an entire chart is footswitchy.
 
@@
 
+Our analysis program can be found at \url{https://github.com/bblum/sigbovik/tree/master/itg/code}, and our full experimental data-set is available at \url{http://tinyurl.com/turniness}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \appendix
-
-\section{Linx}
-
-Our analysis program can be found at \url{https://github.com/bblum/sigbovik/tree/master/itg/code}, and our full experimental data-set is available at \url{http://tinyurl.com/turniness}.
 

### 2016-03-13 itg/paper.tex
write more about future work

@@
 \defcitealias{ode}{Anonymized 2016}
-The most obvious next step for this work is to account for jumps.
+{\bf Jumps.} The most obvious next step for this work is to account for jumps.
 This seems quite easy: compute the two possible facings of the jump, depending which foot lands on which arrow, and use whichever facing produces the least turniness when compared to the previous and next steps in the sequence.
@@
 
-More difficult is the matter of deciding whether stepping the same arrow twice in a row constitutes a jack (same foot) or a footswitch.
+{\bf Footswitches.} More difficult is the matter of deciding whether stepping the same arrow twice in a row constitutes a jack (same foot) or a footswitch.
 As we saw in Figure~\ref{fig:example}, our current approach of assuming it's always a jack can lead to some false-positive $\mathcal{T}$ values.
@@
 
+{\bf Fractional Turns.}
+Experienced ITG players do not step with their feet on the middle of each arrow; rather, they try to step on the edges and ``brackets'' between arrows, hitting each step with heel or toes only to minimize the distance their feet must move.
+This tendency distinguishes further granularity of turning than what we identify in this work.
+Consider the two streams shown in Figure~\ref{fig:future} (a) and (b).
+Although they measure the same $\mathcal{T}$ value of 1337, % TODO
+(b) is considerably more exhausting to step, as (a) allows the player to bracket the entire run.
+Future work could handle this by counting a 1-turn step as the full value only if the player subsequently continues turning farther in the same direction (which would force the player's feet off the brackets), and otherwise assign $\mathcal{T}=1/2$ for such a step.
+
+\begin{figure}[t]
+	% TODO
+	\caption{Future work.}
+	\label{fig:future}
+\end{figure}
+
+{\bf Holds and Rolls.}
+Many charts use holds or rolls to force the player to step all concurrent tap-steps with the other foot, which breaks our alternating-feet assumption.
+When multiple steps are concurrent with a hold, they must be double-stepped, and the facing for each computed using the hold's arrow rather than the most recent one.
+When multiple steps are concurrent with a roll, the facings are determined similarly, but repeatedly stepping the roll results in a stream, as shown in Figure~\ref{fig:future}(c) \cite{botenanna}.
+%Roll-streams have comparatively more steps, which will affect the turniness.
+Judging how hold-steps and roll-streams must be stepped is another exciting opportunity for future work.
+
 \section{Conclusion}
 
-In conclusion, $LURR$ $DURR$!
+Our analysis program can be found at \url{https://github.com/bblum/sigbovik/tree/master/itg/code}, and our full experimental dataset is available at \url{http://tinyurl.com/turniness}.
 
-Our analysis program can be found at \url{https://github.com/bblum/sigbovik/tree/master/itg/code}, and our full experimental data-set is available at \url{http://tinyurl.com/turniness}.
+The value of this work is already abundantly clear, as through it, the author has found several new charts he is interested to play.
+We hope our published dataset can provide similar value to other ITG players worldwide.
+Please accept our paper.
+We worked hard on it.
 

### 2016-03-13 itg/paper.tex
finish future work

@@
 Consider the two streams shown in Figure~\ref{fig:future} (a) and (b).
-Although they measure the same $\mathcal{T}$ value of 1337, % TODO
+Although they measure the same $\mathcal{T}$ value of $1$,
 (b) is considerably more exhausting to step, as (a) allows the player to bracket the entire run.
@@
 \begin{figure}[t]
-	% TODO
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{future-a.png} \\
+	\small (a) Bracketable.
+	\end{tabular}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{future-b.png} \\
+	\small (b) Forced turns.
+	\end{tabular}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{boten.png} \\
+	\small (c) Roll-stream.
+	\end{tabular}
 	\caption{Future work.}

### 2016-03-14 itg/paper.tex
fix some stuff

@@
 Each of the 5 categories shares the same data point at $\mathcal{T}=0$: there are 7 possible streams that never turn at all.
-These are the 7 ``tower'' patterns $LULU$, $LRLR$, $LDLD$, $URUR$, $DRDR$, $UDUD$, and $DUDU$.
+These are the 7 ``tower'' patterns $LRLR$, $LULU$, $LDLD$, $URUR$, $DRDR$, $UDUD$, and $DUDU$.
 We already showed the first of these in Figure~\ref{fig:stream}, left.
@@
 		\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{do-me.png} \\
+		\includegraphics[width=0.169\textwidth]{do-me.png} \\
 		\small
@@
 		\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{mr-saxobeat.png} \\
+		\includegraphics[width=0.169\textwidth]{mr-saxobeat.png} \\
 		\small
@@
 		No wonder it was so hard \\
-		for the author to pass.\\
+		for us to pass this one.\\
 		\end{tabular}
@@
 		\begin{tabular}{c}
-		\includegraphics[width=0.185\textwidth]{flames.png} \\
+		\includegraphics[width=0.169\textwidth]{flames.png} \\
 		\small
@@
 		\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{sick-dubstep-track.png} \\
+		\includegraphics[width=0.169\textwidth]{sick-dubstep-track.png} \\
 		\small

### 2016-03-14 itg/paper.tex
Fix step author turniness table; add related work section; fix other stuff

@@
 We processed every stepchart in our collection, which includes 70+ song packs.
-Most well-known pack series from the ITG community are included, such as {\em Cirque}, {\em ITG}, {\em Mute Sims}, {\em Pendulum}, {\em r21*}, {\em SPEEDCOOOORE}, {\em The Legend of Zim}, and {\em Valex's Magical 4-Arrow Adventure}.
+Most well-known pack series from the ITG community are included, such as {\em Cirque}, {\em ITG}, {\em Mute Sims}, {\em Pendulum}, {\em r21*}, {\em SPEEDCOOOORE}, {\em The Legend of Zim}, {\em Valex's Magical 4-Arrow Adventure}, and {\em Oh Henry!
+Mad Stamina: Streamy $\cdot$ Brutal $\cdot$ Stamina Bar in Milk Chocolate}.
 
@@
 		StepMania 5 & 5 & 0.869 \\
+		Cirque du Zeppelin & 109 & 0.867 \\
+		In The Groove 3 & 448 & 0.858 \\
 		\multicolumn{3}{c}{\normalsize\dots} \\
+		Piece of Cake 5 & 20 & 0.785 \\
+		Cirque du Veyron & 30 & 0.777 \\
 		Stephcharts and Richarts & 81 & 0.770 \\
@@
 		\hline
-		M. Emirzian & 25 & 0.981 \\
-		sssmsm & 40 & 0.980 \\
-		M. Calfin & 16 & 0.947 \\
-		J. Berkowitz (Valex) & 1016 & 0.940 \\
+		M. Emirzian	& 38	& 0.992 \\
+		sssmsm		& 50	& 0.945 \\
+		J.Berkowitz (Valex)	& 1025	& 0.939 \\
 		B. Dinh		& 10	& 0.919 \\
-		J. DeGarmo & 24 & 0.915 \\
-		X. Ythar & 17 & 0.914 \\
-		R. McKanna & 59 & 0.895 \\
-		R. Woods & 18 & 0.892 \\
+		J. DeGarmo	& 26	& 0.906 \\
+		J. Frederick	& 176	& 0.881 \\
 		Fraxtil		& 23	& 0.880 \\
+		M. Calfin	& 26	& 0.880 \\
+		X. Ythar	& 24	& 0.879 \\
+		R. Woods	& 22	& 0.874 \\
+		R. McKanna	& 76	& 0.872 \\
+		MIDI		& 86	& 0.870 \\
+		Dark Luke	& 36	& 0.866 \\
+		@@		& 60	& 0.859 \\
+		D. Bernardone	& 354	& 0.858 \\
+		D. D'Amato	& 154	& 0.857 \\
 		\multicolumn{3}{c}{\normalsize\dots} \\
-		I. Pyles & 175 & 0.754 \\
-		P. Shanklin & 29 & 0.752 \\
-		Gazebo & 355 & 0.751 \\
-		zimlord & 199 & 0.750 \\
-		C. Emirzian & 17 & 0.736 \\
-		Mootz & 22 & 0.734 \\
-		B. Abear & 27 & 0.729 \\
-		Insane Steve & 41 & 0.718 \\
-		Blazing & 15 & 0.696 \\
-		C. Foy & 138 & 0.678 \\
+		%Konami	& 817	& 0.854 \\
+		%NEMORIGINAL	& 63	& 0.85 \\
+		%D. Renzetti	& 26	& 0.846 \\
+		%D. Runberg	& 66	& 0.844 \\
+		%Zaia	& 322	& 0.842 \\
+		%Blank	& 272	& 0.841 \\
+		%Driving55	& 21	& 0.833 \\
+		%JOE2	& 16	& 0.83 \\
+		%Wastum	& 19	& 0.83 \\
+		%mute	& 410	& 0.828 \\
+		%MacGravel	& 55	& 0.827 \\
+		%T. Swag	& 13	& 0.826 \\
+		%NIQ9	& 29	& 0.825 \\
+		%C. Emirzian	& 27	& 0.823 \\
+		%t0ni	& 75	& 0.821 \\
+		%R. Uchima	& 122	& 0.819 \\
+		%BlueChaos	& 104	& 0.818 \\
+		%Brendino	& 14	& 0.816 \\
+		%T. Moore	& 66	& 0.813 \\
+		%Aoreo	& 216	& 0.813 \\
+		%mudkyp	& 35	& 0.811 \\
+		%King of Light	& 39	& 0.809 \\
+		%G. Marchan	& 18	& 0.806 \\
+		%Mad Matt	& 383	& 0.806 \\
+		%Arvin	& 64	& 0.806 \\
+		%Sudzi781	& 99	& 0.803 \\
+		%S. Venkat	& 82	& 0.803 \\
+		%Kindan	& 17	& 0.801 \\
+		%Renard	& 53	& 0.799 \\
+		%J. Nero	& 194	& 0.796 \\
+		%W. Hodge	& 51	& 0.795 \\
+		%Jayce	& 43	& 0.79 \\
+		%WinDEU	& 49	& 0.787 \\
+		M. Simmons	& 334	& 0.787 \\
+		Revolver	& 25	& 0.775 \\
+		R. Konkul (Rynker)	& 125	& 0.774 \\
+		K. Ward	(\smiley) & 600	& 0.773 \\
+		Freyja		& 89	& 0.766 \\
+		P. Mason	& 14	& 0.765 \\
+		M. Puls		& 167	& 0.759 \\
+		P. Shanklin	& 37	& 0.756 \\
+		I. Pyles	& 177	& 0.754 \\
+		Gazebo		& 377	& 0.754 \\
+		zimlord		& 217	& 0.747 \\
+		C. Foy		& 205	& 0.739 \\
+		B. Abear	& 29	& 0.733 \\
+		Blazing		& 21	& 0.729 \\
+		Mootz		& 29	& 0.715 \\
+		Insane Steve	& 49	& 0.714 \\
 	\end{tabular}
@@
 Honestly, we should have implemented that for this paper, but we only realized how it should work after we finished computing our dataset;
-and owing to an inferior conference's deadline coming up in 10 days \citepalias{ode}, we don't have time for re-running experiments \citepalias{ode}.
+and owing to an inferior conference's deadline coming up in 10 days, we don't have time for re-running experiments \citepalias{ode}.
 
@@
 (b) is considerably more exhausting to step, as (a) allows the player to bracket the entire run.
-Future work could handle this by counting a 1-turn step as the full value only if the player subsequently continues turning farther in the same direction (which would force the player's feet off the brackets), and otherwise assign $\mathcal{T}=1/2$ for such a step.
+Future work could handle this by counting a 1-turn step as the full value only if the player subsequently continues turning farther in the same direction (which would force the player's feet off the brackets), and otherwise assign $0<\mathcal{T}<1$ for such a step.
 
@@
 
-Following are the most turny 18-step patterns that are totally vanilla, with turniness of 1.375. (Manual inspection reveals the true turniness to be 4/3.)
+Following are the most turny 18-step patterns that are totally vanilla, with $\mathcal{T}=1.375$. (Manual inspection reveals the true turniness to be 4/3.)
 
@@
 
-Following are the most turny 18-step patterns that may include crossovers only, with turniness of 1.5.
+Following are the most turny 18-step patterns that may include crossovers only, with $\mathcal{T}=1.5$.
 
@@
 
-Following are the most turny 18-step patterns that may include crossovers and laterals, with turniness of 1.625. (Manual inspection reveals the true turniness to be 8/5.)
+Following are the most turny 18-step patterns that may include crossovers and laterals, with $\mathcal{T}=1.625$. (Manual inspection reveals the true turniness to be 8/5.)
 
@@
 
-Following are the most turny 18-step patterns that may include crossovers, laterals, and 270s, with turniness of 1.625.
+Following are the most turny 18-step patterns that may include crossovers, laterals, and 270s, with $\mathcal{T}=1.625$.
 (Manual inspection reveals irregularity; these streams exhibit no core pattern that can be repeated to retain the maximum turniness.)
@@
 
-Following are the most turny 18-step patterns that may include any pattern, with turniness of 2.
+Following are the most turny 18-step patterns that may include any pattern, with $\mathcal{T}=2$.
 
@@
 
+\section{Related Work...?}
+
+\begin{figure}[h]
+	\begin{center}
+	\includegraphics[width=0.48\textwidth]{machine-translation.jpg}
+	\end{center}
+	\caption{We employ the techniques of \cite{deltax} for ITG machine translation.} %All photos credit \cite{alanv}.
+	\label{fig:machine-translation}
+\end{figure}
 \bibliographystyle{abbrvnat}

### 2016-03-15 itg/paper.tex
Add some important new citations
team dragonforce

@@
 Stepcharts are rated in difficulty on a scale from 1 to 20 ``feet'' \cite{simplylove}\footnote{
-Other dance games \cite{konami}, which shall remain unnamed, feature wimpier stepcharts rated only up to 10, but as ITG players got better, they left those charts in the dust and began creating their own charts to challenge themselves with ridiculous difficulties.
+Other dance games, which shall remain unnamed, but whose initials are \cite{konami}, feature wimpier stepcharts rated only up to 10.
+As ITG players got better, they left those charts in the dust and began creating their own charts to challenge themselves with ridiculous difficulties.
 This paper's author plays ITG at the 13-14 level.}.
-To date, difficulty is always judged by a human (typically the step author), and primarily reflects factors such as burstiness (footspeed), prolongedness of streams (stamina), and complexity of jumps, holds, and crossovers (technical).
-These measures of difficulty have been well-studied in prior works \cite{callofthehound,dawgsinthehouse}.
+To date, difficulty is judged by a human (typically the step author), and primarily reflects factors such as burstiness ({\em footspeed}), prolongedness of streams ({\em stamina}), and complexity of jumps, holds, and crossovers ({\em technical}).
+These measures of difficulty have been well-studied in prior works \cite{teamdragonforce,callofthehound,dawgsinthehouse}.
 
@@
 We processed every stepchart in our collection, which includes 70+ song packs.
-Most well-known pack series from the ITG community are included, such as {\em Cirque}, {\em ITG}, {\em Mute Sims}, {\em Pendulum}, {\em r21*}, {\em SPEEDCOOOORE}, {\em The Legend of Zim}, {\em Valex's Magical 4-Arrow Adventure}, and {\em Oh Henry!
+Most well-known pack series from the ITG community are included, such as {\em Cirque}, {\em ITG}, {\em Mute Sims}, {\em Pendulum}, {\em r21*}, {\em SPEEDCOOOORE}, {\em Tachyon}, {\em The Legend of Zim}, {\em Valex's Magical 4-Arrow Adventure}, and {\em Oh Henry!
 Mad Stamina: Streamy $\cdot$ Brutal $\cdot$ Stamina Bar in Milk Chocolate}.
@@
 	\end{center}
-	\caption{We employ the techniques of \cite{deltax} for ITG machine translation.} %All photos credit \cite{alanv}.
+	\caption{We employ the techniques of \cite{deltax} for ITG machine translation \cite{itg-translation1,itg-translation2}.} %All photos credit \cite{alanv}.
 	\label{fig:machine-translation}

### 2017-03-11 itg/code/ITG.hs
Add xover/switch/jack analysis for 2017

@@ -1,0 +1,85 @@
+import Data.List
+import Data.Char
+import Data.Maybe
+
+data Step = L | D | U | R | Jump deriving Eq
+
+data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int, jacks :: Int,
+                         lastStep :: Maybe Step, doubleStep :: Bool, lastFlip :: Bool,
+                         lastFoot :: Bool, stepsL :: [Bool], stepsR :: [Bool] }
+
+commitStream :: AnalysisState -> AnalysisState
+commitStream s = s { xovers   = xovers   s + if f then ns - nx else nx,
+                     switches = switches s + fromEnum (f == lastFlip s && doubleStep s),
+                     jacks    = jacks    s + fromEnum (f /= lastFlip s && doubleStep s),
+                     lastFlip = f, stepsL = [], stepsR = [] }
+    where ns = length $              stepsL s ++ stepsR s
+          nx = length $ filter not $ stepsL s ++ stepsR s
+          -- if more than half the L/R steps in this stream were crossed over,
+          -- then we got the footing backwards and need to flip the stream.
+          -- as a tiebreaker, flip if the stream is already more jacky than
+          -- footswitchy, i.e., if past streams flipped more often than not.
+          f = nx * 2 > ns || nx * 2 == ns && ((jacks s > switches s) /= lastFlip s)
+
+analyzeStep :: AnalysisState -> Step -> AnalysisState
+analyzeStep s step
+    -- a jump resets the footing, so the next step can be stepped with either
+    -- foot. commit the stream so far to treat it separately from what follows.
+    -- bracket-jumps are, of course, future work.
+    | step == Jump = (commitStream s) { lastStep = Nothing, doubleStep = False }
+    -- two steps on the same arrow might be a jack, or might be a footswitch.
+    -- to figure out which, commit the stream so far, and begin a new stream
+    -- whose footing will retroactively determine how to foot this step.
+    -- also, unlike jumps, this step gets counted as part of the next stream.
+    | lastStep s == Just step = stream (commitStream s) { doubleStep = True }
+    -- a normal streamy step.
+    | otherwise = stream s
+    where foot = not $ lastFoot s
+          -- record whether we stepped on a matching or opposite L/R arrow here
+          addStep ft L steps = ft:steps
+          addStep ft R steps = (not ft):steps
[…]

### 2017-03-11 itg/code/run-tests.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+for t in test*.sm; do
+	ans=`head -n 1 "$t"`
+	if ! echo "$ans" | grep ",.*,.*," >/dev/null; then
+		echo -e "\033[01;33m$t not well formed -- $ans not an answer.\033[00m"
+		continue
+	fi
+	out=`cat "$t" | ./ITG`
+	if echo "$out" | grep "$ans" >/dev/null; then
+		echo -e "\033[01;32m[ OK ]\033[00m $t"
+	else
+		echo -e "\033[01;31m[FAIL] $t -- expected: $ans; got: $out\033[00m"
+	fi
+done

### 2017-03-11 itg/2/paper.tex
begin itg2 paper

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

%\documentclass[preprint]{sigplanconf}
\documentclass[10pt]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}

% ____________________________________________________________
% Listings Package Configuration
% \usepackage[scaled]{beramono}

%\renewcommand*\ttdefault{txtt}
\usepackage[T1]{fontenc}

% This Deep Tex Voodoo is from
%   <http://www.latex-community.org/forum/viewtopic.php?f=5&t=2072>
% It's purpose is to make \lstinline normal size, without affecting
% \lstinputlisting.  It seems to work but I have no idea how or why,
% and I rather hope never to learn.
%\makeatletter
%\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
%\makeatother

\begin{document}

\conferenceinfo{SIGBOVIK '17}{Pittsburgh, PA, USA}
\copyrightyear{2017}
\copyrightdata{}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{
A Boring Follow-Up Paper to \\
``Which ITG Stepcharts are Turniest?'' \\
Titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?'' \\
%That `Boring' Stuff Was Part of the Title, BTW. \\
%So was that. And that, and this too. \\
%You got it all, right? \\
%Or Just, ``More Boring Crap about ITG'', for Short. \\
%Oh, That Was Also Part of the Title.
}
% \subtitle{\em The Randomly-Scoped Lambda Calculus}
% \subtitle{Subtitle Text, if any}

\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}

\maketitle

\begin{abstract}
	%ITG is a popular dance game in which players step on arrows while listening to music. The arrow patterns, indicated by a {\em stepchart}, may range among any level of complexity and difficulty. Among the many factors contributing to a stepchart's difficulty is how much the player must turn from side to side.
	%Other more obvious factors, such as raw speed, have been well studied in prior work.
	%This paper presents an analytic study of this {\em turniness} factor.
	%We study the turniness of many existing stepcharts, and present a novel (but unsurprising) approach to automatically generating maximally (or minimally) turny charts.
	%Among real-world songs, we find stepcharts with overall turniness ranging from 0\% to 81.33\% of the theoretical maximum.

\end{abstract}

\category{D.D.R.}{Exercise and Fitness}{Arcade Dance Games}

\keywords
in, the, groove

\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Flashback Scene}

\begin{table}[h]
	\begin{center}
	\begin{tabular}{cc|cccc}
		& & \multicolumn{4}{c}{Right foot} \\
		& & $\leftarrow$ & $\downarrow$ & $\uparrow$ & $\rightarrow$ \\
		\hline
		\multirow{4}{*}{Left foot}
		%                 L   D   U   R
		& $\leftarrow$  & - & $UR$ & $UL$ & $U$ \\
		& $\downarrow$  & $DL$* & - & $L$ & $UL$ \\
		& $\uparrow$    & $DR$* & $R$ & - & $UR$ \\
		& $\rightarrow$ & $D\dagger$ & $DR$* & $DL$* & - \\

	\end{tabular}
	\end{center}
	\caption{Facing directions. ``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
	\label{tab:facing}
\end{table}

% dont forget
% http://dancedancerevolutionddr.wikia.com/wiki/Afronova_walk

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{pink}{RGB}{128,0,192}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,32}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,128,192}

\begin{figure*}[t]
\begin{center}
\begin{tabular}{l}
	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step}~= \hilight{brickred}{L}~| \hilight{brickred}{D}~| \hilight{brickred}{U}~| \hilight{brickred}{R}~| \hilight{brickred}{Jump}~\hilight{orange}{deriving}~\hilight{olivegreen}{Eq}} \\
\texttt{} \\
\texttt{\hilight{orange}{data}~\hilight{olivegreen}{AnalysisState}~= \hilight{brickred}{S}~\{ steps :: \hilight{olivegreen}{Int}, xovers :: \hilight{olivegreen}{Int}, switches :: \hilight{olivegreen}{Int}, jacks :: \hilight{olivegreen}{Int},} \\
\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastStep :: \hilight{olivegreen}{Maybe}~\hilight{olivegreen}{Step}, doubleStep :: \hilight{olivegreen}{Bool}, lastFlip :: \hilight{olivegreen}{Bool},} \\
\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsL :: [\hilight{olivegreen}{Bool}], stepsR :: [\hilight{olivegreen}{Bool}] \}} \\
\texttt{} \\
\texttt{\hilight{pink}{commitStream}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{AnalysisState}} \\
\texttt{\hilight{pink}{commitStream}~s = s \{ xovers~~ = xovers~~ s + \hilight{orange}{if}~f \hilight{orange}{then}~ns - nx \hilight{orange}{else}~nx, } \\
\texttt{~~~~~~~~~~~~~~~~~~~~ switches = switches s + \hilight{orange}{fromEnum}~(f == lastFlip s \&\& doubleStep s), } \\
\texttt{~~~~~~~~~~~~~~~~~~~~ jacks~~~~= jacks~~~~s + \hilight{orange}{fromEnum}~(f /= lastFlip s \&\& doubleStep s), } \\
	\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsL = \hilight{brickred}{[]}, stepsR = \hilight{brickred}{[]}~\}} \\
\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~~~~~~~~~~~~~~stepsL s ++ stepsR s} \\
\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsL s ++ stepsR s} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the stream is already more jacky than}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((jacks s > switches s) /= lastFlip s)} \\
\texttt{} \\
\texttt{\hilight{pink}{analyzeStep}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{Step}~-> \hilight{olivegreen}{AnalysisState}} \\
\texttt{\hilight{pink}{analyzeStep}~s step} \\
\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\
\texttt{~~~~| step == \hilight{brickred}{Jump}~= (\hilight{pink}{commitStream}~s) \{ lastStep = \hilight{brickred}{Nothing}, doubleStep = \hilight{brickred}{False}~\}} \\
\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\
\texttt{~~~~| lastStep s == \hilight{brickred}{Just}~step = stream (\hilight{pink}{commitStream}~s) \{ doubleStep = \hilight{brickred}{True}~\}} \\
\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\
	\texttt{~~~~| \hilight{orange}{otherwise}~= stream s} \\
\texttt{~~~~\hilight{orange}{where}~foot = \hilight{orange}{not}~\$ lastFoot s} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or opposite L/R arrow here}} \\
\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = ft:steps} \\
\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = (\hilight{orange}{not}~ft):steps} \\
\texttt{~~~~~~~~~~addStep ft \_ steps = steps \hilight{commentblue}{-{}- U/D don't help to determine L/R footing.}} \\
\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- add the (matching? xover?) step to the list corresponding to this foot.}} \\
\texttt{~~~~~~~~~~newLs s = \hilight{orange}{if}~foot \hilight{orange}{then}~addStep foot step \$ stepsL s \hilight{orange}{else}~stepsL s} \\
\texttt{~~~~~~~~~~newRs s = \hilight{orange}{if}~foot \hilight{orange}{then}~stepsR s \hilight{orange}{else}~addStep foot step \$ stepsR s} \\
\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step,} \\
\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot = foot, stepsL = newLs s, stepsR = newRs s \} } \\
\texttt{} \\
\texttt{\hilight{pink}{analyze}~:: [\hilight{olivegreen}{Step}] -> \hilight{olivegreen}{AnalysisState}} \\
	\texttt{\hilight{pink}{analyze}~= \hilight{pink}{commitStream}~. \hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[] []}) } \\
\end{tabular}
\end{center}
\caption{it's pseudocode lol}
\label{fig:pseudocode-lol}
\end{figure*}

\section{Existing Chart Analysis}

\section{Conclusion}

Please accept our paper.
We worked hard on it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrvnat}
\bibliography{citations}

\end{document}

### 2017-03-12 itg/code/ITG.hs
itg format w tab instead of , (many song titles have ,)

@@ -56,9 +56,9 @@ stepify = step . map fst . filter snd . zip [L,D,U,R] . map (flip elem "124")
           step [s] = Just s
           step ss = Just Jump
 
--- handles chart metadata and formats the analysis result into CSV
+-- handles chart metadata and formats the analysis result into tsv
 process :: (String, [String]) -> String
-process (title,(name:diff:feet:_:rest)) = intercalate "," $ metadata ++ result
+process (title,(name:diff:feet:_:rest)) = intercalate "\t" $ metadata ++ result
     where -- chop off the trailing ":"
           trim = reverse . tail . dropWhile isSpace . reverse . dropWhile isSpace
           metadata = map trim [title,name,diff,feet]

### 2017-03-12 itg/code/run-tests.sh

@@ -1,15 +1,18 @@
 #!/bin/bash
 
+rv=0
 for t in test*.sm; do
 	ans=`head -n 1 "$t"`
 	if ! echo "$ans" | grep ",.*,.*," >/dev/null; then
 		echo -e "\033[01;33m$t not well formed -- $ans not an answer.\033[00m"
 		continue
 	fi
-	out=`cat "$t" | ./ITG`
+	out=`cat "$t" | ./ITG | sed 's/\t/,/g'`
 	if echo "$out" | grep "$ans" >/dev/null; then
 		echo -e "\033[01;32m[ OK ]\033[00m $t"
 	else
 		echo -e "\033[01;31m[FAIL] $t -- expected: $ans; got: $out\033[00m"
+		rv=1
 	fi
 done
+exit $rv

### 2017-03-12 itg/code/ITG.hs
itg stepsL/stepsR -> stepsLR

@@ -6,15 +6,15 @@ data Step = L | D | U | R | Jump deriving Eq
 
 data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int, jacks :: Int,
                          lastStep :: Maybe Step, doubleStep :: Bool, lastFlip :: Bool,
-                         lastFoot :: Bool, stepsL :: [Bool], stepsR :: [Bool] }
+                         lastFoot :: Bool, stepsLR :: [Bool] }
 
 commitStream :: AnalysisState -> AnalysisState
 commitStream s = s { xovers   = xovers   s + if f then ns - nx else nx,
                      switches = switches s + fromEnum (f == lastFlip s && doubleStep s),
                      jacks    = jacks    s + fromEnum (f /= lastFlip s && doubleStep s),
-                     lastFlip = f, stepsL = [], stepsR = [] }
-    where ns = length $              stepsL s ++ stepsR s
-          nx = length $ filter not $ stepsL s ++ stepsR s
+                     lastFlip = f, stepsLR = [] }
+    where ns = length $ stepsLR s
+          nx = length $ filter not $ stepsLR s
           -- if more than half the L/R steps in this stream were crossed over,
           -- then we got the footing backwards and need to flip the stream.
           -- as a tiebreaker, flip if the stream is already more jacky than
@@ -35,18 +35,15 @@ analyzeStep s step
     -- a normal streamy step.
     | otherwise = stream s
     where foot = not $ lastFoot s
-          -- record whether we stepped on a matching or opposite L/R arrow here
+          -- record whether we stepped on a matching or crossed-over L/R arrow.
           addStep ft L steps = ft:steps
           addStep ft R steps = (not ft):steps
           addStep ft _ steps = steps -- U/D don't help to determine L/R footing.
-          -- add the (matching? xover?) step to the list corresponding to this foot.
-          newLs s = if foot then addStep foot step $ stepsL s else stepsL s
-          newRs s = if foot then stepsR s else addStep foot step $ stepsR s
-          stream s = s { steps = steps s + 1, lastStep = Just step,
-                         lastFoot = foot, stepsL = newLs s, stepsR = newRs s }
+          stream s = s { steps = steps s + 1, lastStep = Just step, lastFoot = foot,
+                         stepsLR = addStep foot step $ stepsLR s }
 
 analyze :: [Step] -> AnalysisState
[…]

### 2017-03-12 itg/code/ITG.hs
fix title stripping bug

@@ -66,9 +66,7 @@ process (title,(name:diff:feet:_:rest)) = intercalate "\t" $ metadata ++ result
 charts :: String -> [String] -> [(String, [String])]
 charts t [] = []
 charts t xs = chart ++ charts title rest
-    where stripTitle t | isPrefixOf "#TITLE:" t = reverse $ tail $ reverse $ drop 7 t
-                       | otherwise = stripTitle $ tail t
-          title = fromMaybe t $ stripTitle <$> find (isInfixOf "#TITLE:") xs
+    where title = fromMaybe t $ drop 7 <$> find (isPrefixOf "#TITLE:") xs
           -- get everything after the first "#NOTES"
           notnotes = not . isInfixOf "#NOTES:"
           stuff = tail $ dropWhile notnotes xs

### 2017-03-12 itg/code/get-chart.sh
performance opt to bad shell script

@@ -55,6 +55,6 @@ function process() {
 
 for DIFF in Beginner Easy Medium Hard Challenge Edit; do
 	if grep "$DIFF:" "$FILE" >/dev/null; then
-		cat "$FILE" | process
+		grep -A1000000000 "$DIFF:" "$FILE" | process
 	fi
 done

### 2017-03-12 itg/code/ITG.hs
Implement stream-splitting heuristics to avoid doublesteps

@@ -5,14 +5,11 @@ import Data.Maybe
 data Step = L | D | U | R | Jump deriving Eq
 
 data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int, jacks :: Int,
-                         lastStep :: Maybe Step, doubleStep :: Bool, lastFlip :: Bool,
-                         lastFoot :: Bool, stepsLR :: [Bool] }
+                         doubles :: Int, lastStep :: Maybe Step, lastJack :: Bool,
+                         lastFlip :: Bool, lastFoot :: Bool, stepsLR :: [Bool] }
 
 commitStream :: AnalysisState -> AnalysisState
-commitStream s = s { xovers   = xovers   s + if f then ns - nx else nx,
-                     switches = switches s + fromEnum (f == lastFlip s && doubleStep s),
-                     jacks    = jacks    s + fromEnum (f /= lastFlip s && doubleStep s),
-                     lastFlip = f, stepsLR = [] }
+commitStream s = maybe s0 splitStream splitIndex
     where ns = length $ stepsLR s
           nx = length $ filter not $ stepsLR s
           -- if more than half the L/R steps in this stream were crossed over,
@@ -20,30 +17,46 @@ commitStream s = s { xovers   = xovers   s + if f then ns - nx else nx,
           -- as a tiebreaker, flip if the stream is already more jacky than
           -- footswitchy, i.e., if past streams flipped more often than not.
           f = nx * 2 > ns || nx * 2 == ns && ((jacks s > switches s) /= lastFlip s)
+          -- if "too much" of the stream is *completely* crossed-over, force
+          -- a double-step there by splitting the stream to stay facing forward.
+          -- heuristic value was chosen by inspection on Subluminal - After Hours.
+          splitIndex = findIndex (isPrefixOf $ replicate 9 f) $ tails $ stepsLR s
+          -- prevent infinite splittage if the fux section starts immediately;
+          -- in that case split instead at fux's end (i.e., first non-crossed step).
+          splitStream 0 = splitStream $ fromJust $ findIndex (/= f) $ stepsLR s
+          splitStream i = s2 { doubles = doubles s2 + 1 }
+              where (steps1, steps2) = splitAt i $ stepsLR s
+                    s1 = commitStream s  { stepsLR = steps1 }
+                    s2 = commitStream s1 { stepsLR = steps2, lastJack = False }
+          -- normal case; consider the whole stream at once
[…]

### 2017-03-12 itg/code/ITG.hs
itg make parsing more forgiving of whitespace

@@ -14,7 +14,7 @@ commitStream s = maybe s0 splitStream splitIndex
           nx = length $ filter not $ stepsLR s
           -- if more than half the L/R steps in this stream were crossed over,
           -- then we got the footing backwards and need to flip the stream.
-          -- as a tiebreaker, flip if the stream is already more jacky than
+          -- as a tiebreaker, flip if the chart is already more jacky than
           -- footswitchy, i.e., if past streams flipped more often than not.
           f = nx * 2 > ns || nx * 2 == ns && ((jacks s > switches s) /= lastFlip s)
           -- if "too much" of the stream is *completely* crossed-over, force
@@ -72,14 +72,14 @@ process (title,(name:diff:feet:_:rest)) = intercalate "\t" $ metadata ++ result
     where -- chop off the trailing ":"
           trim = reverse . tail . dropWhile isSpace . reverse . dropWhile isSpace
           metadata = map trim [title,name,diff,feet]
-          finalState = analyze $ catMaybes $ map stepify rest
+          finalState = analyze $ catMaybes $ map (stepify . dropWhile isSpace) rest
           result = map (show . ($ finalState)) [steps, xovers, switches, jacks, doubles]
 
 -- splits up the input file by the "#NOTES:" lines
 charts :: String -> [String] -> [(String, [String])]
 charts t [] = []
 charts t xs = chart ++ charts title rest
-    where title = fromMaybe t $ drop 7 <$> find (isPrefixOf "#TITLE:") xs
+    where title = fromMaybe t $ drop 7 <$> dropWhile isSpace <$> find (isInfixOf "#TITLE:") xs
           -- get everything after the first "#NOTES"
           notnotes = not . isInfixOf "#NOTES:"
           stuff = tail $ dropWhile notnotes xs

### 2017-03-13 itg/2/paper.tex
write some itg paper

@@
 
+In which I deliver on last year's promise of future work.
+
 \end{abstract}
@@
 \keywords
-in, the, groove
+in, the, groove, 2
 
 \section{Introduction}
+
+Right, let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
+Unlike mainstream conferences, SIGBOVIK doesn't make me waste space repeating all the background material,
+and I can just say go read that paper first and get back to me.
+It's probably a lot funnier than this one anyway, which is gonna be sort of dry, and really of interest only to other ITG players.
+
+\begin{figure}[h]
+	\includegraphics[angle=90,width=0.48\textwidth]{../paper.pdf}
+	\caption{(okay twist your head to read this)}
+	\label{fig:you-stutid-fuckass}
+\end{figure}
 
@@
 \texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastStep :: \hilight{olivegreen}{Maybe}~\hilight{olivegreen}{Step}, doubleStep :: \hilight{olivegreen}{Bool}, lastFlip :: \hilight{olivegreen}{Bool},} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsL :: [\hilight{olivegreen}{Bool}], stepsR :: [\hilight{olivegreen}{Bool}] \}} \\
+\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsLR :: [\hilight{olivegreen}{Bool}] \}} \\
 \texttt{} \\
@@
 \texttt{~~~~~~~~~~~~~~~~~~~~ jacks~~~~= jacks~~~~s + \hilight{orange}{fromEnum}~(f /= lastFlip s \&\& doubleStep s), } \\
-	\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsL = \hilight{brickred}{[]}, stepsR = \hilight{brickred}{[]}~\}} \\
-\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~~~~~~~~~~~~~~stepsL s ++ stepsR s} \\
-\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsL s ++ stepsR s} \\
+	\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsLR = \hilight{brickred}{[]}~\}} \\
+\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~stepsLR s} \\
+\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsLR s} \\
 \texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
 \texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the stream is already more jacky than}} \\
+\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
 \texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
@@
 \texttt{~~~~\hilight{orange}{where}~foot = \hilight{orange}{not}~\$ lastFoot s} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or opposite L/R arrow here}} \\
-\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = ft:steps} \\
-\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = (\hilight{orange}{not}~ft):steps} \\
+\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or crossed-over L/R arrow.}} \\
+\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = steps ++ [ft]} \\
+\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = steps ++ [\hilight{orange}{not}~ft]} \\
 \texttt{~~~~~~~~~~addStep ft \_ steps = steps \hilight{commentblue}{-{}- U/D don't help to determine L/R footing.}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- add the (matching? xover?) step to the list corresponding to this foot.}} \\
-\texttt{~~~~~~~~~~newLs s = \hilight{orange}{if}~foot \hilight{orange}{then}~addStep foot step \$ stepsL s \hilight{orange}{else}~stepsL s} \\
-\texttt{~~~~~~~~~~newRs s = \hilight{orange}{if}~foot \hilight{orange}{then}~stepsR s \hilight{orange}{else}~addStep foot step \$ stepsR s} \\
-\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step,} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot = foot, stepsL = newLs s, stepsR = newRs s \} } \\
+\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step, lastFoot = foot,} \\
+\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ stepsLR = addStep foot step \$ stepsLR s \} } \\
 \texttt{} \\
@@
 	\texttt{\hilight{pink}{analyze}~= \hilight{pink}{commitStream}~.
-\hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[] []}) } \\
+\hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[]}) } \\
 \end{tabular}

### 2017-03-13 itg/code/ITG.hs
itg - make flip biased to jacks over switches in tie between jacks/switches - should fix all the 'switchy' beginner charts

@@ -16,7 +16,7 @@ commitStream s = maybe s0 splitStream splitIndex
           -- then we got the footing backwards and need to flip the stream.
           -- as a tiebreaker, flip if the chart is already more jacky than
           -- footswitchy, i.e., if past streams flipped more often than not.
-          f = nx * 2 > ns || nx * 2 == ns && ((jacks s > switches s) /= lastFlip s)
+          f = nx * 2 > ns || nx * 2 == ns && ((switches s > jacks s) == lastFlip s)
           -- if "too much" of the stream is *completely* crossed-over, force
           -- a double-step there by splitting the stream to stay facing forward.
           -- heuristic value was chosen by inspection on Subluminal - After Hours.

### 2017-03-13 itg/code/ITG.hs
itg - count xover footswitches by remembering the last jack arrow

@@ -4,8 +4,9 @@ import Data.Maybe
 
 data Step = L | D | U | R | Jump deriving Eq
 
-data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int, jacks :: Int,
-                         doubles :: Int, lastStep :: Maybe Step, lastJack :: Bool,
+data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int,
+                         jacks :: Int, doubles :: Int, xoverfs :: Int,
+                         lastStep :: Maybe Step, lastJack :: Maybe Step,
                          lastFlip :: Bool, lastFoot :: Bool, stepsLR :: [Bool] }
 
 commitStream :: AnalysisState -> AnalysisState
@@ -27,11 +28,14 @@ commitStream s = maybe s0 splitStream splitIndex
           splitStream i = s2 { doubles = doubles s2 + 1 }
               where (steps1, steps2) = splitAt i $ stepsLR s
                     s1 = commitStream s  { stepsLR = steps1 }
-                    s2 = commitStream s1 { stepsLR = steps2, lastJack = False }
+                    s2 = commitStream s1 { stepsLR = steps2, lastJack = Nothing }
           -- normal case; consider the whole stream at once
+          jack = lastJack s /= Nothing
+          jackLR = lastJack s == Just L || lastJack s == Just R
           s0 = s { xovers   = xovers   s + if f then ns - nx else nx,
-                   switches = switches s + fromEnum (f == lastFlip s && lastJack s),
-                   jacks    = jacks    s + fromEnum (f /= lastFlip s && lastJack s),
+                   switches = switches s + fromEnum (f == lastFlip s && jack),
+                   jacks    = jacks    s + fromEnum (f /= lastFlip s && jack),
+                   xoverfs  = xoverfs  s + fromEnum (f == lastFlip s && jackLR),
                    lastFlip = f, stepsLR = [] }
 
 analyzeStep :: AnalysisState -> Step -> AnalysisState
@@ -39,12 +43,12 @@ analyzeStep s step
     -- a jump resets the footing, so the next step can be stepped with either
     -- foot. commit the stream so far to treat it separately from what follows.
     -- bracket-jumps are, of course, future work.
-    | step == Jump = (commitStream s) { lastStep = Nothing, lastJack = False }
+    | step == Jump = (commitStream s) { lastStep = Nothing, lastJack = Nothing }
[…]

### 2017-03-14 itg/2/paper.tex
itg paper add a bunch of tables

@@
 and I can just say go read that paper first and get back to me.
-It's probably a lot funnier than this one anyway, which is gonna be sort of dry, and really of interest only to other ITG players.
+It's probably a lot funnier than this one anyway, which is gonna be sort of dry,
+and really of interest only to other ITG players who already know what's going on.
 
 \begin{figure}[h]
-	\includegraphics[angle=90,width=0.48\textwidth]{../paper.pdf}
+	\hspace{-2em}\includegraphics[angle=90,width=0.54\textwidth]{../paper.pdf}
 	\caption{(okay twist your head to read this)}
@@
 \section{Flashback Scene}
+
+Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitch footings.
 
@@
 \texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
-\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((jacks s > switches s) /= lastFlip s)} \\
+\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((switches s > jacks s) == lastFlip s)} \\
 \texttt{} \\
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section{A Bit More on Turniness}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Experimental Results}
+
+% TODO: Put a * on charts where you're going to put a pict of mario man
+% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended
+% TODO: Make a special mention of sandstorm and put a pict of mario man
+
+\begin{figure}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
+		\hline
+		\multicolumn{5}{c}{\em no 8s- with $\ge$100 XOs} \\
+		 9 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 107 & \em 8.7 \\
+		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo MeGaPacK    & \bf 141 & \em 3.9 \\
+		11 & Somebody That I Used To Know & Best of Gazebo    & \bf 114 & \em 11.9 \\
+		12 & Credens Justitiam        & Stuff B Likes         & \bf 136 & \em 12.4 \\
+		13 & Banshee Strikes          & VocaJawnz             & \bf 152 & \em 10.2 \\
+		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
+		15 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 101 & \em 4.2 \\
+		\multicolumn{5}{c}{\em no 16s-20s with $\ge$100 XOs} \\
+		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts with the most total crossovers (XOs).}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
+		\hline
+		 4 & DAM DIRIRAM              & DDR 3rdMIX            & \em  30 & \bf 27.3 \\
+		 5 & STRICTLY BUSINESS        & DDR 1st               & \em  33 & \bf 21.4 \\
+		 6 & STRICTLY BUSINESS        & DDR 1st               & \em  38 & \bf 20.9 \\
+		 7 & MOBO$\star$MOGA          & DDR EXTREME           & \em  37 & \bf 17.3 \\
+		 8 & PARANOiA                 & DDR 1st               & \em  58 & \bf 19.6 \\
+		 9 & Dazzlin Darlin           & r21twins              & \em  76 & \bf 22.0 \\
+		10 & Enchanted Journey        & ITG Rebirth           & \em  78 & \bf 19.3 \\
+		11 & Lune Noir                & r21freak Friendship   & \em  58 & \bf 13.4 \\
+		12 & Winnipeg is Fucking Over & best of r21freak ii   & \em 122 & \bf 14.3 \\
+		13 & The Sampling Paradise    & The Paradise Sampler  & \em 139 & \bf 15.6 \\
+		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
+		\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XO} \\
+		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts with the highest percentage of crossovers among total steps.}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
+		\hline
+		\multicolumn{5}{c}{\em no 5s- with $\ge$50 FSs} \\
+		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
+		\multicolumn{5}{c}{\em no 7s-9s with $\ge$50 FSs} \\
+		10 & Eat 'Em Up!      & Mute Sims 5           & \bf  54 & \em  9.0 \\
+		11 & Nemeton          & The Legend of Zim 4   & \bf  97 & \em 10.3 \\
+		12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
+		13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
+		14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
+		15 & Danse Macabre    & Aoreo's Ariginals     & \bf 201 & \em 12.2 \\
+		16 & Weird Science    & Stamina Showcase      & \bf  61 & \em  1.8 \\
+		\multicolumn{5}{c}{\em no 17s+ with $\ge$50 FSs} \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts with the most total footswitches (FSs).}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
+		\hline
+		%\multicolumn{5}{c}{\em no <6s with >50 FSs} \\
+		 3 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  26 & \bf 15.3 \\
+		 4 & MAKE IT BETTER   & DDR 2ndMIX            & \em  26 & \bf 20.0 \\
+		\multicolumn{5}{c}{\em no 5s with $\ge$10\% FS} \\
+		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
+		 7 & 5.1.1            & DDR MAX               & \em  13 & \bf 12.9 \\
+		\multicolumn{5}{c}{\em no 8s with $\ge$10\% FS} \\
+		 9 & PARANOiA KCET    & DDR 2ndMIX            & \em  33 & \bf 10.2 \\
+		10 & Delhi Ill        & Mute Sims 6           & \em  40 & \bf 11.6 \\
+		11 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  75 & \bf 11.8 \\
+		12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
+		13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
+		14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
+		15 & Flames of the Sky & fort rapids vii      & \em 172 & \bf 16.5 \\
+		\multicolumn{5}{c}{\em no 16s+ with $\ge$10\% FS} \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts with the highest percentage of footswitches among total steps.}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|c|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#XF & \bf XF\% \\
+		\hline
+		\multicolumn{5}{c}{\em no 8s- with $\ge$12 XFs} \\
+		 9 & Dr. Boom-Bombay   & fort rapids vii            & \bf 18 & 4.0 \\
+		10 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.5 \\
+		11 & Heart Shooter     & VocaJawnz                  & \bf 44 & 5.1 \\
+		12 & Beautiful Synergy & Keyboard Collaboration III & \bf 14 & 3.2 \\
+		13 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.2 \\
+		14 & Fancy Footwork    & Cirque du Zeppelin         & \bf 40 & 4.2 \\
+		\multicolumn{5}{c}{\em no 14s+ with $\ge$12 XFs} \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positives from DDR.)}
+\end{figure}
+
 \section{Conclusion}

### 2017-03-14 itg/2/paper.tex
Write about turniness.

@@
 \usepackage{natbib}
+\renewcommand{\UrlBreaks}{\do\/\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t\do\u\do\v\do\w\do\x\do\y\do\z}
 
@@
 \keywords
-in, the, groove, 2
+crossovers, footswitches, jacks, sidefoots
 
@@
 
-Right, let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
+Let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
 Unlike mainstream conferences, SIGBOVIK doesn't make me waste space repeating all the background material,
@@
 
+The TL;DR is that I made a program which figures out how to foot stepcharts in the least crossovery possible way (short of double-stepping everything),
+then found which charts ultimately had the most.
+The algorithm also naturally identifies footswitches and jacks, %, jacks, forced double-steps, and crossover-footswitches;
+and sometimes it's smarter than me in amusing ways.
+I put all the goodies in a giant spreadsheet at \url{http://tinyurl.com/crossoveriest},
+and the program itself is of course freely available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section{Flashback Scene}
+\section{Revisiting Turniness (Flashback Scene)}
 
-Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitch footings.
+Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitches.
+I show a typical $DD$/$UU$ footswitch pattern in Figure~\ref{fig:fs}(a), and typical $LL$/$RR$ switches (henceforth ``crossover footswitches'') in Figure~\ref{fig:fs}(b).
+To step these patterns, the player still alternates feet as usual, but must lift one foot off the repeated arrow before stepping it with her other foot.
+Chart authors will often, but not always, include a ``mine cue'' (shown in the figure) to hint that the second foot should switch onto the same arrow.
 
-\begin{table}[h]
+\begin{table}[t]
 	\begin{center}
@@
 		%                 L   D   U   R
-		& $\leftarrow$  & - & $UR$ & $UL$ & $U$ \\
-		& $\downarrow$  & $DL$* & - & $L$ & $UL$ \\
-		& $\uparrow$    & $DR$* & $R$ & - & $UR$ \\
-		& $\rightarrow$ & $D\dagger$ & $DR$* & $DL$* & - \\
+		& $\leftarrow$  & \bf ? & $UR$ & $UL$ & $U$ \\
+		& $\downarrow$  & $DL$ & \bf ? & $L$ & $UL$ \\
+		& $\uparrow$    & $DR$ & $R$ & \bf ? & $UR$ \\
+		& $\rightarrow$ & $D$ & $DR$ & $DL$ & \bf ? \\
 
@@
 	\end{center}
-	\caption{Facing directions. ``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
+	\caption{Facing directions.
+	%``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
+	}
 	\label{tab:facing}
 \end{table}
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-dd-uu.png} \\
+	(a) DD/UU
+	\end{tabular}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-ll-rr.png} \\
+	(b) LL/RR
+	\end{tabular}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-ll-rr-2.png} \\
+	(c) RR/LL
+	\end{tabular}
+	\caption{Footswitches of various crossoveriness/facing.}
+	\label{fig:fs}
+\end{figure}
+
+It is tempting to assign the facings $L$, $U$, $U$, and $R$ respectively to the $LL$, $DD$, $UU$, and $RR$ footswitches.
+However, Figure~\ref{fig:fs}(c) shows that if a footswitch begins with a crossover on $U$, the facing should be reversed:
+the $RR$ footing should face $L$, and $LL$ should face $R$.
+``Backwards footswitches'' with $D$ facing are also theoretically possible,
+arising from patterns such as $LURDDL$ or $LDRUUL$, although I know of no real-world chart which does this (hmm, tempting...).
+
+Before I realized that, I modified the turniness algorithm \cite{turniness} to face footswitches as above, %the bad way mentioned above,
+%to be faced $L$ for $LL$ and so on, as above,
+and it surprised me with charts of $\mathcal{T}>2$, in excess of the theoretical maximum!
+I show one such chart in Figure~\ref{fig:fs-turniness}(a), in which the step from $LL$ ($\phi=L$) to $UL$ ($\phi=DR$) has individual $\mathcal{T}=3$, and so on for $UL \rightsquigarrow_R UU \rightsquigarrow_L RU \rightsquigarrow_R RR$.
+The steps $RR \rightsquigarrow_L LR \rightsquigarrow_R LL$ are both candles ($\mathcal{T}=2$),
+resulting overall in $\mathcal{T}=8/3$ for the whole chart.
+
+Indeed, when I further modified the algorithm to force $DD$ switches to face $D$ (i.e., always facing the direction of the repeated arrow),
+it produced the chart shown in Figure~\ref{fig:fs-turniness}(b), with overall $\mathcal{T}=3$.
+(Note its resemblance to the basic spin pattern, $LDRU$, whose $\mathcal{T}=2$.)
+
+To fairly represent a human player's desire to step in the least turny of ambiguous ways,
+I extended the algorithm to provide either the assigned facing, from above, or its polar opposite,
+chosen at runtime by whichever is closer to the presequent facing.
+This restores the maximum overall chart turniness to $\mathcal{T}=2$,
+a new example of which is shown in Figure~\ref{fig:fs-turniness}(c).
+
+However, note that individual steps may still have $\mathcal{T}=3$,
+%because after a footswitch's facing is determined, the next step may still ... ?
+as shown in Figure~\ref{fig:fs-turniness}(d).
+In this example, the step $DL \rightsquigarrow_L LL$ assigns $LL$ to face $L$,
+but the subsequent step to $LD$ cannot avoid facing $UR$.
+The reason charts still cannot exceed overall $\mathcal{T}=2$ is that
+setting up such a situation requires a $\mathcal{T}=1$ step,
+which negates the benefit.
+A chart could conceivably end right before such a step,
+sneaking through some small $\epsilon$ extra turniness \cite{epsilon}
+(similar to the case of 270s in \cite{turniness}),
+but {\em sustained} average $\mathcal{T}>2$ remains impossible.
+
+Another approach could assign such a footswitch the opposite footing of the previous facing,
+%regardless of the switch arrow itself;
+so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
+Anyway, remember that all-candle switch stream from (c). We'll see it again soon.
+
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{c}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-turniness-8-3.png} \\
+	(a) $\mathcal{T}=8/3$(?)%\stackrel{?}{=}8/3$
+	\end{tabular}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-turniness-3.png} \\
+	(b) $\mathcal{T}=3$(?)%\stackrel{?}{=}3$
+	\end{tabular}
+	\\
+	\\
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-turniness-2.png} \\
+	(c) $\mathcal{T}=2$.
+	\end{tabular}
+	\begin{tabular}{c}
+	\includegraphics[width=0.13\textwidth]{fs-turniness-2-plus-epsilon.png} \\
+	(d) $\mathcal{T}=2+\epsilon$.
+	\end{tabular}
+	\end{tabular}
+	\end{center}
+	\caption{The turniest footswitch patterns. (a) and (b) are false positives (see prose), while (c) and (d) provide theoretically maximal turniness.}
+	\label{fig:fs-turniness}
+\end{figure}
 
@@
 % TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended
+
 % TODO: Make a special mention of sandstorm and put a pict of mario man
+% TODO: put the above in an honourable mention subsection, and also mention the slow train and sidefoots
 

### 2017-03-14 itg/2/paper.tex
describe xover alg

@@
 	\includegraphics[width=0.13\textwidth]{fs-dd-uu.png} \\
-	(a) DD/UU
+	(a) DD, UU
 	\end{tabular}
@@
 	\includegraphics[width=0.13\textwidth]{fs-ll-rr.png} \\
-	(b) LL/RR
+	(b) LL, RR
 	\end{tabular}
@@
 	\includegraphics[width=0.13\textwidth]{fs-ll-rr-2.png} \\
-	(c) RR/LL
+	(c) RR, LL
 	\end{tabular}
@@
 Another approach could assign such a footswitch the opposite footing of the previous facing,
-%regardless of the switch arrow itself;
+regardless of the arrow itself;
 so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
-Anyway, remember that all-candle switch stream from (c). We'll see it again soon.
+Anyway, remember that all-candle switch stream in (c). We'll see it again soon.
 
@@
 
+\section{Analyzing Crossoveriness}
+
+The major flaw of the turniness algorithm \cite{turniness}
+is that it didn't care whether a stream started with the left or right foot;
+it simply exploited the symmetry of Table~\ref{tab:facing} to find turniness regardless of footing.
+Hence, it could not distinguish technical footing patterns which could affect the way a human would play the chart.
+\cite{turniness} often played charts inhumanly, facing backwards and/or stepping 270s for most of a song.
+
+So, my contribution this year is an algorithm which plays charts more like a human,
+and which consequently can report on a chart's technical patterns beyond simple turniness.
+The algorithm realizes three principles of ITG:
+\begin{enumerate}
+	\item Alternate feet as much as possible.
+	\item Step crossed-over as little as possible.
+	\item Jumps allow the player to reset her footing.
+\end{enumerate}
+
+Figure~\ref{fig:pseudocode} describes the algorithm in pseudocode.
+To summarize it in prose:
+\begin{itemize}
+	\item Split the chart into several units of stream,
+		the boundaries of which occur at every jump and any time an arrow is repeated.
+	\item Step each stream with alternating feet.
+	\item Compare the number of matching steps (i.e., $L$ foot on $L$ arrow or $R$ on $R$) versus crossover steps.
+		If the latter is greater, re-step the stream with opposite feet from before (this kills the crossovers).
+	\item After flipping each stream, if necessary, count the total crossovers in the whole chart.
+\end{itemize}
+
 \newcommand\hilight[2]{\color{#1}#2\color{black}}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[p]
 \begin{center}
@@
 \end{center}
-\caption{it's pseudocode lol}
-\label{fig:pseudocode-lol}
+	\caption{Pseudocode description of the crossoveriness and footswitchiness algorithm.}
+\label{fig:pseudocode}
 \end{figure*}
 
-\section{Existing Chart Analysis}
+\section{Analyzing Footswitchiness}
+
+Because we split the chart whenever an arrow is repeated,
+figuring out whether that arrow is stepped with different feet on either side of the stream boundary
+is a natural consequence of figuring out how to step each stream individually.
+This is also shown in Figure~\ref{fig:pseudocode}'s pseudocode.
+To summarize in prose,
+if neither stream needed to be flipped (or if both did), then the alternating feet assumption holds,
+and the repeat must be a footswitch.
+
+\section{Analyzing Jackiness}
+
+A jack occurs when a repeat arrow is stepped with the same foot,
+rather than alternating (hence the name, from ``jackhammer'').
+You've got the idea by now, right?
 
@@
 
-%\section{A Bit More on Turniness}
+\section{Forced Doublesteps}
 

### 2017-03-15 itg/2/paper.tex
write heuristic algorithm description

@@
 \usepackage{natbib}
+\usepackage[mathcal]{euscript}
+\usepackage[linesnumbered,ruled]{algorithm2e}
+
 \renewcommand{\UrlBreaks}{\do\/\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t\do\u\do\v\do\w\do\x\do\y\do\z}
@@
 The major flaw of the turniness algorithm \cite{turniness}
-is that it didn't care whether a stream started with the left or right foot;
+was that it didn't care whether a stream started with the left or right foot;
 it simply exploited the symmetry of Table~\ref{tab:facing} to find turniness regardless of footing.
 Hence, it could not distinguish technical footing patterns which could affect the way a human would play the chart.
-\cite{turniness} often played charts inhumanly, facing backwards and/or stepping 270s for most of a song.
+It often played charts inhumanly, facing backwards and/or stepping 270s for most of a song.
 
-So, my contribution this year is an algorithm which plays charts more like a human,
+So, my contribution this year is an algorithm which plays more naturally,
 and which consequently can report on a chart's technical patterns beyond simple turniness.
@@
 	\item Step crossed-over as little as possible.
-	\item Jumps allow the player to reset her footing.
+	\item Jumps or jacks allow the player to reset her footing.
 \end{enumerate}
@@
 
+After painstakingly translating the pseudocode from Figure~\ref{fig:pseudocode} into a real implementation,
+I found it vulnerable to false positives when a single double-step could force a long section of stream to be stepped backwards.
+As an extreme example, consider the pattern $LRLRLRLR-D-LRLRLRLR$.
+Because no jumps or jacks allow the footing to be reset, either the first or last 4 pairs of $LR$s must be stepped crossed-over.
+
+Figure~\ref{fig:force-doublestep} shows two examples from real-world charts:
+in (a), the player must not alternate feet across the measure break,
+while in (b), two $L$ arrows are replaced with rolls for an artistic visual accent, which must be stepped twice each.
+On inspection, these charts should be stepped with no crossovers, but were evaluated otherwise %as having very high XO\%
+%(anywhere from 10.6\% to 27.9\%).
+(730 XOs (27.9\%) and 173 XOs (9.6\%), respectively).
+
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{paradise-lost-false-positive.png}
+		\\
+		(a) Paradise Lost \\
+		(16, Cirque du Lykan)
+	\end{tabular}
+	%\begin{tabular}{c}
+	%	\includegraphics[width=0.16\textwidth]{false-positive-toromi-hearts.png}
+	%	\\
+	%	(a) Toromi Hearts 2 \\
+	%	(18, Cirque du Zonda)
+	%\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{hearbeat-rollstream-doublestep.png}
+		\\
+		(b) Heartbeat \\
+		(13, TranceNation)
+	\end{tabular}
+	\end{center}
+	\caption{The doublesteps in some streamy charts must be identified, and the stream split, lest ``too much'' of the following stream appear completely crossed-over.}
+	\label{fig:force-doublestep}
+\end{figure}
+
+\newcommand\Stream{\ensuremath{\mathcal{S}}}
+\newcommand\Doublestep{\ensuremath{\mathcal{DS}}}
+\newcommand\nflip{\ensuremath{n_{\mathfrak{flip}}}}
+\newcommand\percentmatch{\ensuremath{\%_{\mathfrak{match}}}}
+\begin{algorithm}[h!]
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Invariant}
+	\Input{\Stream, a step sequence}
+	\Output{$\forall s_i, s_j \in \Stream, j=i+1 \rightarrow \mathsf{\neg StreamBoundary}(s_i,s_j)$}
+	\Input{\nflip, heuristic minimum length} %of flipped stream to doublestep into}
+	\Input{$\percentmatch$, heuristic percentage, initially 0}
+	\For{$i \in \mathsf{length}(\Stream) \wedge \neg {\sf defined}(i_{\Doublestep})$}{
+		%$\mathcal{S}' \leftarrow \{ s_j | s_j \in \Stream \wedge j \ge i \}$ \\ %i \le j < i + \nflip \})$ \\
+		$\Stream' \leftarrow \{ s_k | s_k \in \mathsf{LRs}(\{ s_j | s_j \in \Stream \wedge j \ge i \}) \wedge k \le \nflip \}$ \\
+		%$\mathcal{X} \leftarrow \mathsf{CrossoverSteps}(\{ s_j | s_j \in \Stream \wedge i \le j < i + \nflip \})$ \\
+		%$\mathcal{M} \leftarrow \mathsf{MatchingSteps} (\{ s_j | s_j \in \Stream \wedge i \le j < i + \nflip \})$
+		\If{$|\Stream'| = \nflip \wedge |\mathsf{Matching}(\Stream')| \le \percentmatch \times (|\mathsf{Crossovers}(\Stream')| + |\mathsf{Matching}(\Stream')|)$}{
+			$i_{\Doublestep} \leftarrow i$
+		}
+	}
+	\eIf{{\sf defined}($i_{\Doublestep})$}{
+		\If{$i_{\Doublestep} = 0$}{
+			% gotta write "tail stream" as complicatedly as possible
+			$i_{\Doublestep} \leftarrow \mathsf{FindUnflippedSection}(\{ s_i | s_i \in \Stream \wedge i \ne 0 \})$
+			%\mathsf{tail}(\Stream))$
+		}
+		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i < i_{\Doublestep} \}$)
+		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i \ge i_{\Doublestep} \}$)
+	}{
+		{\sf CommitStream}(\Stream)
+	}
+	\caption{{\sf HeuristicallyDoublestep}(\Stream)}
+	\label{alg:heuristic}
+\end{algorithm}
+
+To handle such cases,
+I extended the algorithm with a heuristic to identify when a stream becomes ``too crossed-over'' for ``too long'',
+and to force a doublestep by splitting the stream to flip it back.
+Algorithm~\ref{alg:heuristic} shows the implementation.
+I will not summarize how it works due to space limitations, but the description should be intuitive enough.
+The heuristic evaluated Figure~\ref{fig:force-doublestep}'s charts as having 0 crossovers and 1 and 21 doublesteps, respectively.
+In my analysis next section, I will use $\nflip = 9$
+(determined by inspection of a few favourite charts, which should be scientifically rigorous enough for anyone).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XO} \\
+		\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
 		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
@@
 	\end{center}
-	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positives from DDR.)}
+	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}
 \end{figure}
@@
 Please accept our paper.
-We worked hard on it.
+I worked hard on it.
 

### 2017-03-15 itg/code/itg.c
impl footswitches in turniness algorithm

@@ -2,7 +2,8 @@
 #include <stdbool.h>
 #include <assert.h>
 
-#define analysis_main main
+#define synthesis_main main
+#define FOOTSWITCHES
 
 /******************************************************************************
  * Common helpers
@@ -28,30 +29,6 @@ enum facing {
 };
 
 enum arrow { U, D, L, R, NUM_ARROWS };
-
-enum facing whichway(enum arrow left_foot, enum arrow right_foot) {
-	// Foot switches: future work!
-	assert(left_foot != right_foot);
-	static enum facing way_table[NUM_ARROWS][NUM_ARROWS] = {
-		[U][D] = F_R,
-		[U][L] = F_DR, // xover
-		[U][R] = F_UR,
-		//[U][U] = F_U, // footswitch
-		[L][U] = F_UL,
-		[L][D] = F_UR,
-		[L][R] = F_U,
-		//[L][L] = F_L, // xover footswitch
-		[D][U] = F_L,
-		[D][L] = F_DL, // xover
-		[D][R] = F_UL,
-		//[D][D] = F_U, // footswitch (to be theoretically beautiful, this would be F_D, a backwards switch, but who the hell does that?)
-		[R][L] = F_D, // lateral
-		[R][U] = F_DL, // xover
-		[R][D] = F_DR, // xover
-		//[R][R] = F_R, // xover footswitch
-	};
-	return way_table[left_foot][right_foot];
-}
 
 // For any two facings possible with consecutive itg steps (ie, only 1 foot
 // moves), shouldn't return more than 2.
@@ -91,6 +68,41 @@ unsigned int turn_distance(enum facing before, enum facing after)
 	return turn_table[before][after];
 }
 
+enum facing whichway(enum arrow left_foot, enum arrow right_foot, enum facing last_facing) {
+	// Foot switches: future work!
+#ifndef FOOTSWITCHES
+	assert(left_foot != right_foot);
+#endif
+	static enum facing way_table[NUM_ARROWS][NUM_ARROWS] = {
+		[U][D] = F_R,
+		[U][L] = F_DR, // xover
+		[U][R] = F_UR,
+		[L][U] = F_UL,
+		[L][D] = F_UR,
+		[L][R] = F_U,
+		[D][U] = F_L,
+		[D][L] = F_DL, // xover
+		[D][R] = F_UL,
+		[R][L] = F_D, // lateral
+		[R][U] = F_DL, // xover
+		[R][D] = F_DR, // xover
[…]

### 2017-03-15 itg/code/author-stats.sh
add spreadsheet and scripce for author/pack stats

@@ -1,0 +1,33 @@
+#!/bin/bash
+sheet=boring.tsv
+author="$1"
+
+# XXX
+if [ "$author" = "(-[Jayce]-)" ]; then
+	author="...Jayce..."
+fi
+
+# strip filename, song name, filter by author
+content=`cat "$sheet" | sed 's/^[^\t]*\t[^\t]*\t//' | grep "^$author"`
+
+# exclude 1s/2s/3s, strip author, chart name, and difficulty
+# grepping for tabs is horrible :(
+content=`echo "$content" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t//' | grep -ve "^1$(printf '\t')\|^2$(printf '\t')\|^3$(printf '\t')"`
+stats=`echo "$content" | sed 's/^[^\t]*\t//'`
+
+# grab each column and turn into a line for bc
+steps=`echo "$stats" | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+xovers=`echo "$stats" | sed 's/^[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+switches=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+jacks=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+doubles=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+sidefoots=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+
+steps=`echo "$steps" | bc`
+xovers=`echo "$xovers" | bc`
+switches=`echo "$switches" | bc`
+jacks=`echo "$jacks" | bc`
+doubles=`echo "$doubles" | bc`
+sidefoots=`echo "$sidefoots" | bc`
+
+echo -e "$author\t$steps\t$xovers\t$switches\t$jacks\t$doubles\t$sidefoots"

### 2017-03-15 itg/code/pack-stats.sh

@@ -1,0 +1,28 @@
+#!/bin/bash
+sheet=boring.tsv
+pack="$1"
+
+# strip filename, song name, filter by author
+content=`cat "$sheet" | grep "^$pack"`
+
+# exclude 1s/2s/3s, strip author, chart name, and difficulty
+# grepping for tabs is horrible :(
+content=`echo "$content" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | grep -ve "^1$(printf '\t')\|^2$(printf '\t')\|^3$(printf '\t')"`
+stats=`echo "$content" | sed 's/^[^\t]*\t//'`
+
+# grab each column and turn into a line for bc
+steps=`echo "$stats" | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+xovers=`echo "$stats" | sed 's/^[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+switches=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+jacks=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+doubles=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+sidefoots=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
+
+steps=`echo "$steps" | bc`
+xovers=`echo "$xovers" | bc`
+switches=`echo "$switches" | bc`
+jacks=`echo "$jacks" | bc`
+doubles=`echo "$doubles" | bc`
+sidefoots=`echo "$sidefoots" | bc`
+
+echo -e "$pack\t$steps\t$xovers\t$switches\t$jacks\t$doubles\t$sidefoots"

### 2017-03-15 itg/2/paper.tex
make heuristic alg more concise
change some stuff up

@@
 the $RR$ footing should face $L$, and $LL$ should face $R$.
-``Backwards footswitches'' with $D$ facing are also theoretically possible,
+``Spin-switches'' with $D$ facing are also theoretically possible,
 arising from patterns such as $LURDDL$ or $LDRUUL$, although I know of no real-world chart which does this (hmm, tempting...).
@@
 	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{paradise-lost-false-positive.png}
+		\includegraphics[width=0.18\textwidth]{paradise-lost-false-positive.png}
 		\\
@@
 	%\begin{tabular}{c}
-	%	\includegraphics[width=0.16\textwidth]{false-positive-toromi-hearts.png}
+	%	\includegraphics[width=0.18\textwidth]{false-positive-toromi-hearts.png}
 	%	\\
@@
 	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{hearbeat-rollstream-doublestep.png}
+		\includegraphics[width=0.18\textwidth]{hearbeat-rollstream-doublestep.png}
 		\\
@@
 \newcommand\nflip{\ensuremath{n_{\mathfrak{flip}}}}
-\newcommand\percentmatch{\ensuremath{\%_{\mathfrak{match}}}}
+\newcommand\percentflip{\ensuremath{\%_{\mathfrak{flip}}}}
 \begin{algorithm}[h!]
@@
 	\SetKwInOut{Output}{Invariant}
-	\Input{\Stream, a step sequence}
+	\Input{\Stream, a step sequence $s_0 \dots s_n$}
 	\Output{$\forall s_i, s_j \in \Stream, j=i+1 \rightarrow \mathsf{\neg StreamBoundary}(s_i,s_j)$}
 	\Input{\nflip, heuristic minimum length} %of flipped stream to doublestep into}
-	\Input{$\percentmatch$, heuristic percentage, initially 0}
+	\Input{$\percentflip$, heuristic percentage, initially 100}
 	\For{$i \in \mathsf{length}(\Stream) \wedge \neg {\sf defined}(i_{\Doublestep})$}{
-		%$\mathcal{S}' \leftarrow \{ s_j | s_j \in \Stream \wedge j \ge i \}$ \\ %i \le j < i + \nflip \})$ \\
 		$\Stream' \leftarrow \{ s_k | s_k \in \mathsf{LRs}(\{ s_j | s_j \in \Stream \wedge j \ge i \}) \wedge k \le \nflip \}$ \\
-		%$\mathcal{X} \leftarrow \mathsf{CrossoverSteps}(\{ s_j | s_j \in \Stream \wedge i \le j < i + \nflip \})$ \\
-		%$\mathcal{M} \leftarrow \mathsf{MatchingSteps} (\{ s_j | s_j \in \Stream \wedge i \le j < i + \nflip \})$
-		\If{$|\Stream'| = \nflip \wedge |\mathsf{Matching}(\Stream')| \le \percentmatch \times (|\mathsf{Crossovers}(\Stream')| + |\mathsf{Matching}(\Stream')|)$}{
+		\If{$|\Stream'| = \nflip$ $\wedge$ $|\mathsf{Crossovers}(\Stream')| \ge \percentflip \times |\Stream'|$}{
 			$i_{\Doublestep} \leftarrow i$
@@
 I will not summarize how it works due to space limitations, but the description should be intuitive enough.
-The heuristic evaluated Figure~\ref{fig:force-doublestep}'s charts as having 0 crossovers and 1 and 21 doublesteps, respectively.
+The heuristic evaluated Figure~\ref{fig:force-doublestep}'s charts as having 0 crossovers each and 1 and 21 doublesteps, respectively.
 In my analysis next section, I will use $\nflip = 9$
@@
 
+% Stuff to talk about.
+% Dataset size now >10k
+% I reran turniness on fancy footwork and it actually got turnier (double check dis), b/c of all the candles
+
+% By author/by pack: I excluded 1s-3s, admittedly biased against konami, on the pretense of excluding nonsense false positives, but even after that, konami is still the crossover king
+% Step authors are limited to 10 or more charts
+
+% Miscellaneous charts I want to exhibit:
+% Conflict forced doublesteps
+% Heart shooter 'sidefoots'
+% Boom Bombay - triple xover jack footed as sideswitch
+% Toxic - jack+doublestep footed as sideswitch + xover stream instead
+% H'ble mention to sandstorm for jacks
+
 \begin{figure}[t]
@@
 		\small
-	\begin{tabular}{r|l|l|c|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
+	\begin{tabular}{r|l|l|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#XO \\
 		\hline
-		\multicolumn{5}{c}{\em no 8s- with $\ge$100 XOs} \\
-		 9 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 107 & \em 8.7 \\
-		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo MeGaPacK    & \bf 141 & \em 3.9 \\
-		11 & Somebody That I Used To Know & Best of Gazebo    & \bf 114 & \em 11.9 \\
-		12 & Credens Justitiam        & Stuff B Likes         & \bf 136 & \em 12.4 \\
-		13 & Banshee Strikes          & VocaJawnz             & \bf 152 & \em 10.2 \\
-		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		15 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 101 & \em 4.2 \\
-		\multicolumn{5}{c}{\em no 16s-20s with $\ge$100 XOs} \\
-		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
+		\multicolumn{4}{c}{\em no 8s- with $\ge$100 XOs} \\
+		 9 & MAX Forever              & CuoReNeRo & 107 \\
+		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo & 141 \\
+		11 & Somebody I Used To Know    & Best of Gazebo    & 114 \\
+		12 & Credens Justitiam        & Stuff B Likes         & 136 \\
+		13 & Banshee Strikes          & VocaJawnz             & 152 \\
+		14 & Slow Down                & Sexuality Violation 2 & 160 \\
+		15 & MAX Forever              & CuoReNeRo & 101 \\
+		\multicolumn{4}{c}{\em no 16s-20s with $\ge$100 XOs} \\
+		21 & Teenage Dream            & Sexuality Violation 2 & 280 \\
+		%\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
+		%\hline
+		%\multicolumn{5}{c}{\em no 8s- with $\ge$100 XOs} \\
+		% 9 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 107 & \em 8.7 \\
+		%10 & J-PARA SUPER MEGAMIX     & CuoReNeRo MeGaPacK    & \bf 141 & \em 3.9 \\
+		%11 & Somebody That I Used To Know & Best of Gazebo    & \bf 114 & \em 11.9 \\
+		%12 & Credens Justitiam        & Stuff B Likes         & \bf 136 & \em 12.4 \\
+		%13 & Banshee Strikes          & VocaJawnz             & \bf 152 & \em 10.2 \\
+		%14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
+		%15 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 101 & \em 4.2 \\
+		%\multicolumn{5}{c}{\em no 16s-20s with $\ge$100 XOs} \\
+		%21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
 	\end{tabular}
@@
 		\small
-	\begin{tabular}{r|l|l|c|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
+	\begin{tabular}{r|l|l|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf XO\% \\
 		\hline
-		 4 & DAM DIRIRAM              & DDR 3rdMIX            & \em  30 & \bf 27.3 \\
-		 5 & STRICTLY BUSINESS        & DDR 1st               & \em  33 & \bf 21.4 \\
-		 6 & STRICTLY BUSINESS        & DDR 1st               & \em  38 & \bf 20.9 \\
-		 7 & MOBO$\star$MOGA          & DDR EXTREME           & \em  37 & \bf 17.3 \\
-		 8 & PARANOiA                 & DDR 1st               & \em  58 & \bf 19.6 \\
-		 9 & Dazzlin Darlin           & r21twins              & \em  76 & \bf 22.0 \\
-		10 & Enchanted Journey        & ITG Rebirth           & \em  78 & \bf 19.3 \\
-		11 & Lune Noir                & r21freak Friendship   & \em  58 & \bf 13.4 \\
-		12 & Winnipeg is Fucking Over & best of r21freak ii   & \em 122 & \bf 14.3 \\
-		13 & The Sampling Paradise    & The Paradise Sampler  & \em 139 & \bf 15.6 \\
-		14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
-		21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
+		 4 & DAM DIRIRAM              & DDR 3rdMIX            & 27.3 \\
+		 5 & STRICTLY BUSINESS        & DDR 1st               & 21.4 \\
+		 6 & STRICTLY BUSINESS        & DDR 1st               & 20.9 \\
+		 7 & MOBO$\star$MOGA          & DDR EXTREME           & 17.3 \\
+		 8 & PARANOiA                 & DDR 1st               & 19.6 \\
+		 9 & Dazzlin Darlin           & r21twins              & 22.0 \\
+		10 & Enchanted Journey        & ITG Rebirth           & 19.3 \\
+		11 & Lune Noir                & r21freak Friendship   & 13.4 \\
+		12 & W'peg is Fucking Over    & best of r21freak ii   & 14.3 \\
+		13 & The Sampling Paradise    & The Paradise Sampler  & 15.6 \\
+		14 & Slow Down                & Sexuality Violation 2 & 13.3 \\
+		\multicolumn{4}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
+		21 & Teenage Dream            & Sexuality Violation 2 & 11.8 \\
+		%\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
+		%\hline
+		% 4 & DAM DIRIRAM              & DDR 3rdMIX            & \em  30 & \bf 27.3 \\
+		% 5 & STRICTLY BUSINESS        & DDR 1st               & \em  33 & \bf 21.4 \\
+		% 6 & STRICTLY BUSINESS        & DDR 1st               & \em  38 & \bf 20.9 \\
+		% 7 & MOBO$\star$MOGA          & DDR EXTREME           & \em  37 & \bf 17.3 \\
+		% 8 & PARANOiA                 & DDR 1st               & \em  58 & \bf 19.6 \\
+		% 9 & Dazzlin Darlin           & r21twins              & \em  76 & \bf 22.0 \\
+		%10 & Enchanted Journey        & ITG Rebirth           & \em  78 & \bf 19.3 \\
+		%11 & Lune Noir                & r21freak Friendship   & \em  58 & \bf 13.4 \\
+		%12 & Winnipeg is Fucking Over & best of r21freak ii   & \em 122 & \bf 14.3 \\
+		%13 & The Sampling Paradise    & The Paradise Sampler  & \em 139 & \bf 15.6 \\
+		%14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
+		%\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
+		%21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
 	\end{tabular}
 	\end{center}
-	\caption{Charts with the highest percentage of crossovers among total steps.}
+	\caption{Charts with the highest percentage of XOs among total steps.}
 \end{figure}
@@
 		\small
-	\begin{tabular}{r|l|l|c|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
+	\begin{tabular}{r|l|l|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#FS \\
 		\hline
-		\multicolumn{5}{c}{\em no 5s- with $\ge$50 FSs} \\
-		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
-		\multicolumn{5}{c}{\em no 7s-9s with $\ge$50 FSs} \\
-		10 & Eat 'Em Up!      & Mute Sims 5           & \bf  54 & \em  9.0 \\
-		11 & Nemeton          & The Legend of Zim 4   & \bf  97 & \em 10.3 \\
-		12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
-		13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
-		14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
-		15 & Danse Macabre    & Aoreo's Ariginals     & \bf 201 & \em 12.2 \\
-		16 & Weird Science    & Stamina Showcase      & \bf  61 & \em  1.8 \\
-		\multicolumn{5}{c}{\em no 17s+ with $\ge$50 FSs} \\
+		\multicolumn{4}{c}{\em no 5s- with $\ge$50 FSs} \\
+		 6 & Sweat Shop       & ITG Rebirth 2 Beta    &  56 \\
+		\multicolumn{4}{c}{\em no 7s-9s with $\ge$50 FSs} \\
+		10 & Eat 'Em Up!      & Mute Sims 5           &  54 \\
+		11 & Nemeton          & The Legend of Zim 4   &  97 \\
+		12 & Nemeton          & Subluminal            & 140 \\
+		13 & Love Is Eternity & Subluminal            & 140 \\
+		14 & Switch           & Getty                 & 347 \\
+		15 & Danse Macabre    & Aoreo's Ariginals     & 201 \\
+		16 & Weird Science    & Stamina Showcase      &  61 \\
+		\multicolumn{4}{c}{\em no 17s+ with $\ge$50 FSs} \\
+		%\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
+		%\hline
+		%\multicolumn{5}{c}{\em no 5s- with $\ge$50 FSs} \\
+		% 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
+		%\multicolumn{5}{c}{\em no 7s-9s with $\ge$50 FSs} \\
+		%10 & Eat 'Em Up!      & Mute Sims 5           & \bf  54 & \em  9.0 \\
+		%11 & Nemeton          & The Legend of Zim 4   & \bf  97 & \em 10.3 \\
+		%12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
+		%13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
+		%14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
+		%15 & Danse Macabre    & Aoreo's Ariginals     & \bf 201 & \em 12.2 \\
+		%16 & Weird Science    & Stamina Showcase      & \bf  61 & \em  1.8 \\
+		%\multicolumn{5}{c}{\em no 17s+ with $\ge$50 FSs} \\
 	\end{tabular}
@@
 		\small
-	\begin{tabular}{r|l|l|c|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
+	\begin{tabular}{r|l|l|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf FS\% \\
 		\hline
-		%\multicolumn{5}{c}{\em no <6s with >50 FSs} \\
-		 3 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  26 & \bf 15.3 \\
-		 4 & MAKE IT BETTER   & DDR 2ndMIX            & \em  26 & \bf 20.0 \\
-		\multicolumn{5}{c}{\em no 5s with $\ge$10\% FS} \\
-		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
-		 7 & 5.1.1            & DDR MAX               & \em  13 & \bf 12.9 \\
-		\multicolumn{5}{c}{\em no 8s with $\ge$10\% FS} \\
-		 9 & PARANOiA KCET    & DDR 2ndMIX            & \em  33 & \bf 10.2 \\
-		10 & Delhi Ill        & Mute Sims 6           & \em  40 & \bf 11.6 \\
-		11 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  75 & \bf 11.8 \\
-		12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
-		13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
-		14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
-		15 & Flames of the Sky & fort rapids vii      & \em 172 & \bf 16.5 \\
-		\multicolumn{5}{c}{\em no 16s+ with $\ge$10\% FS} \\
+		 3 & Sweat Shop       & ITG Rebirth 2 Beta    & 15.3 \\
+		 4 & MAKE IT BETTER   & DDR 2ndMIX            & 20.0 \\
+		\multicolumn{4}{c}{\em no 5s with $\ge$10\% FS} \\
+		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & 18.2 \\
+		 7 & 5.1.1            & DDR MAX               & 12.9 \\
+		\multicolumn{4}{c}{\em no 8s with $\ge$10\% FS} \\
+		 9 & PARANOiA KCET    & DDR 2ndMIX            & 10.2 \\
+		10 & Delhi Ill        & Mute Sims 6           & 11.6 \\
+		11 & Sweat Shop       & ITG Rebirth 2 Beta    & 11.8 \\
+		12 & Nemeton          & Subluminal            & 14.8 \\
+		13 & Love Is Eternity & Subluminal            & 16.8 \\
+		14 & Switch           & Getty                 & 15.7 \\
+		15 & Flames of the Sky & fort rapids vii      & 16.5 \\
+		\multicolumn{4}{c}{\em no 16s+ with $\ge$10\% FS} \\
+		%\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
+		%\hline
+		% 3 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  26 & \bf 15.3 \\
+		% 4 & MAKE IT BETTER   & DDR 2ndMIX            & \em  26 & \bf 20.0 \\
+		%\multicolumn{5}{c}{\em no 5s with $\ge$10\% FS} \\
+		% 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
+		% 7 & 5.1.1            & DDR MAX               & \em  13 & \bf 12.9 \\
+		%\multicolumn{5}{c}{\em no 8s with $\ge$10\% FS} \\
+		% 9 & PARANOiA KCET    & DDR 2ndMIX            & \em  33 & \bf 10.2 \\
+		%10 & Delhi Ill        & Mute Sims 6           & \em  40 & \bf 11.6 \\
+		%11 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  75 & \bf 11.8 \\
+		%12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
+		%13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
+		%14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
+		%15 & Flames of the Sky & fort rapids vii      & \em 172 & \bf 16.5 \\
+		%\multicolumn{5}{c}{\em no 16s+ with $\ge$10\% FS} \\
 	\end{tabular}
 	\end{center}
-	\caption{Charts with the highest percentage of footswitches among total steps.}
+	\caption{Charts with the highest percentage of FSs among total steps.}
 \end{figure}
@@
 		\small
-	\begin{tabular}{r|l|l|c|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#XF & \bf XF\% \\
+	\begin{tabular}{r|l|l|c}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#XF \\
 		\hline
-		\multicolumn{5}{c}{\em no 8s- with $\ge$12 XFs} \\
-		 9 & Dr. Boom-Bombay   & fort rapids vii            & \bf 18 & 4.0 \\
-		10 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.5 \\
-		11 & Heart Shooter     & VocaJawnz                  & \bf 44 & 5.1 \\
-		12 & Beautiful Synergy & Keyboard Collaboration III & \bf 14 & 3.2 \\
-		13 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.2 \\
-		14 & Fancy Footwork    & Cirque du Zeppelin         & \bf 40 & 4.2 \\
-		\multicolumn{5}{c}{\em no 14s+ with $\ge$12 XFs} \\
+		\multicolumn{4}{c}{\em no 8s- with $\ge$12 XFs} \\
+		 9 & Dr. Boom-Bombay   & fort rapids vii            & 18 \\
+		10 & Toxic             & Sexuality Violation 2      & 12 \\
+		11 & Heart Shooter     & VocaJawnz                  & 44 \\
+		12 & Beautiful Synergy & Keyboard Collaboration III & 14 \\
+		13 & Toxic             & Sexuality Violation 2      & 12 \\
+		14 & Fancy Footwork    & Cirque du Zeppelin         & 40 \\
+		% 9 & Dr. Boom-Bombay   & fort rapids vii            & \bf 18 & 4.0 \\
+		%10 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.5 \\
+		%11 & Heart Shooter     & VocaJawnz                  & \bf 44 & 5.1 \\
+		%12 & Beautiful Synergy & Keyboard Collaboration III & \bf 14 & 3.2 \\
+		%13 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.2 \\
+		%14 & Fancy Footwork    & Cirque du Zeppelin         & \bf 40 & 4.2 \\
+		\multicolumn{4}{c}{\em no 15s+ with $\ge$12 XFs} \\
 	\end{tabular}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Never Work}
+
+Let's be honest: this isn't gonna be a trilogy.
+
+With that out of the way, here are some things that would be cool to implement in a fantasy universe with infinite free time.
+
+% TODO
+% forced doublesteps via holds
+% - bracketjump footswitches - matador from valex 8
+% understanding mine cues for doublesteps
+
 \section{Conclusion}
 
-Please accept our paper.
+Please accept my paper.
 I worked hard on it.

### 2017-03-15 itg/code/author-stats.sh

@@ -23,6 +23,7 @@ jacks=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr
 doubles=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
 sidefoots=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
 
+charts=`echo "$stats" | wc -l`
 steps=`echo "$steps" | bc`
 xovers=`echo "$xovers" | bc`
 switches=`echo "$switches" | bc`
@@ -30,4 +31,4 @@ jacks=`echo "$jacks" | bc`
 doubles=`echo "$doubles" | bc`
 sidefoots=`echo "$sidefoots" | bc`
 
-echo -e "$author\t$steps\t$xovers\t$switches\t$jacks\t$doubles\t$sidefoots"
+echo -e "$author\t$charts\t$steps\t$xovers\t$switches\t$jacks\t$doubles\t$sidefoots"

### 2017-03-15 itg/code/pack-stats.sh

@@ -18,6 +18,7 @@ jacks=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr
 doubles=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
 sidefoots=`echo "$stats" | sed 's/^[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t[^\t]*\t//' | sed 's/\t.*//' | tr '\n' '+' | sed 's/$/0/'`
 
+charts=`echo "$stats" | wc -l`
 steps=`echo "$steps" | bc`
 xovers=`echo "$xovers" | bc`
 switches=`echo "$switches" | bc`
@@ -25,4 +26,4 @@ jacks=`echo "$jacks" | bc`
 doubles=`echo "$doubles" | bc`
 sidefoots=`echo "$sidefoots" | bc`
 
-echo -e "$pack\t$steps\t$xovers\t$switches\t$jacks\t$doubles\t$sidefoots"
+echo -e "$pack\t$charts\t$steps\t$xovers\t$switches\t$jacks\t$doubles\t$sidefoots"

### 2017-03-15 itg/2/paper.tex
update tables

@@
 % I reran turniness on fancy footwork and it actually got turnier (double check dis), b/c of all the candles
+% 9-15 is clearly the sweet spot for techni jawnz
 
@@
 
-\begin{figure}[t]
+% charts
+\begin{table}[t]
 	\begin{center}
 		\small
-	\begin{tabular}{r|l|l|c}
+	\begin{tabular}{r|l|l|l}
 		\bf Ft. & \bf Name & \bf Pack & \bf \#XO \\
 		\hline
-		\multicolumn{4}{c}{\em no 8s- with $\ge$100 XOs} \\
+		 6 & Autoload                 & ITG 3                 &  54 \\
+		 7 & Tetris                   & CuoReNeRo             &  58 \\
+		 8 & Pulse                    & CuoReNeRo             &  94 \\
 		 9 & MAX Forever              & CuoReNeRo             & 107 \\
@@
 		14 & Slow Down                & Sexuality Violation 2 & 160 \\
-		15 & MAX Forever              & CuoReNeRo & 101 \\
-		\multicolumn{4}{c}{\em no 16s-20s with $\ge$100 XOs} \\
+		15 & yoshikawa45 vs siesta45  & Rikame's Simfiles 4   & 111 \\
+		16 & Your Best Nightmare      & Undertale             &  97 \\
+		\multicolumn{4}{c}{\em no 17s-19s with $\ge$50 XOs} \\
+		20 & Rainbow Dimension        & Rikame's Simfiles 2   &  84 \\
 		21 & Teenage Dream            & Sexuality Violation 2 & 280 \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
-		%\hline
-		%\multicolumn{5}{c}{\em no 8s- with $\ge$100 XOs} \\
-		% 9 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 107 & \em 8.7 \\
-		%10 & J-PARA SUPER MEGAMIX     & CuoReNeRo MeGaPacK    & \bf 141 & \em 3.9 \\
-		%11 & Somebody That I Used To Know & Best of Gazebo    & \bf 114 & \em 11.9 \\
-		%12 & Credens Justitiam        & Stuff B Likes         & \bf 136 & \em 12.4 \\
-		%13 & Banshee Strikes          & VocaJawnz             & \bf 152 & \em 10.2 \\
-		%14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		%15 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 101 & \em 4.2 \\
-		%\multicolumn{5}{c}{\em no 16s-20s with $\ge$100 XOs} \\
-		%21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
 	\end{tabular}
@@
 	\caption{Charts with the most total crossovers (XOs).}
-\end{figure}
-\begin{figure}[t]
+\end{table}
+\begin{table}[t]
 	\begin{center}
 		\small
-	\begin{tabular}{r|l|l|c}
+	\begin{tabular}{r|l|l|l}
 		\bf Ft. & \bf Name & \bf Pack & \bf XO\% \\
@@
 		14 & Slow Down                & Sexuality Violation 2 & 13.3 \\
-		\multicolumn{4}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
+		15 & yoshikawa45 vs siesta45  & Rikame's Simfiles 4   & 10.8 \\
+		\multicolumn{4}{c}{\em no 16s-19s with $\ge$10\% XOs} \\
+		20 & Rainbow Dimension        & Rikame's Simfiles 2   & 10.7 \\
 		21 & Teenage Dream            & Sexuality Violation 2 & 11.8 \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
-		%\hline
-		% 4 & DAM DIRIRAM              & DDR 3rdMIX            & \em  30 & \bf 27.3 \\
-		% 5 & STRICTLY BUSINESS        & DDR 1st               & \em  33 & \bf 21.4 \\
-		% 6 & STRICTLY BUSINESS        & DDR 1st               & \em  38 & \bf 20.9 \\
-		% 7 & MOBO$\star$MOGA          & DDR EXTREME           & \em  37 & \bf 17.3 \\
-		% 8 & PARANOiA                 & DDR 1st               & \em  58 & \bf 19.6 \\
-		% 9 & Dazzlin Darlin           & r21twins              & \em  76 & \bf 22.0 \\
-		%10 & Enchanted Journey        & ITG Rebirth           & \em  78 & \bf 19.3 \\
-		%11 & Lune Noir                & r21freak Friendship   & \em  58 & \bf 13.4 \\
-		%12 & Winnipeg is Fucking Over & best of r21freak ii   & \em 122 & \bf 14.3 \\
-		%13 & The Sampling Paradise    & The Paradise Sampler  & \em 139 & \bf 15.6 \\
-		%14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		%\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
-		%21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
 	\end{tabular}
@@
 	\caption{Charts with the highest percentage of XOs among total steps.}
-\end{figure}
-\begin{figure}[t]
+\end{table}
+\begin{table}[t]
 	\begin{center}
 		\small
-	\begin{tabular}{r|l|l|c}
+	\begin{tabular}{r|l|l|l}
 		\bf Ft. & \bf Name & \bf Pack & \bf \#FS \\
 		\hline
-		\multicolumn{4}{c}{\em no 5s- with $\ge$50 FSs} \\
+		%\multicolumn{4}{c}{\em no 5s- with $\ge$50 FSs} \\
 		 6 & Sweat Shop       & ITG Rebirth 2 Beta    &  56 \\
-		\multicolumn{4}{c}{\em no 7s-9s with $\ge$50 FSs} \\
+		 7 & Silent Hill      & DDR 3rdMIX            &  29 \\
+		 8 & Pulse            & CuoReNeRo             &  41 \\
+		 9 & Dr. Boom-Bombay  & fort rapids vii       &  34 \\
 		10 & Eat 'Em Up!      & Mute Sims 5           &  54 \\
@@
 		16 & Weird Science    & Stamina Showcase      &  61 \\
-		\multicolumn{4}{c}{\em no 17s+ with $\ge$50 FSs} \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
-		%\hline
-		%\multicolumn{5}{c}{\em no 5s- with $\ge$50 FSs} \\
-		% 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
-		%\multicolumn{5}{c}{\em no 7s-9s with $\ge$50 FSs} \\
-		%10 & Eat 'Em Up!      & Mute Sims 5           & \bf  54 & \em  9.0 \\
-		%11 & Nemeton          & The Legend of Zim 4   & \bf  97 & \em 10.3 \\
-		%12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
-		%13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
-		%14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
-		%15 & Danse Macabre    & Aoreo's Ariginals     & \bf 201 & \em 12.2 \\
-		%16 & Weird Science    & Stamina Showcase      & \bf  61 & \em  1.8 \\
-		%\multicolumn{5}{c}{\em no 17s+ with $\ge$50 FSs} \\
+		17 & Arcane Apparatus & Tachyon Gamma         &  32 \\
+		18 & Metallic-A-      & Oh Henry! Mad Stamina &  27 \\
+		19 & Geronimo         & Sexuality Violation 3 &  39 \\
+		20 & Scatman's World  & Jimmy Jawns 2         &  22 \\
+		21 & He He He         & Jimmy Jawns 2         &   4 \\
+		22 & Architecture     & SPEEEDCOOOORE 4       &  24 \\
+		23 & Geronimo         & Sexuality Violation 3 &  39 \\
 	\end{tabular}
@@
 	\caption{Charts with the most total footswitches (FSs).}
-\end{figure}
-\begin{figure}[t]
+\end{table}
+\begin{table}[t]
 	\begin{center}
 		\small
-	\begin{tabular}{r|l|l|c}
+	\begin{tabular}{r|l|l|l}
 		\bf Ft. & \bf Name & \bf Pack & \bf FS\% \\
@@
 		 3 & Sweat Shop       & ITG Rebirth 2 Beta    & 15.3 \\
-		 4 & MAKE IT BETTER   & DDR 2ndMIX            & 20.0 \\
-		\multicolumn{4}{c}{\em no 5s with $\ge$10\% FS} \\
+		 4 & DROP THE BOMB    & DDR 3rdMIX            &  9.8 \\
+		 5 & MAKE IT BETTER   & DDR 2ndMIX            & 20.0 \\
 		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & 18.2 \\
 		 7 & 5.1.1            & DDR MAX               & 12.9 \\
-		\multicolumn{4}{c}{\em no 8s with $\ge$10\% FS} \\
+		 8 & La Se\~{n}orita Virtual & DDR 3rdMIX     &  8.4 \\
 		 9 & PARANOiA KCET    & DDR 2ndMIX            & 10.2 \\
@@
 		15 & Flames of the Sky & fort rapids vii      & 16.5 \\
-		\multicolumn{4}{c}{\em no 16s+ with $\ge$10\% FS} \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
-		%\hline
-		% 3 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  26 & \bf 15.3 \\
-		% 4 & MAKE IT BETTER   & DDR 2ndMIX            & \em  26 & \bf 20.0 \\
-		%\multicolumn{5}{c}{\em no 5s with $\ge$10\% FS} \\
-		% 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
-		% 7 & 5.1.1            & DDR MAX               & \em  13 & \bf 12.9 \\
-		%\multicolumn{5}{c}{\em no 8s with $\ge$10\% FS} \\
-		% 9 & PARANOiA KCET    & DDR 2ndMIX            & \em  33 & \bf 10.2 \\
-		%10 & Delhi Ill        & Mute Sims 6           & \em  40 & \bf 11.6 \\
-		%11 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  75 & \bf 11.8 \\
-		%12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
-		%13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
-		%14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
-		%15 & Flames of the Sky & fort rapids vii      & \em 172 & \bf 16.5 \\
-		%\multicolumn{5}{c}{\em no 16s+ with $\ge$10\% FS} \\
+		16 & Mermaid Island   & Tachyon Alpha         &  9.5 \\
+		\multicolumn{4}{c}{\em no 17s+ with $\ge$5\% FS} \\
 	\end{tabular}
@@
 	\caption{Charts with the highest percentage of FSs among total steps.}
-\end{figure}
-\begin{figure}[t]
+\end{table}
+\begin{table}[t]
 	\begin{center}
 		\small
-	\begin{tabular}{r|l|l|c}
+	\begin{tabular}{r|l|l|l}
 		\bf Ft. & \bf Name & \bf Pack & \bf \#XF \\
@@
 		11 & Heart Shooter     & VocaJawnz                  & 44 \\
-		12 & Beautiful Synergy & Keyboard Collaboration III & 14 \\
+		12 & Beautiful Synergy & Keyboard Collab. III & 14 \\
 		13 & Toxic             & Sexuality Violation 2      & 12 \\
 		14 & Fancy Footwork    & Cirque du Zeppelin         & 40 \\
-		% 9 & Dr. Boom-Bombay   & fort rapids vii            & \bf 18 & 4.0 \\
-		%10 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.5 \\
-		%11 & Heart Shooter     & VocaJawnz                  & \bf 44 & 5.1 \\
-		%12 & Beautiful Synergy & Keyboard Collaboration III & \bf 14 & 3.2 \\
-		%13 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.2 \\
-		%14 & Fancy Footwork    & Cirque du Zeppelin         & \bf 40 & 4.2 \\
+		15 & yoshikawa45 vs siesta45 & Rikame's Simfiles 4  & 20 \\
 		\multicolumn{4}{c}{\em no 15s+ with $\ge$12 XFs} \\
@@
 	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}
-\end{figure}
+\end{table}
+
+% authors
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|l}
+		\bf Author & \bf Charts & \bf Total Steps & \bf FS\% \\
+		\hline
+		Konami        & 530 & 114623 & 8.03 \\
+		sssmsm        &  41 &  17219 & 6.55 \\
+		NEMORIGINAL   &  44 &  20165 & 5.36 \\
+		M. Emirzian   &  23 &   9307 & 5.16 \\
+		J. DeGarmo    &  16 &   4362 & 4.86 \\
+		D. Renzetti   &  18 &   7877 & 4.47 \\
+		R. McKanna    &  47 &  14855 & 4.23 \\
+		M. Puls       &  26 &   8379 & 4.18 \\
+		King of Light &  24 &   8817 & 4.13 \\
+		D. Bernardone & 217 &  76290 & 4.10 \\
+		D. D'Amato    & 107 &  36265 & 3.92 \\
+		bblum         &  32 &  32382 & 3.76 \\
+		\multicolumn{4}{c}{\normalsize\dots} \\
+		B. Vergara    &  13 &  15454 & 0.091 \\
+		Aoreo         &  21 &  27990 & 0.089 \\
+		Zaia          & 368 & 448460 & 0.080 \\
+		t0ni          &  85 & 128964 & 0.058 \\
+		Burn          &  27 &  60052 & 0.057 \\
+		Dirk          &  12 &  21996 & 0.055 \\
+		Happy Feet    &  30 &  57888 & 0.040 \\
+		@@            &  63 & 199530 & 0.026 \\
+		Arvin         &  79 & 108612 & 0.023 \\
+		Drazu         & 153 & 221460 & 0.021 \\
+		teejusb       &  11 &  11298 & 0.018 \\
+		Fraxtil       &  19 &  25612 & 0 \\
+	\end{tabular}
+	\end{center}
+	\caption{Chart authors with the highest/lowest XO\%.}
+\end{table}
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|l}
+		\bf Author & \bf Charts & \bf Total Steps & \bf FS\% \\
+		\hline
+		Konami        & 530 & 114623 & 2.29 \\
+		bblum         &  32 &  32382 & 1.29 \\
+		M. Puls       &  26 &   8379 & 1.16 \\
+		R. McKanna    &  47 &  14855 & 1.11 \\
+		mudkyp        &  63 &  42792 & 1.09 \\
+		S. Venkat     &  24 &  11084 & 0.96 \\
+		K. Ward       & 281 &  86475 & 0.89 \\
+		xRGTMx        &  19 &  11734 & 0.88 \\
+		sssmsm        &  41 &  17219 & 0.85 \\
+		D. Bernardone & 217 &  76290 & 0.83 \\
+		ATB           &  31 &  20673 & 0.82 \\
+		Happy Feet    &  30 &  57888 & 0.82 \\
+		\multicolumn{4}{c}{\normalsize\dots} \\
+		Hsarus        &  18 &  70162 & 0.070 \\
+		Drazu         & 153 & 221460 & 0.057 \\
+		@@            &  63 & 199530 & 0.037 \\
+		Revolver      &  11 &   8302 & 0.036 \\
+		T. Swag       &  13 &  22909 & 0.031 \\
+		Dirk          &  12 &  21996 & 0.023 \\
+		B. Vergara    &  13 &  15454 & 0.019 \\
+		warpdr!ve     &  16 &  19090 & 0.016 \\
+		Burn          &  27 &  60052 & 0.015 \\
+		teejusb       &  11 &  11298 & 0.009 \\
+		t0ni          &  85 & 128964 & 0.002 \\
+		S. Tofu       &  26 &  34609 & 0
+	\end{tabular}
+	\end{center}
+	\caption{Chart authors with the highest/lowest FS\%.}
+\end{table}
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|l}
+		\bf Author & \bf Charts & \bf Total Steps & \bf JK\% \\
+		\hline
+		King of Light &  24 &   8817 & 12.94 \\
+		R. McKanna    &  47 &  14855 & 9.95 \\
+		Konami        & 530 & 114623 & 9.63 \\
+		P. Shanklin   &  21 &   9834 & 8.91 \\
+		M. Puls       &  26 &   8379 & 8.90 \\
+		K. Ward       & 281 &  86475 & 8.80 \\
+		ATB           &  31 &  20673 & 8.46 \\
+		J. DeGarmo    &  16 &   4362 & 8.30 \\
+		D. Bernardone & 217 &  76290 & 7.98 \\
+		Renard        &  45 &  15035 & 7.94 \\
+		C. Foy        & 133 &  52418 & 7.72 \\
+		Yoko          &  10 &   4128 & 7.63 \\
+		\multicolumn{4}{c}{\normalsize\dots} \\
+		bblum         &  32 &  32382 & 1.56 \\
+		B. Vergara    &  13 &  15454 & 1.31 \\
+		Arvin         &  79 & 108612 & 1.30 \\
+		Drazu         & 153 & 221460 & 1.06 \\
+		Zaia          & 368 & 448460 & 0.86 \\
+		Aoreo         &  21 &  27990 & 0.78 \\
+		T. Swag       &  13 &  22909 & 0.70 \\
+		@@            &  63 & 199530 & 0.61 \\
+		warpdrive     &  16 &  19090 & 0.49 \\
+		Dirk          &  12 &  21996 & 0.35 \\
+		Burn          &  27 &  60052 & 0.28 \\
+		Hsarus        &  18 &  70162 & 0.27 \\
+	\end{tabular}
+	\end{center}
+	\caption{Chart authors with the highest/lowest JK\%.}
+\end{table}
+
+% packs
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|l}
+		\bf Pack & \bf Charts & \bf Total Steps & \bf XO\% \\
+		\hline
+		DDR 1st Mix to Extreme   &  530 & 114623 & 8.03 \\
+		r2112                    &   47 &  18377 & 4.50 \\
+		the best of r21freak     &  100 &  45156 & 4.22 \\
+		the best of r21freak ii  &   48 &  25024 & 4.16 \\
+		In The Groove 2          &  222 &  66113 & 4.05 \\
+		In The Groove Rebirth+   &  108 &  43758 & 3.99 \\
+		r21twins                 &   52 &  22347 & 3.89 \\
+		In The Groove 3          &  320 & 106363 & 3.61 \\
+		CuoReNeRo MeGaPacK       &  423 & 248625 & 3.45 \\
+		In The Groove            & 1408 & 491819 & 3.27 \\
+		r21freak Friendship Pack &   47 &  20363 & 3.13 \\
+		BemaniBeats 4            &   31 &  18464 & 3.02 \\
+		\multicolumn{4}{c}{\normalsize\dots} \\
+		Tachyon Epsilon          &  150 & 208830 & 0.064 \\
+		SPEEEDCOOOORE 4          &  101 & 123814 & 0.064 \\
+		TranceMania              &   80 & 121415 & 0.062 \\
+		Cirque du Lykan          &  129 & 160312 & 0.059 \\
+		Cirque du Zonda          &   45 &  74890 & 0.057 \\
+		Jimmy Jawns              &  109 & 170894 & 0.044 \\
+		Getty                    &   26 &  53528 & 0.043 \\
+		Tachyon Delta            &   32 &  36712 & 0.038 \\
+		Tachyon Gamma            &   32 &  36134 & 0.033 \\
+		Oh Henry! Mad Stamina    &   46 & 152326 & 0.028 \\
+		Causality Violation      &   10 &  19507 & 0.021 \\
+		Fast Track to Brutetown  &   29 &  46082 & 0.020 \\
+	\end{tabular}
+	\end{center}
+	\caption{Packs with the highest/lowest XO\%.}
+\end{table}
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|l}
+		\bf Pack & \bf Charts & \bf Total Steps & \bf FS\% \\
+		\hline
+		Subluminal               &   17 &  13418 & 4.70 \\
+		Aoreo's Ariginals 2      &   16 &  15222 & 2.42 \\
+		DDR 1st Mix to Extreme   &  530 & 114623 & 2.29 \\
+		Aoreo's Ariginals        &   31 &  26418 & 2.26 \\
+		rocky mount xi           &  113 &  79986 & 0.97 \\
+		In The Groove 2          &  222 &  66113 & 0.93 \\
+		FA and Chill             &   35 &  22045 & 0.86 \\
+		Getty                    &   26 &  53528 & 0.85 \\
+		Undertale                &   19 &  14722 & 0.84 \\
+		r2112                    &   47 &  18377 & 0.82 \\
+		Fort Rapids VI           &   75 &  75563 & 0.77 \\
+		Mute Sims 8              &   72 &  47140 & 0.71 \\
+		\multicolumn{4}{c}{\normalsize\dots} \\
+		SPEEEDCOOOORE 4          &  101 & 123814 & 0.093 \\
+		Stamina Showcase         &   38 & 126146 & 0.088 \\
+		VocaJawnz II             &  128 & 183153 & 0.084 \\
+		Cirque du Zeppelin       &  109 & 102991 & 0.082 \\
+		SPEEEDCOOOORE 3          &   66 &  69236 & 0.071 \\
+		Causality Violation      &   10 &  19507 & 0.051 \\
+		TranceNation             &   41 & 123940 & 0.047 \\
+		Oh Henry! Mad Stamina    &   46 & 152326 & 0.046 \\
+		Tachyon Epsilon          &  150 & 208830 & 0.040 \\
+		Noisiastreamz            &   20 &  41917 & 0.019 \\
+		TranceMania 2            &   40 &  64109 & 0.003 \\
+		TranceMania              &   80 & 121415 & 0.001 \\
+	\end{tabular}
+	\end{center}
+	\caption{Packs with the highest/lowest FS\%.}
+\end{table}
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|l}
+		\bf Pack & \bf Charts & \bf Total Steps & \bf JK\% \\
+		\hline
+		DDR 1st Mix to Extreme   &  530 & 114623 & 9.63 \\
+		r2112                    &   47 &  18377 & 8.80 \\
+		In The Groove 2          &  222 &  66113 & 8.64 \\
+		Gensokyo Holiday         &   87 &  51652 & 7.87 \\
+		r21Freak's Friendship Pack 2 &   32 &  15649 & 7.71 \\
+		Omnifarious              &   10 &   5594 & 7.53 \\
+		r21twins                 &   52 &  22347 & 7.18 \\
+		Piece of Cake 7          &   20 &  11554 & 6.97 \\
+		In The Groove            & 1408 & 491819 & 6.97 \\
+		In The Groove Rebirth+   &  108 &  43758 & 6.97 \\
+		TLOES Chapter 1          &   85 &  42201 & 6.89 \\
+		ITG Rebirth 2 Beta       &  262 &  99078 & 6.79 \\
+		\multicolumn{4}{c}{\normalsize\dots} \\
+		TranceMania 2            &   40 &  64109 & 1.03 \\
+		Causality Violation      &   10 &  19507 & 0.98 \\
+		VocaJawnz II             &  128 & 183153 & 0.98 \\
+		Tachyon Epsilon          &  150 & 208830 & 0.94 \\
+		Tachyon Delta            &   32 &  36712 & 0.87 \\
+		Cirque du Lykan          &  129 & 160312 & 0.87 \\
+		Cirque du Veyron         &   31 &  51545 & 0.79 \\
+		Cirque du Zeppelin       &  109 & 102991 & 0.77 \\
+		Oh Henry! Mad Stamina    &   46 & 152326 & 0.67 \\
+		Stamina Showcase         &   38 & 126146 & 0.56 \\
+		Cirque du Zonda          &   45 &  74890 & 0.47 \\
+		TranceNation             &   41 & 123940 & 0.25 \\
+	\end{tabular}
+	\end{center}
+	\caption{Packs with the highest/lowest JK\%.}
+\end{table}
 

### 2017-03-16 itg/2/paper.tex
finish the paper

@@
 ``Spin-switches'' with $D$ facing are also theoretically possible,
-arising from patterns such as $LURDDL$ or $LDRUUL$, although I know of no real-world chart which does this (hmm, tempting...).
+arising from patterns such as $LURDDL$ or $LDRUUL$,
+%although I know of no real-world chart which does this (hmm, tempting...).
+or similarly, ``270-switches'', as shown in Figure~\ref{fig:real-world-turniness}(a).
 
@@
 a new example of which is shown in Figure~\ref{fig:fs-turniness}(c).
+Figure~\ref{fig:real-world-turniness}(b) also shows a real-world chart exhibiting this pattern.
 
@@
 so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
-Anyway, remember that all-candle switch stream in (c). We'll see it again soon.
+%Anyway, remember that all-candle switch stream in (c). We'll see it again soon.
 
@@
 
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{web33-dd-switch-270.png}
+		\\
+		(a) Web33,260.8 \\
+		(12, Rikame 5)
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{fuego-candle-fs.png}
+		\\
+		(b) Fuego \\
+		(12, Best of Gazebo)
+	\end{tabular}
+	\end{center}
+	\caption{Real-world examples of turny footswitches.}
+	\label{fig:real-world-turniness}
+\end{figure}
+
 % dont forget
@@
 \section{Analyzing Crossoveriness}
+
+\newcommand\hilight[2]{\color{#1}#2\color{black}}
+\definecolor{pink}{RGB}{128,0,192}
+\definecolor{orange}{RGB}{192,96,0}
+\definecolor{olivegreen}{RGB}{0,127,32}
+\definecolor{brickred}{RGB}{192,0,0}
+\definecolor{commentblue}{RGB}{0,128,192}
+
+\begin{figure*}[ht]
+\begin{tabular}{l}
+\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step}~= \hilight{brickred}{L}~| \hilight{brickred}{D}~| \hilight{brickred}{U}~| \hilight{brickred}{R}~| \hilight{brickred}{Jump}~\hilight{orange}{deriving}~\hilight{olivegreen}{Eq}} \\
+\texttt{} \\
+\texttt{\hilight{orange}{data}~\hilight{olivegreen}{AnalysisState}~= \hilight{brickred}{S}~\{ steps :: \hilight{olivegreen}{Int}, xovers :: \hilight{olivegreen}{Int}, switches :: \hilight{olivegreen}{Int}, jacks :: \hilight{olivegreen}{Int},} \\
+\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastStep :: \hilight{olivegreen}{Maybe}~\hilight{olivegreen}{Step}, doubleStep :: \hilight{olivegreen}{Bool}, lastFlip :: \hilight{olivegreen}{Bool},} \\
+\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsLR :: [\hilight{olivegreen}{Bool}] \}} \\
+\texttt{} \\
+\texttt{\hilight{pink}{commitStream}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{AnalysisState}} \\
+\texttt{\hilight{pink}{commitStream}~s = s \{ xovers~~ = xovers~~ s + \hilight{orange}{if}~f \hilight{orange}{then}~ns - nx \hilight{orange}{else}~nx, } \\
+\texttt{~~~~~~~~~~~~~~~~~~~~ switches = switches s + \hilight{orange}{fromEnum}~(f == lastFlip s \&\& doubleStep s), } \\
+\texttt{~~~~~~~~~~~~~~~~~~~~ jacks~~~~= jacks~~~~s + \hilight{orange}{fromEnum}~(f /= lastFlip s \&\& doubleStep s), } \\
+\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsLR = \hilight{brickred}{[]}~\}} \\
+\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~stepsLR s} \\
+\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsLR s} \\
+%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
+%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
+%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
+%\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
+\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- reverse the stream's footing if more L/R steps were crossed over than not.}} \\
+\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((switches s > jacks s) == lastFlip s)} \\
+\texttt{} \\
+\texttt{\hilight{pink}{analyzeStep}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{Step}~-> \hilight{olivegreen}{AnalysisState}} \\
+\texttt{\hilight{pink}{analyzeStep}~s step} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\
+\texttt{~~~~| step == \hilight{brickred}{Jump}~= (\hilight{pink}{commitStream}~s) \{ lastStep = \hilight{brickred}{Nothing}, doubleStep = \hilight{brickred}{False}~\}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\
+\texttt{~~~~| lastStep s == \hilight{brickred}{Just}~step = stream (\hilight{pink}{commitStream}~s) \{ doubleStep = \hilight{brickred}{True}~\}} \\
+%\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\
+\texttt{~~~~| \hilight{orange}{otherwise}~= stream s} \\
+\texttt{~~~~\hilight{orange}{where}~foot = \hilight{orange}{not}~\$ lastFoot s} \\
+\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or crossed-over L/R arrow.}} \\
+\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = steps ++ [ft]} \\
+\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = steps ++ [\hilight{orange}{not}~ft]} \\
+\texttt{~~~~~~~~~~addStep ft \_ steps = steps \hilight{commentblue}{-{}- U/D don't help to determine L/R footing.}} \\
+\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step, lastFoot = foot,} \\
+\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ stepsLR = addStep foot step \$ stepsLR s \} } \\
+\texttt{} \\
+\texttt{\hilight{pink}{analyze}~:: [\hilight{olivegreen}{Step}] -> \hilight{olivegreen}{AnalysisState}} \\
+\texttt{\hilight{pink}{analyze}~= \hilight{pink}{commitStream}~.
+\hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[]}) } \\
+\end{tabular}
+	\caption{Pseudocode description of the crossoveriness and footswitchiness algorithm.}
+\label{fig:pseudocode}
+\end{figure*}
 
@@
 \end{itemize}
-
-\newcommand\hilight[2]{\color{#1}#2\color{black}}
-\definecolor{pink}{RGB}{128,0,192}
-\definecolor{orange}{RGB}{192,96,0}
-\definecolor{olivegreen}{RGB}{0,127,32}
-\definecolor{brickred}{RGB}{192,0,0}
-\definecolor{commentblue}{RGB}{0,128,192}
-
-\begin{figure*}[p]
-\begin{center}
-\begin{tabular}{l}
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step}~= \hilight{brickred}{L}~| \hilight{brickred}{D}~| \hilight{brickred}{U}~| \hilight{brickred}{R}~| \hilight{brickred}{Jump}~\hilight{orange}{deriving}~\hilight{olivegreen}{Eq}} \\
-\texttt{} \\
-\texttt{\hilight{orange}{data}~\hilight{olivegreen}{AnalysisState}~= \hilight{brickred}{S}~\{ steps :: \hilight{olivegreen}{Int}, xovers :: \hilight{olivegreen}{Int}, switches :: \hilight{olivegreen}{Int}, jacks :: \hilight{olivegreen}{Int},} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastStep :: \hilight{olivegreen}{Maybe}~\hilight{olivegreen}{Step}, doubleStep :: \hilight{olivegreen}{Bool}, lastFlip :: \hilight{olivegreen}{Bool},} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsLR :: [\hilight{olivegreen}{Bool}] \}} \\
-\texttt{} \\
-\texttt{\hilight{pink}{commitStream}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{AnalysisState}} \\
-\texttt{\hilight{pink}{commitStream}~s = s \{ xovers~~ = xovers~~ s + \hilight{orange}{if}~f \hilight{orange}{then}~ns - nx \hilight{orange}{else}~nx, } \\
-\texttt{~~~~~~~~~~~~~~~~~~~~ switches = switches s + \hilight{orange}{fromEnum}~(f == lastFlip s \&\& doubleStep s), } \\
-\texttt{~~~~~~~~~~~~~~~~~~~~ jacks~~~~= jacks~~~~s + \hilight{orange}{fromEnum}~(f /= lastFlip s \&\& doubleStep s), } \\
-	\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsLR = \hilight{brickred}{[]}~\}} \\
-\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~stepsLR s} \\
-\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsLR s} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
-\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((switches s > jacks s) == lastFlip s)} \\
-\texttt{} \\
-\texttt{\hilight{pink}{analyzeStep}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{Step}~-> \hilight{olivegreen}{AnalysisState}} \\
-\texttt{\hilight{pink}{analyzeStep}~s step} \\
-\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\
-\texttt{~~~~| step == \hilight{brickred}{Jump}~= (\hilight{pink}{commitStream}~s) \{ lastStep = \hilight{brickred}{Nothing}, doubleStep = \hilight{brickred}{False}~\}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\
-\texttt{~~~~| lastStep s == \hilight{brickred}{Just}~step = stream (\hilight{pink}{commitStream}~s) \{ doubleStep = \hilight{brickred}{True}~\}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\
-	\texttt{~~~~| \hilight{orange}{otherwise}~= stream s} \\
-\texttt{~~~~\hilight{orange}{where}~foot = \hilight{orange}{not}~\$ lastFoot s} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or crossed-over L/R arrow.}} \\
-\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = steps ++ [ft]} \\
-\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = steps ++ [\hilight{orange}{not}~ft]} \\
-\texttt{~~~~~~~~~~addStep ft \_ steps = steps \hilight{commentblue}{-{}- U/D don't help to determine L/R footing.}} \\
-\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step, lastFoot = foot,} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ stepsLR = addStep foot step \$ stepsLR s \} } \\
-\texttt{} \\
-\texttt{\hilight{pink}{analyze}~:: [\hilight{olivegreen}{Step}] -> \hilight{olivegreen}{AnalysisState}} \\
-	\texttt{\hilight{pink}{analyze}~= \hilight{pink}{commitStream}~.
-\hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[]}) } \\
-\end{tabular}
-\end{center}
-	\caption{Pseudocode description of the crossoveriness and footswitchiness algorithm.}
-\label{fig:pseudocode}
-\end{figure*}
 
@@
 	\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{paradise-lost-false-positive.png}
+		\includegraphics[width=0.16\textwidth]{paradise-lost-false-positive.png}
 		\\
@@
 	\end{tabular}
-	%\begin{tabular}{c}
-	%	\includegraphics[width=0.18\textwidth]{false-positive-toromi-hearts.png}
-	%	\\
-	%	(a) Toromi Hearts 2 \\
-	%	(18, Cirque du Zonda)
-	%\end{tabular}
 	\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{hearbeat-rollstream-doublestep.png}
+		\includegraphics[width=0.16\textwidth]{hearbeat-rollstream-doublestep.png}
 		\\
@@
 \newcommand\percentflip{\ensuremath{\%_{\mathfrak{flip}}}}
-\begin{algorithm}[h!]
+\begin{algorithm}[h]
 	\SetKwInOut{Input}{Input}
@@
 
-\section{Experimental Results}
+\section{Evaluation}
 
-% TODO: Put a * on charts where you're going to put a pict of mario man
-% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended
+Our experimental corpus has grown considerably since last year, and now comprises 11,666 stepcharts.
+I ran the crossoveriness/etcetera algorithm on all of them,
+and counted the total steps (not including jumps), crossovers, footswitches, jacks, forced doublesteps, and crossover switches for each.
+I also grouped the charts by author and by song pack to calculate each author's/pack's overall crossoveriness/etc.
+You can view the entire dataset at \url{http://tinyurl.com/crossoveriest}.
 
-% TODO: Make a special mention of sandstorm and put a pict of mario man
-% TODO: put the above in an honourable mention subsection, and also mention the slow train and sidefoots
+Tables
+\ref{tab:chart-xo},
+\ref{tab:chart-xop},
+\ref{tab:chart-fs},
+\ref{tab:chart-fsp},
+\ref{tab:chart-xf},
+\ref{tab:author-xo},
+\ref{tab:author-fs},
+\ref{tab:author-jk},
+\ref{tab:pack-xo},
+\ref{tab:pack-fs},
+and
+\ref{tab:pack-jk}
+summarize the dataset as ``leaderboards'' for each category.
+They present the data the same way as last year \cite{turniness},
+so I won't explain it again.
+Suffice to say that ITG enthusiasts should use their personal chart style preferences to navigate these tables and find song or pack recommendations or feel proud of themselves or whatever.
 
-% Stuff to talk about.
-% Dataset size now >10k
-% I reran turniness on fancy footwork and it actually got turnier (double check dis), b/c of all the candles
-% 9-15 is clearly the sweet spot for techni jawnz
+In the by-author analysis, I excluded authors with fewer than 10 charts, like last time.
+Also, in by-author and by-pack, I excluded 1-, 2-, and 3-foot charts,
+on the pretense that they often ignore the alternating feet assumption
+(though I also admit this biases the analysis against DDR/Konami).
+%(though I also admit some bias in wanting to unseat DDR/Konami as the crossover king).
+Nevertheless, DDR charts generally ruled the single-digits,
+even in footswitchiness,
+showing perhaps more technical depth than I gave them credit for.
+%Also, congratulations to Fraxtil and S. Tofu, who managed to write not a single crossover or footswitch, respectively.
 
-% By author/by pack: I excluded 1s-3s, admittedly biased against konami, on the pretense of excluding nonsense false positives, but even after that, konami is still the crossover king
-% Step authors are limited to 10 or more charts
-
-% Miscellaneous charts I want to exhibit:
-% Conflict forced doublesteps
-% Heart shooter 'sidefoots'
-% Boom Bombay - triple xover jack footed as sideswitch
-% Toxic - jack+doublestep footed as sideswitch + xover stream instead
-% H'ble mention to sandstorm for jacks
+By the way, the theoretical maxima for XO\%, FS\%, and JK\% are 50-$\epsilon$, 100-$\epsilon$, and 100-$\epsilon$, respectively \cite{epsilon}.
 
@@
 		 6 & Autoload                 & ITG 3                 &  54 \\
-		 7 & Tetris                   & CuoReNeRo             &  58 \\
-		 8 & Pulse                    & CuoReNeRo             &  94 \\
-		 9 & MAX Forever              & CuoReNeRo             & 107 \\
-		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo             & 141 \\
+		 7 & Tetris                   & CuoReNeRo M'PacK      &  58 \\
+		 8 & Pulse                    & CuoReNeRo M'PacK      &  94 \\
+		 9 & MAX Forever              & CuoReNeRo M'PacK      & 107 \\
+		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo M'PacK      & 141 \\
 		11 & Somebody I Used To Know  & Best of Gazebo        & 114 \\
@@
 	\caption{Charts with the most total crossovers (XOs).}
+	\label{tab:chart-xo}
 \end{table}
@@
 		15 & yoshikawa45 vs siesta45  & Rikame's Simfiles 4   & 10.8 \\
-		\multicolumn{4}{c}{\em no 16s-19s with $\ge$10\% XOs} \\
+		\multicolumn{4}{c}{\em no 16s-19s with $\ge$10\% XO} \\
 		20 & Rainbow Dimension        & Rikame's Simfiles 2   & 10.7 \\
@@
 	\caption{Charts with the highest percentage of XOs among total steps.}
+	\label{tab:chart-xop}
 \end{table}
@@
 		 7 & Silent Hill      & DDR 3rdMIX            &  29 \\
-		 8 & Pulse            & CuoReNeRo             &  41 \\
+		 8 & Pulse            & CuoReNeRo M'PacK      &  41 \\
 		 9 & Dr. Boom-Bombay  & fort rapids vii       &  34 \\
@@
 	\caption{Charts with the most total footswitches (FSs).}
+	\label{tab:chart-fs}
 \end{table}
@@
 	\end{center}
-	\caption{Charts with the highest percentage of FSs among total steps.}
+	\caption{Charts with the highest percentage of FSs.} %among total steps.}
+	\label{tab:chart-fsp}
 \end{table}
-\begin{table}[t]
+\begin{table}[h!]
 	\begin{center}
@@
 		11 & Heart Shooter     & VocaJawnz                  & 44 \\
-		12 & Beautiful Synergy & Keyboard Collab. III & 14 \\
+		12 & Web 33,260.8      & Rikame's Simfiles 5        & 16 \\
 		13 & Toxic             & Sexuality Violation 2      & 12 \\
@@
 	\end{center}
-	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}
+	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off to exclude a bunch of ambiguously-patterned charts from DDR.)}
+	%because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}
+	\label{tab:chart-xf}
 \end{table}
@@
 % authors
-\begin{table}[t]
+\begin{table}[p]
 	\begin{center}
@@
 	\caption{Chart authors with the highest/lowest XO\%.}
+	\label{tab:author-xo}
 \end{table}
-\begin{table}[t]
+\begin{table}[p]
 	\begin{center}
@@
 	\caption{Chart authors with the highest/lowest FS\%.}
+	\label{tab:author-fs}
 \end{table}
-\begin{table}[t]
+\begin{table}[p]
 	\begin{center}
@@
 	\caption{Chart authors with the highest/lowest JK\%.}
+	\label{tab:author-jk}
 \end{table}
@@
 % packs
-\begin{table}[t]
+\begin{table}[p]
 	\begin{center}
@@
 	\caption{Packs with the highest/lowest XO\%.}
+	\label{tab:pack-xo}
 \end{table}
-\begin{table}[t]
+\begin{table}[p]
 	\begin{center}
@@
 	\caption{Packs with the highest/lowest FS\%.}
+	\label{tab:pack-fs}
 \end{table}
-\begin{table}[t]
+\begin{table}[p]
 	\begin{center}
@@
 	\caption{Packs with the highest/lowest JK\%.}
+	\label{tab:pack-jk}
 \end{table}
@@
 
+\section{Discussion}
+
+To verify the algorithm's accuracy,
+I manually inspected a random (read: not random) sample of the charts at or near the top of the various leaderboards
+(read: I played a lot of Stepmania).
+I also consulted a leading expert in the field of automated ITG chart analysis (read: myself),
+who reported that the algorithm is infinitely more accurate than the prior state-of-the-art.
+
+Honestly though, it works really well.
+Last year's algorithm was often finicky and prone to all sorts of false-positives,
+while this one plays ITG in a recognizably human way (almost) without fail.
+It was a joy to use.
+%Certain rare patterns could still confuse it, while on others...
+
+{\bf Surprises.}
+On occasion, the algorithm surprised me by stepping with crossover footswitches which,
+at first glance, I would probably jack or double-step.
+However, these always proved to be perfectly valid alternative footings,
+in some cases requiring considerable look-ahead.
+In Figure~\ref{fig:surprise-sideswitch}(a),
+the chart repeats $L$ (later $R$) thrice,
+beginning with the right (later, left) foot.
+While a human player would jack these repeated arrows,
+the crossoveriness algorithm performs a double-footswitch,
+effectively reducing the total crossover steps by 1 each time.
+In Figure~\ref{fig:surprise-sideswitch}(b),
+a mine cues the player to double-step with her right foot,
+but the crossoveriness algorithm can begin this section already crossed-over,
+owing to an earlier $L$ jack on which it could switch feet.
+
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{boom-bombay-double-sideswitch-instead-of-jack.png}
+		\\
+		(a) Dr. Boom-Bombay \\
+		(9, fort rapids vii)
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{toxic-xoverfs-avoid-doublestep.png}
+		\\
+		(b) Toxic \\
+		(10/13, Sex'y Violation 2)
+	\end{tabular}
+	\end{center}
+	\caption{Sometimes the algorithm was smarter than me.}
+	\label{fig:surprise-sideswitch}
+\end{figure}
+
+{\bf Honourable Mentions.}
+I omitted a table for the jackiest charts,
+on account of most of them being either trivial beginner charts or extra-long megamixes.
+One deserves a special mention:
+Sandstorm (Jimmy Jawnz 2), shown in Figure~\ref{fig:hble-mention}(a),
+has more than twice as many total jacks as the next jackiest chart,
+clocking in at 1049 (78.5\%) with its 15 and 992 (69.6\%) with its 17.
+And looking at that chart, can't you just hear Sandstorm playing in your head already?
+
+I also wanted to highlight the chart with the most crossover switches,
+shown in Figure~\ref{fig:hble-mention}(b),
+mostly because the skittle notes should add some nice variety of colour to the paper
+(with apologies to the dead-tree SIGBOVIK audience reading in greyscale).
+Figure~\ref{fig:hble-mention}(c), with 2nd place in crossover switches following (b),
+comes with an edit chart titled ``no sidefoots'',
+and to be perfectly honest I just kept saying the word ``sidefoots'' to myself and giggling a lot while writing this paper.
+
+\begin{figure*}[t]
+	\begin{center}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{sandstorm-jacks.png}
+		\\
+		(a) Sandstorm \\
+		(15/17, J. Jawnz II)
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{heart-shooter-xover-fs.png}
+		\\
+		(b) Heart Shooter \\
+		(11, VocaJawnz)
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{slow-train-sidefoots.png}
+		\\
+		(c) '76 (Slow Train) \\
+		(11, Mute Sims X)
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{conflict-cued-doublesteps.png}
+		\\
+		(d) Conflict \\
+		(12, Stephcharts)
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{matador-bracket-switch.png}
+		\\
+		(e) Matador \\
+		(11, Valex 8)
+	\end{tabular}
+	\end{center}
+	\caption{Miscellaneous interesting charts I discovered while browsing the giant spreadsheet.}
+	\label{fig:hble-mention}
+\end{figure*}
+
+{\bf Sweet spot.}
+Finally, in case it wasn't obvious in the tables, I'll point out that 9-15 is clearly the sweet spot of difficulties for technical stepcharts.
+
 \section{Never Work}
 
-Let's be honest: this isn't gonna be a trilogy.
+Let's be honest: this isn't gonna be a paper trilogy.
+Okay, with that said, here are some things that would be cool to implement in a fantasy universe with infinite free time.
+(I have renamed this ``future'' work section accordingly.)
 
-With that out of the way, here are some things that would be cool to implement in a fantasy universe with infinite free time.
+There are a few remaining cases the algorithm doesn't yet understand:
+\begin{itemize}
+	\item Doublesteps forced either by mine cues or by holds;
+	\item Crossover and/or bracket jumps, not usually forced but often way less turny than the alternative;
+	\item Forced footings across stream boundaries arising from bracket jumps or jump-footswitches.
+\end{itemize}
+For example, Figure~\ref{fig:hble-mention}(d) shows many sequential doublesteps, each forced by a mine cue,
+but which the algorithm interprets as spins because the flipped stream length falls below \nflip.
+Figure~\ref{fig:hble-mention}(e) shows an example of jump-footswitches
+which the algorithm fails to count because it ignores the footing of jumps.
 
-% TODO
-% forced doublesteps via holds
-% - bracketjump footswitches - matador from valex 8
-% understanding mine cues for doublesteps
+These patterns would all have to be identified heuristically.
+Apart from that being more work than I wanted to do,
+I also feel that adding too many heuristics to SIGBOVIK research
+compromises the simple and innocent beauty of an implementation unbound by the demands of mainstream conferences.
 

### 2018-03-11 htm/paper.tex
htm intro

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

%\documentclass[preprint]{sigplanconf}
\documentclass[10pt]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}
\usepackage[mathcal]{euscript}
\usepackage[linesnumbered,ruled]{algorithm2e}

\renewcommand{\UrlBreaks}{\do\/\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t\do\u\do\v\do\w\do\x\do\y\do\z}

% ____________________________________________________________
% Listings Package Configuration
% \usepackage[scaled]{beramono}

%\renewcommand*\ttdefault{txtt}
\usepackage[T1]{fontenc}

% This Deep Tex Voodoo is from
%   <http://www.latex-community.org/forum/viewtopic.php?f=5&t=2072>
% It's purpose is to make \lstinline normal size, without affecting
% \lstinputlisting.  It seems to work but I have no idea how or why,
% and I rather hope never to learn.
%\makeatletter
%\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
%\makeatother

\begin{document}

\conferenceinfo{SIGBOVIK '17}{Pittsburgh, PA, USA}
\copyrightyear{2017}
\copyrightdata{}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{
Transactional Memory Concurrency Verification with Landslide
%That `Boring' Stuff Was Part of the Title, BTW. \\
%So was that. And that, and this too. \\
%You got it all, right? \\
%Or Just, ``More Boring Crap about ITG'', for Short. \\
%Oh, That Was Also Part of the Title.
}
% \subtitle{\em The Randomly-Scoped Lambda Calculus}
% \subtitle{Subtitle Text, if any}

\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}

\maketitle

\begin{abstract}
Hardware transactional memory is a recently-introduced concurrent programming paradigm
%backed by a set of new instructions on recent x86 CPUs,
which allows programmers to elide locks for performance in low-contention workloads.
However,
%using this feature
it
comes at a cost in implementation complexity:
%programmers must support their
fast-path code must be accompanied by backup paths to handle transaction failure.

We extend Landslide, a popular stateless model checker,
with a concurrency model for transactional memory
and evaluate it on several real-world transactional
benchmarks and data structure implementations.
\end{abstract}

\category{D.1.3}{Programming Techniques}{Concurrent Programming}
\category{D.2.4}{Software Engineering}{Software/Program Verification}

\keywords
landslide terminal, baggage claim, ground transportation, ticketing

\section{Introduction}
Transactional Synchronization Extensions (TSX) \cite{transactional-memory}
is an instruction set extension for x86 CPUs which adds hardware-based transactional memory.
The processor uses its existing cache coherence algorithm to
check for memory conflicts with other threads while temporarily staging a sequence of memory accesses.
If no other CPU accesses the same memory during the transaction,
the access sequence is committed to main memory atomically (with respect to visibility to other CPUs).
Otherwise, the operations are discarded and the transaction returns a failure code.

This feature can be used to replace conventional locking in performance-critical concurrent programs.
When a concurrent workload largely accesses thread-local data,
or disjoint sections of a shared data structure,
the contention rate between threads is low,
and transactions will often succeed.
Compared to programs which use conventional locks,
which use bus-locking atomic accesses even in the fastest code path,
TSX provides substantial performance improvements in such programs \cite{htm-experience, htm-performance, htm-mario}.
However, the possibility of transaction failure introduces additional implementation complexity:
programmers must also provide a backup plan
to safely resolve contention between threads,
usually involving conventional synchronization.
These backup paths must coordinate not only with other backup paths
but also with other fast paths which another thread may begin after the original transaction failed,
which even in the simplest transactions requires complex synchronization sequences
\cite{htm-mario}.
This introduces an additional dimension of nondeterminism into an already concurrent program,
and moreover, because transaction failure is expected to be rare,
% だからこそ
obscure interleavings between failure paths are difficult to expose during testing.

This motivates the use of stateless model checking (MC) \cite{verisoft}
to comprehensively verify these transactional programs,
fast paths failure paths and all.
MC aims to force the system to execute all possible thread interleavings under a given test case,
exhaustively checking for bugs or verifying their absence in the corresponding state space.
Many such model checkers exist, of varying flavours of interleaving granularity, memory analysis, and types of programs checked
\cite{chess,dbug-ssv,spin,inspect,r4}.
This work builds upon Landslide \cite{landslide,quicksand},
a simulator-based checker which checks both user- and kernel-level programs at the granularity of individual memory accesses.
Our contributions are as follows:

\begin{enumerate}
	\item We extend Landslide's concurrency model to include transaction failure as an additional source of nondeterminism;
	\item We provide a proof sketch that our implementation matches HTM's execution semantics,
	\item We evaluate the extended Landslide on several transactional programs, analyzing both its bug-finding and verification performance.
\end{enumerate}

The paper is organized as follows.
Section \ref{sec:intro} introduces the problem domain and motivates our research.
The rest of the sections state the rest of the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Background}

TSX was supported on consumer hardware for the first time on Intel's Haswell architecture \cite{htm-haswell},
which provides the {\tt xbegin}, {\tt xend}, and {\tt xabort} instructions for beginning, committing, and aborting
transactions, respectively.

Under software transactional memory (STM) \cite{stm-pldi06},
memory conflicts with other threads and explicit aborts by the programmer are the only reason for transaction failure;
hence, depending on program semantics, some transactions may be guaranteed to succeed.
However, hardware implementations (HTM) may also fail transactions
for several other reasons such as random system interrupts or exhausting the CPU's cache capacity.
Because timer interrupts can in principle occur at any moment,
and with arbitrary frequency (observable by the program, perhaps as a result of a heavily-loaded system),
in this paper we will simplify the failure model by saying that HTM transactions can fail for any reason.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

Please accept my paper.
I worked hard on it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{abbrv}
\bibliography{citations}

\end{document}

### 2018-03-12 htm/paper.tex
more txn

@@
 \usepackage[usenames,dvipsnames]{xcolor}
-\usepackage[utopia]{mathdesign}
+\usepackage[charter]{mathdesign}
 \usepackage{natbib}
@@
 
-\conferenceinfo{SIGBOVIK '17}{Pittsburgh, PA, USA}
-\copyrightyear{2017}
+\conferenceinfo{SIGBOVIK '18}{Pittsburgh, PA, USA}
+\copyrightyear{2018}
 \copyrightdata{}
@@
 \section{Introduction}
+\label{sec:intro}
+
 Transactional Synchronization Extensions (TSX) \cite{transactional-memory}
@@
 The processor uses its existing cache coherence algorithm to
-check for memory conflicts with other threads while temporarily staging a sequence of memory accesses.
+check for memory conflicts with other cores while temporarily staging a sequence of memory accesses.
 If no other CPU accesses the same memory during the transaction,
 the access sequence is committed to main memory atomically (with respect to visibility to other CPUs).
-Otherwise, the operations are discarded and the transaction returns a failure code.
+Otherwise, the accesses are discarded, the CPU's local state is reverted, and the transaction returns a failure code.
 
 This feature can be used to replace conventional locking in performance-critical concurrent programs.
-When a concurrent workload largely accesses thread-local data,
+When a concurrent workload accesses largely thread-local data,
 or disjoint sections of a shared data structure,
@@
 exhaustively checking for bugs or verifying their absence in the corresponding state space.
-Many such model checkers exist, of varying flavours of interleaving granularity, memory analysis, and types of programs checked
-\cite{chess,dbug-ssv,spin,inspect,r4}.
+Many such model checkers exist, varying in
+interleaving granularity, memory analysis, types of programs checked, and search ordering strategy
+\cite{chess,dbug-ssv,spin,inspect,r4,portend,samc,mcr,randomized-scheduler}.
 This work builds upon Landslide \cite{landslide,quicksand},
-a simulator-based checker which checks both user- and kernel-level programs at the granularity of individual memory accesses.
+a simulator-based tester which checks both user- and kernel-level programs
+and incorporates data-race analysis \cite{tsan,fasttrack} to find new preemption points
+at memory access granularity.
 Our contributions are as follows:
@@
 	\item We extend Landslide's concurrency model to include transaction failure as an additional source of nondeterminism;
-	\item We provide a proof sketch that our implementation matches HTM's execution semantics,
+	\item We provide a proof sketch that our implementation matches TSX's execution semantics,
 	\item We evaluate the extended Landslide on several transactional programs, analyzing both its bug-finding and verification performance.
@@
 Section \ref{sec:intro} introduces the problem domain and motivates our research.
-The rest of the sections state the rest of the paper.
+The other sections state the rest of the paper.
 
@@
 
-TSX was supported on consumer hardware for the first time on Intel's Haswell architecture \cite{htm-haswell},
-which provides the {\tt xbegin}, {\tt xend}, and {\tt xabort} instructions for beginning, committing, and aborting
+{\bf Hardware transactional memory.}
+TSX was supported on consumer hardware for the first time by Intel's Haswell architecture \cite{htm-haswell},
+which extends the x86 instruction set to provide
+{\tt xbegin}, {\tt xend}, and {\tt xabort} for beginning, committing, and aborting
 transactions, respectively.
+Higher-level programming languages or compilers may offer libraries or intrinsics to access these instructions;
+for C and C++ GCC provides intrinsics named {\tt \_xbegin()} and so on \cite{htm-gcc}.
+Figure~\ref{fig:htm-example} shows an example program using TSX to synchronize access to a shared counter,
+including a failure path which defaults to a conventional lock.
+This example actually has a bug, which we will discuss in the next section;
+the reader is encouraged to try to spot it before then.
+Several related works formally prove the correctness of transactional memory {\em implementations}
+\cite{specifying-verifying-tm,tm-correctness,tm-completeness,mc-tm-with-spin},
+but verifying the client programs written to use transactions remains an open problem.
+
+\begin{figure}[t]
+	\begin{center}
+		\begin{tabular}{ll}
+		%\texttt{void count() \{} \\
+		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
+		1 & \texttt{if ((status = \_xbegin()) == SUCCESS) \{} \\
+		2 & \texttt{~~~~x++;} \\
+		3 & \texttt{~~~~\_xend();} \\
+		4 & \texttt{\} else \{} \\
+		5 & \texttt{~~~~mutex\_lock(\&m);} \\
+		6 & \texttt{~~~~x++;} \\
+		7 & \texttt{~~~~mutex\_unlock(\&m);} \\
+		8 & \texttt{\}} \\
+		%\texttt{~~~~\}} \\
+		%\texttt{\}} \\
+		\end{tabular}
+	\end{center}
+	\caption{Example transactional program.
+		%The example {\tt count} routine from Figure~\ref{fig:mutex}, rewritten to use HTM.
+		If the
+		%transaction in the
+		top branch aborts,
+		%whether from a memory conflict or random system interrupt,
+		%from the programmer's intention,
+		execution will revert to the return of {\tt \_xbegin()}
+		%{\tt status} will be assigned an error code
+		and control will drop into the {\tt else} branch.
+		The programmer can then use explicit synchronization
+		%, such as a mutex,
+		to resolve the conflict.}
+	\label{fig:htm-example}
+\end{figure}
+
+% TODO
 
 Under software transactional memory (STM) \cite{stm-pldi06},
-memory conflicts with other threads and explicit aborts by the programmer are the only reason for transaction failure;
+memory conflicts with other threads are the only reason for transaction failure
+(apart from programmer-supplied explicit aborts);
 hence, depending on program semantics, some transactions may be guaranteed to succeed.
@@
 in this paper we will simplify the failure model by saying that HTM transactions can fail for any reason.
+We defer discussion of programs which distinguish the reason for aborts through the failure code to Section~\ref{sec:warpzone}.
+
+{\bf Stateless model checking.}
+Model checking (MC) \cite{verisoft} is a testing technique for systematically executing and verifying
+the possible thread interleavings of a concurrent program.
+The main research challenge is to cope with exponential explosion of the state space,
+which is $O(n^k)$ for a program with $n$ operations and $k$ threads.
+Some {\em stateful} MCs explicitly store and compare all visited states of the program being tested \cite{spin},
+which both keeps track of test coverage
+and allows identifying identical states to avoid testing redundant interleavings.
+%
+By contrast, {\em stateless} MC (henceforth abbreviated simply as MC)
+stores only the current sequence of execution events to avoid a prohibitive memory footprint.
+Reduction algorithms \cite{dpor,optimal-dpor,satcheck,mcr,mcr,tsopso}
+can then analyze the memory accesses in that sequence to identify
+interleavings observationally-equivalent under Mazurkiewicz trace theory \cite{mazurkiewicz} and hence safe to skip.
+The resulting state spaces are still exponentially-sized,
+but only in the number of conflicting operations rather than all operations.
+Of these, Landslide uses Dynamic Partial Order Reduction \cite{dpor} to prune its state spaces.
+
+MCs may instrument programs to introduce thread switches at varying granularity, which affects the number of operations $n$.
+%The number of operations depends on how finely-grained the MC instruments the program to introduce thread switches.
+Some target distributed systems, instrumenting only message-passing events \cite{modist};
+some run multithreaded programs natively, instrumenting only the pthread API for performance \cite{dbug-ssv};
+and some insert compiler instrumentation on statically-identified memory accesses \cite{chess,inspect}.
+Landslide traces every memory access through the use of a simulated environment \cite{bochs},
+which is important for identifying data races to use as new preemption points \cite{fasttrack,djit,quicksand},
+as well as for identifying when a memory conflict may cause transaction aborts.
+With regard to checking for bugs,
+the ``model'' the name refers to being checked may be
+an external formal specification,
+the program's own internal consistency checks,
+or a set of expected properties encoded in the tool itself.
+Landslide uses the latter two cases, checking for assertion failures
+as well as deadlocks, use-after-frees, and segfaults.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Design}
+
+Let $I = TN_1@L_1, TN_2@L_2, ... TN_n@L_n$,
+with $N_i$ a thread ID and $L_i$ a code line number,
+denote the execution sequence of a program as it runs according to the specified thread interleaving.%
+\footnote{This serialization of concurrent execution is told from the perspective of main memory
+and hence assumes sequential consistency.
+For discussion of relaxed memory models refer to Section~\ref{sec:warpzone}.}
+
+{\bf Example.}
+Consider again the program in Figure~\ref{fig:htm-example}.
+Note that the C-style {\tt x++} operations, when compiled into assembly,
+results in multiple memory accesses which can be interleaved with other threads.
+\[
+\begin{tabular}{ll}
+	$2a$ & \texttt{temp <- x;} \\
+	$2b$ & \texttt{temp <- temp + 1;} \\
+	$2c$ & \texttt{x <- temp;} \\
+\end{tabular}
+\]
+
+\newcommand\hilight[2]{\color{#1}#2\color{black}}
+\definecolor{orange}{RGB}{192,96,0}
+\definecolor{olivegreen}{RGB}{0,127,0}
+\definecolor{brickred}{RGB}{192,0,0}
+\definecolor{commentblue}{RGB}{0,0,192}
+%\definecolor{lavender}{RGB}{238,85,238}
+%\definecolor{seafoam}{RGB}{85,238,102}
+%\definecolor{salmon}{RGB}{238,102,85} % ee6655
+
+\definecolor{lavender}{RGB}{178,65,178} % * 3/4
+\definecolor{seafoam}{RGB}{65,178,76} % * 3/4
+\definecolor{salmon}{RGB}{208,89,74} % * 7/8
+
+\newcommand\ti{\ensuremath{\hilight{lavender}{\mathbf{T1}}}}
+\newcommand\tj{\ensuremath{\hilight{seafoam}{\mathbf{T2}}}}
+\newcommand\tk{\ensuremath{\hilight{salmon}{\mathbf{T3}}}}
+
+If these instructions from the {\tt x++} in the transaction are preempted,
+with another thread's access to {\tt x} interleaved in between,
+the transaction will abort.
+So, the interleaving
+\[
+	\ti@1, \ti@2a, \ti@2b, \tj@1, \tj@2, \tj@3, \ti@2c, \ti@3
+\]
+or, henceforth abbreviated for clarity:
+\[
+	\ti@1-2b, \tj@1-3, \ti@2c-3
+\]
+is not possible; rather, \ti will fall into the backup path:
+\[
+	\ti@1-2b, \tj@1-3, \ti@4-7
+\]
+However, the {\tt x++} operation from the failure path (correspondingly $6a$, $6b$, $6c$)
+can be thusly separated with conflicting accesses interleaved in between,
+since the mutex only protects the failure path against other failure paths,
+but not against the transaction itself.
+So (assuming {\tt x} is intended to be a precise counter rather than a sloppy one),
+%losing one of the increments to which constitutes a bug),
+we observe a bug in the following interleaving.%
+\footnote{Note also that this bug requires either at least 3 threads or at least 2 iterations between 2 threads to expose;
+this highlights MC's dependence on its test cases to produce meaningful state spaces in the first place.}
+\[
+	\ti@1-2b, \tj@1-3, \ti@4-6b, \tk@1-3, \ti@6c-7
+\]
+Prior work \cite{htm-mario} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
+to exclude this family of interleavings,
+showing that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
+%further motivating our research.
+
+\begin{figure}[t]
+	\begin{center}
+		\begin{tabular}{ll}
+		%\texttt{void count() \{} \\
+		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
+		  & \texttt{prevent\_transactions = false;} \\
+		\\
+		0 & \texttt{while (prevent\_transactions) continue;} \\
+		1 & \texttt{if ((status = \_xbegin()) == SUCCESS) \{} \\
+		2 & \texttt{~~~~if (prevent\_transactions)} \\
+		3 & \texttt{~~~~~~~~\_xabort();} \\
+		4 & \texttt{~~~~x++;} \\
+		5 & \texttt{~~~~\_xend();} \\
+		6 & \texttt{\} else \{} \\
+		7 & \texttt{~~~~mutex\_lock(\&m);} \\
+		8 & \texttt{~~~~prevent\_transactions = true;} \\
+		9 & \texttt{~~~~x++;} \\
+		A & \texttt{~~~~prevent\_transactions = false;} \\
+		B & \texttt{~~~~mutex\_unlock(\&m);} \\
+		C & \texttt{\}} \\
+		%\texttt{~~~~\}} \\
+		%\texttt{\}} \\
+		\end{tabular}
+	\end{center}
+	\caption{Variant of the program in Figure~\ref{fig:htm-example},
+		with additional synchronization to protect the failure path from the transactional path.
+		(The optional line 0 serves to prevent a cascade of failure paths
+		for the sake of performance
+		by allowing threads to wait until transacting is safe again.)}
+	\label{fig:htm-fixed}
+\end{figure}
+
+{\bf Modeling transaction failure.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Evaluation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Limitations}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Future Work}
+\label{sec:warpzone}
 
@@
 
-Please accept my paper.
-I worked hard on it.
+I hope I can get away with citing this in my thesis.
 

### 2018-03-12 htm/paper.tex
txn structure deisgn section and write about tso/pso

@@
 \usepackage{natbib}
+\usepackage{xspace}
 \usepackage[mathcal]{euscript}
@@
 
-% TODO
-
 Under software transactional memory (STM) \cite{stm-pldi06},
@@
 but only in the number of conflicting operations rather than all operations.
-Of these, Landslide uses Dynamic Partial Order Reduction \cite{dpor} to prune its state spaces.
+Of these, Landslide uses Dynamic Partial Order Reduction (DPOR) \cite{dpor} to prune its state spaces.
 
@@
 \section{Design}
+\label{sec:design}
 
-Let $I = TN_1@L_1, TN_2@L_2, ... TN_n@L_n$,
+This section presents our formalization of transactional memory in Landslide's framework of thread concurrency.
+We make two major simplifications:
+simulating transaction aborts as failure injections,
+and treating transaction atomicity as a global mutex during data-race analysis;
+and provide corresponding equivalence proofs.
+
+{\bf Notation.} Let $I = TN_1@L_1, TN_2@L_2, ... TN_n@L_n$,
 with $N_i$ a thread ID and $L_i$ a code line number,
@@
 
-{\bf Example.}
+\subsection{Example}
+
 Consider again the program in Figure~\ref{fig:htm-example}.
@@
 
-\newcommand\ti{\ensuremath{\hilight{lavender}{\mathbf{T1}}}}
-\newcommand\tj{\ensuremath{\hilight{seafoam}{\mathbf{T2}}}}
-\newcommand\tk{\ensuremath{\hilight{salmon}{\mathbf{T3}}}}
+\newcommand\ti{\ensuremath{\hilight{lavender}{\mathbf{T1}}}\xspace}
+\newcommand\tj{\ensuremath{\hilight{seafoam}{\mathbf{T2}}}\xspace}
+\newcommand\tk{\ensuremath{\hilight{salmon}{\mathbf{T3}}}\xspace}
 
@@
 However, the {\tt x++} operation from the failure path (correspondingly $6a$, $6b$, $6c$)
-can be thusly separated with conflicting accesses interleaved in between,
+{\em can} be thusly separated with conflicting accesses interleaved in between,
 since the mutex only protects the failure path against other failure paths,
@@
 		with additional synchronization to protect the failure path from the transactional path.
-		(The optional line 0 serves to prevent a cascade of failure paths
+		The optional line 0 serves to prevent a cascade of failure paths
 		for the sake of performance
-		by allowing threads to wait until transacting is safe again.)}
+		by allowing threads to wait until transacting is safe again.}
 	\label{fig:htm-fixed}
@@
 
-{\bf Modeling transaction failure.}
+\subsection{Modeling Transaction Failure}
+
+Left unstated in interleavings such as $\ti@1-2c, \tj@1-3, \ti@4-7$%
+\footnote{For the sake of example I reorder \ti's write to {\tt x} before \tj's part here.}
+are HTM's execution semantics, namely:
+\begin{enumerate}
+	\item any modifications to shared state (such as $2c$) by \ti are not visible to \tj during its execution,
+		despite \tj being executed afterwards, and
+	\item all local and global state changes by \ti between lines $1$ and $2c$ are discarded when jumping to line $4$.
+\end{enumerate}
+
+\subsection{Modeling Transaction Atomicity}
 
@@
 \section{Limitations}
+\label{sec:warpzone}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+{\bf Relaxed memory orderings.}
+Our formalization of thread interleavings in Section~\ref{sec:design} does not account for read/write reorderings
+possible on relaxed memory consistency architectures \cite{memory-consistency-models}.
+In fact,
+even after \cite{htm-mario}'s proposed fix,
+our running example program is still incorrect on Total Store Ordering (TSO) architectures such as x86.
+Despite stores being totally-ordered, x86 may still reorder stores after subsequent loads.
+Accordingly, an execution of $8,9a,9b,9c$ in Figure~\ref{fig:htm-fixed}
+may be locally visible to another thread as $9a,8,9b,9c$,
+and hence an apparent interleaving of
+\[
+	\ti@1,\tj@1-5,\ti@7,\mathbf{\ti@9a},\tk@1-5,\mathbf{\ti@8},\ti@9b-B
+\]
+is possible
+(reordered accesses highlighted in bold).
+An {\tt mfence} barrier is needed between lines 8 and 9 to solve this problem on TSO \cite{tsx-need-barrier}.
+On Partial Store Ordering (PSO) architectures, even more barriers may be necessary.
 
-\section{Future Work}
-\label{sec:warpzone}
+Because Landslide's concurrency model includes only instruction-level thread nondeterminism,
+not per-CPU memory buffer reorderings,
+our current HTM implementation in Landslide cannot find this bug.
+In fact, it erroneously verifies the corresponding test {\tt htm2} in 7 CPU-minutes,
+with 294 distinct interleavings in total.
+Recent work extended DPOR to incorporate TSO and PSO memory nondeterminism \cite{tsopso};
+if incorporated into Landslide, we could find or verify the absence of such bugs.
+
+% TODO should this come 1st or 2nd
+{\bf Transaction failure codes.}
 

### 2018-03-12 htm/paper.tex
impl and xabort codes

@@
 
+\subsection{Implementation}
+
+Our implementation of HTM has been incorporated by the Landslide maintainers upstream.
+The repository is available open-source at \url{https://github.com/bblum/landslide}.
+Programs should be ported to the Pebbles userland \cite{thrlib,kspec},
+their use of compiler HTM intrinsics should be replaced with the Landslide stubs provided in {\tt 410user/inc/htm.h},
+and HTM nondeterminism can then be enabled with the {\tt -X} command-line flag.
+Several example tests are also provided:
+{\tt htm1} demonstrates the bug in Figure~\ref{fig:htm-example},
+and {\tt htm2} demonstrates Figure~\ref{fig:htm-fixed}'s fix.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+{\bf Transaction failure codes.}
+When a transaction fails, {\tt \_xbegin()} returns a failure code
+denoting the reason (or combination of reasons) therefor \cite{htm-gcc}.
+If a program then cases on that failure code to select between different backup paths,
+model checking it by simply injecting a single type of failure may be unsound.
+For example, a program executing a transaction which is guaranteed to never conflict with any other threads,
+and hence never abort without {\tt \_XABORT\_RETRY},
+could legally {\tt assert(false);} in its failure path,
+while our approach in Section~\ref{sec:design} would erroneously trigger that assertion and report a bug.
+Likewise, the transactional data structures from \cite{htm-mario}
+abstract away any spurious {\tt \_XABORT\_RETRY} aborts behind a retry loop;
+a MC unwise to that idiom would call it an infinite loop bug.
+
+Our HTM implementation in Landslide includes an experimental feature
+to track the set of abort codes possible for each transaction.
+It harnesses DPOR's existing memory analysis to identify when {\tt \_XABORT\_CONFLICT} is possible,
+and instruments {\tt \_xabort()} calls to record any user-supplied codes.
+This comes at a cost of even more state space explosion,
+increasing the exponent at each {\tt \_xbegin()} preemption point
+by (usually) 1 plus however many distinct {\tt \_xabort()} codes the program uses.
+Future work could use static or dynamic flow analysis to
+identify which transactions actually branch on their failure codes
+to prune equivalent backup paths;
+for now, this feature is disabled by default but accessible via the {\tt -A} command-line option
+to Landslide (in addition to {\tt -X}).
+
 {\bf Relaxed memory orderings.}
-Our formalization of thread interleavings in Section~\ref{sec:design} does not account for read/write reorderings
+Section~\ref{sec:design}'s formalization of thread interleavings does not account for read/write reorderings
 possible on relaxed memory consistency architectures \cite{memory-consistency-models}.
@@
 In fact, it erroneously verifies the corresponding test {\tt htm2} in 7 CPU-minutes,
-with 294 distinct interleavings in total.
+with 294 distinct interleavings in total,
+none of which include the above-listed sequence.
 Recent work extended DPOR to incorporate TSO and PSO memory nondeterminism \cite{tsopso};
 if incorporated into Landslide, we could find or verify the absence of such bugs.
-
-% TODO should this come 1st or 2nd
-{\bf Transaction failure codes.}
 

### 2018-03-12 htm/paper.tex
state first thm

@@
 \section{Background}
+
+This section introduces the fundamental concepts and prior work
+in both hardware transactional memory and stateless model checking,
+which we propose to combine.
 
@@
 \end{enumerate}
+% there shouldbe a better word than 'requires' here.. benefits from.. "the point of it is"..
+While use of TSX in production requires the performance advantage
+of temporarily staging such accesses in local CPU cache,
+model checking such programs need be concerned only with the program's {\em observable} behaviours.
+We claim that MCing the interleaving $\ti@1, \tj@1-3, \ti@4-7$
+is an equivalent verification as MCing the one above;
+in fact, this interleaving suffices to check
+all observable behaviours
+of all interleavings
+of all subsets of $\tj@1-3$
+with all subsets of $\ti@2a-2c$,
+whether they share a memory conflict or not.
+Stated formally:
+
+\renewcommand\proofname{Proof Sketch}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+
+\newcommand\tii{\ensuremath{\hilight{lavender}{\mathbf{Ti}}}\xspace}
+\newcommand\tjj{\ensuremath{\hilight{seafoam}{\mathbf{Tj}}}\xspace}
+\newcommand\tkk{\ensuremath{\hilight{salmon}{\mathbf{Tk}}}\xspace}
+
+\begin{theorem}[Equivalence of Aborts]
+	\label{thm:equiv}
+	Let:%
+	\footnote{Arbitrary code may not be structured to distinguish these as nicely as in our examples;
+	e.g., more code may exist in the success branch after {\tt \_xend()};
+	such would be considered part of $\omega$ here.}
+	\begin{itemize}
+		\item $\tii@\alpha$ be an HTM begin operation,
+		\item $\tii@\beta_1\dots\tii@\beta_n$ be the transaction body (with $\beta_n$ the HTM end call),
+		\item $\tii@\phi_1\dots\tii@\phi_m$ be the failure path, and
+		\item $\tii@\omega_1\dots\tii@\omega_l$ be the subsequent code executed unconditionally.
+	\end{itemize}
+	Then, for any interleaving prefix
+	\[
+		\tii@\alpha,\tii@\beta_1\dots\tii@\beta_b,\tjj@\gamma_1\dots\tjj@\gamma_j,\tkk@\kappa_1\dots\tkk@\kappa_k
+	\]
+	with $b<n$, $j \ne i$, $k \ne i$, etc., either:
+	\begin{enumerate}
+		\item $\tii@\alpha,\tii@\beta_1\dots\tii@\beta_b\dots\tii@\beta_n,\tjj@\gamma_1\dots\tjj@\gamma_j,$ \\
+			$\tkk@\kappa_1\dots\tkk@\kappa_k$
+			(independent case), or
+		\item $\tii@\alpha,\tjj@\gamma_1\dots\tjj@\gamma_j,\tkk@\kappa_1\dots\tkk@\kappa_k,\tii@\phi_1\dots$
+			(conflicting case)
+	\end{enumerate}
+	is observationally equivalent.
+\end{theorem}
+
+\begin{proof}
+	asdfasdf
+\end{proof}
 
 \subsection{Modeling Transaction Atomicity}
+
+Next, we address the memory accesses within transactions with regard to data-race analysis.
+From Theorem~\ref{thm:equiv} we have that the body of all transactions may be executed atomically within the MC environment.
+While they may interleave within other non-transactional sequences,
+no other operations (whether transactional or not) will interrupt them.
+We claim this level of atomicity is equivalent to that provided by a global lock,
+and hence abstracting it as such in Landslide's data-race analysis is sound.
+
+\begin{theorem}[Transactions are a Global Lock]
+	asdfasdfaf
+\end{theorem}
+
+\begin{proof}
+	asdfasdf
+\end{proof}
 
@@
 
+This section discusses opportunities for future work.
+
 {\bf Transaction failure codes.}
 When a transaction fails, {\tt \_xbegin()} returns a failure code
-denoting the reason (or combination of reasons) therefor \cite{htm-gcc}.
+denoting the reason, or combination thereof, therefor \cite{htm-gcc}.
 If a program then cases on that failure code to select between different backup paths,

### 2018-03-13 htm/paper.tex
add txxat macros and prove thm 1

@@
 \definecolor{commentblue}{RGB}{0,0,192}
+\definecolor{black}{RGB}{0,0,0}
 %\definecolor{lavender}{RGB}{238,85,238}
@@
 
+% T1 T2 T3 but latex doesn't allow #s in cmd names
 \newcommand\ti{\ensuremath{\hilight{lavender}{\mathbf{T1}}}\xspace}
@@
 \newcommand\tk{\ensuremath{\hilight{salmon}{\mathbf{T3}}}\xspace}
+
+\newcommand\tiat[1]{\ensuremath{\hilight{lavender}{\mathbf{T1}@#1}}\xspace}
+\newcommand\tjat[1]{\ensuremath{\hilight{seafoam}{\mathbf{T2}@#1}}\xspace}
+\newcommand\tkat[1]{\ensuremath{\hilight{salmon}{\mathbf{T3}@#1}}\xspace}
 
@@
 \[
-	\ti@1, \ti@2a, \ti@2b, \tj@1, \tj@2, \tj@3, \ti@2c, \ti@3
+	\tiat{1}, \tiat{2a}, \tiat{2b}, \tjat{1}, \tjat{2}, \tjat{3}, \tiat{2c}, \tiat{3}
 \]
@@
 \[
-	\ti@1-2b, \tj@1-3, \ti@2c-3
+	\tiat{1-2b}, \tjat{1-3}, \tiat{2c-3}
 \]
@@
 \[
-	\ti@1-2b, \tj@1-3, \ti@4-7
+	\tiat{1-2b}, \tjat{1-3}, \tiat{4-7}
 \]
@@
 \[
-	\ti@1-2b, \tj@1-3, \ti@4-6b, \tk@1-3, \ti@6c-7
+	\tiat{1-2b}, \tjat{1-3}, \tiat{4-6b}, \tkat{1-3}, \tiat{6c-7}
 \]
@@
 
-Left unstated in interleavings such as $\ti@1-2c, \tj@1-3, \ti@4-7$%
+Left unstated in interleavings such as $\tiat{1-2c}, \tjat{1-3}, \tiat{4-7}$%
 \footnote{For the sake of example I reorder \ti's write to {\tt x} before \tj's part here.}
@@
 model checking such programs need be concerned only with the program's {\em observable} behaviours.
-We claim that MCing the interleaving $\ti@1, \tj@1-3, \ti@4-7$
+We claim that MCing the interleaving $\tiat{1}, \tjat{1-3}, \tiat{4-7}$
 is an equivalent verification as MCing the one above;
@@
 of all interleavings
-of all subsets of $\tj@1-3$
-with all subsets of $\ti@2a-2c$,
+of all subsets of $\tjat{1-3}$
+with all subsets of $\tiat{2a-2c}$,
 whether they share a memory conflict or not.
@@
 
-\begin{theorem}[Equivalence of Aborts]
-	\label{thm:equiv}
+\newcommand\tiiat[1]{\ensuremath{\hilight{lavender}{\mathbf{Ti}@#1}}\xspace}
+\newcommand\tjjat[1]{\ensuremath{\hilight{seafoam} {\mathbf{Tj}@#1}}\xspace}
+\newcommand\tkkat[1]{\ensuremath{\hilight{salmon}  {\mathbf{Tk}@#1}}\xspace}
+
+\begin{lemma}[Equivalence of Aborts]
+	\label{lem:equiv}
 	Let:%
@@
 	\begin{itemize}
-		\item $\tii@\alpha$ be an HTM begin operation,
-		\item $\tii@\beta_1\dots\tii@\beta_n$ be the transaction body (with $\beta_n$ the HTM end call),
-		\item $\tii@\phi_1\dots\tii@\phi_m$ be the failure path, and
-		\item $\tii@\omega_1\dots\tii@\omega_l$ be the subsequent code executed unconditionally.
+		\item $\tiiat{\alpha}$ be an HTM begin operation,
+		\item $\tiiat{\beta_1}\dots\tiiat{\beta_n}$ be the transaction body (with $\beta_n$ the HTM end call),
+		\item $\tiiat{\phi_1}\dots\tiiat{\phi_m}$ be the failure path, and
+		\item $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.
 	\end{itemize}
-	Then, for any interleaving prefix
+	Then, for any interleaving prefix%
+	\footnote{Without loss of generality: for any number of other threads \tjj/\tkk,
+	and for any number of thread switches away from \tii during the transaction.}
 	\[
-		\tii@\alpha,\tii@\beta_1\dots\tii@\beta_b,\tjj@\gamma_1\dots\tjj@\gamma_j,\tkk@\kappa_1\dots\tkk@\kappa_k
+	\begin{tabular}{c}
+		$\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$\\
+		$\tjjat{\gamma_1}\dots\tjjat{\gamma_j},$ \\
+		$\tkkat{\kappa_1}\dots\tkkat{\kappa_k},$ \\
+		$\tiiat{\beta_{b+1}}$ % $\dots\tiiat{\beta_{n-1}}$ -- excluded bc might abort
+	\end{tabular}
 	\]
@@
 	\begin{enumerate}
-		\item $\tii@\alpha,\tii@\beta_1\dots\tii@\beta_b\dots\tii@\beta_n,\tjj@\gamma_1\dots\tjj@\gamma_j,$ \\
-			$\tkk@\kappa_1\dots\tkk@\kappa_k$
-			(independent case), or
-		\item $\tii@\alpha,\tjj@\gamma_1\dots\tjj@\gamma_j,\tkk@\kappa_1\dots\tkk@\kappa_k,\tii@\phi_1\dots$
-			(conflicting case)
+		\item $\tiiat{\alpha},\tjjat{\gamma_1}\dots\tjjat{\gamma_j},\tkkat{\kappa_1}\dots\tkkat{\kappa_k},\tiiat{\phi_1}\dots$
+			(conflicting case), or
+		\item $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b}\dots\tiiat{\beta_n},\tjjat{\gamma_1}\dots\tjjat{\gamma_j},$ \\
+			$\tkkat{\kappa_1}\dots\tkkat{\kappa_k}$
+			(independent case)
 	\end{enumerate}
 	is observationally equivalent.
+\end{lemma}
+
+\begin{proof}
+	We case on whether the operations by \tjj and/or \tkk have any memory conflicts (read/write or write/write)
+	with $\tiiat{\beta_1}\dots\tiiat{\beta_n}$.
+	If so, then the hardware will abort \tii's transaction, discarding the effects of $\tiiat{\beta_1}\dots\tiiat{\beta_n}$
+	and jumping to $\tiiat{\phi_1}$,
+	satisfying case 1.
+	Otherwise, by DPOR's definition of transition dependence \cite{dpor,landslide-phdthesis},
+	$\tiiat{\beta_{b+1}}\dots\tiiat{\beta_n}$ is independent with the transitions of \tjj and \tkk,
+	may be successfully executed until transaction commit,
+	and reordering them produces an equivalent interleaving,
+	satisfying case 2.
+\end{proof}
+
+The second part of our claim follows naturally.
+
+\begin{theorem}[Atomicity of Transactions]
+	\label{thm:atom}
+	For any state space $U$ of a transactionally-concurrent program,
+	an equivalent state space exists in which all transactions are either executed atomically or aborted immediately.
 \end{theorem}
@@
 \begin{proof}
-	asdfasdf
+	For every $I \in U$ with $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$ $\tjjat{\dots},\tkkat{\dots},\tiiat{\beta_{b+1}}... \in I$,
+	apply Lemma~\ref{lem:equiv} to obtain an equivalent interleaving $I'$ satisfying the theorem condition.
+	The resulting $U'$ can then be MCed without simulating HTM rollbacks.
 \end{proof}
 
-\subsection{Modeling Transaction Atomicity}
+\subsection{Memory Access Analysis}
 
 Next, we address the memory accesses within transactions with regard to data-race analysis.
-From Theorem~\ref{thm:equiv} we have that the body of all transactions may be executed atomically within the MC environment.
+From Theorem~\ref{thm:atom} we have that the body of all transactions may be executed atomically within the MC environment.
 While they may interleave within other non-transactional sequences,
@@
 \[
-	\ti@1,\tj@1-5,\ti@7,\mathbf{\ti@9a},\tk@1-5,\mathbf{\ti@8},\ti@9b-B
+	\tiat{1},\tjat{1-5},\tiat{7},\underline{\tiat{9a}},\tkat{1-5},\underline{\tiat{8}},\tiat{9b-B}
 \]
 is possible
-(reordered accesses highlighted in bold).
+(reordered accesses underlined for emphasis).
 An {\tt mfence} barrier is needed between lines 8 and 9 to solve this problem on TSO \cite{tsx-need-barrier}.

### 2018-03-13 htm/paper.tex
wax poetic

@@
 
-I hope I can get away with citing this in my thesis.
+Stateless model checking research is a perpetual existence of staring up the sheer cliff face that is the exponential curve.
+As new concurrency paradigms emerge,
+we make our living by adapting our reduction algorithms and search strategies to them to climb ever higher on that curve,
+eking out a few more loop iterations or slightly higher thread counts in our verification guarantees.
+Whether that is beautiful in its imperfection or cause for despair is merely a matter of perspective.
+Also, the author hopes their committee won't mind the publication venue when they cite this in their thesis.
 

### 2018-03-13 htm/paper.tex
eval stub

@@
 \cite{chess,dbug-ssv,spin,inspect,r4,portend,samc,mcr,randomized-scheduler}.
-This work builds upon Landslide \cite{landslide,quicksand},
+This work builds upon Landslide \cite{landslide,quicksand,landslide-phdthesis},
 a simulator-based tester which checks both user- and kernel-level programs
@@
 \begin{proof}
-	For every $I \in U$ with $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$ $\tjjat{\dots},\tkkat{\dots},\tiiat{\beta_{b+1}}... \in I$,
+	For every $I \in U$ with $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$ $\tjjat{\dots},\tkkat{\dots},\tiiat{\beta_{b+1}} \in I$,
 	apply Lemma~\ref{lem:equiv} to obtain an equivalent interleaving $I'$ satisfying the theorem condition.
@@
 \section{Evaluation}
+
+To the best of our knowledge, this is the first work to test transactional programs in a model-checking environment,
+%To the best of the author's knowledge, Landslide is the first MC to check transactional concurrency,
+so no other MC State-of-the-Art%
+\footnote{My DJ name.}
+exists to compare to in controlled experiments.
+Nevertheless, we pose the following evaluation questions.
+
+\begin{enumerate}
+	\item How quickly does Landslide find bugs in incorrect transactional programs of varying sizes?
+	\item How quickly does Landslide verify correct transactional programs of varying sizes?
+	\item Does Landslide find any previously-unknown bugs in real-world transactional code?
+\end{enumerate}
+
+Our evaluation suite comprises several hand-written unit tests,
+as well as \cite{htm-mario}'s microbenchmarks and transactional AVL tree and separate-chaining hashmap.
+All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
+
+...We don't actually have any results here yet.
+We intended to update our submission before camera-ready,
+but if somehow this becomes the final conference version of the paper,
+please look to \cite{landslide-phdthesis}, coming soon.
 

### 2018-03-13 htm/paper.tex
minor fixes

@@
 as well as deadlocks, use-after-frees, and segfaults.
+For this work we also detect use of {\tt xend} outside of a transaction as a bug.
 
@@
 	\label{thm:atom}
-	For any state space $U$ of a transactionally-concurrent program,
+	For any state space $S$ of a transactionally-concurrent program,
 	an equivalent state space exists in which all transactions are either executed atomically or aborted immediately.
@@
 \begin{proof}
-	For every $I \in U$ with $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$ $\tjjat{\dots},\tkkat{\dots},\tiiat{\beta_{b+1}} \in I$,
+	For every $I \in S$ with $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$ $\tjjat{\dots},\tkkat{\dots},\tiiat{\beta_{b+1}} \in I$,
 	apply Lemma~\ref{lem:equiv} to obtain an equivalent interleaving $I'$ satisfying the theorem condition.
-	The resulting $U'$ can then be MCed without simulating HTM rollbacks.
+	The resulting $S'$ can then be MCed without ever simulating HTM rollbacks.
 \end{proof}

### 2018-03-13 htm/paper.tex
state the data race theorem

@@
 
+Let $\tiiat{\mu},\tjjat{\nu}$ be a pair of memory accesses to the same address, at least one a write,
+in some transactional execution $I$ normalized under Lemma~\ref{lem:equiv}.
+Then let $\mathsf{lockify}_m(\tkkat{L})$ denote a function over instructions in $I$,
+which replaces $\tkkat{L}$ with $\tkkat{\mathsf{lock}(m)}$ if $L$ is a successful HTM begin,
+with a no-op if $L$ is a transaction abort,
+or with $\tkkat{\mathsf{unlock}(m)}$ if $L$ is an HTM end,
+or no replacement otherwise.
+Finally, let $I' = \exists m. \mathsf{lockify}_m(I)$,
+the execution with all successful transactions replaced by an abstract global lock.
+Lemma~\ref{lem:equiv} guarantees mutual exclusion of $m$.
+
 \begin{theorem}[Transactions are a Global Lock]
-	asdfasdfaf
+	$\tiiat{\mu},\tjjat{\nu}$ is a data race in $I$ iff it is a data race in $I'$.
 \end{theorem}
@@
 \begin{proof}
-	asdfasdf
+We prove one case for each variant definiton for data races \cite{quicksand}.
+
+\begin{itemize}
+	\item {\bf Limited Happens-Before.}
+	\item {\bf Pure Happens-Before.}
+\end{itemize}
+
 \end{proof}

### 2018-03-13 htm/paper.tex
cuidado!

@@
 
+\newcommand\llama[1]{\includegraphics[width=#1]{llama.pdf}}
+\newcommand\llitem{\item[\raisebox{-0.15em}{\llama{1.2em}}]}
+
 \begin{lemma}[Equivalence of Aborts]
@@
 	\begin{itemize}
-		\item $\tiiat{\alpha}$ be an HTM begin operation,
-		\item $\tiiat{\beta_1}\dots\tiiat{\beta_n}$ be the transaction body (with $\beta_n$ the HTM end call),
-		\item $\tiiat{\phi_1}\dots\tiiat{\phi_m}$ be the failure path, and
-		\item $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.
+		\llitem $\tiiat{\alpha}$ be an HTM begin operation,
+		\llitem $\tiiat{\beta_1}\dots\tiiat{\beta_n}$ be the transaction body (with $\beta_n$ the HTM end call),
+		\llitem $\tiiat{\phi_1}\dots\tiiat{\phi_m}$ be the failure path, and
+		\llitem $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.
 	\end{itemize}
@@
 \begin{itemize}
-	\item {\bf Limited Happens-Before.}
-	\item {\bf Pure Happens-Before.}
+	\llitem {\bf Limited Happens-Before.}
+	\llitem {\bf Pure Happens-Before.}
 \end{itemize}

### 2018-03-13 htm/paper.tex
prove that sucker

--- 2018-03-13 htm/paper.tex
@@
 Finally, let $I' = \exists m. \mathsf{lockify}_m(I)$,
-the execution with all successful transactions replaced by an abstract global lock.
+the execution with the boundaries of all successful transactions replaced by an abstract global lock.
 Lemma~\ref{lem:equiv} guarantees mutual exclusion of $m$.
@@
 \begin{proof}
-We prove one case for each variant definiton for data races \cite{quicksand}.
+We prove one case for each variant definiton for data races supported in Landslide \cite{quicksand}.
+For each, we semiformally state what it means to race in an execution with synchronizing HTM instructions.
 
@@
 	\item {\bf Limited Happens-Before.}
+		Under this definition, to be a data race in $I$ they must be reorderable at instruction granularity.
+		\begin{itemize}
+			\llitem $I \Rightarrow I'$:
+				If $\tiiat{\mu},\tjjat{\nu}$ race in $I$,
+				then they cannot both be in successful transactions,
+				or else placing \tiiat{\mu} within the boundaries of \tjjat{\nu}'s transaction
+				would cause the latter to abort, invalidating \tjjat{\nu}, or vice versa.
+				Hence they will not both hold $m$ in $I'$.
+				Otherwise their lock-sets and DPOR dependence relation remain unchanged.
+			\llitem $I' \Rightarrow I$:
+				If $\tiiat{\mu},\tjjat{\nu}$ race in $I'$,
+				both corresponding threads cannot hold $m$;
+				WLOG let $\tii$ not hold $m$ during $\tiiat{\mu}$.
+				Then in $I$, $\tiiat{\mu}$ is not in a transaction.
+				With the remainder of their lock-sets still disjoint,
+				and still not DPOR-dependent, $\tjjat{\nu}$ (or its containing transaction)
+				can then be reordered directly before or after $\tiiat{\mu}$.
+		\end{itemize}
 	\item {\bf Pure Happens-Before.}
+		WLOG fix $\tiiat{\mu} \prec \tjjat{\nu} \in I$.
+		Then under this definition, to be a data race in $I$ there must be no pair of synchronizing instructions
+		$\tiiat{\epsilon}$ and $\tjjat{\chi}$ such that
+		\[
+			\tiiat{\mu} \prec \tiiat{\epsilon} \prec \tjjat{\chi} \prec \tjjat{\nu}
+		\]
+		$\in I$ to update the vector clock epoch between \tiiat{\mu} and \tjjat{\nu}.
+		\begin{itemize}
+			\llitem $I \Rightarrow I'$:
+				If $\tiiat{\mu},\tjjat{\nu}$ race in $I$,
+				then they cannot both be in successful transactions,
+				or else Lemma~\ref{lem:equiv} normalization would provide
+				the corresponding HTM end and begin for $\tii$ and $\tjj$ respectively.
+				Hence there will be no unlock/lock pair on $m$ in $I'$ to satisfy the above sequence.
+			\llitem $I' \Rightarrow I$:
+				If $\tiiat{\mu},\tjjat{\nu}$ race in $I'$,
+				then they cannot both hold $m$,
+				or else $\mathsf{lockify}_m$ would provide the corresponding
+				unlock and lock for $\tii$ and $\tjj$ respectively.
+				Hence there will be no HTM end/begin pair in $I$ to satisfy the above sequence.
 		\end{itemize}
-
+		%In both cases $I$ and $I'$ otherwise have the same synchronizing instructions.
+\end{itemize}
+Hence, data-race analysis is sound with transactions abstracted as an abstract global lock.
 \end{proof}

### 2018-03-13 htm/paper.tex
touchups

@@
 	\item {\bf Limited Happens-Before.}
-		Under this definition, to be a data race in $I$ they must be reorderable at instruction granularity.
+		Under this definition, to be a data race in $I$ they must be reorderable at instruction granularity
+		\cite{tsan,hybriddatarace}.
 		\begin{itemize}
@@
 		Then under this definition, to be a data race in $I$ there must be no pair of synchronizing instructions
-		$\tiiat{\epsilon}$ and $\tjjat{\chi}$ such that
+		$\tiiat{\epsilon}$ (a release edge) and $\tjjat{\chi}$ (an acquire edge) such that
 		\[
@@
 		\]
-		$\in I$ to update the vector clock epoch between \tiiat{\mu} and \tjjat{\nu}.
+		$\in I$ to update the vector clock epoch between \tiiat{\mu} and \tjjat{\nu} \cite{djit,fasttrack}.
 		\begin{itemize}

### 2018-03-13 htm/paper.tex
more minor touchup

@@
 
-\definecolor{lavender}{RGB}{178,65,178} % * 3/4
+\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
 \definecolor{seafoam}{RGB}{65,178,76} % * 3/4
@@
 	\label{lem:equiv}
-	Let:%
-	\footnote{Arbitrary code may not be structured to distinguish these as nicely as in our examples;
-	e.g., more code may exist in the success branch after {\tt \_xend()};
-	such would be considered part of $\omega$ here.}
+	Let:
 	\begin{itemize}
@@
 		\item $\tiiat{\phi_1}\dots\tiiat{\phi_m}$ be the failure path, and
-		\item $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.
+		\item $\tiiat{\omega_1}\dots\tiiat{\omega_l}$ be the subsequent code executed unconditionally.%
+			\footnote{Arbitrary code may not be structured to distinguish these as nicely as in our examples;
+			e.g., more code may exist in the success branch after {\tt \_xend()};
+			such would be considered part of $\omega$ here.}
 	\end{itemize}
@@
 		\[
-			\tiiat{\mu} \prec \tiiat{\epsilon} \prec \tjjat{\chi} \prec \tjjat{\nu}
+			\tiiat{\mu} \prec \tiiat{\epsilon} \prec \tjjat{\chi} \prec \tjjat{\nu} \in I
 		\]
-		$\in I$ to update the vector clock epoch between \tiiat{\mu} and \tjjat{\nu} \cite{djit,fasttrack}.
+		to update the vector clock epoch between \tiiat{\mu} and \tjjat{\nu} \cite{djit,fasttrack}.
 		\begin{itemize}
@@
 \end{itemize}
-Hence, data-race analysis is sound with transactions abstracted as an abstract global lock.
+Hence, data-race analysis is sound when transaction boundaries are replaced by an abstract global lock.
 \end{proof}

### 2018-03-13 htm/paper.tex
variants

@@
 \section{Background}
+\label{sec:background}
 
@@
 	\end{enumerate}
-	is observationally equivalent.
+	exists and is observationally equivalent.
 \end{lemma}
@@
 				or else Lemma~\ref{lem:equiv} normalization would provide
-				the corresponding HTM end and begin for $\tii$ and $\tjj$ respectively.
+				the corresponding HTM end and begin for $\tiiat{\epsilon}$ and $\tjjat{\chi}$ respectively.
 				Hence there will be no unlock/lock pair on $m$ in $I'$ to satisfy the above sequence.
@@
 				or else $\mathsf{lockify}_m$ would provide the corresponding
-				unlock and lock for $\tii$ and $\tjj$ respectively.
+				unlock and lock for $\tiiat{\epsilon}$ and $\tjjat{\chi}$ respectively.
 				Hence there will be no HTM end/begin pair in $I$ to satisfy the above sequence.
@@
 	\item Does Landslide find any previously-unknown bugs in real-world transactional code?
+	\item By the way, should the evaluations of MC research papers quantify variance in their CPU-time performance?
 \end{enumerate}
@@
 
-...We don't actually have any results here yet.
-We intended to update our submission before camera-ready,
-but if somehow this becomes the final conference version of the paper,
-please look to \cite{landslide-phdthesis}, coming soon.
+\subsection{Bugs}
+
+\subsection{Verification}
+
+For verification tests, we configured Landslide %'s Iterative Deepening algorithm
+to optimize for completion time by prioritizing the maximal state space job and cancelling all others
+({\tt -M} command-line flag),
+a feature newly added since \cite{quicksand} and hence evaluated for the first time here.
+
+\subsection{Variance}
+
+Because many of the verification tests are long-running,
+and we are writing this too close to the deadline (who doesn't),
+we regret being unable to present every performance measurement above
+as an average of multiple samples with error bars.
+Nevertheless, we make some effort to address variance.
+
+We noticed significant slowdowns when varying several factors of our experimental environment.
+For example, multiple Landslides running at once on our test machine slow each other down,
+%(including multiple single state spaces in a single test when not using {\tt -M}),
+likely reuslting from kernel resource contention as Landslide uses {\tt fork()} to save simulation state,
+Table~\ref{tab:variants} shows the impact of running a single Landslide instance with {\tt -M} on
+{\tt counter(2,2)} (384 distinct interleavings)
+with various programs also running, despite never saturating our test system's 8 CPUs.
+
+\begin{table}[h]
+	\begin{center}
+	\begin{tabular}{l|c|c}
+		workload & cputime (s) & vs baseline \\
+		\hline
+		{\sf none	} & 14.95 $\pm$ 0.17	& 1	\\
+		{\sf vid-L	} & 15.13 $\pm$ 0.10	& 1.01x	\\
+		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 1.04x	\\
+		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 1.14x	\\
+		{\sf ls		} & 19.11 $\pm$ 0.32	& 1.28x	\\
+		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 1.31x	\\
+	\end{tabular}
+	\end{center}
+	\caption{Performance variance on {\tt counter(2,2)} with other programs running on the test machine.
+		{\sf vid-L} is full-screen video (played locally with {\tt mplayer}),
+		{\sf ff/c} is Firefox and Chrome (idle, $\le$20 tabs),
+		{\sf vid-S} is full-screen video (streamed via Crunchyroll),
+		{\sf ls} is a 2nd instance of Landslide.
+		Average of 10 samples, $\pm$ is 1 stddev.
+	}
+	\label{tab:variants}
+\end{table}
+
+We also show
+in Table~\ref{tab:variants2}
+the impact of Landslide needing to automatically annotate a test being run for the first time,
+rather than reusing existing instrumentation,
+as well as the variance of Landslide in bug-finding mode (i.e., without {\tt -M}).
+%(2128 interleavings total).
+Surprisingly, the {\tt bugfind} distribution was bimodal,
+with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142$\pm$1.67,
+suggesting two distinct scheduling patterns for its workqueue threads.
+Future work should figure out why.
+
+\begin{table}[h]
+	\begin{center}
+	\begin{tabular}{l|r|r}
+		landslide mode & total int's & cputime (s) \\
+		\hline
+		{\sf verif} ({\tt -M})		& 403 & 14.95 $\pm$ 0.17	\\
+		{\sf reinstrument}		& 403 & 24.72 $\pm$ 0.10	\\
+		{\sf bugfind} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
+	\end{tabular}
+	\end{center}
+	\caption{Performance variance on {\tt counter(2,2)} in various modes.
+		Average of 10 samples, $\pm$ is 1 stddev.
+	}
+	\label{tab:variants2}
+\end{table}
+
+We conclude that MC performance evaluations must address
+variables in the experimental environment
+at least enough to ensure consistent performance between runs.
+So doing, multiple samples and error bars are then necessary
+only when using nondeterministic search ordering strategies
+such as Iterative Deepening.
+Otherwise, they should be shown only for a token small test
+to assure the reader the environment is consistent in the larger tests,
+(especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
+% We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper
+
+In this evaluation, all other performance numbers were collected with
+with Firefox and Chrome as the only other machine load.
+(we gotta stay amused somehow).
+The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.
 
@@
 to track the set of abort codes possible for each transaction.
-It harnesses DPOR's existing memory analysis to identify when {\tt \_XABORT\_CONFLICT} is possible,
-and instruments {\tt \_xabort()} calls to record any user-supplied codes.
+%According to the any-reason principle in Section~\ref{sec:background},
+{\tt \_XABORT\_RETRY} is always enabled;
+then,
+it harnesses DPOR's existing memory analysis to identify when {\tt \_XABORT\_CONFLICT} is possible,
+and instruments {\tt \_xabort()} calls to record any user-supplied codes for {\tt \_XABORT\_EXPLICIT}.
 This comes at a cost of even more state space explosion,
@@
 none of which include the above-listed sequence.
-Recent work extended DPOR to incorporate TSO and PSO memory nondeterminism \cite{tsopso};
+Recent work extended DPOR to support TSO and PSO memory nondeterminism \cite{tsopso};
 if incorporated into Landslide, we could find or verify the absence of such bugs.

### 2018-03-14 htm/paper.tex
show some sick buges

@@
 and HTM nondeterminism can then be enabled with the {\tt -X} command-line flag.
-Several example tests are also provided:
-{\tt htm1} demonstrates the bug in Figure~\ref{fig:htm-example},
-and {\tt htm2} demonstrates Figure~\ref{fig:htm-fixed}'s fix.
+We have also extended its Iterative Deepening implementation \cite{quicksand}
+with an option ({\tt -M}) to
+to optimize for completion time by prioritizing the maximal state space job and cancelling all others,
+which we will use in our evaluation following.
+All test cases therein are also available in the repository linked above.
+%Several example tests are also provided:
+%{\tt htm1} demonstrates the bug in Figure~\ref{fig:htm-example},
+%and {\tt htm2} demonstrates Figure~\ref{fig:htm-fixed}'s fix.
 
@@
 \section{Evaluation}
+
+\begin{table*}[t]
+	\begin{center}
+		% TODO: highlight the winning statistic for each test
+	\begin{tabular}{cc||r|r|r||r|r|r|r}
+			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
+		buggy test	& params&cputime&walltime&int's&cputime&walltime&int's& SS size (est.) \\
+		\hline
+		\hline
+		{\tt htm1}
+			& 2,1	& 45.78	& 9.70	& 21	& 9.47	& 6.40	& 21	& 213 \\
+			& 2,2	& 84.14	& 13.59	& 33	& 10.39	& 7.70	& 49	& 1536 \\
+			& 2,3	& 131.91& 20.44	& 73	& 12.83	& 9.67	& 113	& 10752 \\
+			& 2,4	& 255.75& 37.56	& 257	& 18.63	& 15.86	& 257	& 73728 \\
+		\cline{2-9}
+			& 3,1	& 114.06& 17.45	& 15	& 9.50	& 6.79	& 21	& 13653 \\
+			& 3,2	& 109.60& 26.16	& 49	& 10.72	& 7.97	& 49	& 393216 \\
+			& 3,3	& 124.80& 20.40	& 73	& 13.84	& 11.01	& 113	& 11010048 \\
+			& 3,4	& 227.49& 35.15	& 161	& 31.37	& 28.53	& 257	& 301989888 \\
+		\cline{2-9}
+			& 4,1	& 53.08	& 9.79	& 15	& 9.82	& 7.00	& 21	& 873813 \\
+		\hline
+		{\tt swapbug}
+			& 2,1	& 70.95	& 13.45	& 16	& 38.96	& 13.15	& 109	& 194 \\
+			& 2,2	& 107.28& 17.45	& 146	& 44.73	& 19.47	& 281	& 1620 \\
+			& 2,3	& 280.05& 38.70	& 352	& 60.30	& 35.55	& 718	& 12748 \\
+			& 2,4	& 617.94& 81.50	& 834	& 108.58& 82.60	& 1820	& 97823 \\
+		\cline{2-9}
+			& 3,1	& 11.19	& 3.46	& 9	& 3.13	& 3.13	& 9	& 432 \\
+			& 3,2	& 12.41	& 3.76	& 19	& 3.33	& 3.33	& 19	& 5832 \\
+			& 3,3	& 15.77	& 4.58	& 41	& 3.95	& 3.95	& 41	& 77759 \\
+			& 3,4	& 24.82	& 6.85	& 89	& 5.20	& 5.20	& 89	& 1026431 \\
+		\cline{2-9}
+			& 4,1	& 702.33& 98.06	& 361	& 90.94	& 64.27	& 1607	& 138758 \\
+			& 4,2	&	&	&	& 674.87& 650.58& 16511	& 13755188 \\
+	\end{tabular}
+	\end{center}
+	\caption{Landslide's bug-finding performance on various test configurations.
+		Quicksand's workqueue approach optimized for fast bug-finding
+		is compared against our maximum-state-space-prioritizing approach for fast verification.
+		For each, we list the CPU-time and wall-clock time (both in seconds)
+		elapsed until the bug was found,
+		plus the number of interleavings of the ultimately buggy state space.
+		Landslide's state space size estimation \cite{estimation},
+		though approximate at best,
+		confers a sense of the exponential explosion.
+	}
+	\label{tab:buges}
+\end{table*}
 
@@
 	\item How quickly does Landslide verify correct transactional programs of varying sizes?
-	\item Does Landslide find any previously-unknown bugs in real-world transactional code?
-	\item By the way, should the evaluations of MC research papers quantify variance in their CPU-time performance?
+	%\item Does Landslide find any previously-unknown bugs in real-world transactional code?
+	\item By the way, should MC research papers quantify variance in their CPU-time performance experiments?
 \end{enumerate}
 
-Our evaluation suite comprises several hand-written unit tests,
-as well as \cite{htm-mario}'s microbenchmarks and transactional AVL tree and separate-chaining hashmap.
+Our evaluation suite comprises several hand-written unit tests
+and \cite{htm-mario}'s microbenchmarks and transactional AVL tree and separate-chaining hashmap,
+as follows.
+
+\begin{itemize}
+	\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example} (assertion failure on {\tt count}).
+	\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
+	\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex failure path with a simple {\tt xadd}.
+	\item {\tt swap}: Microbenchmark that swaps values in an array.
+	\item {\tt swapbug}: {\tt swap} modified to introduce a bug in the failure path (deadlock).
+\end{itemize}
+
+The notation {\tt testname(K,N)} will denote a test configuration of $K$ threads, each running $N$ iterations of the test logic.
 All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
@@
 
+Table~\ref{tab:buges}
+presents our bug-finding results.
+We configured Landslide to run the Quicksand algorithm \cite{quicksand}
+with a time limit of 30 minutes,%
+\footnote{This was necessary for {\tt swapbug(4,1)} to trigger a switch to the buggy state space.}
+shown left,
+as well as to prioritize the maximal state space as discussed above,
+shown right.
+We make the following observations/conclusions:
+\begin{enumerate}
+	\item The multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
+		than that of the total number of interleavings (10-30x)
+		as the test parameters increase.
+		In other words,
+		should they exist,
+		Landslide find bugs reasonably quickly in these transactional programs
+		despite prohibitive exponential explosion in total state space size.
+	\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
+		this does not correlate with better CPU-time or even better wall-clock time.
+		While possibly a consequence of the modest test size,
+		this suggests future work should extend the Quicksand algorithm
+		to prioritize by state space maximality as well as by state space estimate.
+\end{enumerate}
+
 \subsection{Verification}
-
-For verification tests, we configured Landslide %'s Iterative Deepening algorithm
-to optimize for completion time by prioritizing the maximal state space job and cancelling all others
-({\tt -M} command-line flag),
-a feature newly added since \cite{quicksand} and hence evaluated for the first time here.
 
@@
 %(including multiple single state spaces in a single test when not using {\tt -M}),
-likely reuslting from kernel resource contention as Landslide uses {\tt fork()} to save simulation state,
+likely reuslting from kernel resource contention as Landslide uses {\tt fork()} to save simulation state.
 Table~\ref{tab:variants} shows the impact of running a single Landslide instance with {\tt -M} on
@@
 
-\begin{table}[h]
+\begin{table}[t]
 	\begin{center}
-	\begin{tabular}{l|c|c}
-		workload & cputime (s) & vs baseline \\
+	\begin{tabular}{l|c|c|c}
+		workload & cputime (s) & vs self avg & vs baseline \\
 		\hline
-		{\sf none	} & 14.95 $\pm$ 0.17	& 1	\\
-		{\sf vid-L	} & 15.13 $\pm$ 0.10	& 1.01x	\\
-		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 1.04x	\\
-		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 1.14x	\\
-		{\sf ls		} & 19.11 $\pm$ 0.32	& 1.28x	\\
-		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 1.31x	\\
+		{\sf none	} & 14.95 $\pm$ 0.17	& 0.99-1.02x	& (baseline)	\\
+		{\sf vid-L	} & 15.13 $\pm$ 0.10	& 0.99-1.01x	& 1.01x	\\
+		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 0.99-1.02x	& 1.04x	\\
+		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 0.98-1.05x	& 1.14x	\\
+		{\sf ls		} & 19.11 $\pm$ 0.32	& 0.97-1.03x	& 1.28x	\\
+		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 0.96-1.02x	& 1.31x	\\
 	\end{tabular}
@@
 
-\begin{table}[h]
+\begin{table}[t]
 	\begin{center}
 	\begin{tabular}{l|r|r}
-		landslide mode & total int's & cputime (s) \\
+		Landslide mode & total int's & cputime (s) \\
 		\hline
-		{\sf verif} ({\tt -M})		& 403 & 14.95 $\pm$ 0.17	\\
-		{\sf reinstrument}		& 403 & 24.72 $\pm$ 0.10	\\
+		{\sf verif} ({\tt -M})		& 403 & 14.95 $\pm$ 00.17	\\
+		{\sf reinstrument}		& 403 & 24.72 $\pm$ 00.10	\\
 		{\sf bugfind} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
@@
 such as Iterative Deepening.
-Otherwise, they should be shown only for a token small test
-to assure the reader the environment is consistent in the larger tests,
+Otherwise, considering the maximum variance of 5\% in Table~\ref{tab:variants},
+%(2\% in reasonable loads),
+they should be shown only for a token small test
+to assure the reader the environment is consistent for the larger tests
 (especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
@@
 
-In this evaluation, all other performance numbers were collected with
-with Firefox and Chrome as the only other machine load.
+Outside of Table~\ref{tab:variants}, we collected all performance numbers
+with Firefox and Chrome as the only other significant machine load.
 (we gotta stay amused somehow).

### 2018-03-14 htm/paper.tex
more touchups including epsilon cite

@@
 			& 4,1	& 53.08	& 9.79	& 15	& 9.82	& 7.00	& 21	& 873813 \\
+			& 4,2	&117.07	& 19.09	& 33	& 11.54	& 8.55	& 49	& 100663296 \\
 		\hline
@@
 			& 4,1	& 702.33& 98.06	& 361	& 90.94	& 64.27	& 1607	& 138758 \\
-			& 4,2	&	&	&	& 674.87& 650.58& 16511	& 13755188 \\
+			& 4,2	&9738.43&1257.20& 3023	& 674.87& 650.58& 16511	& 13755188 \\
 	\end{tabular}
@@
 		For each, we list the CPU-time and wall-clock time (both in seconds)
-		elapsed until the bug was found,
-		plus the number of interleavings of the ultimately buggy state space.
-		Landslide's state space size estimation \cite{estimation},
+		elapsed, %until the bug was found,
+		plus the number of interleavings of the ultimately buggy state space tested,
+		before the bug was found.
+		Lastly, Landslide's state space size estimation \cite{estimation},
 		though approximate at best,
@@
 so no other MC State-of-the-Art%
-\footnote{My DJ name.}
+\footnote{The author's DJ name.}
 exists to compare to in controlled experiments.
@@
 	\item {\tt swapbug}: {\tt swap} modified to introduce a bug in the failure path (deadlock).
+		% TODO
+	\item Those data structure tests are coming soon. Please stay tuned.
 \end{itemize}
@@
 with a time limit of 30 minutes,%
-\footnote{This was necessary for {\tt swapbug(4,1)} to trigger a switch to the buggy state space.}
+\footnote{This was necessary in {\tt swapbug}s (4,1) and (4,2) to trigger a switch to the buggy state space.}
 shown left,
@@
 	\item The multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
-		than that of the total number of interleavings (10-30x)
+		than that of the total number of interleavings (10-100x)
 		as the test parameters increase.
@@
 
+% TODO
+This section is coming soon. Please stay tuned.
+
 \subsection{Variance}
@@
 Because many of the verification tests are long-running,
-and we are writing this too close to the deadline (who doesn't),
+and we are writing this too close to the submission deadline (who doesn't),
 we regret being unable to present every performance measurement above
 as an average of multiple samples with error bars.
-Nevertheless, we make some effort to address variance.
+Nevertheless, we make some effort to address variance \cite{epsilon}.
 
@@
 %(including multiple single state spaces in a single test when not using {\tt -M}),
-likely reuslting from kernel resource contention as Landslide uses {\tt fork()} to save simulation state.
+likely arising from kernel resource contention as Landslide uses {\tt fork()} to save simulation state.
 Table~\ref{tab:variants} shows the impact of running a single Landslide instance with {\tt -M} on
@@
 	\begin{tabular}{l|c|c|c}
-		workload & cputime (s) & vs self avg & vs baseline \\
+		load & cputime (s) & vs self avg & vs baseline \\
 		\hline
@@
 		{\sf ls} is a 2nd instance of Landslide.
-		Average of 10 samples, $\pm$ is 1 stddev.
+		Average of 10 samples, $\pm N$ is 1 stddev.
 	}
@@
 %(2128 interleavings total).
-Surprisingly, the {\tt bugfind} distribution was bimodal,
-with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142$\pm$1.67,
+The {\sf bugfind} variance was surprisingly bimodal,
+with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142.11$\pm$1.67,
 suggesting two distinct scheduling patterns for its workqueue threads.
@@
 	\caption{Performance variance on {\tt counter(2,2)} in various modes.
-		Average of 10 samples, $\pm$ is 1 stddev.
+		Average of 10 samples, $\pm N$ is 1 stddev.
 	}

### 2018-03-14 htm/paper.tex
fix swapbug(3,*), measure stepmania, discuss spectre/meltdown

@@
 to optimize for completion time by prioritizing the maximal state space job and cancelling all others,
-which we will use in our evaluation following.
+while maintaining its soundness guarantee,
+which we will use in our evaluation.
 All test cases therein are also available in the repository linked above.
@@
 		\cline{2-9}
-			& 3,1	& 11.19	& 3.46	& 9	& 3.13	& 3.13	& 9	& 432 \\
-			& 3,2	& 12.41	& 3.76	& 19	& 3.33	& 3.33	& 19	& 5832 \\
-			& 3,3	& 15.77	& 4.58	& 41	& 3.95	& 3.95	& 41	& 77759 \\
-			& 3,4	& 24.82	& 6.85	& 89	& 5.20	& 5.20	& 89	& 1026431 \\
-		\cline{2-9}
-			& 4,1	& 702.33& 98.06	& 361	& 90.94	& 64.27	& 1607	& 138758 \\
-			& 4,2	&9738.43&1257.20& 3023	& 674.87& 650.58& 16511	& 13755188 \\
+			& 3,1	&1275.04& 163.42& 771	&{\em >30m}&{\em >30m}&{\em unk.}& 184984 \\
+			& 3,2	&{\em >30m}&{\em >30m}&{\em unk.}&{\em >30m}&{\em >30m}&{\em unk.}& 3099225 \\
+		% whoops, this was with the dumb lock ordering
+		%	& 3,1	& 11.19	& 3.46	& 9	& 3.13	& 3.13	& 9	& 432 \\
+		%	& 3,2	& 12.41	& 3.76	& 19	& 3.33	& 3.33	& 19	& 5832 \\
+		%	& 3,3	& 15.77	& 4.58	& 41	& 3.95	& 3.95	& 41	& 77759 \\
+		%	& 3,4	& 24.82	& 6.85	& 89	& 5.20	& 5.20	& 89	& 1026431 \\
+		%\cline{2-9}
+		%	& 4,1	& 702.33& 98.06	& 361	& 90.94	& 64.27	& 1607	& 138758 \\
+		%	& 4,2	&9738.43&1257.20& 3023	& 674.87& 650.58& 16511	& 13755188 \\
 	\end{tabular}
@@
 
-The notation {\tt testname(K,N)} will denote a test configuration of $K$ threads, each running $N$ iterations of the test logic.
+The notation {\tt testname}$(K,N)$ will denote a test configuration of $K$ threads, each running $N$ iterations of the test logic.
 All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
@@
 We configured Landslide to run the Quicksand algorithm \cite{quicksand}
-with a time limit of 30 minutes,%
-\footnote{This was necessary in {\tt swapbug}s (4,1) and (4,2) to trigger a switch to the buggy state space.}
+% this note obsolete since swapbug was fixed and ends at 3,2
+%\footnote{This was necessary in {\tt swapbug}s (4,1) and (4,2) to trigger a switch to the buggy state space.}
 shown left,
 as well as to prioritize the maximal state space as discussed above,
-shown right.
-We make the following observations/conclusions:
-\begin{enumerate}
+shown right,
+each with a time limit of 30 minutes.
+We make two observations/conclusions:
+\begin{itemize}
 	\item The multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
@@
 		despite prohibitive exponential explosion in total state space size.
-	\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
-		this does not correlate with better CPU-time or even better wall-clock time.
-		While possibly a consequence of the modest test size,
-		this suggests future work should extend the Quicksand algorithm
-		to prioritize by state space maximality as well as by state space estimate.
-\end{enumerate}
+	%\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
+	%	this does not correlate with better CPU-time or even better wall-clock time.
+	%	While possibly a consequence of the modest test size,
+	%	this suggests future work should extend the Quicksand algorithm
+	%	to prioritize by state space maximality as well as by state space estimate.
+	\item Quicksand's ability to find bugs in fewer distinct interleavings %overall smaller state spaces
+		does not necessarily correlate with better performance. %CPU- or even wall-clock time.
+		%Comparing to the break-even point in Quicksand's evaluation \cite{quicksand},
+		%most of these test configurations fall on the side favouring the single-state-space approach,
+		Most of our tests are too small for its approach to pay off,
+		with {\tt swapbug}(3,1) as its lone win.
+		While the prior work showed plenty more wins \cite{quicksand},
+		%an upper bound also exists:
+		%in {\tt swapbug}(3,2),
+		%not even the {\em minimal} state space was completed in time.
+		%this suggests
+		future MCs could prioritize state spaces using not just size estimation
+		but state space maximality as well
+		%but a hybrid approach also conisdering state space maximality and preemption bounds \cite{chess-icb},
+		to soften the trade-off
+		both for smaller tests and for verification
+\end{itemize}
+Speaking of which...
 
@@
 we regret being unable to present every performance measurement above
-as an average of multiple samples with error bars.
-Nevertheless, we make some effort to address variance \cite{epsilon}.
+as an average of multiple samples with error bars \cite{epsilon}.
+Nevertheless, we make some effort to address variance.
 
@@
 Table~\ref{tab:variants} shows the impact of running a single Landslide instance with {\tt -M} on
-{\tt counter(2,2)} (384 distinct interleavings)
+{\tt counter}(2,2) (384 distinct interleavings)
 with various programs also running, despite never saturating our test system's 8 CPUs.
+% for i in `seq 1 10`; do ./landslide -X -v -p counter -M 2>&1 | grep 'total CPU time'; done
 
@@
 		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 0.99-1.02x	& 1.04x	\\
+		{\sf sm5	} & 16.63 $\pm$ 0.07	& 0.99-1.01x	& 1.11x \\
 		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 0.98-1.05x	& 1.14x	\\
 		{\sf ls		} & 19.11 $\pm$ 0.32	& 0.97-1.03x	& 1.28x	\\
+		% simply love title screen
+		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& 0.97-1.02x	& 1.28x \\
 		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 0.96-1.02x	& 1.31x	\\
@@
 	\end{center}
-	\caption{Performance variance on {\tt counter(2,2)} with other programs running on the test machine.
+	\caption{Performance variance on {\tt counter}(2,2) with other programs running on the test machine.
 		{\sf vid-L} is full-screen video (played locally with {\tt mplayer}),
 		{\sf ff/c} is Firefox and Chrome (idle, $\le$20 tabs),
+		{\sf sm5} is StepMania 5.1 (during gameplay \cite{itg2}),
 		{\sf vid-S} is full-screen video (streamed via Crunchyroll),
@@
 
-We also show
-in Table~\ref{tab:variants2}
+Table~\ref{tab:variants2} shows
 the impact of Landslide needing to automatically annotate a test being run for the first time,
 rather than reusing existing instrumentation,
-as well as the variance of Landslide in bug-finding mode (i.e., without {\tt -M}).
+as well as the variance of Quicksand mode (i.e., without {\tt -M}).
 %(2128 interleavings total).
-The {\sf bugfind} variance was surprisingly bimodal,
+The {\sf quicksand} variance was surprisingly bimodal,
 with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142.11$\pm$1.67,
@@
 		{\sf reinstrument}		& 403 & 24.72 $\pm$ 00.10	\\
-		{\sf bugfind} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
+		{\sf quicksand} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
 	\end{tabular}
 	\end{center}
-	\caption{Performance variance on {\tt counter(2,2)} in various modes.
+	\caption{Performance variance on {\tt counter}(2,2) in various modes.
 		Average of 10 samples, $\pm N$ is 1 stddev.
@@
 We conclude that MC performance evaluations must address
-variables in the experimental environment
-at least enough to ensure consistent performance between runs.
+experimental environment variables
+%at least enough
+to ensure consistent performance between runs.
+Considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
+we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
 So doing, multiple samples and error bars are then necessary
@@
 such as Iterative Deepening.
-Otherwise, considering the maximum variance of 5\% in Table~\ref{tab:variants},
-%(2\% in reasonable loads),
-they should be shown only for a token small test
-to assure the reader the environment is consistent for the larger tests
+Otherwise, considering the low variance shown in Table~\ref{tab:variants},
+they need be shown only for a token small test
+to promise the reader the same consistency in the larger tests
 (especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
@@
 
-Outside of Table~\ref{tab:variants}, we collected all performance numbers
+Outside of Table~\ref{tab:variants}, we measured all performance numbers
 with Firefox and Chrome as the only other significant machine load.

### 2018-03-15 htm/paper.tex
verif

@@
 			& 2,1	& 45.78	& 9.70	& 21	& 9.47	& 6.40	& 21	& 213 \\
+		(assertion)
 			& 2,2	& 84.14	& 13.59	& 33	& 10.39	& 7.70	& 49	& 1536 \\
@@
 			& 2,1	& 70.95	& 13.45	& 16	& 38.96	& 13.15	& 109	& 194 \\
+		(deadlock)
 			& 2,2	& 107.28& 17.45	& 146	& 44.73	& 19.47	& 281	& 1620 \\
@@
 \begin{itemize}
-	\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example} (assertion failure on {\tt count}).
+	\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}. %(assertion failure on {\tt count}).
 	\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
@@
 	\item {\tt swap}: Microbenchmark that swaps values in an array.
-	\item {\tt swapbug}: {\tt swap} modified to introduce a bug in the failure path (deadlock).
-		% TODO
-	\item Those data structure tests are coming soon. Please stay tuned.
+	\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
+	\item {\tt map\_basic}: Concurrent insertion test for the separate-chaining hashmap.
+	\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
+		% TODO more
 \end{itemize}
@@
 		to soften the trade-off
-		both for smaller tests and for verification
+		both for smaller tests and for verification.
 \end{itemize}
@@
 
-% TODO
-This section is coming soon. Please stay tuned.
+Landslide proved the following tests correct using {\tt -M} mode.
+With no bugs to find, comparing the different testing modes against each other is meaningless;
+we simply present their state space sizes and runtime in Table~\ref{tab:verifs}.
+%
+Also shown are several
+larger tests we were unable to complete before the submission deadline,
+each given a timeout of 1 hour. % TODO: ??
+Their completion times are listed as {\em unk.}
+and their listed state space sizes represent Landslide's estimate at the moment of timeout.
+%For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
+Future work may expand our coverage upon this frontier \cite{landslide-phdthesis}.
+
+\begin{table}[h]
+	\begin{center}
+		\begin{tabular}{cc|r|r}
+			test & params & cputime (s) & SS size \\
+			\hline
+			{\tt counter}
+			& (2,1) & 5.57	& 30	\\
+			& (2,2) & 15.53	& 384	\\
+			& (2,3) & 20.00	& 5280	\\
+			& (2,4) &2211.10& 75264	\\
+			& (3,1) & 57.9	& 1960	\\
+			& (3,2) &	&	\\
+			\hline
+			{\tt htm2}
+			& (2,1) & 18.57	& 294	\\
+			& (2,2) & 133.78& 4902	\\
+			& (2,3) &1986.98& 79017	\\
+			& (3,1) &11672.15& 467730	\\
+			& (3,2) &	&	\\
+			\hline
+			{\tt swap}
+			& (2,1) & 38.77	& 228	\\
+			& (2,2) &	&	\\
+			\hline
+			{\tt map\_basic}
+			& (2,1) &	&	\\
+			\hline
+			{\tt map\_basicer}
+			& (2,1) & 877.44& 28635	\\
+			& (2,2) &	&	\\
+		\end{tabular}
+	\end{center}
+	\caption{Transactional tests verified by Landslide.}
+	\label{tab:verifs}
+\end{table}
+
+%\begin{table}[h]
+%	\begin{center}
+%		\begin{tabular}{cc|r}
+%			test & params & SS size (est.) \\
+%			\hline
+%		\end{tabular}
+%	\end{center}
+%	\caption{Transactional test verification timeouts.}
+%	\label{tab:timeouts}
+%\end{table}
 
@@
 
-Our HTM implementation in Landslide includes an experimental feature
+Our HTM implementation includes an experimental feature
 to track the set of abort codes possible for each transaction.
@@
 by (usually) 1 plus however many distinct {\tt \_xabort()} codes the program uses.
+
+Many such branches may be equivalent;
+for example,
+explicit and conflict aborts need not be tested separately in transactions
+whose failure paths do not distinguish the cause,
+and the perennial {\tt \_XABORT\_RETRY} abort can be skipped entirely
+if the client abstracts it away behind a retry loop.
+In fact, applying both reductions simultaneously
+produces a state space smaller than the original;
+testing them by hand (trusting their soundness by visual inspection)
+on {\tt map\_basic}(2,1) reduced its 28635 interleavings to 11577.
+%We tested both reductions by hand, trusting their soundness by visual inspection,
+%on {\tt map\_basic}(2,1),
+%and preview the resulting state space sizes in Table~\ref{tab:abort-codes}.
+%
+%\begin{table}[h]
+%	\caption{...}
+%	\label{tab:abort-codes}
+%\end{table}
+%
 Future work could use static or dynamic flow analysis to
-identify which transactions actually branch on their failure codes
-to prune equivalent backup paths;
+identify such reduction opportunities automatically;
 for now, this feature is disabled by default but accessible via the {\tt -A} command-line option
@@
 if incorporated into Landslide, we could find or verify the absence of such bugs.
+Visual inspection of \cite{htm-mario}'s HTM data structures found the same implementation pattern with no barriers;
+we would urge any reader interested in using them to add them in by hand first.
 

### 2018-03-15 htm/paper.tex
minor reorganization of bunch of stuff

@@
 
-\begin{table}[t]
+\begin{table}[h]
 	\begin{center}
@@
 to ensure consistent performance between runs.
-Considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
-we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
 So doing, multiple samples and error bars are then necessary
@@
 % We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper
+%% i'm not actually sure this fits..
+%Finally, considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
+%we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
 
-Outside of Table~\ref{tab:variants}, we measured all performance numbers
+Outside of Table~\ref{tab:variants}, we fixed our environment by measuring all performance numbers
 with Firefox and Chrome as the only other significant machine load.
-(we gotta stay amused somehow).
+%(we gotta stay amused somehow).
+We believe the exponential differences among completion times justifies the absence of error bars,
+which would be expected to show 2\% variance if extrapolating from this experiment.
 The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.
@@
 testing them by hand (trusting their soundness by visual inspection)
-on {\tt map\_basic}(2,1) reduced its 28635 interleavings to 11577.
+on {\tt map\_basicer}(2,1) reduced its 28635 interleavings to 11577
+(although {\tt map\_basic}(2,1) lies still beyond reach).
 %We tested both reductions by hand, trusting their soundness by visual inspection,
@@
 Section~\ref{sec:design}'s formalization of thread interleavings does not account for read/write reorderings
-possible on relaxed memory consistency architectures \cite{memory-consistency-models}.
+possible on relaxed consistency architectures \cite{memory-consistency-models}.
 In fact,

### 2018-03-15 htm/paper.tex
the next step is to colored it.

@@
 We have also extended its Iterative Deepening implementation \cite{quicksand}
-with an option ({\tt -M}) to
+with an option ({\tt -M} flag) to
 to optimize for completion time by prioritizing the maximal state space job and cancelling all others,
@@
 
+\definecolor{darkcyan}{RGB}{0,128,128}
+\definecolor{lime}{RGB}{48,128,0}
+\definecolor{pinkish}{RGB}{128,34,102} % ff4488
+\definecolor{brownish}{RGB}{128,96,0}
+\newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
+\newcommand\cpu[1]{\hilight{darkcyan}{{#1}}\xspace}
+\newcommand\wtm[1]{\hilight{lime}{{#1}}\xspace}
+\newcommand\ints[1]{\hilight{pinkish}{{#1}}\xspace}
+
 \begin{table*}[t]
@@
 			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
-		buggy test	& params&cputime&walltime&int's&cputime&walltime&int's& SS size (est.) \\
+		\bf buggy test	& \bf params&\cpu{\bf cputime}&\wtm{\bf walltime}&\ints{\bf int's}&\cpu{\bf cputime}&\wtm{\bf walltime}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
 		\hline
@@
 		{\tt htm1}
-			& 2,1	& 45.78	& 9.70	& 21	& 9.47	& 6.40	& 21	& 213 \\
+			& 2,1	&\cpu{  45.78}&\wtm{9.70}&\ints{21}& \cpu{9.47}& \wtm{6.40}& \ints{21	}& \ETA{213} \\
 		(assertion)
-			& 2,2	& 84.14	& 13.59	& 33	& 10.39	& 7.70	& 49	& 1536 \\
-			& 2,3	& 131.91& 20.44	& 73	& 12.83	& 9.67	& 113	& 10752 \\
-			& 2,4	& 255.75& 37.56	& 257	& 18.63	& 15.86	& 257	& 73728 \\
+			& 2,2	&\cpu{  84.14}&\wtm{13.59}&\ints{33}& \cpu{10.39}& \wtm{7.70}& \ints{49	}& \ETA{1536} \\
+			& 2,3	&\cpu{ 131.91}&\wtm{20.44}&\ints{73}& \cpu{12.83}& \wtm{9.67}& \ints{113	}& \ETA{10752} \\
+			& 2,4	&\cpu{ 255.75}&\wtm{37.56}&\ints{257}& \cpu{18.63}& \wtm{15.86}& \ints{257	}& \ETA{73728} \\
 		\cline{2-9}
-			& 3,1	& 114.06& 17.45	& 15	& 9.50	& 6.79	& 21	& 13653 \\
-			& 3,2	& 109.60& 26.16	& 49	& 10.72	& 7.97	& 49	& 393216 \\
-			& 3,3	& 124.80& 20.40	& 73	& 13.84	& 11.01	& 113	& 11010048 \\
-			& 3,4	& 227.49& 35.15	& 161	& 31.37	& 28.53	& 257	& 301989888 \\
+			& 3,1	&\cpu{ 114.06}&\wtm{17.45}&\ints{15}& \cpu{9.50}& \wtm{6.79}& \ints{21	}& \ETA{13653} \\
+			& 3,2	&\cpu{ 109.60}&\wtm{26.16}&\ints{49}& \cpu{10.72}& \wtm{7.97}& \ints{49	}& \ETA{393216} \\
+			& 3,3	&\cpu{ 124.80}&\wtm{20.40}&\ints{73}& \cpu{13.84}& \wtm{11.01}& \ints{113	}& \ETA{11010048} \\
+			& 3,4	&\cpu{ 227.49}&\wtm{35.15}&\ints{161}& \cpu{31.37}& \wtm{28.53}& \ints{257	}& \ETA{301989888} \\
 		\cline{2-9}
-			& 4,1	& 53.08	& 9.79	& 15	& 9.82	& 7.00	& 21	& 873813 \\
-			& 4,2	&117.07	& 19.09	& 33	& 11.54	& 8.55	& 49	& 100663296 \\
+			& 4,1	&\cpu{ 53.08}&\wtm{9.79}&\ints{15}& \cpu{9.82}& \wtm{7.00}& \ints{21	}& \ETA{873813} \\
+			& 4,2	&\cpu{117.07}&\wtm{19.09}&\ints{33}& \cpu{11.54}& \wtm{8.55}& \ints{49	}& \ETA{100663296} \\
 		\hline
 		{\tt swapbug}
-			& 2,1	& 70.95	& 13.45	& 16	& 38.96	& 13.15	& 109	& 194 \\
+			& 2,1	&\cpu{ 70.95}&\wtm{13.45}&\ints{16}& \cpu{38.96}& \wtm{13.15}& \ints{109	}& \ETA{194} \\
 		(deadlock)
-			& 2,2	& 107.28& 17.45	& 146	& 44.73	& 19.47	& 281	& 1620 \\
-			& 2,3	& 280.05& 38.70	& 352	& 60.30	& 35.55	& 718	& 12748 \\
-			& 2,4	& 617.94& 81.50	& 834	& 108.58& 82.60	& 1820	& 97823 \\
+			& 2,2	&\cpu{ 107.28}&\wtm{17.45}&\ints{146}& \cpu{44.73}& \wtm{19.47}& \ints{281	}& \ETA{1620} \\
+			& 2,3	&\cpu{ 280.05}&\wtm{38.70}&\ints{352}& \cpu{60.30}& \wtm{35.55}& \ints{718	}& \ETA{12748} \\
+			& 2,4	&\cpu{ 617.94}&\wtm{81.50}&\ints{834}& \cpu{108.58}& \wtm{82.60}& \ints{1820	}& \ETA{97823} \\
 		\cline{2-9}
-			& 3,1	&1275.04& 163.42& 771	&{\em >30m}&{\em >30m}&{\em unk.}& 184984 \\
-			& 3,2	&{\em >30m}&{\em >30m}&{\em unk.}&{\em >30m}&{\em >30m}&{\em unk.}& 3099225 \\
+			& 3,1	&\cpu{1275.04}&\wtm{163.42}&\ints{771}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{184984} \\
+			& 3,2	&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{3099225} \\
 		% whoops, this was with the dumb lock ordering
@@
 we simply present their state space sizes and runtime in Table~\ref{tab:verifs}.
-%
-Also shown are several
-larger tests we were unable to complete before the submission deadline,
-each given a timeout of 1 hour. % TODO: ??
-Their completion times are listed as {\em unk.}
-and their listed state space sizes represent Landslide's estimate at the moment of timeout.
-%For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
-Future work may expand our coverage upon this frontier \cite{landslide-phdthesis}.
 
+\newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}\xspace}
 \begin{table}[h]
@@
 			test & params & cputime (s) & SS size \\
+			\hline
 			\hline
@@
 			& (2,4) &2211.10& 75264	\\
-			& (3,1) & 57.9	& 1960	\\
+			& (3,1) & 57.90	& 1960	\\
 			& (3,2) &	&	\\
@@
 			{\tt map\_basic}
-			& (2,1) &	&	\\
+			& (2,1) & \ETAdag{10d 16h} & \ETAdag{16388977} \\
+			%[JOB 15] progress: 70330/16388977 brs (0.429130%), ETA 10d 16h 34m 15s (elapsed 43m 21s)
+			%[JOB 15] TIMED OUT (0.429130%; ETA 10d 16h 34m 15s)
 			\hline
@@
 			& (2,1) & 877.44& 28635	\\
-			& (2,2) &	&	\\
+			& (2,2) &	&	\\ % currently running
 		\end{tabular}
 	\end{center}
-	\caption{Transactional tests verified by Landslide.}
+	\caption{Transactional tests verified (or not) by Landslide.}
 	\label{tab:verifs}
@@
 
+Several larger tests we were unable to complete before the submission deadline are also shown.
+%each timing out after 1 hour. % TODO: ??
+%Their completion times are listed as {\em unk.}
+These are indicated with \ETAdag{}
+and their listed ETAs and state space sizes represent Landslide's estimate %at the moment of timeout.
+after a timeout of 1 hour
+(hopefully enough for data-race PPs to saturate and estimates to stablize).
+%For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
+Future work may expand our coverage upon this frontier \cite{landslide-phdthesis}.
 %\begin{table}[h]

### 2018-03-15 htm/paper.tex
can you colored this picture?

@@
 			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
-		\bf buggy test	& \bf params&\cpu{\bf cputime}&\wtm{\bf walltime}&\ints{\bf int's}&\cpu{\bf cputime}&\wtm{\bf walltime}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
+		\bf buggy test	& \bf params&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
 		\hline
@@
 		is compared against our maximum-state-space-prioritizing approach for fast verification.
-		For each, we list the CPU-time and wall-clock time (both in seconds)
+		For each, we list the CPU-time and wall-clock time %(both in seconds)
 		elapsed, %until the bug was found,
@@
 
-\newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}\xspace}
+\newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}}
 \begin{table}[h]
 	\begin{center}
-		\begin{tabular}{cc|r|r}
-			test & params & cputime (s) & SS size \\
+		\begin{tabular}{cc||r|r}
+			\bf test & \bf params & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
 			\hline
@@
 			{\tt counter}
-			& (2,1) & 5.57	& 30	\\
-			& (2,2) & 15.53	& 384	\\
-			& (2,3) & 20.00	& 5280	\\
-			& (2,4) &2211.10& 75264	\\
-			& (3,1) & 57.90	& 1960	\\
+			& (2,1) & \cpu{5.57}	& \ints{30}	\\
+			& (2,2) & \cpu{15.53}	& \ints{384}	\\
+			& (2,3) &\cpu{20.00}	& \ints{5280}	\\
+			& (2,4) &\cpu{2211.10}	& \ints{75264}	\\
+			& (3,1) & \cpu{57.90}	& \ints{1960}	\\
 			& (3,2) &		& 	\\
@@
 			{\tt htm2}
-			& (2,1) & 18.57	& 294	\\
-			& (2,2) & 133.78& 4902	\\
-			& (2,3) &1986.98& 79017	\\
-			& (3,1) &11672.15& 467730	\\
+			& (2,1) & \cpu{18.57}	& \ints{294}	\\
+			& (2,2) & \cpu{133.78}	& \ints{4902}	\\
+			& (2,3) &\cpu{1986.98}	& \ints{79017}	\\
+			& (3,1) &\cpu{11672.15}	& \ints{467730}	\\
 			& (3,2) &	&	\\
@@
 			{\tt swap}
-			& (2,1) & 38.77	& 228	\\
+			& (2,1) & \cpu{38.77}	& \ints{228}	\\
 			& (2,2) &	&	\\
+			& (3,1) &	&	\\ % running atm
 			\hline
@@
 			{\tt map\_basicer}
-			& (2,1) & 877.44& 28635	\\
-			& (2,2) &	&	\\ % currently running
+			& (2,1) & \cpu{877.44}& \ints{28635}	\\
+			& (2,2) & \ETAdag{2d 6h} & \ETAdag{5925634} \\
+			%[JOB 6] progress: 35042/5925634 brs (0.591363%), ETA 2d 6h 40m 30s (elapsed 21m 12s)
+			%[JOB 6] TIMED OUT (0.591363%; ETA 2d 6h 40m 30s)
 		\end{tabular}
@@
 after a timeout of 1 hour
-(hopefully enough for data-race PPs to saturate and estimates to stablize).
+(hopefully enough for data-race PPs to saturate and estimates to stablize,
+although note the two estimate types use different algorithms so may disagree significantly).
 %For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
@@
 	\begin{tabular}{l|c|c|c}
-		load & cputime (s) & vs self avg & vs baseline \\
+		\bf load & \cpu{\bf cpu (s)} & \bf vs self avg & \bf vs baseline \\
 		\hline
-		{\sf none	} & 14.95 $\pm$ 0.17	& 0.99-1.02x	& (baseline)	\\
-		{\sf vid-L	} & 15.13 $\pm$ 0.10	& 0.99-1.01x	& 1.01x	\\
-		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 0.99-1.02x	& 1.04x	\\
-		{\sf sm5	} & 16.63 $\pm$ 0.07	& 0.99-1.01x	& 1.11x \\
-		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 0.98-1.05x	& 1.14x	\\
-		{\sf ls		} & 19.11 $\pm$ 0.32	& 0.97-1.03x	& 1.28x	\\
+		{\sf none	} & \cpu{14.95 $\pm$ 0.17} & 0.99-1.02x	& (baseline)	\\
+		{\sf vid-L	} & \cpu{15.13 $\pm$ 0.10} & 0.99-1.01x	& 1.01x	\\
+		{\sf ff/c	} & \cpu{15.55 $\pm$ 0.16} & 0.99-1.02x	& 1.04x	\\
+		{\sf sm5	} & \cpu{16.63 $\pm$ 0.07} & 0.99-1.01x	& 1.11x \\
+		{\sf ff/c+vid-S	} & \cpu{17.09 $\pm$ 0.34} & 0.98-1.05x	& 1.14x	\\
+		{\sf ls		} & \cpu{19.11 $\pm$ 0.32} & 0.97-1.03x	& 1.28x	\\
 		% simply love title screen
 		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& 0.97-1.02x	& 1.28x \\
-		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 0.96-1.02x	& 1.31x	\\
+		{\sf ff/c+ls	} & \cpu{19.66 $\pm$ 0.50} & 0.96-1.02x	& 1.31x	\\
 	\end{tabular}
@@
 	\begin{center}
-	\begin{tabular}{l|r|r}
-		Landslide mode & total int's & cputime (s) \\
+	\begin{tabular}{l|r|c}
+		\bf Landslide mode & \bf total int's & \cpu{\bf cpu (s)} \\
 		\hline
-		{\sf verif} ({\tt -M})		& 403 & 14.95 $\pm$ 00.17	\\
-		{\sf reinstrument}		& 403 & 24.72 $\pm$ 00.10	\\
-		{\sf quicksand} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
+		{\sf verif} ({\tt -M})		& 403  & \cpu{14.95 $\pm$ 00.17} \\
+		{\sf reinstrument}		& 403  & \cpu{24.72 $\pm$ 00.10}	\\
+		{\sf quicksand} (no {\tt -M})	& 2128 & \cpu{120.53 $\pm$ 18.62}	\\
 	\end{tabular}

### 2018-03-15 htm/paper.tex
acks
avl_insert bug!!

@@
 	\begin{center}
-		% TODO: highlight the winning statistic for each test
 	\begin{tabular}{cc||r|r|r||r|r|r|r}
@@
 			& 3,2	&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{3099225} \\
+		\hline
+		{\tt avl\_insert}
+			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{83}	&\cpu{81.00}&\wtm{40.30}&\ints{336}&\ETA{379982} \\
+		(see text)
+			% XXX: this 2,3 -M-mode bug was found *mostly* running in isolation,
+			% but for the last few minutes i had another landslide running in parallel bc i thought itd time out.
+			% according to variance expt, it should be rerun for a *minor* speedup, but i can't be bothered
+			& 2,3	&\cpu{2670.87}&\wtm{330.45}&\ints{3066}	&\cpu{1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
+		\cline{2-9}
+			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{28}	&\cpu{69.99}&\wtm{24.25}&\ints{78}&\ETA{1572107} \\
 		% whoops, this was with the dumb lock ordering
@@
 	\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
-	\item {\tt map\_basic}: Concurrent insertion test for the separate-chaining hashmap.
+	\item {\tt avl\_insert}: AVL tree concurrent insertion test.
+	\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed. %(spoilers!!)
+	\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test.
 	\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
-		% TODO more
 \end{itemize}
@@
 each with a time limit of 30 minutes.
-We make two observations/conclusions:
-\begin{itemize}
-	\item The multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
+We draw three main conclusions from this data.
+%\begin{itemize}
+	%\item
+
+		{\bf Finding bugs quickly.}
+		The multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
 		than that of the total number of interleavings (10-100x)
@@
 		despite prohibitive exponential explosion in total state space size.
+		This corroborates the prior work \cite{quicksand},
+		extending its good news to the world of HTM.
+	%\item
+
+		{\bf New bugs.}
+		In addition to the bugs we intentionally wrote in {\tt htm1} and {\tt swapbug},
+		to our pleasant surprise
+		Landslide also found a previously-unknown bug in the transactional AVL tree,
+		exposed by {\tt avl\_insert} with any parameters higher than (2,1).
+		Figure~\ref{fig:avlbug} shows the root cause, essentially the {\tt htm1} bug in disguise.
+		% landslide-trace-1521093835.18927.html
+		This manifested
+		alternately
+		%in (2,2) and (3,1)
+		as a segfault
+		% landslide-trace-1521092404.23772.html
+		and
+		%in (2,3)
+		as a consistency-check assertion failure.
+		The presence of {\tt while (\_retry);} makes the necessary preemption window extremely small
+		in practice (between it and {\tt \_xbegin()}),
+		whereas MC is blind to such matters of chance.%
+		\footnote{Considering the loop does not affect the test's possible behaviours,
+		only its {\em likely} ones,
+		we removed it in Landslide's version of the test to keep the state space size manageable.}
+		%We removed the loop in Landslide's version of the test
+		%%To make this test Landslide-friendly, we removed that loop,
+		%which reduces the state space size, but maintains the program's semantics.}
+		Moreover, we conclude that even Figure~\ref{fig:htm-fixed}'s protocol's very proposer % こそさえ
+		getting it wrong motivates the need for MC on such programs,
+		and perhaps also means TSX primitives should be used only behind higher-level abstractions
+		such as lock elision \cite{lock-elision}.
+
+% TODO syntax hilighting
+\begin{figure}[t]
+	\begin{center}
+		\begin{tabular}{l}
+		%\texttt{void count() \{} \\
+		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
+		\texttt{while (\_retry);} \\
+		\texttt{if (\_xbegin() == SUCCESS) \{} \\
+		\texttt{~~~~tie(\_root,inserted) = \_insert(\_root,n); } \\
+		\texttt{~~~~\_xend();} \\
+		\texttt{\} else \{} \\
+		\texttt{~~~~pthread\_mutex\_lock(\&\_tree\_lock);} \\
+		\texttt{~~~~\_retry = true;} \\
+		\texttt{~~~~tie(\_root,inserted) = \_insert(\_root,n); } \\
+		\texttt{~~~~\_retry = false;} \\
+		\texttt{~~~~pthread\_mutex\_unlock(\&\_tree\_lock);} \\
+		\texttt{\}} \\
+		\end{tabular}
+	\end{center}
+	%\caption{mario man had a very large house. or a castle.}
+	\caption{Unmodified code from {\tt htmavl.cpp} with the previously-unknown bug found by Landslide.
+		The transaction path fails to check {\tt \_retry},
+		leading to data races and corruption just as in {\tt htm1}.
+		}
+	\label{fig:avlbug}
+\end{figure}
 	%\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
@@
 	%	to prioritize by state space maximality as well as by state space estimate.
-	\item Quicksand's ability to find bugs in fewer distinct interleavings %overall smaller state spaces
+	%\item
+
+		{\bf Performance comparison.}
+		Quicksand's ability to find bugs in fewer distinct interleavings %overall smaller state spaces
 		does not necessarily correlate with better performance. %CPU- or even wall-clock time.
@@
 		Most of our tests are too small for its approach to pay off,
-		with {\tt swapbug}(3,1) as its lone win.
+		with {\tt swapbug}(3,1) and {\tt avl\_insert}(2,3) as its notable wins.
 		While the prior work showed plenty more wins \cite{quicksand},
@@
 		both for smaller tests and for verification.
-\end{itemize}
+%\end{itemize}
 Speaking of which...
@@
 
-Landslide proved the following tests correct using {\tt -M} mode.
+Landslide proved the following tests correct.
 With no bugs to find, comparing the different testing modes against each other is meaningless;
-we simply present their state space sizes and runtime in Table~\ref{tab:verifs}.
+we simply present their state space sizes and runtime (using {\tt -M}) in Table~\ref{tab:verifs}.
 
@@
 			{\tt counter}
-			& (2,1) & \cpu{5.57}	& \ints{30}	\\
-			& (2,2) & \cpu{15.53}	& \ints{384}	\\
-			& (2,3) &\cpu{20.00}	& \ints{5280}	\\
-			& (2,4) &\cpu{2211.10}	& \ints{75264}	\\
-			& (3,1) & \cpu{57.90}	& \ints{1960}	\\
-			& (3,2) &		& 	\\
+			& 2,1 & \cpu{5.57}	& \ints{30}	\\
+			& 2,2 & \cpu{15.53}	& \ints{384}	\\
+			& 2,3 &\cpu{20.00}	& \ints{5280}	\\
+			& 2,4 &\cpu{2211.10}	& \ints{75264}	\\
+			& 3,1 & \cpu{57.90}	& \ints{1960}	\\
+			%& 3,2 &		& 	\\
 			\hline
 			{\tt htm2}
-			& (2,1) & \cpu{18.57}	& \ints{294}	\\
-			& (2,2) & \cpu{133.78}	& \ints{4902}	\\
-			& (2,3) &\cpu{1986.98}	& \ints{79017}	\\
-			& (3,1) &\cpu{11672.15}	& \ints{467730}	\\
-			& (3,2) &	&	\\
+			& 2,1 & \cpu{18.57}	& \ints{294}	\\
+			& 2,2 & \cpu{133.78}	& \ints{4902}	\\
+			& 2,3 &\cpu{1986.98}	& \ints{79017}	\\
+			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
+			%& 3,2 &	&	\\
 			\hline
 			{\tt swap}
-			& (2,1) & \cpu{38.77}	& \ints{228}	\\
-			& (2,2) &	&	\\
-			& (3,1) &	&	\\ % running atm
+			& 2,1 & \cpu{38.77}	& \ints{228}	\\
+			& 2,2 &	&	\\ % TODO: fix, or use timeout if you HAVE to
+			%[JOB 8] progress: 97913/161897 brs (60.478362%), ETA 1h 29m 30s (elapsed 59m 52s)
+			% [JOB 8] TIMED OUT (60.478362%; ETA 1h 29m 30s)
+
+			& 3,1 & \ETAdag{8d 11h} & \ETAdag{411312}	\\
+			%[JOB 8] progress: 103873/411312 brs (25.254021%), ETA 8d 11h 36m 0s (elapsed 59m 42s)
+			%[JOB 8] TIMED OUT (25.254021%; ETA 8d 11h 36m 0s)
+			\hline
+			{\tt avl\_insert}
+			& 2,1 & \cpu{1083.35}	& \ints{40062}	\\
+			%& 2,2 & \multicolumn{2}{c}{\em (bug!)} \\
 			\hline
 			{\tt map\_basic}
-			& (2,1) & \ETAdag{10d 16h} & \ETAdag{16388977} \\
+			& 2,1 & \ETAdag{10d 16h} & \ETAdag{16388977} \\
 			%[JOB 15] progress: 70330/16388977 brs (0.429130%), ETA 10d 16h 34m 15s (elapsed 43m 21s)
@@
 			{\tt map\_basicer}
-			& (2,1) & \cpu{877.44}& \ints{28635}	\\
-			& (2,2) & \ETAdag{2d 6h} & \ETAdag{5925634} \\
+			& 2,1 & \cpu{877.44}& \ints{28635}	\\
+			& 2,2 & \ETAdag{2d 6h} & \ETAdag{5925634} \\
 			%[JOB 6] progress: 35042/5925634 brs (0.591363%), ETA 2d 6h 40m 30s (elapsed 21m 12s)
@@
 Several larger tests we were unable to complete before the submission deadline are also shown.
-%each timing out after 1 hour. % TODO: ??
 %Their completion times are listed as {\em unk.}
-These are indicated with \ETAdag{}
+These are indicated with $\dagger$ %\ETAdag{}
 and their listed ETAs and state space sizes represent Landslide's estimate %at the moment of timeout.
@@
 
-\begin{table}[h]
+\begin{table}[t]
 	\begin{center}
@@
 Table~\ref{tab:variants2} shows
-the impact of Landslide needing to automatically annotate a test being run for the first time,
+the impact of Landslide needing to automatically annotate {\tt counter}(2,2) being run for the first time,
 rather than reusing existing instrumentation,
-as well as the variance of Quicksand mode (i.e., without {\tt -M}).
+as well as the variance of Quicksand mode. %(i.e., without {\tt -M}).
 %(2128 interleavings total).
@@
 		\hline
-		{\sf verif} ({\tt -M})		& 403  & \cpu{14.95 $\pm$ 00.17} \\
-		{\sf reinstrument}		& 403  & \cpu{24.72 $\pm$ 00.10}	\\
+		{\sf verif} ({\tt -M})		& 403  & \cpu{~~14.95 $\pm$ 00.17} \\
+		{\sf reinstrument}		& 403  & \cpu{~~24.72 $\pm$ 00.10}	\\
 		{\sf quicksand} (no {\tt -M})	& 2128 & \cpu{120.53 $\pm$ 18.62}	\\
@@
 
-Outside of Table~\ref{tab:variants}, we fixed our environment by measuring all performance numbers
+Outside of this experiment, we fixed our environment by measuring all performance numbers
 with Firefox and Chrome as the only other significant machine load.
@@
 and the perennial {\tt \_XABORT\_RETRY} abort can be skipped entirely
-if the client abstracts it away behind a retry loop.
+if the client abstracts it away behind a retry loop.%
+\footnote{Proof left to future work. We've proved enough here already.}
 In fact, applying both reductions simultaneously
 produces a state space smaller than the original;
-testing them by hand (trusting their soundness by visual inspection)
-on {\tt map\_basicer}(2,1) reduced its 28635 interleavings to 11577
+testing them by hand
+on {\tt map\_basicer}(2,1)%
+\footnote{Trusting the reductions' soundness from visual inspection of the test.}
+reduced its 28635 interleavings to 11577
 (although {\tt map\_basic}(2,1) lies still beyond reach).
@@
 
+\acks
+
+We thank Mario Dehesa-Azuara for generously providing the HTM data structures and benchmarks used in our evaluation.
+%We also thank the member companies of the PDL Consortium (Broadcom,
+%Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
+%Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
+%Tintri, Two Sigma, Uber, Veritas and Western Digital).
+This work was supported in part by the U.S. Army Research Office under grant number W911NF0910273.
+
 \bibliographystyle{abbrv}

### 2018-03-15 htm/paper.tex
buncha minor prose stuff

@@
 \cite{chess,dbug-ssv,spin,inspect,r4,portend,samc,mcr,randomized-scheduler}.
-This work builds upon Landslide \cite{landslide,quicksand,landslide-phdthesis},
+This work builds upon Landslide \cite{landslide}, %,quicksand,landslide-phdthesis},
 a simulator-based tester which checks both user- and kernel-level programs
@@
 Landslide traces every memory access through the use of a simulated environment \cite{bochs},
-which is important for identifying data races to use as new preemption points \cite{fasttrack,djit,quicksand},
+which is important for identifying data races to use as new preemption points \cite{quicksand},
 as well as for identifying when a memory conflict may cause transaction aborts.
@@
 	satisfying case 1.
-	Otherwise, by DPOR's definition of transition dependence \cite{dpor,landslide-phdthesis},
+	Otherwise, by DPOR's definition of transition dependence \cite{dpor}, %,landslide-phdthesis},
 	$\tiiat{\beta_{b+1}}\dots\tiiat{\beta_n}$ is independent with the transitions of \tjj and \tkk,
@@
 
-Our implementation of HTM has been incorporated by the Landslide maintainers upstream.
-The repository is available open-source at \url{https://github.com/bblum/landslide}.
+Our implementation of HTM-equivalent semantics has been incorporated by the Landslide maintainers upstream.
+It is available open-source at \url{https://github.com/bblum/landslide}.
 Programs should be ported to the Pebbles userland \cite{thrlib,kspec},
@@
 We have also extended its Iterative Deepening implementation \cite{quicksand}
-with an option ({\tt -M} flag) to
-to optimize for completion time by prioritizing the maximal state space job and cancelling all others,
-while maintaining its soundness guarantee,
+with a new option ({\tt -M} flag) to
+to optimize for completion time by prioritizing the maximal state space job and cancelling all others
+(which maintains its soundness guarantee)
 which we will use in our evaluation.
@@
 		whereas MC is blind to such matters of chance.%
+		% NB. In case a reader doesn't believe this footnote,
+		% I put the loop back in (with thr_yield(-1) therein) to restore the original state space,
+		% and found the bug with these statistics:
+		% [JOB 28] progress: 1957/3005897 brs (0.065105%), ETA 6414673y 121d 18h 28m 59s (elapsed 3m 15s)
+		% [JOB 28] BUG FOUND: landslide-trace-1521099044.76395.html (1958 interleavings tested; 3 preemptions; job time 3m 15s; pldi time 746832861; new-fixed pldi cputime 823243872)
 		\footnote{Considering the loop does not affect the test's possible behaviours,
@@
 \begin{figure}[t]
-	\begin{center}
+	%\begin{center}
 		\begin{tabular}{l}
@@
 		\end{tabular}
-	\end{center}
+	%\end{center}
 	%\caption{mario man had a very large house. or a castle.}
@@
 			%& 2,2 & \multicolumn{2}{c}{\em (bug!)} \\
+			\hline
+			{\tt avl\_fixed} % TODO
+			& 2,1 & 	& 	\\
+			& 2,2 & 	& 	\\
 			\hline
@@
 	\begin{center}
-	\begin{tabular}{l|r|c}
-		\bf Landslide mode & \bf total int's & \cpu{\bf cpu (s)} \\
-		\hline
-		{\sf verif} ({\tt -M})		& 403  & \cpu{~~14.95 $\pm$ 00.17} \\
-		{\sf reinstrument}		& 403  & \cpu{~~24.72 $\pm$ 00.10}	\\
-		{\sf quicksand} (no {\tt -M})	& 2128 & \cpu{120.53 $\pm$ 18.62}	\\
+	\begin{tabular}{l|c|r}
+		\bf Landslide mode & \cpu{\bf cpu (s)} & \ints{\bf total int's} \\
+		\hline
+		{\sf verif} ({\tt -M})		& \cpu{~~14.95 $\pm$ 00.17} &  \ints{403} \\
+		{\sf reinstrument}		& \cpu{~~24.72 $\pm$ 00.10} &  \ints{403}     \\
+		{\sf quicksand} (no {\tt -M})	& \cpu{120.53 $\pm$ 18.62}  & \ints{2128}     \\
 	\end{tabular}

### 2018-03-15 htm/paper.tex
improve discussion

@@
 \definecolor{brownish}{RGB}{128,96,0}
+\definecolor{darklavender}{RGB}{68,44,138}
 \newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
@@
 			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{83}	&\cpu{81.00}&\wtm{40.30}&\ints{336}&\ETA{379982} \\
-		(see text)
+		(segfault+)
 			% XXX: this 2,3 -M-mode bug was found *mostly* running in isolation,
@@
 		getting it wrong motivates the need for MC on such programs,
-		and perhaps also means TSX primitives should be used only behind higher-level abstractions
-		such as lock elision \cite{lock-elision}.
+		%and perhaps also means TSX primitives should be used only behind higher-level abstractions
+		%such as lock elision \cite{lock-elision}.
+		and suggests TSX primitives should be encapsulated behind higher-level abstractions
+		such as lock elision \cite{lock-elision},
+		which can be verified in isolation with smaller state spaces
+		then trusted in turn when checking their client programs \cite{dbug-phdthesis}.
 
@@
 \newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}}
-\begin{table}[h]
+\begin{table}[t]
 	\begin{center}
@@
 %For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
+Though these tests proved beyond our reach,
+in contrast with the high speed expected of bug-finding results,
+we believe the verification guarantee may be worth the estimated time requirement
+for widely-used industrial implementations.
 Future work may expand our coverage upon this frontier \cite{landslide-phdthesis}.
-%\begin{table}[h]
-%	\begin{center}
-%		\begin{tabular}{cc|r}
-%			test & params & SS size (est.) \\
-%			\hline
-%		\end{tabular}
-%	\end{center}
-%	\caption{Transactional test verification timeouts.}
-%	\label{tab:timeouts}
-%\end{table}
 
@@
 
+\newcommand\vsx[1]{\hilight{darklavender}{#1}\xspace}
 \begin{table}[t]
@@
 	\begin{tabular}{l|c|c|c}
-		\bf load & \cpu{\bf cpu (s)} & \bf vs self avg & \bf vs baseline \\
-		\hline
-		{\sf none	} & \cpu{14.95 $\pm$ 0.17} & 0.99-1.02x	& (baseline)	\\
-		{\sf vid-L	} & \cpu{15.13 $\pm$ 0.10} & 0.99-1.01x	& 1.01x	\\
-		{\sf ff/c	} & \cpu{15.55 $\pm$ 0.16} & 0.99-1.02x	& 1.04x	\\
-		{\sf sm5	} & \cpu{16.63 $\pm$ 0.07} & 0.99-1.01x	& 1.11x \\
-		{\sf ff/c+vid-S	} & \cpu{17.09 $\pm$ 0.34} & 0.98-1.05x	& 1.14x	\\
-		{\sf ls		} & \cpu{19.11 $\pm$ 0.32} & 0.97-1.03x	& 1.28x	\\
+		\bf load & \cpu{\bf cpu (s)} & \vsx{\bf vs self avg} & \vsx{\bf vs baseline} \\
+		\hline
+		{\sf none	} & \cpu{14.95 $\pm$ 0.17} & \vsx{0.99-1.02x}	& \vsx{(baseline)}	\\
+		{\sf vid-L	} & \cpu{15.13 $\pm$ 0.10} & \vsx{0.99-1.01x}	& \vsx{1.01x}	\\
+		{\sf ff/c	} & \cpu{15.55 $\pm$ 0.16} & \vsx{0.99-1.02x}	& \vsx{1.04x}	\\
+		{\sf sm5	} & \cpu{16.63 $\pm$ 0.07} & \vsx{0.99-1.01x}	& \vsx{1.11x} \\
+		{\sf ff/c+vid-S	} & \cpu{17.09 $\pm$ 0.34} & \vsx{0.98-1.05x}	& \vsx{1.14x}	\\
+		{\sf ls		} & \cpu{19.11 $\pm$ 0.32} & \vsx{0.97-1.03x}	& \vsx{1.28x}	\\
 		% simply love title screen
-		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& 0.97-1.02x	& 1.28x \\
-		{\sf ff/c+ls	} & \cpu{19.66 $\pm$ 0.50} & 0.96-1.02x	& 1.31x	\\
+		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& \vsx{0.97-1.02x   } & \vsx{1.28x} \\
+		{\sf ff/c+ls	} & \cpu{19.66 $\pm$ 0.50} & \vsx{0.96-1.02x}	& \vsx{1.31x}	\\
 	\end{tabular}
@@
 
-This section discusses opportunities for future work.
+%This section discusses opportunities for future work.
+This section should be mandatory in all systems papers.
 
@@
 testing them by hand
-on {\tt map\_basicer}(2,1)%
-\footnote{Trusting the reductions' soundness from visual inspection of the test.}
-reduced its 28635 interleavings to 11577
-(although {\tt map\_basic}(2,1) lies still beyond reach).
+\footnote{Using visual inspection of the test to trust the reductions' soundness.}
+reduced {\tt map\_basicer}(2,1)'s 28635 interleavings to 11577.
+% TODO: try map_basic and get a 1 hour estimate
+%(although {\tt map\_basic}(2,1) lies still beyond reach).
 %We tested both reductions by hand, trusting their soundness by visual inspection,
@@
 if incorporated into Landslide, we could find or verify the absence of such bugs.
-Visual inspection of \cite{htm-mario}'s HTM data structures found the same implementation pattern with no barriers;
+Visual inspection of \cite{htm-mario}'s HTM data structures found no barriers used in this implementation pattern;
 we would urge any reader interested in using them to add them in by hand first.

### 2018-03-15 htm/paper.tex
pleease look at this picture. see? you colored it! q d e

@@
 
+\newcommand\hilight[2]{\color{#1}#2\color{black}}
+\definecolor{orange}{RGB}{192,96,0}
+\definecolor{olivegreen}{RGB}{0,127,0}
+\definecolor{brickred}{RGB}{192,0,0}
+\definecolor{commentblue}{RGB}{0,0,192}
+\definecolor{black}{RGB}{0,0,0}
+%\definecolor{lavender}{RGB}{238,85,238}
+%\definecolor{seafoam}{RGB}{85,238,102}
+%\definecolor{salmon}{RGB}{238,102,85} % ee6655
+
+% threads
+\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
+\definecolor{seafoam}{RGB}{65,178,76} % * 3/4
+\definecolor{salmon}{RGB}{208,89,74} % * 7/8
+
+% eval data
+\definecolor{darkcyan}{RGB}{0,128,128}
+\definecolor{lime}{RGB}{48,128,0}
+\definecolor{darklavender}{RGB}{68,44,138}
+\definecolor{brownish}{RGB}{128,96,0}
+\definecolor{pinkish}{RGB}{128,34,102} % ff4488
+
+% syntax hilighting
+\definecolor{brown}{RGB}{170,85,0}
+\newcommand\flow[1]{\hilight{brown}{#1}\xspace}
+\newcommand\const[1]{\hilight{brickred}{#1}\xspace}
+\newcommand\call[1]{\hilight{pinkish}{#1}}
+
 \begin{figure}[t]
@@
 		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
-		1 & \texttt{if ((status = \_xbegin()) == SUCCESS) \{} \\
+		1 & \texttt{\flow{if} ((status = \call{\_xbegin}()) == \const{SUCCESS}) \{} \\
 		2 & \texttt{~~~~x++;} \\
-		3 & \texttt{~~~~\_xend();} \\
-		4 & \texttt{\} else \{} \\
-		5 & \texttt{~~~~mutex\_lock(\&m);} \\
+		3 & \texttt{~~~~\call{\_xend}();} \\
+		4 & \texttt{\} \flow{else} \{} \\
+		5 & \texttt{~~~~\call{mutex\_lock}(\&m);} \\
 		6 & \texttt{~~~~x++;} \\
-		7 & \texttt{~~~~mutex\_unlock(\&m);} \\
+		7 & \texttt{~~~~\call{mutex\_unlock}(\&m);} \\
 		8 & \texttt{\}} \\
@@
 \]
-
-\newcommand\hilight[2]{\color{#1}#2\color{black}}
-\definecolor{orange}{RGB}{192,96,0}
-\definecolor{olivegreen}{RGB}{0,127,0}
-\definecolor{brickred}{RGB}{192,0,0}
-\definecolor{commentblue}{RGB}{0,0,192}
-\definecolor{black}{RGB}{0,0,0}
-%\definecolor{lavender}{RGB}{238,85,238}
-%\definecolor{seafoam}{RGB}{85,238,102}
-%\definecolor{salmon}{RGB}{238,102,85} % ee6655
-
-\definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
-\definecolor{seafoam}{RGB}{65,178,76} % * 3/4
-\definecolor{salmon}{RGB}{208,89,74} % * 7/8
 
@@
 		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
-		  & \texttt{prevent\_transactions = false;} \\
+		  & \texttt{prevent\_transactions = \const{false};} \\
 		\\
-		0 & \texttt{while (prevent\_transactions) continue;} \\
-		1 & \texttt{if ((status = \_xbegin()) == SUCCESS) \{} \\
-		2 & \texttt{~~~~if (prevent\_transactions)} \\
-		3 & \texttt{~~~~~~~~\_xabort();} \\
+		0 & \texttt{\flow{while} (prevent\_transactions) \flow{continue};} \\
+		1 & \texttt{\flow{if} ((status = \call{\_xbegin}()) == \const{SUCCESS}) \{} \\
+		2 & \texttt{~~~~\flow{if} (prevent\_transactions)} \\
+		3 & \texttt{~~~~~~~~\call{\_xabort}();} \\
 		4 & \texttt{~~~~x++;} \\
-		5 & \texttt{~~~~\_xend();} \\
-		6 & \texttt{\} else \{} \\
-		7 & \texttt{~~~~mutex\_lock(\&m);} \\
-		8 & \texttt{~~~~prevent\_transactions = true;} \\
+		5 & \texttt{~~~~\call{\_xend}();} \\
+		6 & \texttt{\} \flow{else} \{} \\
+		7 & \texttt{~~~~\call{mutex\_lock}(\&m);} \\
+		8 & \texttt{~~~~prevent\_transactions = \const{true};} \\
 		9 & \texttt{~~~~x++;} \\
-		A & \texttt{~~~~prevent\_transactions = false;} \\
-		B & \texttt{~~~~mutex\_unlock(\&m);} \\
+		A & \texttt{~~~~prevent\_transactions = \const{false};} \\
+		B & \texttt{~~~~\call{mutex\_unlock}(\&m);} \\
 		C & \texttt{\}} \\
@@
 
-\definecolor{darkcyan}{RGB}{0,128,128}
-\definecolor{lime}{RGB}{48,128,0}
-\definecolor{pinkish}{RGB}{128,34,102} % ff4488
-\definecolor{brownish}{RGB}{128,96,0}
-\definecolor{darklavender}{RGB}{68,44,138}
 \newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
@@
 		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
-		\texttt{while (\_retry);} \\
-		\texttt{if (\_xbegin() == SUCCESS) \{} \\
-		\texttt{~~~~tie(\_root,inserted) = \_insert(\_root,n); } \\
-		\texttt{~~~~\_xend();} \\
-		\texttt{\} else \{} \\
-		\texttt{~~~~pthread\_mutex\_lock(\&\_tree\_lock);} \\
-		\texttt{~~~~\_retry = true;} \\
-		\texttt{~~~~tie(\_root,inserted) = \_insert(\_root,n); } \\
-		\texttt{~~~~\_retry = false;} \\
-		\texttt{~~~~pthread\_mutex\_unlock(\&\_tree\_lock);} \\
+		\texttt{\flow{while} (\_retry);} \\
+		\texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
+		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
+		\texttt{~~~~\call{\_xend}();} \\
+		\texttt{\} \flow{else} \{} \\
+		\texttt{~~~~\call{pthread\_mutex\_lock}(\&\_tree\_lock);} \\
+		\texttt{~~~~\_retry = \const{true};} \\
+		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
+		\texttt{~~~~\_retry = \const{false};} \\
+		\texttt{~~~~\call{pthread\_mutex\_unlock}(\&\_tree\_lock);} \\
 		\texttt{\}} \\

### 2018-03-15 htm/paper.tex
few more verifs, some minors, spoilers!!, clarify remaining todos (down to 4!)

@@
 \definecolor{brown}{RGB}{170,85,0}
+% \definecolor{pink}{RGB}{170,45,136} % pinkish * 4/3 % but actually looks better darker
 \newcommand\flow[1]{\hilight{brown}{#1}\xspace}
@@
 as well as deadlocks, use-after-frees, and segfaults.
-For this work we also detect use of {\tt xend} outside of a transaction as a bug.
+For this work we also detect use of {\tt xend} outside of a transaction
+or {\tt xbegin} within one as a bug.
 
@@
 			& 3,2	&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{3099225} \\
-		\hline
-		{\tt avl\_insert}
-			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{83}	&\cpu{81.00}&\wtm{40.30}&\ints{336}&\ETA{379982} \\
-		(segfault+)
-			% XXX: this 2,3 -M-mode bug was found *mostly* running in isolation,
-			% but for the last few minutes i had another landslide running in parallel bc i thought itd time out.
-			% according to variance expt, it should be rerun for a *minor* speedup, but i can't be bothered
-			& 2,3	&\cpu{2670.87}&\wtm{330.45}&\ints{3066}	&\cpu{1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
-		\cline{2-9}
-			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{28}	&\cpu{69.99}&\wtm{24.25}&\ints{78}&\ETA{1572107} \\
 		% whoops, this was with the dumb lock ordering
@@
 		%	& 4,2	&9738.43&1257.20& 3023	& 674.87& 650.58& 16511	& 13755188 \\
+		\hline
+		{\tt avl\_insert}
+			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{83}	&\cpu{81.00}&\wtm{40.30}&\ints{336}&\ETA{379982} \\
+		(segfault+)
+			% XXX: this 2,3 -M-mode bug was found *mostly* running in isolation,
+			% but for the last few minutes i had another landslide running in parallel bc i thought itd time out.
+			% according to variance expt, it should be rerun for a *minor* speedup, but i can't be bothered
+			& 2,3	&\cpu{2670.87}&\wtm{330.45}&\ints{3066}	&\cpu{1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
+			& 2,4	&\cpu{3259.37}&\wtm{436.50}&\ints{1639}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{???} \\
+			% TODO
+		\cline{2-9}
+			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{28}	&\cpu{69.99}&\wtm{24.25}&\ints{78}&\ETA{1572107} \\
+			& 3,2	&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{???} \\
+			% TODO
 	\end{tabular}
@@
 	\item {\tt avl\_insert}: AVL tree concurrent insertion test.
-	\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed. %(spoilers!!)
+	\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed (spoilers!!).
 	\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test.
@@
 
-% TODO syntax hilighting
 \begin{figure}[t]
@@
 \newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}}
-\begin{table}[t]
+\begin{table}[h]
 	\begin{center}
-		\begin{tabular}{cc||r|r}
-			\bf test & \bf params & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
-			\hline
+		\begin{tabular}{cc|r|r}
+			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
+			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
+			\hline
+			\hline
+			{\tt htm2}
+			& 2,1 & \cpu{18.57}	& \ints{294}	\\
+			& 2,2 & \cpu{133.78}	& \ints{4902}	\\
+			& 2,3 &\cpu{1986.98}	& \ints{79017}	\\
+			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
+			%& 3,2 &	&	\\
 			\hline
@@
 			\hline
-			{\tt htm2}
-			& 2,1 & \cpu{18.57}	& \ints{294}	\\
-			& 2,2 & \cpu{133.78}	& \ints{4902}	\\
-			& 2,3 &\cpu{1986.98}	& \ints{79017}	\\
-			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
-			%& 3,2 &	&	\\
-			\hline
 			{\tt swap}
-			& 2,1 & \cpu{38.77}	& \ints{228}	\\
-			& 2,2 & &       \\ % TODO: fix, or use timeout if you HAVE to
+			%& 2,1 & \cpu{38.77}	& \ints{228}	\\ % pre-ignore_thr_func fix
+			& 2,1 & \cpu{32.98}	& \ints{193}	\\
+			& 2,2 & \cpu{3652.91}	& \ints{101150}	\\
 			%[JOB 8] progress: 97913/161897 brs (60.478362%), ETA 1h 29m 30s (elapsed 59m 52s)
@@
 
-			& 3,1 & \ETAdag{8d 11h} & \ETAdag{411312}	\\
+			& 3,1 & \ETAdag{8d 12h} & \ETAdag{411312}	\\
 			%[JOB 8] progress: 103873/411312 brs (25.254021%), ETA 8d 11h 36m 0s (elapsed 59m 42s)
@@
 			\hline
-			{\tt avl\_fixed} % TODO
-			& 2,1 & 	& 	\\
-			& 2,2 & 	& 	\\
+			{\tt avl\_fixed}
+			& 2,1 & \cpu{1079.03} 	& \ints{45078} 	\\
+			& 2,2 & \ETAdag{2762y}	& \ETAdag{8714863} 	\\
+			%[JOB 19] progress: 45863/8714863 brs (0.526262%), ETA 2762y 28d 9h 57m 14s (elapsed 59m 25s)
+			%[JOB 19] TIMED OUT (0.526262%; ETA 2762y 28d 9h 57m 14s)
+			% the below is from avlfixed2,2 after running it overnight
+			% eta is ever worse but i'm using the 1hr one as promised
+			%[JOB 19] progress: 315672/29992026 brs (1.052520%), ETA 321008y 136d 18h 22m 12s (elapsed 6h 58m 15s)
+			& 3,1 & \ETAdag{12d 2h}	& \ETAdag{11498545}	\\
+
 			\hline
 			{\tt map\_basic}
-			& 2,1 & \ETAdag{10d 16h} & \ETAdag{16388977} \\
+			& 2,1 & \ETAdag{10d 17h} & \ETAdag{16388977} \\
 			%[JOB 15] progress: 70330/16388977 brs (0.429130%), ETA 10d 16h 34m 15s (elapsed 43m 21s)
@@
 			& 2,1 & \cpu{877.44}& \ints{28635}	\\
-			& 2,2 & \ETAdag{2d 6h} & \ETAdag{5925634} \\
+			& 2,2 & \ETAdag{2d 7h} & \ETAdag{5925634} \\
+			& 3,1 & 	& 	\\ % TODO
+			% ^^ running now, certified not recompiled
 			%[JOB 6] progress: 35042/5925634 brs (0.591363%), ETA 2d 6h 40m 30s (elapsed 21m 12s)
@@
 in contrast with the high speed expected of bug-finding results,
-we believe the verification guarantee may be worth the estimated time requirement
+%we believe
+the verification guarantee may in some cases be worth the estimated time requirement
 for widely-used industrial implementations.
-Future work may expand our coverage upon this frontier \cite{landslide-phdthesis}.
+Future work may also expand our coverage upon this frontier \cite{landslide-phdthesis}.
 

### 2018-03-15 htm/paper.tex
move the variants tables so latex lays em out nicely
mark the winner in bugperf

@@
 		{\tt htm1}
-			& 2,1	&\cpu{  45.78}&\wtm{9.70}&\ints{21}& \cpu{9.47}& \wtm{6.40}& \ints{21	}& \ETA{213} \\
+			& 2,1	&\cpu{  45.78}&\wtm{9.70}&\ints{21}& \cpu{*9.47}& \wtm{*6.40}& \ints{21	}& \ETA{213} \\
 		(assertion)
-			& 2,2	&\cpu{  84.14}&\wtm{13.59}&\ints{33}& \cpu{10.39}& \wtm{7.70}& \ints{49	}& \ETA{1536} \\
-			& 2,3	&\cpu{ 131.91}&\wtm{20.44}&\ints{73}& \cpu{12.83}& \wtm{9.67}& \ints{113	}& \ETA{10752} \\
-			& 2,4	&\cpu{ 255.75}&\wtm{37.56}&\ints{257}& \cpu{18.63}& \wtm{15.86}& \ints{257	}& \ETA{73728} \\
+			& 2,2	&\cpu{  84.14}&\wtm{13.59}&\ints{*33}& \cpu{*10.39}& \wtm{*7.70}& \ints{49	}& \ETA{1536} \\
+			& 2,3	&\cpu{ 131.91}&\wtm{20.44}&\ints{*73}& \cpu{*12.83}& \wtm{*9.67}& \ints{113	}& \ETA{10752} \\
+			& 2,4	&\cpu{ 255.75}&\wtm{37.56}&\ints{257}& \cpu{*18.63}& \wtm{*15.86}& \ints{257	}& \ETA{73728} \\
 		\cline{2-9}
-			& 3,1	&\cpu{ 114.06}&\wtm{17.45}&\ints{15}& \cpu{9.50}& \wtm{6.79}& \ints{21	}& \ETA{13653} \\
-			& 3,2	&\cpu{ 109.60}&\wtm{26.16}&\ints{49}& \cpu{10.72}& \wtm{7.97}& \ints{49	}& \ETA{393216} \\
-			& 3,3	&\cpu{ 124.80}&\wtm{20.40}&\ints{73}& \cpu{13.84}& \wtm{11.01}& \ints{113	}& \ETA{11010048} \\
-			& 3,4	&\cpu{ 227.49}&\wtm{35.15}&\ints{161}& \cpu{31.37}& \wtm{28.53}& \ints{257	}& \ETA{301989888} \\
+			& 3,1	&\cpu{ 114.06}&\wtm{17.45}&\ints{*15}& \cpu{*9.50}& \wtm{*6.79}& \ints{21	}& \ETA{13653} \\
+			& 3,2	&\cpu{ 109.60}&\wtm{26.16}&\ints{49}& \cpu{*10.72}& \wtm{*7.97}& \ints{49	}& \ETA{393216} \\
+			& 3,3	&\cpu{ 124.80}&\wtm{20.40}&\ints{*73}& \cpu{*13.84}& \wtm{*11.01}& \ints{113	}& \ETA{11010048} \\
+			& 3,4	&\cpu{ 227.49}&\wtm{35.15}&\ints{*161}& \cpu{*31.37}& \wtm{*28.53}& \ints{257	}& \ETA{301989888} \\
 		\cline{2-9}
-			& 4,1	&\cpu{ 53.08}&\wtm{9.79}&\ints{15}& \cpu{9.82}& \wtm{7.00}& \ints{21	}& \ETA{873813} \\
-			& 4,2	&\cpu{117.07}&\wtm{19.09}&\ints{33}& \cpu{11.54}& \wtm{8.55}& \ints{49	}& \ETA{100663296} \\
+			& 4,1	&\cpu{ 53.08}&\wtm{9.79}&\ints{*15}& \cpu{*9.82}& \wtm{*7.00}& \ints{21	}& \ETA{873813} \\
+			& 4,2	&\cpu{117.07}&\wtm{19.09}&\ints{*33}& \cpu{*11.54}& \wtm{*8.55}& \ints{49	}& \ETA{100663296} \\
 		\hline
 		{\tt swapbug}
-			& 2,1	&\cpu{ 70.95}&\wtm{13.45}&\ints{16}& \cpu{38.96}& \wtm{13.15}& \ints{109	}& \ETA{194} \\
+			& 2,1	&\cpu{ 70.95}&\wtm{13.45}&\ints{*16}& \cpu{*38.96}& \wtm{*13.15}& \ints{109	}& \ETA{194} \\
 		(deadlock)
-			& 2,2	&\cpu{ 107.28}&\wtm{17.45}&\ints{146}& \cpu{44.73}& \wtm{19.47}& \ints{281	}& \ETA{1620} \\
-			& 2,3	&\cpu{ 280.05}&\wtm{38.70}&\ints{352}& \cpu{60.30}& \wtm{35.55}& \ints{718	}& \ETA{12748} \\
-			& 2,4	&\cpu{ 617.94}&\wtm{81.50}&\ints{834}& \cpu{108.58}& \wtm{82.60}& \ints{1820	}& \ETA{97823} \\
+			& 2,2	&\cpu{ 107.28}&\wtm{*17.45}&\ints{*146}& \cpu{*44.73}& \wtm{19.47}& \ints{281	}& \ETA{1620} \\
+			& 2,3	&\cpu{ 280.05}&\wtm{38.70}&\ints{*352}& \cpu{*60.30}& \wtm{*35.55}& \ints{718	}& \ETA{12748} \\
+			& 2,4	&\cpu{ 617.94}&\wtm{*81.50}&\ints{*834}& \cpu{*108.58}& \wtm{82.60}& \ints{1820	}& \ETA{97823} \\
 		\cline{2-9}
-			& 3,1	&\cpu{1275.04}&\wtm{163.42}&\ints{771}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{184984} \\
-			& 3,2	&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{3099225} \\
+			& 3,1	&\cpu{*1275.04}&\wtm{*163.42}&\ints{*771}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{184984} \\
+			& 3,2	&\ETA{--}&\ETA{>30m}&\ETA{--}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{3099225} \\
 		% whoops, this was with the dumb lock ordering
@@
 		{\tt avl\_insert}
-			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{83}	&\cpu{81.00}&\wtm{40.30}&\ints{336}&\ETA{379982} \\
+			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{*83}	&\cpu{*81.00}&\wtm{*40.30}&\ints{336}&\ETA{379982} \\
 		(segfault+)
@@
 			% according to variance expt, it should be rerun for a *minor* speedup, but i can't be bothered
-			& 2,3	&\cpu{2670.87}&\wtm{330.45}&\ints{3066}	&\cpu{1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
-			& 2,4	&\cpu{3259.37}&\wtm{436.50}&\ints{1639}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{???} \\
-			% TODO
+			& 2,3	&\cpu{2670.87}&\wtm{*330.45}&\ints{*3066}&\cpu{*1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
+			& 2,4	&\cpu{*3259.37}&\wtm{436.50}&\ints{*1639}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{36019973} \\
+			% smaller than above? estimate's just fickle i guess; they can go up or down with time
 		\cline{2-9}
-			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{28}	&\cpu{69.99}&\wtm{24.25}&\ints{78}&\ETA{1572107} \\
-			& 3,2	&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}&\ETA{>30m}&\ETA{>30m}&\ETA{unk.}& \ETA{???} \\
-			% TODO
+			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{*28}	&\cpu{*69.99}&\wtm{*24.25}&\ints{78}&\ETA{1572107} \\
+			& 3,2	&\cpu{*1569.09}&\wtm{*216.85}&\ints{*209}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{1402363529} \\
 	\end{tabular}
@@
 		before the bug was found.
+		* marks the winning measurements between each series.
 		Lastly, Landslide's state space size estimation \cite{estimation},
@@
 		Most of our tests are too small for its approach to pay off,
-		with {\tt swapbug}(3,1) and {\tt avl\_insert}(2,3) as its notable wins.
+		with {\tt swapbug}(3,1) and {\tt avl\_insert} as its notable wins.
 		While the prior work showed plenty more wins \cite{quicksand},
@@
 % for i in `seq 1 10`; do ./landslide -X -v -p counter -M 2>&1 | grep 'total CPU time'; done
+
+Table~\ref{tab:variants2} shows
+the impact of Landslide needing to automatically annotate {\tt counter}(2,2) being run for the first time,
+rather than reusing existing instrumentation,
+as well as the variance of Quicksand mode. %(i.e., without {\tt -M}).
+%(2128 interleavings total).
+The {\sf quicksand} variance was surprisingly bimodal,
+with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142.11$\pm$1.67,
+suggesting two distinct scheduling patterns for its workqueue threads.
+Future work should figure out why.
+
+We conclude that MC performance evaluations must address
+experimental environment variables
+%at least enough
+to ensure consistent performance between runs.
+So doing, multiple samples and error bars are then necessary
+only when using nondeterministic search ordering strategies
+such as Iterative Deepening.
+Otherwise, considering the low variance shown in Table~\ref{tab:variants},
+they need be shown only for a token small test
+to promise the reader the same consistency in the larger tests
+(especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
+% We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper
+%% i'm not actually sure this fits..
+%Finally, considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
+%we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
+
+Outside of this experiment, we fixed our environment by measuring all performance numbers
+with Firefox and Chrome as the only other significant machine load.
+%(we gotta stay amused somehow).
+We believe the exponential differences among completion times justifies the absence of error bars,
+which would be expected to show 2\% variance if extrapolating from this experiment.
+The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Limitations}
+\label{sec:warpzone}
+
+%This section discusses opportunities for future work.
+This section should be mandatory in all systems papers.
+
+{\bf Transaction failure codes.}
+When a transaction fails, {\tt \_xbegin()} returns a failure code
+denoting the reason, or combination thereof, therefor \cite{htm-gcc}.
+If a program then cases on that failure code to select between different backup paths,
+model checking it by simply injecting a single type of failure may be unsound.
+For example, a program executing a transaction which is guaranteed to never conflict with any other threads,
+and hence never abort without {\tt \_XABORT\_RETRY},
+could legally {\tt assert(false);} in its failure path,
+while our approach in Section~\ref{sec:design} would erroneously trigger that assertion and report a bug.
+Likewise, the transactional data structures from \cite{htm-mario}
+abstract away any spurious {\tt \_XABORT\_RETRY} aborts behind a retry loop;
+a MC unwise to that idiom would call it an infinite loop bug.
+
+Our HTM implementation includes an experimental feature
+to track the set of abort codes possible for each transaction.
+%According to the any-reason principle in Section~\ref{sec:background},
+{\tt \_XABORT\_RETRY} is always enabled;
+then,
+it harnesses DPOR's existing memory analysis to identify when {\tt \_XABORT\_CONFLICT} is possible,
+and instruments {\tt \_xabort()} calls to record any user-supplied codes for {\tt \_XABORT\_EXPLICIT}.
+This comes at a cost of even more state space explosion,
+increasing the exponent at each {\tt \_xbegin()} preemption point
+by (usually) 1 plus however many distinct {\tt \_xabort()} codes the program uses.
 
@@
 
-Table~\ref{tab:variants2} shows
-the impact of Landslide needing to automatically annotate {\tt counter}(2,2) being run for the first time,
-rather than reusing existing instrumentation,
-as well as the variance of Quicksand mode. %(i.e., without {\tt -M}).
-%(2128 interleavings total).
-The {\sf quicksand} variance was surprisingly bimodal,
-with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142.11$\pm$1.67,
-suggesting two distinct scheduling patterns for its workqueue threads.
-Future work should figure out why.
-
 \begin{table}[t]
@@
 \end{table}
-
-We conclude that MC performance evaluations must address
-experimental environment variables
-%at least enough
-to ensure consistent performance between runs.
-So doing, multiple samples and error bars are then necessary
-only when using nondeterministic search ordering strategies
-such as Iterative Deepening.
-Otherwise, considering the low variance shown in Table~\ref{tab:variants},
-they need be shown only for a token small test
-to promise the reader the same consistency in the larger tests
-(especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
-% We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper
-%% i'm not actually sure this fits..
-%Finally, considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
-%we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
-
-Outside of this experiment, we fixed our environment by measuring all performance numbers
-with Firefox and Chrome as the only other significant machine load.
-%(we gotta stay amused somehow).
-We believe the exponential differences among completion times justifies the absence of error bars,
-which would be expected to show 2\% variance if extrapolating from this experiment.
-The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Limitations}
-\label{sec:warpzone}
-
-%This section discusses opportunities for future work.
-This section should be mandatory in all systems papers.
-
-{\bf Transaction failure codes.}
-When a transaction fails, {\tt \_xbegin()} returns a failure code
-denoting the reason, or combination thereof, therefor \cite{htm-gcc}.
-If a program then cases on that failure code to select between different backup paths,
-model checking it by simply injecting a single type of failure may be unsound.
-For example, a program executing a transaction which is guaranteed to never conflict with any other threads,
-and hence never abort without {\tt \_XABORT\_RETRY},
-could legally {\tt assert(false);} in its failure path,
-while our approach in Section~\ref{sec:design} would erroneously trigger that assertion and report a bug.
-Likewise, the transactional data structures from \cite{htm-mario}
-abstract away any spurious {\tt \_XABORT\_RETRY} aborts behind a retry loop;
-a MC unwise to that idiom would call it an infinite loop bug.
-
-Our HTM implementation includes an experimental feature
-to track the set of abort codes possible for each transaction.
-%According to the any-reason principle in Section~\ref{sec:background},
-{\tt \_XABORT\_RETRY} is always enabled;
-then,
-it harnesses DPOR's existing memory analysis to identify when {\tt \_XABORT\_CONFLICT} is possible,
-and instruments {\tt \_xabort()} calls to record any user-supplied codes for {\tt \_XABORT\_EXPLICIT}.
-This comes at a cost of even more state space explosion,
-increasing the exponent at each {\tt \_xbegin()} preemption point
-by (usually) 1 plus however many distinct {\tt \_xabort()} codes the program uses.
 

### 2018-03-15 htm/paper.tex
finished, minus proofreading

@@
 			& 2,2 & \ETAdag{2d 7h} & \ETAdag{5925634} \\
-			& 3,1 & 	& 	\\ % TODO
-			% ^^ running now, certified not recompiled
+			& 3,1 & \ETAdag{468d 13h} & \ETAdag{35893653} \\
+			% [JOB 5] progress: 100181/35893653 brs (0.279105%), ETA 1y 103d 12h 50m 47s (elapsed 59m 24s)
+			%[JOB 5] TIMED OUT (0.279105%; ETA 1y 103d 12h 50m 47s)
+			% i don't know what this is but it was hanging out under 3,1 here :<
 			%[JOB 6] progress: 35042/5925634 brs (0.591363%), ETA 2d 6h 40m 30s (elapsed 21m 12s)
@@
 Though these tests proved beyond our reach,
-in contrast with the high speed expected of bug-finding results,
+in contrast with the instant gratification expected of bug-finding,
 %we believe
@@
 
-We noticed significant slowdowns when varying several factors of our experimental environment.
-For example, multiple Landslides running at once on our test machine slow each other down,
+We noticed significant slowdowns when varying several aspects of our experimental environment.
+For example, multiple Landslides running at once
+%on our test machine
+slow each other down,
 %(including multiple single state spaces in a single test when not using {\tt -M}),
@@
 Table~\ref{tab:variants} shows the impact of running a single Landslide instance with {\tt -M} on
-{\tt counter}(2,2) (384 distinct interleavings)
+{\tt counter}(2,2) %(384 distinct interleavings)
 with various programs also running, despite never saturating our test system's 8 CPUs.
@@
 as well as the variance of Quicksand mode. %(i.e., without {\tt -M}).
+Because Quicksand repeats more work across jobs,
+rather than comparing it to a baseline,
+we show in the right column how many total interleavings each approach executed.
+(The maximal state space alone comprises 384 in all cases.)
 %(2128 interleavings total).
@@
 and the perennial {\tt \_XABORT\_RETRY} abort can be skipped entirely
-if the client abstracts it away behind a retry loop.%
+if the client abstracts it away behind a retry loop.
+In fact, applying both reductions simultaneously
+amounts to the STM concurrency model \cite{stm-pldi06}%
 \footnote{Proof left to future work. We've proved enough here already.}
-In fact, applying both reductions simultaneously
-produces a state space smaller than the original;
-testing them by hand
+and may reduce the state space even smaller than the original.
+Testing them by hand
 \footnote{Using visual inspection of the test to trust the reductions' soundness.}
-reduced {\tt map\_basicer}(2,1)'s 28635 interleavings to 11577.
-% TODO: try map_basic and get a 1 hour estimate
+reduced {\tt map\_basicer}(2,1)'s 28635 interleavings to 11577,
+and produced the same 384-interleaving state space on {\tt counter}(2,2)
+(in which DPOR triggers conflict aborts on every transaction anyway).
+%and {\tt map\_basic}(2,1)'s 1-hour size estimate from 16388977 to 16126960. % about the same after all...
 %(although {\tt map\_basic}(2,1) lies still beyond reach).
@@
 for now, this feature is disabled by default but accessible via the {\tt -A} command-line option
-to Landslide (in addition to {\tt -X}).
+%to Landslide
+(in addition to {\tt -X}).
 
@@
 even after \cite{htm-mario}'s proposed fix,
-our running example program is still incorrect on Total Store Ordering (TSO) architectures such as x86.
+our running example program is still incorrect on Total Store Order (TSO) architectures such as x86.
 Despite stores being totally-ordered, x86 may still reorder stores after subsequent loads.
@@
 An {\tt mfence} barrier is needed between lines 8 and 9 to solve this problem on TSO \cite{tsx-need-barrier}.
-On Partial Store Ordering (PSO) architectures, even more barriers may be necessary.
+On Partial Store Order (PSO) architectures, even more barriers may be necessary.
 
@@
 This work was supported in part by the U.S. Army Research Office under grant number W911NF0910273.
+\vspace{1em}
 

### 2018-03-16 htm/paper.tex
proofreading pass

@@
 fast-path code must be accompanied by backup paths to handle transaction failure.
-
+%
 We extend Landslide, a popular stateless model checker,
@@
 If no other CPU accesses the same memory during the transaction,
-the access sequence is committed to main memory atomically (with respect to visibility to other CPUs).
+the access sequence is committed to main memory atomically (with respect to visibility by other CPUs).
 Otherwise, the accesses are discarded, the CPU's local state is reverted, and the transaction returns a failure code.
@@
 % だからこそ
-obscure interleavings between failure paths are difficult to expose during testing.
+obscure interleavings between failure paths are difficult to expose during stress testing.
 
@@
 {\bf Hardware transactional memory.}
-TSX was supported on consumer hardware for the first time by Intel's Haswell architecture \cite{htm-haswell},
+TSX was implemented on consumer hardware for the first time by Intel's Haswell architecture \cite{htm-haswell},
 which extends the x86 instruction set to provide
@@
 \newcommand\call[1]{\hilight{pinkish}{#1}}
+
+Under software transactional memory (STM) \cite{stm-pldi06},
+memory conflicts with other threads are the only reason for transaction failure
+(apart from programmer-supplied explicit aborts);
+hence, depending on program semantics, some transactions may be guaranteed to succeed.
+However, hardware transactions (HTM) may also fail
+for several other reasons such as random system interrupts or exhausting the CPU's cache capacity.
+Because timer interrupts can in principle occur at any moment,
+and with arbitrary frequency (observable by the program, perhaps as a result of a heavily-loaded system),
+in this paper we will simplify the failure model by saying that HTM transactions can fail for any reason.
+We defer discussion of programs which distinguish the reason for aborts through the failure code to Section~\ref{sec:warpzone}.
+
+{\bf Stateless model checking.}
+Model checking (MC) \cite{verisoft} is a testing technique for systematically executing and verifying
+the possible thread interleavings of a concurrent program.
+The main research challenge is to cope with exponential explosion of the state space,
+which is sized $O(n^k)$ for a program with $n$ operations and $k$ threads.
+Some {\em stateful} MCs explicitly store and compare all visited states of the program being tested \cite{spin},
+which both keeps track of test coverage
+and allows identifying identical states to avoid testing redundant interleavings.
+%
+By contrast, {\em stateless} MC (henceforth abbreviated simply as MC)
+stores only the current sequence of execution events to avoid a prohibitive memory footprint.
+Reduction algorithms \cite{dpor,optimal-dpor,satcheck,mcr,mcr,tsopso}
+can then analyze the memory accesses in that sequence to identify
+interleavings observationally-equivalent under Mazurkiewicz trace theory \cite{mazurkiewicz} and hence safe to skip.
+The resulting state spaces are still exponentially-sized,
+but only in the number of conflicting operations rather than all operations.
+Of these, Landslide uses Dynamic Partial Order Reduction (DPOR) \cite{dpor} to prune its state spaces.
 
@@
 
-Under software transactional memory (STM) \cite{stm-pldi06},
-memory conflicts with other threads are the only reason for transaction failure
-(apart from programmer-supplied explicit aborts);
-hence, depending on program semantics, some transactions may be guaranteed to succeed.
-However, hardware implementations (HTM) may also fail transactions
-for several other reasons such as random system interrupts or exhausting the CPU's cache capacity.
-Because timer interrupts can in principle occur at any moment,
-and with arbitrary frequency (observable by the program, perhaps as a result of a heavily-loaded system),
-in this paper we will simplify the failure model by saying that HTM transactions can fail for any reason.
-We defer discussion of programs which distinguish the reason for aborts through the failure code to Section~\ref{sec:warpzone}.
-
-{\bf Stateless model checking.}
-Model checking (MC) \cite{verisoft} is a testing technique for systematically executing and verifying
-the possible thread interleavings of a concurrent program.
-The main research challenge is to cope with exponential explosion of the state space,
-which is $O(n^k)$ for a program with $n$ operations and $k$ threads.
-Some {\em stateful} MCs explicitly store and compare all visited states of the program being tested \cite{spin},
-which both keeps track of test coverage
-and allows identifying identical states to avoid testing redundant interleavings.
-%
-By contrast, {\em stateless} MC (henceforth abbreviated simply as MC)
-stores only the current sequence of execution events to avoid a prohibitive memory footprint.
-Reduction algorithms \cite{dpor,optimal-dpor,satcheck,mcr,mcr,tsopso}
-can then analyze the memory accesses in that sequence to identify
-interleavings observationally-equivalent under Mazurkiewicz trace theory \cite{mazurkiewicz} and hence safe to skip.
-The resulting state spaces are still exponentially-sized,
-but only in the number of conflicting operations rather than all operations.
-Of these, Landslide uses Dynamic Partial Order Reduction (DPOR) \cite{dpor} to prune its state spaces.
-
 MCs may instrument programs to introduce thread switches at varying granularity, which affects the number of operations $n$.
@@
 We make two major simplifications:
-simulating transaction aborts as failure injections,
+simulating transaction aborts as immediate failure injections,
 and treating transaction atomicity as a global mutex during data-race analysis;
@@
 denote the execution sequence of a program as it runs according to the specified thread interleaving.%
-\footnote{This serialization of concurrent execution is told from the perspective of main memory
+\footnote{This serialization of concurrent execution is told from the perspective of all CPUs at once
 and hence assumes sequential consistency.
@@
 to exclude this family of interleavings,
-showing that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
+which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
 %further motivating our research.
@@
 Left unstated in interleavings such as $\tiat{1-2c}, \tjat{1-3}, \tiat{4-7}$%
-\footnote{For the sake of example I reorder \ti's write to {\tt x} before \tj's part here.}
+\footnote{For a clearer example we reorder \ti's write to {\tt x} before \tj's part here.}
 are HTM's execution semantics, namely:
@@
 model checking such programs need be concerned only with the program's {\em observable} behaviours.
-We claim that MCing the interleaving $\tiat{1}, \tjat{1-3}, \tiat{4-7}$
+We claim that MCing the simpler interleaving $\tiat{1}, \tjat{1-3}, \tiat{4-7}$
 is an equivalent verification as MCing the one above;
@@
 From Theorem~\ref{thm:atom} we have that the body of all transactions may be executed atomically within the MC environment.
-While they may interleave within other non-transactional sequences,
+While they may interleave between other non-transactional sequences,
 no other operations (whether transactional or not) will interrupt them.
@@
 	\item {\bf Limited Happens-Before.}
-		Under this definition, to be a data race in $I$ they must be reorderable at instruction granularity
+		%Under this definition,
+		To race in $I$ they must be reorderable at instruction granularity,
+		at least one with a thread switch immediately before or after.
 		\cite{tsan,hybriddatarace}.
@@
 		WLOG fix $\tiiat{\mu} \prec \tjjat{\nu} \in I$.
-		Then under this definition, to be a data race in $I$ there must be no pair of synchronizing instructions
+		Then to race in $I$ there must be no pair of synchronizing instructions
 		$\tiiat{\epsilon}$ (a release edge) and $\tjjat{\chi}$ (an acquire edge) such that
@@
 			& 2,3	&\cpu{2670.87}&\wtm{*330.45}&\ints{*3066}&\cpu{*1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
-			& 2,4	&\cpu{*3259.37}&\wtm{436.50}&\ints{*1639}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{36019973} \\
+			& 2,4	&\cpu{*3259.37}&\wtm{*436.50}&\ints{*1639}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{36019973} \\
 			% smaller than above? estimate's just fickle i guess; they can go up or down with time
@@
 %To the best of the author's knowledge, Landslide is the first MC to check transactional concurrency,
-so no other MC State-of-the-Art%
+so no other MC State of the Art%
 \footnote{The author's DJ name.}
@@
 		{\bf Finding bugs quickly.}
-		The multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
-		than that of the total number of interleavings (10-100x)
-		as the test parameters increase.
+		As the test parameters increase,
+		the multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
+		than that of the total number of interleavings (10-100x).
 		In other words,
@@
 	%\caption{mario man had a very large house. or a castle.}
-	\caption{Unmodified code from {\tt htmavl.cpp} with the previously-unknown bug found by Landslide.
+	\caption{Unmodified code from {\tt htmavl.cpp} showing the previously-unknown bug found by Landslide.
 		The transaction path fails to check {\tt \_retry},
@@
 			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
-			%& 3,2 &	&	\\
+			& 3,2 &	&	\\ % TODO currently running
 			\hline
@@
 			& 3,1 & \cpu{57.90}	& \ints{1960}	\\
-			%& 3,2 &		& 	\\
+			& 3,2 &		& 	\\ % TODO
 			\hline
@@
 they need be shown only for a token small test
-to promise the reader the same consistency in the larger tests
+to provide reader some assurance of similar consistency in the larger tests
 (especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
@@
 
-Outside of this experiment, we fixed our environment by measuring all performance numbers
+%Outside of this experiment,
+For all tests outside of Table~\ref{tab:variants},
+we fixed our environment by measuring all performance numbers
 with Firefox and Chrome as the only other significant machine load.
@@
 We believe the exponential differences among completion times justifies the absence of error bars,
-which would be expected to show 2\% variance if extrapolating from this experiment.
+which one would expect to show 2\% variance if extrapolating from these results. %this experiment.
 The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.
@@
 not per-CPU memory buffer reorderings,
-our current HTM implementation in Landslide cannot find this bug.
-In fact, it erroneously verifies the corresponding test {\tt htm2} in 7 CPU-minutes,
-with 294 distinct interleavings in total,
+our current HTM implementation cannot find this bug.
+In fact, it erroneously verifies the corresponding test {\tt htm2}(3,1)
+% lol this is both old (quicksand) AND wrong (testing (2,1) which can't expose the bug)
+%in 7 CPU-minutes,
+%with 294 distinct interleavings in total,
+in 3 CPU-hours,
+with 467730 distinct interleavings in total,
 none of which include the above-listed sequence.
@@
 Visual inspection of \cite{htm-mario}'s HTM data structures found no barriers used in this implementation pattern;
-we would urge any reader interested in using them to add them in by hand first.
+we would urge any reader interested in using those to add them in by hand first.
 
@@
 
-We thank Mario Dehesa-Azuara for generously providing the HTM data structures and benchmarks used in our evaluation.
+We thank Mario Dehesa-Azuara for generously providing the HTM data structures used in our evaluation,
+% no
+%and the anonymous SIGBOVIK reviewers for their helpful comments.
+the anonymous SIGBOVIK program committee,
+and anyone who actually read this meta-joke of a non-joke paper in entirety.
 %We also thank the member companies of the PDL Consortium (Broadcom,
@@
 This work was supported in part by the U.S. Army Research Office under grant number W911NF0910273.
-\vspace{1em}
 

### 2018-03-16 htm/paper.tex
done

@@
 \definecolor{lavender}{RGB}{102,65,208} % (?? i don't know what happened here anymore but it looks good)
-\definecolor{seafoam}{RGB}{65,178,76} % * 3/4
+\definecolor{seafoam}{RGB}{56,158,68} % * 2/3
 \definecolor{salmon}{RGB}{208,89,74} % * 7/8
@@
 Consider again the program in Figure~\ref{fig:htm-example}.
-Note that the C-style {\tt x++} operations, when compiled into assembly,
+Note that the C-style {\tt x++} operations, when compiled into assembly \cite{printy},
 results in multiple memory accesses which can be interleaved with other threads.
@@
 			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
-			& 3,2 &	&	\\ % TODO currently running
+			& 3,2 &	\ETAdag{10d 14h}& \ETAdag{13763999}	\\
+			%[JOB 4] progress: 127032/13763999 brs (0.922929%), ETA 10d 13h 52m 44s (elapsed 59m 56s)
+			%[JOB 4] TIMED OUT (0.922929%; ETA 10d 13h 52m 44s)
 			\hline
@@
 			& 2,2 & \cpu{15.53}	& \ints{384}	\\
-			& 2,3 &\cpu{20.00}	& \ints{5280}	\\
+			& 2,3 &\cpu{155.00}	& \ints{5280}	\\
 			& 2,4 &\cpu{2211.10}	& \ints{75264}	\\
 			& 3,1 & \cpu{57.90}	& \ints{1960}	\\
-			& 3,2 &		& 	\\ % TODO
+			& 3,2 & \cpu{10028.93}	& \ints{329888}	\\ % TODO running atm
 			\hline
@@
 the anonymous SIGBOVIK program committee,
-and anyone who actually read this meta-joke of a non-joke paper in entirety.
+and anyone who actually read this unusual paper all the way through.
 %We also thank the member companies of the PDL Consortium (Broadcom,

### 2018-03-29 htm/for-garth/paper.tex
add for-garth version

--- 2018-03-16 htm/paper.tex
@@
 %\documentclass[preprint]{sigplanconf}
-\documentclass[10pt]{sigplanconf}
+\documentclass{article}
 
@@
 \usepackage{xspace}
+\usepackage{fullpage}
 \usepackage[mathcal]{euscript}
@@
 
-\conferenceinfo{SIGBOVIK '18}{Pittsburgh, PA, USA}
-\copyrightyear{2018}
-\copyrightdata{}
-
-\titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{short description of paper}   % 'preprint' option specified.
+%\conferenceinfo{SIGBOVIK '18}{Pittsburgh, PA, USA}
+%\copyrightyear{2018}
+%\copyrightdata{}
+%
+%\titlebanner{banner above paper title}        % These are ignored unless
+%\preprintfooter{short description of paper}   % 'preprint' option specified.
 
@@
 
-\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}
+\author{Ben Blum \\ bblum@cs.cmu.edu}
+\date{}
 
@@
 
-\begin{abstract}
-Hardware transactional memory is a recently-introduced concurrent programming paradigm
-%backed by a set of new instructions on recent x86 CPUs,
-which allows programmers to elide locks for performance in low-contention workloads.
-However,
-%using this feature
-it
-comes at a cost in implementation complexity:
-%programmers must support their
-fast-path code must be accompanied by backup paths to handle transaction failure.
-%
-We extend Landslide, a popular stateless model checker,
-with a concurrency model for transactional memory
-and evaluate it on several real-world transactional
-benchmarks and data structure implementations.
-\end{abstract}
-
-\category{D.1.3}{Programming Techniques}{Concurrent Programming}
-\category{D.2.4}{Software Engineering}{Software/Program Verification}
-
-\keywords
-landslide terminal, baggage claim, ground transportation, ticketing
-
-\section{Introduction}
-\label{sec:intro}
-
-Transactional Synchronization Extensions (TSX) \cite{transactional-memory}
-is an instruction set extension for x86 CPUs which adds hardware-based transactional memory.
-The processor uses its existing cache coherence algorithm to
-check for memory conflicts with other cores while temporarily staging a sequence of memory accesses.
-If no other CPU accesses the same memory during the transaction,
-the access sequence is committed to main memory atomically (with respect to visibility by other CPUs).
-Otherwise, the accesses are discarded, the CPU's local state is reverted, and the transaction returns a failure code.
-
-This feature can be used to replace conventional locking in performance-critical concurrent programs.
-When a concurrent workload accesses largely thread-local data,
-or disjoint sections of a shared data structure,
-the contention rate between threads is low,
-and transactions will often succeed.
-Compared to programs which use conventional locks,
-which use bus-locking atomic accesses even in the fastest code path,
-TSX provides substantial performance improvements in such programs \cite{htm-experience, htm-performance, htm-mario}.
-However, the possibility of transaction failure introduces additional implementation complexity:
-programmers must also provide a backup plan
-to safely resolve contention between threads,
-usually involving conventional synchronization.
-These backup paths must coordinate not only with other backup paths
-but also with other fast paths which another thread may begin after the original transaction failed,
-which even in the simplest transactions requires complex synchronization sequences
-\cite{htm-mario}.
-This introduces an additional dimension of nondeterminism into an already concurrent program,
-and moreover, because transaction failure is expected to be rare,
-% だからこそ
-obscure interleavings between failure paths are difficult to expose during stress testing.
-
-This motivates the use of stateless model checking (MC) \cite{verisoft}
-to comprehensively verify these transactional programs,
-fast paths failure paths and all.
-MC aims to force the system to execute all possible thread interleavings under a given test case,
-exhaustively checking for bugs or verifying their absence in the corresponding state space.
-Many such model checkers exist, varying in
-interleaving granularity, memory analysis, types of programs checked, and search ordering strategy
-\cite{chess,dbug-ssv,spin,inspect,r4,portend,samc,mcr,randomized-scheduler}.
-This work builds upon Landslide \cite{landslide}, %,quicksand,landslide-phdthesis},
-a simulator-based tester which checks both user- and kernel-level programs
-and incorporates data-race analysis \cite{tsan,fasttrack} to find new preemption points
-at memory access granularity.
-Our contributions are as follows:
-
-\begin{enumerate}
-	\item We extend Landslide's concurrency model to include transaction failure as an additional source of nondeterminism;
-	\item We provide a proof sketch that our implementation matches TSX's execution semantics,
-	\item We evaluate the extended Landslide on several transactional programs, analyzing both its bug-finding and verification performance.
-\end{enumerate}
-
-The paper is organized as follows.
-Section \ref{sec:intro} introduces the problem domain and motivates our research.
-The other sections state the rest of the paper.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\category{D.1.3}{Programming Techniques}{Concurrent Programming}
+%\category{D.2.4}{Software Engineering}{Software/Program Verification}
+%
+%\keywords
+%landslide terminal, baggage claim, ground transportation, ticketing
 
@@
 
-This section introduces the fundamental concepts and prior work
-in both hardware transactional memory and stateless model checking,
-which we propose to combine.
-
-{\bf Hardware transactional memory.}
-TSX was implemented on consumer hardware for the first time by Intel's Haswell architecture \cite{htm-haswell},
+\subsection{Hardware transactional memory}
+
+Transactional Synchronization Extensions (TSX) was implemented on consumer hardware for the first time by Intel's Haswell architecture \cite{htm-haswell},
 which extends the x86 instruction set to provide
@@
 including a failure path which defaults to a conventional lock.
-This example actually has a bug, which we will discuss in the next section;
-the reader is encouraged to try to spot it before then.
+%This example actually has a bug, which we will discuss in the next section;
+%the reader is encouraged to try to spot it before then.
 Several related works formally prove the correctness of transactional memory {\em implementations}
@@
 in this paper we will simplify the failure model by saying that HTM transactions can fail for any reason.
-We defer discussion of programs which distinguish the reason for aborts through the failure code to Section~\ref{sec:warpzone}.
-
-{\bf Stateless model checking.}
+%We defer discussion of programs which distinguish the reason for aborts through the failure code to Section~\ref{sec:warpzone}.
+
+\subsection{Stateless model checking}
+
 Model checking (MC) \cite{verisoft} is a testing technique for systematically executing and verifying
@@
 but only in the number of conflicting operations rather than all operations.
-Of these, Landslide uses Dynamic Partial Order Reduction (DPOR) \cite{dpor} to prune its state spaces.
+%Of these, Landslide uses Dynamic Partial Order Reduction (DPOR) \cite{dpor} to prune its state spaces.
 
@@
 		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
-		1 & \texttt{\flow{if} ((status = \call{\_xbegin}()) == \const{SUCCESS}) \{} \\
+		1 & \texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
 		2 & \texttt{~~~~x++;} \\
@@
 	\end{center}
-	\caption{Example transactional program.
-		%The example {\tt count} routine from Figure~\ref{fig:mutex}, rewritten to use HTM.
+	\caption{Example transactional programs.
 			If the
@@
 			%, such as a mutex,
-		to resolve the conflict.}
+			to resolve the conflict.
+		}
 	\label{fig:htm-example}
 \end{figure}
-
 MCs may instrument programs to introduce thread switches at varying granularity, which affects the number of operations $n$.
@@
 and some insert compiler instrumentation on statically-identified memory accesses \cite{chess,inspect}.
-Landslide traces every memory access through the use of a simulated environment \cite{bochs},
+Our model checker, Landslide \cite{landslide},
+traces every memory access through the use of a simulated environment \cite{bochs},
 which is important for identifying data races to use as new preemption points \cite{quicksand},
@@
 as well as deadlocks, use-after-frees, and segfaults.
-For this work we also detect use of {\tt xend} outside of a transaction
-or {\tt xbegin} within one as a bug.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Design}
-\label{sec:design}
-
-This section presents our formalization of transactional memory in Landslide's framework of thread concurrency.
-We make two major simplifications:
-simulating transaction aborts as immediate failure injections,
-and treating transaction atomicity as a global mutex during data-race analysis;
-and provide corresponding equivalence proofs.
-
-{\bf Notation.} Let $I = TN_1@L_1, TN_2@L_2, ... TN_n@L_n$,
-with $N_i$ a thread ID and $L_i$ a code line number,
-denote the execution sequence of a program as it runs according to the specified thread interleaving.%
-\footnote{This serialization of concurrent execution is told from the perspective of all CPUs at once
-and hence assumes sequential consistency.
-For discussion of relaxed memory models refer to Section~\ref{sec:warpzone}.}
-
-\subsection{Example}
+%For this work we also detect use of {\tt xend} outside of a transaction
+%or {\tt xbegin} within one as a bug.
+
+\subsection{Example bug}
+\label{sec:example}
+
+\newcommand\tii{\ensuremath{\hilight{lavender}{\mathbf{Ti}}}\xspace}
+\newcommand\tjj{\ensuremath{\hilight{seafoam}{\mathbf{Tj}}}\xspace}
+\newcommand\tkk{\ensuremath{\hilight{salmon}{\mathbf{Tk}}}\xspace}
+
+\newcommand\tiiat[1]{\ensuremath{\hilight{lavender}{\mathbf{Ti}@#1}}\xspace}
+\newcommand\tjjat[1]{\ensuremath{\hilight{seafoam} {\mathbf{Tj}@#1}}\xspace}
+\newcommand\tkkat[1]{\ensuremath{\hilight{salmon}  {\mathbf{Tk}@#1}}\xspace}
 
 Consider again the program in Figure~\ref{fig:htm-example}.
-Note that the C-style {\tt x++} operations, when compiled into assembly \cite{printy},
+Note that the C-style {\tt x++} operations, when compiled into assembly.
 results in multiple memory accesses which can be interleaved with other threads.
@@
 
-% T1 T2 T3 but latex doesn't allow #s in cmd names
-\newcommand\ti{\ensuremath{\hilight{lavender}{\mathbf{T1}}}\xspace}
-\newcommand\tj{\ensuremath{\hilight{seafoam}{\mathbf{T2}}}\xspace}
-\newcommand\tk{\ensuremath{\hilight{salmon}{\mathbf{T3}}}\xspace}
-
-\newcommand\tiat[1]{\ensuremath{\hilight{lavender}{\mathbf{T1}@#1}}\xspace}
-\newcommand\tjat[1]{\ensuremath{\hilight{seafoam}{\mathbf{T2}@#1}}\xspace}
-\newcommand\tkat[1]{\ensuremath{\hilight{salmon}{\mathbf{T3}@#1}}\xspace}
-
-If these instructions from the {\tt x++} in the transaction are preempted,
-with another thread's access to {\tt x} interleaved in between,
-the transaction will abort.
-So, the interleaving
-\[
-	\tiat{1}, \tiat{2a}, \tiat{2b}, \tjat{1}, \tjat{2}, \tjat{3}, \tiat{2c}, \tiat{3}
-\]
-or, henceforth abbreviated for clarity:
-\[
-	\tiat{1-2b}, \tjat{1-3}, \tiat{2c-3}
-\]
-is not possible; rather, \ti will fall into the backup path:
-\[
-	\tiat{1-2b}, \tjat{1-3}, \tiat{4-7}
-\]
-However, the {\tt x++} operation from the failure path (correspondingly $6a$, $6b$, $6c$)
-{\em can} be thusly separated with conflicting accesses interleaved in between,
+The {\tt x++} operation from the failure path
+%(correspondingly $6a$, $6b$, $6c$)
+can by interrupted with conflicting accesses from the transaction path interleaved in between,
 since the mutex only protects the failure path against other failure paths,
 but not against the transaction itself.
-So (assuming {\tt x} is intended to be a precise counter rather than a sloppy one),
-%losing one of the increments to which constitutes a bug),
-we observe a bug in the following interleaving.%
-\footnote{Note also that this bug requires either at least 3 threads or at least 2 iterations between 2 threads to expose;
-this highlights MC's dependence on its test cases to produce meaningful state spaces in the first place.}
+Hence an interleaving such as the following is possible,
+in which only two increments of {\tt x} are observed while three are expected.
+
 \[
-	\tiat{1-2b}, \tjat{1-3}, \tiat{4-6b}, \tkat{1-3}, \tiat{6c-7}
+\begin{tabular}{ll|ll|ll}
+	\multicolumn{2}{c|}{\hilight{lavender}{Thread 1}} &
+	\multicolumn{2}{c|}{\hilight{seafoam}{Thread 2}} &
+	\multicolumn{2}{c}{\hilight{salmon}{Thread 3}} \\
+	\hline
+	\hilight{lavender}{$1$} & \hilight{lavender}{\texttt{\_xbegin();}} & & \\
+	%
+	& & \hilight{seafoam}{$1$} & \hilight{seafoam}{\texttt{\_xbegin();}} \\
+	& & \hilight{seafoam}{$2$} & \hilight{seafoam}{\texttt{x++;}} \\
+	& & \hilight{seafoam}{$3$} & \hilight{seafoam}{\texttt{\_xend();}} \\
+	%
+	\hilight{lavender}{$5$} & \hilight{lavender}{\texttt{mutex\_lock(m);}} & & \\
+	\hilight{lavender}{$6a$} & \hilight{lavender}{\texttt{temp <- x;}} & & \\
+	%
+	& & & & \hilight{salmon}{$1$} & \hilight{salmon}{\texttt{\_xbegin();}} \\
+	& & & & \hilight{salmon}{$2$} & \hilight{salmon}{\texttt{x++;}} \\
+	& & & & \hilight{salmon}{$3$} & \hilight{salmon}{\texttt{\_xend();}} \\
+	%
+	\hilight{lavender}{$6b$} & \hilight{lavender}{\texttt{temp <- temp + 1;}} & & \\
+	\hilight{lavender}{$6c$} & \hilight{lavender}{\texttt{x <- temp;}} & & \\
+	\hilight{lavender}{$7$} & \hilight{lavender}{\texttt{mutex\_unlock(m);}} & & \\
+\end{tabular}
 \]
+
 Prior work \cite{htm-mario} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
-to exclude this family of interleavings,
-which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
+to exclude interleavings which violate the failure path's atomicity.
+Thread 1 would set the {\tt prevent\_transactions} flag across its {\tt x++} accesses,
+and
+Thread 3 would have to check it and trigger an abort. %(Figure \ref{fig:htm-fixed} lines 2-3).
+%which shows that correctly synchronizing even the simplest transactions may be surprisingly difficult or complex.
 %further motivating our research.
@@
 			0 & \texttt{\flow{while} (prevent\_transactions) \flow{continue};} \\
-		1 & \texttt{\flow{if} ((status = \call{\_xbegin}()) == \const{SUCCESS}) \{} \\
+			1 & \texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
 			2 & \texttt{~~~~\flow{if} (prevent\_transactions)} \\
@@
 			C & \texttt{\}} \\
-		%\texttt{~~~~\}} \\
-		%\texttt{\}} \\
 		\end{tabular}
 	\end{center}
-	\caption{Variant of the program in Figure~\ref{fig:htm-example},
+	\caption{Variant %of the program in Figure~\ref{fig:htm-example},
 		with additional synchronization to protect the failure path from the transactional path.
@@
 		for the sake of performance
-		by allowing threads to wait until transacting is safe again.}
+		by allowing threads to wait until transacting is safe again.
+		}
 	\label{fig:htm-fixed}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Implementation}
+
+To model check a transactional program it suffices to test a set of interleavings
+corresponding to all {\em observable} behaviours of the program.
+Because incomplete transactions are rolled back upon abort,
+we claim that interleaving other threads within the body of a transaction is never necessary,
+and hence model checking can achieve full coverage simply by conisdering each {\tt xbegin} invocation
+to be a preemption point with nondeterministic failure injection.
+Appendix~\ref{sec:proofs} provides several proof sketches to justify this claim.
+
+Our implementation of HTM-equivalent semantics has been incorporated by the Landslide maintainers upstream.
+It is available open-source at \url{https://github.com/bblum/landslide}.
+Programs should be ported to the Pebbles userland \cite{thrlib,kspec},
+their use of compiler HTM intrinsics should be replaced with the Landslide stubs provided in {\tt 410user/inc/htm.h},
+and HTM nondeterminism can then be enabled with the {\tt -X} command-line flag.
+%We have also extended its Iterative Deepening implementation \cite{quicksand}
+%with a new option ({\tt -M} flag) to
+%to optimize for completion time by prioritizing the maximal state space job and cancelling all others
+%(which maintains its soundness guarantee)
+%which we will use in our evaluation.
+%All test cases therein are also available in the repository linked above.
+All test cases in our evaluation to follow are also available in the repository linked above.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Evaluation}
+
+\newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
+\newcommand\cpu[1]{\hilight{darkcyan}{{#1}}\xspace}
+\newcommand\wtm[1]{\hilight{lime}{{#1}}\xspace}
+\newcommand\ints[1]{\hilight{pinkish}{{#1}}\xspace}
+
+%To the best of our knowledge, this is the first work to test transactional programs in a model-checking environment,
+%%To the best of the author's knowledge, Landslide is the first MC to check transactional concurrency,
+%so no other MC State of the Art%
+%\footnote{The author's DJ name.}
+%exists to compare to in controlled experiments.
+%Nevertheless, we pose the following evaluation questions.
+
+%\begin{enumerate}
+%	\item How quickly does Landslide find bugs in incorrect transactional programs of varying sizes?
+%	\item How quickly does Landslide verify correct transactional programs of varying sizes?
+%	%\item Does Landslide find any previously-unknown bugs in real-world transactional code?
+%	\item By the way, should MC research papers quantify variance in their CPU-time performance experiments?
+%\end{enumerate}
+
+Our preliminary evaluation evaluates Landslide's bug-finding performance on tests with subtle concurrency errors,
+as well as its verification performance on (as far as we know) correct tests.
+So far we have several hand-written unit tests
+and \cite{htm-mario}'s microbenchmarks and transactional AVL tree and separate-chaining hashmap,
+as follows.
+
+\begin{itemize}
+	\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}. %(assertion failure on {\tt count}).
+	\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
+	\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex failure path with a simple {\tt xadd}.
+	\item {\tt swap}: Microbenchmark that swaps values in an array.
+	\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
+	\item {\tt avl\_insert}: AVL tree concurrent insertion test.
+	\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed. %(spoilers!!).
+	\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test.
+	\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
+\end{itemize}
+
+The notation {\tt testname}$(K,N)$ will denote a test configuration of $K$ threads, each running $N$ iterations of the test logic.
+All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
+
+\subsection{Bugs}
+
+Table~\ref{tab:buges}
+presents our bug-finding results.
+%We configured Landslide to run the Quicksand algorithm \cite{quicksand}
+%% this note obsolete since swapbug was fixed and ends at 3,2
+%%\footnote{This was necessary in {\tt swapbug}s (4,1) and (4,2) to trigger a switch to the buggy state space.}
+%shown left,
+%as well as to prioritize the maximal state space as discussed above,
+%shown right,
+%each with a time limit of 30 minutes.
+%
+Most notably,
+in addition to the bugs we intentionally wrote in {\tt htm1} and {\tt swapbug},
+to our pleasant surprise
+Landslide also found a previously-unknown bug in the transactional AVL tree,
+exposed by {\tt avl\_insert} with any parameters higher than (2,1).
+Figure~\ref{fig:avlbug} shows the root cause, essentially the {\tt htm1} bug in disguise.
+% landslide-trace-1521093835.18927.html
+This manifested
+alternately
+%in (2,2) and (3,1)
+as a segfault
+% landslide-trace-1521092404.23772.html
+and
+%in (2,3)
+as a consistency-check assertion failure.
+The presence of {\tt while (\_retry);} makes the necessary preemption window extremely small
+in practice (between it and {\tt \_xbegin()}),
+whereas MC is blind to such matters of chance.%
+% NB. In case a reader doesn't believe this footnote,
+% I put the loop back in (with thr_yield(-1) therein) to restore the original state space,
+% and found the bug with these statistics:
+% [JOB 28] progress: 1957/3005897 brs (0.065105%), ETA 6414673y 121d 18h 28m 59s (elapsed 3m 15s)
+% [JOB 28] BUG FOUND: landslide-trace-1521099044.76395.html (1958 interleavings tested; 3 preemptions; job time 3m 15s; pldi time 746832861; new-fixed pldi cputime 823243872)
+%\footnote{Considering the loop does not affect the test's possible behaviours,
+%only its {\em likely} ones,
+%we removed it in Landslide's version of the test to keep the state space size manageable.}
+%We removed the loop in Landslide's version of the test
+%%To make this test Landslide-friendly, we removed that loop,
+%which reduces the state space size, but maintains the program's semantics.}
+The fact that even Figure~\ref{fig:htm-fixed}'s protocol's very proposer % こそさえ
+didn't get it right every time %motivates the need for MC on such programs,
+%and perhaps also means TSX primitives should be used only behind higher-level abstractions
+%such as lock elision \cite{lock-elision}.
+%and
+suggests TSX primitives should be encapsulated behind higher-level abstractions
+such as lock elision \cite{lock-elision},
+which can be verified in isolation with smaller state spaces
+then trusted in turn when checking their client programs \cite{dbug-phdthesis}.
+
+\begin{table*}[h]
+	\begin{center}
+	\begin{tabular}{cc||r|r|r|r}
+		\bf buggy test	& \bf params&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
+		\hline
+		\hline
+		{\tt htm1}
+			& 2,1	&\cpu{*9.47}& \wtm{*6.40}& \ints{21	}& \ETA{213} \\
+		(assertion)
+			& 2,2	&\cpu{*10.39}& \wtm{*7.70}& \ints{49	}& \ETA{1536} \\
+			& 2,3	&\cpu{*12.83}& \wtm{*9.67}& \ints{113	}& \ETA{10752} \\
+			& 2,4	&\cpu{*18.63}& \wtm{*15.86}& \ints{257	}& \ETA{73728} \\
+		\cline{2-6}
+			& 3,1	&\cpu{*9.50}& \wtm{*6.79}& \ints{21	}& \ETA{13653} \\
+			& 3,2	&\cpu{*10.72}& \wtm{*7.97}& \ints{49	}& \ETA{393216} \\
+			& 3,3	&\cpu{*13.84}& \wtm{*11.01}& \ints{113	}& \ETA{11010048} \\
+			& 3,4	&\cpu{*31.37}& \wtm{*28.53}& \ints{257	}& \ETA{301989888} \\
+		\cline{2-6}
+			& 4,1	&\cpu{*9.82}& \wtm{*7.00}& \ints{21	}& \ETA{873813} \\
+			& 4,2	&\cpu{*11.54}& \wtm{*8.55}& \ints{49	}& \ETA{100663296} \\
+		\hline
+		{\tt swapbug}
+			& 2,1	&\cpu{*38.96}& \wtm{*13.15}& \ints{109	}& \ETA{194} \\
+		(deadlock)
+			& 2,2	&\cpu{*44.73}& \wtm{19.47}& \ints{281	}& \ETA{1620} \\
+			& 2,3	&\cpu{*60.30}& \wtm{*35.55}& \ints{718	}& \ETA{12748} \\
+			& 2,4	&\cpu{*108.58}& \wtm{82.60}& \ints{1820	}& \ETA{97823} \\
+		\cline{2-6}
+			& 3,1	&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{184984} \\
+			& 3,2	&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{3099225} \\
+		\hline
+		{\tt avl\_insert}
+			& 2,2	&\cpu{*81.00}&\wtm{*40.30}&\ints{336}&\ETA{379982} \\
+		(segfault+)
+			& 2,3	&\cpu{*1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
+			& 2,4	&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{36019973} \\
+		\cline{2-6}
+			& 3,1	&\cpu{*69.99}&\wtm{*24.25}&\ints{78}&\ETA{1572107} \\
+			& 3,2	&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{1402363529} \\
+	\end{tabular}
+	\end{center}
+	\caption{Landslide's bug-finding performance on various test configurations.
+		We list the CPU-time and wall-clock time %(both in seconds)
+		elapsed, %until the bug was found,
+		plus the number of interleavings of the ultimately buggy state space tested,
+		before the bug was found.
+		Lastly, Landslide's state space size estimation \cite{estimation},
+		though approximate at best,
+		confers a sense of the exponential explosion.
+	}
+	\label{tab:buges}
+\end{table*}
+%\begin{table*}[t]
+%	\begin{center}
+%	\begin{tabular}{cc||r|r|r||r|r|r|r}
+%			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
+%		\bf buggy test	& \bf params&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
+%		\hline
+%		\hline
+%		{\tt htm1}
+%			& 2,1	&\cpu{  45.78}&\wtm{9.70}&\ints{21}& \cpu{*9.47}& \wtm{*6.40}& \ints{21	}& \ETA{213} \\
+%		(assertion)
+%			& 2,2	&\cpu{  84.14}&\wtm{13.59}&\ints{*33}& \cpu{*10.39}& \wtm{*7.70}& \ints{49	}& \ETA{1536} \\
+%			& 2,3	&\cpu{ 131.91}&\wtm{20.44}&\ints{*73}& \cpu{*12.83}& \wtm{*9.67}& \ints{113	}& \ETA{10752} \\
+%			& 2,4	&\cpu{ 255.75}&\wtm{37.56}&\ints{257}& \cpu{*18.63}& \wtm{*15.86}& \ints{257	}& \ETA{73728} \\
+%		\cline{2-9}
+%			& 3,1	&\cpu{ 114.06}&\wtm{17.45}&\ints{*15}& \cpu{*9.50}& \wtm{*6.79}& \ints{21	}& \ETA{13653} \\
+%			& 3,2	&\cpu{ 109.60}&\wtm{26.16}&\ints{49}& \cpu{*10.72}& \wtm{*7.97}& \ints{49	}& \ETA{393216} \\
+%			& 3,3	&\cpu{ 124.80}&\wtm{20.40}&\ints{*73}& \cpu{*13.84}& \wtm{*11.01}& \ints{113	}& \ETA{11010048} \\
+%			& 3,4	&\cpu{ 227.49}&\wtm{35.15}&\ints{*161}& \cpu{*31.37}& \wtm{*28.53}& \ints{257	}& \ETA{301989888} \\
+%		\cline{2-9}
+%			& 4,1	&\cpu{ 53.08}&\wtm{9.79}&\ints{*15}& \cpu{*9.82}& \wtm{*7.00}& \ints{21	}& \ETA{873813} \\
+%			& 4,2	&\cpu{117.07}&\wtm{19.09}&\ints{*33}& \cpu{*11.54}& \wtm{*8.55}& \ints{49	}& \ETA{100663296} \\
+%		\hline
+%		{\tt swapbug}
+%			& 2,1	&\cpu{ 70.95}&\wtm{13.45}&\ints{*16}& \cpu{*38.96}& \wtm{*13.15}& \ints{109	}& \ETA{194} \\
+%		(deadlock)
+%			& 2,2	&\cpu{ 107.28}&\wtm{*17.45}&\ints{*146}& \cpu{*44.73}& \wtm{19.47}& \ints{281	}& \ETA{1620} \\
+%			& 2,3	&\cpu{ 280.05}&\wtm{38.70}&\ints{*352}& \cpu{*60.30}& \wtm{*35.55}& \ints{718	}& \ETA{12748} \\
+%			& 2,4	&\cpu{ 617.94}&\wtm{*81.50}&\ints{*834}& \cpu{*108.58}& \wtm{82.60}& \ints{1820	}& \ETA{97823} \\
+%		\cline{2-9}
+%			& 3,1	&\cpu{*1275.04}&\wtm{*163.42}&\ints{*771}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{184984} \\
+%			& 3,2	&\ETA{--}&\ETA{>30m}&\ETA{--}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{3099225} \\
+%		% whoops, this was with the dumb lock ordering
+%		%	& 3,1	& 11.19	& 3.46	& 9	& 3.13	& 3.13	& 9	& 432 \\
+%		%	& 3,2	& 12.41	& 3.76	& 19	& 3.33	& 3.33	& 19	& 5832 \\
+%		%	& 3,3	& 15.77	& 4.58	& 41	& 3.95	& 3.95	& 41	& 77759 \\
+%		%	& 3,4	& 24.82	& 6.85	& 89	& 5.20	& 5.20	& 89	& 1026431 \\
+%		%\cline{2-9}
+%		%	& 4,1	& 702.33& 98.06	& 361	& 90.94	& 64.27	& 1607	& 138758 \\
+%		%	& 4,2	&9738.43&1257.20& 3023	& 674.87& 650.58& 16511	& 13755188 \\
+%		\hline
+%		{\tt avl\_insert}
+%			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{*83}	&\cpu{*81.00}&\wtm{*40.30}&\ints{336}&\ETA{379982} \\
+%		(segfault+)
+%			% XXX: this 2,3 -M-mode bug was found *mostly* running in isolation,
+%			% but for the last few minutes i had another landslide running in parallel bc i thought itd time out.
+%			% according to variance expt, it should be rerun for a *minor* speedup, but i can't be bothered
+%			& 2,3	&\cpu{2670.87}&\wtm{*330.45}&\ints{*3066}&\cpu{*1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
+%			& 2,4	&\cpu{*3259.37}&\wtm{*436.50}&\ints{*1639}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{36019973} \\
+%			% smaller than above? estimate's just fickle i guess; they can go up or down with time
+%		\cline{2-9}
+%			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{*28}	&\cpu{*69.99}&\wtm{*24.25}&\ints{78}&\ETA{1572107} \\
+%			& 3,2	&\cpu{*1569.09}&\wtm{*216.85}&\ints{*209}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{1402363529} \\
+%	\end{tabular}
+%	\end{center}
+%	\caption{Landslide's bug-finding performance on various test configurations.
+%		Quicksand's workqueue approach optimized for fast bug-finding
+%		is compared against our maximum-state-space-prioritizing approach for fast verification.
+%		For each, we list the CPU-time and wall-clock time %(both in seconds)
+%		elapsed, %until the bug was found,
+%		plus the number of interleavings of the ultimately buggy state space tested,
+%		before the bug was found.
+%		* marks the winning measurements between each series.
+%		Lastly, Landslide's state space size estimation \cite{estimation},
+%		though approximate at best,
+%		confers a sense of the exponential explosion.
+%	}
+%	\label{tab:buges}
+%\end{table*}
+
+\begin{figure}[t]
+	\begin{center}
+		\begin{tabular}{l}
+		%\texttt{void count() \{} \\
+		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
+		\texttt{\flow{while} (\_retry);} \\
+		\texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
+		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
+		\texttt{~~~~\call{\_xend}();} \\
+		\texttt{\} \flow{else} \{} \\
+		\texttt{~~~~\call{pthread\_mutex\_lock}(\&\_tree\_lock);} \\
+		\texttt{~~~~\_retry = \const{true};} \\
+		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
+		\texttt{~~~~\_retry = \const{false};} \\
+		\texttt{~~~~\call{pthread\_mutex\_unlock}(\&\_tree\_lock);} \\
+		\texttt{\}} \\
+		\end{tabular}
+	\end{center}
+	%\caption{mario man had a very large house. or a castle.}
+	\caption{Unmodified code from {\tt htmavl.cpp} showing the previously-unknown bug found by Landslide.
+		The transaction path fails to check {\tt \_retry},
+		leading to data races and corruption just as in {\tt htm1}.
+		}
+	\label{fig:avlbug}
+\end{figure}
+
+\subsection{Verification}
+
+Landslide proved several of our test configurations correct, as shown in Table~\ref{tab:verifs}
+%With no bugs to find, comparing the different testing modes against each other is meaningless;
+%we simply present their state space sizes and runtime (using {\tt -M}) in Table~\ref{tab:verifs}.
+Several larger tests we were unable to complete. %within several hours.
+% バレちゃった
+%before the submission deadline are also shown.
+%Their completion times are listed as {\em unk.}
+These are indicated with $\dagger$ %\ETAdag{}
+and their listed ETAs and state space sizes represent Landslide's estimate %at the moment of timeout.
+after a timeout of 1 hour
+(hopefully enough for data-race PPs to saturate and estimates to stablize,
+although note the two estimate types use different algorithms so may disagree significantly).
+
+\newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}}
+\begin{table}[h!]
+	\begin{center}
+		\begin{tabular}{cc|r|r}
+			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
+			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
+			\hline
+			\hline
+			{\tt htm2}
+			& 2,1 & \cpu{18.57}	& \ints{294}	\\
+			& 2,2 & \cpu{133.78}	& \ints{4902}	\\
+			& 2,3 &\cpu{1986.98}	& \ints{79017}	\\
+			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
+			& 3,2 &	\ETAdag{10d 14h}& \ETAdag{13763999}	\\
+			%[JOB 4] progress: 127032/13763999 brs (0.922929%), ETA 10d 13h 52m 44s (elapsed 59m 56s)
+			%[JOB 4] TIMED OUT (0.922929%; ETA 10d 13h 52m 44s)
+			\hline
+			{\tt counter}
+			& 2,1 & \cpu{5.57}	& \ints{30}	\\
+			& 2,2 & \cpu{15.53}	& \ints{384}	\\
+			& 2,3 &\cpu{155.00}	& \ints{5280}	\\
+			& 2,4 &\cpu{2211.10}	& \ints{75264}	\\
+			& 3,1 & \cpu{57.90}	& \ints{1960}	\\
+			& 3,2 & \cpu{10028.93}	& \ints{329888}	\\ % TODO running atm
+			\hline
+			{\tt swap}
+			%& 2,1 & \cpu{38.77}	& \ints{228}	\\ % pre-ignore_thr_func fix
+			& 2,1 & \cpu{32.98}	& \ints{193}	\\
+			& 2,2 & \cpu{3652.91}	& \ints{101150}	\\
+			%[JOB 8] progress: 97913/161897 brs (60.478362%), ETA 1h 29m 30s (elapsed 59m 52s)
+			% [JOB 8] TIMED OUT (60.478362%; ETA 1h 29m 30s)
+
+			& 3,1 & \ETAdag{8d 12h} & \ETAdag{411312}	\\
+			%[JOB 8] progress: 103873/411312 brs (25.254021%), ETA 8d 11h 36m 0s (elapsed 59m 42s)
+			%[JOB 8] TIMED OUT (25.254021%; ETA 8d 11h 36m 0s)
+			\hline
+			{\tt avl\_insert}
+			& 2,1 & \cpu{1083.35}	& \ints{40062}	\\
+			%& 2,2 & \multicolumn{2}{c}{\em (bug!)} \\
+			\hline
+			{\tt avl\_fixed}
+			& 2,1 & \cpu{1079.03} 	& \ints{45078} 	\\
+			& 2,2 & \ETAdag{2762y}	& \ETAdag{8714863} 	\\
+			%[JOB 19] progress: 45863/8714863 brs (0.526262%), ETA 2762y 28d 9h 57m 14s (elapsed 59m 25s)
+			%[JOB 19] TIMED OUT (0.526262%; ETA 2762y 28d 9h 57m 14s)
+			% the below is from avlfixed2,2 after running it overnight
+			% eta is ever worse but i'm using the 1hr one as promised
+			%[JOB 19] progress: 315672/29992026 brs (1.052520%), ETA 321008y 136d 18h 22m 12s (elapsed 6h 58m 15s)
+			& 3,1 & \ETAdag{12d 2h}	& \ETAdag{11498545}	\\
+
+			\hline
+			{\tt map\_basic}
+			& 2,1 & \ETAdag{10d 17h} & \ETAdag{16388977} \\
+			%[JOB 15] progress: 70330/16388977 brs (0.429130%), ETA 10d 16h 34m 15s (elapsed 43m 21s)
+			%[JOB 15] TIMED OUT (0.429130%; ETA 10d 16h 34m 15s)
+			\hline
+			{\tt map\_basicer}
+			& 2,1 & \cpu{877.44}& \ints{28635}	\\
+			& 2,2 & \ETAdag{2d 7h} & \ETAdag{5925634} \\
+			& 3,1 & \ETAdag{468d 13h} & \ETAdag{35893653} \\
+			% [JOB 5] progress: 100181/35893653 brs (0.279105%), ETA 1y 103d 12h 50m 47s (elapsed 59m 24s)
+			%[JOB 5] TIMED OUT (0.279105%; ETA 1y 103d 12h 50m 47s)
+			% i don't know what this is but it was hanging out under 3,1 here :<
+			%[JOB 6] progress: 35042/5925634 brs (0.591363%), ETA 2d 6h 40m 30s (elapsed 21m 12s)
+			%[JOB 6] TIMED OUT (0.591363%; ETA 2d 6h 40m 30s)
+		\end{tabular}
+	\end{center}
+	\caption{Transactional tests verified (or not) by Landslide.}
+	\label{tab:verifs}
+\end{table}
+
+%For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
+%Though these tests proved beyond our reach,
+%in contrast with the instant gratification expected of bug-finding,
+%%we believe
+%the verification guarantee may in some cases be worth the estimated time requirement
+%for widely-used industrial implementations.
+%Future work may also expand our coverage upon this frontier \cite{landslide-phdthesis}.
+
+\bibliographystyle{abbrv}
+\bibliography{citations}
+
+\appendix
+
+\section{Soundness proofs}
+\label{sec:proofs}
+
+\newcommand\ti{\ensuremath{\hilight{lavender}{\mathbf{T1}}}\xspace}
+\newcommand\tj{\ensuremath{\hilight{seafoam}{\mathbf{T2}}}\xspace}
+\newcommand\tk{\ensuremath{\hilight{salmon}{\mathbf{T3}}}\xspace}
+
+\newcommand\tiat[1]{\ensuremath{\hilight{lavender}{\mathbf{T1}@#1}}\xspace}
+\newcommand\tjat[1]{\ensuremath{\hilight{seafoam}{\mathbf{T2}@#1}}\xspace}
+\newcommand\tkat[1]{\ensuremath{\hilight{salmon}{\mathbf{T3}@#1}}\xspace}
+
+To formalize transactional memory in Landslide's framework of thread concurrency,
+we have made two major simplifications:
+simulating transaction aborts as immediate failure injections,
+and treating transaction atomicity as a global mutex during data-race analysis.
+This appendix provides the corresponding equivalence proofs.
+
+{\bf Notation.} Let $I = TN_1@L_1, TN_2@L_2, ... TN_n@L_n$,
+with $N_i$ a thread ID and $L_i$ a code line number,
+denote the execution sequence of a program as it runs according to the specified thread interleaving.%
+\footnote{This serialization of concurrent execution is told from the perspective of all CPUs at once
+and hence assumes sequential consistency.
+For a treatment of relaxed memory models see \cite{tsopso}.}
+
 \subsection{Modeling Transaction Failure}
 
-Left unstated in interleavings such as $\tiat{1-2c}, \tjat{1-3}, \tiat{4-7}$%
-\footnote{For a clearer example we reorder \ti's write to {\tt x} before \tj's part here.}
+Left unstated in interleavings such as $\tiat{1-2c}, \tjat{1-3}, \tiat{4-7}$ in Section~\ref{sec:example}'s example
+%\footnote{For a clearer example we reorder \ti's write to {\tt x} before \tj's part here.}
 are HTM's execution semantics, namely:
@@
 
-\newcommand\tii{\ensuremath{\hilight{lavender}{\mathbf{Ti}}}\xspace}
-\newcommand\tjj{\ensuremath{\hilight{seafoam}{\mathbf{Tj}}}\xspace}
-\newcommand\tkk{\ensuremath{\hilight{salmon}{\mathbf{Tk}}}\xspace}
-
-\newcommand\tiiat[1]{\ensuremath{\hilight{lavender}{\mathbf{Ti}@#1}}\xspace}
-\newcommand\tjjat[1]{\ensuremath{\hilight{seafoam} {\mathbf{Tj}@#1}}\xspace}
-\newcommand\tkkat[1]{\ensuremath{\hilight{salmon}  {\mathbf{Tk}@#1}}\xspace}
-
-\newcommand\llama[1]{\includegraphics[width=#1]{llama.pdf}}
+\newcommand\llama[1]{\includegraphics[width=#1]{../llama.pdf}}
 \newcommand\llitem{\item[\raisebox{-0.15em}{\llama{1.2em}}]}
@@
 	\[
-	\begin{tabular}{c}
-		$\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},$\\
-		$\tjjat{\gamma_1}\dots\tjjat{\gamma_j},$ \\
-		$\tkkat{\kappa_1}\dots\tkkat{\kappa_k},$ \\
-		$\tiiat{\beta_{b+1}}$ % $\dots\tiiat{\beta_{n-1}}$ -- excluded bc might abort
-	\end{tabular}
+		\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b},
+		\tjjat{\gamma_1}\dots\tjjat{\gamma_j},
+		\tkkat{\kappa_1}\dots\tkkat{\kappa_k},
+		\tiiat{\beta_{b+1}} % $\dots\tiiat{\beta_{n-1}}$ -- excluded bc might abort
 	\]
@@
 			(conflicting case), or
-		\item $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b}\dots\tiiat{\beta_n},\tjjat{\gamma_1}\dots\tjjat{\gamma_j},$ \\
-			$\tkkat{\kappa_1}\dots\tkkat{\kappa_k}$
+		\item $\tiiat{\alpha},\tiiat{\beta_1}\dots\tiiat{\beta_b}\dots\tiiat{\beta_n},\tjjat{\gamma_1}\dots\tjjat{\gamma_j},\tkkat{\kappa_1}\dots\tkkat{\kappa_k}$
 			(independent case)
@@
 
-\subsection{Implementation}
-
-Our implementation of HTM-equivalent semantics has been incorporated by the Landslide maintainers upstream.
-It is available open-source at \url{https://github.com/bblum/landslide}.
-Programs should be ported to the Pebbles userland \cite{thrlib,kspec},
-their use of compiler HTM intrinsics should be replaced with the Landslide stubs provided in {\tt 410user/inc/htm.h},
-and HTM nondeterminism can then be enabled with the {\tt -X} command-line flag.
-We have also extended its Iterative Deepening implementation \cite{quicksand}
-with a new option ({\tt -M} flag) to
-to optimize for completion time by prioritizing the maximal state space job and cancelling all others
-(which maintains its soundness guarantee)
-which we will use in our evaluation.
-All test cases therein are also available in the repository linked above.
-%Several example tests are also provided:
-%{\tt htm1} demonstrates the bug in Figure~\ref{fig:htm-example},
-%and {\tt htm2} demonstrates Figure~\ref{fig:htm-fixed}'s fix.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Evaluation}
-
-\newcommand\ETA[1]{\hilight{brownish}{{\em #1}}\xspace}
-\newcommand\cpu[1]{\hilight{darkcyan}{{#1}}\xspace}
-\newcommand\wtm[1]{\hilight{lime}{{#1}}\xspace}
-\newcommand\ints[1]{\hilight{pinkish}{{#1}}\xspace}
-
-\begin{table*}[t]
-	\begin{center}
-	\begin{tabular}{cc||r|r|r||r|r|r|r}
-			&	&\multicolumn{3}{c||}{\bf Quicksand mode}&\multicolumn{4}{c}{{\bf Maximal state space mode} ({\tt -M})} \\
-		\bf buggy test	& \bf params&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}&\cpu{\bf cpu (s)}&\wtm{\bf wall (s)}&\ints{\bf int's}& \ETA{\bf SS size (est.)} \\
-		\hline
-		\hline
-		{\tt htm1}
-			& 2,1	&\cpu{  45.78}&\wtm{9.70}&\ints{21}& \cpu{*9.47}& \wtm{*6.40}& \ints{21	}& \ETA{213} \\
-		(assertion)
-			& 2,2	&\cpu{  84.14}&\wtm{13.59}&\ints{*33}& \cpu{*10.39}& \wtm{*7.70}& \ints{49	}& \ETA{1536} \\
-			& 2,3	&\cpu{ 131.91}&\wtm{20.44}&\ints{*73}& \cpu{*12.83}& \wtm{*9.67}& \ints{113	}& \ETA{10752} \\
-			& 2,4	&\cpu{ 255.75}&\wtm{37.56}&\ints{257}& \cpu{*18.63}& \wtm{*15.86}& \ints{257	}& \ETA{73728} \\
-		\cline{2-9}
-			& 3,1	&\cpu{ 114.06}&\wtm{17.45}&\ints{*15}& \cpu{*9.50}& \wtm{*6.79}& \ints{21	}& \ETA{13653} \\
-			& 3,2	&\cpu{ 109.60}&\wtm{26.16}&\ints{49}& \cpu{*10.72}& \wtm{*7.97}& \ints{49	}& \ETA{393216} \\
-			& 3,3	&\cpu{ 124.80}&\wtm{20.40}&\ints{*73}& \cpu{*13.84}& \wtm{*11.01}& \ints{113	}& \ETA{11010048} \\
-			& 3,4	&\cpu{ 227.49}&\wtm{35.15}&\ints{*161}& \cpu{*31.37}& \wtm{*28.53}& \ints{257	}& \ETA{301989888} \\
-		\cline{2-9}
-			& 4,1	&\cpu{ 53.08}&\wtm{9.79}&\ints{*15}& \cpu{*9.82}& \wtm{*7.00}& \ints{21	}& \ETA{873813} \\
-			& 4,2	&\cpu{117.07}&\wtm{19.09}&\ints{*33}& \cpu{*11.54}& \wtm{*8.55}& \ints{49	}& \ETA{100663296} \\
-		\hline
-		{\tt swapbug}
-			& 2,1	&\cpu{ 70.95}&\wtm{13.45}&\ints{*16}& \cpu{*38.96}& \wtm{*13.15}& \ints{109	}& \ETA{194} \\
-		(deadlock)
-			& 2,2	&\cpu{ 107.28}&\wtm{*17.45}&\ints{*146}& \cpu{*44.73}& \wtm{19.47}& \ints{281	}& \ETA{1620} \\
-			& 2,3	&\cpu{ 280.05}&\wtm{38.70}&\ints{*352}& \cpu{*60.30}& \wtm{*35.55}& \ints{718	}& \ETA{12748} \\
-			& 2,4	&\cpu{ 617.94}&\wtm{*81.50}&\ints{*834}& \cpu{*108.58}& \wtm{82.60}& \ints{1820	}& \ETA{97823} \\
-		\cline{2-9}
-			& 3,1	&\cpu{*1275.04}&\wtm{*163.42}&\ints{*771}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{184984} \\
-			& 3,2	&\ETA{--}&\ETA{>30m}&\ETA{--}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{3099225} \\
-		% whoops, this was with the dumb lock ordering
-		%	& 3,1	& 11.19	& 3.46	& 9	& 3.13	& 3.13	& 9	& 432 \\
-		%	& 3,2	& 12.41	& 3.76	& 19	& 3.33	& 3.33	& 19	& 5832 \\
-		%	& 3,3	& 15.77	& 4.58	& 41	& 3.95	& 3.95	& 41	& 77759 \\
-		%	& 3,4	& 24.82	& 6.85	& 89	& 5.20	& 5.20	& 89	& 1026431 \\
-		%\cline{2-9}
-		%	& 4,1	& 702.33& 98.06	& 361	& 90.94	& 64.27	& 1607	& 138758 \\
-		%	& 4,2	&9738.43&1257.20& 3023	& 674.87& 650.58& 16511	& 13755188 \\
-		\hline
-		{\tt avl\_insert}
-			& 2,2	&\cpu{488.07}&\wtm{64.77}&\ints{*83}	&\cpu{*81.00}&\wtm{*40.30}&\ints{336}&\ETA{379982} \\
-		(segfault+)
-			% XXX: this 2,3 -M-mode bug was found *mostly* running in isolation,
-			% but for the last few minutes i had another landslide running in parallel bc i thought itd time out.
-			% according to variance expt, it should be rerun for a *minor* speedup, but i can't be bothered
-			& 2,3	&\cpu{2670.87}&\wtm{*330.45}&\ints{*3066}&\cpu{*1331.79}&\wtm{1274.36}&\ints{13926}&\ETA{96248131} \\
-			& 2,4	&\cpu{*3259.37}&\wtm{*436.50}&\ints{*1639}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{36019973} \\
-			% smaller than above? estimate's just fickle i guess; they can go up or down with time
-		\cline{2-9}
-			& 3,1	&\cpu{222.02}&\wtm{40.04}&\ints{*28}	&\cpu{*69.99}&\wtm{*24.25}&\ints{78}&\ETA{1572107} \\
-			& 3,2	&\cpu{*1569.09}&\wtm{*216.85}&\ints{*209}&\ETA{--}&\ETA{>30m}&\ETA{--}& \ETA{1402363529} \\
-	\end{tabular}
-	\end{center}
-	\caption{Landslide's bug-finding performance on various test configurations.
-		Quicksand's workqueue approach optimized for fast bug-finding
-		is compared against our maximum-state-space-prioritizing approach for fast verification.
-		For each, we list the CPU-time and wall-clock time %(both in seconds)
-		elapsed, %until the bug was found,
-		plus the number of interleavings of the ultimately buggy state space tested,
-		before the bug was found.
-		* marks the winning measurements between each series.
-		Lastly, Landslide's state space size estimation \cite{estimation},
-		though approximate at best,
-		confers a sense of the exponential explosion.
-	}
-	\label{tab:buges}
-\end{table*}
-
-To the best of our knowledge, this is the first work to test transactional programs in a model-checking environment,
-%To the best of the author's knowledge, Landslide is the first MC to check transactional concurrency,
-so no other MC State of the Art%
-\footnote{The author's DJ name.}
-exists to compare to in controlled experiments.
-Nevertheless, we pose the following evaluation questions.
-
-\begin{enumerate}
-	\item How quickly does Landslide find bugs in incorrect transactional programs of varying sizes?
-	\item How quickly does Landslide verify correct transactional programs of varying sizes?
-	%\item Does Landslide find any previously-unknown bugs in real-world transactional code?
-	\item By the way, should MC research papers quantify variance in their CPU-time performance experiments?
-\end{enumerate}
-
-Our evaluation suite comprises several hand-written unit tests
-and \cite{htm-mario}'s microbenchmarks and transactional AVL tree and separate-chaining hashmap,
-as follows.
-
-\begin{itemize}
-	\item {\tt htm1}: The bug from Figure~\ref{fig:htm-example}. %(assertion failure on {\tt count}).
-	\item {\tt htm2}: The fixed version as in Figure~\ref{fig:htm-fixed}.
-	\item {\tt counter}: Microbenchmark version of {\tt htm2} which replaces the complex failure path with a simple {\tt xadd}.
-	\item {\tt swap}: Microbenchmark that swaps values in an array.
-	\item {\tt swapbug}: {\tt swap} modified to introduce circular locking in the failure path. %(deadlock).
-	\item {\tt avl\_insert}: AVL tree concurrent insertion test.
-	\item {\tt avl\_fixed}: {\tt avl\_insert} with the AVL bug fixed (spoilers!!).
-	\item {\tt map\_basic}: Separate-chaining hashmap concurrent insertion test.
-	\item {\tt map\_basicer}: {\tt map\_basic} modified with a larger initial size to skip the resizing step.
-\end{itemize}
-
-The notation {\tt testname}$(K,N)$ will denote a test configuration of $K$ threads, each running $N$ iterations of the test logic.
-All tests were run on an 8-core 2.7GHz Core i7 with 32 GB RAM.
-
-\subsection{Bugs}
-
-Table~\ref{tab:buges}
-presents our bug-finding results.
-We configured Landslide to run the Quicksand algorithm \cite{quicksand}
-% this note obsolete since swapbug was fixed and ends at 3,2
-%\footnote{This was necessary in {\tt swapbug}s (4,1) and (4,2) to trigger a switch to the buggy state space.}
-shown left,
-as well as to prioritize the maximal state space as discussed above,
-shown right,
-each with a time limit of 30 minutes.
-We draw three main conclusions from this data.
-%\begin{itemize}
-	%\item
-
-		{\bf Finding bugs quickly.}
-		As the test parameters increase,
-		the multiplicative factor in bug-finding speed (2-4x, eyeballing) is generally smaller
-		than that of the total number of interleavings (10-100x).
-		In other words,
-		should they exist,
-		Landslide find bugs reasonably quickly in these transactional programs
-		despite prohibitive exponential explosion in total state space size.
-		This corroborates the prior work \cite{quicksand},
-		extending its good news to the world of HTM.
-	%\item
-
-		{\bf New bugs.}
-		In addition to the bugs we intentionally wrote in {\tt htm1} and {\tt swapbug},
-		to our pleasant surprise
-		Landslide also found a previously-unknown bug in the transactional AVL tree,
-		exposed by {\tt avl\_insert} with any parameters higher than (2,1).
-		Figure~\ref{fig:avlbug} shows the root cause, essentially the {\tt htm1} bug in disguise.
-		% landslide-trace-1521093835.18927.html
-		This manifested
-		alternately
-		%in (2,2) and (3,1)
-		as a segfault
-		% landslide-trace-1521092404.23772.html
-		and
-		%in (2,3)
-		as a consistency-check assertion failure.
-		The presence of {\tt while (\_retry);} makes the necessary preemption window extremely small
-		in practice (between it and {\tt \_xbegin()}),
-		whereas MC is blind to such matters of chance.%
-		% NB. In case a reader doesn't believe this footnote,
-		% I put the loop back in (with thr_yield(-1) therein) to restore the original state space,
-		% and found the bug with these statistics:
-		% [JOB 28] progress: 1957/3005897 brs (0.065105%), ETA 6414673y 121d 18h 28m 59s (elapsed 3m 15s)
-		% [JOB 28] BUG FOUND: landslide-trace-1521099044.76395.html (1958 interleavings tested; 3 preemptions; job time 3m 15s; pldi time 746832861; new-fixed pldi cputime 823243872)
-		\footnote{Considering the loop does not affect the test's possible behaviours,
-		only its {\em likely} ones,
-		we removed it in Landslide's version of the test to keep the state space size manageable.}
-		%We removed the loop in Landslide's version of the test
-		%%To make this test Landslide-friendly, we removed that loop,
-		%which reduces the state space size, but maintains the program's semantics.}
-		Moreover, we conclude that even Figure~\ref{fig:htm-fixed}'s protocol's very proposer % こそさえ
-		getting it wrong motivates the need for MC on such programs,
-		%and perhaps also means TSX primitives should be used only behind higher-level abstractions
-		%such as lock elision \cite{lock-elision}.
-		and suggests TSX primitives should be encapsulated behind higher-level abstractions
-		such as lock elision \cite{lock-elision},
-		which can be verified in isolation with smaller state spaces
-		then trusted in turn when checking their client programs \cite{dbug-phdthesis}.
-
-\begin{figure}[t]
-	%\begin{center}
-		\begin{tabular}{l}
-		%\texttt{void count() \{} \\
-		%\texttt{~~~~for (int i = 0; i < 1000; i++) \{} \\
-		\texttt{\flow{while} (\_retry);} \\
-		\texttt{\flow{if} (\call{\_xbegin}() == \const{SUCCESS}) \{} \\
-		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
-		\texttt{~~~~\call{\_xend}();} \\
-		\texttt{\} \flow{else} \{} \\
-		\texttt{~~~~\call{pthread\_mutex\_lock}(\&\_tree\_lock);} \\
-		\texttt{~~~~\_retry = \const{true};} \\
-		\texttt{~~~~\hilight{brickred}{tie}(\_root,inserted) = \call{\_insert}(\_root,n); } \\
-		\texttt{~~~~\_retry = \const{false};} \\
-		\texttt{~~~~\call{pthread\_mutex\_unlock}(\&\_tree\_lock);} \\
-		\texttt{\}} \\
-		\end{tabular}
-	%\end{center}
-	%\caption{mario man had a very large house. or a castle.}
-	\caption{Unmodified code from {\tt htmavl.cpp} showing the previously-unknown bug found by Landslide.
-		The transaction path fails to check {\tt \_retry},
-		leading to data races and corruption just as in {\tt htm1}.
-		}
-	\label{fig:avlbug}
-\end{figure}
-	%\item Despite Quicksand's ability to optimize for finding bugs in overall smaller state spaces,
-	%	this does not correlate with better CPU-time or even better wall-clock time.
-	%	While possibly a consequence of the modest test size,
-	%	this suggests future work should extend the Quicksand algorithm
-	%	to prioritize by state space maximality as well as by state space estimate.
-	%\item
-
-		{\bf Performance comparison.}
-		Quicksand's ability to find bugs in fewer distinct interleavings %overall smaller state spaces
-		does not necessarily correlate with better performance. %CPU- or even wall-clock time.
-		%Comparing to the break-even point in Quicksand's evaluation \cite{quicksand},
-		%most of these test configurations fall on the side favouring the single-state-space approach,
-		Most of our tests are too small for its approach to pay off,
-		with {\tt swapbug}(3,1) and {\tt avl\_insert} as its notable wins.
-		While the prior work showed plenty more wins \cite{quicksand},
-		%an upper bound also exists:
-		%in {\tt swapbug}(3,2),
-		%not even the {\em minimal} state space was completed in time.
-		%this suggests
-		future MCs could prioritize state spaces using not just size estimation
-		but state space maximality as well
-		%but a hybrid approach also conisdering state space maximality and preemption bounds \cite{chess-icb},
-		to soften the trade-off
-		both for smaller tests and for verification.
-%\end{itemize}
-Speaking of which...
-
-\subsection{Verification}
-
-Landslide proved the following tests correct.
-With no bugs to find, comparing the different testing modes against each other is meaningless;
-we simply present their state space sizes and runtime (using {\tt -M}) in Table~\ref{tab:verifs}.
-
-\newcommand\ETAdag[1]{\ETA{\ensuremath{\dagger}#1}}
-\begin{table}[h]
-	\begin{center}
-		\begin{tabular}{cc|r|r}
-			& & \cpu{\bf cpu (s)} & \ints{\bf SS size} \\
-			\bf test & \bf params & \em (or \ETAdag{\bf \em ETA}) & \em (or \ETAdag{\bf \em est.}) \\
-			\hline
-			\hline
-			{\tt htm2}
-			& 2,1 & \cpu{18.57}	& \ints{294}	\\
-			& 2,2 & \cpu{133.78}	& \ints{4902}	\\
-			& 2,3 &\cpu{1986.98}	& \ints{79017}	\\
-			& 3,1 &\cpu{11672.15}	& \ints{467730}	\\
-			& 3,2 &	\ETAdag{10d 14h}& \ETAdag{13763999}	\\
-			%[JOB 4] progress: 127032/13763999 brs (0.922929%), ETA 10d 13h 52m 44s (elapsed 59m 56s)
-			%[JOB 4] TIMED OUT (0.922929%; ETA 10d 13h 52m 44s)
-			\hline
-			{\tt counter}
-			& 2,1 & \cpu{5.57}	& \ints{30}	\\
-			& 2,2 & \cpu{15.53}	& \ints{384}	\\
-			& 2,3 &\cpu{155.00}	& \ints{5280}	\\
-			& 2,4 &\cpu{2211.10}	& \ints{75264}	\\
-			& 3,1 & \cpu{57.90}	& \ints{1960}	\\
-			& 3,2 & \cpu{10028.93}	& \ints{329888}	\\ % TODO running atm
-			\hline
-			{\tt swap}
-			%& 2,1 & \cpu{38.77}	& \ints{228}	\\ % pre-ignore_thr_func fix
-			& 2,1 & \cpu{32.98}	& \ints{193}	\\
-			& 2,2 & \cpu{3652.91}	& \ints{101150}	\\
-			%[JOB 8] progress: 97913/161897 brs (60.478362%), ETA 1h 29m 30s (elapsed 59m 52s)
-			% [JOB 8] TIMED OUT (60.478362%; ETA 1h 29m 30s)
-
-			& 3,1 & \ETAdag{8d 12h} & \ETAdag{411312}	\\
-			%[JOB 8] progress: 103873/411312 brs (25.254021%), ETA 8d 11h 36m 0s (elapsed 59m 42s)
-			%[JOB 8] TIMED OUT (25.254021%; ETA 8d 11h 36m 0s)
-			\hline
-			{\tt avl\_insert}
-			& 2,1 & \cpu{1083.35}	& \ints{40062}	\\
-			%& 2,2 & \multicolumn{2}{c}{\em (bug!)} \\
-			\hline
-			{\tt avl\_fixed}
-			& 2,1 & \cpu{1079.03} 	& \ints{45078} 	\\
-			& 2,2 & \ETAdag{2762y}	& \ETAdag{8714863} 	\\
-			%[JOB 19] progress: 45863/8714863 brs (0.526262%), ETA 2762y 28d 9h 57m 14s (elapsed 59m 25s)
-			%[JOB 19] TIMED OUT (0.526262%; ETA 2762y 28d 9h 57m 14s)
-			% the below is from avlfixed2,2 after running it overnight
-			% eta is ever worse but i'm using the 1hr one as promised
-			%[JOB 19] progress: 315672/29992026 brs (1.052520%), ETA 321008y 136d 18h 22m 12s (elapsed 6h 58m 15s)
-			& 3,1 & \ETAdag{12d 2h}	& \ETAdag{11498545}	\\
-
-			\hline
-			{\tt map\_basic}
-			& 2,1 & \ETAdag{10d 17h} & \ETAdag{16388977} \\
-			%[JOB 15] progress: 70330/16388977 brs (0.429130%), ETA 10d 16h 34m 15s (elapsed 43m 21s)
-			%[JOB 15] TIMED OUT (0.429130%; ETA 10d 16h 34m 15s)
-			\hline
-			{\tt map\_basicer}
-			& 2,1 & \cpu{877.44}& \ints{28635}	\\
-			& 2,2 & \ETAdag{2d 7h} & \ETAdag{5925634} \\
-			& 3,1 & \ETAdag{468d 13h} & \ETAdag{35893653} \\
-			% [JOB 5] progress: 100181/35893653 brs (0.279105%), ETA 1y 103d 12h 50m 47s (elapsed 59m 24s)
-			%[JOB 5] TIMED OUT (0.279105%; ETA 1y 103d 12h 50m 47s)
-			% i don't know what this is but it was hanging out under 3,1 here :<
-			%[JOB 6] progress: 35042/5925634 brs (0.591363%), ETA 2d 6h 40m 30s (elapsed 21m 12s)
-			%[JOB 6] TIMED OUT (0.591363%; ETA 2d 6h 40m 30s)
-		\end{tabular}
-	\end{center}
-	\caption{Transactional tests verified (or not) by Landslide.}
-	\label{tab:verifs}
-\end{table}
-
-Several larger tests we were unable to complete before the submission deadline are also shown.
-%Their completion times are listed as {\em unk.}
-These are indicated with $\dagger$ %\ETAdag{}
-and their listed ETAs and state space sizes represent Landslide's estimate %at the moment of timeout.
-after a timeout of 1 hour
-(hopefully enough for data-race PPs to saturate and estimates to stablize,
-although note the two estimate types use different algorithms so may disagree significantly).
-%For these, Table~\ref{tab:timeouts} presents their size estimates at the moment of timeout.
-Though these tests proved beyond our reach,
-in contrast with the instant gratification expected of bug-finding,
-%we believe
-the verification guarantee may in some cases be worth the estimated time requirement
-for widely-used industrial implementations.
-Future work may also expand our coverage upon this frontier \cite{landslide-phdthesis}.
-
-\subsection{Variance}
-
-Because many of the verification tests are long-running,
-and we are writing this too close to the submission deadline (who doesn't),
-we regret being unable to present every performance measurement above
-as an average of multiple samples with error bars \cite{epsilon}.
-Nevertheless, we make some effort to address variance.
-
-We noticed significant slowdowns when varying several aspects of our experimental environment.
-For example, multiple Landslides running at once
-%on our test machine
-slow each other down,
-%(including multiple single state spaces in a single test when not using {\tt -M}),
-likely arising from kernel resource contention as Landslide uses {\tt fork()} to save simulation state.
-Table~\ref{tab:variants} shows the impact of running a single Landslide instance with {\tt -M} on
-{\tt counter}(2,2) %(384 distinct interleavings)
-with various programs also running, despite never saturating our test system's 8 CPUs.
-% for i in `seq 1 10`; do ./landslide -X -v -p counter -M 2>&1 | grep 'total CPU time'; done
-
-Table~\ref{tab:variants2} shows
-the impact of Landslide needing to automatically annotate {\tt counter}(2,2) being run for the first time,
-rather than reusing existing instrumentation,
-as well as the variance of Quicksand mode. %(i.e., without {\tt -M}).
-Because Quicksand repeats more work across jobs,
-rather than comparing it to a baseline,
-we show in the right column how many total interleavings each approach executed.
-(The maximal state space alone comprises 384 in all cases.)
-%(2128 interleavings total).
-The {\sf quicksand} variance was surprisingly bimodal,
-with 6 samples in a distribution of 106.14$\pm$1.22 and 4 in 142.11$\pm$1.67,
-suggesting two distinct scheduling patterns for its workqueue threads.
-Future work should figure out why.
-
-We conclude that MC performance evaluations must address
-experimental environment variables
-%at least enough
-to ensure consistent performance between runs.
-So doing, multiple samples and error bars are then necessary
-only when using nondeterministic search ordering strategies
-such as Iterative Deepening.
-Otherwise, considering the low variance shown in Table~\ref{tab:variants},
-they need be shown only for a token small test
-to provide reader some assurance of similar consistency in the larger tests
-(especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
-% We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper
-%% i'm not actually sure this fits..
-%Finally, considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
-%we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
-
-%Outside of this experiment,
-For all tests outside of Table~\ref{tab:variants},
-we fixed our environment by measuring all performance numbers
-with Firefox and Chrome as the only other significant machine load.
-%(we gotta stay amused somehow).
-We believe the exponential differences among completion times justifies the absence of error bars,
-which one would expect to show 2\% variance if extrapolating from these results. %this experiment.
-The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Limitations}
-\label{sec:warpzone}
-
-%This section discusses opportunities for future work.
-This section should be mandatory in all systems papers.
-
-{\bf Transaction failure codes.}
-When a transaction fails, {\tt \_xbegin()} returns a failure code
-denoting the reason, or combination thereof, therefor \cite{htm-gcc}.
-If a program then cases on that failure code to select between different backup paths,
-model checking it by simply injecting a single type of failure may be unsound.
-For example, a program executing a transaction which is guaranteed to never conflict with any other threads,
-and hence never abort without {\tt \_XABORT\_RETRY},
-could legally {\tt assert(false);} in its failure path,
-while our approach in Section~\ref{sec:design} would erroneously trigger that assertion and report a bug.
-Likewise, the transactional data structures from \cite{htm-mario}
-abstract away any spurious {\tt \_XABORT\_RETRY} aborts behind a retry loop;
-a MC unwise to that idiom would call it an infinite loop bug.
-
-Our HTM implementation includes an experimental feature
-to track the set of abort codes possible for each transaction.
-%According to the any-reason principle in Section~\ref{sec:background},
-{\tt \_XABORT\_RETRY} is always enabled;
-then,
-it harnesses DPOR's existing memory analysis to identify when {\tt \_XABORT\_CONFLICT} is possible,
-and instruments {\tt \_xabort()} calls to record any user-supplied codes for {\tt \_XABORT\_EXPLICIT}.
-This comes at a cost of even more state space explosion,
-increasing the exponent at each {\tt \_xbegin()} preemption point
-by (usually) 1 plus however many distinct {\tt \_xabort()} codes the program uses.
-
-\newcommand\vsx[1]{\hilight{darklavender}{#1}\xspace}
-\begin{table}[t]
-	\begin{center}
-	\begin{tabular}{l|c|c|c}
-		\bf load & \cpu{\bf cpu (s)} & \vsx{\bf vs self avg} & \vsx{\bf vs baseline} \\
-		\hline
-		{\sf none	} & \cpu{14.95 $\pm$ 0.17} & \vsx{0.99-1.02x}	& \vsx{(baseline)}	\\
-		{\sf vid-L	} & \cpu{15.13 $\pm$ 0.10} & \vsx{0.99-1.01x}	& \vsx{1.01x}	\\
-		{\sf ff/c	} & \cpu{15.55 $\pm$ 0.16} & \vsx{0.99-1.02x}	& \vsx{1.04x}	\\
-		{\sf sm5	} & \cpu{16.63 $\pm$ 0.07} & \vsx{0.99-1.01x}	& \vsx{1.11x} \\
-		{\sf ff/c+vid-S	} & \cpu{17.09 $\pm$ 0.34} & \vsx{0.98-1.05x}	& \vsx{1.14x}	\\
-		{\sf ls		} & \cpu{19.11 $\pm$ 0.32} & \vsx{0.97-1.03x}	& \vsx{1.28x}	\\
-		% simply love title screen
-		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& \vsx{0.97-1.02x   } & \vsx{1.28x} \\
-		{\sf ff/c+ls	} & \cpu{19.66 $\pm$ 0.50} & \vsx{0.96-1.02x}	& \vsx{1.31x}	\\
-	\end{tabular}
-	\end{center}
-	\caption{Performance variance on {\tt counter}(2,2) with other programs running on the test machine.
-		{\sf vid-L} is full-screen video (played locally with {\tt mplayer}),
-		{\sf ff/c} is Firefox and Chrome (idle, $\le$20 tabs),
-		{\sf sm5} is StepMania 5.1 (during gameplay \cite{itg2}),
-		{\sf vid-S} is full-screen video (streamed via Crunchyroll),
-		{\sf ls} is a 2nd instance of Landslide.
-		Average of 10 samples, $\pm N$ is 1 stddev.
-	}
-	\label{tab:variants}
-\end{table}
-
-\begin{table}[t]
-	\begin{center}
-	\begin{tabular}{l|c|r}
-		\bf Landslide mode & \cpu{\bf cpu (s)} & \ints{\bf total int's} \\
-		\hline
-		{\sf verif} ({\tt -M})		& \cpu{~~14.95 $\pm$ 00.17} &  \ints{403} \\
-		{\sf reinstrument}		& \cpu{~~24.72 $\pm$ 00.10} &  \ints{403}     \\
-		{\sf quicksand} (no {\tt -M})	& \cpu{120.53 $\pm$ 18.62}  & \ints{2128}     \\
-	\end{tabular}
-	\end{center}
-	\caption{Performance variance on {\tt counter}(2,2) in various modes.
-		Average of 10 samples, $\pm N$ is 1 stddev.
-	}
-	\label{tab:variants2}
-\end{table}
-
-Many such branches may be equivalent;
-for example,
-explicit and conflict aborts need not be tested separately in transactions
-whose failure paths do not distinguish the cause,
-and the perennial {\tt \_XABORT\_RETRY} abort can be skipped entirely
-if the client abstracts it away behind a retry loop.
-In fact, applying both reductions simultaneously
-amounts to the STM concurrency model \cite{stm-pldi06}%
-\footnote{Proof left to future work. We've proved enough here already.}
-and may reduce the state space even smaller than the original.
-Testing them by hand
-\footnote{Using visual inspection of the test to trust the reductions' soundness.}
-reduced {\tt map\_basicer}(2,1)'s 28635 interleavings to 11577,
-and produced the same 384-interleaving state space on {\tt counter}(2,2)
-(in which DPOR triggers conflict aborts on every transaction anyway).
-%and {\tt map\_basic}(2,1)'s 1-hour size estimate from 16388977 to 16126960. % about the same after all...
-%(although {\tt map\_basic}(2,1) lies still beyond reach).
-%We tested both reductions by hand, trusting their soundness by visual inspection,
-%on {\tt map\_basic}(2,1),
-%and preview the resulting state space sizes in Table~\ref{tab:abort-codes}.
-%
-%\begin{table}[h]
-%	\caption{...}
-%	\label{tab:abort-codes}
-%\end{table}
-%
-Future work could use static or dynamic flow analysis to
-identify such reduction opportunities automatically;
-for now, this feature is disabled by default but accessible via the {\tt -A} command-line option
-%to Landslide
-(in addition to {\tt -X}).
-
-{\bf Relaxed memory orderings.}
-Section~\ref{sec:design}'s formalization of thread interleavings does not account for read/write reorderings
-possible on relaxed consistency architectures \cite{memory-consistency-models}.
-In fact,
-even after \cite{htm-mario}'s proposed fix,
-our running example program is still incorrect on Total Store Order (TSO) architectures such as x86.
-Despite stores being totally-ordered, x86 may still reorder stores after subsequent loads.
-Accordingly, an execution of $8,9a,9b,9c$ in Figure~\ref{fig:htm-fixed}
-may be locally visible to another thread as $9a,8,9b,9c$,
-and hence an apparent interleaving of
-\[
-	\tiat{1},\tjat{1-5},\tiat{7},\underline{\tiat{9a}},\tkat{1-5},\underline{\tiat{8}},\tiat{9b-B}
-\]
-is possible
-(reordered accesses underlined for emphasis).
-An {\tt mfence} barrier is needed between lines 8 and 9 to solve this problem on TSO \cite{tsx-need-barrier}.
-On Partial Store Order (PSO) architectures, even more barriers may be necessary.
-
-Because Landslide's concurrency model includes only instruction-level thread nondeterminism,
-not per-CPU memory buffer reorderings,
-our current HTM implementation cannot find this bug.
-In fact, it erroneously verifies the corresponding test {\tt htm2}(3,1)
-% lol this is both old (quicksand) AND wrong (testing (2,1) which can't expose the bug)
-%in 7 CPU-minutes,
-%with 294 distinct interleavings in total,
-in 3 CPU-hours,
-with 467730 distinct interleavings in total,
-none of which include the above-listed sequence.
-Recent work extended DPOR to support TSO and PSO memory nondeterminism \cite{tsopso};
-if incorporated into Landslide, we could find or verify the absence of such bugs.
-Visual inspection of \cite{htm-mario}'s HTM data structures found no barriers used in this implementation pattern;
-we would urge any reader interested in using those to add them in by hand first.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Conclusion}
-
-Stateless model checking research is a perpetual existence of staring up the sheer cliff face that is the exponential curve.
-As new concurrency paradigms emerge,
-we make our living by adapting our reduction algorithms and search strategies to them to climb ever higher on that curve,
-eking out a few more loop iterations or slightly higher thread counts in our verification guarantees.
-Whether that is beautiful in its imperfection or cause for despair is merely a matter of perspective.
-Also, the author hopes their committee won't mind the publication venue when they cite this in their thesis.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\acks
-
-We thank Mario Dehesa-Azuara for generously providing the HTM data structures used in our evaluation,
-% no
-%and the anonymous SIGBOVIK reviewers for their helpful comments.
-the anonymous SIGBOVIK program committee,
-and anyone who actually read this unusual paper all the way through.
-%We also thank the member companies of the PDL Consortium (Broadcom,
-%Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
-%Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
-%Tintri, Two Sigma, Uber, Veritas and Western Digital).
-This work was supported in part by the U.S. Army Research Office under grant number W911NF0910273.
-
-\bibliographystyle{abbrv}
-\bibliography{citations}
-
 \end{document}

### 2018-03-30 htm/for-garth/paper.tex
update forg-arth

@@
 	%
+	& \hilight{lavender}{\em transaction fails} & & & \\
 	\hilight{lavender}{$5$} & \hilight{lavender}{\texttt{mutex\_lock(m);}} & & \\
@@
 Prior work \cite{htm-mario} proposed the idiom shown in Figure~\ref{fig:htm-fixed}
-to exclude interleavings which violate the failure path's atomicity.
+to exclude interleavings which violate the failure path's atomicity such as shown above.
 Thread 1 would set the {\tt prevent\_transactions} flag across its {\tt x++} accesses,
@@
 
+\newpage
 \section{Implementation}
@@
 \subsection{Bugs}
+\label{sec:bugs}
 
@@
 	\caption{Landslide's bug-finding performance on various test configurations.
+		A parameter of $K,N$ denotes $K$ threads each running the test logic for $N$ iterations.
 		We list the CPU-time and wall-clock time %(both in seconds)
@@
 	%\caption{mario man had a very large house. or a castle.}
-	\caption{Unmodified code from {\tt htmavl.cpp} showing the previously-unknown bug found by Landslide.
+	\caption{Unmodified code from {\tt htmavl.hpp} showing the previously-unknown bug found by Landslide.
 		The transaction path fails to check {\tt \_retry},
@@
 
+\newpage
 \subsection{Verification}
@@
 after a timeout of 1 hour
-(hopefully enough for data-race PPs to saturate and estimates to stablize,
-although note the two estimate types use different algorithms so may disagree significantly).
+(hopefully enough for data-race PPs to saturate and estimates to stablize).
+Note that the two estimate types use different underlying algorithms, so may disagree significantly,
+such as in {\tt avl\_fixed}$(2,2)$.
+
+While we verified {\tt avl\_fixed} for 2 threads and 1 iteration,
+the buggy {\tt avl\_insert} actually required higher parameters ($2,2$ or $3,1$) to expose the bug in the first place,
+either of which produced too large of a state space under {\tt avl\_fixed},
+so we were unable to formally verify that our fix actually removed the bug.
+Likewise, {\tt map\_basic}, which includes a concurrent map-resizing step,
+could not be verified even under the minimum parameters ($2,1$);
+at best we were able to verify a modified version {\tt map\_basicer}
+which started with a pre-sized map to test inserts only.
+Possible approaches to bring the sizes of these state spaces within reach
+include testing the synchronization mechanism and the data structure logic separately, as discussed in Section~\ref{sec:bugs},
+upgrading Landslide's exploration algorithms to achieve better reduction,
+\cite{optimal-dpor,mcr}.
+and identifying new types of equivalent interleavings under the new concurrency model \cite{thesprop}.
 

### 2019-03-11 itg/3/paper.tex
begin writing itg3

--- 2017-03-15 itg/2/paper.tex
@@
 %\usepackage{mathpartir}
-\usepackage{hyperref}
+\usepackage[colorlinks=true,
+citecolor=citec,
+linkcolor=linkc,
+urlcolor=urlc,
+]{hyperref}
 \usepackage{url}
@@
 \usepackage{multirow}
+\usepackage[nameinlink,nosort]{cleveref}
+\usepackage{xeCJK}
 \usepackage[usenames,dvipsnames]{xcolor}
@@
 \usepackage{natbib}
+\usepackage{ulem}
 \usepackage[mathcal]{euscript}
@@
 
+\definecolor{citec}{RGB}{128,0,64}
+\definecolor{linkc}{RGB}{0,64,128}
+\definecolor{urlc} {RGB}{128,64,0}
+
+\setCJKmainfont{HeiseiMinStd-W5}[Path = ./]
 % This Deep Tex Voodoo is from
@@
 \title{
-A Boring Follow-Up Paper to \\
+Which ITG Stepcharts are Bracket-Jumpiest?: \\
+In Which They Milk the \\
+「A Boring Follow-Up Paper to \\
 ``Which ITG Stepcharts are Turniest?'' \\
-Titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?'' \\
+Titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?''」\\
+Series for All Its Worth in Publication Count After All \\
 %That `Boring' Stuff Was Part of the Title, BTW. \\
@@
 
-In which I deliver on last year's promise of future work.
+In which I break last last year's promise of no future work.
 
@@
 \keywords
-crossovers, footswitches, jacks, sidefoots
+brackets, jumps, bracket jumps
 
@@
 
-Let's resume right where I left off in my last paper \cite{turniness}, shown in Figure~\ref{fig:you-stutid-fuckass}.
-Unlike mainstream conferences, SIGBOVIK doesn't make me waste space repeating all the background material,
-and I can just say go read that paper first and get back to me.
-It's probably a lot funnier than this one anyway, which is gonna be sort of dry,
-and really of interest only to other ITG players who already know what's going on.
+Recent work by \cite{dril}
+%[W. Dril et al., 2019]
+proposed the hypothesis that
+recent stepchart authors have grown bored with the array of technical ITG step patterns
+documented to date \cite{turniness,crossoveriness},
+and have moved on to break the old model's one-foot-per-arrow assumption to allow for even more technical patterns yet.
+I paraphrase their main conclusion as follows:
+
+\begin{center}
+	\includegraphics[width=0.42\textwidth]{using-my-turn.png}
+\end{center}
+
+All good researchers know that when rise the standards for software or hardware performance
+(or stepchart trickiness, as the case may be),
+they must revisit their own work to prove its ongoing relevance to the research (dance) community at large.
+Thus I must regrettably break the promise \sout{I} the authors set forth in \cite{crossoveriness}
+(see \Cref{fig:you-stutid-fuckass})
+and revisit \sout{my} their old future work section
+(XXX: they said to change any first-person ``our prior work''
+stuff like this for double-blind review but like cmon this makes no sense?
+theyll totally see through this
+(TODO: maybe email the PC chair for advice?
+(FIXME: make sure to remove these comments before the camera-ready deadline!!))),
+extending it to handle these new so-called ``bracket'' jumps.
 
 \begin{figure}[h]
-	\hspace{-2em}\includegraphics[angle=90,width=0.54\textwidth]{../paper.pdf}
-	\caption{(okay twist your head to read this)}
+	\hspace{-2em}\includegraphics[angle=270,width=0.54\textwidth]{../2/paper.pdf}
+	\caption{(yeah i reused this joke from last time ok deal)}
 	\label{fig:you-stutid-fuckass}
-\end{figure}
-
-The TL;DR is that I made a program which figures out how to foot stepcharts in the least crossovery possible way (short of double-stepping everything),
-then found which charts ultimately had the most.
-The algorithm also naturally identifies footswitches and jacks, %, jacks, forced double-steps, and crossover-footswitches;
-and sometimes it's smarter than me in amusing ways.
-I put all the goodies in a giant spreadsheet at \url{http://tinyurl.com/crossoveriest},
-and the program itself is of course freely available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Revisiting Turniness (Flashback Scene)}
-
-Recall Table~\ref{tab:facing} from the last paper, in which I left undefined the facings for $LL$, $DD$, $UU$, and $RR$, the four footswitches.
-I show a typical $DD$/$UU$ footswitch pattern in Figure~\ref{fig:fs}(a), and typical $LL$/$RR$ switches (henceforth ``crossover footswitches'') in Figure~\ref{fig:fs}(b).
-To step these patterns, the player still alternates feet as usual, but must lift one foot off the repeated arrow before stepping it with her other foot.
-Chart authors will often, but not always, include a ``mine cue'' (shown in the figure) to hint that the second foot should switch onto the same arrow.
-
-\begin{table}[t]
-	\begin{center}
-	\begin{tabular}{cc|cccc}
-		& & \multicolumn{4}{c}{Right foot} \\
-		& & $\leftarrow$ & $\downarrow$ & $\uparrow$ & $\rightarrow$ \\
-		\hline
-		\multirow{4}{*}{Left foot}
-		%                 L   D   U   R
-		& $\leftarrow$  & \bf ? & $UR$ & $UL$ & $U$ \\
-		& $\downarrow$  & $DL$ & \bf ? & $L$ & $UL$ \\
-		& $\uparrow$    & $DR$ & $R$ & \bf ? & $UR$ \\
-		& $\rightarrow$ & $D$ & $DR$ & $DL$ & \bf ? \\
-
-	\end{tabular}
-	\end{center}
-	\caption{Facing directions.
-	%``Crossover'' facings are marked (*), and the ``lateral'' facing is marked ($\dagger$). Note the appealing diagonal symmetry.}
-	}
-	\label{tab:facing}
-\end{table}
-
-\begin{figure}[t]
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-dd-uu.png} \\
-	(a) DD, UU
-	\end{tabular}
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-ll-rr.png} \\
-	(b) LL, RR
-	\end{tabular}
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-ll-rr-2.png} \\
-	(c) RR, LL
-	\end{tabular}
-	\caption{Footswitches of various crossoveriness/facing.}
-	\label{fig:fs}
-\end{figure}
-
-It is tempting to assign the facings $L$, $U$, $U$, and $R$ respectively to the $LL$, $DD$, $UU$, and $RR$ footswitches.
-However, Figure~\ref{fig:fs}(c) shows that if a footswitch begins with a crossover on $U$, the facing should be reversed:
-the $RR$ footing should face $L$, and $LL$ should face $R$.
-``Spin-switches'' with $D$ facing are also theoretically possible,
-arising from patterns such as $LURDDL$ or $LDRUUL$, although I know of no real-world chart which does this (hmm, tempting...).
-
-Before I realized that, I modified the turniness algorithm \cite{turniness} to face footswitches as above, %the bad way mentioned above,
-%to be faced $L$ for $LL$ and so on, as above,
-and it surprised me with charts of $\mathcal{T}>2$, in excess of the theoretical maximum!
-I show one such chart in Figure~\ref{fig:fs-turniness}(a), in which the step from $LL$ ($\phi=L$) to $UL$ ($\phi=DR$) has individual $\mathcal{T}=3$, and so on for $UL \rightsquigarrow_R UU \rightsquigarrow_L RU \rightsquigarrow_R RR$.
-The steps $RR \rightsquigarrow_L LR \rightsquigarrow_R LL$ are both candles ($\mathcal{T}=2$),
-resulting overall in $\mathcal{T}=8/3$ for the whole chart.
-
-Indeed, when I further modified the algorithm to force $DD$ switches to face $D$ (i.e., always facing the direction of the repeated arrow),
-it produced the chart shown in Figure~\ref{fig:fs-turniness}(b), with overall $\mathcal{T}=3$.
-(Note its resemblance to the basic spin pattern, $LDRU$, whose $\mathcal{T}=2$.)
-
-To fairly represent a human player's desire to step in the least turny of ambiguous ways,
-I extended the algorithm to provide either the assigned facing, from above, or its polar opposite,
-chosen at runtime by whichever is closer to the presequent facing.
-This restores the maximum overall chart turniness to $\mathcal{T}=2$,
-a new example of which is shown in Figure~\ref{fig:fs-turniness}(c).
-
-However, note that individual steps may still have $\mathcal{T}=3$,
-%because after a footswitch's facing is determined, the next step may still ... ?
-as shown in Figure~\ref{fig:fs-turniness}(d).
-In this example, the step $DL \rightsquigarrow_L LL$ assigns $LL$ to face $L$,
-but the subsequent step to $LD$ cannot avoid facing $UR$.
-The reason charts still cannot exceed overall $\mathcal{T}=2$ is that
-setting up such a situation requires a $\mathcal{T}=1$ step,
-which negates the benefit.
-A chart could conceivably end right before such a step,
-sneaking through some small $\epsilon$ extra turniness \cite{epsilon}
-(similar to the case of 270s in \cite{turniness}),
-but {\em sustained} average $\mathcal{T}>2$ remains impossible.
-
-Another approach could assign such a footswitch the opposite footing of the previous facing,
-regardless of the arrow itself;
-so in this case the $LL$ would face $UL$, and each step would have exactly $\mathcal{T}=2$.
-Anyway, remember that all-candle switch stream in (c). We'll see it again soon.
-
-\begin{figure}[t]
-	\begin{center}
-	\begin{tabular}{c}
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-turniness-8-3.png} \\
-	(a) $\mathcal{T}=8/3$(?)%\stackrel{?}{=}8/3$
-	\end{tabular}
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-turniness-3.png} \\
-	(b) $\mathcal{T}=3$(?)%\stackrel{?}{=}3$
-	\end{tabular}
-	\\
-	\\
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-turniness-2.png} \\
-	(c) $\mathcal{T}=2$.
-	\end{tabular}
-	\begin{tabular}{c}
-	\includegraphics[width=0.13\textwidth]{fs-turniness-2-plus-epsilon.png} \\
-	(d) $\mathcal{T}=2+\epsilon$.
-	\end{tabular}
-	\end{tabular}
-	\end{center}
-	\caption{The turniest footswitch patterns. (a) and (b) are false positives (see prose), while (c) and (d) provide theoretically maximal turniness.}
-	\label{fig:fs-turniness}
-\end{figure}
-
-% dont forget
-% http://dancedancerevolutionddr.wikia.com/wiki/Afronova_walk
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Analyzing Crossoveriness}
-
-The major flaw of the turniness algorithm \cite{turniness}
-was that it didn't care whether a stream started with the left or right foot;
-it simply exploited the symmetry of Table~\ref{tab:facing} to find turniness regardless of footing.
-Hence, it could not distinguish technical footing patterns which could affect the way a human would play the chart.
-It often played charts inhumanly, facing backwards and/or stepping 270s for most of a song.
-
-So, my contribution this year is an algorithm which plays more naturally,
-and which consequently can report on a chart's technical patterns beyond simple turniness.
-The algorithm realizes three principles of ITG:
-\begin{enumerate}
-	\item Alternate feet as much as possible.
-	\item Step crossed-over as little as possible.
-	\item Jumps or jacks allow the player to reset her footing.
-\end{enumerate}
-
-Figure~\ref{fig:pseudocode} describes the algorithm in pseudocode.
-To summarize it in prose:
-\begin{itemize}
-	\item Split the chart into several units of stream,
-		the boundaries of which occur at every jump and any time an arrow is repeated.
-	\item Step each stream with alternating feet.
-	\item Compare the number of matching steps (i.e., $L$ foot on $L$ arrow or $R$ on $R$) versus crossover steps.
-		If the latter is greater, re-step the stream with opposite feet from before (this kills the crossovers).
-	\item After flipping each stream, if necessary, count the total crossovers in the whole chart.
-\end{itemize}
-
-\newcommand\hilight[2]{\color{#1}#2\color{black}}
-\definecolor{pink}{RGB}{128,0,192}
-\definecolor{orange}{RGB}{192,96,0}
-\definecolor{olivegreen}{RGB}{0,127,32}
-\definecolor{brickred}{RGB}{192,0,0}
-\definecolor{commentblue}{RGB}{0,128,192}
-
-\begin{figure*}[p]
-\begin{center}
-\begin{tabular}{l}
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step}~= \hilight{brickred}{L}~| \hilight{brickred}{D}~| \hilight{brickred}{U}~| \hilight{brickred}{R}~| \hilight{brickred}{Jump}~\hilight{orange}{deriving}~\hilight{olivegreen}{Eq}} \\
-\texttt{} \\
-\texttt{\hilight{orange}{data}~\hilight{olivegreen}{AnalysisState}~= \hilight{brickred}{S}~\{ steps :: \hilight{olivegreen}{Int}, xovers :: \hilight{olivegreen}{Int}, switches :: \hilight{olivegreen}{Int}, jacks :: \hilight{olivegreen}{Int},} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastStep :: \hilight{olivegreen}{Maybe}~\hilight{olivegreen}{Step}, doubleStep :: \hilight{olivegreen}{Bool}, lastFlip :: \hilight{olivegreen}{Bool},} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ lastFoot :: \hilight{olivegreen}{Bool}, stepsLR :: [\hilight{olivegreen}{Bool}] \}} \\
-\texttt{} \\
-\texttt{\hilight{pink}{commitStream}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{AnalysisState}} \\
-\texttt{\hilight{pink}{commitStream}~s = s \{ xovers~~ = xovers~~ s + \hilight{orange}{if}~f \hilight{orange}{then}~ns - nx \hilight{orange}{else}~nx, } \\
-\texttt{~~~~~~~~~~~~~~~~~~~~ switches = switches s + \hilight{orange}{fromEnum}~(f == lastFlip s \&\& doubleStep s), } \\
-\texttt{~~~~~~~~~~~~~~~~~~~~ jacks~~~~= jacks~~~~s + \hilight{orange}{fromEnum}~(f /= lastFlip s \&\& doubleStep s), } \\
-	\texttt{~~~~~~~~~~~~~~~~~~~~ lastFlip = f, stepsLR = \hilight{brickred}{[]}~\}} \\
-\texttt{~~~~\hilight{orange}{where}~ns = \hilight{orange}{length}~\$~stepsLR s} \\
-\texttt{~~~~~~~~~~nx = \hilight{orange}{length}~\$ \hilight{orange}{filter not}~\$ stepsLR s} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- if more than half the L/R steps in this stream were crossed over,}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- then we got the footing backwards and need to flip the stream.}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- as a tiebreaker, flip if the chart is already more jacky than}} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- footswitchy, i.e., if past streams flipped more often than not.}} \\
-\texttt{~~~~~~~~~~f = nx * 2 > ns || nx * 2 == ns \&\& ((switches s > jacks s) == lastFlip s)} \\
-\texttt{} \\
-\texttt{\hilight{pink}{analyzeStep}~:: \hilight{olivegreen}{AnalysisState}~-> \hilight{olivegreen}{Step}~-> \hilight{olivegreen}{AnalysisState}} \\
-\texttt{\hilight{pink}{analyzeStep}~s step} \\
-\texttt{~~~~\hilight{commentblue}{-{}- a jump resets the footing, so the next step can be stepped with either}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- foot. commit the stream so far to treat it separately from what follows.}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- bracket-jumps are, of course, future work.}} \\
-\texttt{~~~~| step == \hilight{brickred}{Jump}~= (\hilight{pink}{commitStream}~s) \{ lastStep = \hilight{brickred}{Nothing}, doubleStep = \hilight{brickred}{False}~\}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- two steps on the same arrow might be a jack, or might be a footswitch.}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- to figure out which, commit the stream so far, and begin a new stream}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- whose footing will retroactively determine how to foot this step.}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- also, unlike jumps, this step gets counted as part of the next stream.}} \\
-\texttt{~~~~| lastStep s == \hilight{brickred}{Just}~step = stream (\hilight{pink}{commitStream}~s) \{ doubleStep = \hilight{brickred}{True}~\}} \\
-\texttt{~~~~\hilight{commentblue}{-{}- a normal streamy step.}} \\
-	\texttt{~~~~| \hilight{orange}{otherwise}~= stream s} \\
-\texttt{~~~~\hilight{orange}{where}~foot = \hilight{orange}{not}~\$ lastFoot s} \\
-\texttt{~~~~~~~~~~\hilight{commentblue}{-{}- record whether we stepped on a matching or crossed-over L/R arrow.}} \\
-\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{L}~steps = steps ++ [ft]} \\
-\texttt{~~~~~~~~~~addStep ft \hilight{brickred}{R}~steps = steps ++ [\hilight{orange}{not}~ft]} \\
-\texttt{~~~~~~~~~~addStep ft \_ steps = steps \hilight{commentblue}{-{}- U/D don't help to determine L/R footing.}} \\
-\texttt{~~~~~~~~~~stream s = s \{ steps = steps s + \hilight{brickred}{1}, lastStep = \hilight{brickred}{Just}~step, lastFoot = foot,} \\
-\texttt{~~~~~~~~~~~~~~~~~~~~~~~~ stepsLR = addStep foot step \$ stepsLR s \} } \\
-\texttt{} \\
-\texttt{\hilight{pink}{analyze}~:: [\hilight{olivegreen}{Step}] -> \hilight{olivegreen}{AnalysisState}} \\
-	\texttt{\hilight{pink}{analyze}~= \hilight{pink}{commitStream}~.
-\hilight{orange}{foldl}~\hilight{pink}{analyzeStep}~(\hilight{brickred}{S 0 0 0 0}~\hilight{brickred}{Nothing}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{False}~\hilight{brickred}{[]}) } \\
-\end{tabular}
-\end{center}
-	\caption{Pseudocode description of the crossoveriness and footswitchiness algorithm.}
-\label{fig:pseudocode}
-\end{figure*}
-
-\section{Analyzing Footswitchiness}
-
-Because we split the chart whenever an arrow is repeated,
-figuring out whether that arrow is stepped with different feet on either side of the stream boundary
-is a natural consequence of figuring out how to step each stream individually.
-This is also shown in Figure~\ref{fig:pseudocode}'s pseudocode.
-To summarize in prose,
-if neither stream needed to be flipped (or if both did), then the alternating feet assumption holds,
-and the repeat must be a footswitch.
-
-\section{Analyzing Jackiness}
-
-A jack occurs when a repeat arrow is stepped with the same foot,
-rather than alternating (hence the name, from ``jackhammer'').
-You've got the idea by now, right?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Forced Doublesteps}
-
-After painstakingly translating the pseudocode from Figure~\ref{fig:pseudocode} into a real implementation,
-I found it vulnerable to false positives when a single double-step could force a long section of stream to be stepped backwards.
-As an extreme example, consider the pattern $LRLRLRLR-D-LRLRLRLR$.
-Because no jumps or jacks allow the footing to be reset, either the first or last 4 pairs of $LR$s must be stepped crossed-over.
-
-Figure~\ref{fig:force-doublestep} shows two examples from real-world charts:
-in (a), the player must not alternate feet across the measure break,
-while in (b), two $L$ arrows are replaced with rolls for an artistic visual accent, which must be stepped twice each.
-On inspection, these charts should be stepped with no crossovers, but were evaluated otherwise %as having very high XO\%
-%(anywhere from 10.6\% to 27.9\%).
-(730 XOs (27.9\%) and 173 XOs (9.6\%), respectively).
-
-\begin{figure}[t]
-	\begin{center}
-	\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{paradise-lost-false-positive.png}
-		\\
-		(a) Paradise Lost \\
-		(16, Cirque du Lykan)
-	\end{tabular}
-	%\begin{tabular}{c}
-	%	\includegraphics[width=0.18\textwidth]{false-positive-toromi-hearts.png}
-	%	\\
-	%	(a) Toromi Hearts 2 \\
-	%	(18, Cirque du Zonda)
-	%\end{tabular}
-	\begin{tabular}{c}
-		\includegraphics[width=0.18\textwidth]{hearbeat-rollstream-doublestep.png}
-		\\
-		(b) Heartbeat \\
-		(13, TranceNation)
-	\end{tabular}
-	\end{center}
-	\caption{The doublesteps in some streamy charts must be identified, and the stream split, lest ``too much'' of the following stream appear completely crossed-over.}
-	\label{fig:force-doublestep}
-\end{figure}
-
-\newcommand\Stream{\ensuremath{\mathcal{S}}}
-\newcommand\Doublestep{\ensuremath{\mathcal{DS}}}
-\newcommand\nflip{\ensuremath{n_{\mathfrak{flip}}}}
-\newcommand\percentflip{\ensuremath{\%_{\mathfrak{flip}}}}
-\begin{algorithm}[h!]
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Invariant}
-	\Input{\Stream, a step sequence $s_0 \dots s_n$}
-	\Output{$\forall s_i, s_j \in \Stream, j=i+1 \rightarrow \mathsf{\neg StreamBoundary}(s_i,s_j)$}
-	\Input{\nflip, heuristic minimum length} %of flipped stream to doublestep into}
-	\Input{$\percentflip$, heuristic percentage, initially 100}
-	\For{$i \in \mathsf{length}(\Stream) \wedge \neg {\sf defined}(i_{\Doublestep})$}{
-		$\Stream' \leftarrow \{ s_k | s_k \in \mathsf{LRs}(\{ s_j | s_j \in \Stream \wedge j \ge i \}) \wedge k \le \nflip \}$ \\
-		\If{$|\Stream'| = \nflip$ $\wedge$ $|\mathsf{Crossovers}(\Stream')| \ge \percentflip \times |\Stream'|$}{
-			$i_{\Doublestep} \leftarrow i$
-		}
-	}
-	\eIf{{\sf defined}($i_{\Doublestep})$}{
-		\If{$i_{\Doublestep} = 0$}{
-			% gotta write "tail stream" as complicatedly as possible
-			$i_{\Doublestep} \leftarrow \mathsf{FindUnflippedSection}(\{ s_i | s_i \in \Stream \wedge i \ne 0 \})$
-			%\mathsf{tail}(\Stream))$
-		}
-		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i < i_{\Doublestep} \}$)
-		{\sf HeuristicallyDoublestep}($\{ s_i | s_i \in \Stream \wedge i \ge i_{\Doublestep} \}$)
-	}{
-		{\sf CommitStream}(\Stream)
-	}
-	\caption{{\sf HeuristicallyDoublestep}(\Stream)}
-	\label{alg:heuristic}
-\end{algorithm}
-
-To handle such cases,
-I extended the algorithm with a heuristic to identify when a stream becomes ``too crossed-over'' for ``too long'',
-and to force a doublestep by splitting the stream to flip it back.
-Algorithm~\ref{alg:heuristic} shows the implementation.
-I will not summarize how it works due to space limitations, but the description should be intuitive enough.
-The heuristic evaluated Figure~\ref{fig:force-doublestep}'s charts as having 0 crossovers each and 1 and 21 doublesteps, respectively.
-In my analysis next section, I will use $\nflip = 9$
-(determined by inspection of a few favourite charts, which should be scientifically rigorous enough for anyone).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experimental Results}
-
-% TODO: Put a * on charts where you're going to put a pict of mario man
-% TODO: Put a dagger on footswitch charts where the footing is a clever reinterpretation, rather than intended
-
-% TODO: Make a special mention of sandstorm and put a pict of mario man
-% TODO: put the above in an honourable mention subsection, and also mention the slow train and sidefoots
-
-% Stuff to talk about.
-% Dataset size now >10k
-% I reran turniness on fancy footwork and it actually got turnier (double check dis), b/c of all the candles
-
-% By author/by pack: I excluded 1s-3s, admittedly biased against konami, on the pretense of excluding nonsense false positives, but even after that, konami is still the crossover king
-% Step authors are limited to 10 or more charts
-
-% Miscellaneous charts I want to exhibit:
-% Conflict forced doublesteps
-% Heart shooter 'sidefoots'
-% Boom Bombay - triple xover jack footed as sideswitch
-% Toxic - jack+doublestep footed as sideswitch + xover stream instead
-% H'ble mention to sandstorm for jacks
-
-\begin{figure}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#XO \\
-		\hline
-		\multicolumn{4}{c}{\em no 8s- with $\ge$100 XOs} \\
-		 9 & MAX Forever              & CuoReNeRo & 107 \\
-		10 & J-PARA SUPER MEGAMIX     & CuoReNeRo & 141 \\
-		11 & Somebody I Used To Know    & Best of Gazebo    & 114 \\
-		12 & Credens Justitiam        & Stuff B Likes         & 136 \\
-		13 & Banshee Strikes          & VocaJawnz             & 152 \\
-		14 & Slow Down                & Sexuality Violation 2 & 160 \\
-		15 & MAX Forever              & CuoReNeRo & 101 \\
-		\multicolumn{4}{c}{\em no 16s-20s with $\ge$100 XOs} \\
-		21 & Teenage Dream            & Sexuality Violation 2 & 280 \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
-		%\hline
-		%\multicolumn{5}{c}{\em no 8s- with $\ge$100 XOs} \\
-		% 9 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 107 & \em 8.7 \\
-		%10 & J-PARA SUPER MEGAMIX     & CuoReNeRo MeGaPacK    & \bf 141 & \em 3.9 \\
-		%11 & Somebody That I Used To Know & Best of Gazebo    & \bf 114 & \em 11.9 \\
-		%12 & Credens Justitiam        & Stuff B Likes         & \bf 136 & \em 12.4 \\
-		%13 & Banshee Strikes          & VocaJawnz             & \bf 152 & \em 10.2 \\
-		%14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		%15 & MAX Forever              & CuoReNeRo MeGaPacK    & \bf 101 & \em 4.2 \\
-		%\multicolumn{5}{c}{\em no 16s-20s with $\ge$100 XOs} \\
-		%21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts with the most total crossovers (XOs).}
-\end{figure}
-\begin{figure}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf XO\% \\
-		\hline
-		 4 & DAM DIRIRAM              & DDR 3rdMIX            & 27.3 \\
-		 5 & STRICTLY BUSINESS        & DDR 1st               & 21.4 \\
-		 6 & STRICTLY BUSINESS        & DDR 1st               & 20.9 \\
-		 7 & MOBO$\star$MOGA          & DDR EXTREME           & 17.3 \\
-		 8 & PARANOiA                 & DDR 1st               & 19.6 \\
-		 9 & Dazzlin Darlin           & r21twins              & 22.0 \\
-		10 & Enchanted Journey        & ITG Rebirth           & 19.3 \\
-		11 & Lune Noir                & r21freak Friendship   & 13.4 \\
-		12 & W'peg is Fucking Over    & best of r21freak ii   & 14.3 \\
-		13 & The Sampling Paradise    & The Paradise Sampler  & 15.6 \\
-		14 & Slow Down                & Sexuality Violation 2 & 13.3 \\
-		\multicolumn{4}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
-		21 & Teenage Dream            & Sexuality Violation 2 & 11.8 \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#XO & \bf XO\% \\
-		%\hline
-		% 4 & DAM DIRIRAM              & DDR 3rdMIX            & \em  30 & \bf 27.3 \\
-		% 5 & STRICTLY BUSINESS        & DDR 1st               & \em  33 & \bf 21.4 \\
-		% 6 & STRICTLY BUSINESS        & DDR 1st               & \em  38 & \bf 20.9 \\
-		% 7 & MOBO$\star$MOGA          & DDR EXTREME           & \em  37 & \bf 17.3 \\
-		% 8 & PARANOiA                 & DDR 1st               & \em  58 & \bf 19.6 \\
-		% 9 & Dazzlin Darlin           & r21twins              & \em  76 & \bf 22.0 \\
-		%10 & Enchanted Journey        & ITG Rebirth           & \em  78 & \bf 19.3 \\
-		%11 & Lune Noir                & r21freak Friendship   & \em  58 & \bf 13.4 \\
-		%12 & Winnipeg is Fucking Over & best of r21freak ii   & \em 122 & \bf 14.3 \\
-		%13 & The Sampling Paradise    & The Paradise Sampler  & \em 139 & \bf 15.6 \\
-		%14 & Slow Down                & Sexuality Violation 2 & \bf 160 & \bf 13.3 \\
-		%\multicolumn{5}{c}{\em no 15s-20s with $\ge$10\% XOs} \\
-		%21 & Teenage Dream            & Sexuality Violation 2 & \bf 280 & \bf 11.8 \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts with the highest percentage of XOs among total steps.}
-\end{figure}
-\begin{figure}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#FS \\
-		\hline
-		\multicolumn{4}{c}{\em no 5s- with $\ge$50 FSs} \\
-		 6 & Sweat Shop       & ITG Rebirth 2 Beta    &  56 \\
-		\multicolumn{4}{c}{\em no 7s-9s with $\ge$50 FSs} \\
-		10 & Eat 'Em Up!      & Mute Sims 5           &  54 \\
-		11 & Nemeton          & The Legend of Zim 4   &  97 \\
-		12 & Nemeton          & Subluminal            & 140 \\
-		13 & Love Is Eternity & Subluminal            & 140 \\
-		14 & Switch           & Getty                 & 347 \\
-		15 & Danse Macabre    & Aoreo's Ariginals     & 201 \\
-		16 & Weird Science    & Stamina Showcase      &  61 \\
-		\multicolumn{4}{c}{\em no 17s+ with $\ge$50 FSs} \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
-		%\hline
-		%\multicolumn{5}{c}{\em no 5s- with $\ge$50 FSs} \\
-		% 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
-		%\multicolumn{5}{c}{\em no 7s-9s with $\ge$50 FSs} \\
-		%10 & Eat 'Em Up!      & Mute Sims 5           & \bf  54 & \em  9.0 \\
-		%11 & Nemeton          & The Legend of Zim 4   & \bf  97 & \em 10.3 \\
-		%12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
-		%13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
-		%14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
-		%15 & Danse Macabre    & Aoreo's Ariginals     & \bf 201 & \em 12.2 \\
-		%16 & Weird Science    & Stamina Showcase      & \bf  61 & \em  1.8 \\
-		%\multicolumn{5}{c}{\em no 17s+ with $\ge$50 FSs} \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts with the most total footswitches (FSs).}
-\end{figure}
-\begin{figure}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf FS\% \\
-		\hline
-		 3 & Sweat Shop       & ITG Rebirth 2 Beta    & 15.3 \\
-		 4 & MAKE IT BETTER   & DDR 2ndMIX            & 20.0 \\
-		\multicolumn{4}{c}{\em no 5s with $\ge$10\% FS} \\
-		 6 & Sweat Shop       & ITG Rebirth 2 Beta    & 18.2 \\
-		 7 & 5.1.1            & DDR MAX               & 12.9 \\
-		\multicolumn{4}{c}{\em no 8s with $\ge$10\% FS} \\
-		 9 & PARANOiA KCET    & DDR 2ndMIX            & 10.2 \\
-		10 & Delhi Ill        & Mute Sims 6           & 11.6 \\
-		11 & Sweat Shop       & ITG Rebirth 2 Beta    & 11.8 \\
-		12 & Nemeton          & Subluminal            & 14.8 \\
-		13 & Love Is Eternity & Subluminal            & 16.8 \\
-		14 & Switch           & Getty                 & 15.7 \\
-		15 & Flames of the Sky & fort rapids vii      & 16.5 \\
-		\multicolumn{4}{c}{\em no 16s+ with $\ge$10\% FS} \\
-		%\bf Ft. & \bf Name & \bf Pack & \bf \#FS & \bf FS\% \\
-		%\hline
-		% 3 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  26 & \bf 15.3 \\
-		% 4 & MAKE IT BETTER   & DDR 2ndMIX            & \em  26 & \bf 20.0 \\
-		%\multicolumn{5}{c}{\em no 5s with $\ge$10\% FS} \\
-		% 6 & Sweat Shop       & ITG Rebirth 2 Beta    & \bf  56 & \bf 18.2 \\
-		% 7 & 5.1.1            & DDR MAX               & \em  13 & \bf 12.9 \\
-		%\multicolumn{5}{c}{\em no 8s with $\ge$10\% FS} \\
-		% 9 & PARANOiA KCET    & DDR 2ndMIX            & \em  33 & \bf 10.2 \\
-		%10 & Delhi Ill        & Mute Sims 6           & \em  40 & \bf 11.6 \\
-		%11 & Sweat Shop       & ITG Rebirth 2 Beta    & \em  75 & \bf 11.8 \\
-		%12 & Nemeton          & Subluminal            & \bf 140 & \bf 14.8 \\
-		%13 & Love Is Eternity & Subluminal            & \bf 140 & \bf 16.8 \\
-		%14 & Switch           & Getty                 & \bf 347 & \bf 15.7 \\
-		%15 & Flames of the Sky & fort rapids vii      & \em 172 & \bf 16.5 \\
-		%\multicolumn{5}{c}{\em no 16s+ with $\ge$10\% FS} \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts with the highest percentage of FSs among total steps.}
-\end{figure}
-\begin{figure}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|c}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#XF \\
-		\hline
-		\multicolumn{4}{c}{\em no 8s- with $\ge$12 XFs} \\
-		 9 & Dr. Boom-Bombay   & fort rapids vii            & 18 \\
-		10 & Toxic             & Sexuality Violation 2      & 12 \\
-		11 & Heart Shooter     & VocaJawnz                  & 44 \\
-		12 & Beautiful Synergy & Keyboard Collaboration III & 14 \\
-		13 & Toxic             & Sexuality Violation 2      & 12 \\
-		14 & Fancy Footwork    & Cirque du Zeppelin         & 40 \\
-		% 9 & Dr. Boom-Bombay   & fort rapids vii            & \bf 18 & 4.0 \\
-		%10 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.5 \\
-		%11 & Heart Shooter     & VocaJawnz                  & \bf 44 & 5.1 \\
-		%12 & Beautiful Synergy & Keyboard Collaboration III & \bf 14 & 3.2 \\
-		%13 & Toxic             & Sexuality Violation 2      & \bf 12 & 1.2 \\
-		%14 & Fancy Footwork    & Cirque du Zeppelin         & \bf 40 & 4.2 \\
-		\multicolumn{4}{c}{\em no 15s+ with $\ge$12 XFs} \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts with the most crossover footswitches (XFs). (Here I chose 12 as the cut-off because below that are a bunch of false-positive nonsense charts from DDR. Sheesh, Konami.)}
 \end{figure}
@@
 
-\section{Never Work}
+\section{Analyzing Bracket-Jumpiness}
 
-Let's be honest: this isn't gonna be a trilogy.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-With that out of the way, here are some things that would be cool to implement in a fantasy universe with infinite free time.
+\section{Evaluation}
 
-% TODO
-% forced doublesteps via holds
-% - bracketjump footswitches - matador from valex 8
-% understanding mine cues for doublesteps
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+
+\section{Future Work}
+
+It would be cool to integrate this (and the preceding two algorithms \cite{turniness,crossoveriness})
+into Simply Love \cite{simplylove}, or whatever the prevailing stepmania theme may be in the future.
+Currently, as shown in \Cref{fig:songpreview}, % TODO
+the song preview screen wastes considerable space displaying the count of more irrelevant chart aspects
+such as mines (now used primarily for signaling the presence of footswitches)
+and hands (now mostly stepped with the feet by bracketing anyways).
+I bet the community might actually pay attention to my work -- the holy grail of research, honestly --
+if the premier dance game theme could tell precisely how technical a chart was in such modern terms.
 
 \section{Conclusion}
-
-Please accept my paper.
-I worked hard on it.
 

### 2019-03-12 itg/code/ITG.hs
refactor ITG.hs to accommodate diff types of jumps, in prep for brackets. no algorithmic changes yet

@@ -2,16 +2,28 @@ import Data.List
 import Data.Char
 import Data.Maybe
 
-data Step = L | D | U | R | Jump deriving Eq
+data Step = L | D | U | R | J Jump deriving (Ord, Eq)
 
+data Jump = LD | LU | DR | UR | Other deriving (Ord, Eq)
+
+data Foot = LeftFoot | RightFoot deriving (Ord, Eq)
+
+alternate LeftFoot = RightFoot
+alternate RightFoot = LeftFoot
+
+-- TODO add a field to track what the last arrow the left foot and right foot were on
+-- if you commitstream flipped, you need to flip those around too
 data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int,
                          jacks :: Int, doubles :: Int, xoverfs :: Int,
+                         brackets :: Int, stepsAndJumps :: Int,
                          lastStep :: Maybe Step, lastJack :: Maybe Step,
-                         lastFlip :: Bool, lastFoot :: Bool, stepsLR :: [Bool] }
+                         lastFlip :: Bool, lastFoot :: Foot, stepsLR :: [Bool] }
 
 commitStream :: AnalysisState -> AnalysisState
 commitStream s = maybe s0 splitStream splitIndex
-    where ns = length $ stepsLR s
+    where -- number of steps in the chunk of stream
+          ns = length $ stepsLR s
+          -- number of (seemingly) crossed-over steps
           nx = length $ filter not $ stepsLR s
           -- if more than half the L/R steps in this stream were crossed over,
           -- then we got the footing backwards and need to flip the stream.
@@ -39,11 +51,15 @@ commitStream s = maybe s0 splitStream splitIndex
                    lastFlip = f, stepsLR = [] }
 
 analyzeStep :: AnalysisState -> Step -> AnalysisState
-analyzeStep s step
+analyzeStep s (J _) = -- TODO change to Other only and handle bracketables below
     -- a jump resets the footing, so the next step can be stepped with either
     -- foot. commit the stream so far to treat it separately from what follows.
     -- bracket-jumps are, of course, future work.
-    | step == Jump = (commitStream s) { lastStep = Nothing, lastJack = Nothing }
+    -- NB. old version! still holds for LR and DU jumps, as well as 3+s
+    (commitStream s) { lastStep = Nothing, lastJack = Nothing,
+                       stepsAndJumps = stepsAndJumps s + 1 }
+-- analyzeStep s (J jump) = undefined -- TODO
+analyzeStep s step
[…]

### 2019-03-12 itg/code/ITG.hs
more refactor - make Arrow its own datatype

@@ -2,9 +2,11 @@ import Data.List
 import Data.Char
 import Data.Maybe
 
-data Step = L | D | U | R | J Jump deriving (Ord, Eq)
+data Arrow = L| D | U | R deriving (Ord, Eq)
 
 data Jump = LD | LU | DR | UR | Other deriving (Ord, Eq)
+
+data Step = A Arrow | J Jump deriving (Ord, Eq)
 
 data Foot = LeftFoot | RightFoot deriving (Ord, Eq)
 
@@ -16,7 +18,7 @@ alternate RightFoot = LeftFoot
 data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int,
                          jacks :: Int, doubles :: Int, xoverfs :: Int,
                          brackets :: Int, stepsAndJumps :: Int,
-                         lastStep :: Maybe Step, lastJack :: Maybe Step,
+                         lastStep :: Maybe Arrow, lastJack :: Maybe Arrow,
                          lastFlip :: Bool, lastFoot :: Foot, stepsLR :: [Bool] }
 
 commitStream :: AnalysisState -> AnalysisState
@@ -59,12 +61,12 @@ analyzeStep s (J _) = -- TODO change to Other only and handle bracketables below
     (commitStream s) { lastStep = Nothing, lastJack = Nothing,
                        stepsAndJumps = stepsAndJumps s + 1 }
 -- analyzeStep s (J jump) = undefined -- TODO
-analyzeStep s step
+analyzeStep s (A arrow)
     -- two steps on the same arrow might be a jack, or might be a footswitch.
     -- to figure out which, commit the stream so far, and begin a new stream
     -- whose footing will retroactively determine how to foot this step.
     -- also, unlike jumps, this step gets counted as part of the next stream.
-    | lastStep s == Just step = stream (commitStream s) { lastJack = Just step }
+    | lastStep s == Just arrow = stream (commitStream s) { lastJack = Just arrow }
     -- a normal streamy step.
     | otherwise = stream s
     where foot = alternate $ lastFoot s
@@ -72,8 +74,8 @@ analyzeStep s step
           addStep ft L steps = steps ++ [ft == LeftFoot]
           addStep ft R steps = steps ++ [ft == RightFoot]
           addStep ft _ steps = steps -- U/D don't help to determine L/R footing.
-          stream s = s { steps = steps s + 1, lastStep = Just step, lastFoot = foot,
-                         stepsLR = addStep foot step $ stepsLR s,
[…]

### 2019-03-13 itg/code/ITG.hs
first pass at implementing brackets... it passes a bare majority of the tests at least. kinda sucks -- not only not feature complete but also buggy, but it's bedtime

@@ -1,6 +1,7 @@
 import Data.List
 import Data.Char
 import Data.Maybe
+import Data.Tuple
 
 data Arrow = L| D | U | R deriving (Ord, Eq)
 
@@ -19,7 +20,12 @@ data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int,
                          jacks :: Int, doubles :: Int, xoverfs :: Int,
                          brackets :: Int, stepsAndJumps :: Int,
                          lastStep :: Maybe Arrow, lastJack :: Maybe Arrow,
-                         lastFlip :: Bool, lastFoot :: Foot, stepsLR :: [Bool] }
+                         lastFlip :: Bool, lastFoot :: Foot, stepsLR :: [Bool],
+                         -- used for bracket jumps e.g. if left foot's last
+                         -- step was on U, right foot cannot bracket UR
+                         lastArrowLR :: ([Arrow], [Arrow]), -- (left, right)
+                         lastTrueLastArrowLR :: ([Arrow], [Arrow]) -- ugh...
+                       }
 
 commitStream :: AnalysisState -> AnalysisState
 commitStream s = maybe s0 splitStream splitIndex
@@ -46,21 +52,89 @@ commitStream s = maybe s0 splitStream splitIndex
           -- normal case; consider the whole stream at once
           jack = lastJack s /= Nothing
           jackLR = lastJack s == Just L || lastJack s == Just R
+          -- this nonsense merges the lastArrow lists with a known good state
+          -- snapshotted from the previous stream commit, because a 1-step long
+          -- stream e.g. may end up with an empty list on the other foot, and
+          -- you want neither to clobber the old value with [], or clobber it by
+          -- writing willy nilly while you're still not sure if you'll flip.
+          -- analyze doesn't need to worry about this; the correct thing will
+          -- show up in lastArrowLR each time regardless.
+          unifySideArrows []     old = old
+          unifySideArrows recent old = recent
+          unifyArrows (ls,rs) = (unifySideArrows ls $ fst $ lastTrueLastArrowLR s,
+                                 unifySideArrows rs $ snd $ lastTrueLastArrowLR s)
+          -- update the true lastFoot and lastArrows (if nonzero stream)
+          -- this way the value always shows our true position on the pad after
+          -- the stream ended, to see if bracket jumps are possible
[…]

### 2019-03-14 itg/code/run-tests.sh
allow run-tests.sh argument

@@ -1,7 +1,18 @@
 #!/bin/bash
 
 rv=0
-for t in test*.sm; do
+
+if [ -z "$1" ]; then
+	TESTS=`ls test*.sm`
+else
+	TESTS=`ls *$1* 2>/dev/null | grep '\.sm$'`
+	if [ -z "$TESTS" ]; then
+		echo -e "\033[01;33mno such test matching *$1*(.sm)\033[00m"
+		exit 1
+	fi
+fi
+
+for t in $TESTS; do
 	ans=`head -n 1 "$t"`
 	if ! echo "$ans" | grep ",.*,.*," >/dev/null; then
 		echo -e "\033[01;33m$t not well formed -- $ans not an answer.\033[00m"

### 2019-03-14 itg/code/ITG.hs
make brackets work, afaik!!

@@ -2,14 +2,15 @@ import Data.List
 import Data.Char
 import Data.Maybe
 import Data.Tuple
+import Debug.Trace
 
-data Arrow = L| D | U | R deriving (Ord, Eq)
+data Arrow = L| D | U | R deriving (Ord, Eq, Show)
 
-data Jump = LD | LU | DR | UR | Other deriving (Ord, Eq)
+data Jump = LD | LU | DR | UR | LR | DU | Other deriving (Ord, Eq, Show)
 
-data Step = A Arrow | J Jump deriving (Ord, Eq)
+data Step = A Arrow | J Jump deriving (Ord, Eq, Show)
 
-data Foot = LeftFoot | RightFoot deriving (Ord, Eq)
+data Foot = LeftFoot | RightFoot deriving (Ord, Eq, Show)
 
 alternate LeftFoot = RightFoot
 alternate RightFoot = LeftFoot
@@ -20,15 +21,18 @@ data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int,
                          jacks :: Int, doubles :: Int, xoverfs :: Int,
                          brackets :: Int, stepsAndJumps :: Int,
                          lastStep :: Maybe Arrow, lastJack :: Maybe Arrow,
-                         lastFlip :: Bool, lastFoot :: Foot, stepsLR :: [Bool],
+                         lastFlip :: Bool, lastFoot :: Maybe Foot, stepsLR :: [Bool],
                          -- used for bracket jumps e.g. if left foot's last
                          -- step was on U, right foot cannot bracket UR
                          lastArrowLR :: ([Arrow], [Arrow]), -- (left, right)
                          lastTrueLastArrowLR :: ([Arrow], [Arrow]) -- ugh...
-                       }
+                       } deriving Show
 
-commitStream :: AnalysisState -> AnalysisState
-commitStream s = maybe s0 splitStream splitIndex
+commitStream s = s' -- traceShow ("commit", s, "=>", s') s'
+    where s' = commitStream' s
+
+commitStream' :: AnalysisState -> AnalysisState
+commitStream' s = maybe s0 splitStream splitIndex
     where -- number of steps in the chunk of stream
           ns = length $ stepsLR s
           -- number of (seemingly) crossed-over steps
@@ -66,15 +70,11 @@ commitStream s = maybe s0 splitStream splitIndex
           -- update the true lastFoot and lastArrows (if nonzero stream)
           -- this way the value always shows our true position on the pad after
           -- the stream ended, to see if bracket jumps are possible
[…]

### 2019-03-15 itg/code/ITG.hs
add a bunch more bracker jump tests and fix all the bugs in the alg

@@ -21,27 +21,34 @@ data AnalysisState = S { steps :: Int, xovers :: Int, switches :: Int,
                          jacks :: Int, doubles :: Int, xoverfs :: Int,
                          brackets :: Int, stepsAndJumps :: Int,
                          lastStep :: Maybe Arrow, lastJack :: Maybe Arrow,
-                         lastFlip :: Bool, lastFoot :: Maybe Foot, stepsLR :: [Bool],
+                         lastFlip :: Bool, lastFoot :: Maybe Foot,
+                         justBracketed :: Bool, stepsLR :: [Bool],
                          -- used for bracket jumps e.g. if left foot's last
                          -- step was on U, right foot cannot bracket UR
                          lastArrowLR :: ([Arrow], [Arrow]), -- (left, right)
                          lastTrueLastArrowLR :: ([Arrow], [Arrow]) -- ugh...
                        } deriving Show
 
-commitStream s = s' -- traceShow ("commit", s, "=>", s') s'
-    where s' = commitStream' s
+commitStream tiebreakFoot s = s' -- traceShow ("commit", s, "=>", s') s'
+    where s' = commitStream' tiebreakFoot s
 
-commitStream' :: AnalysisState -> AnalysisState
-commitStream' s = maybe s0 splitStream splitIndex
+commitStream' :: Maybe Foot -> AnalysisState -> AnalysisState
+commitStream' tiebreakFoot s = maybe s0 splitStream splitIndex
     where -- number of steps in the chunk of stream
           ns = length $ stepsLR s
           -- number of (seemingly) crossed-over steps
           nx = length $ filter not $ stepsLR s
           -- if more than half the L/R steps in this stream were crossed over,
           -- then we got the footing backwards and need to flip the stream.
-          -- as a tiebreaker, flip if the chart is already more jacky than
+          -- as tiebreaker #1, don't flip if immediately preceded by a bracket
+          flipTieBreak _ | justBracketed s = False
+          -- as tiebreaker #2, if the stream is followed by a bracketable jump,
+          -- choose whichever flip-or-no-flip lets us bracket that jump.
+          flipTieBreak (Just bracketFoot) = lastFoot s == Just bracketFoot
+          -- as tiebreaker #3, flip if the chart is already more jacky than
           -- footswitchy, i.e., if past streams flipped more often than not.
[…]

### 2019-03-15 itg/3/paper.tex
start writing overview

@@
 
-\conferenceinfo{SIGBOVIK '17}{Pittsburgh, PA, USA}
-\copyrightyear{2017}
+\conferenceinfo{SIGBOVIK '19}{Pittsburgh, PA, USA}
+\copyrightyear{2019}
 \copyrightdata{}
@@
 
-\authorinfo{Ben Blum}{}{bblum@cs.cmu.edu}
+\authorinfo{Ben Blum}{}{bblum@alumni.cmu.edu}
 
@@
 \keywords
-brackets, jumps, bracket jumps
+bracket, groove, in, jumps, the
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\hspace{-2em}\includegraphics[angle=270,width=0.54\textwidth]{../2/paper.pdf}
@@
 	\label{fig:you-stutid-fuckass}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Overview}
+
+% TODO: go through this paper and make every *other* instance of "bracket-jump" hyphenated
+
+What, then, is a bracket jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
+Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
+(see \cite{turniness}),
+while a novice player might think they must step with both feet at once, one for each arrow,
+experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
+to hit both arrows by stepping across the metal bracket enclosing the corners of each panel
+(thereby triggering one sensor with the heel and one with the little bitty toesies), hence ``bracket jump''.
+In case my verbal explanation is not up to snuff, I have also included in \Cref{fig:how-2-bracket}
+a high-quality graphics render of a player's typical foot positions during a down+right bracket jump (henceforth ``DR'', et cetera).
+
+\begin{figure}[h]
+	\begin{center}
+		\includegraphics[width=0.48\textwidth]{how-2-bracket.jpg}
+	\end{center}
+	\caption{Down+right bracket jump real-world example.}
+	\label{fig:how-2-bracket}
 \end{figure}
@@
 \section{Discussion}
+
+% TODO tings to talk about
+
+% authors seem to know it's easy to overdo -- see, "Stupid" and "I'm sorry" as chart authors in some of the top charts
+
+% many old charts written without regard to jump patterning can be bracker anyway, see OWA (which just makes the list bc it's long)
+% - have a figure of OWA in the strema that shows the *non* bracketable stream jumps
+% - make a test case that captures these and shows indeed they are not brackets
+
+% temporal analysis of pack age vs bracker density
+
+% other charts of note
+% - chicago timing / electrical paradise: bracketiest chart where *ALL* jumps are brackets (BJJ%=100) -- quite tasteful actually!
+% - encore is 2nd place lmao
+
+% conclusion: BJJ% is sa better measure of chart funness than BJS%?
 

### 2019-03-16 itg/3/paper.tex
write a ubnch more of the paper.. no design/eval yet

@@
 \usepackage{multirow}
+\usepackage{xspace}
 \usepackage[nameinlink,nosort]{cleveref}
@@
 \setCJKmainfont{HeiseiMinStd-W5}[Path = ./]
+\setmonofont{Inconsolata-Regular}[Path = ./]
 % This Deep Tex Voodoo is from
@@
 experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
-to hit both arrows by stepping across the metal bracket enclosing the corners of each panel
-(thereby triggering one sensor with the heel and one with the little bitty toesies), hence ``bracket jump''.
-In case my verbal explanation is not up to snuff, I have also included in \Cref{fig:how-2-bracket}
+to hit both arrows by
+triggering one arrow's sensor with the heel and the other with the little bitty toesies.
+Pads are typically constructed with a small triangular metal bracket at the corner of each arrow panel,
+as shown in \Cref{fig:bracker-detail},
+which the bridge of the foot must cross to achieve this, hence the name ``bracket jump''.
+In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
 a high-quality graphics render of a player's typical foot positions during a down+right bracket jump (henceforth ``DR'', et cetera).
+
+The reader, or stepper,
+may notice the extreme angle of footing depicted in the latter figure,
+which is necessary to reliably trigger both pad sensors.
+Accordingly, this maneuver is comfortable (and hence preferable to a normal jump)
+only if the player is already facing in roughly the same direction \cite{turniness}.
+%after stepping any preceding patterns.
+The next section will attempt to codify (ahem) when
+preceding patterns encourage the player to bracket rather than jump,
+and thereby identify how ``bracket jumpy'' each chart is.
+
+\begin{figure}[t]
+	\begin{center}
+		\includegraphics[width=0.48\textwidth]{jims-pix.jpg}
+	\end{center}
+	\caption{Detail of metal corner brackets, this paper's namesake. Photo credit {\tt JIM}.}
+	\label{fig:bracker-detail}
+\end{figure}
 
@@
 \section{Analyzing Bracket-Jumpiness}
+\label{sec:analyzing}
+
+I extended the crossoveriness et cetera algorithm from \cite{crossoveriness} to reason about jumps,
+which previously it treated all identically,
+ignoring the arrows involved and allowing the player to reset her footing as desired.
+Now, it considers LD, LU, DR, and UR jumps as potentially bracketable,
+allowing the player to continue a stream of alternating feet uninterrupted through.
+
+\textit{Confession.}
+When I was first brainstorming this project,
+I had some grand visions of unifying the turniness algorithm \cite{turniness}---which
+accounts for U/D steps to figure out how far the player must turn each step,
+but has no idea which way the player is actually facing at any given time---with
+the crossoveriness one
+(cited just above; cmon how much do you want me to repeat these, gimme a break)---which
+totally ignores U/D steps and just figures out which arrows the left and right feet must each step---in
+some theoretically beautiful way to produce the unquestionably perfect footing sequence for each jump.
+%
+However, as I was considering how to incorporate features from last time's algorithm,
+potentially allowing crossover brackets (\Cref{sec:xover-brackers})
+and footswitch brackets (\Cref{sec:fs-brackers}),
+I realized that ultimately there would be no restrictions on what jumps were bracketable;
+the algorithm would oops simply twist and turn as much as necessary to bracket everything,
+and this paper would become ``Which ITG Stepcharts Have the Most Jumps?'', and like who wants to read that.
+
+One may think to simply try either bracketing each jump or not
+and seeing what combination gives the minimum turniness result,
+but since whether to bracket each jump or not affects subsequent steps' footing,
+each choice cannot in general be solved independently,
+and I hope you can see where this is going.
+Now the last time I tried to solve an exponentially-sized problem, it took me 7 years
+(and I still ended up with a pile of heuristics after all anyway)
+\cite{landslide-thesis},
+so considering I started this project a week before the deadline,
+I instead to just code up a bunch of ad-hoc rules to handle all the different bracket jump patterns
+that occurred to me to write test cases for.
+
+As before, the code is available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
+To give an idea of how much its elegance has been despoiled % hi william
+since the last version:
+98 lines of Haskell (which computed crossovers, footswitches, jacks, doubleswitches, \textit{and} crossover footswitches)
+has grown to 248 lines (just to handle this one new feature),
+and the once-simple datatype definition of
+
+\newcommand\hilight[2]{\color{#1}#2\color{black}\xspace}
+\definecolor{pink}{RGB}{128,0,192}
+\definecolor{orange}{RGB}{192,96,0}
+\definecolor{olivegreen}{RGB}{0,127,32}
+\definecolor{brickred}{RGB}{192,0,0}
+\definecolor{commentblue}{RGB}{0,128,192}
+
+\begin{center}
+	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step} =
+	\hilight{brickred}{L} |
+	\hilight{brickred}{D} |
+	\hilight{brickred}{U} |
+	\hilight{brickred}{R} |
+	\hilight{brickred}{Jump}}
+\end{center}
+
+has become:
+
+\begin{center}
+	\begin{tabular}{l}
+	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Arrow} =
+	\hilight{brickred}{L} |
+	\hilight{brickred}{D} |
+	\hilight{brickred}{U} |
+	\hilight{brickred}{R}} \\
+	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Jump} =
+	\hilight{brickred}{LD} |
+	\hilight{brickred}{LU} |
+	\hilight{brickred}{DR} |
+	\hilight{brickred}{UR} |
+	\hilight{brickred}{LR} |
+	\hilight{brickred}{DU} |
+	\hilight{brickred}{Other}} \\
+	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step} =
+	\hilight{brickred}{A} \hilight{olivegreen}{Arrow} |
+	\hilight{brickred}{J} \hilight{olivegreen}{Jump}} \\
+	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Foot} =
+	\hilight{brickred}{LeftFoot} |
+	\hilight{brickred}{RightFoot}} \\
+	\end{tabular}
+\end{center}
+
+In lieu of an algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how I handle them.
+
+\subsection{Basic Examples}
+
+Figure \Cref{fig:bracket-example}(a) shows a simple example where every jump (blue)
+can be bracketed by the right foot,
+alternating with
 
@@
 
-\section{Future Work}
+\section{Never Work}
+
+The only thing that impresses the research community more than overdelivering on your future work promises
+is to overdeliver twice; hence, I shamelessly reuse the joke of this section name from last time.
+
+\subsection{Crossover Brackets}
+\label{sec:xover-brackers}
+
+\subsection{Footswitch Brackets}
+\label{sec:fs-brackers}
+
+I put this section here only just so I could have something to forward-reference back in \Cref{sec:analyzing}.
+I always wanted to do that, you know?
+% TODO: put a pict anyway? if you find one
+
+\subsection{UI}
 
 It would be cool to integrate this (and the preceding two algorithms \cite{turniness,crossoveriness})
-into Simply Love \cite{simplylove}, or whatever the prevailing stepmania theme may be in the future.
-Currently, as shown in \Cref{fig:songpreview}, % TODO
+into
+%Simply Love \cite{simplylove}, or whatever
+the community's prevailing stepmania theme.
+%may be in the future.
+Currently, as shown in \Cref{fig:songpreview}(a),
 the song preview screen wastes considerable space displaying the count of more irrelevant chart aspects
@@
 and hands (now mostly stepped with the feet by bracketing anyways).
-I bet the community might actually pay attention to my work -- the holy grail of research, honestly --
-if the premier dance game theme could tell precisely how technical a chart was in such modern terms.
+I bet the community might actually pay attention to my work---the holy grail of research, honestly---if
+the premier dance game theme showed precisely how technical a chart was, as in (b).
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+	\includegraphics[width=0.46\textwidth]{song-info.png} \\
+		(a) Current state-of-the-art.
+		\\
+		\\
+	\includegraphics[width=0.46\textwidth]{song-info-better.png} \\
+		(b) Artist's conception.
+	\end{tabular}
+	\caption{Song info preview panel for a recent bracket jump-heavy stepchart \cite{divine}.}
+	\label{fig:songpreview}
+\end{figure}
+
+\subsection{Obligatory Machine Learning Section}
+
+I guess you could skip all this fiddly ``algorithm'' stuff by
+hopping on the pads yourself for a few rounds, playing a variety of technical charts,
+and just training a neural network based on how you stepped the patterns.
+It would probably even learn to fake crossovers more at the ends of songs than at the beginnings,
+where you're more likely to be physically tired.
+But like, do you really want that kind of bias in your dataset?
+
+Remember kids, friends don't let friends use ML for problems that are more fun to solve by hand.
 
 \section{Conclusion}
+
+I never know what to put down here. Much love to the community..!
 

### 2019-03-16 itg/3/paper.tex
more edits

@@
 98 lines of Haskell (which computed crossovers, footswitches, jacks, doubleswitches, \textit{and} crossover footswitches)
-has grown to 248 lines (just to handle this one new feature),
+has grown to 248 lines (just to handle this one. new. feature),
 and the once-simple datatype definition of
@@
 I always wanted to do that, you know?
-% TODO: put a pict anyway? if you find one
+I mean, footswitch brackets are also a real thing, but still. $\bar{~~}$\textbackslash\_(ツ)\_/$\bar{~~}$
+% not-to-to: put a pict anyway? if you find one
+% actually, joke comes off better w/o
 
@@
 Currently, as shown in \Cref{fig:songpreview}(a),
-the song preview screen wastes considerable space displaying the count of more irrelevant chart aspects
+the song preview screen wastes considerable space displaying the count of irrelevant chart aspects
 such as mines (now used primarily for signaling the presence of footswitches)
@@
 I bet the community might actually pay attention to my work---the holy grail of research, honestly---if
-the premier dance game theme showed precisely how technical a chart was, as in (b).
+popular themes used it to show precisely how technical a chart was, as in (b).
 

### 2019-03-16 itg/3/paper.tex
xb

@@
 %after stepping any preceding patterns.
+Note also that the two ``candle jumps'', LR and DU, are not possible to bracket,
+unless your foot size is (physically) beyond the scope of this work.
 The next section will attempt to codify (ahem) when
@@
 and I hope you can see where this is going.
-Now the last time I tried to solve an exponentially-sized problem, it took me 7 years
-(and I still ended up with a pile of heuristics after all anyway)
-\cite{landslide-thesis},
+Now, the last time I tried to solve an exponentially-sized problem, it took me 7 years
+\cite{landslide-thesis}
+(and I still ended up with a pile of heuristics after all anyway),
 so considering I started this project a week before the deadline,
@@
 since the last version:
-98 lines of Haskell (which computed crossovers, footswitches, jacks, doubleswitches, \textit{and} crossover footswitches)
+% TODO: figure out these numbers without counting comments
+98 lines of Haskell (which computed crossovers, footswitches, jacks, doublesteps, \textit{and} crossover footswitches)
 has grown to 248 lines (just to handle this one. new. feature),
@@
 
+\vspace{-1em} % dont know why this is necessary
 \begin{center}
@@
 
+% TODO: be like, we set out to prove that ppl are using bracker jumps more often lately
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+% TODO: make sure to say somewhere, step-cartography
+
 \section{Never Work}
@@
 
+Whereas with normal crossovers (e.g., in the sequence L-U-R, stepping the R with the left foot),
+the player's foot can pretty much point whichever way she finds most comfortable;
+however,
+crossover brackets (e.g., in the sequence L-U-DR...),
+the player's foot must point backwards (...stepping the DR's R with her left heel and D with toes),
+inducing extra turniness.%
+\footnote{The other way, e.g. stepping the D with the left heel and R with toes, while your right foot is fixed on U, is extremely uncomfortable at speed. Trust me on this one.}
+Such jumps could instead just be stepped with both feet as normal,
+possibly inducing doublesteps or jacks,
+and indeed that is how the algorithm presently handles them.
+
+However, ambitious step-cartographers have recently experimented with encouraging the player to bracket jumps while crossed-over.
+Such charts attempt to force, or at least hint, these jumps to be bracketed
+via additional stepchart elements: in \Cref{fig:xb}(a), mines on the left arrow force the player to remove her left foot in preparation for the crossover (and being in the middle of fast stream, further discourages doublestepping),
+and in \Cref{fig:xb}(b), extending one of the arrows as a hold encourages the player to bracket the subsequent jump with the other foot.
+In (b)'s case, the player must also switch feet on each of the non-hold arrows,
+effectively making these crossover-footswitch-brackets,
+which would truly be a wonder to identify programmatically.
+% these can be cheated anyway
+
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{xb-encore.png}
+		\\
+		(a) Encore \cite{encore} %\\
+		%(12, Tech. Showcase 1)
+	\end{tabular}
+	\quad
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{xb-atariwave.png}
+		\\
+		(a) Atariwave \cite{atariwave} %\\
+		%(10, Tech. Showcase 3)
+	\end{tabular}
+	\end{center}
+	\caption{Crossover bracket jumps must be signaled via either (a) mines or (b) holds to remove ambiguity.}
+	\label{fig:xb}
+\end{figure}
+
 \subsection{Footswitch Brackets}
@@
 I always wanted to do that, you know?
-I mean, footswitch brackets are also a real thing, but still. $\bar{~~}$\textbackslash\_(ツ)\_/$\bar{~~}$
+I mean, footswitch brackets are %also
+a real thing, but still. $\bar{~~}$\textbackslash\_(ツ)\_/$\bar{~~}$
 % not-to-to: put a pict anyway? if you find one
@@
 where you're more likely to be physically tired.
-But like, do you really want that kind of bias in your dataset?
+And like, do you really want that kind of bias in your dataset?
 

### 2019-03-16 itg/3/paper.tex
add subtitle; write eval intro blurb; write discussion (epara, owa)

@@
 Titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?''」\\
-Series for All Its Worth in Publication Count After All \\
+Series for All Its Worth in Publication Count After All, or: \\
 %That `Boring' Stuff Was Part of the Title, BTW. \\
@@
 %Oh, That Was Also Part of the Title.
+Hit Me With An Encore
 }
@@
 \begin{center}
-	\includegraphics[width=0.42\textwidth]{using-my-turn.png}
+	\includegraphics[width=0.38\textwidth]{using-my-turn.png}
 \end{center}
+
+\begin{figure}[t]
+	\hspace{-2em}\includegraphics[angle=270,width=0.54\textwidth]{../2/paper.pdf}
+	\caption{(yeah i reused this joke from last time ok deal)}
+	\label{fig:you-stutid-fuckass}
+\end{figure}
 
@@
 extending it to handle these new so-called ``bracket'' jumps.
-
-\begin{figure}[t]
-	\hspace{-2em}\includegraphics[angle=270,width=0.54\textwidth]{../2/paper.pdf}
-	\caption{(yeah i reused this joke from last time ok deal)}
-	\label{fig:you-stutid-fuckass}
-\end{figure}
 
@@
 
+The experimental corpus has grown even more since last time, now comprising 17253 stepcharts from 181 packs.
+Of note, all three Technical Showcase packs
+(collaborative community packs where all authors were encouraged to submit their freshest beats and judged on complexity)
+were released in the last two years,
+which have proved to be important sources of diverse bracket-jump patterns for this research.
+I would also have liked to re-rank the crossoveriest and footswitchiest charts from last time to include these submissions,
+but I frankly don't have time.
+The results spreadsheet of bracket jump counts
+(as well as more crossovers et cetera for packs released since last time)
+is available for your browsing pleasure at
+\url{https://tinyurl.com/bracketiest}.
+
+I pose two evaluation questions, to be answered in the following subsections respectively.
+\begin{enumerate}
+	\item Which songs, packs, and/or step-cartographers are bracket jumpiest?
+	\item Is bracket jumps the new crossovers or something?
+\end{enumerate}
+
+\subsection{Bracket Jumpiness}
+
+% TODO: define BJJ% and BJS%
+
+\subsection{Historical Trends}
+
 % TODO: be like, we set out to prove that ppl are using bracker jumps more often lately
+
+% temporal analysis of pack age vs bracker density
 
@@
 
-% TODO tings to talk about
+Many old charts written without regard to jump patterning allow for many of their jumps to be bracketed anyway.
+I noticed an old stepchart for One-Winged Angel \cite{owa} ranking high in the list of overall total bracket jumps,
+with 182,
+simply because it's long and has a lot of steps,
+but upon playing, the patterns definitely do not feel intentional,
+and it even includes a section of stream with wholly unbracketable jumps interspersed.
+\Cref{fig:discussion}(a) shows this section;
+in this stream, the player would expect to step the red and blue arrows with her right foot and the lime ones with her left,
+but note how the jumps (hint: all in blue)
+occur as a mix of left and right arrows, with even a UD ``candle'' jump making an appearance.
+However, it measures an unremarkable 9 BJS\% and 38 BJJ\%---I just wanted to feature it
+as an example of what non-bracketable jump stream looks like.
 
-% authors seem to know it's easy to overdo -- see, "Stupid" and "I'm sorry" as chart authors in some of the top charts
+Authors seem to realize when their charts are too bracket jumpy for comfort:
+several of the charts appearing in the highest ranks of the spreadsheet had their step author field filled in as ``Stupid'' or ``I'm sorry''.
+In fact, I personally found the charts with 100 BJJ\%, i.e., \textit{all} jumps appearing therein were bracketable,
+to be more tasteful than those with more total BJS\% but a few normal jumps as well.
+I'd hypothesize this is because to achieve 100 BJJ\% requires a certain intentionality,
+resulting in better chart design overall
+(note that such charts necessarily include no LR or DU jumps whatsoever).
+Either that, or the more uncomfortable ones include certain extremely turny/candley/doublesteppy patterns
+around their brackets that causes the algorithm to count them as normal jumps instead.
+Anyway, \Cref{fig:discussion}(b) shows the chart with most bracket jumps (35) among ones with 100 BJJ\%,
+which I tried out for myself and found quite enjoyable \cite{e-paradise}.
+Amusingly, Encore \cite{encore}
+features a Hard chart in addition to its Challenge,
+which former ranks second place to Electrical Paradise among 100 BJJ\%s with 29 brackets.
+Upon inspection, I found this chart identical to the Challenge (shown in \Cref{fig:xb}(a)),
+only the crossover brackets
+(to be discussed later in \Cref{sec:xover-brackers})
+had been replaced by easier, normal ones.
 
-% many old charts written without regard to jump patterning can be bracker anyway, see OWA (which just makes the list bc it's long)
-% - have a figure of OWA in the strema that shows the *non* bracketable stream jumps
-% - make a test case that captures these and shows indeed they are not brackets
-
-% temporal analysis of pack age vs bracker density
-
-% other charts of note
-% - chicago timing / electrical paradise: bracketiest chart where *ALL* jumps are brackets (BJJ%=100) -- quite tasteful actually!
-% - encore is 2nd place lmao
-
-% conclusion: BJJ% is sa better measure of chart funness than BJS%?
-
-% TODO: make sure to say somewhere, step-cartography
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{cc}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-owa.png}
+		\\
+		(a) One-Winged Angel, \\ 14 \cite{owa} %\\
+		%(10, Tech. Showcase 3)
+	\end{tabular}
+		&
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-paradise.png}
+		\\
+		(b) Electrical Paradise, \\ 12 \cite{e-paradise} %\\
+		%(12, Tech. Showcase 1)
+	\end{tabular}
+	\end{tabular}
+	\end{center}
+	%\caption{Why I gotta write a overall caption, just read the prose on this one.}
+	\caption{Example charts with
+	(a) ``retro'' style jump-stream which, largely, cannot be bracketed,
+	and
+	(b) very intentionally many bracket jumps.}
+	\label{fig:discussion}
+\end{figure}
 
@@
 
-However, ambitious step-cartographers have recently experimented with encouraging the player to bracket jumps while crossed-over.
+However, ambitious stepchart authors have recently experimented with encouraging the player to bracket jumps while crossed-over.
 Such charts attempt to force, or at least hint, these jumps to be bracketed
@@
 	\begin{center}
+	\begin{tabular}{cc}
 	\begin{tabular}{c}
@@
 		\\
-		(a) Encore \cite{encore} %\\
+		(a) Encore, 12 \\ \cite{encore} %\\
 		%(12, Tech. Showcase 1)
 	\end{tabular}
-	\quad
+		&
 	\begin{tabular}{c}
@@
 		\\
-		(a) Atariwave \cite{atariwave} %\\
+		(a) Atariwave, 10 \\ \cite{atariwave} %\\
 		%(10, Tech. Showcase 3)
+	\end{tabular}
 	\end{tabular}

### 2019-03-17 itg/3/paper.tex
write a lot

@@
 accounts for U/D steps to figure out how far the player must turn each step,
-but has no idea which way the player is actually facing at any given time---with
+but has no idea which way she is actually facing at any given time---with
 the crossoveriness one
-(cited just above; cmon how much do you want me to repeat these, gimme a break)---which
+(cited just above; cmon how much do you want me to repeat these same two citations, gimme a break)---which
 totally ignores U/D steps and just figures out which arrows the left and right feet must each step---in
@@
 the algorithm would oops simply twist and turn as much as necessary to bracket everything,
-and this paper would become ``Which ITG Stepcharts Have the Most Jumps?'', and like who wants to read that.
+and this paper would become just ``Which ITG Stepcharts Have the Most Jumps?'', and like who wants to read that.
 
@@
 so considering I started this project a week before the deadline,
-I instead to just code up a bunch of ad-hoc rules to handle all the different bracket jump patterns
+I opted instead to just code up a bunch of ad-hoc rules to handle all the different bracket jump patterns
 that occurred to me to write test cases for.
@@
 As before, the code is available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
-To give an idea of how much its elegance has been despoiled % hi william
+To give a sense of how much its elegance has been despoiled % hi william
 since the last version:
@@
 
-has become:
+has become the unwieldy:
 
@@
 
+\subsection{Wackier Examples}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The experimental corpus has grown even more since last time, now comprising 17253 stepcharts from 181 packs.
+The experimental corpus has grown even more since last time, now comprising %17253
+17340 % with ups 4
+stepcharts from 182 packs.
 Of note, all three Technical Showcase packs
-(collaborative community packs where all authors were encouraged to submit their freshest beats and judged on complexity)
+(collaborative packs where the community at large was encouraged to submit their freshest beats and judged on complexity)
 were released in the last two years,
@@
 The results spreadsheet of bracket jump counts
-(as well as more crossovers et cetera for packs released since last time)
+(as well as more crossovers, footswitches, et cetera for charts newer than the last paper)
 is available for your browsing pleasure at
 \url{https://tinyurl.com/bracketiest}.
+
+% command used to generate the sheet
+% used "|" as sed delimiter bc others appear in song names
+% IFS=$'\n'; for i in `find ~/stepmania/Songs/ | grep '\.sm$'`; do cat "$i" | ./ITG 2>&1 | sed "s|^|$i\t|"; done > ../3/results/results.txt
 
@@
 \subsection{Bracket Jumpiness}
+\label{sec:bjness}
 
-% TODO: define BJJ% and BJS%
+I measured the overal bracketiness of each chart in two ways,
+first,
+by comparing the number of bracket jumps against the total number of steps,
+or the \textit{bracket jump step percentage} (BJS\%);
+second,
+by comparing against the total number of jumps only,
+or the \textit{bracket jump jump percentage} (BJJ\%).
+The BJS\% indicates a chart's overall density of steps requiring the player to step across the brackets
+(which depends on how jumpy the chart is to begin with),
+whereas the BJJ\% measures perhaps the author's intentionality in patterning their jumps either as brackets,
+or as random where just some of them happen to be bracketable.
+
+\Cref{tab:bjs} shows the leaderboard for the former metric, and \Cref{tab:bjj} the latter.%
+\footnote{When measuring BJJ\% I excluded charts with fewer than 10 jumps in total,
+which put up a handful of false positives in the 100\% ranks.}
+Note that the aforementioned recent Technical Showcase pack series put up three bracketiest charts in the BJJ\% category;
+meanwhile, the UPS packs, known for their gimmicks and general lack of respect for player comfort,
+secure several top spots on the BJS\% board.
+For comparison's sake, the overall BJJ\% of the entire corpus is 20\% (93k/470k),
+meaning that roughly 1 in 5 randomly-patterned jumps are by chance bracketable.
+The overall BJS\% (a thoroughly less meaningful statistic) is 0.7\%.
+
+% TODO check margins
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|r|r}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJS\% \\
+		\hline
+		10 & SOBA			& Squeaky Beds \&c.	& 366 & 83\% \\
+		11 & Get Off of My Way		& UPS 3			& 115 & 28\% \\
+		12 & Hardware Store		& Keyboard Coll. III	& 207 & 23\% \\
+		13 & Firestorm			& BemaniBeats 3		& 153 & 31\% \\
+		14 & Bounce			& UPS 2			& 126 & 21\% \\
+		15 & Ikaros Dynamite!!!!	& UPS 2			& 238 & 34\% \\
+		16 & Mermaid Island		& Tachyon Alpha		& 346 & 49\% \\
+		17 & Toccata \& Fugue		& CuoReNeRo M.P.	& 255 & 24\% \\
+		% this chart is ALL jumps so it's kinda false positive
+		%420 & Paa			& UPS 4			& 263 & 34\% \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts of each difficulty with the highest bracket jump density among all steps (BJS\%).}
+	\label{tab:bjs}
+\end{table}
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{r|l|l|r|r}
+		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJJ\% \\
+		\hline
+		 9 & Drifting Away		& UPS 4			&  26 &  88\% \\
+		10 & SOBA			& Squeaky Beds \&c.	& 366 &  99\% \\
+		11 & Save Miracles		& ECFA 2019		&  42 &  95\% \\
+		12 & Electrical Paradise	& Chic. Timing Auth.	&  35 & 100\% \\
+		12 & Encore			& Tech. Showcase	&  29 & 100\% \\
+		12 & Nemeton			& Subluminal		&  12 & 100\% \\
+		13 & Decadent Dandy		& Tech. Showcase 3	&  12 & 100\% \\
+		14 & Nageki no Ki		& Valex's M.4-A.A. 8	& 116 &  94\% \\
+		15 & Beach Party		& Tech. Showcase 2	& 162 &  80\% \\
+		16 & Mermaid Island		& Tachyon Alpha		& 346 &  79\% \\
+		17 & Zombie Sunset		& Jummy Jawns 2		& 325 &  88\% \\
+		69 & koopa bling		& UPS 4			&  39 &  98\% \\
+	\end{tabular}
+	\end{center}
+	\caption{Charts of each difficulty with the highest percentage of their total jumps bracketable (BJJ\%).
+	All participants of the 3-way tie for 12-footers are listed.}
+	%filtered by having at least 10 jumps to begin with.}
+	\label{tab:bjj}
+\end{table}
+
+Next up is bracket jumpiness by pack and by author.
+\Cref{tab:pack-bjs,tab:pack-bjj} show the top 10 packs in both bracketiness metrics,
+and \Cref{tab:author-bjs,tab:author-bjj} the top 10 authors
+(filtered by having written at least 10 charts).
+For the packs, I also list the release year of each;
+note how despite the relatively even spread of years in BJS\%,
+2018 dominates the BJJ\% leaderboard,
+suggesting that older charts' bracket jumps arose by chance simply from having a lot of jumps to begin with.
+I'll come back to this point in \Cref{sec:eval-years}.
+In honorable mentionth place is chart author Halogen-, whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
+What a purist!
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|r}
+		\bf Pack name & \bf Year & \bf \#charts & \bf BJS\% \\
+		\hline
+		Keyboard Collaboration III		& $\le$2012	&  17 & 12.4\% \\
+		Squeaky Beds and Leaky Faucets		& 2018	& 125 &  4.9\% \\
+		Keyboard Collaboration I		& $\le$2012	&  12 &  3.6\% \\
+		CuoReNeRo MeGaPacK			& N/A	& 460 &  3.3\% \\
+		Mute Sims X2 WIP			& 2018	&  12 &  3.0\% \\
+		r2112					& 2007	&  60 &  2.7\% \\
+		Technical Showcase 3			& 2018	& 196 &  2.7\% \\
+		FoxyMix 4 - Nuclear Overdrive		& $\le$2012	&  67 &  2.6\% \\
+		Technical Showcase 2			& 2017	&  80 &  2.6\% \\
+		Gensokyo Midnight			& $\le$2012 	&  77 &  2.5\% \\
+	\end{tabular}
+	\end{center}
+	\caption{Song packs densest in bracket jumps.}
+	\label{tab:pack-bjs}
+\end{table}
+
+% sorry about my sigfig discrepancies between this and the by-chart table but like, latex and margins yknow
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r|r}
+		\bf Pack name & \bf Year & \bf \#charts & \bf BJJ\% \\
+		\hline
+		Feelin' Rusty 3			& 2018	&  51 & 50.2\% \\
+		Squeaky Beds and Leaky Faucets	& 2018	& 125 & 46.7\% \\
+		TYLR's Technical Difficulties	& 2018	&  33 & 44.9\% \\
+		Technical Showcase 3		& 2018	& 196 & 41.7\% \\
+		Mute Sims X2 WIP		& 2018	&  12 & 40.9\% \\
+		Jimmy Jawns 2			& 2015	&  55 & 39.7\% \\
+		Technical Showcase 2		& 2017	&  80 & 38.2\% \\
+		Keyboard Collaboration III	& $\le$2012	&  17 & 38.1\% \\
+		Chicago Timing Authority	& 2018	&  33 & 36.3\% \\
+		DVogan's Tech Support 2		& 2018	&  32 & 36.1\% \\
+	\end{tabular}
+	\end{center}
+	\caption{Song packs with the most jumps bracketable.}
+	\label{tab:pack-bjj}
+\end{table}
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r}
+		\bf Author name & \bf \#charts & \bf BJS\% \\
+		\hline
+		Paul J Kim & 45 & 9.9 \\
+		sssmsm & 45 & 4.4 \\
+		Liam & 11 & 3.4 \\
+		Snooze & 22 & 2.5 \\
+		M. Emirzian & 26 & 2.2 \\
+		bblum & 40 & 1.9 \\
+		Ninevolt & 10 & 1.7 \\
+		B. Vergara & 95 & 1.6 \\
+		C. Emirzian & 18 & 1.6 \\
+		B. Dinh & 10 & 1.6 \\
+		Renard & 51 & 1.5 \\
+	\end{tabular}
+	\end{center}
+	\caption{Step cartographers who write charts densest in bracket jumps.}
+	\label{tab:author-bjs}
+\end{table}
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r}
+		\bf Author name & \bf \#charts & \bf BJJ\% \\
+		\hline
+		Paul J Kim & 45 & 71.8 \\
+		Rust & 97 & 51.7 \\
+		bblum & 40 & 49.3 \\
+		Liam & 11 & 44.5 \\
+		Snooze & 22 & 38.1 \\
+		Rems & 10 & 36.5 \\
+		sssmsm & 45 & 32.9 \\
+		Loak & 19 & 31.7 \\
+		Little Matt & 42 & 31.4 \\
+		Paparazzi & 10 & 29.6 \\
+	\end{tabular}
+	\end{center}
+	\caption{Step cartographers whose jumps are most bracketable.}
+	\label{tab:author-bjj}
+\end{table}
 
 \subsection{Historical Trends}
+\label{sec:eval-years}
 
-% TODO: be like, we set out to prove that ppl are using bracker jumps more often lately
-
-% temporal analysis of pack age vs bracker density
+I sought to prove \cite{dril}'s claim that bracket jumps have skyrocketed in popularity in the most recent years of stepchart-making,
+%I decided to do this
+by finding the progression of bracketiness across past years of stepchart packs.
+However,
+actually assigning a firm release date to every pack on my hard drive turned out to be a feat of internet archaeology unto itself,
+even involving archive.org for one step.
+I'll spare you the details, but you can peruse them in the second spreadsheet of the dataset linked at the start of this section.
+Ultimately, all packs older than 2013 had to be grouped together,
+as no pack collection website, facebook group, or laptop filesystem could accurately date enough packs
+before then to give meaningful sample sizes for each year.
+I also decided to exclude tournament packs such as ECFA from this analysis,
+which curate existing stepcharts written possibly long ago.
 
@@
 \section{Discussion}
+
+% TODO: check figure placement
+\begin{figure*}[t]
+	\begin{center}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-beachparty.png}
+		\\
+		(a) Beach Party, \\ 15 \cite{beachparty} %\\
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-koopa.png}
+		\\
+		(b) koopa bling, \\ 69 \cite{koopa} %\\
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-zombiesunet.png}
+		\\
+		(c) Zombie Sunset, \\ 17 \cite{zombie} %\\
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-owa.png}
+		\\
+		(d) 1-Winged Angel, \\ 14 \cite{owa} %\\
+	\end{tabular}
+	\begin{tabular}{c}
+		\includegraphics[width=0.16\textwidth]{disco-paradise.png}
+		\\
+		(e) Electrical Paradise, \\ 12 \cite{e-paradise} %\\
+	\end{tabular}
+	\end{center}
+	%\caption{Why I gotta write a overall caption, just read the prose on this one.}
+	\caption{Example real-world charts with
+	(a) false negative jumps that mines render bracketable after all,
+	(b) jumps whose footing is affected by preceding holds,
+	%false negative jumps forced to bracket by preceding holds, % XXX: this is a lie!! bonus pts if anyone spots it. i don't actually know why the alg didn't classify
+	(c) brackets the chart author was sorry about,
+	(d) ``retro'' style unbracketable jump-stream, %which, largely, cannot be bracketed,
+	and
+	(e) very intentionally many true bracket jumps from the BJJ\% category winner.}
+	\label{fig:discussion}
+\end{figure*}
+
+I confirmed in stepmania the bracketiness of each of \Cref{sec:bjness}'s high scorers,
+and happily observed no false positives (i.e., all charts seemed to intend the player to bracket),
+surprising myself for the 3rd time running at the accuracy of the algorithm.
+I did observe some false negatives, i.e., jumps seeming intended to be bracketed that the algorithm wouldn't catch.
+In Beach Party \cite{beachparty}, shown in \Cref{fig:discussion}(a),
+the ``mine jump'' forces the player to reset her footing,
+removing the right foot's initial presence on U to allow LU to be bracketed.
+This would actually be pretty easy to fix (just have the algorithm parse mines like, at all)
+but it's the day of the deadline and I already ran the experiment, so.
+Another class of false positives showed up in koopa bling \cite{koopa} (\Cref{fig:discussion}(b)),
+where one foot is fixed to a hold, affecting the footing of subsequent jumps.%
+\footnote{The algorithm actually foots them correctly in this chart by sheer luck (an even number of preceding doublesteps maintains footing parity),
+but these would be harder to handle in general.}
+For kicks, I also show the source of the ``I'm sorry'' stepchart description in \Cref{fig:discussion}(c),
+from Zombie Sunset \cite{zombie} (\Cref{tab:bjj});
+note here the DUR bracket-triple-jumps in the third measure,
+which the algorithm currently cannot process.
 
@@
 and it even includes a section of stream with wholly unbracketable jumps interspersed.
-\Cref{fig:discussion}(a) shows this section;
+\Cref{fig:discussion}(d) shows this section;
 in this stream, the player would expect to step the red and blue arrows with her right foot and the lime ones with her left,
-but note how the jumps (hint: all in blue)
+but note how the jumps (hint: all in blue arrows, unless you're reading this on dead trees)
 occur as a mix of left and right arrows, with even a UD ``candle'' jump making an appearance.
@@
 around their brackets that causes the algorithm to count them as normal jumps instead.
-Anyway, \Cref{fig:discussion}(b) shows the chart with most bracket jumps (35) among ones with 100 BJJ\%,
+Anyway, \Cref{fig:discussion}(e) shows the chart with most bracket jumps (35) among ones with 100 BJJ\%,
 which I tried out for myself and found quite enjoyable \cite{e-paradise}.
@@
 which former ranks second place to Electrical Paradise among 100 BJJ\%s with 29 brackets.
-Upon inspection, I found this chart identical to the Challenge (shown in \Cref{fig:xb}(a)),
+Upon inspection, I found this chart identical to the Challenge (upcoming in \Cref{fig:xb}(a)),
 only the crossover brackets
 (to be discussed later in \Cref{sec:xover-brackers})
-had been replaced by easier, normal ones.
-
-\begin{figure}[t]
-	\begin{center}
-	\begin{tabular}{cc}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-owa.png}
-		\\
-		(a) One-Winged Angel, \\ 14 \cite{owa} %\\
-		%(10, Tech. Showcase 3)
-	\end{tabular}
-		&
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-paradise.png}
-		\\
-		(b) Electrical Paradise, \\ 12 \cite{e-paradise} %\\
-		%(12, Tech. Showcase 1)
-	\end{tabular}
-	\end{tabular}
-	\end{center}
-	%\caption{Why I gotta write a overall caption, just read the prose on this one.}
-	\caption{Example charts with
-	(a) ``retro'' style jump-stream which, largely, cannot be bracketed,
-	and
-	(b) very intentionally many bracket jumps.}
-	\label{fig:discussion}
-\end{figure}
+having been replaced by easier, normal ones.
 
@@
 % these can be cheated anyway
+Note however the high measure counter in both, meaning these occur quite late in the song;
+both charts first spend some time ``teaching'' the player both to expect bracket jumps
+and how it intends to signal crossovers in general---so if a human is not expected to understand them at first glance,
+I think it is fair to leave out of the algorithm too for now.
 
@@
 
-I put this section here only just so I could have something to forward-reference back in \Cref{sec:analyzing}.
+I put this section here only just so I could have something to forward-reference back in \Cref{sec:analyzing},
+but not to actually write anything.
 I always wanted to do that, you know?
@@
 
-I never know what to put down here. Much love to the community..!
+% TODO: is it true.
+Bracket jumps is, in fact, the new crossovers or something.
+%I never know what to put down here. Much love to the community..!
+
+\acks
+
+neuropantser (Michael T. Lawson) provided invaluable eleventh-hour (literally) advice
+on \Cref{sec:eval-years}'s statistical analysis, up to and including making the graphs for me.
+And much love to the SIGBOVIK community.
 

### 2019-03-17 itg/3/paper.tex
finish eval secn

@@
 
+\textit{Statistical significance.}
+Stepcharts thus dated, I then summed the total steps, jumps, and brackets published in each year,
+and analyzed the resulting overall BJS\% and BJJ\% as a linear regression.
+\Cref{fig:stats} shows the best-fit and 95\% confidence intervals,
+plotted in R \cite{r-lang} with ggplot \cite{ggplot}.
+For BJS\%, the $\le$2012 bucket actually produced the highest data point (at 11.8\%);
+I suspect this simply represents the fact that jumps were more popular overall during ITG's nascency
+(as corroborated by \Cref{tab:pack-bjs}).
+Owing to this and also to its nature as an aggregation over nearly a decade (which spoils the linear model anyway),
+I chose to exclude it from BJS\%,
+but keep it for BJJ\%, which overall jump count has no bearing on.
+(Its linear fit has $\beta$ = 0.012\%, CI = \{-0.08,+0.10\}, not significant.)
+For the two pictured distributions, their confidence intervals do not include 0,
+i.e., a flat line representing no growth,
+so I conclude bracket jumps' increased popularity is statistically significant.
+
+%% more sigfigs values provided
+% the 95% CI for step (excluding <=2012) is _(0.032,0.139)_
+% the 95% CI for jump (including all years) is _(0.396,2.208)_
+% step (not all years): 0.085
+% jump: 1.302
+% step (all years): 0.012 with a 95% CI of (-0.080, 0.104) (edited)
+
+%%%% ggplot code provided by neuropantser
+% # set up data vectors
+% bracket_jump = 100*c(0.1816545127,0.1614241831,0.1486753945,0.1755102041,0.1593747466,0.2210173322,0.2676218299,0.2333917477)
+% year_jump = 2012:2019
+% 
+% bracket_step = 100*c(0.005698341736,0.003965042926,0.00499281212,0.005327880519,0.006349085019,0.007480904624,0.01087723206)
+% year_step = 2013:2019
+% 
+% # run linear models (and manually output the 95% CIs)
+% step_lin = summary(lm(bracket_step~year_step))
+% print(sprintf("95 pct CI for step: (%s,%s)",step_lin$coefficients[2,1]-1.96*step_lin$coefficients[2,2],
+%               step_lin$coefficients[2,1]+1.96*step_lin$coefficients[2,2]))
+% jump_lin = summary(lm(bracket_jump~year_jump))
+% print(sprintf("95 pct CI for jump: (%s,%s)",jump_lin$coefficients[2,1]-1.96*jump_lin$coefficients[2,2],
+%               jump_lin$coefficients[2,1]+1.96*jump_lin$coefficients[2,2]))
+% 
+% # plot the linear trends
+% library(ggplot2)
+% 
+% # directory to store plots -- CHANGE THIS to where you want them stored
+% plotdir = "C:/Users/mtlawson/Pictures/stuff"
+% 
+% # create temp dataset with bracket_jump variables
+% plotdata = data.frame(year_jump,bracket_jump)
+% # construct plot
+% jump = (ggplot()
+%         + geom_point(data=plotdata,aes(x=year_jump,y=bracket_jump),alpha=0.7,col='red')
+%         + geom_smooth(data=plotdata,aes(x=year_jump,y=bracket_jump),method='lm',alpha=0.4)
+%         + xlab("Year")
+%         + ylab("Brackets-Vs-Jumps Pct")
+%         + scale_x_continuous(breaks=year_jump)
+%         + theme_classic())
+% # output to pdf (can switch to PNG or other desired filetype)
+% pdf(file=sprintf("%s/jump_plot.pdf",plotdir))
+% print(jump)
+% dev.off()
+% # create temp dataset with bracket_step variables
+% plotdata = data.frame(year_step,bracket_step)
+% # construct plot
+% step = (ggplot()
+%         + geom_point(data=plotdata,aes(x=year,y=bracket_step),alpha=0.7,col='red')
+%         + geom_smooth(data=plotdata,aes(x=year,y=bracket_step),method='lm',alpha=0.4)
+%         + xlab("Year")
+%         + ylab("Brackets-Vs-Steps Pct")
+%         + scale_x_continuous(breaks=year_step)
+%         + theme_classic())
+% # output to pdf (can switch to PNG or other desired filetype)
+% pdf(file=sprintf("%s/step_plot.pdf",plotdir))
+% print(step)
+% dev.off()
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+		\includegraphics[width=0.42\textwidth]{step_plot.pdf}
+		\\
+		(a) BJS\% across years ($\beta$ = 0.085\%, CI = \{0.032,0.14\}).
+		\\
+		\\
+		\includegraphics[width=0.42\textwidth]{jump_plot.pdf}
+		\\
+		(a) BJJ\% across years ($\beta$ = 1.30\%, CI = \{0.40,2.21\}).
+	\end{tabular}
+	\caption{Linear best fit of bracket jumps' popularity across the years of ITG's history.
+	Grey area depicts 95\% CI.}
+	\label{fig:stats}
+\end{figure}
+
+\textit{Bias.}
+This analysis is prone to selection bias in my own pack downloading habits:
+supposing I suddenly became more interested in playing brackety charts recently,
+that would certainly influence how many bracket jumps appeared each year in this corpus.
+However, I believe the BJJ\% measurement adequately compensates for this,
+for if I suddenly became interested in stamina instead of technical and added a bunch of 10-minute trance packs,
+that would impact only BJS\%.
+As far as I know, I show no preference for more or less jumpy charts overall than the community average,
+leading me to trust the BJJ\% test.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-03-17 itg/3/results/by-pack-etc.sh
add grouping by pack script

@@ -1,0 +1,27 @@
+#!/bin/bash
+
+IFS=$'\n'
+
+input=by-pack-etc.txt
+
+# field 2 = pack
+# field 3 = year
+# field 4 = author
+key=4
+
+# extract the key
+keys=`cat $input | cut -d'|' -f$key | sort | uniq`
+
+#echo "there are `echo "$keys" | wc -l` keys"
+for key in $keys; do
+	matches=`cat $input | grep "|$key|"`
+	entries=`echo "$matches" | wc -l`
+	step_bc=`echo "$matches" | cut -d'|' -f5 | tr '\n' '+' | sed 's/+$//'` # dont know why i need to separate this; IFS maybe?
+	steps=`echo "$step_bc" | bc`
+	jump_bc=`echo "$matches" | cut -d'|' -f6 | tr '\n' '+' | sed 's/+$//'`
+	jumps=`echo "$jump_bc" | bc`
+	bj_bc=`echo "$matches" | cut -d'|' -f7 | tr '\n' '+' | sed 's/+$//'`
+	bjs=`echo "$bj_bc" | bc`
+	# output keyname - num_entries - total_steps - total_jumps - total_bjs - bjs% - bjj%
+	echo -e "$key\t$entries\t$steps\t$jumps\t$bjs"
+done

### 2019-03-17 itg/3/paper.tex
finished i think

@@
 
-What, then, is a bracket jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
+What, then, is a bracket-jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
 Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
@@
 In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
-a high-quality graphics render of a player's typical foot positions during a down+right bracket jump (henceforth ``DR'', et cetera).
+a high-quality graphics render of a player's typical foot positions during a down+right bracket-jump (henceforth ``DR'', et cetera).
 
@@
 	\begin{center}
-		\includegraphics[width=0.48\textwidth]{how-2-bracket.jpg}
+		\includegraphics[width=0.3625\textwidth]{how-2-bracket.jpg}
 	\end{center}
-	\caption{Down+right bracket jump real-world example.}
+	\caption{Down+right bracket-jump real-world example.}
 	\label{fig:how-2-bracket}
@@
 
-\section{Analyzing Bracket-Jumpiness}
+\section{Algorithm Design}
 \label{sec:analyzing}
@@
 since the last version:
-% TODO: figure out these numbers without counting comments
-98 lines of Haskell (which computed crossovers, footswitches, jacks, doublesteps, \textit{and} crossover footswitches)
-has grown to 248 lines (just to handle this one. new. feature),
+62 lines of Haskell (which computed crossovers, footswitches, jacks, doublesteps, \textit{and} crossover footswitches)
+has grown to 156 lines (just to handle this one. new. feature),
 and the once-simple datatype definition of
@@
 
-In lieu of an algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how I handle them.
+The general gist is that in addition to tracking which foot stepped the last arrow and optimizing for alternating feet,
+we also track which arrow(s) each foot was last on,
+and whenever we encounter a LD/LU/DR/UR jump, bracket it if both:
+\begin{enumerate}
+	\item The last foot is opposite the foot required for the jump (e.g. L for LD), and
+	\item The last foot's last note(s) does not intersect the arrows involved in the jump (i.e., R not on D preceding LD).
+\end{enumerate}
+In lieu of a detailed algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how the implementation must handle them,
+which was pretty much my implementation strategy to begin with.
 
-\subsection{Basic Examples}
+\Cref{fig:tests} shows a bunch of patterns with a bracketable jump,
+and each subfigure in \Cref{fig:tests2} shows a corresponding unbracketable case.
+%with a similar pattern.
+Whenever two jumps are pictured in the same unit test, we are concerned with whether the latter one is bracketable.
+I'll now explain how the algorithm handles each case.
 
-Figure \Cref{fig:bracket-example}(a) shows a simple example where every jump (blue)
-can be bracketed by the right foot,
-alternating with
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{ccc}
+		\includegraphics[width=0.14\textwidth]{test-basic-yes.png} & 
+		\includegraphics[width=0.14\textwidth]{test-footswitch-bracket-yes.png} &
+		\includegraphics[width=0.14\textwidth]{test-reset-footing-yes.png} \\
+		(a) All 4 OK
+		&
+		(b) LD, UR OK
+		&
+		(c) UR OK
+		\\
+		\\
+		\includegraphics[width=0.14\textwidth]{test-propagate-through-ud-yes.png} &
+		\includegraphics[width=0.14\textwidth]{test-reset-subsequent-stream-yes.png} &
+		\includegraphics[width=0.14\textwidth]{test-footswitch-jack-yes.png} \\
+		(d) LD OK
+		&
+		(e) UR, LU OK
+		&
+		(f) DR, LD OK
+	\end{tabular}
+	\end{center}
+	\caption{Examples of bracketable jumps.}
+	\label{fig:tests}
+\end{figure}
+\begin{figure}[t]
+	\begin{center}
+	\begin{tabular}{ccc}
+		\includegraphics[width=0.14\textwidth]{test-basic-no.png} &
+		\includegraphics[width=0.14\textwidth]{test-footswitch-bracket-no.png} &
+		\includegraphics[width=0.14\textwidth]{test-reset-footing-no.png} \\
+		(a) 2nd, 4th NG
+		&
+		(b) DR NG
+		&
+		(c) UR NG
+		\\
+		\\
+		\includegraphics[width=0.14\textwidth]{test-propagate-through-ud-no.png} &
+		\includegraphics[width=0.14\textwidth]{test-reset-subsequent-stream-no.png} &
+		\includegraphics[width=0.14\textwidth]{test-footswitch-jack-also-yes.png} \\
+		(d) LU NG
+		&
+		(e) 2nd UR NG
+		&
+		(f) Trick question
+	\end{tabular}
+	\end{center}
+	\caption{Examples of unbracketable jumps,
+	each corresponding to the similar pattern in \Cref{fig:tests}.}
+	\label{fig:tests2}
+\end{figure}
 
-\subsection{Wackier Examples}
+\renewcommand{\theenumi}{\alph{enumi}}
+\begin{enumerate}
+	\item In the OK case, the right foot can bracket every jump (condition 1),
+		as the left foot is always on a different arrow than the jump (condition 2).
+		In the NG case, condition 2 is violated for the second and fourth jumps.
+	\item Version of (a) testing that preceding bracket-jumps also contribute to condition 2.
+		The NG case would be a footswitch bracket requiring rather uncomfortable footing angles.
+	\item In OK, LR cannot be bracketed, but clears the prior footings so UR can.
+		In NG, LR can be bracketed, meaning UR cannot while still alternating feet (condition 1).
+	\item DU jumps can never be bracketed, but can be stepped facing either direction, and this must respect prior footings.
+		In both cases DU is stepped with right foot U, which allows bracketing LD but not LU thereafter.
+	\item In OK, the crossover does not interfere with either bracket and the player can alternate feet throughout.
+		In NG, either doublestepping or jumping normally is required
+		(or a future-work crossover bracket; see \Cref{sec:xover-brackers}).
+	\item In OK, the player easily footswitches on U.
+		In the corresponding case, the player can actually just jack U instead, stepping it twice with her left foot
+		before bracketing DR again with her right.
+\end{enumerate}
+\renewcommand{\theenumi}{\arabic{enumi}}
+
+These examples are available as {\tt test-bracket-*.sm} unit tests in the code repository.
+Hopefully armed with this understanding of the bracket jumping rules, let's now move on to real-world stepcharts.
+
+% non bracket jumps reset the lastarrowLR
+
+% no footswitch bracket - angles of yr ankles hurt
+% propagate thru UD
+% tiebreak preceding stream
+% retroactively cancel if last stream hadta switch
 
@@
 \begin{enumerate}
-	\item Which songs, packs, and/or step-cartographers are bracket jumpiest?
+	\item Which songs, packs, and/or step-cartographers are bracket-jumpiest?
 	\item Is bracket jumps the new crossovers or something?
@@
 
-\subsection{Bracket Jumpiness}
+\subsection{Bracket-Jumpiness}
 \label{sec:bjness}
@@
 by comparing the number of bracket jumps against the total number of steps,
-or the \textit{bracket jump step percentage} (BJS\%);
+or the \textit{bracket-jump step percentage} (BJS\%);
 second,
@@
 secure several top spots on the BJS\% board.
-For comparison's sake, the overall BJJ\% of the entire corpus is 20\% (93k/470k),
+For comparison's sake, the overall BJJ\% of the entire corpus is 19.8\% (93k/470k),
 meaning that roughly 1 in 5 randomly-patterned jumps are by chance bracketable.
@@
 
-% TODO check margins
 \begin{table}[t]
@@
 	\end{center}
-	\caption{Charts of each difficulty with the highest bracket jump density among all steps (BJS\%).}
+	\caption{Charts of each difficulty with the highest bracket-jump density among all steps (BJS\%).}
 	\label{tab:bjs}
@@
 2018 dominates the BJJ\% leaderboard,
-suggesting that older charts' bracket jumps arose by chance simply from having a lot of jumps to begin with.
+suggesting that older charts' bracket-jumps arose by chance simply from having a lot of jumps to begin with.
 I'll come back to this point in \Cref{sec:eval-years}.
@@
 
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{l|r|r}
-		\bf Author name & \bf \#charts & \bf BJS\% \\
-		\hline
-		Paul J Kim & 45 & 9.9 \\
-		sssmsm & 45 & 4.4 \\
-		Liam & 11 & 3.4 \\
-		Snooze & 22 & 2.5 \\
-		M. Emirzian & 26 & 2.2 \\
-		bblum & 40 & 1.9 \\
-		Ninevolt & 10 & 1.7 \\
-		B. Vergara & 95 & 1.6 \\
-		C. Emirzian & 18 & 1.6 \\
-		B. Dinh & 10 & 1.6 \\
-		Renard & 51 & 1.5 \\
-	\end{tabular}
-	\end{center}
-	\caption{Step cartographers who write charts densest in bracket jumps.}
-	\label{tab:author-bjs}
-\end{table}
-
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{l|r|r}
-		\bf Author name & \bf \#charts & \bf BJJ\% \\
-		\hline
-		Paul J Kim & 45 & 71.8 \\
-		Rust & 97 & 51.7 \\
-		bblum & 40 & 49.3 \\
-		Liam & 11 & 44.5 \\
-		Snooze & 22 & 38.1 \\
-		Rems & 10 & 36.5 \\
-		sssmsm & 45 & 32.9 \\
-		Loak & 19 & 31.7 \\
-		Little Matt & 42 & 31.4 \\
-		Paparazzi & 10 & 29.6 \\
-	\end{tabular}
-	\end{center}
-	\caption{Step cartographers whose jumps are most bracketable.}
-	\label{tab:author-bjj}
-\end{table}
-
 \subsection{Historical Trends}
@@
 
-I sought to prove \cite{dril}'s claim that bracket jumps have skyrocketed in popularity in the most recent years of stepchart-making,
+I sought to prove \cite{dril}'s claim that bracket-jumps have skyrocketed in popularity in the most recent years of stepchart-making,
 %I decided to do this
@@
 so I conclude bracket jumps' increased popularity is statistically significant.
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r}
+		\bf Author name & \bf \#charts & \bf BJS\% \\
+		\hline
+		Paul J Kim & 45 & 9.9 \\
+		sssmsm & 45 & 4.4 \\
+		Liam & 11 & 3.4 \\
+		Snooze & 22 & 2.5 \\
+		M. Emirzian & 26 & 2.2 \\
+		bblum & 40 & 1.9 \\
+		Ninevolt & 10 & 1.7 \\
+		B. Vergara & 95 & 1.6 \\
+		C. Emirzian & 18 & 1.6 \\
+		B. Dinh & 10 & 1.6 \\
+		Renard & 51 & 1.5 \\
+	\end{tabular}
+	\end{center}
+	\caption{Step cartographers who write charts densest in bracket-jumps.}
+	\label{tab:author-bjs}
+\end{table}
+
+\begin{table}[t]
+	\begin{center}
+		\small
+	\begin{tabular}{l|r|r}
+		\bf Author name & \bf \#charts & \bf BJJ\% \\
+		\hline
+		Paul J Kim & 45 & 71.8 \\
+		Rust & 97 & 51.7 \\
+		bblum & 40 & 49.3 \\
+		Liam & 11 & 44.5 \\
+		Snooze & 22 & 38.1 \\
+		Rems & 10 & 36.5 \\
+		sssmsm & 45 & 32.9 \\
+		Loak & 19 & 31.7 \\
+		Little Matt & 42 & 31.4 \\
+		Paparazzi & 10 & 29.6 \\
+	\end{tabular}
+	\end{center}
+	\caption{Step cartographers whose jumps are most bracketable.}
+	\label{tab:author-bjj}
+\end{table}
 
@@
 		\\
-		(a) BJS\% across years ($\beta$ = 0.085\%, CI = \{0.032,0.14\}).
+		(a) BJS\% across years ($\beta$ = 0.085\%/yr, CI = \{0.032,0.14\}).
 		\\
@@
 		\\
-		(a) BJJ\% across years ($\beta$ = 1.30\%, CI = \{0.40,2.21\}).
+		(a) BJJ\% across years ($\beta$ = 1.30\%/yr, CI = \{0.40,2.21\}).
 	\end{tabular}
@@
 supposing I suddenly became more interested in playing brackety charts recently,
-that would certainly influence how many bracket jumps appeared each year in this corpus.
+that would certainly influence how many bracket-jumps appeared each year in this corpus.
 However, I believe the BJJ\% measurement adequately compensates for this,
@@
 	(a) false negative jumps that mines render bracketable after all,
-	(b) jumps whose footing is affected by preceding holds,
+	(b) jumps whose footing is affected by preceding holds fixing one foot in place,
 	%false negative jumps forced to bracket by preceding holds, % XXX: this is a lie!! bonus pts if anyone spots it. i don't actually know why the alg didn't classify
@@
 	and
-	(e) very intentionally many true bracket jumps from the BJJ\% category winner.}
+	(e) intentionally many true bracket jumps from the BJJ\% category winner.}
 	\label{fig:discussion}
@@
 Many old charts written without regard to jump patterning allow for many of their jumps to be bracketed anyway.
-I noticed an old stepchart for One-Winged Angel \cite{owa} ranking high in the list of overall total bracket jumps,
+I noticed an old stepchart for One-Winged Angel \cite{owa} ranking high in the list of overall total bracket-jumps,
 with 182,
@@
 around their brackets that causes the algorithm to count them as normal jumps instead.
-Anyway, \Cref{fig:discussion}(e) shows the chart with most bracket jumps (35) among ones with 100 BJJ\%,
+Anyway, \Cref{fig:discussion}(e) shows the chart with most bracket-jumps (35) among ones with 100 BJJ\%,
 which I tried out for myself and found quite enjoyable \cite{e-paradise}.
@@
 is to overdeliver twice; hence, I shamelessly reuse the joke of this section name from last time.
+
+\subsection{UI}
+
+It would be cool to integrate this (and the preceding two algorithms \cite{turniness,crossoveriness})
+into
+%Simply Love \cite{simplylove}, or whatever
+the community's prevailing stepmania theme.
+%may be in the future.
+Currently, as shown in \Cref{fig:songpreview}(a),
+the song preview screen wastes considerable space displaying the count of irrelevant chart aspects
+such as mines (now used primarily for signaling the presence of footswitches)
+and hands (now mostly stepped with the feet by bracketing anyways).
+I bet the community might actually pay attention to my work---the holy grail of research, honestly---if
+popular themes used it to show precisely how technical a chart was, as in (b).
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+	\includegraphics[width=0.46\textwidth]{song-info.png} \\
+		(a) Current state-of-the-art.
+		\\
+		\\
+	\includegraphics[width=0.46\textwidth]{song-info-better.png} \\
+		(b) Artist's conception.
+	\end{tabular}
+	\caption{Song info preview panel for a recent bracket jump-heavy stepchart \cite{divine}.}
+	\label{fig:songpreview}
+\end{figure}
 
@@
 
-However, ambitious stepchart authors have recently experimented with encouraging the player to bracket jumps while crossed-over.
+However, ambitious stepchart authors have recently experimented with encouraging the player to bracket-jumps while crossed-over.
 Such charts attempt to force, or at least hint, these jumps to be bracketed
@@
 	\end{center}
-	\caption{Crossover bracket jumps must be signaled via either (a) mines or (b) holds to remove ambiguity.}
+	\caption{Crossover bracket-jumps must be signaled via either (a) mines or (b) holds to remove ambiguity.}
 	\label{fig:xb}
@@
 
-\subsection{UI}
-
-It would be cool to integrate this (and the preceding two algorithms \cite{turniness,crossoveriness})
-into
-%Simply Love \cite{simplylove}, or whatever
-the community's prevailing stepmania theme.
-%may be in the future.
-Currently, as shown in \Cref{fig:songpreview}(a),
-the song preview screen wastes considerable space displaying the count of irrelevant chart aspects
-such as mines (now used primarily for signaling the presence of footswitches)
-and hands (now mostly stepped with the feet by bracketing anyways).
-I bet the community might actually pay attention to my work---the holy grail of research, honestly---if
-popular themes used it to show precisely how technical a chart was, as in (b).
-
-\begin{figure}[t]
-	\begin{tabular}{c}
-	\includegraphics[width=0.46\textwidth]{song-info.png} \\
-		(a) Current state-of-the-art.
-		\\
-		\\
-	\includegraphics[width=0.46\textwidth]{song-info-better.png} \\
-		(b) Artist's conception.
-	\end{tabular}
-	\caption{Song info preview panel for a recent bracket jump-heavy stepchart \cite{divine}.}
-	\label{fig:songpreview}
-\end{figure}
-
 \subsection{Obligatory Machine Learning Section}
@@
 
-% TODO: is it true.
-Bracket jumps is, in fact, the new crossovers or something.
+Bracket jumps is, in fact, statistically significantly, the new crossovers or something.
 %I never know what to put down here. Much love to the community..!
@@
 on \Cref{sec:eval-years}'s statistical analysis, up to and including making the graphs for me.
+Greg Hanneman unknowingly helped copy-edit my hyphenation.
 And much love to the SIGBOVIK community.

### 2019-03-17 itg/3/paper.tex
proofread

@@
 Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
-(see \cite{turniness}),
+\cite{turniness},
 while a novice player might think they must step with both feet at once, one for each arrow,
@@
 Now, it considers LD, LU, DR, and UR jumps as potentially bracketable,
-allowing the player to continue a stream of alternating feet uninterrupted through.
+allowing the player to continue a stream of alternating feet uninterrupted therethrough.
 
@@
 \end{enumerate}
-In lieu of a detailed algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how the implementation must handle them,
+In lieu of a detailed algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how a player would typically step them,
 which was pretty much my implementation strategy to begin with.
@@
 		The NG case would be a footswitch bracket requiring rather uncomfortable footing angles.
-	\item In OK, LR cannot be bracketed, but clears the prior footings so UR can.
+	\item In OK, DR cannot be bracketed, but clears the prior footings so UR can.
 		In NG, LR can be bracketed, meaning UR cannot while still alternating feet (condition 1).
@@
 I'll come back to this point in \Cref{sec:eval-years}.
-In honorable mentionth place is chart author Halogen-, whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
+In honorable mentionth place is chart author \mbox{Halogen--,} whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
 What a purist!
@@
 
-I sought to prove \cite{dril}'s claim that bracket-jumps have skyrocketed in popularity in the most recent years of stepchart-making,
+I sought to prove \cite{dril}'s claim that bracket-jumps' popularity has skyrocketed in recent years of stepchart-making,
 %I decided to do this
@@
 Ultimately, all packs older than 2013 had to be grouped together,
-as no pack collection website, facebook group, or laptop filesystem could accurately date enough packs
+as no pack collection website, facebook group, or laptop filesystem metadata could accurately date enough packs
 before then to give meaningful sample sizes for each year.
@@
 I chose to exclude it from BJS\%,
-but keep it for BJJ\%, which overall jump count has no bearing on.
+but kept it for BJJ\%, which overall jump count has no bearing on.
 (Its linear fit has $\beta$ = 0.012\%, CI = \{-0.08,+0.10\}, not significant.)

### 2020-03-09 asdf/src/main.rs
implement bayes rule, nbsearch (me irl), and determinims tests

@@ -5,11 +5,11 @@ use naive::NaiveBinarySearch;
 use sim::SimulationState;
 
 fn main() {
-    let s = SimulationState::new(65536, 0.0);
-    let res = s.simulate_n_steps(NaiveBinarySearch::new(), 1_000_000);
-    println!("{:?}", res);
+    // let s = SimulationState::new(65536, 0.0);
+    // let res = s.simulate_n_steps(NaiveBinarySearch::new(), 1_000_000);
+    // println!("{:?}", res);
 
-    let s = SimulationState::new(65536, 0.0);
+    let s = SimulationState::new(65536, 0.5);
     let res = s.simulate_til_confident(NaiveBinarySearch::new(), 0.99);
     println!("{:?}", res);
 }

### 2020-03-09 asdf/src/naive.rs

@@ -1,23 +1,64 @@
 use crate::sim::{
+    BisectAttempt,
     BisectStrategy,
     SimulationState,
 };
 
+// TODO handle contradiction -- keep a history of past "known good" his and los
 pub struct NaiveBinarySearch {
+    lo: usize,
+    hi: Option<usize>,
 }
 
 impl NaiveBinarySearch {
     pub fn new() -> Self {
-        Self { }
+        Self { lo: 0, hi: None }
     }
 }
 
 impl BisectStrategy for NaiveBinarySearch {
-    fn select_commit(&mut self, _state: &SimulationState) -> usize {
-        unimplemented!()
+    fn select_commit(&mut self, state: &SimulationState) -> usize {
+        if self.hi == None {
+            self.hi = Some(state.pdf.len() - 1);
+        }
+        (self.lo + self.hi.unwrap()) / 2
     }
 
-    fn notify_result(&mut self, _commit: usize, _bug_repros: bool) {
-        unimplemented!()
+    fn notify_result(&mut self, result: BisectAttempt) {
+        if result.bug_repros {
+            self.hi = Some(result.commit);
+        } else {
+            self.lo = result.commit + 1;
+        }
     }
+}
+
+#[cfg(test)]
+mod tests {
+    use super::*;
+
+    #[test]
+    fn test_naive_basic() {
+        let mut b = NaiveBinarySearch::new();
+        // we'll say the bug is secretly at 5
+        // we just need this to make `b` initialize its `hi`
+        let s = SimulationState::new(8, 0.0);
+
+        assert_eq!(b.select_commit(&s), 3);
+        b.notify_result(BisectAttempt { commit: 3, bug_repros: false });
+        assert_eq!(b.lo, 4);
+        assert_eq!(b.hi, Some(7));
+
+        assert_eq!(b.select_commit(&s), 5);
+        b.notify_result(BisectAttempt { commit: 5, bug_repros: true });
+        assert_eq!(b.lo, 4);
+        assert_eq!(b.hi, Some(5));
+
+        assert_eq!(b.select_commit(&s), 4);
+        b.notify_result(BisectAttempt { commit: 4, bug_repros: false });
+        assert_eq!(b.lo, 5);
+        assert_eq!(b.hi, Some(5));
+    }
+
+    // TODO: test contradiction
 }

### 2020-03-09 asdf/src/sim.rs

@@ -1,21 +1,25 @@
+use float_cmp::approx_eq;
 use rand::Rng;
+use std::f64::EPSILON;
+
 
 pub trait BisectStrategy {
     fn select_commit(&mut self, state: &SimulationState) -> usize;
-    fn notify_result(&mut self, commit: usize, bug_repros: bool);
+    fn notify_result(&mut self, result: BisectAttempt);
 }
 
 pub struct SimulationState {
-    pub probabilities: Vec<f64>,
-    pub false_positive_rate: f64, // \elem [0,1)
+    pub pdf: Vec<f64>,
+    pub cdf: Vec<f64>,
+    pub false_negative_rate: f64, // \elem [0,1)
     buggy_commit: usize, // shh, its a himitsu!!
-    history: Vec<Attempt>,
+    history: Vec<BisectAttempt>,
 }
 
-struct Attempt {
-    commit: usize,
-    bug_present: bool,
-    bug_repros: bool,
+#[derive(Clone, Copy)]
+pub struct BisectAttempt {
+    pub commit: usize,
+    pub bug_repros: bool,
 }
 
 #[derive(Debug)]
@@ -26,50 +30,130 @@ pub struct SimulateResult {
 }
 
 impl SimulationState {
-    pub fn new(num_commits: usize, false_positive_rate: f64) -> Self {
+    pub fn new(num_commits: usize, false_negative_rate: f64) -> Self {
+        let buggy_commit = rand::thread_rng().gen_range(0, num_commits);
+        Self::new_with_bug_at(num_commits, false_negative_rate, buggy_commit)
+    }
+
+    fn new_with_bug_at(num_commits: usize, false_negative_rate: f64, buggy_commit: usize) -> Self {
         assert!(num_commits > 0, "excuse me?");
-        assert!(false_positive_rate >= 0.0, "you clown");
-        assert!(false_positive_rate < 1.0, "you double clown"); // get it??
-        let initial_p = 1.0 / num_commits as f64;
-        let res = Self {
-            probabilities: std::iter::repeat(initial_p).take(num_commits).collect(),
-            false_positive_rate,
-            buggy_commit: rand::thread_rng().gen_range(0, num_commits),
+        assert!(buggy_commit < num_commits, "bug not in range");
+        assert!(false_negative_rate >= 0.0, "you clown");
+        assert!(false_negative_rate < 1.0, "you double clown"); // get it??
+        let mut res = Self {
[…]

### 2020-03-11 asdf/src/cdfbisect.rs
implement cdfbisect and assert it makes progress

@@ -1,0 +1,120 @@
+use crate::sim::{
+    BisectAttempt,
+    BisectStrategy,
+    SimulationState,
+};
+
+// it just finds where the cdf crosses 0.5 or whatever
+pub struct CdfBisect {
+    bisect_point: f64,
+    earliest_bug_seen: Option<usize>,
+}
+
+impl CdfBisect {
+    // TODO: make these `new()`s take a &SimulationState
+    // so you can initialize using pdf.len and falsepos prob and so on
+    pub fn new() -> Self {
+        Self::new_with_bisect_point(0.5) // probably suboptimal
+    }
+
+    pub fn new_with_bisect_point(bisect_point: f64) -> Self {
+        Self { bisect_point, earliest_bug_seen: None }
+    }
+}
+
+impl BisectStrategy for CdfBisect {
+    fn select_commit(&mut self, state: &SimulationState) -> usize {
+        let res = find_first_eq_or_greater(&state.cdf, self.bisect_point);
+        if Some(res) == self.earliest_bug_seen {
+            assert!(res > 0, "saw bug on commit 0 -> should be 100% already");
+            res - 1
+        } else if res == state.cdf.len() - 1 {
+            // never test the last commit
+            // TODO: simplify this by just initializing earliest_bug_seen to len.
+            // pending the refactor of the TODO above
+            res - 1
+        } else {
+            res
+        }
+    }
+
+    fn notify_result(&mut self, result: BisectAttempt) {
+        if result.bug_repros {
+            match self.earliest_bug_seen {
+                Some(i) if i < result.commit => {}, // progress
+                Some(i) if i == result.commit => panic!("tested known bug twice"),
+                _ => {
+                    self.earliest_bug_seen = Some(result.commit);
+                }
+            }
+        }
+    }
+}
+
+#[cfg_attr(not(test), allow(unused))]
+fn find_last_not_exceeding<T: PartialOrd>(v: &[T], target_value: T) -> usize {
+    assert!(v.len() > 0);
+    let mut lo = 0;
+    let mut hi = v.len();
+    loop {
+        let mid = (lo + hi) / 2;
+        if v[mid] > target_value {
+            hi = mid;
[…]

### 2020-03-11 asdf/src/main.rs

@@ -1,6 +1,8 @@
+mod cdfbisect;
 mod sim;
 mod naive;
 
+use cdfbisect::CdfBisect;
 use naive::NaiveBinarySearch;
 use sim::SimulationState;
 
@@ -9,7 +11,11 @@ fn main() {
     // let res = s.simulate_n_steps(NaiveBinarySearch::new(), 1_000_000);
     // println!("{:?}", res);
 
-    let s = SimulationState::new(65536, 0.5);
+    let s = SimulationState::new(1024, 0.0);
     let res = s.simulate_til_confident(NaiveBinarySearch::new(), 0.99);
+    println!("{:?}", res);
+
+    let s = SimulationState::new(8, 0.5);
+    let res = s.simulate_til_confident(CdfBisect::new(), 0.99);
     println!("{:?}", res);
 }

### 2020-03-11 asdf/src/sim.rs

@@ -12,11 +12,11 @@ pub struct SimulationState {
     pub pdf: Vec<f64>,
     pub cdf: Vec<f64>,
     pub false_negative_rate: f64, // \elem [0,1)
+    pub history: Vec<BisectAttempt>,
     buggy_commit: usize, // shh, its a himitsu!!
-    history: Vec<BisectAttempt>,
 }
 
-#[derive(Clone, Copy)]
+#[derive(Clone, Copy, Debug)]
 pub struct BisectAttempt {
     pub commit: usize,
     pub bug_repros: bool,
@@ -27,6 +27,7 @@ pub struct SimulateResult {
     pub steps: usize,
     pub confidence: f64,
     pub suspected_buggy_commit: usize,
+    pub actual_buggy_commit: usize,
 }
 
 impl SimulationState {
@@ -67,11 +68,13 @@ impl SimulationState {
         for (i, &p) in self.pdf.iter().enumerate() {
             sum += p;
             assert!(p >= 0.0);
-            assert!(p <= 1.0);
-            assert!(self.cdf[i] <= 1.0);
+            // god i hate floating point
+            // TODO: renormalize every so often
+            // assert!(p <= 1.0 + eps);
+            // assert!(self.cdf[i] <= 1.0 + eps);
             assert!(approx_eq!(f64, sum, self.cdf[i], epsilon = eps));
         }
-        assert!(approx_eq!(f64, sum, 1.0, epsilon = eps));
+        // assert!(approx_eq!(f64, sum, 1.0, epsilon = eps));
     }
 
     fn simulate_step(&mut self, commit: usize) -> BisectAttempt {
@@ -140,10 +143,12 @@ impl SimulationState {
 
             self.refresh_cdf();
             // check consistency of the first half
-            assert!(approx_eq!(f64, self.cdf[commit], p_bug_before, epsilon = EPSILON));
+            // let eps = commit as f64 * EPSILON;
+            // assert!(approx_eq!(f64, self.cdf[commit], p_bug_before, epsilon = eps));
             // also this
             let total_p = p_bug_before + p_bug_in_or_after;
-            assert!(approx_eq!(f64, total_p, 1.0, epsilon = EPSILON));
+            let eps = self.pdf.len() as f64 * EPSILON;
+            assert!(approx_eq!(f64, total_p, 1.0, epsilon = eps));
         }
         let res = BisectAttempt { commit, bug_repros };
         self.history.push(res);
[…]

### 2020-03-11 asdf/src/cdfbisect.rs
refactor the new()s

@@ -7,31 +7,30 @@ use crate::sim::{
 // it just finds where the cdf crosses 0.5 or whatever
 pub struct CdfBisect {
     bisect_point: f64,
-    earliest_bug_seen: Option<usize>,
+    earliest_bug_seen: usize,
 }
 
 impl CdfBisect {
     // TODO: make these `new()`s take a &SimulationState
     // so you can initialize using pdf.len and falsepos prob and so on
-    pub fn new() -> Self {
-        Self::new_with_bisect_point(0.5) // probably suboptimal
+    pub fn new(s: &SimulationState) -> Self {
+        Self::new_with_bisect_point(s, 0.5) // probably suboptimal
     }
 
-    pub fn new_with_bisect_point(bisect_point: f64) -> Self {
-        Self { bisect_point, earliest_bug_seen: None }
+    pub fn new_with_bisect_point(s: &SimulationState, bisect_point: f64) -> Self {
+        Self {
+            bisect_point,
+            // never test the last commit
+            earliest_bug_seen: s.pdf.len() - 1,
+        }
     }
 }
 
 impl BisectStrategy for CdfBisect {
-    fn select_commit(&mut self, state: &SimulationState) -> usize {
-        let res = find_first_eq_or_greater(&state.cdf, self.bisect_point);
-        if Some(res) == self.earliest_bug_seen {
+    fn select_commit(&mut self, s: &SimulationState) -> usize {
+        let res = find_first_eq_or_greater(&s.cdf, self.bisect_point);
+        if res == self.earliest_bug_seen {
             assert!(res > 0, "saw bug on commit 0 -> should be 100% already");
-            res - 1
-        } else if res == state.cdf.len() - 1 {
-            // never test the last commit
-            // TODO: simplify this by just initializing earliest_bug_seen to len.
-            // pending the refactor of the TODO above
             res - 1
         } else {
             res
@@ -40,12 +39,10 @@ impl BisectStrategy for CdfBisect {
 
     fn notify_result(&mut self, result: BisectAttempt) {
         if result.bug_repros {
-            match self.earliest_bug_seen {
-                Some(i) if i < result.commit => {}, // progress
-                Some(i) if i == result.commit => panic!("tested known bug twice"),
-                _ => {
-                    self.earliest_bug_seen = Some(result.commit);
-                }
+            if result.commit < self.earliest_bug_seen {
+                self.earliest_bug_seen = result.commit;
+            } else if result.commit == self.earliest_bug_seen {
[…]

### 2020-03-11 asdf/src/main.rs

@@ -12,10 +12,12 @@ fn main() {
     // println!("{:?}", res);
 
     let s = SimulationState::new(1024, 0.0);
-    let res = s.simulate_til_confident(NaiveBinarySearch::new(), 0.99);
+    let nb = NaiveBinarySearch::new(&s);
+    let res = s.simulate_til_confident(nb, 0.99);
     println!("{:?}", res);
 
     let s = SimulationState::new(8, 0.5);
-    let res = s.simulate_til_confident(CdfBisect::new(), 0.99);
+    let c = CdfBisect::new(&s);
+    let res = s.simulate_til_confident(c, 0.99);
     println!("{:?}", res);
 }

### 2020-03-11 asdf/src/naive.rs

@@ -7,26 +7,26 @@ use crate::sim::{
 // TODO handle contradiction -- keep a history of past "known good" his and los
 pub struct NaiveBinarySearch {
     lo: usize,
-    hi: Option<usize>,
+    hi: usize,
 }
 
 impl NaiveBinarySearch {
-    pub fn new() -> Self {
-        Self { lo: 0, hi: None }
+    pub fn new(s: &SimulationState) -> Self {
+        Self {
+            lo: 0,
+            hi: s.pdf.len() - 1,
+        }
     }
 }
 
 impl BisectStrategy for NaiveBinarySearch {
-    fn select_commit(&mut self, state: &SimulationState) -> usize {
-        if self.hi == None {
-            self.hi = Some(state.pdf.len() - 1);
-        }
-        (self.lo + self.hi.unwrap()) / 2
+    fn select_commit(&mut self, _s: &SimulationState) -> usize {
+        (self.lo + self.hi) / 2
     }
 
     fn notify_result(&mut self, result: BisectAttempt) {
         if result.bug_repros {
-            self.hi = Some(result.commit);
+            self.hi = result.commit;
         } else {
             self.lo = result.commit + 1;
         }
@@ -39,25 +39,25 @@ mod tests {
 
     #[test]
     fn test_naive_basic() {
-        let mut b = NaiveBinarySearch::new();
         // we'll say the bug is secretly at 5
         // we just need this to make `b` initialize its `hi`
         let s = SimulationState::new(8, 0.0);
+        let mut b = NaiveBinarySearch::new(&s);
 
         assert_eq!(b.select_commit(&s), 3);
         b.notify_result(BisectAttempt { commit: 3, bug_repros: false });
         assert_eq!(b.lo, 4);
-        assert_eq!(b.hi, Some(7));
+        assert_eq!(b.hi, 7);
 
         assert_eq!(b.select_commit(&s), 5);
         b.notify_result(BisectAttempt { commit: 5, bug_repros: true });
         assert_eq!(b.lo, 4);
-        assert_eq!(b.hi, Some(5));
+        assert_eq!(b.hi, 5);
 
         assert_eq!(b.select_commit(&s), 4);
         b.notify_result(BisectAttempt { commit: 4, bug_repros: false });
         assert_eq!(b.lo, 5);
-        assert_eq!(b.hi, Some(5));
+        assert_eq!(b.hi, 5);
     }
 
     // TODO: test contradiction

### 2020-03-11 asdf/src/sim.rs

@@ -254,7 +254,8 @@ mod tests {
         let n = 1 << logn;
         for buggy_commit in 0..n {
             let s = SimulationState::new_with_bug_at(n, 0.0, buggy_commit);
-            let res = s.simulate_til_confident(NaiveBinarySearch::new(), 1.0);
+            let nb = NaiveBinarySearch::new(&s);
+            let res = s.simulate_til_confident(nb, 1.0);
             assert_eq!(res.steps, logn);
             assert_eq!(res.confidence, 1.0);
             assert_eq!(res.suspected_buggy_commit, buggy_commit);
@@ -268,7 +269,7 @@ mod tests {
             for &bisect_point in &[0.01, 0.1, 0.5, 0.9, 0.99] {
                 for &fp_prob in &[0.0, 0.5, 0.9] {
                     let s = SimulationState::new_with_bug_at(n, fp_prob, buggy_commit);
-                    let c = CdfBisect::new_with_bisect_point(bisect_point);
+                    let c = CdfBisect::new_with_bisect_point(&s, bisect_point);
                     let _res = s.simulate_til_confident(c, 0.99);
                     // lol no! the algorithm could be wrong :))
                     // assert_eq!(res.suspected_buggy_commit, buggy_commit);

### 2020-03-11 asdf/src/sim.rs
Curious strategy

@@ -1,13 +1,14 @@
 use float_cmp::approx_eq;
 use rand::Rng;
-use std::f64::EPSILON;
 
+const EPSILON: f64 = 0.00001;
 
 pub trait BisectStrategy {
     fn select_commit(&mut self, state: &SimulationState) -> usize;
     fn notify_result(&mut self, result: BisectAttempt);
 }
 
+#[derive(Clone)]
 pub struct SimulationState {
     pub pdf: Vec<f64>,
     pub cdf: Vec<f64>,
@@ -63,7 +64,7 @@ impl SimulationState {
     fn assert_consistent(&self) {
         assert_eq!(self.pdf.len(), self.cdf.len());
         // i guess? i don't really understand this stuff
-        let eps = self.pdf.len() as f64 * EPSILON;
+        let eps = EPSILON;
         let mut sum = 0.0;
         for (i, &p) in self.pdf.iter().enumerate() {
             sum += p;
@@ -77,6 +78,74 @@ impl SimulationState {
         // assert!(approx_eq!(f64, sum, 1.0, epsilon = eps));
     }
 
+    fn adjust_pdf_repro(&mut self, commit: usize) {
+        // Test failed! bug definitely present here.
+        // TODO: rewrite this part using bayes' rule, prove it's equivalent
+        // this is the amount of the probability space that will go to 1
+        let bisected_cdf = self.cdf[commit];
+        for i in 0..self.pdf.len() {
+            if i <= commit {
+                // adjust commits before and including the buggy commit
+                // any of them might have introduced the bug
+                // all pdfs before the test point currently sum to bisected_cdf
+                // now we want it to sum to 1. so multiply by 1/bisected_cdf
+                self.pdf[i] /= bisected_cdf;
+            } else {
+                // adjust commits after this buggy commit down to 0
+                // they won't have introduced it
+                self.pdf[i] = 0.0;
+            }
+        }
+        self.refresh_cdf();
+    }
+
+    fn adjust_pdf_no_repro(&mut self, commit: usize) {
+        // Test passed... bug maybe absent, maybe not.
+
+        // this is the amount of the probability space probably withOUT the bug now
+        let bisected_cdf = self.cdf[commit];
+        // apply bayes' rule.
+        // P(A|B) = P(B|A) P(A) / P(B) where
+        // P(bug <= commit | test passed) =
+        // P(test passed | bug <= commit) --> false negative rate
[…]

### 2020-03-11 asdf/src/sim.rs
clean up

@@ -193,15 +193,12 @@ impl SimulationState {
         mut strat: impl BisectStrategy,
         f: impl Fn(&Self) -> bool,
     ) -> SimulateResult {
-        println!("=======================");
         println!("simulating; himitsu @ {}", self.buggy_commit);
         while f(&self) {
             let prev_pdf = self.pdf.clone();
 
             let commit = strat.select_commit(&self);
-            println!("{:?}", commit);
             let result = self.simulate_step(commit);
-            println!("{:?}", self.pdf);
             strat.notify_result(result);
 
             println!("simulating; {:?} -> pdf {:?}", result, self.pdf);

### 2020-03-12 itg/2/retraction/retraction.tex
retract xoveriness

--- 2019-03-17 itg/3/paper.tex
@@
 %\documentclass[preprint]{sigplanconf}
-\documentclass[10pt]{sigplanconf}
+\documentclass[onecolumn]{sigplanconf}
 
@@
 %\usepackage{mathpartir}
-\usepackage[colorlinks=true,
-citecolor=citec,
-linkcolor=linkc,
-urlcolor=urlc,
-]{hyperref}
+\usepackage{hyperref}
 \usepackage{url}
@@
 \usepackage{multirow}
-\usepackage{xspace}
-\usepackage[nameinlink,nosort]{cleveref}
-\usepackage{xeCJK}
 \usepackage[usenames,dvipsnames]{xcolor}
@@
 \usepackage{natbib}
-\usepackage{ulem}
 \usepackage[mathcal]{euscript}
@@
 
-\definecolor{citec}{RGB}{128,0,64}
-\definecolor{linkc}{RGB}{0,64,128}
-\definecolor{urlc} {RGB}{128,64,0}
-
-\setCJKmainfont{HeiseiMinStd-W5}[Path = ./]
-\setmonofont{Inconsolata-Regular}[Path = ./]
 % This Deep Tex Voodoo is from
@@
 
-\conferenceinfo{SIGBOVIK '19}{Pittsburgh, PA, USA}
-\copyrightyear{2019}
+\conferenceinfo{SIGBOVIK '18}{Pittsburgh, PA, USA}
+\copyrightyear{2018}
 \copyrightdata{}
@@
 \title{
-Which ITG Stepcharts are Bracket-Jumpiest?: \\
-In Which They Milk the \\
-「A Boring Follow-Up Paper to \\
+Retraction of \\
+a boring follow-up paper to \\
 ``Which ITG Stepcharts are Turniest?'' \\
-Titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?''」\\
-Series for All Its Worth in Publication Count After All, or: \\
+titled, ``Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?'' \\
 %That `Boring' Stuff Was Part of the Title, BTW. \\
@@
 %Oh, That Was Also Part of the Title.
-Hit Me With An Encore
 }
@@
 
-\begin{abstract}
-	%ITG is a popular dance game in which players step on arrows while listening to music.
-The arrow patterns, indicated by a {\em stepchart}, may range among any level of complexity and difficulty.
-Among the many factors contributing to a stepchart's difficulty is how much the player must turn from side to side.
-	%Other more obvious factors, such as raw speed, have been well studied in prior work.
-	%This paper presents an analytic study of this {\em turniness} factor.
-	%We study the turniness of many existing stepcharts, and present a novel (but unsurprising) approach to automatically generating maximally (or minimally) turny charts.
-	%Among real-world songs, we find stepcharts with overall turniness ranging from 0\% to 81.33\% of the theoretical maximum.
-
-In which I break last last year's promise of no future work.
-
-\end{abstract}
-
-\category{D.D.R.}{Exercise and Fitness}{Arcade Dance Games}
-
-\keywords
-bracket, groove, in, jumps, the
-
-\section{Introduction}
-
-Recent work by \cite{dril}
-%[W. Dril et al., 2019]
-proposed the hypothesis that
-recent stepchart authors have grown bored with the array of technical ITG step patterns
-documented to date \cite{turniness,crossoveriness},
-and have moved on to break the old model's one-foot-per-arrow assumption to allow for even more technical patterns yet.
-I paraphrase their main conclusion as follows:
+In my 2017 paper, a boring follow-up paper to {\em Which ITG Stepcharts are Turniest?} titled,
+{\em Which ITG Stepcharts are Crossoveriest and/or Footswitchiest?} \cite{crossoveriness},
+I wrote of maximum $\mathcal{T}$, XO\%, FS\%, and JK\% values as follows:
 
 \begin{center}
-	\includegraphics[width=0.38\textwidth]{using-my-turn.png}
-\end{center}
-
-\begin{figure}[t]
-	\hspace{-2em}\includegraphics[angle=270,width=0.54\textwidth]{../2/paper.pdf}
-	\caption{(yeah i reused this joke from last time ok deal)}
-	\label{fig:you-stutid-fuckass}
-\end{figure}
-
-All good researchers know that when rise the standards for software or hardware performance
-(or stepchart trickiness, as the case may be),
-they must revisit their own work to prove its ongoing relevance to the research (dance) community at large.
-Thus I must regrettably break the promise \sout{I} the authors set forth in \cite{crossoveriness}
-(see \Cref{fig:you-stutid-fuckass})
-and revisit \sout{my} their old future work section
-(XXX: they said to change any first-person ``our prior work''
-stuff like this for double-blind review but like cmon this makes no sense?
-theyll totally see through this
-(TODO: maybe email the PC chair for advice?
-(FIXME: make sure to remove these comments before the camera-ready deadline!!))),
-extending it to handle these new so-called ``bracket'' jumps.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Overview}
-
-% TODO: go through this paper and make every *other* instance of "bracket-jump" hyphenated
-
-What, then, is a bracket jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
-Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
-(see \cite{turniness}),
-while a novice player might think they must step with both feet at once, one for each arrow,
-experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
-to hit both arrows by
-triggering one arrow's sensor with the heel and the other with the little bitty toesies.
-Pads are typically constructed with a small triangular metal bracket at the corner of each arrow panel,
-as shown in \Cref{fig:bracker-detail},
-which the bridge of the foot must cross to achieve this, hence the name ``bracket jump''.
-In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
-a high-quality graphics render of a player's typical foot positions during a down+right bracket jump (henceforth ``DR'', et cetera).
-
-The reader, or stepper,
-may notice the extreme angle of footing depicted in the latter figure,
-which is necessary to reliably trigger both pad sensors.
-Accordingly, this maneuver is comfortable (and hence preferable to a normal jump)
-only if the player is already facing in roughly the same direction \cite{turniness}.
-%after stepping any preceding patterns.
-Note also that the two ``candle jumps'', LR and DU, are not possible to bracket,
-unless your foot size is (physically) beyond the scope of this work.
-The next section will attempt to codify (ahem) when
-preceding patterns encourage the player to bracket rather than jump,
-and thereby identify how ``bracket jumpy'' each chart is.
-
-\begin{figure}[t]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth]{jims-pix.jpg}
-	\end{center}
-	\caption{Detail of metal corner brackets, this paper's namesake. Photo credit {\tt JIM}.}
-	\label{fig:bracker-detail}
-\end{figure}
-
-\begin{figure}[h]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth]{how-2-bracket.jpg}
-	\end{center}
-	\caption{Down+right bracket jump real-world example.}
-	\label{fig:how-2-bracket}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Analyzing Bracket-Jumpiness}
-\label{sec:analyzing}
-
-I extended the crossoveriness et cetera algorithm from \cite{crossoveriness} to reason about jumps,
-which previously it treated all identically,
-ignoring the arrows involved and allowing the player to reset her footing as desired.
-Now, it considers LD, LU, DR, and UR jumps as potentially bracketable,
-allowing the player to continue a stream of alternating feet uninterrupted through.
-
-\textit{Confession.}
-When I was first brainstorming this project,
-I had some grand visions of unifying the turniness algorithm \cite{turniness}---which
-accounts for U/D steps to figure out how far the player must turn each step,
-but has no idea which way she is actually facing at any given time---with
-the crossoveriness one
-(cited just above; cmon how much do you want me to repeat these same two citations, gimme a break)---which
-totally ignores U/D steps and just figures out which arrows the left and right feet must each step---in
-some theoretically beautiful way to produce the unquestionably perfect footing sequence for each jump.
-%
-However, as I was considering how to incorporate features from last time's algorithm,
-potentially allowing crossover brackets (\Cref{sec:xover-brackers})
-and footswitch brackets (\Cref{sec:fs-brackers}),
-I realized that ultimately there would be no restrictions on what jumps were bracketable;
-the algorithm would oops simply twist and turn as much as necessary to bracket everything,
-and this paper would become just ``Which ITG Stepcharts Have the Most Jumps?'', and like who wants to read that.
-
-One may think to simply try either bracketing each jump or not
-and seeing what combination gives the minimum turniness result,
-but since whether to bracket each jump or not affects subsequent steps' footing,
-each choice cannot in general be solved independently,
-and I hope you can see where this is going.
-Now, the last time I tried to solve an exponentially-sized problem, it took me 7 years
-\cite{landslide-thesis}
-(and I still ended up with a pile of heuristics after all anyway),
-so considering I started this project a week before the deadline,
-I opted instead to just code up a bunch of ad-hoc rules to handle all the different bracket jump patterns
-that occurred to me to write test cases for.
-
-As before, the code is available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
-To give a sense of how much its elegance has been despoiled % hi william
-since the last version:
-% TODO: figure out these numbers without counting comments
-98 lines of Haskell (which computed crossovers, footswitches, jacks, doublesteps, \textit{and} crossover footswitches)
-has grown to 248 lines (just to handle this one. new. feature),
-and the once-simple datatype definition of
-
-\newcommand\hilight[2]{\color{#1}#2\color{black}\xspace}
-\definecolor{pink}{RGB}{128,0,192}
-\definecolor{orange}{RGB}{192,96,0}
-\definecolor{olivegreen}{RGB}{0,127,32}
-\definecolor{brickred}{RGB}{192,0,0}
-\definecolor{commentblue}{RGB}{0,128,192}
-
-\begin{center}
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step} =
-	\hilight{brickred}{L} |
-	\hilight{brickred}{D} |
-	\hilight{brickred}{U} |
-	\hilight{brickred}{R} |
-	\hilight{brickred}{Jump}}
-\end{center}
-
-has become the unwieldy:
-
-\vspace{-1em} % dont know why this is necessary
-\begin{center}
-	\begin{tabular}{l}
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Arrow} =
-	\hilight{brickred}{L} |
-	\hilight{brickred}{D} |
-	\hilight{brickred}{U} |
-	\hilight{brickred}{R}} \\
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Jump} =
-	\hilight{brickred}{LD} |
-	\hilight{brickred}{LU} |
-	\hilight{brickred}{DR} |
-	\hilight{brickred}{UR} |
-	\hilight{brickred}{LR} |
-	\hilight{brickred}{DU} |
-	\hilight{brickred}{Other}} \\
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Step} =
-	\hilight{brickred}{A} \hilight{olivegreen}{Arrow} |
-	\hilight{brickred}{J} \hilight{olivegreen}{Jump}} \\
-	\texttt{\hilight{orange}{data}~\hilight{olivegreen}{Foot} =
-	\hilight{brickred}{LeftFoot} |
-	\hilight{brickred}{RightFoot}} \\
+\begin{tabular}{p{0.87\textwidth}}
+	A chart could conceivably end right before such a step,
+	sneaking through some small $\epsilon$ extra turniness \cite{epsilon}
+	(similar to the case of 270s in \cite{turniness}),
+	\\
+	\\
+	{[\dots]}
+	\\
+	\\
+	By the way, the theoretical maxima for XO\%, FS\%, and JK\%
+	are 50-$\epsilon$, 100-$\epsilon$, and 100-$\epsilon$, respectively \cite{epsilon}.
 \end{tabular}
@@
 
-In lieu of an algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how I handle them.
+\noindent
+However,
+in the experimental results, {\em Tachyon Epsilon} \cite{tachyon-epsilon}
+placed among the lowest-ranking stepchart packs in every category,
+yet I neglected to properly cite Dr. VII's landmark paper from 2014 at that time.
+Though we may never know know what, if anything, is Epsilon?
+we now know at least what it is not: crossovery and/or footswitchy.
 
-\subsection{Basic Examples}
-
-Figure \Cref{fig:bracket-example}(a) shows a simple example where every jump (blue)
-can be bracketed by the right foot,
-alternating with
-
-\subsection{Wackier Examples}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Evaluation}
-
-The experimental corpus has grown even more since last time, now comprising %17253
-17340 % with ups 4
-stepcharts from 182 packs.
-Of note, all three Technical Showcase packs
-(collaborative packs where the community at large was encouraged to submit their freshest beats and judged on complexity)
-were released in the last two years,
-which have proved to be important sources of diverse bracket-jump patterns for this research.
-I would also have liked to re-rank the crossoveriest and footswitchiest charts from last time to include these submissions,
-but I frankly don't have time.
-The results spreadsheet of bracket jump counts
-(as well as more crossovers, footswitches, et cetera for charts newer than the last paper)
-is available for your browsing pleasure at
-\url{https://tinyurl.com/bracketiest}.
-
-% command used to generate the sheet
-% used "|" as sed delimiter bc others appear in song names
-% IFS=$'\n'; for i in `find ~/stepmania/Songs/ | grep '\.sm$'`; do cat "$i" | ./ITG 2>&1 | sed "s|^|$i\t|"; done > ../3/results/results.txt
-
-I pose two evaluation questions, to be answered in the following subsections respectively.
-\begin{enumerate}
-	\item Which songs, packs, and/or step-cartographers are bracket jumpiest?
-	\item Is bracket jumps the new crossovers or something?
-\end{enumerate}
-
-\subsection{Bracket Jumpiness}
-\label{sec:bjness}
-
-I measured the overal bracketiness of each chart in two ways,
-first,
-by comparing the number of bracket jumps against the total number of steps,
-or the \textit{bracket jump step percentage} (BJS\%);
-second,
-by comparing against the total number of jumps only,
-or the \textit{bracket jump jump percentage} (BJJ\%).
-The BJS\% indicates a chart's overall density of steps requiring the player to step across the brackets
-(which depends on how jumpy the chart is to begin with),
-whereas the BJJ\% measures perhaps the author's intentionality in patterning their jumps either as brackets,
-or as random where just some of them happen to be bracketable.
-
-\Cref{tab:bjs} shows the leaderboard for the former metric, and \Cref{tab:bjj} the latter.%
-\footnote{When measuring BJJ\% I excluded charts with fewer than 10 jumps in total,
-which put up a handful of false positives in the 100\% ranks.}
-Note that the aforementioned recent Technical Showcase pack series put up three bracketiest charts in the BJJ\% category;
-meanwhile, the UPS packs, known for their gimmicks and general lack of respect for player comfort,
-secure several top spots on the BJS\% board.
-For comparison's sake, the overall BJJ\% of the entire corpus is 20\% (93k/470k),
-meaning that roughly 1 in 5 randomly-patterned jumps are by chance bracketable.
-The overall BJS\% (a thoroughly less meaningful statistic) is 0.7\%.
-
-% TODO check margins
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|r|r}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJS\% \\
-		\hline
-		10 & SOBA			& Squeaky Beds \&c.	& 366 & 83\% \\
-		11 & Get Off of My Way		& UPS 3			& 115 & 28\% \\
-		12 & Hardware Store		& Keyboard Coll. III	& 207 & 23\% \\
-		13 & Firestorm			& BemaniBeats 3		& 153 & 31\% \\
-		14 & Bounce			& UPS 2			& 126 & 21\% \\
-		15 & Ikaros Dynamite!!!!	& UPS 2			& 238 & 34\% \\
-		16 & Mermaid Island		& Tachyon Alpha		& 346 & 49\% \\
-		17 & Toccata \& Fugue		& CuoReNeRo M.P.	& 255 & 24\% \\
-		% this chart is ALL jumps so it's kinda false positive
-		%420 & Paa			& UPS 4			& 263 & 34\% \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts of each difficulty with the highest bracket jump density among all steps (BJS\%).}
-	\label{tab:bjs}
-\end{table}
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{r|l|l|r|r}
-		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJJ\% \\
-		\hline
-		 9 & Drifting Away		& UPS 4			&  26 &  88\% \\
-		10 & SOBA			& Squeaky Beds \&c.	& 366 &  99\% \\
-		11 & Save Miracles		& ECFA 2019		&  42 &  95\% \\
-		12 & Electrical Paradise	& Chic. Timing Auth.	&  35 & 100\% \\
-		12 & Encore			& Tech. Showcase	&  29 & 100\% \\
-		12 & Nemeton			& Subluminal		&  12 & 100\% \\
-		13 & Decadent Dandy		& Tech. Showcase 3	&  12 & 100\% \\
-		14 & Nageki no Ki		& Valex's M.4-A.A. 8	& 116 &  94\% \\
-		15 & Beach Party		& Tech. Showcase 2	& 162 &  80\% \\
-		16 & Mermaid Island		& Tachyon Alpha		& 346 &  79\% \\
-		17 & Zombie Sunset		& Jummy Jawns 2		& 325 &  88\% \\
-		69 & koopa bling		& UPS 4			&  39 &  98\% \\
-	\end{tabular}
-	\end{center}
-	\caption{Charts of each difficulty with the highest percentage of their total jumps bracketable (BJJ\%).
-	All participants of the 3-way tie for 12-footers are listed.}
-	%filtered by having at least 10 jumps to begin with.}
-	\label{tab:bjj}
-\end{table}
-
-Next up is bracket jumpiness by pack and by author.
-\Cref{tab:pack-bjs,tab:pack-bjj} show the top 10 packs in both bracketiness metrics,
-and \Cref{tab:author-bjs,tab:author-bjj} the top 10 authors
-(filtered by having written at least 10 charts).
-For the packs, I also list the release year of each;
-note how despite the relatively even spread of years in BJS\%,
-2018 dominates the BJJ\% leaderboard,
-suggesting that older charts' bracket jumps arose by chance simply from having a lot of jumps to begin with.
-I'll come back to this point in \Cref{sec:eval-years}.
-In honorable mentionth place is chart author Halogen-, whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
-What a purist!
-
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{l|r|r|r}
-		\bf Pack name & \bf Year & \bf \#charts & \bf BJS\% \\
-		\hline
-		Keyboard Collaboration III		& $\le$2012	&  17 & 12.4\% \\
-		Squeaky Beds and Leaky Faucets		& 2018	& 125 &  4.9\% \\
-		Keyboard Collaboration I		& $\le$2012	&  12 &  3.6\% \\
-		CuoReNeRo MeGaPacK			& N/A	& 460 &  3.3\% \\
-		Mute Sims X2 WIP			& 2018	&  12 &  3.0\% \\
-		r2112					& 2007	&  60 &  2.7\% \\
-		Technical Showcase 3			& 2018	& 196 &  2.7\% \\
-		FoxyMix 4 - Nuclear Overdrive		& $\le$2012	&  67 &  2.6\% \\
-		Technical Showcase 2			& 2017	&  80 &  2.6\% \\
-		Gensokyo Midnight			& $\le$2012 	&  77 &  2.5\% \\
-	\end{tabular}
-	\end{center}
-	\caption{Song packs densest in bracket jumps.}
-	\label{tab:pack-bjs}
-\end{table}
-
-% sorry about my sigfig discrepancies between this and the by-chart table but like, latex and margins yknow
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{l|r|r|r}
-		\bf Pack name & \bf Year & \bf \#charts & \bf BJJ\% \\
-		\hline
-		Feelin' Rusty 3			& 2018	&  51 & 50.2\% \\
-		Squeaky Beds and Leaky Faucets	& 2018	& 125 & 46.7\% \\
-		TYLR's Technical Difficulties	& 2018	&  33 & 44.9\% \\
-		Technical Showcase 3		& 2018	& 196 & 41.7\% \\
-		Mute Sims X2 WIP		& 2018	&  12 & 40.9\% \\
-		Jimmy Jawns 2			& 2015	&  55 & 39.7\% \\
-		Technical Showcase 2		& 2017	&  80 & 38.2\% \\
-		Keyboard Collaboration III	& $\le$2012	&  17 & 38.1\% \\
-		Chicago Timing Authority	& 2018	&  33 & 36.3\% \\
-		DVogan's Tech Support 2		& 2018	&  32 & 36.1\% \\
-	\end{tabular}
-	\end{center}
-	\caption{Song packs with the most jumps bracketable.}
-	\label{tab:pack-bjj}
-\end{table}
-
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{l|r|r}
-		\bf Author name & \bf \#charts & \bf BJS\% \\
-		\hline
-		Paul J Kim & 45 & 9.9 \\
-		sssmsm & 45 & 4.4 \\
-		Liam & 11 & 3.4 \\
-		Snooze & 22 & 2.5 \\
-		M. Emirzian & 26 & 2.2 \\
-		bblum & 40 & 1.9 \\
-		Ninevolt & 10 & 1.7 \\
-		B. Vergara & 95 & 1.6 \\
-		C. Emirzian & 18 & 1.6 \\
-		B. Dinh & 10 & 1.6 \\
-		Renard & 51 & 1.5 \\
-	\end{tabular}
-	\end{center}
-	\caption{Step cartographers who write charts densest in bracket jumps.}
-	\label{tab:author-bjs}
-\end{table}
-
-\begin{table}[t]
-	\begin{center}
-		\small
-	\begin{tabular}{l|r|r}
-		\bf Author name & \bf \#charts & \bf BJJ\% \\
-		\hline
-		Paul J Kim & 45 & 71.8 \\
-		Rust & 97 & 51.7 \\
-		bblum & 40 & 49.3 \\
-		Liam & 11 & 44.5 \\
-		Snooze & 22 & 38.1 \\
-		Rems & 10 & 36.5 \\
-		sssmsm & 45 & 32.9 \\
-		Loak & 19 & 31.7 \\
-		Little Matt & 42 & 31.4 \\
-		Paparazzi & 10 & 29.6 \\
-	\end{tabular}
-	\end{center}
-	\caption{Step cartographers whose jumps are most bracketable.}
-	\label{tab:author-bjj}
-\end{table}
-
-\subsection{Historical Trends}
-\label{sec:eval-years}
-
-I sought to prove \cite{dril}'s claim that bracket jumps have skyrocketed in popularity in the most recent years of stepchart-making,
-%I decided to do this
-by finding the progression of bracketiness across past years of stepchart packs.
-However,
-actually assigning a firm release date to every pack on my hard drive turned out to be a feat of internet archaeology unto itself,
-even involving archive.org for one step.
-I'll spare you the details, but you can peruse them in the second spreadsheet of the dataset linked at the start of this section.
-Ultimately, all packs older than 2013 had to be grouped together,
-as no pack collection website, facebook group, or laptop filesystem could accurately date enough packs
-before then to give meaningful sample sizes for each year.
-I also decided to exclude tournament packs such as ECFA from this analysis,
-which curate existing stepcharts written possibly long ago.
-
-\textit{Statistical significance.}
-Stepcharts thus dated, I then summed the total steps, jumps, and brackets published in each year,
-and analyzed the resulting overall BJS\% and BJJ\% as a linear regression.
-\Cref{fig:stats} shows the best-fit and 95\% confidence intervals,
-plotted in R \cite{r-lang} with ggplot \cite{ggplot}.
-For BJS\%, the $\le$2012 bucket actually produced the highest data point (at 11.8\%);
-I suspect this simply represents the fact that jumps were more popular overall during ITG's nascency
-(as corroborated by \Cref{tab:pack-bjs}).
-Owing to this and also to its nature as an aggregation over nearly a decade (which spoils the linear model anyway),
-I chose to exclude it from BJS\%,
-but keep it for BJJ\%, which overall jump count has no bearing on.
-(Its linear fit has $\beta$ = 0.012\%, CI = \{-0.08,+0.10\}, not significant.)
-For the two pictured distributions, their confidence intervals do not include 0,
-i.e., a flat line representing no growth,
-so I conclude bracket jumps' increased popularity is statistically significant.
-
-%% more sigfigs values provided
-% the 95% CI for step (excluding <=2012) is _(0.032,0.139)_
-% the 95% CI for jump (including all years) is _(0.396,2.208)_
-% step (not all years): 0.085
-% jump: 1.302
-% step (all years): 0.012 with a 95% CI of (-0.080, 0.104) (edited)
-
-%%%% ggplot code provided by neuropantser
-% # set up data vectors
-% bracket_jump = 100*c(0.1816545127,0.1614241831,0.1486753945,0.1755102041,0.1593747466,0.2210173322,0.2676218299,0.2333917477)
-% year_jump = 2012:2019
-% 
-% bracket_step = 100*c(0.005698341736,0.003965042926,0.00499281212,0.005327880519,0.006349085019,0.007480904624,0.01087723206)
-% year_step = 2013:2019
-% 
-% # run linear models (and manually output the 95% CIs)
-% step_lin = summary(lm(bracket_step~year_step))
-% print(sprintf("95 pct CI for step: (%s,%s)",step_lin$coefficients[2,1]-1.96*step_lin$coefficients[2,2],
-%               step_lin$coefficients[2,1]+1.96*step_lin$coefficients[2,2]))
-% jump_lin = summary(lm(bracket_jump~year_jump))
-% print(sprintf("95 pct CI for jump: (%s,%s)",jump_lin$coefficients[2,1]-1.96*jump_lin$coefficients[2,2],
-%               jump_lin$coefficients[2,1]+1.96*jump_lin$coefficients[2,2]))
-% 
-% # plot the linear trends
-% library(ggplot2)
-% 
-% # directory to store plots -- CHANGE THIS to where you want them stored
-% plotdir = "C:/Users/mtlawson/Pictures/stuff"
-% 
-% # create temp dataset with bracket_jump variables
-% plotdata = data.frame(year_jump,bracket_jump)
-% # construct plot
-% jump = (ggplot()
-%         + geom_point(data=plotdata,aes(x=year_jump,y=bracket_jump),alpha=0.7,col='red')
-%         + geom_smooth(data=plotdata,aes(x=year_jump,y=bracket_jump),method='lm',alpha=0.4)
-%         + xlab("Year")
-%         + ylab("Brackets-Vs-Jumps Pct")
-%         + scale_x_continuous(breaks=year_jump)
-%         + theme_classic())
-% # output to pdf (can switch to PNG or other desired filetype)
-% pdf(file=sprintf("%s/jump_plot.pdf",plotdir))
-% print(jump)
-% dev.off()
-% # create temp dataset with bracket_step variables
-% plotdata = data.frame(year_step,bracket_step)
-% # construct plot
-% step = (ggplot()
-%         + geom_point(data=plotdata,aes(x=year,y=bracket_step),alpha=0.7,col='red')
-%         + geom_smooth(data=plotdata,aes(x=year,y=bracket_step),method='lm',alpha=0.4)
-%         + xlab("Year")
-%         + ylab("Brackets-Vs-Steps Pct")
-%         + scale_x_continuous(breaks=year_step)
-%         + theme_classic())
-% # output to pdf (can switch to PNG or other desired filetype)
-% pdf(file=sprintf("%s/step_plot.pdf",plotdir))
-% print(step)
-% dev.off()
-
-\begin{figure}[t]
-	\begin{tabular}{c}
-		\includegraphics[width=0.42\textwidth]{step_plot.pdf}
-		\\
-		(a) BJS\% across years ($\beta$ = 0.085\%, CI = \{0.032,0.14\}).
-		\\
-		\\
-		\includegraphics[width=0.42\textwidth]{jump_plot.pdf}
-		\\
-		(a) BJJ\% across years ($\beta$ = 1.30\%, CI = \{0.40,2.21\}).
-	\end{tabular}
-	\caption{Linear best fit of bracket jumps' popularity across the years of ITG's history.
-	Grey area depicts 95\% CI.}
-	\label{fig:stats}
-\end{figure}
-
-\textit{Bias.}
-This analysis is prone to selection bias in my own pack downloading habits:
-supposing I suddenly became more interested in playing brackety charts recently,
-that would certainly influence how many bracket jumps appeared each year in this corpus.
-However, I believe the BJJ\% measurement adequately compensates for this,
-for if I suddenly became interested in stamina instead of technical and added a bunch of 10-minute trance packs,
-that would impact only BJS\%.
-As far as I know, I show no preference for more or less jumpy charts overall than the community average,
-leading me to trust the BJJ\% test.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Discussion}
-
-% TODO: check figure placement
-\begin{figure*}[t]
-	\begin{center}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-beachparty.png}
-		\\
-		(a) Beach Party, \\ 15 \cite{beachparty} %\\
-	\end{tabular}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-koopa.png}
-		\\
-		(b) koopa bling, \\ 69 \cite{koopa} %\\
-	\end{tabular}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-zombiesunet.png}
-		\\
-		(c) Zombie Sunset, \\ 17 \cite{zombie} %\\
-	\end{tabular}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-owa.png}
-		\\
-		(d) 1-Winged Angel, \\ 14 \cite{owa} %\\
-	\end{tabular}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{disco-paradise.png}
-		\\
-		(e) Electrical Paradise, \\ 12 \cite{e-paradise} %\\
-	\end{tabular}
-	\end{center}
-	%\caption{Why I gotta write a overall caption, just read the prose on this one.}
-	\caption{Example real-world charts with
-	(a) false negative jumps that mines render bracketable after all,
-	(b) jumps whose footing is affected by preceding holds,
-	%false negative jumps forced to bracket by preceding holds, % XXX: this is a lie!! bonus pts if anyone spots it. i don't actually know why the alg didn't classify
-	(c) brackets the chart author was sorry about,
-	(d) ``retro'' style unbracketable jump-stream, %which, largely, cannot be bracketed,
-	and
-	(e) very intentionally many true bracket jumps from the BJJ\% category winner.}
-	\label{fig:discussion}
-\end{figure*}
-
-I confirmed in stepmania the bracketiness of each of \Cref{sec:bjness}'s high scorers,
-and happily observed no false positives (i.e., all charts seemed to intend the player to bracket),
-surprising myself for the 3rd time running at the accuracy of the algorithm.
-I did observe some false negatives, i.e., jumps seeming intended to be bracketed that the algorithm wouldn't catch.
-In Beach Party \cite{beachparty}, shown in \Cref{fig:discussion}(a),
-the ``mine jump'' forces the player to reset her footing,
-removing the right foot's initial presence on U to allow LU to be bracketed.
-This would actually be pretty easy to fix (just have the algorithm parse mines like, at all)
-but it's the day of the deadline and I already ran the experiment, so.
-Another class of false positives showed up in koopa bling \cite{koopa} (\Cref{fig:discussion}(b)),
-where one foot is fixed to a hold, affecting the footing of subsequent jumps.%
-\footnote{The algorithm actually foots them correctly in this chart by sheer luck (an even number of preceding doublesteps maintains footing parity),
-but these would be harder to handle in general.}
-For kicks, I also show the source of the ``I'm sorry'' stepchart description in \Cref{fig:discussion}(c),
-from Zombie Sunset \cite{zombie} (\Cref{tab:bjj});
-note here the DUR bracket-triple-jumps in the third measure,
-which the algorithm currently cannot process.
-
-Many old charts written without regard to jump patterning allow for many of their jumps to be bracketed anyway.
-I noticed an old stepchart for One-Winged Angel \cite{owa} ranking high in the list of overall total bracket jumps,
-with 182,
-simply because it's long and has a lot of steps,
-but upon playing, the patterns definitely do not feel intentional,
-and it even includes a section of stream with wholly unbracketable jumps interspersed.
-\Cref{fig:discussion}(d) shows this section;
-in this stream, the player would expect to step the red and blue arrows with her right foot and the lime ones with her left,
-but note how the jumps (hint: all in blue arrows, unless you're reading this on dead trees)
-occur as a mix of left and right arrows, with even a UD ``candle'' jump making an appearance.
-However, it measures an unremarkable 9 BJS\% and 38 BJJ\%---I just wanted to feature it
-as an example of what non-bracketable jump stream looks like.
-
-Authors seem to realize when their charts are too bracket jumpy for comfort:
-several of the charts appearing in the highest ranks of the spreadsheet had their step author field filled in as ``Stupid'' or ``I'm sorry''.
-In fact, I personally found the charts with 100 BJJ\%, i.e., \textit{all} jumps appearing therein were bracketable,
-to be more tasteful than those with more total BJS\% but a few normal jumps as well.
-I'd hypothesize this is because to achieve 100 BJJ\% requires a certain intentionality,
-resulting in better chart design overall
-(note that such charts necessarily include no LR or DU jumps whatsoever).
-Either that, or the more uncomfortable ones include certain extremely turny/candley/doublesteppy patterns
-around their brackets that causes the algorithm to count them as normal jumps instead.
-Anyway, \Cref{fig:discussion}(e) shows the chart with most bracket jumps (35) among ones with 100 BJJ\%,
-which I tried out for myself and found quite enjoyable \cite{e-paradise}.
-Amusingly, Encore \cite{encore}
-features a Hard chart in addition to its Challenge,
-which former ranks second place to Electrical Paradise among 100 BJJ\%s with 29 brackets.
-Upon inspection, I found this chart identical to the Challenge (upcoming in \Cref{fig:xb}(a)),
-only the crossover brackets
-(to be discussed later in \Cref{sec:xover-brackers})
-having been replaced by easier, normal ones.
-
-\section{Never Work}
-
-The only thing that impresses the research community more than overdelivering on your future work promises
-is to overdeliver twice; hence, I shamelessly reuse the joke of this section name from last time.
-
-\subsection{Crossover Brackets}
-\label{sec:xover-brackers}
-
-Whereas with normal crossovers (e.g., in the sequence L-U-R, stepping the R with the left foot),
-the player's foot can pretty much point whichever way she finds most comfortable;
-however,
-crossover brackets (e.g., in the sequence L-U-DR...),
-the player's foot must point backwards (...stepping the DR's R with her left heel and D with toes),
-inducing extra turniness.%
-\footnote{The other way, e.g. stepping the D with the left heel and R with toes, while your right foot is fixed on U, is extremely uncomfortable at speed. Trust me on this one.}
-Such jumps could instead just be stepped with both feet as normal,
-possibly inducing doublesteps or jacks,
-and indeed that is how the algorithm presently handles them.
-
-However, ambitious stepchart authors have recently experimented with encouraging the player to bracket jumps while crossed-over.
-Such charts attempt to force, or at least hint, these jumps to be bracketed
-via additional stepchart elements: in \Cref{fig:xb}(a), mines on the left arrow force the player to remove her left foot in preparation for the crossover (and being in the middle of fast stream, further discourages doublestepping),
-and in \Cref{fig:xb}(b), extending one of the arrows as a hold encourages the player to bracket the subsequent jump with the other foot.
-In (b)'s case, the player must also switch feet on each of the non-hold arrows,
-effectively making these crossover-footswitch-brackets,
-which would truly be a wonder to identify programmatically.
-% these can be cheated anyway
-Note however the high measure counter in both, meaning these occur quite late in the song;
-both charts first spend some time ``teaching'' the player both to expect bracket jumps
-and how it intends to signal crossovers in general---so if a human is not expected to understand them at first glance,
-I think it is fair to leave out of the algorithm too for now.
-
-\begin{figure}[t]
-	\begin{center}
-	\begin{tabular}{cc}
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{xb-encore.png}
-		\\
-		(a) Encore, 12 \\ \cite{encore} %\\
-		%(12, Tech. Showcase 1)
-	\end{tabular}
-		&
-	\begin{tabular}{c}
-		\includegraphics[width=0.16\textwidth]{xb-atariwave.png}
-		\\
-		(a) Atariwave, 10 \\ \cite{atariwave} %\\
-		%(10, Tech. Showcase 3)
-	\end{tabular}
-	\end{tabular}
-	\end{center}
-	\caption{Crossover bracket jumps must be signaled via either (a) mines or (b) holds to remove ambiguity.}
-	\label{fig:xb}
-\end{figure}
-
-\subsection{Footswitch Brackets}
-\label{sec:fs-brackers}
-
-I put this section here only just so I could have something to forward-reference back in \Cref{sec:analyzing},
-but not to actually write anything.
-I always wanted to do that, you know?
-I mean, footswitch brackets are %also
-a real thing, but still. $\bar{~~}$\textbackslash\_(ツ)\_/$\bar{~~}$
-% not-to-to: put a pict anyway? if you find one
-% actually, joke comes off better w/o
-
-\subsection{UI}
-
-It would be cool to integrate this (and the preceding two algorithms \cite{turniness,crossoveriness})
-into
-%Simply Love \cite{simplylove}, or whatever
-the community's prevailing stepmania theme.
-%may be in the future.
-Currently, as shown in \Cref{fig:songpreview}(a),
-the song preview screen wastes considerable space displaying the count of irrelevant chart aspects
-such as mines (now used primarily for signaling the presence of footswitches)
-and hands (now mostly stepped with the feet by bracketing anyways).
-I bet the community might actually pay attention to my work---the holy grail of research, honestly---if
-popular themes used it to show precisely how technical a chart was, as in (b).
-
-\begin{figure}[t]
-	\begin{tabular}{c}
-	\includegraphics[width=0.46\textwidth]{song-info.png} \\
-		(a) Current state-of-the-art.
-		\\
-		\\
-	\includegraphics[width=0.46\textwidth]{song-info-better.png} \\
-		(b) Artist's conception.
-	\end{tabular}
-	\caption{Song info preview panel for a recent bracket jump-heavy stepchart \cite{divine}.}
-	\label{fig:songpreview}
-\end{figure}
-
-\subsection{Obligatory Machine Learning Section}
-
-I guess you could skip all this fiddly ``algorithm'' stuff by
-hopping on the pads yourself for a few rounds, playing a variety of technical charts,
-and just training a neural network based on how you stepped the patterns.
-It would probably even learn to fake crossovers more at the ends of songs than at the beginnings,
-where you're more likely to be physically tired.
-And like, do you really want that kind of bias in your dataset?
-
-Remember kids, friends don't let friends use ML for problems that are more fun to solve by hand.
-
-\section{Conclusion}
-
-% TODO: is it true.
-Bracket jumps is, in fact, the new crossovers or something.
-%I never know what to put down here. Much love to the community..!
-
-\acks
-
-neuropantser (Michael T. Lawson) provided invaluable eleventh-hour (literally) advice
-on \Cref{sec:eval-years}'s statistical analysis, up to and including making the graphs for me.
-And much love to the SIGBOVIK community.
+In conclusion, please reject my paper.
+I messed up on it.
 

### 2020-03-12 asdf/src/sim.rs
is_known_buggy_commit() and use it for some assertions/optmz

@@ -55,9 +55,28 @@ impl SimulationState {
 
     fn refresh_cdf(&mut self) {
         self.cdf[0] = self.pdf[0];
-        for i in 1..self.pdf.len() {
+        let n = self.pdf.len();
+        for i in 1..n {
+            // TODO figure out why this assertion/shortcut doesn't work
+            // if i < n - 1 && self.cdf[i] /* old value */ == self.cdf[n-1] {
+            //     assert_eq!(self.pdf[i+1], 0.0);
+            //     assert_eq!(self.cdf[i], self.cdf[self.cdf.len()-1]);
+            //     // optimization
+            //     break;
+            // }
             self.cdf[i] = self.cdf[i-1] + self.pdf[i];
         }
+    }
+
+    fn is_known_buggy_commit(&self, i: usize) -> bool {
+        // if false negatives are possible, the only way we can be sure commit `i`
+        // doesn't introduce the bug is if a test on commit `j < i` repros it. so we
+        // should never expect to see the pdf go from 0 back up to nonzero -- if it's 0,
+        // it means the bug was encountered before, and all subsequent commits' pdfs
+        // will also be 0. `assert_consistent` checks this property; see `bug_seen`.
+        // n.b. this is used for assertions and optimizations (not functionality) so
+        // when we're in deterministic mode we just don't ever return true here.
+        self.false_negative_rate != 0.0 && self.pdf[i] == 0.0
     }
 
     fn assert_consistent(&self) {
@@ -65,9 +84,15 @@ impl SimulationState {
         // i guess? i don't really understand this stuff
         let eps = self.pdf.len() as f64 * EPSILON;
         let mut sum = 0.0;
+        let mut bug_seen = false;
         for (i, &p) in self.pdf.iter().enumerate() {
             sum += p;
             assert!(p >= 0.0);
+            if bug_seen {
+                assert!(self.is_known_buggy_commit(i));
+            } else if self.is_known_buggy_commit(i) {
+                bug_seen = true;
+            }
             // god i hate floating point
             // TODO: renormalize every so often
             // assert!(p <= 1.0 + eps);
@@ -92,11 +117,15 @@ impl SimulationState {
             let bisected_cdf = self.cdf[commit];
             for i in 0..self.pdf.len() {
                 if i <= commit {
+                    assert!(!self.is_known_buggy_commit(i), "tested known buggy commit");
                     // adjust commits before and including the buggy commit
                     // any of them might have introduced the bug
                     // all pdfs before the test point currently sum to bisected_cdf
[…]

### 2020-03-12 asdf/src/sim.rs
disable sim prince

@@ -206,7 +206,7 @@ impl SimulationState {
         mut strat: impl BisectStrategy,
         f: impl Fn(&Self) -> bool,
     ) -> SimulateResult {
-        println!("simulating; himitsu @ {}", self.buggy_commit);
+        // println!("simulating; himitsu @ {}", self.buggy_commit);
         while f(&self) {
             let prev_pdf = self.pdf.clone();
 
@@ -214,7 +214,7 @@ impl SimulationState {
             let result = self.simulate_step(commit);
             strat.notify_result(result);
 
-            println!("simulating; {:?} -> pdf {:?}", result, self.pdf);
+            // println!("simulating; {:?} -> pdf {:?}", result, self.pdf);
             assert!(self.pdf != prev_pdf, "no progress");
         }
 

### 2020-03-12 asdf/src/cdfbisect.rs
implement fpprob-vs-bisect-point experiment in main

@@ -11,13 +11,7 @@ pub struct CdfBisect {
 }
 
 impl CdfBisect {
-    // TODO: make these `new()`s take a &SimulationState
-    // so you can initialize using pdf.len and falsepos prob and so on
-    pub fn new(s: &SimulationState) -> Self {
-        Self::new_with_bisect_point(s, 0.5) // probably suboptimal
-    }
-
-    pub fn new_with_bisect_point(s: &SimulationState, bisect_point: f64) -> Self {
+    pub fn new(s: &SimulationState, bisect_point: f64) -> Self {
         Self {
             bisect_point,
             // never test the last commit

### 2020-03-12 asdf/src/main.rs

@@ -1,23 +1,66 @@
+use std::io::Write;
+use std::fs::File;
+use rand::Rng;
+
 mod cdfbisect;
 mod sim;
 mod naive;
 
 use cdfbisect::CdfBisect;
-use naive::NaiveBinarySearch;
 use sim::SimulationState;
 
+fn output(f: &mut File, s: String) {
+    println!("{}", s);
+    f.write_all(s.as_bytes()).expect("couldnt write to file");
+}
+
 fn main() {
-    // let s = SimulationState::new(65536, 0.0);
-    // let res = s.simulate_n_steps(NaiveBinarySearch::new(), 1_000_000);
-    // println!("{:?}", res);
+    let filename = format!("simulate-{}.txt", rand::thread_rng().gen::<usize>());
+    println!("(emitting output to {})", filename);
+    let mut f = File::create(filename).expect("couldn't make ouptput file");
 
-    let s = SimulationState::new(1024, 0.0);
-    let nb = NaiveBinarySearch::new(&s);
-    let res = s.simulate_til_confident(nb, 0.99);
-    println!("{:?}", res);
+    let n = 1024;
+    let conf_thresh = 0.99999;
+    let trials = 65536;
 
-    let s = SimulationState::new(8, 0.5);
-    let c = CdfBisect::new(&s);
-    let res = s.simulate_til_confident(c, 0.99);
-    println!("{:?}", res);
+    output(&mut f, format!("searching for best bisect_poince~~"));
+    output(&mut f, format!("n = {}, confidence goal = {}, trials = {}",
+                           n, conf_thresh, trials));
+    for &false_negative_prob in &[
+        0.0,
+        0.00001, 0.0001, 0.001, 0.01,
+        0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9,
+        0.99, 0.999, 0.9999, 0.99999, // these will take FOREVER
+    ] {
+        output(&mut f, format!("==== trial for P(false negative) = {} ====",
+                               false_negative_prob));
+        for &bisect_point in &[
+            0.30, 0.31, 0.32, 0.33, 0.34, 0.35, 0.36, 0.37, 0.38, 0.39,
+            0.40, 0.41, 0.42, 0.43, 0.44, 0.45, 0.46, 0.47, 0.48, 0.49,
+            0.50, 0.51, 0.52, 0.53, 0.54, 0.55,
+        ] {
+            let mut result_steps = vec![];
+            let mut wrongs = 0;
[…]

### 2020-03-12 asdf/src/naive.rs

@@ -11,6 +11,7 @@ pub struct NaiveBinarySearch {
 }
 
 impl NaiveBinarySearch {
+    #[cfg_attr(not(test), allow(unused))]
     pub fn new(s: &SimulationState) -> Self {
         Self {
             lo: 0,

### 2020-03-12 asdf/src/sim.rs

@@ -303,7 +303,7 @@ mod tests {
             for &bisect_point in &[0.01, 0.1, 0.5, 0.9, 0.99] {
                 for &fp_prob in &[0.0, 0.5, 0.9] {
                     let s = SimulationState::new_with_bug_at(n, fp_prob, buggy_commit);
-                    let c = CdfBisect::new_with_bisect_point(&s, bisect_point);
+                    let c = CdfBisect::new(&s, bisect_point);
                     let _res = s.simulate_til_confident(c, 0.99);
                     // lol no! the algorithm could be wrong :))
                     // assert_eq!(res.suspected_buggy_commit, buggy_commit);

### 2020-03-26 bisect/src/main.rs
make main run a bunch of simulations

@@ -7,16 +7,19 @@ mod strategies;
 mod sim;
 mod test;
 
-use strategies::CdfBisect;
-use sim::SimulationState;
+use strategies::*;
+use sim::*;
 
 fn output(f: &mut File, s: String) {
     println!("{}", s);
     f.write_all(s.as_bytes()).expect("couldnt write to file");
+    f.write_all(b"\n").expect("couldnt write newline to file");
 }
 
-fn main() {
-    let filename = format!("simulate-{}.txt", rand::thread_rng().gen::<usize>());
+fn run_trial<S: BisectStrategy>(strat_fn: impl Fn(&SimulationState) -> S) {
+    let strat_name = strat_fn(&SimulationState::new(1, 0.0)).name();
+    let filename_suffix: usize = rand::thread_rng().gen();
+    let filename = format!("trial-{}-{}.txt", strat_name, filename_suffix);
     println!("(emitting output to {})", filename);
     let mut f = File::create(filename).expect("couldn't make ouptput file");
 
@@ -24,44 +27,64 @@ fn main() {
     let conf_thresh = 0.99999;
     let trials = 65536;
 
-    output(&mut f, format!("searching for best bisect_poince~~"));
-    output(&mut f, format!("n = {}, confidence goal = {}, trials = {}",
-                           n, conf_thresh, trials));
+    output(&mut f, format!("==== searching for best bisect_poince~~ ===="));
+    output(&mut f, format!("==== {}; n = {}, confidence goal = {}, trials = {} ====",
+                           strat_name, n, conf_thresh, trials));
+
     for &false_negative_prob in &[
         0.0,
         0.00001, 0.0001, 0.001, 0.01,
         0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9,
         0.99, 0.999, 0.9999, 0.99999, // these will take FOREVER
     ] {
-        output(&mut f, format!("==== trial for P(false negative) = {} ====",
-                               false_negative_prob));
-        for &bisect_point in &[
-            0.30, 0.31, 0.32, 0.33, 0.34, 0.35, 0.36, 0.37, 0.38, 0.39,
-            0.40, 0.41, 0.42, 0.43, 0.44, 0.45, 0.46, 0.47, 0.48, 0.49,
-            0.50, 0.51, 0.52, 0.53, 0.54, 0.55,
-        ] {
-            let mut result_steps = vec![];
-            let mut wrongs = 0;
-            for _ in 0..trials {
-                'inner: loop {
[…]

### 2020-03-26 bisect/src/strategies/naive.rs

@@ -7,6 +7,7 @@ pub struct NaiveBinarySearch {
     memory: Option<Vec<usize>>,
 }
 
+// TODO add mistrustful mode, of a certain N number of reruns
 pub enum ConfusedHumanMode { ForgetEverything, UsePreviousLow }
 
 impl NaiveBinarySearch {

### 2020-03-26 bisect/src/entropy.rs
fix the build

@@ -82,6 +82,7 @@ impl SimulationState {
 mod tests {
     use crate::sim::test_helpers::run_pdf_invariant_test;
     use super::*;
+    use std::f64::EPSILON;
 
     #[test]
     fn test_entropy() {

### 2020-03-26 bisect/src/strategies/naive.rs
mistrustful mode (but it boken)

@@ -1,14 +1,32 @@
 use crate::sim::*;
 
-#[derive(Debug)]
 pub struct NaiveBinarySearch {
     lo: usize,
     hi: usize,
-    memory: Option<Vec<usize>>,
+    memory: Memory,
 }
 
-// TODO add mistrustful mode, of a certain N number of reruns
-pub enum ConfusedHumanMode { ForgetEverything, UsePreviousLow }
+struct Retry {
+    commit: usize,
+    remaining_retries: usize,
+}
+
+enum Memory {
+    PreviousLows(Vec<usize>),
+    Retry {
+        max_retries: usize,
+        active_retry: Option<Retry>,
+    },
+    BlissfulIgnorance,
+}
+
+// mistrustful(3) == will run each passing commit 4 times before moving on
+// mistrustful(0) is equivalent to usepreviouslow
+pub enum ConfusedHumanMode {
+    ForgetEverything,
+    UsePreviousLow,
+    Mistrustful(usize)
+}
 
 impl NaiveBinarySearch {
     pub fn new(s: &SimulationState, how: ConfusedHumanMode) -> Self {
@@ -16,8 +34,12 @@ impl NaiveBinarySearch {
             lo: 0,
             hi: s.pdf.len() - 1,
             memory: match how {
-                ConfusedHumanMode::ForgetEverything => None,
-                ConfusedHumanMode::UsePreviousLow => Some(vec![]),
+                ConfusedHumanMode::ForgetEverything => Memory::BlissfulIgnorance,
+                ConfusedHumanMode::UsePreviousLow => Memory::PreviousLows(vec![]),
+                ConfusedHumanMode::Mistrustful(max_retries) => {
+                    assert_ne!(max_retries, 0, "just use ForgetEverything in this case");
+                    Memory::Retry { max_retries, active_retry: None }
+                },
             },
         }
     }
@@ -26,13 +48,16 @@ impl NaiveBinarySearch {
 impl BisectStrategy for NaiveBinarySearch {
     fn name(&self) -> String {
         match self.memory {
-            Some(_) => "naive_memory".to_string(),
-            None => "naive_forgetful".to_string(),
+            Memory::PreviousLows(_) => "naive_memory".to_string(),
+            Memory::Retry { max_retries, .. } => format!("naive_mistrustful_{}", max_retries),
+            Memory::BlissfulIgnorance => "naive_forgetful".to_string(),
         }
     }
 
[…]

### 2020-03-26 bisect/src/test.rs

@@ -59,8 +59,10 @@ fn test_cdfbisect_progress() {
 
 #[test]
 fn test_naive_progress() {
-    run_progress_test(64, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::ForgetEverything));
-    run_progress_test(64, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::UsePreviousLow));
+    run_progress_test(16, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::ForgetEverything));
+    run_progress_test(16, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::UsePreviousLow));
+    // TODO this boken
+    // run_progress_test(16, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::Mistrustful(1)));
 }
 
 #[test]

### 2020-03-26 bisect/src/main.rs
make mistrustful be trustful sometimes :))

@@ -17,7 +17,7 @@ fn output(f: &mut File, s: String) {
 }
 
 fn run_trial<S: BisectStrategy>(strat_fn: impl Fn(&SimulationState) -> S) {
-    let strat_name = strat_fn(&SimulationState::new(1, 0.0)).name();
+    let strat_name = strat_fn(&SimulationState::new(1, 0.5)).name();
     let filename_suffix: usize = rand::thread_rng().gen();
     let filename = format!("trial-{}-{}.txt", strat_name, filename_suffix);
     println!("(emitting output to {})", filename);

### 2020-03-26 bisect/src/strategies/naive.rs

@@ -21,7 +21,7 @@ enum Memory {
 }
 
 // mistrustful(3) == will run each passing commit 4 times before moving on
-// mistrustful(0) is equivalent to usepreviouslow
+// mistrustful(0) is equivalent to forgeteverything
 pub enum ConfusedHumanMode {
     ForgetEverything,
     UsePreviousLow,
@@ -38,7 +38,12 @@ impl NaiveBinarySearch {
                 ConfusedHumanMode::UsePreviousLow => Memory::PreviousLows(vec![]),
                 ConfusedHumanMode::Mistrustful(max_retries) => {
                     assert_ne!(max_retries, 0, "just use ForgetEverything in this case");
-                    Memory::Retry { max_retries, active_retry: None }
+                    if s.false_negative_rate == 0.0 {
+                        // don't bother retrying commits in this mode, obviously
+                        Memory::BlissfulIgnorance
+                    } else {
+                        Memory::Retry { max_retries, active_retry: None }
+                    }
                 },
             },
         }
@@ -175,7 +180,8 @@ mod tests {
         -> (SimulationState, NaiveBinarySearch)
     {
         let retries = if let ConfusedHumanMode::Mistrustful(n) = how { n } else { 0 };
-        let s = SimulationState::new(8, 0.0);
+        // need to have nonzero fnrate here for mistrustful to do its retries
+        let s = SimulationState::new(8, 0.5);
         let mut b = NaiveBinarySearch::new(&s, how);
 
         // push the human to the end of the range with some negative flakes.

### 2020-03-26 bisect/src/test.rs

@@ -59,10 +59,11 @@ fn test_cdfbisect_progress() {
 
 #[test]
 fn test_naive_progress() {
-    run_progress_test(16, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::ForgetEverything));
-    run_progress_test(16, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::UsePreviousLow));
-    // TODO this boken
-    // run_progress_test(16, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::Mistrustful(1)));
+    run_progress_test(64, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::ForgetEverything));
+    run_progress_test(64, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::UsePreviousLow));
+    for &num_retries in &[1, 8, 64] {
+        run_progress_test(64, |s| NaiveBinarySearch::new(s, ConfusedHumanMode::Mistrustful(num_retries)));
+    }
 }
 
 #[test]

### 2020-03-29 bisect/src/sim.rs
add example test

@@ -495,4 +495,30 @@ mod tests {
             }
         })
     }
+
+    #[test]
+    fn test_figure_1() {
+        let mut s = SimulationState::new_with_bug_at(16, 0.5, 11);
+
+        s.adjust_pdf_no_repro(7);
+        s.adjust_pdf_bug_repros(11);
+        assert_eq!(s.pdf[0], 0.0625);
+        assert_eq!(s.pdf[11], 0.125);
+
+        s.adjust_pdf_no_repro(9);
+        assert_eq!(s.pdf[0], 0.05);
+        assert_eq!(s.pdf[9], 0.1);
+        assert_eq!(s.pdf[10], 0.2);
+
+        s.adjust_pdf_no_repro(10);
+        assert!(s.pdf[11] < 1.0 / 3.0 + s.epsilon());
+        assert!(s.pdf[11] > 1.0 / 3.0 - s.epsilon());
+
+        for _ in 0..17 {
+            s.adjust_pdf_no_repro(10);
+        }
+        assert!(s.pdf[11] < 0.99999 - s.epsilon());
+        s.adjust_pdf_no_repro(10);
+        assert!(s.pdf[11] > 0.99999 + s.epsilon());
+    }
 }

### 2020-03-29 bisect/paper/paper.tex
write a lot of paper

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

%\documentclass[preprint]{sigplanconf}
\documentclass[11pt]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage[colorlinks=true,
citecolor=citec,
linkcolor=linkc,
urlcolor=urlc,
]{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage{xspace}
\usepackage[nameinlink,nosort]{cleveref}
\usepackage{xeCJK}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}
\usepackage{ulem}
\usepackage[mathcal]{euscript}
\usepackage[linesnumbered,ruled]{algorithm2e}

\renewcommand{\UrlBreaks}{\do\/\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t\do\u\do\v\do\w\do\x\do\y\do\z}

% ____________________________________________________________
% Listings Package Configuration
% \usepackage[scaled]{beramono}

%\renewcommand*\ttdefault{txtt}
\usepackage[T1]{fontenc}
% \usepackage{inconsolata}

\definecolor{citec}{RGB}{128,0,64}
\definecolor{linkc}{RGB}{0,64,128}
\definecolor{urlc} {RGB}{128,64,0}

\setCJKmainfont{HeiseiMinStd-W5}[Path = ./]
\setmonofont{Inconsolata-Regular}[Path = ./]

\newfontfamily\birbaslo{BirbasloText}[Path = ./]

% This Deep Tex Voodoo is from
%   <http://www.latex-community.org/forum/viewtopic.php?f=5&t=2072>
% It's purpose is to make \lstinline normal size, without affecting
% \lstinputlisting.  It seems to work but I have no idea how or why,
% and I rather hope never to learn.
%\makeatletter
%\lst@AddToHook{TextStyle}{\let\lst@basicstyle\ttfamily\normalsize}
%\makeatother

\begin{document}

\conferenceinfo{{\birbaslo sigbovik~'20}}{pittsburgh, PA, USA}
\copyrightyear{2020}
\copyrightdata{}

\title{
% how to \texttt{git bisect} when your test suite hates you
% how to \texttt{git bisect} when your test suite is a hater
how to \texttt{git bisect} when your test suite is having a bad day
% how to {\tt git bisect} when yr test suite is just, trying its best ok % there we go
}
% \subtitle{\em The Randomly-Scoped Lambda Calculus}
% \subtitle{Subtitle Text, if any}
% \subtitle{a probability puzzle}

\authorinfo{ben blum}{}{bblum@alumni.cmu.edu}

\maketitle

\begin{abstract}
	I like probability puzzles but don't know enough stats to actually solve them properly
	so I threw some thinking sand at it and learned some interesting stuff anyway.
\end{abstract}

% \category{D.D.R.}{Exercise and Fitness}{Arcade Dance Games}

\keywords pdflatex
% bracket, groove, in, jumps, the

\newcommand\confidents{\ensuremath{\mathcal{Z}}\xspace}
\newcommand\pdf{\ensuremath{\mathsf{pdf}}\xspace}
\newcommand\cdf{\ensuremath{\mathsf{cdf}}\xspace}

\section{problem statement}

Let's say you're trying to bisect
% a bug in
a given range of $n$ commits.%
\footnote{To use binary search to find a commit that introduced a bug.}
Call them $c_0 \dots c_{n-1}$, where $c_0$ is known safe and $c_n$ is known to have the bug.
You'd probably start by testing $c_{n/2}$, right?
And you'd expect to pinpoint the buggy comit in $\log_2(n)$ steps.
That's math.

Ok, but what if the bug reprodues nondeterministically with some probability $p<1$.
You can't even pinpoint the bug at some $c_{b}$ for sure anymore;
you can at best know
that it \textasciitilde{}prooobably\textasciitilde~won't reproduce in any $c_{i<b}$,
with some confidence $\confidents$.
Now evaluate your strategy by the expected number of steps to achieve, let's say for tradition's sake, $\confidents \ge 0.99999$.
Is it still optimal to bisect at the midpoint?%
\footnote{Spoiler: No, or I wouldn't have bothered with this paper.}
What would be a better strategy, as a function of $p$?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{intermission}

Put the paper on pause now and think about it.
No, really, give it a go!
I mean, if you don't think math puzzles like this are cool, just stop reading, no worries.
I'm sure there's a paper about, like, hygienic image macros or empire machines or something for you just a few page-turns away.
% I'm sure tom7 is up to some wacky hijinks or something just a few page-turns away.

If you're feeling adventurous,
implement a strategy and throw it in this simulator I implemented to see how it stacks up:
\url{https://github.com/bblum/sigbovik/blob/master/bisect}.
You just gotta implement {\tt trait BisectStrategy},
and it even does all the hard work of applying Bayes's rule
%computing the
%distribution of
%priors with Bayes's rule
for you and letting you see the PDF and everything.
Check it out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{pdfs, but not the portable document format kind, and our friend rev. bayes}

Ok, so let's model the problem as a sequence of steps on a probability distribution function
(henceforth, PDF; and also, CDF for the cumulative kind).
Initially, $\pdf(i) = 1/n$ for all $0 \le i < n$.
When the bug reproduces at some commit $b$,
you're certain no $c_{i>b}$ introduced the bug, so $\pdf(i>b) = 0$,
and $\pdf(i\le b)$ renormalizes as $1/b$, to satisfy the $\int \pdf(i) \text{d} i = 1$ invariant.%
\footnote{Implemented as {\tt fn adjust\_pdf\_bug\_repros()} in the simulator.}

In the deterministic case ($p=1$),
the symmetric thing happens when the test passes at some $c_j$: $\pdf(i\le j) = 0$.
But when $p<1$, we must generalized it \cite{mario3}. Bayes's rule provides that:
\[
P(A|B) = \frac{P(B|A) \times P(A)}{P(B)}
\]

\newcommand\renorm[1]{\ensuremath{\mathcal{R}_{#1}}\xspace}

In this case, $B$ is that the test passed at $c_j$, and $A$ is that bug be present at or before $c_j$ after all.
$P(B|A)$ is simply $1-p$.
$P(A)$ is the prior on $c_j$ containing the bug, i.e., $\cdf(j)$.
And $P(B)$ is the false negative probability weighted by the bug existing, i.e., $(1-p)\cdf(j) + 1(1-\cdf(j))$.
% simplifies to: cdf(j) - p*cdf(j) + 1 - cdf(j)
% simplifies to: 1 - p*cdf(j)
To update our priors on commits up to $j$, we renormalize by dividing out the old $\cdf(j)$
and multiplying by the new $P(A|B)$, i.e.,
%To update our priors on commits up to $j$, we divide out the old $\cdf(j)$
%and multiply by the new $P(A|B)$.
%Call this renormalization factor $\renorm{j}{i}$.
%\[
%	\forall i \le j \Rightarrow \renorm{i}{i} = 
%	\frac{1}{\cdf(j)}
%	\frac{(1-p)\cdf(j)}{(1-p)\cdf(j) + (1-\cdf(j))}
%\]
\[
	\forall i \le j, \pdf'(i)
	\leftarrow
	\pdf(i)
	\frac{1}{\cdf(j)}
	\frac{(1-p)\cdf(j)}{(1-p)\cdf(j) + (1-\cdf(j))}
\]
Which simplifies to:
\[
	\forall i \le j, \pdf'(i)
	\leftarrow
	\pdf(i)
	\frac{1-p}{1 - p\cdf(j)}
\]
Call this renormalization factor $\renorm{}$.
As a sanity check, $p\cdf(j)$ is less than $p$, so $\renorm{i\le j} < 1$.

Likewise, for commits above $j$, we have $P(B|A) = 1$, $P(A) = 1-\cdf(j)$, and $P(B)$ the same as before.
Renormalizing from $1-\cdf(j)$ this time (and skipping the unsimplified version), we get:
\[
	\forall i > j, \pdf'(i)
	\leftarrow
	\pdf(i)
	\frac{1}{1 - p\cdf(j)}
\]
As a sanity check, $p\cdf(j)$ is positive, so $\renorm{i > j} > 1$.
If you like pen-and-paper algebra, you can also see that
$
\cdf(j)
\renorm{i \le j}
+
(1-\cdf(j))
\renorm{i>j}
= 1$.%
\footnote{Implemented as {\tt fn adjust\_pdf\_no\_repro()} in the simulator.}

Let's do a nice concrete example.
Say $n=16$, the test passes at $j=7$, and then the bug repros at $j=11$.
In the deterministic case, all the probability mass will be concentrated uniformly in the range $[8,11]$.
However, if the bug repros only half the time,
%the renormalization factors around $c_7$ come out to $2/3$ and $1/3$
$\renorm{i \le 7} = 2/3$
and
$\renorm{i > 7} = 4/3$,
and we get probability mass scattered all the way down to $c_0$,
as shown in \Cref{fig:example}(a).
Yuck, someone clean that up!

Now let's say the test passes at $j=9$, then at $j=10$.
\Cref{fig:example}(b) shows the updated PDFs/CDFs:
for $p=1$, this pinpoints the bug at $j=11$, and the search is over.
But for $p=0.5$, there's still 2/3 odds we'd be wrong!
In fact, from here it takes {\it eighteen} further probes at $j=10$ until we are at least five 9's confident that $c_{11}$ is the culprit.%
\footnote{See {\tt fn test\_figure\_1()}!}

% cdf 9 -- 0.75
% 
% R- = 1/2 / 1-((3/4)/2)
% R- = 1/2 / 5/8
% R- = 8/10 = 4/5
% 
% R+ = 1 / 5/8
% R+ = 8/5 = 16/10

% cdf 10 - 0.8
% 
% R- = 1/2 / 1-((4/5)/2)
% R- = 1/2 / 6/10
% R- = 10/12 = 5/6
% 
% R+ = 1 / 6/10
% R+ = 10/6

\begin{figure}[t]
	\begin{tabular}{c}
	\includegraphics[width=0.45\textwidth]{example.pdf}
	\\
		(a) After two tests.
		\\
		\\

	\includegraphics[width=0.45\textwidth]{example2.pdf}
	\\
		(b) After ``classical`` search terminates.
	\end{tabular}
	%\caption{Example deterministic and nondeterministic PDFs and CDFs.}
	\caption{Example \{,non\}deterministic \{P,C\}DFs.}
	\label{fig:example}
\end{figure}

A noteworthy invariant here is that the PDF is always monotonically nondecreasing in its nonzero range:
each passing test always shifts probability mass to the right of the bisect point,
but past the earliest known bug repro, nothing can ever revive it back above 0.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{prior work}

I was kinda surprised to find no existing mathy solution to this problem lying around on the online.
Wikipedia has a brief little subsection on ``noisy binary search'', which links a few papers older than I am.
In one \cite{noisy2}, they bound the number of erroneous answers by a fixed factor of the number of queries,
so it's more like ``twenty questions with lies'' than bisect.
In another \cite{noisy1}, they do fix the error rate $p$,
but they allow for symmetric false negatives and false positives, both with the same $p$.
This too changes the nature of the problem;
notably, if $p=0.5$, you can't make any progress whatsoever.

Dropbox has a CI service called Athena \cite{athena}
which automatically searches for flaky tests, when otherwise no bug exists.
In this case the goal is to keep the build green, but if you consider the flaky test itself to be the bug, it's the same problem.%
\footnote{Incidentally, the symmetric case
-- where a bug repros with $p=1$, but the test also flakes with some $q<1$ --
is also the same problem.}
Athena's ``deflakes'' the test at each commit by running 10 times,
and treats the combined result as ``basically as good as $p=1$''.%
\footnote{Upcoming, I will call this strategy $\mathsf{Mistrustful(10)}$.}
In this setting, $p$ is not known in advance, so using Bayes's rule is off the table.
But I will show that even without access to the PDF, a better strategy exists. % at least by my definition

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{strategies}

Ok, so how do we
%cope with the fact that we'll never have a proper lower bound?
%How do we
make progress,
i.e., concentrate probability mass til there's $\confidents$ of it on one commit,
as quickly as possible?
%In principle, testing anywhere in the range $[0,b)$,
%where $b$ is the earliest known buggy commit,
%will make progress,
%
%% ghhh this sentence only feels good if i know what i'm talking about
%Let's deconstruct the motives of classical binary search.
Let $c_b$ be the earliest known buggy commit, and $c_a$ be the latest known safe commit.
In Determinism World, we would bisect at $c_{(a+b)/2}$
because that minimizes the maximum range remaining in the worst case outcome
(i.e., the two possible outcome ranges are the same length).
%But in Nondeterminism World,
%the worst case is a big constant times linear:
%an adversarial RNG god could pass your test a thousand times on $c_{n-1}$, then fail,
%then pass it a thousand times on $c_{n-2}$, then fail, and so on,
%before ultimately arriving at buggy commit $c_0$.
%So thinking about the worst case is not productive for a sublinear strategy,
%and also,
%we shouldn't be thinking about the range length,
%because a pass is still progress even though $a$ doesn't ever budge,
But in Nondeterminism World, $a$ doesn't ever budge from $0$,
so bisecting at $c_{(0+b)/2}$ will not even terminate.
%will eventually fail to make progress.
Sure, hitting the PDF with
%Bayes's rule at $b/2$
$\renorm{b/2}$
will always move {\it some} mass rightward,
but once five 9's of it is already over there,
%to the right of that point,
it can't concentrate it onto one point.
So let's not think about the range.

\subsection{bisect probability mass}

Another way to think about Determinism World is the $c_{(a+b)/2}$ bisect point
% maximizes the expected probability mass that moves from one side to the other.
cuts the probability mass, rather than the range, in half,
% How about in Nondeterminism World,
% we ignore the range length, and instead bisect at the midpoint of probability mass,
i.e., $\mathsf{last}_i \cdf(i) \le 0.5$.
This approach is compatible with Nondeterminism World,
fits the ``binary search'' spirit,
and will at least terminate
(if $b$ is the solution, it converges to repeatedly probing $b-1$),
which is more than you can say for the range-based $c_{(0+b)/2}$.
%It is intuitively obvious that this will make progress faster than bisecting at $c_{(b+0)/2}$,
%But is it fast?
However, it doesn't maximize either the minimum or the expected amount of mass that moves:
when $p<1$, it's intuitively easy to see that CDF threshold should actually be less than $0.5$.
\footnote{I did some pen-and-paper algebra to find a closed form for this,
but I actually trust the simulator more than myself at this point,
so I stopped spending brains on this part of the problem. I kind of regret that.}
% TODO TODO: actually do the paperwork.
Is maximizing the expected moved mass as a function of $p$ the fastest we can go,
and if so, why?

\subsection{bisect entropy}

\newcommand\entropy{\ensuremath{\mathcal{H}}\xspace}

A PDF's information content is measured in {\it entropy}:
$\entropy = \sum_i \pdf(i)\mathsf{ln}(\pdf(i))$.
Indeed,
in Determinism World the search terminates when $\entropy = 0$.
I thought for a while about how to link ``minimum expected entropy''
to the stated goal %in Nondeterminism World:
of
%minimizing the expected steps until
$\confidents = \mathsf{max}_i \pdf(i) \ge 0.99999$,
but couldn't really formalize the idea.
The goal of five 9's is fairly arbitrary anyway,
and also not necessarily stable under the entropy measurement,
since it doesn't care how the remaining $0.00001$ is distributed.
%
Also, a Very Smart Math Friend of mine was unable to find a simple closed form for
the expected resulting $\entropy(\renorm{} \circ \pdf)$
%\[
%(\cdf(j) + (1-\cdf(j))(1-p))\entropy(\renorm \circ \pdf)
%+
%p(1-\cdf(j))\entropy(\renorm{\text{bug}} \circ \pdf)
%\]
or whatever (bleh, my notation is kinda reaching its limits here).
%so how would you write a proof about it anyway.
Nevertheless, it seems promising, and easy to implement.

\subsection{bisect randomly}

Maybe a random world calls for a random algorithm!
It's obvious this will terminate, but of course it won't be fast.
Intuitively, if we use the PDF to weight the random choice, it will be faster than choosing uniformly in $[0,b)$.
But how much faster?

You can tell it's at this point I'm starting to get statistics fatigue.

\subsection{human}

Humans are known to be less patient than computers (citation: see previous sentence).
If a human is unwilling to compute $\renorm{} \circ \pdf$ by hand every step,
or, more prohibitively,
simply doesn't know $p$ in advance so can't compute $\renorm{}$ at all,
what should they do?

Think back to Athena from the prior work section.
The human

Regardless, it's easy to

%(
% hmmm...

% In Determinism World, the search terminates when

% Let's think of bisecting at $c_{(a+b)/2}$, in Determinism World, 
% as maximizing the expected information gained from either test result.
% On PDFs, information gain is measured as

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{simulating it}

At this point I threw in the towel on the maths front,
and wrote a bunch of code to let the law of large numbers do my work for me.

\cite{drowningpoint}
\cite{epsilon}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{evaluation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{future work}

% weigh cost of running test suite vs cost of git checkout -- mistrustful() gets a whole lot better! but, where is the break even point when you should git checkout?

% p not known, figur eit out on the fly

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{conclusion}

\acks

Thanks to Jason Reed, Lynn Chordbug, and Sujay Jayakar for thinking about this problem with me.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\refname}{references}
\bibliographystyle{abbrvnat}
\bibliography{citations}

\end{document}

### 2020-03-29 bisect/src/entropy.rs
test hstab

@@ -102,6 +102,22 @@ mod tests {
         assert!(entropy(&v4d) < entropy(&v4c));
     }
 
+    #[test]
+    fn test_entropy_stability() {
+        let v4e = vec![0.00000_00, 0.00000_00, 0.00001_00, 0.99999_00];
+        let v4f = vec![0.00000_00, 0.00000_00, 0.00000_99, 0.99999_01];
+        let v4g = vec![0.00000_33, 0.00000_33, 0.00000_33, 0.99999_01];
+        assert_eq!(v4e.iter().sum::<f64>(), 1.0);
+        assert_eq!(v4f.iter().sum::<f64>(), 1.0);
+        assert_eq!(v4g.iter().sum::<f64>(), 1.0);
+
+        // as you'd expect, v4f's max is higher
+        assert!(entropy(&v4e) > entropy(&v4f));
+        // but! v4e tells you more about the first 3 than v4g does, despite lower max.
+        assert!(entropy(&v4e) < entropy(&v4g));
+    }
+
+
     // this test validates `strategies::MinExpectedEntropy`'s reliance on binary search
     // to find the bisect point of minimum expected entropy, instead of linear
     #[test]

### 2020-03-29 bisect/paper/paper.tex
write more

@@
 You'd probably start by testing $c_{n/2}$, right?
-And you'd expect to pinpoint the buggy comit in $\log_2(n)$ steps.
+And you'd expect to pinpoint the buggy comit in $\mathsf{log} (n)$ steps.
 That's math.
@@
 When the bug reproduces at some commit $b$,
-you're certain no $c_{i>b}$ introduced the bug, so $\pdf(i>b) = 0$,
-and $\pdf(i\le b)$ renormalizes as $1/b$, to satisfy the $\int \pdf(i) \text{d} i = 1$ invariant.%
+you're certain no $c_{i>b}$ introduced the bug, so each $\pdf(i>b)$ becomes $0$,
+and each $\pdf(i\le b)$ renormalizes by a factor of $n/b$, to preserve the $\int \pdf(i) \text{d} i = 1$ invariant.%
 \footnote{Implemented as {\tt fn adjust\_pdf\_bug\_repros()} in the simulator.}
@@
 the symmetric thing happens when the test passes at some $c_j$: $\pdf(i\le j) = 0$.
-But when $p<1$, we must generalized it \cite{mario3}. Bayes's rule provides that:
+But when $p<1$, we must generalized it \cite{mario3}. Here's Bayes's rule:
 \[
@@
 
-In this case, $B$ is that the test passed at $c_j$, and $A$ is that bug be present at or before $c_j$ after all.
-$P(B|A)$ is simply $1-p$.
+In this case, $B$ is that the test passes at $c_j$, and $A$ is that bug exists at or before $c_j$ after all.
+$P(B|A)$ is the false negative rate, i.e. $1-p$.
 $P(A)$ is the prior on $c_j$ containing the bug, i.e., $\cdf(j)$.
-And $P(B)$ is the false negative probability weighted by the bug existing, i.e., $(1-p)\cdf(j) + 1(1-\cdf(j))$.
+And $P(B)$ is the false negative rate weighted by the bug existing, i.e., $(1-p)\cdf(j) + 1(1-\cdf(j))$.
 % simplifies to: cdf(j) - p*cdf(j) + 1 - cdf(j)
@@
 But for $p=0.5$, there's still 2/3 odds we'd be wrong!
-In fact, from here it takes {\it eighteen} further probes at $j=10$ until we are at least five 9's confident that $c_{11}$ is the culprit.%
+In fact, from here it takes 18 further probes at $j=10$ until we are at least five 9's confident that $c_{11}$ is the culprit.%
 \footnote{See {\tt fn test\_figure\_1()}!}
+Bayes's rule gonna get ya.
 
@@
 Athena's ``deflakes'' the test at each commit by running 10 times,
-and treats the combined result as ``basically as good as $p=1$''.%
-\footnote{Upcoming, I will call this strategy $\mathsf{Mistrustful(10)}$.}
+treating the combined result as ``basically as good as $p=1$'',
+and then runs an otherwise classical binary search.%
+\footnote{Upcoming, I will call this strategy \textsf{Human-Mistrustful}$(10)$.}
 In this setting, $p$ is not known in advance, so using Bayes's rule is off the table.
@@
 make progress,
-i.e., concentrate probability mass til there's $\confidents$ of it on one commit,
+i.e., concentrate probability mass til there's $\confidents$ of it in one place,
 as quickly as possible?
@@
 %% ghhh this sentence only feels good if i know what i'm talking about
-%Let's deconstruct the motives of classical binary search.
+Let's deconstruct the motives of classical binary search.
 Let $c_b$ be the earliest known buggy commit, and $c_a$ be the latest known safe commit.
-In Determinism World, we would bisect at $c_{(a+b)/2}$
-because that minimizes the maximum range remaining in the worst case outcome
-(i.e., the two possible outcome ranges are the same length).
+In Determinism World, bisecting at $c_{(a+b)/2}$
+minimizes the worst case maximum range remaining,
+as the two possible outcome ranges are the same length.
 %But in Nondeterminism World,
@@
 
-Another way to think about Determinism World is the $c_{(a+b)/2}$ bisect point
+Another way to think about Determinism World is that the $c_{(a+b)/2}$ bisect point
 % maximizes the expected probability mass that moves from one side to the other.
@@
 % we ignore the range length, and instead bisect at the midpoint of probability mass,
-i.e., $\mathsf{last}_i \cdf(i) \le 0.5$.
-This approach is compatible with Nondeterminism World,
+i.e., $\mathsf{max}_j (i)$ s.t. $\cdf(j) \le 0.5)$.
+This approach
 fits the ``binary search'' spirit,
-and will at least terminate
-(if $b$ is the solution, it converges to repeatedly probing $b-1$),
-which is more than you can say for the range-based $c_{(0+b)/2}$.
+and will terminate in Nondeterminism World,
+i.e.,
+if $b$ is the solution, it converges to repeatedly probing $b-1$,
+so it can pass any fixed $\confidents$ threshold.
+%(which is more than you can say for the range-based $c_{(0+b)/2}$).
 %It is intuitively obvious that this will make progress faster than bisecting at $c_{(b+0)/2}$,
 %But is it fast?
-However, it doesn't maximize either the minimum or the expected amount of mass that moves:
-when $p<1$, it's intuitively easy to see that CDF threshold should actually be less than $0.5$.
-\footnote{I did some pen-and-paper algebra to find a closed form for this,
-but I actually trust the simulator more than myself at this point,
-so I stopped spending brains on this part of the problem. I kind of regret that.}
-% TODO TODO: actually do the paperwork.
-Is maximizing the expected moved mass as a function of $p$ the fastest we can go,
-and if so, why?
+% However, it doesn't maximize either the minimum or the expected amount of mass that moves:
+% when $p<1$, it's intuitively apparent that this CDF threshold should actually be less than $0.5$.%
+%\footnote{I did some pen-and-paper algebra to find a closed form for this,
+%but I actually trust the simulator more than myself at this point,
+%so I stopped spending brains on this part of the problem. I kind of regret that.}
+%
+%Is maximizing the expected moved mass as a function of $p$ the fastest we can go,
+%and if so, why?
+But is $0.5$ still the best bisect point even when $p<1$?
+I wondered if this could be expressed as the amount of probability mass moved from one side to the other,
+i.e., $\sum_i \mathsf{abs}(\pdf(i) - \pdf'(i))$.
+In the case where the bug repros this is:
+\[
+	p\cdf(j) \times (1-\cdf(j))
+\]
+and in the case where the test passes:
+\[
+	(1-p\cdf(j)) \times \cdf(j) \times (1-\renorm{i\le j})
+\]
+which, surprisingly, simplifies to exactly the same thing as the bug repros case.
+The maximum occurs where $\partial/\partial \cdf(j)$ is 0,
+which turns out to be at $\cdf(j)=0.5$ after all, and independent of $p$.%
+\footnote{I also worked out the {\it minimum} moved mass between the pass and fail case,
+which is almost always the pass case,
+and comes out to
+%$\frac{p\cdf(j)(1-\cdf(j))}{1-p\cdf()}$
+$\frac{\cdf(j)(1-\cdf(j))}{1/p-\cdf(j)}$,
+which experiences its maximum at
+$\frac{1-\sqrt{1-p}}{p}$ (thanks wolframalpha).
+But this is worst-case thinking, which doesn't seem appropriate.
+Let's have some optimism!}
+But it's not clear that the amount of mass moved
+necessarily corresponds to reaching $\confidents$ the fastest.
+%It just sounds kinda nice?
+%What is the best threshold, perhaps as a function of $p$, and why?
+
+%expected moved entropy
+%	call cdfj 'z' for the threshold we're looking for
+%
+%moved entropy at j in pass case
+%	new area is z (1-p)/(1-pz)
+%	old area is z
+%	so:
+%	z - z R
+%	z (1-R)
+%
+%	z [ 1-pz - (1-p) ] / (1-pz)
+%	z [ p-pz ] / (1-pz)
+% moved entropy to the right of j, should be equal
+% 	new area is (1-z)R
+% 		is (1-z)/(1-pz)
+% 	old area is 1-z
+% 	moved area is (1-z)(R-1)
+% 		is (1-z)(pz)/(1-pz)
+% 		is (pz-pzz)/(1-pz) yup, checks out
+% 
+%
+%probability that it moves = p test passes
+%	1(1-z) + (1-p)z (that's B)
+%	= 1-z + z-pz
+%	= 1-pz (still B)
+%
+%so this term:
+%	(1-pz) (z) (p-pz) / (1-pz)
+%	= z(p-pz)
+%	= pz-pzz
+%
+%moved entropy at j in fail case
+%	new area is 1
+%	old area is z
+%	of course, all 1-z of that shit moves
+%
+%probability that it moves = p test fails
+%	pz
+%
+%so this term:
+%	pz(1-z)
+%	= pz-pzz
+%	(wow, that's the same?)
+%
+%total expected moved: 2pz-2pzz
+%
+%solve for derivative dz = 0 to find argmax z
+%	p - 2pz = 0
 
@@
 
-A PDF's information content is measured in {\it entropy}:
+A PDF's information content is measured in entropy:
 $\entropy = \sum_i \pdf(i)\mathsf{ln}(\pdf(i))$.
+Alone, this value is fairly meaningless, but to compare them,
+a spikier PDF has lower entropy than a flatter one.
 Indeed,
-in Determinism World the search terminates when $\entropy = 0$.
+when the search terminates in Determinism World, $\entropy = 0$.
 I thought for a while about how to link ``minimum expected entropy''
@@
 %minimizing the expected steps until
-$\confidents = \mathsf{max}_i \pdf(i) \ge 0.99999$,
+$\confidents = \mathsf{max}_i (\pdf(i)) \ge 0.99999$,
 but couldn't really formalize the idea.
@@
 and also not necessarily stable under the entropy measurement,
-since it doesn't care how the remaining $0.00001$ is distributed.
-%
-Also, a Very Smart Math Friend of mine was unable to find a simple closed form for
-the expected resulting $\entropy(\renorm{} \circ \pdf)$
-%\[
-%(\cdf(j) + (1-\cdf(j))(1-p))\entropy(\renorm \circ \pdf)
-%+
-%p(1-\cdf(j))\entropy(\renorm{\text{bug}} \circ \pdf)
-%\]
-or whatever (bleh, my notation is kinda reaching its limits here).
-%so how would you write a proof about it anyway.
-Nevertheless, it seems promising, and easy to implement.
+since it doesn't care how the remaining $0.00001$ is distributed.%
+\footnote{See {\tt fn test\_entropy\_stability()}.}
+
+This strategy is expensive to compute.
+Whereas bisecting at some fixed threshold of probability mass merely requires a $O(\mathsf{log}  n)$ query of the CDF,
+computing the expected entropy is $O(n)$ for {\it each} bisect point.
+A Very Smart Math Friend of mine \cite{primer} analyzed the easy case of the initial test,
+when the PDF is still uniform (i.e., $\forall i, \cdf(i) = i$),
+and found the closed form
+\[
+	\entropy_{j,0} = p \mathsf{ln}(j)
+	+
+	(1-pj) \mathsf{ln} (1-pj)
+	-
+	j(1-p) \mathsf{ln} (j-p)
+\]
+whose derivative,
+in his words,
+``looks like a giant mess that does not admit an analytic solution for $j$.''
+
+% even the easy case of the inn
+% %the expected resulting
+% $\mathsf{EV}(\entropy(\renorm{} \circ \pdf))$
+% %\[
+% %(\cdf(j) + (1-\cdf(j))(1-p))\entropy(\renorm \circ \pdf)
+% %+
+% %p(1-\cdf(j))\entropy(\renorm{\text{bug}} \circ \pdf)
+% %\]
+% or whatever (bleh, my notation is kinda reaching its limits here).
+% %so how would you write a proof about it anyway.
+% Nevertheless, it seems promising, and easy to implement.
+
+There is a ray of hope, however:
+thanks to the nondecreasing-PDF invariant I mentioned on the previous page,
+the expected entropy has a unique local minimum.%
+\footnote{See {\tt fn test\_expected\_entropy\_has\_unique\_local\_minimum()}.}
+Thus we can binary search for the global minimum, making this at least $O(n\mathsf{log}n)$
+instead of $O(n^2)$.
 
@@
 
-You can tell it's at this point I'm starting to get statistics fatigue.
+You can tell at this point I'm starting to get statistics fatigue.
 
@@
 
+% parameterize over 10
+% what should the human do when they think they're done? -> probe repeatedly at b-1
+% what should they do when
+
 Regardless, it's easy to
-
-%(
-% hmmm...
-
-% In Determinism World, the search terminates when
-
-% Let's think of bisecting at $c_{(a+b)/2}$, in Determinism World, 
-% as maximizing the expected information gained from either test result.
-% On PDFs, information gain is measured as
 
@@
 
-\cite{drowningpoint}
-\cite{epsilon}
+To save you some page-turning, here's the source code link again:
+\url{https://github.com/bblum/sigbovik/blob/master/bisect}.
+A {\tt BisectStrategy}
+implements a function over $p$, the PDF, and/or its own internal state
+to choose the next bisect point.
+The {\tt SimulationState} is initialized with fixed $n$ and $p$,
+and repeatedly invokes a given {\tt BisectStrategy},
+hitting the PDF with Bayes's rule accordingly,
+until $\confidents \ge 0.99999$.
+I provide implementations for all the strategies from the previous section.
+It's written in Rust so it's self-documenting.
+
+I learned a lot about floating point.
+I don't mean like NaNs and mantissa bits,
+I mean that when your problem is fundamentally a long multiplication sequence of increasingly high rationals,
+your precious $\cdf(n)$ inevitably drifts ever farther from 1, and you start drowning in imprecision \cite{drowningpoint}.
+I suppose I could have used the {\tt BigRational} library,
+but I chose Rust over Haskell for this so I could {\it avoid} runaway memory consumption,
+so I resigned myself to explicitly renormalizing my PDFs by $1/\cdf(n)$ each time I updated them.
+After that, I was able to write some assertions
+to bound the imprecision drift within $k(\text{\tt std::f64::EPSILON})$ \cite{epsilon}.
+But yeah, I definitely spent some sanity points writing a function named {\tt fn assert\_kinda\_equals\_one(\&self)}.
 
@@
 
-% weigh cost of running test suite vs cost of git checkout -- mistrustful() gets a whole lot better! but, where is the break even point when you should git checkout?
-
-% p not known, figur eit out on the fly
+I'm not ashamed to admit I left a lot of loose ends in this paper.
+The biggest open questions for me are:
+\begin{enumerate}
+	\item What relates $p$ to {\sf cdfbisect}'s optimal argument?
+	\item Why is {\sf entropy} {\it so tantalizingly close} to {\sf cdfbisect}? Is there some way to show they're the same?
+\end{enumerate}
+I have also prepared a set of backup questions in case those aren't enough for you:
+\begin{enumerate}
+	\setcounter{enumi}{3}
+	\item If you don't know $p$ in advance, is there an efficient way to multiplex the search with figuring it out on the fly?
+	\item What if false positives are also possible, with some $q \ne p$?
+	\item How do you account for the cost of {\tt git checkout} being nonzero (i.e., having to recompile your tests on new code)?
+		Clearly {\sf human-mistrustful} becomes better. Where is the break-even point?
+\end{enumerate}
 
@@
 \section{conclusion}
+
+If you are a computer, you can do pretty well by bisecting the probability mass somewhere between 1/3 and 1/2.
+%
+If you are a human, you should forget everything you know as soon as you see new evidence that contradicts your priors.
+
+And remember to be patient and kind. Your test suite is just trying its best!
 

### 2020-03-30 bisect/paper/paper.tex
finish writing body of paper

@@
 \usepackage[nameinlink,nosort]{cleveref}
+\usepackage{mathspec}
 \usepackage{xeCJK}
@@
 
+\renewcommand{\figurename}{figure}
+
+\crefformat{section}{#2{section~#1}#3}
+\Crefformat{section}{#2{Section~#1}#3}
+\Crefformat{figure}{#2{Figure~#1}#3}
+\crefformat{figure}{#2{figure~#1}#3}
+%\Crefformat{Figure}{#2{figure~#1}#3}
+
 \begin{document}
@@
 
-\keywords pdflatex
+\keywords pdflatex, cdflatex
 % bracket, groove, in, jumps, the
@@
 with some confidence $\confidents$.
-Now evaluate your strategy by the expected number of steps to achieve, let's say for tradition's sake, $\confidents \ge 0.99999$.
+Now evaluate your strategy by the expected number of steps to achieve, let's say for tradition's sake, $\confidents > 0.99999$.
 Is it still optimal to bisect at the midpoint?%
@@
 When the bug reproduces at some commit $b$,
-you're certain no $c_{i>b}$ introduced the bug, so each $\pdf(i>b)$ becomes $0$,
-and each $\pdf(i\le b)$ renormalizes by a factor of $n/b$, to preserve the $\int \pdf(i) \text{d} i = 1$ invariant.%
+you're certain no $c_{i>b}$ introduced the bug, so each $\pdf(i>b) \leftarrow 0$
+and each $\pdf(i\le b)$ renormalizes by a factor of $n/b$, to preserve the
+%$\int \pdf(i) \text{d} i = 1$
+$\sum_i \pdf(i) = 1$
+invariant.%
 \footnote{Implemented as {\tt fn adjust\_pdf\_bug\_repros()} in the simulator.}
 
-In the deterministic case ($p=1$),
-the symmetric thing happens when the test passes at some $c_j$: $\pdf(i\le j) = 0$.
+In the deterministic case, $p=1$,
+the symmetric thing happens when the test passes at some $c_j$: $\pdf(i\le j) \leftarrow 0$.
 But when $p<1$, we must generalized it \cite{mario3}. Here's Bayes's rule:
@@
 Let's do a nice concrete example.
-Say $n=16$, the test passes at $j=7$, and then the bug repros at $j=11$.
+Say $n=16$, the test passes at $j=7$, and then fails at $j=11$.
 In the deterministic case, all the probability mass will be concentrated uniformly in the range $[8,11]$.
@@
 and we get probability mass scattered all the way down to $c_0$,
-as shown in \Cref{fig:example}(a).
+as shown in \cref{fig:example}(a).
 Yuck, someone clean that up!
@@
 Now let's say the test passes at $j=9$, then at $j=10$.
-\Cref{fig:example}(b) shows the updated PDFs/CDFs:
+\cref{fig:example}(b) shows the updated PDFs/CDFs:
 for $p=1$, this pinpoints the bug at $j=11$, and the search is over.
@@
 	\\
-		(a) After two tests.
+		(a) after two tests
 		\\
@@
 	\\
-		(b) After ``classical`` search terminates.
+		(b) after ``classical`` search terminates
 	\end{tabular}
 	%\caption{Example deterministic and nondeterministic PDFs and CDFs.}
-	\caption{Example \{,non\}deterministic \{P,C\}DFs.}
+	\caption{example \{,non\}deterministic \{P,C\}DFs}
 	\label{fig:example}
@@
 \section{prior work}
+\label{sec:prior}
 
@@
 Dropbox has a CI service called Athena \cite{athena}
-which automatically searches for flaky tests, when otherwise no bug exists.
-In this case the goal is to keep the build green, but if you consider the flaky test itself to be the bug, it's the same problem.%
+which automatically searches for flaky tests. %, when otherwise no bug exists.
+In this case the goal is to keep the build green,
+but if you consider the flaky test itself to be the bug, it's the same problem.%
 \footnote{Incidentally, the symmetric case
@@
 treating the combined result as ``basically as good as $p=1$'',
-and then runs an otherwise classical binary search.%
-\footnote{Upcoming, I will call this strategy \textsf{Human-Mistrustful}$(10)$.}
+and then runs an otherwise classical binary search.
+%\footnote{Upcoming, I will call this strategy \textsf{Human-Mistrustful}$(10)$.}
 In this setting, $p$ is not known in advance, so using Bayes's rule is off the table.
@@
 \section{strategies}
+\label{sec:strats}
 
@@
 Let's deconstruct the motives of classical binary search.
-Let $c_b$ be the earliest known buggy commit, and $c_a$ be the latest known safe commit.
+Let $c_b$ denote the earliest known buggy commit, and $c_a$ be the latest known safe commit.
 In Determinism World, bisecting at $c_{(a+b)/2}$
@@
 %because a pass is still progress even though $a$ doesn't ever budge,
-But in Nondeterminism World, $a$ doesn't ever budge from $0$,
+But in Nondeterminism World, $a$ doesn't ever budge from 0,
 so bisecting at $c_{(0+b)/2}$ will not even terminate.
@@
 \subsection{bisect probability mass}
+\label{sec:cdfbisect}
 
-Another way to think about Determinism World is that the $c_{(a+b)/2}$ bisect point
+Another way to frame it is that the $c_{(a+b)/2}$ bisect point
 % maximizes the expected probability mass that moves from one side to the other.
@@
 \footnote{I also worked out the {\it minimum} moved mass between the pass and fail case,
-which is almost always the pass case,
-and comes out to
+which is almost always the pass case.
+It comes out to
 %$\frac{p\cdf(j)(1-\cdf(j))}{1-p\cdf()}$
@@
 %What is the best threshold, perhaps as a function of $p$, and why?
+I show my work in \cref{fig:let-me-shown-that-my-work},
+because that's what they taught me to do in high school.
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+		\includegraphics[width=0.45\textwidth]{horses.jpg}
+	\end{tabular}
+	\caption{derivation(?) of optimal(??) bisect point}
+	\label{fig:let-me-shown-that-my-work}
+\end{figure}
 
@@
 \subsection{bisect entropy}
+\label{sec:entropy}
 
@@
 %minimizing the expected steps until
-$\confidents = \mathsf{max}_i (\pdf(i)) \ge 0.99999$,
+$\confidents = \mathsf{max}_i (\pdf(i)) > 0.99999$,
 but couldn't really formalize the idea.
@@
 when the PDF is still uniform (i.e., $\forall i, \cdf(i) = i$),
-and found the closed form
+and found the closed form:
 \[
-	\entropy_{j,0} = p \mathsf{ln}(j)
+	\entropy_{0,j} = p \mathsf{ln}(j)
 	+
-	(1-pj) \mathsf{ln} (1-pj)
+	(1-pj) (\mathsf{ln} (1-pj))
 	-
-	j(1-p) \mathsf{ln} (j-p)
+	j(1-p) (\mathsf{ln} (j-p))
 \]
@@
 \subsection{bisect randomly}
+\label{sec:random}
 
@@
 You can tell at this point I'm starting to get statistics fatigue.
+Not unprecedented in these hallowed proceedings \cite{batch-myass}.
 
 \subsection{human}
+\label{sec:human}
 
-Humans are known to be less patient than computers (citation: see previous sentence).
+Humans are known to be less patient than computers (citation: \cref{sec:random}).
 If a human is unwilling to compute $\renorm{} \circ \pdf$ by hand every step,
@@
 what should they do?
+Let's say a ``human strategy'' is any that doesn't use the PDF/CDF when deciding where to bisect.%
+%i.e., one that's easy for a human to do without bothering with Bayes's rule.%
+\footnote{In fact {\sf rand-uniform} from last section counts as a human strategy. Did you bring your dice?}
 
-Think back to Athena from the prior work section.
-The human
+The most straightforward thing for a human to do is to assume $p=1$ until they encounter evidence to the contrary.
+They'll conduct a normal binary search, and when they think they've found the bug at $c_b$,
+they'll just keep probing $c_{b-1}$ until enough probability mass (invisible to them)
+moves to $b$ and the simulator says they can stop.
+If this instead repros the bug at $c_{b-1}$ after all, the human gets confused!
+They'll assume the bug might have been present as early as $c_0$,
+forgets their lower bound, and begins binary searching anew in the range $[0,b-1]$.
+It's easy to see this makes progress.
+Let's call this the ``confused, forgetful human'' strategy.
 
-% parameterize over 10
-% what should the human do when they think they're done? -> probe repeatedly at b-1
-% what should they do when
+Another confused human variant is for them to remember their previous lower bounds.
+After all, just because their best lower bound $b-1$ was contradicted
+doesn't say anything about some earlier bound $a<b-1$ they might have observed.
+So this human will maintain a stack of lower bounds, and when $b-1$ is contradicted,
+they'll pop off the stack and resume searching in $[a,b-1]$ instead.
+Let's call this the ``confused human who remembers''.
+%
+Incidentally, while implementing this strategy, I accidentally wrote the comment,
+\begin{center}
+	\begin{tabular}{l}
+		{\tt // the human, who now has perfect memory,} \\
+		{\tt // walks backwards in time}
+	\end{tabular}
+\end{center}
+and giggled a lot to myself as I imagined {\birbaslo sigbovik}~fanfiction.
 
-Regardless, it's easy to
+Finally, let's imagine the human knows maybe something is up with this $p$ stuff in advance. %before they start searching.
+Wishing to carry on classically as though $p$ were 1,
+they'll try to emulate $p$ being higher by retrying any passing test ``just to be sure'', before moving on.
+If they retry the test $r$ times,
+the probability it fails given the bug is present is then $1-(1-p)^{1+r}$.
+%In general, for $r$ retries, that's $1-(1-p)^{r+1}$.
+But it comes at the cost of $r$ more steps for each new lower bound!
+As before, when this human thinks they're done, they'll repeatedly probe $c_{b-1}$ until contradicted
+(and we'll make them forgetful, to keep things simple).
+Let's call this the ``suspicious human of $r$ retries''.
 
@@
 
-\section{simulating it}
+\section{simulated it}
 
@@
 hitting the PDF with Bayes's rule accordingly,
-until $\confidents \ge 0.99999$.
-I provide implementations for all the strategies from the previous section.
+until $\confidents > 0.99999$.
+I provide implementations for all the strategies from the previous section in {\tt src/strategies/}.
 It's written in Rust so it's self-documenting.
@@
 I don't mean like NaNs and mantissa bits,
-I mean that when your problem is fundamentally a long multiplication sequence of increasingly high rationals,
-your precious $\cdf(n)$ inevitably drifts ever farther from 1, and you start drowning in imprecision \cite{drowningpoint}.
+I mean that when your project is fundamentally a long multiplication chain of increasingly high rationals,
+your precious $\cdf(n)$ inevitably drifts farther from 1 than you can bound with any arbitrary $\epsilon$, and you start drowning in imprecision \cite{drowningpoint}.
 I suppose I could have used the {\tt BigRational} library,
-but I chose Rust over Haskell for this so I could {\it avoid} runaway memory consumption,
+but I chose Rust over Haskell for this so I could {\it avoid} runaway memory consumption...
 so I resigned myself to explicitly renormalizing my PDFs by $1/\cdf(n)$ each time I updated them.
@@
 
-\section{evaluation}
+\section{experiments}
+
+Let's get right to the good stuff.
+\cref{fig:resulce} shows the overall results,
+plotting the average performance of each of the strategies from \cref{sec:strats}.
+
+\begin{figure}[ht]
+	\begin{tabular}{c}
+		\includegraphics[width=0.45\textwidth]{resulce.pdf}
+		\\
+		(a) full view
+		\\
+		\\
+		\includegraphics[width=0.45\textwidth]{resulce-detail.pdf}
+		\\
+		(b) detail view of high $p$ range; linear scale
+	\end{tabular}
+	\caption{perf results.
+	how many bisect steps each strategy takes to reach five 9s of confidence} %,
+	%simulated on a 1024-long {\tt git} {\tt log}; 65536 trials}
+	\label{fig:resulce}
+\end{figure}
+
+Each data point is simulated with $n=1024$ and averaged over 65536 trials.
+I chose a power of two for $n$ to minimize floating point drift (handwave),
+and a multiple of $n$ for trials so that {\tt buggy\_commit = trial \% n} would be uniform.
+
+Of course, each trial takes exponentially
+%\footnote{Yes, I am using this word in the literal sense.}
+longer and longer as $p \rightarrow 0$, so I didn't bother testing past 0.1.
+I also didn't bother measuring execution time,
+as prior work has shown that one's personal laptop is a fraught environment for stable performance evaluation \cite{htm},
+but it was still very obvious that {\sf entropy} was far, far slower per step than all other strategies.
+%I guess I'd say it took minutes to run {\sf cdfbisect} up to $p=0.01$, and hours/overnight for the same for {\sf entropy}.
+
+\subsection{computer strategies}
+
+Minimizing expected entropy
+%({\sf entropy}; \Cref{sec:entropy})
+turned out to be the best strategy, globally across all $p$.
+Bisecting probability mass
+%({\sf cdfbisect}; \Cref{sec:cdf})
+turns out to be preeetty close,
+although its performance varies depending on its bisect-point parameter $j$.
+Obviously, when $p=1$, $j$ should be 0.5 to reproduce classical binary search,
+but at around $p=0.8$, $j=0.333$ overtakes it.
+Ultimately at $p=0.1$, the former is about 4\% slower compared to the latter and to entropy,
+contradicting the ``independent of $p$'' hypothesis from \cref{sec:cdfbisect}.
+I investigated a little more to see how $p$ affected the optimal $j$, testing various $j$s in increments of 0.01:
+\begin{center}
+	\begin{tabular}{c|cc}
+		$p$ & best $j$ & {\sf cdfbisect}$(j)$ \\
+		\hline
+		1.0 & 0.50 & 10     \\
+		0.9 & 0.45 & 17.4   \\
+		0.8 & 0.41 & 21.9   \\
+		0.7 & 0.39 & 27.5   \\
+		0.6 & 0.40 & 34.6   \\
+		0.5 & 0.36 & 44.1   \\
+		0.4 & 0.35 & 58.3   \\
+		0.3 & 0.34 & 81.6   \\
+		0.2 & 0.36 & 127.9  \\
+		0.1 & 0.35 & 266.6  \\
+	\end{tabular}
+	%\begin{tabular}{l|c ccccc ccccc}
+	%	$p$				& 1 & 0.9 & 0.8 & 0.7 & 0.6
+	%						& 0.5 & 0.4 & 0.3 & 0.2 & 0.1 \\
+	%	\hline
+	%	best $j$			& 0.5 & 0.45 & 0.41 & 0.39 & 0.40
+	%						& 0.36 & 0.35 & 0.34 & 0.36 & 0.35 \\
+	%	best {\sf cdfbisect}$(j)$	& 10 & 17.4 & 21.9 & 27.5 & 34.6
+	%						& 44.1 & 58.3 & 81.6 & 127.9 & 266.6
+	%\end{tabular}
+\end{center}
+65536 trials here wasn't really enough to get variance under control,
+but it definitely seems to converge towards 0.333 or maybe 0.35 or something as $p \rightarrow 0$.
+Open question why for warp zone, I guess.
+
+{\sf rand-uniform} is obviously terrible but I threw it in there for kicks.
+{\sf rand-cdf} is not so terrible, doing even better than all the human strategies.
+
+\subsection{human strategies}
+
+I investigated how the suspicious human's number-of-retries parameter $r$ affects their performance.
+In this notation, $r=1$ means they retry once whenever the test passes, meaning two test runs on every passing commit.
+$r=0$ is equivalent to the forgetful strategy, and $r=9$ is what Athena does (\cref{sec:prior}).%
+% actually it's an enhanced version of athena so that it will terminate
+
+\cref{fig:resulce} was getting a little crowded so I've plotted all the human-friendly strategies on \cref{fig:human}.
+It's pretty self-explanatory imo.
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+		\includegraphics[width=0.45\textwidth]{resulce-human.pdf}
+	\end{tabular}
+	\caption{version of \cref{fig:resulce} for human strategies}
+	\label{fig:human}
+\end{figure}
+
+One thing that surprised me was that the human who forgets all their past known lower bounds when {\it any} of them is contradicted
+performed better than the human who tried
+to rewind to their last not-yet-contradicted lower bound.
+I guess there's a moral here about not clinging to the past, moving on with life, or something.
+
+\subsection{why five 9s?}
+
+Let's talk about the arbitrary confidence threshold.
+There's a lot of knobs to tweak here and only so much time til April 1 (not to mention reader patience),
+but I did at least want to show what happens if you terminate the search when you're only 50\% confident.
+Compared to \cref{fig:resulce},
+the strategies all terminate in about half as many steps, with the same relative ranking and break-even points.
+What's more interesting is the number of times they were wrong.
+
+I had the simulator refuse to count a wrong result, %towards the 65536 trial count,
+restarting the search until it correctly identified the buggy commit 65536 times,
+and counting the number of restarts.
+With $\confidents > 0.99999$ this was not very interesting:
+the strategies collectively made 0 errors 81\% of the time, and were wrong more than twice only 0.6\%.
+But $\confidents > 0.5$ was a different story.
+As \cref{fig:wrongs} shows, most strategies
+-- even the comically slow {\sf rand-uniform} --
+converged to being wrong about half the time, i.e., 65536 wrongs per 65536 rights.
+{\sf athena} was most resilient, at least for higher $p$.
+
+\begin{figure}[t]
+	\begin{tabular}{c}
+		\includegraphics[width=0.45\textwidth]{wrongs.pdf}
+	\end{tabular}
+	\caption{how often each strategy was wrong, when terminating with only 50\% confidence.
+	the dashed line marks the approximate convergence point at 65536, equal to the number of right answers}
+	\label{fig:wrongs}
+\end{figure}
+
+{\sf confused-human-who-remembers}, alone, did not converge.
+They became more and more unable to cope with the repro rate approaching 0,
+utlimately being wrong 8 times as often as they were right.
+Oh, human... please learn to let go and love yourself for who you are.
+% japanese speakers should be able to hear the untranslated "kagayaiteru" here
+There's a better world out there, a sparkling radiant future waiting just for you!
 
@@
 \begin{enumerate}
-	\setcounter{enumi}{3}
+	\setcounter{enumi}{3} % intentional
 	\item If you don't know $p$ in advance, is there an efficient way to multiplex the search with figuring it out on the fly?
-	\item What if false positives are also possible, with some $q \ne p$?
-	\item How do you account for the cost of {\tt git checkout} being nonzero (i.e., having to recompile your tests on new code)?
-		Clearly {\sf human-mistrustful} becomes better. Where is the break-even point?
+	\item What if false positives are also possible, with some flake rate $q \ne p$?
+		% TODO make this convey the sigbovik spirit more
+	\item What if the cost of {\tt git} {\tt checkout} is nonzero (e.g., having to recompile your tests on new code)?
+		Clearly {\sf human-mistrustful} becomes better.
+		% Where is the break-even point?
+		What about some wacky hybrid {\sf cdfbisect-mistrustful}?
 \end{enumerate}
@@
 
+\setcounter{section}{6} % wee
 \section{conclusion}
@@
 
-\renewcommand{\refname}{references}
+\renewcommand{\refname}{referents}
 \bibliographystyle{abbrvnat}

### 2020-03-30 bisect/paper/paper.tex
finish this dang bisect project

@@
 
-\keywords pdflatex, cdflatex
+\keywords pdflatex, cdflatex, epsilon
 % bracket, groove, in, jumps, the
 
-\newcommand\confidents{\ensuremath{\mathcal{Z}}\xspace}
+\newcommand\confidents{\ensuremath{z}\xspace}
 \newcommand\pdf{\ensuremath{\mathsf{pdf}}\xspace}
@@
 If you're feeling adventurous,
-implement a strategy and throw it in this simulator I implemented to see how it stacks up:
+implement a strategy and throw it in this simulator I made to see how it fares:
 \url{https://github.com/bblum/sigbovik/blob/master/bisect}.
@@
 \section{pdfs, but not the portable document format kind, and our friend rev. bayes}
+\label{sec:pdfs}
 
@@
 In fact, from here it takes 18 further probes at $j=10$ until we are at least five 9's confident that $c_{11}$ is the culprit.%
-\footnote{See {\tt fn test\_figure\_1()}!}
+\footnote{See {\tt fn test\_figure\_1()}.}
 Bayes's rule gonna get ya.
@@
 is also the same problem.}
-Athena's ``deflakes'' the test at each commit by running 10 times,
+Athena ``deflakes'' the test at each commit by running it 10 times,
 treating the combined result as ``basically as good as $p=1$'',
@@
 % we ignore the range length, and instead bisect at the midpoint of probability mass,
-i.e., $\mathsf{max}_j (i)$ s.t. $\cdf(j) \le 0.5)$.
+i.e., $\mathsf{max}_j (j$; $\cdf(j) \le 0.5)$.
 This approach
 fits the ``binary search'' spirit,
-and will terminate in Nondeterminism World,
-i.e.,
+and will also terminate in Nondeterminism World:
+%i.e.,
 if $b$ is the solution, it converges to repeatedly probing $b-1$,
@@
 I show my work in \cref{fig:let-me-shown-that-my-work},
-because that's what they taught me to do in high school.
+because that's what they taught me to do in high school algebra class.
 
@@
 There is a ray of hope, however:
-thanks to the nondecreasing-PDF invariant I mentioned on the previous page,
+thanks to the nondecreasing-PDF invariant I mentioned in \cref{sec:pdfs}, %on the previous page,
 the expected entropy has a unique local minimum.%
 \footnote{See {\tt fn test\_expected\_entropy\_has\_unique\_local\_minimum()}.}
-Thus we can binary search for the global minimum, making this at least $O(n\mathsf{log}n)$
+Thus we can binary search for the global minimum, making this at worst $O(n\mathsf{log}n)$
 instead of $O(n^2)$.
@@
 They'll assume the bug might have been present as early as $c_0$,
-forgets their lower bound, and begins binary searching anew in the range $[0,b-1]$.
+forgets their lower bound, and begins binary searching anew in the range $[0,b)$.
 It's easy to see this makes progress.
@@
 So this human will maintain a stack of lower bounds, and when $b-1$ is contradicted,
-they'll pop off the stack and resume searching in $[a,b-1]$ instead.
+they'll pop off the stack and resume searching in $[a,b)$ instead.
 Let's call this the ``confused human who remembers''.
@@
 
-Finally, let's imagine the human knows maybe something is up with this $p$ stuff in advance. %before they start searching.
+Finally, let's imagine the human knows in advance that maybe something is up with this $p$ stuff.
+%in advance. %before they start searching.
 Wishing to carry on classically as though $p$ were 1,
@@
 until $\confidents > 0.99999$.
-I provide implementations for all the strategies from the previous section in {\tt src/strategies/}.
+I provide implementations for all \cref{sec:strats}'s strategies in {\tt src/strategies/}.
 It's written in Rust so it's self-documenting.
@@
 but I chose Rust over Haskell for this so I could {\it avoid} runaway memory consumption...
-so I resigned myself to explicitly renormalizing my PDFs by $1/\cdf(n)$ each time I updated them.
+so, I resigned myself to explicitly renormalizing my PDFs by $1/\cdf(n)$ each time I updated them.
 After that, I was able to write some assertions
@@
 %I guess I'd say it took minutes to run {\sf cdfbisect} up to $p=0.01$, and hours/overnight for the same for {\sf entropy}.
+It was even slower than {\sf rand-uniform}!
 
@@
 
-\cref{fig:resulce} was getting a little crowded so I've plotted all the human-friendly strategies on \cref{fig:human}.
+\cref{fig:resulce} was getting a little crowded so I've plotted all the human-friendly strategies separately on \cref{fig:human}.
 It's pretty self-explanatory imo.
@@
 the strategies collectively made 0 errors 81\% of the time, and were wrong more than twice only 0.6\%.
-But $\confidents > 0.5$ was a different story.
+But $\confidents > 0.5$ is a different story.
 As \cref{fig:wrongs} shows, most strategies
@@
 converged to being wrong about half the time, i.e., 65536 wrongs per 65536 rights.
-{\sf athena} was most resilient, at least for higher $p$.
+For higher $p$,
+{\sf athena} was most resilient.
 

### 2021-01-10 formina/read-jmdict/main.rs
get started on formina

@@ -1,0 +1,110 @@
+use kanji::is_katakana;
+
+use std::env::args;
+use std::fs::File;
+use std::io::BufReader;
+
+use xml::reader::{EventReader, XmlEvent};
+
+// SIGBOVIK logic
+
+fn process_word(readings: &[String], meanings: &[String]) {
+    // TODO: do the thing w/essential consonants
+    // Probably good to strip out parentheses in the definitions
+    // TODO: allow morphological suffixes to be dropped, eg "ソロキャンプ" -> "solo campING"
+    // This probably requires a special case whitelist otws it makes the essential-char logic
+    // vacuous lol
+    println!("word: {:?} => {:?}", readings, meanings);
+}
+
+// Misc. japanese stuff
+
+fn is_katakana_word(word: &str) -> bool {
+    let mut nonempty = false;
+    // TODO: Might need to add stuff like wi, we, hiragana repetition marks?
+    // Otoh, essential-consonants filter logic might subsume that.
+    for c in word.chars() {
+        nonempty = true;
+        // This just tests if it's between u30A0 and u30FF.
+        if !is_katakana(&c) {
+            return false;
+        }
+    }
+    nonempty
+}
+
+// XML stuff
+
+fn expect_chars(parser: &mut impl Iterator<Item = xml::reader::Result<XmlEvent>>, msg: &str) -> String {
+    match parser.next().expect("xml stream ended").expect("parse xml") {
+        XmlEvent::Characters(s) => s,
+        e => panic!("expected characters, got {:?}, {}", e, msg),
+    }
+}
+
+fn main() {
+    let filename = args().skip(1).next().expect("expected a filename as first arg");
+    let file = File::open(filename).expect("open file");
+    // let mut contents = String::new();
+    // file.read_to_string(&mut contents).expect("read file");
+    let mut parser = EventReader::new(BufReader::new(file)).into_iter();
+
+    let mut cur_word_rebs = vec![];
+    let mut cur_word_meanings = vec![];
+    let mut cur_word_katakana_only = true;
+
+    while let Some(e) = parser.next() {
+        match e {
+            Ok(XmlEvent::StartElement { name, .. }) => {
+                if name.local_name == "entry" {
+                    assert!(cur_word_rebs.is_empty(), "nested entry?");
+                } else if name.local_name == "reb" {
+                    // Phonetic reading for the word.
+                    let reb = expect_chars(&mut parser, "reb text");
+                    if is_katakana_word(&reb) {
[…]

### 2021-01-10 bisect/src/main.rs
remaining bisect change

@@ -25,21 +25,26 @@ fn run_trial<S: BisectStrategy>(strat_fn: impl Fn(&SimulationState) -> S) {
 
     let n = 1024;
     let conf_thresh = 0.99999;
+    let conf_thresh = 0.50001;
     let trials = 65536;
+    let trials = 16384;
 
-    output(&mut f, format!("==== searching for best bisect_poince~~ ===="));
     output(&mut f, format!("==== {}; n = {}, confidence goal = {}, trials = {} ====",
                            strat_name, n, conf_thresh, trials));
 
     for &false_negative_prob in &[
-        0.0,
-        0.00001, 0.0001, 0.001, 0.01,
-        0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9,
-        0.99, 0.999, 0.9999, 0.99999, // these will take FOREVER
+        0.9,
+        // 0.0,
+        // 0.00001, 0.0001, 0.001, 0.01,
+        // 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9,
+        // 0.99, // 0.999, 0.9999, 0.99999, // these will take FOREVER
     ] {
         let mut result_steps = vec![];
         let mut wrongs = 0;
         for i in 0..trials {
+            if i % 1024 == 0 {
+                println!("making progress; trial {}", i);
+            }
             'inner: loop {
                 // TODO: time these suckas
                 let s = SimulationState::new_with_bug_at(n, false_negative_prob, i % n);
@@ -78,13 +83,15 @@ fn main() {
         run_trial(LinearSearch::new);
         run_trial(|s| NaiveBinarySearch::new(s, ConfusedHumanMode::UsePreviousLow));
         run_trial(|s| NaiveBinarySearch::new(s, ConfusedHumanMode::ForgetEverything));
+        run_trial(|s| NaiveBinarySearch::new(s, ConfusedHumanMode::Mistrustful(0)));
         run_trial(|s| ChooseRandomly::new(s, RandomMode::Uniformly));
         run_trial(|s| ChooseRandomly::new(s, RandomMode::WeightedByCDF));
     }
 
+    run_trial(|s| NaiveBinarySearch::new(s, ConfusedHumanMode::UsePreviousLow));
     // run_trial(|s| ChooseRandomly::new(s, RandomMode::Uniformly));
[…]

### 2021-01-10 bisect/src/sim.rs

@@ -272,7 +272,17 @@ impl SimulationState {
             let prev_pdf = self.pdf.clone();
 
             let commit = strat.select_commit(&self);
-            let result = self.simulate_step(commit);
+            let result = if self.false_negative_rate == 0.0 && commit < self.lower_bound {
+                let result = BisectAttempt {
+                    commit,
+                    bug_repros: false,
+                };
+                // count it against mistrustful's number of steps!
+                self.history.push(result);
+                result
+            } else {
+                self.simulate_step(commit)
+            };
             strat.notify_result(result);
 
             if self.pdf[commit] < EPSILON {

### 2021-01-10 bisect/src/strategies/naive.rs

@@ -40,7 +40,8 @@ impl NaiveBinarySearch {
                     assert_ne!(max_retries, 0, "just use ForgetEverything in this case");
                     if s.false_negative_rate == 0.0 {
                         // don't bother retrying commits in this mode, obviously
-                        Memory::BlissfulIgnorance
+                        // Memory::BlissfulIgnorance
+                        Memory::Retry { max_retries, active_retry: None }
                     } else {
                         Memory::Retry { max_retries, active_retry: None }
                     }

### 2023-03-27 seitan/seitan.tex
begin writing about seitan

\documentclass[12pt]{article}

\usepackage[charter]{mathdesign}

\setlength{\parskip}{0.5em}

\newcommand\chapter[2]{{\thispagestyle{empty} \Large \sc #1. \quad #2

\vspace{1.5em}}}

\begin{document}

\chapter{1}{Homesick}

It's summer 2020, and you are the city of Seattle PNW USA.

Within you, in upper Queen Anne, lives this story's protagonist.
They recently moved into you, far too recently to have deep enough social roots for the epidemiological circumstances of the year in question.
Queen Anne is a beautiful neighborhood,
full of cherry blossom trees stunning in the spring,
radio towers that one can stand at the base of and crane their neck to see the top in a dizzying twist of perspective
(except during wildfire season, when, incidentally, one's newly-acquired KN95s find a secondary use),
and artfully manicured flower gardens afore multimillion-dollar homes,
which make our protagonist feel keenly alienated in their 800sqft apartment.
The protagonist copes with quarantine, in part, by attending ThursDz, now occurring virtually,
and they begin to long for a particular food of home.
%
Said home, Pittsburgh Appalachia USA, is famous for many foods,
but they long in particular for a less famous one: the humble seitan wing.

You, city of Seattle, do not know what a seitan wing is.

Seitan wings is quite simple, in concept: Take chicken wings, ideally buffalo-sauced with a side of blue cheese dip, and replace the meat with seasoned dough of vital wheat gluten.
You are no stranger to seitan itself, of course, boasting a bustling vegan scene, including at least two Thai joints whose menus' every single meat entr\'ee is actually seitan.
But--as our protagonist will only later theorize--the preparation with buffalo sauce is an innovation unique to a city where the hipster foodie and blue-collar bar food scenes intersect.
Presently, said theory yet an inkling in their mind, they search countless delivery menus.
Their final attempt, upon recommendation of a CMU acquaintance they thought understood their struggle, a place from Capitol Hill;
whose seitan ``appetizer'' turns out to be a long, tough stick most akin to a Sl*m Jim, tossed in a thick coating of bread crumbs more voluminous than the inner stick jimself.

Resignedly, the protagonist asks Google Trends what, exactly, the fuck. Google-{\em sama} replies with a color-coded map of interest by US state: only Pennsylvania, New York, and Colorado are shaded at all.

You, city of Seattle, are invited to pause reading, issue the same Google Trends query, and reflect upon your sins.

\chapter{2}{Secrets}

It's fall 2020, and you are Google's search results for ``seitan wings recipe''.

% TODO

\chapter{3}{Reprieve}

It's early June 2021, and you are M*d M*x Big Burrito Restaurant Group restaurant.

You serve the canonical implementation of seitan wings in Shadyside Pittsburgh Appalachia USA. If the ANSI standards committee filed an entry for the food, they would surely enshrine one of yours.

% TODO

\chapter{4}{Texture}

It's late June 2021, and you are the ``Seitan Reloaded'' recipe posted by author {\sf @simrob} on Blogging Website M*dium Dot Com.

% TODO

\newpage \thispagestyle{empty}

\chapter{5}{Wing}

It's August 2021, and you are visiting your parents for the first time since quarantine in Santa Barbara Socal USA.

%It's the first time you've seen them since quarantine, and
You have a lot to catch up on.
Where there was border collie Louie, there is now border collie Junior.
You are teaching your mom to play Heaven's Vault,
your new favorite conlang-learning archaeology VN,
trying to tell her that the helplessness the game makes her feel is in fact diagetic,
and she (also a language nerd) is kinda actually getting it.
%
Your psychological association is still tenuous,
because your gender is new to them and your body is new to all three of you,
but your parents will get there someday, beyond the scope of this work.

You tell them the preceding 4 chapters,
and they agree to share with you the honor of putting the pieces together,
and that is the story of how seitan wings were made correctly in Santa Barbara for the first time ever.

\vspace{24em}

\noindent You acknowledge Will Scott, William Lovas, Rob Simmons, Chris Martens, Eve Blum, David Blum, and Jenny Lin for their invaluable roles in the story.

\newpage \thispagestyle{empty}

\begin{center}
	% {\sc \Large ben blum's seitan wings}
	{\sc \Large trans rights seitan wings}
\end{center}

Based on Chapter 4's viewpoint character.
You will need a steamer and either an immersion blender or a lot of wrist stamina.
Makes 4-6 servings.

\vspace{1em}

\noindent \begin{tabular}{p{0.95\textwidth}}
	{\bf Ingredients (Wing; dry)} \\
	\hline
	{\em 14 oz} -- Vital wheat gluten \\
	{\em 1 oz} -- Nutritional yeast (or parmesan) \\
	{\em 2 tsp} -- Salt \\
	{\em 3 Tbsp} -- Spices: rosemary, thyme, onion or garlic powder, and white pepper \\
	\\
	{\bf Ingredients (Wing; wet)} \\
	\hline
	{\em 12 oz} -- Water (or chicken stock, for irony) \\
	{\em 6 oz} -- Tomato paste \\
	{\em 2 oz} -- Olive oil \\
	\\
	{\bf Ingredients (Buffalo)} \\
	\hline
	{\em 2 Tbsp} -- Butter (omit for vegans) \\
	{\em 2 Tbsp} -- Hot sauce (Frank's is canonical, but they prefer Cholula) \\
	\\
	{\bf Ingredients (Bluech)} \\
	\hline
	{\em 1/4 cup} -- Blue cheese (a nice, creamy pungent one from Penn Mac) \\
	{\em 1/4 cup} -- Buttermilk (you can substitute but it really makes a difference) \\
	{\em Spoonful} -- Plain greek yogurt or sour cream, for consistency, optional \\
	\\
	{\bf Method} \\
	\hline
\setlength{\parskip}{0.5em}
	In two bowls, separately mix the dry and wet wingredients. Combine and knead for about 5 minutes, until coherent but also tears easily. Add more liquid or gluten as necessary, a little at a time.

	Tear by hand (scraggly is good) into 1" chunks and steam for 30 minutes.

	Meanwhile, chongle the blue cheese dip: Reserve half the cheese, and immersion blend everything else. Add reserved cheese and mash together with a fork until pleasantly chunky.

	Melt butter and then mix in hot sauce. Do not microwave your hot sauce.

	Fry steamed wings in neutral oil at 350$^{\circ}$F for 3-5 minutes, until just crispy. Pat dry with a paper towel, toss in sauce, and serve over math puzzles.
\end{tabular}

\end{document}

### 2023-03-27 seitan/seitan.tex
ch2

@@
 which make our protagonist feel keenly alienated in their 800sqft apartment.
-The protagonist copes with quarantine, in part, by attending ThursDz, now occurring virtually,
+They cope with quarantine, in part, by attending the now-virtual ThursDz,
+%, now occurring virtually,
 and they begin to long for a particular food of home.
@@
 
+\newpage \thispagestyle{empty}
+
 \chapter{2}{Secrets}
@@
 It's fall 2020, and you are Google's search results for ``seitan wings recipe''.
+You know what a seitan wing is.
 
-% TODO
+You profess to contain one and a half million results.
+The protagonist doesn't have quite that much time on their hands,
+but they do have a lot, so they try (or at least read) everything on your first page.
+You don't seem to agree with yourself very much, they notice:
+you sometimes suggest breading the wing (which the protagonist knows to be wrong),
+you sometimes suggest BBQ or garlic sauce instead of buffalo (which is noncanonical, but fine),
+but most bafflingly, you suggest as many different ways of cooking the seitan as there are recipes.
+Most often, you suggest to boil it,
+which the protagonist gamely tries,
+all the while remembering a story they heard long ago about how their mom's dad was so bad a cook she once caught him boiling a steak.
+It sucks.
+You say they should boil it then bake it;
+you say they should boil it then saut\'e it;
+you say they should saut\'e it then boil it;
+you say they should fry it;
+you say they should not just fry it, because it will leave the inside raw and gummy (this, for once, is true),
+you say they should bake it;
+you say they should bake it in broth;
+you say they should bake it dry tossed in oil.
+
+Among these, the last idea at least produces something both edible and texturally close to the crispy exterior the protagonist remembers, but it is only ``close'' like standing on the next tallest mountain peak and looking across the way at the one you {\em meant} to climb is ``close''.
+Everything you say is touched by the spectre of health food writing,
+because your reader must be into \textasciitilde{}vegan alternatives\textasciitilde{}.
+Drowning in your million useless results the protagonist knows one truth deep in their heart: Seitan wings is not a health food.
+
+Months pass, and the protagonist does not talk to you anymore.
+They play Breath of the Wild in Japanese (which had long been a goal of theirs),
+they play lots of gay VNs on itch dot io,
+they practice Touhou music on the piano,
+%they play Touhou,
+and they play Netrunner worlds online over an enchanting coq au vin their elderly neighbor Ellen prepared for their nascent no-contact food relationship.
+
+It's December 25, and thinking of you again, they pose your unanswered mystery to Tw*tter.
+A Seattle friend replies, at long last, with the secret:
+
+Steam, then fry.
+
+\newpage \thispagestyle{empty}
 
@@
 
-\chapter{4}{Texture}
+\newpage \thispagestyle{empty}
+
+\chapter{4}{Tenderness}
 
@@
 	{\em 1 oz} -- Nutritional yeast (or parmesan) \\
+	{\em 3 Tbsp} -- Spices: rosemary, thyme, onion or garlic powder, and white pepper \\
 	{\em 2 tsp} -- Salt \\
-	{\em 3 Tbsp} -- Spices: rosemary, thyme, onion or garlic powder, and white pepper \\
 	\\
@@
 	\hline
-	{\em 1/4 cup} -- Blue cheese (a nice, creamy pungent one from Penn Mac) \\
-	{\em 1/4 cup} -- Buttermilk (you can substitute but it really makes a difference) \\
+	{\em 1/3 cup} -- Blue cheese (a nice, creamy pungent one from Penn Mac) \\
+	{\em 1/3 cup} -- Buttermilk (you can substitute but it really makes a difference) \\
 	{\em Spoonful} -- Plain greek yogurt or sour cream, for consistency, optional \\

### 2023-03-27 seitan/seitan.tex
finished

@@
 You are no stranger to seitan itself, of course, boasting a bustling vegan scene, including at least two Thai joints whose menus' every single meat entr\'ee is actually seitan.
-But--as our protagonist will only later theorize--the preparation with buffalo sauce is an innovation unique to a city where the hipster foodie and blue-collar bar food scenes intersect.
+But -- as our protagonist will only later theorize -- the preparation with buffalo sauce is an innovation unique to a city where the hipster foodie and blue-collar bar food scenes intersect.
 Presently, said theory yet an inkling in their mind, they search countless delivery menus.
@@
 you say they should fry it;
-you say they should not just fry it, because it will leave the inside raw and gummy (this, for once, is true),
+you say they should not fry it, because it will leave the inside raw and gummy (this, for once, is true),
 you say they should bake it;
@@
 %they play Touhou,
-and they play Netrunner worlds online over an enchanting coq au vin their elderly neighbor Ellen prepared for their nascent no-contact food relationship.
+and they play Netrunner worlds online over an enchanting coq au vin their elderly neighbor prepared for their nascent no-contact food relationship.
 
@@
 
-\chapter{3}{Reprieve}
+\chapter{3}{Respite}
 
-It's early June 2021, and you are M*d M*x Big Burrito Restaurant Group restaurant.
+It's early June 2021, and you are M*d Mex Big Burrito Restaurant Group restaurant.
+%
+In Shadyside Pittsburgh Appalachia USA
+you serve the canonical implementation of seitan wings.
+If the NIST standards committee filed an entry for the food, they would surely enshrine one of yours.
 
-You serve the canonical implementation of seitan wings in Shadyside Pittsburgh Appalachia USA. If the ANSI standards committee filed an entry for the food, they would surely enshrine one of yours.
+The protagonist, vaccinated, visits your city. Their city.
 
-% TODO
+You are one of the first stops they make,
+a high honor in the book of a foodie ex-grad student who knows the East End inside and out
+(friends in tow, of course).
+They have practiced Chapter 2's secret lost art, but something is still wrong,
+and sitting at your table in the summer heat they realize what it is:
+technique aside, even fully cooked, their dough recipe produces a chewy, glutinous inside,
+whereas your wings are perfectly soft inside;
+one's teeth pierce the fried shell and sink effortlessly through to the other side
+in a delightful textural contrast.
+They do not know how to replicate it, and at this point are too invested in the narrative to just ask.
+
+They stay for three weeks in a ThursDz friend's guest room, only a little jealous of his homeownership.
+They tell him the preceding 2 chapters, and together they make what is known so far.
+He agrees it's too chewy.
+
+They play board games; they attend art festivals; they share a ride up to the observatory on a borrowed bike nicer than their own.
+
+They realize they need to move back.
 
@@
 
-\chapter{4}{Tenderness}
+\chapter{4}{Tender}
 
@@
 
-% TODO
+With the last week of their month-long vax celebration trip,
+the protagonist visits two other ThursDz friends in Raleigh.
+They speedrun Celeste (a nervous cat who never shows affection to strangers, except nonbinary people).
+Said friends are themselves a little like cats, visibly changed by their own quarantine experiences.
+All three humans are eager to share whatever fruits said experiences have borne.
+One of them is you.
+
+You, Seitan Reloaded, are not seitan wings.
+You are sausage, strongly spiced, in need of no sauce,
+to be put on scrambled eggs or pizza or simply to be eaten unadorned and cold from the fridge.
+Your author demonstrates you to the protagonist.
+He puts half a cup of paprika into you.
+He pours soy sauce into you, and keeps pouring -- the protagonist agog -- for impossibly long.
+He kneads you,
+and the protagonist remarks that they were afraid to knead their own because it already came out so tough,
+and he remarks that the olive oil and the nooch inside you help disrupt gluten formation,
+% so even with all the kneading it comes out pretty tender,
+and they think ah, this might be the final piece of the puzzle.
+
+June being Pride month,
+the protagonist's employer attempts some trans-themed programming,
+% (new from the generic corporate rainbows they usually do),
+but they dramatically neglect the ``nothing about us without us'' test
+and commit several terrible faux pas.
+Several trans employees complain in slack,
+and the brass tone police them into silence.
+One is nearly fired.
+The protagonist, already having been silenced previously,
+can only try to help behind the scenes,
+and it's good they're traveling so they have someone to weakly say ``hold me'' to
+while they work through their trauma response.
+Like always in protests, the surface problems get fixed, and the rest are left to become scars in the sands of time -- almost as long as it takes to knead you.
+
+With only a week in this visit, between all that and all the video games the humans need to nerd over, there isn't time to synthesize your wisdom into the main quest.
+Instead they fry up slices of you, toss you in buffalo sauce (as if you need it),
+and produce, speaking extremely charitably, a Scientific Byproduct.
+
+Cooking, baking, working matter with one's hands, is a wonderful grounding activity,
+and your texture, your tenderness is perfect.
+
+% Your texture is perfect.
 

### 2023-03-27 seitan/seitan.tex
some important changes

@@
 %
-In Shadyside Pittsburgh Appalachia USA
-you serve the canonical implementation of seitan wings.
+You serve the canonical implementation of seitan wings
+in Shadyside Pittsburgh Appalachia USA.
 If the NIST standards committee filed an entry for the food, they would surely enshrine one of yours.
@@
 
-You are one of the first stops they make,
-a high honor in the book of a foodie ex-grad student who knows the East End inside and out
-(friends in tow, of course).
+You are one of the first stops they make
+(friends in tow, of course),
+a high honor in the book of a foodie ex-grad-student who knows the East End inside and out.
 They have practiced Chapter 2's secret lost art, but something is still wrong,
 and sitting at your table in the summer heat they realize what it is:
-technique aside, even fully cooked, their dough recipe produces a chewy, glutinous inside,
-whereas your wings are perfectly soft inside;
+technique aside, even fully cooked, their dough produces a chewy, glutinous inside,
+whereas your wings are perfectly soft;
 one's teeth pierce the fried shell and sink effortlessly through to the other side
@@
 They tell him the preceding 2 chapters, and together they make what is known so far.
-He agrees it's too chewy.
+He likes it, but agrees it's too chewy.
 
@@
 
-Cooking, baking, working matter with one's hands, is a wonderful grounding activity,
+But cooking, baking, working matter with one's hands, is a wonderful grounding activity,
 and your texture, your tenderness is perfect.
@@
 
-\chapter{5}{Wing}
+\chapter{5}{Synthesis}
 
@@
 Where there was border collie Louie, there is now border collie Junior.
+You will never know if Louie could tell you smell different now.
 You are teaching your mom to play Heaven's Vault,
@@
 
+They take you on their traditional morning walk. % along the beach.
+You always preferred the snow to the beach,
+%but you come to understand why your mom does,
+but you still relish the feeling of sand in your toes and saltwater lapping at your calves.
+You had not felt the ocean on them since you made them yours.
+You walk Junior and you get seafood together from a stall along the boardwalk
+and suddenly the city of your birth feels like a bubble of immunity to that disaffected-millenial grief you carry everywhere else you go.
+
 You tell them the preceding 4 chapters,
@@
 
-\vspace{24em}
+\vspace{15em}
 
@@
 	{\em 2 Tbsp} -- Butter (omit for vegans) \\
-	{\em 2 Tbsp} -- Hot sauce (Frank's is canonical, but they prefer Cholula) \\
+	{\em 2 Tbsp} -- Hot sauce (Frank's is canonical, but you prefer Cholula) \\
 	\\
@@
 	\hline
-	{\em 1/3 cup} -- Blue cheese (a nice, creamy pungent one from Penn Mac) \\
+	{\em 1/3 cup} -- Blue cheese (a nice, creamy, pungent one from Penn Mac) \\
 	{\em 1/3 cup} -- Buttermilk (you can substitute but it really makes a difference) \\
@@
 \setlength{\parskip}{0.5em}
-	In two bowls, separately mix the dry and wet wingredients. Combine and knead for about 5 minutes, until coherent but also tears easily. Add more liquid or gluten as necessary, a little at a time.
+	In two bowls, separately mix the dry and wet wingredients. Combine and knead for 5-10 minutes, until coherent but also tears easily. Add more liquid or gluten as necessary, a little at a time.
 

### 2023-03-27 seitan/seitan.tex
more important changes

@@
 Everything you say is touched by the spectre of health food writing,
-because your reader must be into \textasciitilde{}vegan alternatives\textasciitilde{}.
-Drowning in your million useless results the protagonist knows one truth deep in their heart: Seitan wings is not a health food.
+because your think your reader must be into \textasciitilde{}vegan alternatives\textasciitilde{}.
+Drowning in your million useless faces the protagonist knows one truth deep in their heart: Seitan wings is not a health food.
 
@@
 and sitting at your table in the summer heat they realize what it is:
-technique aside, even fully cooked, their dough produces a chewy, glutinous inside,
-whereas your wings are perfectly soft;
+technique aside, even fully cooked, their dough yields a chewy, glutinous inside,
+whereas your insides are perfectly soft;
 one's teeth pierce the fried shell and sink effortlessly through to the other side
@@
 
-\chapter{4}{Tender}
+\chapter{4}{Tenderness}
 
@@
 and commit several terrible faux pas.
-Several trans employees complain in slack,
+Trans employees complain in slack,
 and the brass tone police them into silence.
@@
 
-With only a week in this visit, between all that and all the video games the humans need to nerd over, there isn't time to synthesize your wisdom into the main quest.
+With only a week in this visit, between that and all the video games the humans need to nerd over, there isn't time to synthesize your wisdom into the main quest.
 Instead they fry up slices of you, toss you in buffalo sauce (as if you need it),
@@
 
-\chapter{5}{Synthesis}
+\chapter{5}{Wings}
 
@@
 because your gender is new to them and your body is new to all three of you,
-but your parents will get there someday, beyond the scope of this work.
+but your parents will get there later, beyond the scope of this work.
 
@@
 
-\noindent You acknowledge Will Scott, William Lovas, Rob Simmons, Chris Martens, Eve Blum, David Blum, and Jenny Lin for their invaluable roles in the story.
+\noindent You thank Ellen Scherb, Will Scott, William Lovas, Rob Simmons, Chris Martens, Eve Blum, David Blum, and Jenny Lin for their invaluable roles in the story.
 
@@
 	% {\sc \Large ben blum's seitan wings}
-	{\sc \Large trans rights seitan wings}
+	% {\sc \Large trans rights seitan wings}
+	{\sc \large Quarantine seitan wings}
 \end{center}

### 2023-03-27 seitan/seitan.tex
last important changes

@@
 With the last week of their month-long vax celebration trip,
-the protagonist visits two other ThursDz friends in Raleigh.
+the protagonist visits two other ThursDz friends in Raleigh RTP USA.
 They speedrun Celeste (a nervous cat who never shows affection to strangers, except nonbinary people).
@@
 
-You, Seitan Reloaded, are not seitan wings.
-You are sausage, strongly spiced, in need of no sauce,
+You, Seitan Reloaded, are not wings.
+You are sausage, strongly spiced, fierce, in need of no sauce,
 to be put on scrambled eggs or pizza or simply to be eaten unadorned and cold from the fridge.
@@
 and they agree to share with you the honor of putting the pieces together,
+and you even make half with noncanonical sauce because your dad can't have spicy anymore,
 and that is the story of how seitan wings were made correctly in Santa Barbara for the first time ever.
 
-\vspace{15em}
+\vspace{14em}
 
